### 2018-01-02 article.tex
initial commit

\documentclass[a4paper]{article}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{bussproofs}
\usepackage{mathtools}
\usepackage{verbatim}
\usepackage{dsfont}
\usepackage[all, 2cell]{xy}
\usepackage[all]{xy}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{defin}{Definition}
\newtheorem{ex}{Example}
\newtheorem{col}{Corollary}
\newtheorem{con}{Convention}
\usepackage{listings} 		% for source code
\date{}
\title{Modal type theory based on the intuitionistic epistemic logic}

\begin{document}
\maketitle

\begin{abstract}
Modal intuitionistic epistemic logic IEL$^{-}$ was proposed by S.Artemov and T. Protopopescu as the formal
foundation for the intuitionistic theory of knowledge. We construct a modal simply typed lambda-calculus
which is Curry-Howard isomorphic to IEL$^{-}$ as formal theory of calculations with applicative functors in
functional programming languages like Haskell or Idris. We prove that this typed lambda-calculus has the
strong normalization and Church-Rosser properties.
\end{abstract}

  % This is a comment, not shown in final output.
  % The following shows typesetting  power of LaTeX:

  \section{Introduction}

  Modal intutionistic epistemic logic IEL was proposed by S. Artemov and T. Proropopescu \cite{Artemov}. IEL
  provides the epistimology and the theory of knowledge as based on BHK-semantics of intuitionistic logic.
  $IEL^{-}$  is a variant of IEL, that corresponds to intuitionistic belief.
  Informally, $\textbf{K}A$ denotes that $A$ is verified intuitionistically.

  Intuitionistic epistemic logic IEL$^{-}$ is defined with by following axioms and derivation rules:

  \begin{defin} Intuitionistic epistemic logic IEL:

    1) IPC axioms;

    2) $\textbf{K}(A \to B) \to (\textbf{K}A \to \textbf{K}B)$ (normality);

    3) $A \to \textbf{K}A$ (co-reflection);

    Rule: MP.

  \end{defin}

  We have the deduction theorem and necessitation rule which is derivable.

  V. Krupski and  A. Yatmanov provided the sequential calculus for IEL and proved that this calculus is
  PSPACE-complete \cite{Krupski}.

  It's not difficult to see that modal axioms in $IEL^{-}$ and types of the methods of Applicative class in
  Haskell-like languages (which is described below) are syntactically similar and we are going to show that
  this coincidence has a
  non-trivial computational meaning.

  Functional programming languages such as Haskell \cite{Haskell}, Idris \cite{Idris}, Purescript \cite{Purs}
  or Elm \cite{Elm} have special type classes\footnote{Type class in Haskell is a general interface for
  special group of datatypes.} for calculations with container types like \verb"Functor" and
  \verb"Applicative" \footnote{Reader may read more about container types in the Haskell standard library documentation\cite{Base} or in the next one textbook
  \cite{Miran}}:

  \begin{lstlisting}[language=Haskell]
    class Functor f where
      fmap :: (a -> b) -> f a -> f b

    class Functor f => Applicative f where
      pure :: a -> f a
      (<*>) :: f (a -> b) -> f a -> f b
  \end{lstlisting}

  By \emph{container} (or \emph{computational context}) type we mean some type-operator $f$, where $f$ is a
  ``function'' from $*$ to $*$: type operator takes a simple type (which has kind $*$) and returns another
  simple type type with kind \verb"*". For more detailed description of the type system with kinds used in
  Haskell see \cite{Morten}.

  The main goal of our research is a relationship between intuitionistic epistemic logic $IEL^{-}$ and
  functional programming with effects. We show that relationship by building the type system (which is called $\lambda_{\textbf{K}}$) which is Curry-Howard isomorphic to $IEL^{-}$. So we will consider $\textbf{K}$-modality as an arbitrary applicative functor.

  $\lambda K$ consists of the rules for simply typed lambda-calculus and special typing rules for
  lifting types into the applicative functor $\textbf{K}$. We assume that our type system will
  axiomatize the simplest case of computation with effects with one container. We provide proof-theoretical
  view on
  this kind of computations in functional programming and prove strong normalization and confluence.

  \section{Typed lambda-calculus based on IEL$^{-}$}

  At first we define the natural deduction for IEL$^{-}$ with $\textbf{K}$-modality and binary connectives
  $\to$ and $\land$ (we call that calculus NIEL$^{-}_{\land, \to}$):

  \begin{defin} Natural deduction NIEL$^{-}_{\land, \to}$ for IEL$^{-}$ with $\to$ and $\wedge$:

  \begin{center}
  \begin{prooftree}
  \AxiomC{}
  \RightLabel{\scriptsize{ax}}
  \UnaryInfC{$\Gamma , \alpha \vdash \alpha$}
  \end{prooftree}
  \end{center}

  \begin{minipage}{0.5\textwidth}
    \begin{flushleft}
  	\begin{prooftree}
  	\AxiomC{$\Gamma, \alpha \vdash \beta$}
  	\RightLabel{$\rightarrow_i$}
  	\UnaryInfC{$\Gamma \vdash \alpha \to \beta$}
    	\end{prooftree}

  	\begin{prooftree}
  	\AxiomC{ $\Gamma \vdash \alpha$ }
  	\AxiomC{ $\Gamma \vdash \beta$ }
  	\RightLabel{$\land_i$}
  	\BinaryInfC{$\Gamma \vdash \alpha \land \beta$}
  	\end{prooftree}

  	\begin{prooftree}
    	\AxiomC{$\Gamma \vdash \alpha$}
    	\RightLabel{$\textbf{K}_I$}
    	\UnaryInfC{$\Gamma \vdash \textbf{K}\alpha$}
  \end{prooftree}
    \end{flushleft}
  \end{minipage}
  \begin{minipage}{0.5\textwidth}
    \begin{flushright}
    	\begin{prooftree}
  	\AxiomC{$\Gamma \vdash \alpha \to \beta$}
  	\AxiomC{$\Gamma \vdash \alpha$}
  	\RightLabel{$\rightarrow_i$}
  	\BinaryInfC{$\Gamma \vdash \beta$}
    	\end{prooftree}

  	\begin{prooftree}
  	\AxiomC{ $\Gamma \vdash \alpha_1 \land \alpha_2$ }
  	\RightLabel{$\land_e, i \in \{ 1, 2 \}$}
  	\UnaryInfC{$\Gamma \vdash \alpha_i$}
  	\end{prooftree}

  	\begin{prooftree}
    	\AxiomC{$\Gamma \vdash \textbf{K}(\alpha \to \beta)$}
    	\AxiomC{$\Gamma \vdash \textbf{K}\alpha$}
    	\RightLabel{$\textbf{K}_{MP}$}
          \BinaryInfC{$\Gamma \vdash \textbf{K}\beta$}
  	\end{prooftree}
    \end{flushright}
  \end{minipage}
  \end{defin}

  \vspace{\baselineskip}

  \begin{lemma}
    NIEL$^{-}_{\land, \to} \vdash \alpha \Rightarrow$ IEL$^{-}$ $\vdash \alpha$.
  \end{lemma}

  \begin{proof}
Induction on the derivation.

  \vspace{\baselineskip}

Let us consider cases with modality.

\vspace{\baselineskip}

1) If $\Gamma \vdash_{NIEL^{-}_{\land, \to}} \alpha$, then $IEL^{-} \vdash \wedge \Gamma \rightarrow \textbf{K}\alpha$.

$\begin{array}{lll}
(1)& \wedge \Gamma \rightarrow \alpha & \text{assumption}\\
(2)& \alpha \rightarrow \textbf{K}\alpha &\text{co-reflection.}\\
(3)& (\wedge \Gamma \rightarrow \alpha) \rightarrow ((\alpha \rightarrow \textbf{K}\alpha) \rightarrow (\wedge \Gamma \rightarrow \textbf{K}\alpha))&\text{IPC theorem}\\
(4) &(\alpha \rightarrow \textbf{K}\alpha) \rightarrow (\wedge \Gamma \rightarrow \textbf{K}\alpha) &\text{1, 3, MP}\\
(5) &\wedge \Gamma \rightarrow \textbf{K}\alpha &\text{2,4, MP}\\
\end{array}$

\vspace{\baselineskip}

2) If $\Gamma \vdash_{NIEL^{-}_{\land, \to}} \textbf{K}(\alpha \to \beta)$ and $\Gamma \vdash_{NIEL^{-}_{\land, \to}} \textbf{K}\alpha$, then $IEL^{-} \vdash \wedge \Gamma \rightarrow \textbf{K}\beta$.

$\begin{array}{lll}
(1)& \Gamma \vdash_{NIEL^{-}_{\land, \to}} \textbf{K}(\alpha \to \beta) & \text{condition}\\
(2)& \Gamma \vdash_{NIEL^{-}_{\land, \to}} \textbf{K}\alpha & \text{condition}\\
(3)& \Gamma \vdash_{NIEL^{-}_{\land, \to}} \textbf{K}(\alpha \to \beta) \land \textbf{K}\alpha & \land_i \\
(4)& \wedge \Gamma \rightarrow \textbf{K}(\alpha \to \beta) \land \textbf{K}\alpha & \text{induction
hypothesis} \\
(5)& (\textbf{K}(\alpha \rightarrow \beta) \rightarrow (\textbf{K}\alpha \to \textbf{K}\beta)) \rightarrow
((\textbf{K}(\alpha \rightarrow \beta) \land \textbf{K}\alpha) \rightarrow \textbf{K}\beta)& \text{IPC
theorem} \\
(6)&\textbf{K}(\alpha \rightarrow \beta) \rightarrow (\textbf{K}\alpha \to \textbf{K}\beta)&
\text{normality}\\
(7)&(\textbf{K}(\alpha \rightarrow \beta) \land \textbf{K}\alpha) \rightarrow \textbf{K}\beta)& \text{5,6,
MP}\\
(8)&(\wedge \Gamma \rightarrow \textbf{K}(\alpha \to \beta) \land \textbf{K}\alpha) \rightarrow & \\
&\quad\rightarrow ((\textbf{K}(\alpha \rightarrow \beta) \land \textbf{K}\alpha) \rightarrow
\textbf{K}\beta) \rightarrow
(\wedge \Gamma \rightarrow \textbf{K}\beta))& \text{IPC theorem} \\
(9) &(\textbf{K}(\alpha \rightarrow \beta) \land \textbf{K}\alpha) \rightarrow \textbf{K}\beta) \rightarrow
(\wedge \Gamma \rightarrow \textbf{K}\beta)& \text{4, 8, MP} \\
(10) &\wedge \Gamma \rightarrow \textbf{K}\beta& \text{7, 9, MP}
\end{array}$

  \end{proof}

  \vspace{\baselineskip}

  At the next step we build the typed lambda-calculus based on NIEL$^{-}_{\land, \to}$ by proof-assingment in
  rules.

  At first, we define lambda-terms and types for this lambda-calculus.

  \vspace{\baselineskip}

  \begin{defin} The set of terms:

  Let $\mathbb{V}$ be the set of variables. The set $\Lambda_{\textbf{K}}$ of terms is defined by the grammar:
  \begin{equation}
    \Lambda_{\textbf{K}} ::= \mathbb{V} \: | \:  (\lambda \Lambda.\Lambda_{\textbf{K}}) \: | \:
    (\Lambda_{\textbf{K}}\Lambda_{\textbf{K}}) \: | \: (\Lambda_{\textbf{K}} , \Lambda_{\textbf{K}}) \: | \:
    (\pi_i \Lambda_{\textbf{K}}) \:
    | \: ({\bf pure \: } \: \Lambda_{\textbf{K}}) \: | \: (\Lambda_{\textbf{K}} \star
    \Lambda_{\textbf{K}})
  \end{equation}
  where $i \in \{ 1, 2 \}$.
  \end{defin}

  \begin{defin} The set of types:

  Let $\mathbb{T}$ be the set of atomic types. The set $\mathbb{T}_{\textbf{K}}$ of types with applicative
  functor
  $\textbf{K}$ is generated by the grammar:
  \begin{equation}
    \mathbb{T}_{\textbf{K}} ::= \mathbb{T} \: | \: (\mathbb{T}_{\textbf{K}} \to \mathbb{T}_{\textbf{K}}) \: |
    \:
    (\mathbb{T}_{\textbf{K}} \times \mathbb{T}_{\textbf{K}}) \: | \: (\textbf{K}\mathbb{T}_{\textbf{K}})
  \end{equation}
  \end{defin}

  Context, domain of context and range of context are defined standardly \cite{Neder}\cite{Morten}.

  Our type system is based on the Curry-style typing rules:

  \begin{defin} Modal typed lambda calculus $\lambda \textbf{K}$ based on NIEL$^{-}_{\land, \to}$:

  \begin{center}
  \begin{prooftree}
  \AxiomC{$ $}
  \RightLabel{\scriptsize{ax}}
  \UnaryInfC{$\Gamma , x : \alpha \vdash x : \alpha$}
  \end{prooftree}
  \end{center}

  \begin{minipage}{0.5\textwidth}
    \begin{flushleft}
  \begin{prooftree}
  \AxiomC{$\Gamma, x : \alpha \vdash M : \beta$}
  \RightLabel{$\rightarrow_i$}
  \UnaryInfC{$\Gamma \vdash \lambda x. M : \alpha \to \beta$}
  \end{prooftree}

  \begin{prooftree}
  \AxiomC{ $\Gamma \vdash x : \alpha$ }
  \AxiomC{ $\Gamma \vdash y : \beta$ }
  \RightLabel{$\times_i$}
  \BinaryInfC{$\Gamma \vdash (x, y) : \alpha \times \beta$}
  \end{prooftree}

  \begin{prooftree}
    \AxiomC{$\Gamma \vdash x : \alpha$}
    \RightLabel{$\textbf{K}_I$}
    \UnaryInfC{$\Gamma \vdash {\bf pure \: } \: x : \textbf{K}\alpha $}
  \end{prooftree}
    \end{flushleft}
  \end{minipage}

  \begin{minipage}{0.5\textwidth}
    \begin{flushright}
  \begin{prooftree}
  \AxiomC{$\Gamma \vdash f : \alpha \to \beta$}
  \AxiomC{$\Gamma \vdash x : \alpha$}
  \RightLabel{$\rightarrow_e$}
  \BinaryInfC{$\Gamma \vdash fx : \beta$}
  \end{prooftree}

  \begin{prooftree}
  \AxiomC{ $\Gamma \vdash p : \alpha_1 \times \alpha_2$ }
  \RightLabel{$\times_e$, $i \in \{ 1, 2 \}$}
  \UnaryInfC{$\Gamma \vdash \pi_i p : \alpha_i$}
  \end{prooftree}

  \begin{prooftree}
    \AxiomC{$\Gamma \vdash f : \textbf{K}(\alpha \to \beta)$}
    \AxiomC{$\Gamma \vdash x : \textbf{K}\alpha$}
    \RightLabel{$\textbf{K}_{app}$}
    \BinaryInfC{$\Gamma \vdash f \star x : \textbf{K}\beta$}
  \end{prooftree}
    \end{flushright}
  \end{minipage}
  \end{defin}

  $\textbf{K}_I$-typing rule is the same as $\bigcirc$-introduction in lax logic (also known as monadic
  metalanguage \cite{Lax}) and in typed lambda-calculus which is derived by proof-assignment for lax-logic
  proofs.
  $\textbf{K}_I$ allows to inject an object of type $\alpha$ into the functor. $\textbf{K}_I$  reflects the
  Haskell method {\bf pure} for Applicative class. It plays the same role as the {\bf return} method in Monad class.

  $\textbf{K}_{app}$ is a typing rule for lifted application. This rule is similar to
  $\rightarrow_e$-rule, therefore we introduce the infix operator $\star$ for lifted application.

  \begin{con}

  \emph{$\star$ is a left-associative, so $M \star N \star P = (M \star N) \star P$,  unless otherwise specified.}

  \end{con}

  \vspace{\baselineskip}

Here are some examples of derivation trees.

  \begin{prooftree}
  \AxiomC{$x : \alpha \vdash x : \alpha$}
  \RightLabel{$\textbf{K}_I$}
  \UnaryInfC{$x : \alpha \vdash {\bf pure \: } x : \textbf{K} \alpha$}
  \RightLabel{$\rightarrow_i$}
  \UnaryInfC{$\vdash (\lambda x. {\bf pure \: } x) : \alpha \to \textbf{K}\alpha$}
  \end{prooftree}

  \begin{prooftree}
    \AxiomC{$f : \textbf{K}(\alpha \to \beta) \vdash f : \textbf{K}(\alpha \to \beta)$}
    \AxiomC{$x : \textbf{K}\alpha \vdash x : \textbf{K}\alpha$}
    \RightLabel{$\textbf{K}_{app}$}
    \BinaryInfC{$f : \textbf{K}(\alpha \to \beta), x : \textbf{K}\alpha \vdash f \star x :
    \textbf{K}\beta$}
    \RightLabel{$\rightarrow_i$}
    \UnaryInfC{$f : \textbf{K}(\alpha \to \beta) \vdash \lambda x. f \star x : \textbf{K}\alpha \to \textbf{K}\beta$}
    \RightLabel{$\rightarrow_i$}
    \UnaryInfC{$\vdash \lambda f. \lambda x. f \star x: \textbf{K}(\alpha \to \beta) \to
    \textbf{K}\alpha \to \textbf{K}\beta$}
  \end{prooftree}

  \vspace{\baselineskip}

  Now we define free variables and substitutions. $\beta$-reduction, multi-step $\beta$-reduction and $\beta$-equality are defined standardly:

  \begin{defin} Set $FV(M)$ of free variables for arbitrary term $M$:

  1) $FV(x) = \{ x \}$;

  2) $FV(\lambda x. M) = FV(M) \setminus \{ x\}$;

  3) $FV(M N) = FV(M) \cup FV(N);$

  4) $FV((M,N)) = FV(M) \cup FV(N)$;

  5) $FV(\pi_i p) \subseteq FV(p)$, $i \in \{ 1, 2\}$;

  6) $FV(\text{pure } M) = FV(M)$;

  7) $FV(M \star N) = FV(M) \cup FV(N)$.
  \end{defin}

  \begin{defin} Substitution:

  1) $x [x := N] = N$, $x [y := N] = x$;

  2) $(M N) [x := N] = M[x := N] N [x := N]$;

  3) $(\lambda x. M) [x := N] = \lambda x. M [x := N]$;

  4) $(M, N)[x := P] = (M[x := P], N [x := P])$;

  5) $(\pi_i M) [x := P] = \pi_i (M[x := P])$, $i \in \{ 1, 2\}$;

  6) $({\bf pure \: } M) [x := P] = {\bf pure \: } (M [x := P])$;

  7) $(M \star N) [x := P] = (M [x := P]) \star (N [x := P])$.
  \end{defin}

  In $\lambda \textbf{K}$ we have the following computational rules. We will define them for terms with
  ${\bf pure \: }$ or $\star$. Reduction rules for $(,)$ and $\pi_i$ are described, for example, in
  \cite{Pierce}.

  \begin{defin} $\beta$-reduction rules for \emph{$\lambda \textbf{K}$}.

  1) ${\bf pure \: } (\lambda x. x) \star M \rightarrow_{\beta} M$;

  2) ${\bf pure \: } (\lambda f g x. f (g x)) \star M \star N \star P \rightarrow_{\beta} M \star (N \star
  P)$;

  3) $({\bf pure \: } M) \star (\text{pure } N) \rightarrow_{\beta} \text{pure } (M N)$;

  4) $M \star {\bf pure \: } N \rightarrow_{\beta} (\lambda f. f N) \star M$;

  5) $M_1 \rightarrow_{\beta} M_2 \Leftrightarrow {\bf pure \: } \: M_1 \rightarrow_{\beta} {\bf pure \: } \: M_2$;

  6) $M_1 \rightarrow_{\beta} M_2 \Rightarrow M_1 \star N \rightarrow_{\beta} M_2 \star N$;

  7) $N_1 \rightarrow_{\beta} N_2 \Rightarrow M \star N_1 \rightarrow_{\beta} M \star N_2$;
  \end{defin}

  In Haskell any implementation of the Applicative instance should preserve the laws proposed by McBride and
  Paterson \cite{McP}. We preserve these laws by the introduction of corresponding reduction rules (rules
  (1)-(4)). These rules will be called respectively identity rule, composition rule, homomorphism rule and
  interchange rule.

  Rules (5)-(7) are would be useful for strong normalization proof.

  \section{Basic lemmas}

  Now we will prove standard lemmas for contexts in type systems\footnote{We will not prove cases with
  $\to$-constructor, they are proved standardly in the same lemmas for simply typed lambda calculus, for
  example see \cite{Neder}\cite{Morten}\cite{Girard}. We will consider only modal cases}:

  \begin{defin} The domain of a context $\Gamma$:

  Let $\Gamma = \{ x_1 : \alpha_1, ..., x_n : \alpha_n \}$. Then the domain of $\Gamma$, or $dom(\Gamma)$, is a set $\{  x_1, ..., x_n\}$.

  \end{defin}

  \begin{lemma}

  If $\Gamma \vdash M : \alpha$, then $FV(M) \subseteq dom(\Gamma)$

  \end{lemma}

  \begin{proof}

  Induction on the derivation of $\Gamma \vdash M : \alpha$.

  \end{proof}

  \vspace{\baselineskip}

  \begin{lemma} Generation for $\lambda \textbf{K}$.

  1) $\Gamma \vdash {\bf pure \: } M : \textbf{K}\alpha$ implies that $\Gamma \vdash M : \alpha$;

  2) $\Gamma \vdash M \star N : \textbf{K}\beta$ implies that $\Gamma \vdash M : \textbf{K}(\alpha \to \beta)$ and
  $\Gamma \vdash N : \textbf{K}\alpha$.
  \end{lemma}

  \begin{proof}

  $ $

  Simple induction on the derivation of $\Gamma \vdash {\bf pure \: } M : \textbf{K}\alpha$ and $\Gamma \vdash M \star N : \textbf{K}(\alpha \to \beta)$ respectively.
  \end{proof}

  \vspace{\baselineskip}

  The next one lemma allows that weakening structural rule is admissable.

  \begin{lemma} Weakening for $\lambda \textbf{K}$.

  Let $\Gamma \vdash M : \alpha$ and $\Gamma \subseteq \Delta$, then $\Delta \vdash M : \alpha$.

  \end{lemma}

  \begin{proof}

  $ $

  Induction on derivation of $\Gamma \vdash M : \alpha$. Let us assume $\Gamma \subseteq \Delta$.

  1) Let $\Gamma \vdash x : \alpha$, such that $\Gamma = \Delta, x : \alpha$ and $\Theta \subseteq \Gamma$.
  Let $\Sigma = \Theta \setminus \Gamma$, or, which is the same, $\Sigma = \Theta \setminus \Delta, x :
  \alpha$, then $\Sigma,
  \Delta, x : \alpha \vdash x : \alpha$, or, $\Theta \vdash x : \alpha$.

  2) Let $\Gamma \vdash {\bf pure \: } M : \textbf{K}\alpha$ and $\Gamma \subseteq \Theta$.

  If $\Gamma \vdash {\bf pure \: } M : \textbf{K}\alpha$, then $\Gamma \vdash M : \alpha$ by generation and, by
  hypothesis, $\Theta \vdash M : \alpha$, so $\Theta \vdash {\bf pure \: } M : \textbf{K}\alpha$ by applying
  $\textbf{K}_I$-rule.

  3) Let $\Gamma \vdash M \star N : \textbf{K}\beta$. So $\Gamma \vdash M : \textbf{K}(\alpha \to \beta)$ and $\Gamma
  \vdash N : \textbf{K}\alpha$. By hypothesis $\Delta \vdash M : \textbf{K}(\alpha \to \beta)$ and $\Delta \vdash N :
  \textbf{K}\alpha$. Then $\Delta \vdash M \star N : \textbf{K}\beta$.

  \end{proof}

  \begin{lemma} Considering for $\lambda \textbf{K}$.

  If $\Gamma \vdash M : \alpha$, then $\Gamma \uparrow FV(M) \vdash M : \alpha$, where $\Gamma \uparrow
  FV(M)$ is a subcontext of $\Gamma$, such that $dom(\Gamma \uparrow FV(M)) = dom(\Gamma) \cap FV(M)$.

  \end{lemma}

  \begin{proof}

  Induction by derivation. We consider the base of induction and the case with $\textbf{K}_{app}$. The rest
  cases are
  proven by the same way.

  $ $

  1) Let $\Gamma \vdash x : \alpha$, where $\Gamma = \Delta, x : \alpha$, $x \in \mathbb{V}$.

  $FV(x) = \{ x\}$, then $dom(\Gamma) \cap \{ x\} = \{ x\}. $ So $(\Delta, x : \alpha) \uparrow FV(x) = \{ x
  : \alpha\}$, then $x : \alpha \vdash x : \alpha$ by axiom.

  2) Let $\Gamma \vdash M \star N : \textbf{K}\beta$.

  By generation $\Gamma \vdash M : \textbf{K}(\alpha \to \beta)$ and $\Gamma \vdash N : \textbf{K}\alpha$.

  By induction hypothesis $\Gamma \uparrow FV(M) \vdash M : \textbf{K}(\alpha \to \beta)$ and $\Gamma
  \uparrow FV(N) \vdash N : \textbf{K}\beta$.

  But $\Gamma \uparrow FV(M) \subseteq \Gamma \uparrow FV(M \star N)$ and $\Gamma
  \uparrow FV(N) \subseteq \Gamma \uparrow FV(M \star N)$, so $\Gamma \uparrow FV(M \star N) \vdash M :
  \textbf{K}(\alpha \to \beta)$ and $\Gamma \uparrow FV(M \star N) \vdash N : \textbf{K}\alpha$ by weakening.

  Then $\Gamma \uparrow FV(M \star N) \vdash M \star N : \textbf{K}\beta$ by $\textbf{K}_{app}$.

  \end{proof}

  \vspace{\baselineskip}

  \begin{lemma}

  If $\Gamma, x : \alpha \vdash M : \beta$  and $\Gamma \vdash N : \alpha$, then $\Gamma \vdash (M [x := N]) :
  \beta $

  \end{lemma}

  \begin{proof}

  $ $

  1) Let $\Gamma, x : \alpha \vdash {\bf pure \: } M : \textbf{K}\beta$ and $\Gamma \vdash N : \alpha$.

  If $\Gamma, x : \alpha \vdash {\bf pure \: } M : \textbf{K}\beta$, then, by generation, $\Gamma, x : A \vdash
  M : \beta$. So, by induction hypothesis, $\Gamma \vdash (M [x := N]) : \beta $, then $\Gamma \vdash
  {\bf pure \: } (M [x := N]) : \textbf{K}\beta$ by $\textbf{K}_I$, but ${\bf pure \: } (M [x := N]) =
  ({\bf pure \: } M) (M [x := N])$
  by substitution definition, so $\Gamma \vdash ({\bf pure \: } M) (M [x := N]) : \textbf{K}\beta$

  2) Let $\Gamma, x : \gamma \vdash M \star N : \textbf{K}\beta$, and $\Gamma \vdash y : \gamma$.

  So, by generation, $\Gamma, x : \gamma \vdash M : \textbf{K}(\alpha \to \beta)$ and $\Gamma, x : \gamma
  \vdash N : \textbf{K}\alpha$.

  Hence $\Gamma, x : \gamma \vdash (M [x := y]) : \textbf{K}(\alpha \to \beta)$ and
  $\Gamma \vdash (N [x := y]): \textbf{K}\alpha$ by hypothesis.

  So $\Gamma \vdash (M [x := y]) \star (N [x := y]) : \textbf{K}\beta$, or, $\Gamma \vdash (M \star N) ([x
  := y]) : \textbf{K}\beta$.

  \end{proof}

  \vspace{\baselineskip}

  \begin{theorem} Subject reduction

  $ $

  Let $\Gamma \vdash M : \alpha$ and $M \twoheadrightarrow_{\beta} N$, then $\Gamma \vdash N : \alpha$

  \end{theorem}

  We consider cases with reduction rules which are applicative laws. The general statement for $\twoheadrightarrow_{\beta}$ follows from transitivity of multi-step $\beta$-reduction.

  \begin{proof}

  $ $

\vspace{\baselineskip}

  1) Let $\Gamma \vdash {\bf pure \: } (\lambda x. x) \star M : \textbf{K}\alpha$. Then $\Gamma \vdash
  {\bf pure \: } (\lambda x.x) : \textbf{K}(\alpha \to \alpha)$ and $\Gamma \vdash M : \textbf{K}\alpha$ by
  generation. Then $\Gamma \vdash M : \textbf{K}\alpha$ trivially.

\vspace{\baselineskip}

  2) Let $\Gamma \vdash {\bf pure \: }(\lambda f g x. f (g x)) \star M \star N \star P : \textbf{K}\gamma$.

  Then $\Gamma \vdash {\bf pure \: } (\lambda f g x. f (g x)) : \textbf{K}((\beta \to \gamma) \to (\alpha \to \beta)
  \to \alpha \to \gamma)$, $\Gamma \vdash M : \textbf{K}(\beta \to \gamma)$, $\Gamma \vdash N :
  \textbf{K}(\alpha \to \beta)$ and $\Gamma \vdash P : \textbf{K}\alpha$ by generation.

  If $\Gamma \vdash N : \textbf{K}(\alpha \to \beta)$ and $\Gamma \vdash P : \textbf{K}\alpha$, then $\Gamma
  \vdash N \star P : \textbf{K}\beta$ by $\textbf{K}_{app}$.

  Hence, if $\Gamma \vdash M : \textbf{K}(\beta \to \gamma)$, then $\Gamma \vdash M \star (N \star P) :
  \textbf{K}\gamma$ by $\textbf{K}_{app}$.

\vspace{\baselineskip}

  3) Let $\Gamma \vdash ({\bf pure \: } M) \star ({\bf pure \: } N) : \textbf{K}\beta$. Then $\Gamma \vdash
  {\bf pure \: } M : \textbf{K}(\alpha \to \beta)$ and $\Gamma \vdash {\bf pure \: } N : \textbf{K}\alpha$ by
  generation. Moreover, $\Gamma \vdash M : \alpha \to \beta$ and $\Gamma \vdash N : \alpha$.

  Then $\Gamma \vdash M N : \beta$ by application.

  Hence, $\Gamma \vdash {\bf pure \: } (M N) : \textbf{K}\beta$ by $K_I$.

\vspace{\baselineskip}

  4) Let $\Gamma \vdash M \star ({\bf pure \: } N) : \textbf{K}\beta$.

  Then $\Gamma \vdash M : \textbf{K}(\alpha \to \beta)$ and $\Gamma \vdash {\bf pure \: } N : \textbf{K}\alpha$.

  Moreover, $\Gamma \vdash N : \alpha$ by generation.

  Let $\Gamma, f : \alpha \to \beta \vdash f : \alpha \to \beta$ and $\Gamma, f : \alpha \to \beta \vdash N :
  \alpha$ by weakening.

  So $\Gamma, f : \alpha \to \beta \vdash f N : \beta$ by application, so $\Gamma \vdash \lambda f. f N :
  (\alpha \to \beta) \to \beta$ by abstraction.

  Then $\Gamma \vdash {\bf pure \: } (\lambda f. f N) : \textbf{K}((\alpha \to \beta) \to \beta)$ by $\textbf{K}_I$.

  Hence, $\Gamma \vdash {\bf pure \: } (\lambda f. f N) \star M : \textbf{K}\beta$.

  \end{proof}

  \vspace{\baselineskip}

\section{Strong normalization}

  We modify and apply Tait's technique of logical relation for modalities. Strong normalization proof with Tait's method for simply typed lambda calculus is described here \cite{Pierce}.

  \begin{theorem}

    Let $M \in \Lambda_{\textbf{K}}$, then any sequence of reduction $M \rightarrow_{\beta} M_1 \dots$ terminates.

  \end{theorem}

  \begin{proof}

  We build the smallest of subset of strongly normalazing terms of modal types and show that an arbitrary
  term belongs to this subset.

  \begin{defin} The set of strongly computable terms of type $\phi \in \mathbb{T}_{\textbf{K}}$, $SC_{\phi}$:

  \begin{itemize}
  \item Let $\phi = \textbf{K}\alpha$ and $\alpha \in \mathbb{T}$ , then:
  \begin{equation}
  SC_{\textbf{K}\alpha} = \{ M : \textbf{K}\alpha \: | \: \text{$M$ is strongly normalizing} \}
  \end{equation}

  \item Let $\phi = \textbf{K}(\tau \to \psi)$ and $\tau, \psi \in \mathbb{T}_{\textbf{K}}$, then:
  \begin{equation}
  SC_{\textbf{K}(\tau \to \psi)} = \{ M : \textbf{K}(\tau \to \psi) \: | \: \forall N \in SC_{\textbf{K}\tau}, M \star N \in SC_{\textbf{K}\psi}\}
  \end{equation}

  \item Let $\phi = \textbf{K}(\tau_1 \times \tau_2)$ and $\tau_1, \tau_2 \in \mathbb{T}_{\textbf{K}}$, then:
  \begin{equation}
  SC_{\textbf{K}(\tau_1 \times \tau_2)} = \{ P : \textbf{K}(\tau_1 \times \tau_2) \: |
  \: {\bf pure \:} (\lambda x. \pi_i x) \star P \in SC_{\textbf{K}\tau_i}, i \in \{1,2\} \}
  \end{equation}
  \end{itemize}
  \end{defin}

  \begin{lemma}

  $ $

  If $M \in SC_{\alpha}$, then $M$ is strongly normalizing.

  \end{lemma}

  \begin{proof}

$ $

  1) If $M \in SC_{{\bf K} \alpha}$ and $\alpha \in \mathbb{T}$, then $M$ is strongly normalizing by the
  definition of $SC_{{\bf K} \alpha}$.

  2) Let $M \in SC_{\textbf{K}(\tau \to \psi)}$, so by every $N \in SC_{\textbf{K}\tau}$,
  $M \star N \in SC_{\textbf{K}\psi}$, which is strongly normalizing by hypothesis. So $M$ is
  strongly normalizing.

  3) Let $M \in SC_{\textbf{K}(\tau_1 \times \tau_2)}$, so ${\bf pure \:} (\lambda x. \pi_i x) \star M \in
  SC_{\textbf{K}\tau_i}$, $i \in \{ 1,2 \}$, which are strongly normalizing. So $M$ is
  strongly normalizing.

  \end{proof}

  \begin{lemma}

  $ $

  Let $M \rightarrow_{\beta} M'$ and $M \in SC_{\alpha}$, then $M' \in SC_{\alpha}$.

  \end{lemma}

  \begin{proof}

$ $

  1) Let $M \rightarrow_{\beta} M'$ and $M \in SC_{\textbf{K}\alpha}$, where $\alpha \in \mathbb{T}$.

  $M$ has the longest reduction path (which we denote as $p(M)$). So $p(M') < p(M)$, then $M' \in SC_{\textbf{K}\alpha}$.

  2) Let $M \in SC_{\textbf{K}(\alpha \to \beta)}$ and $M \rightarrow_{\beta} M'$. Let $N \in
  SC_{\textbf{K}\alpha}$. So $M \star N \in SC_{\textbf{K}\beta}$.

  If $M \rightarrow_{\beta} M'$, then $M \star N \rightarrow_{\beta} M' \star N$ by reduction
  rule, so $M' \star N \in SC_{\textbf{K}\beta}$ and $M' \in SC_{\textbf{K}(\alpha \to \beta)}$ by
  hypothesis.

  3) Let $M \in SC_{\textbf{K}(\tau_1 \times \tau_2)}$ and $M \rightarrow_{\beta} M'$.

  So ${\bf pure \: } (\lambda x. \pi_i x) \star M \rightarrow_{\beta} {\bf pure \: } (\lambda x. \pi_i x)
  \star M'$, $i \in \{ 1, 2 \}$ by reduction rule. So ${\bf pure \: } (\lambda x. \pi_i x) \star M' \in
  SC_{\textbf{K} \tau_i }$ and $M' \in SC_{\textbf{K}(\tau_1 \times \tau_2)}$.

  \end{proof}

  \begin{defin} Neutral term:

  We define a term $M$ to be neutral if it has of the next forms:

  1) $M = x$, where $x \in \mathbb{V}$;

  2) $M = (P Q)$;

  3) $M = \pi_i M$, $i \in \{ 1,2 \}$;

  4) $M = P \star Q$;

  5) If $M$ is a neutral, then ${\bf pure \: } M$ is a neutral.

  \end{defin}

  \begin{lemma}

  Let $M \rightarrow_{\beta} M'$ and $M' \in SC_{\alpha}$ for every one-step reduction. So if $M'$ is a
  neutral, then $M \in SC_{\alpha}$.

  \end{lemma}

  \begin{proof}

  $ $

  Simple induction on the structure of $M'$.

  \end{proof}

  \begin{lemma}

  $ $

  Let $x_1 : \phi_1, \dots, x_n : \phi_n \vdash M : \phi$ and for all $i \in \{ 1, \dots, n\}$, $N_i
  \in SC_{\phi_i}$, then $(M [x_1 := N_1, \dots, x_n := N_n]) \in SC_{\phi}$.

  \end{lemma}

  \begin{proof}

$ $

  1) If $\phi$ is an atomic and $M$ is a variable, then this condition holds straightforwardly.

  2) Let $\Gamma = \{ x_1 : \phi_1, \dots, x_n : \phi_n \}$, $\Gamma \vdash {\bf pure \: } M :
  \textbf{K}\alpha$ and for all $i \in \{1, \dots, n\}, N_i \in SC_{\phi_i}$.

  Then by $\Gamma \vdash M : \alpha$ by generation
  and $(M [x_1 := N_1, \dots, x_n := N_n]) \in SC_{\alpha}$ by induction hypothesis.

  Hence, $\Gamma \vdash {\bf pure \: } M : \textbf{K}\alpha$ and $({\bf pure \: } M ([x_1 := N_1, \dots, x_n := N_n])) \in SC_{\textbf{K}\alpha}$ by definition of $SC_{\textbf{K}\alpha}$.

  3) Let $\Gamma = \{ x_1 : \phi_1, \dots, x_n : \phi_n \}$, $\Gamma : \phi_n \vdash M \star P : \textbf{K}\beta$ and forall $i \in \{ 1, \dots, n\}, N_i \in SC_{\phi_i}$.

  Then $\Gamma \vdash M : \textbf{K}(\alpha \to \beta)$, $\Gamma \vdash P : \textbf{K}\alpha$ by generation.

  But by induction hypothesis $M [x_1 := N_1, \dots, x_n := N_n] \in SC_{\textbf{K}(\alpha \to
  \beta)}$ and $P [x_1 := N_1, \dots, x_n := N_n] \in SC_{\textbf{K}\alpha}$.

  Then, by definition of $SC_{\textbf{K}\beta}$, $((M [x_1 := N_1, \dots, x_n := N_n]) \star (P [x_1 := N_1, \dots, x_n := N_n])) \in SC_{\textbf{K}\beta}$, i.e. $(M \star N ([x_1 := N_1, \dots, x_n := N_n])) \in
  SC_{\textbf{K}\beta}$.
  \end{proof}

  \begin{col}

  $ $

  If $\vdash M : \alpha$, then $M$ is strongly normalizing.

  \end{col}

  \begin{proof}

  $M \in SC_{\alpha}$ by Lemma 10, so $M$ is strongly normalizing.

  \end{proof}

  \end{proof}

\section{Confluence}

In the confluence proof (below) we treat the cases with ${\bf pure}$ and $\star$ similar to \cite{Baren} \cite{Baren2}.

\begin{defin} Alphabet for the labelled terms:

variables: $x, y, z, x_1, y_1, z_1, ...$;

lambdas: $\lambda, \lambda_0, \lambda_1, \lambda_2, ...$;

constructors for an applicative functor: ${\bf pure \: }, \star$;

parentheses (,).

\end{defin}

\begin{defin} The set of labelled terms $\Lambda^{'}_{\textbf{K}}$ inductively defined as a set of words on the alphabet
described above:

1) $x \in \Lambda^{'}$;

2) If $M \in \Lambda^{'}_{\textbf{K}}$, then $(\lambda x. M) \in \Lambda^{'}_{\textbf{K}}$;

3) If $M, N \in \Lambda^{'}_{\textbf{K}}$, then $(MN) \in \Lambda^{'}_{\textbf{K}}$;

4) If $M \in \Lambda^{'}_{\textbf{K}}$, then ${\bf pure \: } M \in \Lambda^{'}_{\textbf{K}}$;

5) If $M, N \in \Lambda^{'}_{\textbf{K}}$, then $M \star N \in \Lambda^{'}_{\textbf{K}}$;

6) If $M, N \in \Lambda^{'}_{\textbf{K}}$, then for all $i \in \mathbb{N}$, $((\lambda_i x. M) N) \in
\Lambda^{'}_{\textbf{K}}$.

\end{defin}

\begin{defin} Erasing map

Erasing map is a map $|.| : \Lambda^{'}_{\textbf{K}} \to \Lambda_{\textbf{K}}$, such that:

1) $|x| = x$;

2) $|(\lambda x. M)| = \lambda x. |M|$;

3) $|(M N)| = |M| |N|$;

4) $|({\bf pure \: } M)| = {\bf pure \: } |M|$;

5) $|M \star N| = |M| \star |N|$;

6) $|((\lambda_i x. M) N)| = (\lambda x. |M|) |N|$

\end{defin}

\begin{ex}

$ $

$|{\bf pure \: } ((\lambda_i x. M) N) \star P| = {\bf pure \: } (\lambda x. |M|) |N|) \star |P|$

\end{ex}

\begin{defin} Substitution for $\Lambda^{'}_{\textbf{K}}$:

1) $x [x := N] = N$, $x [y := N] = x$;

2) $(M N) [x := N] = M[x := N] N [x := N]$;

3) $(\lambda x. M) [x := N] = \lambda x. M [x := N]$;

4) $({\bf pure \: } M) [x := P] = {\bf pure \: } \: (M [x := P])$;

5) $(M \star N) [x := P] = (M [x := P]) \star (N [x := P])$;

6) $(\lambda_i x. M) N [y := P] = (\lambda_i x. M [y := P]) (N [y := P])$.

\end{defin}

\begin{defin} One-step reduction $\rightarrow_{\beta^{'}}$ for $\Lambda^{'}_{\textbf{K}}$:

1) $(\lambda x. M) N \rightarrow_{\beta^{'}} M [x := N]$;

2) ${\bf pure \: } (\lambda x. x) \star M \rightarrow_{\beta^{'}} M$;

3) ${\bf pure \: } (\lambda f g x. f (g x)) \star M \star N \star P \rightarrow_{\beta^{'}} M \star (N \star
P)$;

4) $({\bf pure \: } M) \star ({\bf pure \: } N) \rightarrow_{\beta^{'}} {\bf pure \: } (M N)$;

5) $M \star ({\bf pure \: } N) \rightarrow_{\beta^{'}} {\bf pure \: } (\lambda f. f N) \star M$;

6) $(\lambda_i x. M) N \rightarrow_{\beta^{'}} M [x := N]$.

Multi-step reduction $\twoheadrightarrow_{\beta^{'}}$ is a reflexive-transitive closure of
$\rightarrow_{\beta^{'}}$.

\end{defin}

\begin{defin} Let us define a map $\phi : \Lambda^{'}_{\textbf{K}} \to \Lambda_{\textbf{K}}$ inductively as
follows:

1) $\phi(x) = x$;

2) $\phi(M N) = \phi(M)\phi(N)$;

3) $\phi(\lambda x. M) = \lambda x. \phi(M)$;

4) $\phi({\bf pure \: } M) = {\bf pure \: } (\phi(M))$;

5) $\phi(M \star N) = \phi(M) \star \phi(N)$;

6) $\phi((\lambda_i x. M) N) = \phi(M) [x := \phi(N)]$.

\end{defin}

\begin{ex}

$ $

$\phi({\bf pure \: } ((\lambda_i x. M) N) \star P) = {\bf pure \: } (\phi(M) [x := \phi(N)]) \star \phi(P)$

\end{ex}

\begin{lemma}

$ $

1) Let $M, N \in \Lambda^{'}_{\textbf{K}}$ and $|M| \twoheadrightarrow_{\beta} |N|$, then $M \twoheadrightarrow_{\beta^{'}} N$.

2) Let $M, N \in \Lambda^{'}_{\textbf{K}}$ and $M \twoheadrightarrow_{\beta^{'}} N$, then $|M| \twoheadrightarrow_{\beta} |N|$.

\end{lemma}

\begin{proof}

$ $

Induction on the generation of $\twoheadrightarrow_{\beta}$ ($\twoheadrightarrow_{\beta^{'}}$).

1) Let us consider homomorphism rule. The rest applicative reduction rules are considered similary.

Let $({\bf pure \: } M') \star ({\bf pure \: } N'), {\bf pure \: } (M' N') \in \Lambda^{'}_{\textbf{K}}$.

So $|({\bf pure \: } M') \star ({\bf pure \: } N')| = ({\bf pure \: } |M'|) \star ({\bf pure \: } |N'|)$ and
$|{\bf pure \: } (M' N')| = {\bf pure \: } (|M'| |N'|)$.

By reduction rule, $({\bf pure \: } |M'|) \star ({\bf pure \: } |N'|) \rightarrow_{\beta} {\bf pure \: }
(|M'||N'|)$.

But $({\bf pure \: } M') \star ({\bf pure \: } N') \rightarrow_{\beta^{'}} {\bf pure \: } (M' N')$ by
reduction rule for $\rightarrow_{\beta^{'}}$.

  \vspace{\baselineskip}

2) Let us consider interchange rule.

Let $M \star ({\bf pure \: } N), {\bf pure \: } (\lambda f. f N) \star M \in \Lambda^{'}_{\textbf{K}}$ and
$M \star ({\bf pure \: } N) \rightarrow_{\beta^{'}} {\bf pure \: } (\lambda f. f N) \star M$.

But $|M \star ({\bf pure \: } N)| = |M| \star ({\bf pure \: } |N|)$ and $|{\bf pure \: } (\lambda f. f N)
\star M| = {\bf pure \: } (\lambda f. f |N|) \star |M|$.

So $|M| \star ({\bf pure \: } |N|) \rightarrow_{\beta} {\bf pure \: } (\lambda f. f |N|) \star |M|$ by
$\beta$-reduction rule.

\vspace{\baselineskip}

It is easy to see, that the statement for $\twoheadrightarrow_{\beta^{'}}$ and $\twoheadrightarrow_{\beta}$
immedeatly follows from transitivity of multi-step rediction for labelled terms and for usual terms
respectively.

\end{proof}

\begin{lemma}

$ $

$\phi(M [x := N]) = \phi(M) [x := \phi(N)]$.

\end{lemma}

\begin{proof} Induction on $M$.

1) Let $M = x$. Then $\phi(x [x := N]) = \phi(N)$.

On the other hand, $\phi(x) [x := \phi(N)] = x [x := \phi(N)] = \phi(N)$.

So $\phi(x [x := N]) = \phi(x) [x := \phi(N)]$.

  \vspace{\baselineskip}

2) Let $M = y$ and $y \neq x$. Then $\phi(y [x := N]) = \phi(y) = y$.

But $\phi(y) [x := \phi(N)] = y [x := \phi(N)] = y$.

Therefore $\phi(y [x := N]) = \phi(y) [x := \phi(N)]$.

  \vspace{\baselineskip}

3) Let $M = {\bf pure \: } M'$. Then $\phi({\bf pure \: } M' [x := N]) = {\bf pure \: } \phi(M' [x := N])$.

By hypothesis, ${\bf pure \: } (\phi(M' [x := N])) = {\bf pure \: } (\phi(M') [x := \phi(N)])$, which is
$({\bf pure \: } \phi(M')) [x := \phi(N)]$ by substitution definition.

\vspace{\baselineskip}

4) Let $M = M' \star N'$. So $\phi((M' \star N') [x := N])) = \phi(M' [x := N] \star N' [x := N])$.

By definition of $\phi$,

\quad $\phi(M' [x := N] \star N' [x := N]) = \phi(M' [x := N]) \star \phi(N' [x := N])$.

But by induction hypothesis,

\quad $\phi(M' [x := N]) = \phi(M') [x := \phi(N)]$ and

\quad $\phi(N' [x := N]) = \phi(N') [x := \phi(N)]$.

Hence,

\quad $\phi(M' [x := N]) \star \phi(N' [x := N]) = \phi(M') [x := \phi(N)] \star \phi(N') [x := \phi(N)]$.

So,

\quad $\phi(M') [x := \phi(N)] \star \phi(N')[x := \phi(N)] = (\phi(M') \star \phi(N')) [x := \phi(N)]$.

And by definition of $\phi$, $(\phi(M') \star \phi(N')) [x := \phi(N)] = \phi(M' \star N') [x := \phi(N)]$.

\end{proof}

\begin{lemma}

$ $

  Let $M, N \in \Lambda^{'}_{\textbf{K}}$ and $M \twoheadrightarrow_{\beta^{'}} N$, then $\phi(M)
  \twoheadrightarrow_{\beta} \phi(N)$.

\end{lemma}

\begin{proof}

$ $

1) Let ${\bf pure \: } (\lambda x. x) \star M, M \in \Lambda^{'}_{\textbf{K}}$ and ${\bf pure \: } (\lambda
x. x) \star M \rightarrow_{\beta^{'}} M$.

But $\phi ({\bf pure \: } (\lambda x. x) \star M) = {\bf pure \: } (\lambda x. x) \star \phi(M)$.\

So ${\bf pure \: } (\lambda x. x) \star \phi(M) \rightarrow_{\beta} \phi(M)$ by $\beta$-reduction rule.

\vspace{\baselineskip}

2) Let ${\bf pure \: } (\lambda f g x. f (g x)) \star M \star N \star P, M \star (N \star P) \in
\Lambda^{'}_{\textbf{K}}$ and ${\bf pure \: } (\lambda f g x. f (g x)) \star M \star N \star P
\rightarrow_{\beta^{'}} M \star (N \star P)$.

By the definition of $\phi$:

$\phi({\bf pure \: } (\lambda f g x. f (g x)) \star M \star N \star P) = {\bf pure \: } (\lambda f g x. f (g
x)) \star \phi(M) \star \phi(N) \star \phi(P)$;

$M \star (N \star P) = \phi(M) \star (\phi(N) \star \phi(P))$.

Hence, ${\bf pure \: } (\lambda f g x. f (g x))
\star \phi(M) \star \phi(N) \star \phi(P) \rightarrow_{\beta} \phi(M) \star (\phi(N) \star \phi(P))$ by
$\beta$-reduction rule.

\vspace{\baselineskip}

3) Let $({\bf pure \: } M) \star ({\bf pure \: } N), {\bf pure \: } (M N) \in \Lambda^{'}_{\textbf{K}}$ and
$({\bf pure \: } M) \star ({\bf pure \: } N) \rightarrow_{\beta} {\bf pure \: } (M N)$.

By the definition of $\phi$:

$\phi(({\bf pure \: } M) \star ({\bf pure \: } N)) = ({\bf pure \: } \phi(M)) \star ({\bf pure \: }
\phi(N))$;

$\phi({\bf pure \: } (M N)) = {\bf pure \: } (\phi(M) \phi(N))$.

So, by reduction rule, $({\bf pure \: } \phi(M)) \star ({\bf pure \: } \phi(N)) \rightarrow_{\beta} {\bf
pure \: } (\phi(M) \phi(N))$.

\vspace{\baselineskip}

4) Let $M \star ({\bf pure \: }), {\bf pure \: } (\lambda f. f N) \star M$ and $M \star ({\bf pure \: } N)
\rightarrow_{\beta^{'}} (\lambda f. f N) \star M$.

$\phi(M \star ({\bf pure \: } N)) = \phi(M) \star ({\bf pure \: } \phi(N))$

$\phi((\lambda f. f N) \star M) = (\lambda f. f \phi(N)) \star \phi(M)$.

So, $\phi(M) \star ({\bf pure \: } \phi(N)) \rightarrow_{\beta} {\bf pure \: } (\lambda f. f \phi(N)) \star
\phi(M)$.
\end{proof}

\begin{lemma}

$ $

Let $M \in \Lambda^{'}_{\textbf{K}}$. Then $|M| \twoheadrightarrow_{\beta} \phi(M)$.

\end{lemma}

\begin{proof}

Induction on the structure of $M$.

\end{proof}

\begin{lemma} Strip lemma.

$ $

If $M \rightarrow_{\beta} N$ and $M \twoheadrightarrow_{\beta} P$. Then there exists some term $Q$, such that
$N \twoheadrightarrow_{\beta} Q$ and $P \twoheadrightarrow_{\beta} Q$.

\end{lemma}

\begin{proof}

  Proof is similar to \cite{Baren} \cite{Baren2}. We build the following diagram

\vspace{\baselineskip}

\xymatrix{
&&& M \ar[dd]_{\beta} \ar@{->>}[rrr] &&& P \ar@{-->>}[dd]^{\beta} \\
&&&& N' \ar[ul]^{|.|} \ar[dl]_{\phi} \ar@{-->>}[rrr]^{\beta^{'}} &&& P' \ar[ul]^{|.|} \ar[dl]_{\phi} \\
&&& N \ar@{-->>}[rrr]_{\beta} &&& Q
}

\end{proof}

which is commutes by lemmas 11 -- 14.

\begin{theorem} Confluence.

$ $

If $M \twoheadrightarrow_{\beta} N$ and $M \twoheadrightarrow_{\beta} P$. Then there exists some term $Q$,
such that $N \twoheadrightarrow_{\beta} Q$ and $P \twoheadrightarrow_{\beta} Q$.

\end{theorem}

\begin{proof}

$ $

  By unfolding $M \twoheadrightarrow_{\beta} N$ as the sequence of one-step reductions $M \rightarrow_{\beta}
  M_1 \rightarrow_{\beta} ... \rightarrow_{\beta} M_n \rightarrow_{\beta} N$ and applying strip lemma on every step.

\end{proof}

  \section{Acknowledgement.}

  Author would like to thank his supervisior
  V.L.Vasukov, V.N. Krupski for general idea and wise
  advice, V. de Paiva, V.I. Shalack, A.V. Rodin and M.
  Taldykin for discussing, critics and consulting.

  \addcontentsline{toc}{section}{References}

  %далее сам список используевой литературы
  \begin{thebibliography}{}

  \bibitem{Artemov} Artemov S. and Protopopescu T., \/ ``Intuitionistic Epistemic Logic'', \textit{The
  Review of Symbolic Logic}, 2016, vol. 9, no 2. pp. 266-298.\parskip=1mm

  \bibitem{Krupski} Krupski V. N. and Yatmanov A., \/ ``Sequent Calculus for Intuitionistic Epistemic Logic
  IEL'', \textit{Logical Foundations of Computer Science: International Symposium, LFCS 2016, Deerfield
  Beach, FL, USA, January 4-7, 2016. Proceedings}, 2016, pp. 187-201.\parskip=1mm

  \bibitem{Haskell} Haskell Language. // URL: https://www.haskell.org. (Date: 1.08.2017) \parskip=1mm

  \bibitem{Idris} Idris. A Language with Dependent Types.// URL:https://www.idris-lang.org. (Date:
  1.08.2017) \parskip=1mm

  \bibitem{Purs} Purescript. A strongly-typed functional programming language that compiles to JavaScript.
  URL: http://www.purescript.org. (Date: 1.08.2017) \parskip=1mm

  \bibitem{Elm} Elm. A delightful language for reliable webapps. // URL: http://elm-lang.org. (Date:
  1.08.2017) \parskip=1mm

  \bibitem{Base} Hackage, \/ ``The base package'' // URL: https://hackage.haskell.org/package/base-4.10.0.0
  (Date: 1.08.2017) \parskip=1mm

  \bibitem{Miran} Lipovaca M, \/ ``Learn you a Haskell for Great Good!''. //URL:
  http://learnyouahaskell.com/chapters (Date: 1.08.2017) \parskip=1mm

  \bibitem{McP} McBride C. and Paterson R., ``Applicative programming with effects", \textit{Journal of
  Functional Programming}, 2008, vol. 18, no 01. pp 1-13. \parskip=1mm

  \bibitem{McP2} McBride C. and Paterson R, ``Functional Pearl. Idioms: applicative programming with
  effects'', \textit{Journal of Functional Programming}, 2005. vol. 18, no 01. pp 1-20. \parskip=1mm

  \bibitem{Neder} R. Nederpelt and H. Geuvers, ``Type Theory and Formal Proof: An Introduction''.
  \textit{Cambridge University Press}, New York, NY, USA, 2014. pp. 436. \parskip=1mm

  \bibitem{Morten} Sorensen M. H. and Urzyczyn P, ``Lectures on the Curry-Howard isomorphism'',
  \textit{Studies in Logic and the Foundations of Mathematics}, vol. 149, \textit{Elsevier Science}, 1998.
  pp 261. \parskip=1mm

  \bibitem{Pierce} Pierce B. C., ``Types and Programming Languages''. \textit{Cambridge, Mass: The MIT
  Press}, 2002. pp. 605. \parskip=1mm

  \bibitem{Girard} Girard J.-Y., Taylor P. and  Lafont Y, ``Proofs and Types'', \textit{Cambridge University
  Press}, New York, NY, USA, 1989. pp. 175. \parskip=1mm

  \bibitem{Baren} Barendregt. H. P., ``Lambda calculi with types" // Abramsky S., Gabbay Dov M., and S. E.
  Maibaum, ``Handbook of logic in computer science (vol. 2), Osborne Handbooks Of Logic In Computer
  Science'', Vol. 2. \textit{Oxford University Press, Inc.}, New York, NY, USA, 1993. pp 117-309.
  \parskip=1mm

  \bibitem{Hindley} Hindley J. Roger, ``Basic Simple Type Theory''. \textit{Cambridge University Press}, New
  York, NY, USA, 1997. pp. 185. \parskip=1mm

  \bibitem{Lax} Pfenning F. and Davies R., ``A judgmental reconstruction of modal logic'',
  \textit{Mathematical Structures in Computer Science}, vol. 11, no 4, 2001, pp. 511-540. \parskip=1mm

  \bibitem{Baren2} H.P. Barendregt. The Lambda Calculus --- Its Syntax and Semantics. Studies in Logic and
  the Foundations of Mathematics, vol. 103. Amsterdam: North-Holland, 1985.
  \end{thebibliography}

\end{document}

### 2018-01-13 article.tex
start to rewrite

@@
   \RightLabel{\scriptsize{ax}}
-  \UnaryInfC{$\Gamma , \alpha \vdash \alpha$}
+  \UnaryInfC{$\Gamma , \alpha \vdash A$}
   \end{prooftree}
@@
   	\begin{prooftree}
-  	\AxiomC{$\Gamma, \alpha \vdash \beta$}
+  	\AxiomC{$\Gamma, A \vdash B$}
   	\RightLabel{$\rightarrow_i$}
-  	\UnaryInfC{$\Gamma \vdash \alpha \to \beta$}
-    	\end{prooftree}
-
-  	\begin{prooftree}
-  	\AxiomC{ $\Gamma \vdash \alpha$ }
-  	\AxiomC{ $\Gamma \vdash \beta$ }
+  	\UnaryInfC{$\Gamma \vdash A \to B$}
+    	\end{prooftree}
+
+  	\begin{prooftree}
+  	\AxiomC{ $\Gamma \vdash A$ }
+  	\AxiomC{ $\Gamma \vdash B$ }
   	\RightLabel{$\land_i$}
-  	\BinaryInfC{$\Gamma \vdash \alpha \land \beta$}
-  	\end{prooftree}
-
-  	\begin{prooftree}
-    	\AxiomC{$\Gamma \vdash \alpha$}
+  	\BinaryInfC{$\Gamma \vdash A \land B$}
+  	\end{prooftree}
+
+  	\begin{prooftree}
+    	\AxiomC{$\Gamma \vdash A$}
     	\RightLabel{$\textbf{K}_I$}
-    	\UnaryInfC{$\Gamma \vdash \textbf{K}\alpha$}
+    	\UnaryInfC{$\Gamma \vdash \textbf{K}A$}
   \end{prooftree}
@@
     	\begin{prooftree}
-  	\AxiomC{$\Gamma \vdash \alpha \to \beta$}
-  	\AxiomC{$\Gamma \vdash \alpha$}
+  	\AxiomC{$\Gamma \vdash A \to B$}
+  	\AxiomC{$\Gamma \vdash A$}
   	\RightLabel{$\rightarrow_i$}
-  	\BinaryInfC{$\Gamma \vdash \beta$}
-    	\end{prooftree}
-
-  	\begin{prooftree}
-  	\AxiomC{ $\Gamma \vdash \alpha_1 \land \alpha_2$ }
+  	\BinaryInfC{$\Gamma \vdash B$}
+    	\end{prooftree}
+
+  	\begin{prooftree}
+  	\AxiomC{ $\Gamma \vdash A_1 \land A_2$ }
   	\RightLabel{$\land_e, i \in \{ 1, 2 \}$}
-  	\UnaryInfC{$\Gamma \vdash \alpha_i$}
-  	\end{prooftree}
-
-  	\begin{prooftree}
-    	\AxiomC{$\Gamma \vdash \textbf{K}(\alpha \to \beta)$}
-    	\AxiomC{$\Gamma \vdash \textbf{K}\alpha$}
-    	\RightLabel{$\textbf{K}_{MP}$}
-          \BinaryInfC{$\Gamma \vdash \textbf{K}\beta$}
+  	\UnaryInfC{$\Gamma \vdash A_i$}
+  	\end{prooftree}
+
+  	\begin{prooftree}
+    \AxiomC{$\Gamma \vdash {\bf K} \vec{A}$}
+    \AxiomC{$\vec{A} \vdash B$}
+    \BinaryInfC{$\Gamma \vdash {\bf K} B$}
   	\end{prooftree}
@@
 
+Where $\Gamma \vdash {\bf K} \vec{A}$ is a syntax sugar for $\Gamma \vdash {\bf K} A_1 ,\dots, \Gamma \vdash {\bf K} A_n$.
+
   \vspace{\baselineskip}
@@
   \begin{lemma}
-    NIEL$^{-}_{\land, \to} \vdash \alpha \Rightarrow$ IEL$^{-}$ $\vdash \alpha$.
+    $\Gamma \vdash_{\text{NIEL}^{-}_{\land, \to}} A \Rightarrow$ IEL$^{-} \vdash \bigwedge \Gamma \rightarrow A$.
   \end{lemma}
@@
 
-1) If $\Gamma \vdash_{NIEL^{-}_{\land, \to}} \alpha$, then $IEL^{-} \vdash \wedge \Gamma \rightarrow \textbf{K}\alpha$.
+1) If $\Gamma \vdash_{NIEL^{-}_{\land, \to}} A$, then $IEL^{-} \vdash \bigwedge \Gamma \rightarrow \textbf{K}A$.
 
 $\begin{array}{lll}
-(1)& \wedge \Gamma \rightarrow \alpha & \text{assumption}\\
-(2)& \alpha \rightarrow \textbf{K}\alpha &\text{co-reflection.}\\
-(3)& (\wedge \Gamma \rightarrow \alpha) \rightarrow ((\alpha \rightarrow \textbf{K}\alpha) \rightarrow (\wedge \Gamma \rightarrow \textbf{K}\alpha))&\text{IPC theorem}\\
-(4) &(\alpha \rightarrow \textbf{K}\alpha) \rightarrow (\wedge \Gamma \rightarrow \textbf{K}\alpha) &\text{1, 3, MP}\\
-(5) &\wedge \Gamma \rightarrow \textbf{K}\alpha &\text{2,4, MP}\\
+(1) & \bigwedge \Gamma \rightarrow A & \text{assumption}\\
+(2) & A \rightarrow \textbf{K}A &\text{co-reflection}\\
+(3) & (\bigwedge \Gamma \rightarrow A) \rightarrow ((A \rightarrow \textbf{K}A) \rightarrow (\bigwedge \Gamma \rightarrow \textbf{K}A))&\text{IPC theorem}\\
+(4) & (A \rightarrow \textbf{K}A) \rightarrow (\bigwedge \Gamma \rightarrow \textbf{K}A) &\text{from (1), (3) and MP}\\
+(5) & \bigwedge \Gamma \rightarrow \textbf{K}A &\text{from (2), (4) and MP}\\
 \end{array}$
@@
 
-2) If $\Gamma \vdash_{NIEL^{-}_{\land, \to}} \textbf{K}(\alpha \to \beta)$ and $\Gamma \vdash_{NIEL^{-}_{\land, \to}} \textbf{K}\alpha$, then $IEL^{-} \vdash \wedge \Gamma \rightarrow \textbf{K}\beta$.
+2) If $\Gamma \vdash_{NIEL^{-}_{\land, \to}} {\bf K} \vec{A}$ and $\vec{A} \vdash B$, then $IEL^{-} \vdash \bigwedge \Gamma \rightarrow \textbf{K}B$.
 
 $\begin{array}{lll}
-(1)& \Gamma \vdash_{NIEL^{-}_{\land, \to}} \textbf{K}(\alpha \to \beta) & \text{condition}\\
-(2)& \Gamma \vdash_{NIEL^{-}_{\land, \to}} \textbf{K}\alpha & \text{condition}\\
-(3)& \Gamma \vdash_{NIEL^{-}_{\land, \to}} \textbf{K}(\alpha \to \beta) \land \textbf{K}\alpha & \land_i \\
-(4)& \wedge \Gamma \rightarrow \textbf{K}(\alpha \to \beta) \land \textbf{K}\alpha & \text{induction
-hypothesis} \\
-(5)& (\textbf{K}(\alpha \rightarrow \beta) \rightarrow (\textbf{K}\alpha \to \textbf{K}\beta)) \rightarrow
-((\textbf{K}(\alpha \rightarrow \beta) \land \textbf{K}\alpha) \rightarrow \textbf{K}\beta)& \text{IPC
-theorem} \\
-(6)&\textbf{K}(\alpha \rightarrow \beta) \rightarrow (\textbf{K}\alpha \to \textbf{K}\beta)&
-\text{normality}\\
-(7)&(\textbf{K}(\alpha \rightarrow \beta) \land \textbf{K}\alpha) \rightarrow \textbf{K}\beta)& \text{5,6,
-MP}\\
-(8)&(\wedge \Gamma \rightarrow \textbf{K}(\alpha \to \beta) \land \textbf{K}\alpha) \rightarrow & \\
-&\quad\rightarrow ((\textbf{K}(\alpha \rightarrow \beta) \land \textbf{K}\alpha) \rightarrow
-\textbf{K}\beta) \rightarrow
-(\wedge \Gamma \rightarrow \textbf{K}\beta))& \text{IPC theorem} \\
-(9) &(\textbf{K}(\alpha \rightarrow \beta) \land \textbf{K}\alpha) \rightarrow \textbf{K}\beta) \rightarrow
-(\wedge \Gamma \rightarrow \textbf{K}\beta)& \text{4, 8, MP} \\
-(10) &\wedge \Gamma \rightarrow \textbf{K}\beta& \text{7, 9, MP}
+(1) &\bigwedge \Gamma \rightarrow \bigwedge \limits_{i = 1}^{n} {\bf K} A_i & \text{assumption} \\
+(2) &\bigwedge \limits_{i = 1}^{n} {\bf K} A_i \rightarrow {\bf K} \bigwedge \limits_{i = 1}^{n} A_i& \text{IEL theorem} \\
+(3) &\bigwedge \Gamma \rightarrow {\bf K} \bigwedge \limits_{i = 1}^{n} A_i & \text{from (1), (2) and transitivity} \\
+(4) &\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B& \text{assumption} \\
+(5) &(\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B) \rightarrow {\bf K} (\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B)& \text{co-reflection}\\
+(6) &{\bf K} (\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B)& \text{from (2), (3) and MP} \\
+(7) &{\bf K} \bigwedge \limits_{i = 1}^{n} A_i \rightarrow {\bf K} B & \text{from (6) and normality} \\
+(8) &\bigwedge \Gamma \rightarrow {\bf K} B & \text{from (3), (7) and transitivity}
 \end{array}$
@@
   Let $\mathbb{V}$ be the set of variables. The set $\Lambda_{\textbf{K}}$ of terms is defined by the grammar:
-  \begin{equation}
-    \Lambda_{\textbf{K}} ::= \mathbb{V} \: | \:  (\lambda \Lambda.\Lambda_{\textbf{K}}) \: | \:
-    (\Lambda_{\textbf{K}}\Lambda_{\textbf{K}}) \: | \: (\Lambda_{\textbf{K}} , \Lambda_{\textbf{K}}) \: | \:
-    (\pi_i \Lambda_{\textbf{K}}) \:
-    | \: ({\bf pure \: } \: \Lambda_{\textbf{K}}) \: | \: (\Lambda_{\textbf{K}} \star
-    \Lambda_{\textbf{K}})
-  \end{equation}
-  where $i \in \{ 1, 2 \}$.
+
+$\begin{array}{lll}
+& \Lambda_{\textbf{K}} ::= \mathbb{V} \: | \:  (\lambda \Lambda.\Lambda_{\textbf{K}}) \: | \: (\Lambda_{\textbf{K}}\Lambda_{\textbf{K}}) \: | \: (\Lambda_{\textbf{K}} , \Lambda_{\textbf{K}}) \: | \: (\pi_1 \Lambda_{\textbf{K}}) \: | \: (\pi_2 \Lambda_{\textbf{K}}) \: | & \\
+& \quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad ({\bf pure \: } \: \Lambda_{\textbf{K}}) \: | \: ({\bf let \:\: pure} \:\: \Lambda_{\textbf{K}} = \Lambda_{\textbf{K}} \:\: {\bf in} \:\: \Lambda_{\textbf{K}})
+\end{array}$
+
   \end{defin}
@@
   \begin{prooftree}
-    \AxiomC{$\Gamma \vdash f : \textbf{K}(\alpha \to \beta)$}
-    \AxiomC{$\Gamma \vdash x : \textbf{K}\alpha$}
-    \RightLabel{$\textbf{K}_{app}$}
-    \BinaryInfC{$\Gamma \vdash f \star x : \textbf{K}\beta$}
+    \AxiomC{$\Gamma \vdash \vec{x} : {\bf K} \vec{A}$}
+    \AxiomC{$\vec{y} : \vec{A} \vdash M : B$}
+    \BinaryInfC{$\Gamma \vdash {\bf let \:\: pure} \:\: \vec{y} = \vec{x} \:\: {\bf in} \:\: M : {\bf K} B$}
   \end{prooftree}
@@
 
-  $\textbf{K}_{app}$ is a typing rule for lifted application. This rule is similar to
-  $\rightarrow_e$-rule, therefore we introduce the infix operator $\star$ for lifted application.
-
-  \begin{con}
-
-  \emph{$\star$ is a left-associative, so $M \star N \star P = (M \star N) \star P$,  unless otherwise specified.}
-
-  \end{con}
-
   \vspace{\baselineskip}
@@
   \begin{prooftree}
-  \AxiomC{$x : \alpha \vdash x : \alpha$}
+  \AxiomC{$x : A \vdash x : A$}
   \RightLabel{$\textbf{K}_I$}
-  \UnaryInfC{$x : \alpha \vdash {\bf pure \: } x : \textbf{K} \alpha$}
+  \UnaryInfC{$x : A \vdash {\bf pure \: } x : \textbf{K} A$}
   \RightLabel{$\rightarrow_i$}
-  \UnaryInfC{$\vdash (\lambda x. {\bf pure \: } x) : \alpha \to \textbf{K}\alpha$}
-  \end{prooftree}
-
-  \begin{prooftree}
-    \AxiomC{$f : \textbf{K}(\alpha \to \beta) \vdash f : \textbf{K}(\alpha \to \beta)$}
-    \AxiomC{$x : \textbf{K}\alpha \vdash x : \textbf{K}\alpha$}
-    \RightLabel{$\textbf{K}_{app}$}
-    \BinaryInfC{$f : \textbf{K}(\alpha \to \beta), x : \textbf{K}\alpha \vdash f \star x :
-    \textbf{K}\beta$}
-    \RightLabel{$\rightarrow_i$}
-    \UnaryInfC{$f : \textbf{K}(\alpha \to \beta) \vdash \lambda x. f \star x : \textbf{K}\alpha \to \textbf{K}\beta$}
-    \RightLabel{$\rightarrow_i$}
-    \UnaryInfC{$\vdash \lambda f. \lambda x. f \star x: \textbf{K}(\alpha \to \beta) \to
-    \textbf{K}\alpha \to \textbf{K}\beta$}
+  \UnaryInfC{$\vdash (\lambda x. {\bf pure \: } x) : A \to \textbf{K}A$}
+  \end{prooftree}
+
+  \begin{prooftree}
+    \AxiomC{$f : {\bf K} (A \to B) \vdash f : {\bf K} (A \to B)$}
+    \AxiomC{$x : {\bf K} A \vdash x : {\bf K} A $}
+    \AxiomC{$g : A \to B$}
+    \AxiomC{$y : A$}
+    \BinaryInfC{$g : A \to B, y : A \vdash g y : B$}
+    \TrinaryInfC{$f : {\bf K} (A \to B), x : {\bf K} A \vdash {\bf let \:\: pure} \:\: \langle g, y \rangle = \langle f, x \rangle \:\: {\bf in } \:\: g y : {\bf K} B$}
+    \UnaryInfC{$f : {\bf K} (A \to B) \vdash \lambda x. {\bf let \:\: pure} \:\: \langle g, y \rangle = \langle f, x \rangle \:\: {\bf in } \:\: g y : {\bf K} A \to {\bf K} B$}
+    \UnaryInfC{$\vdash \lambda f. \lambda x. {\bf let \:\: pure} \:\: \langle g, y \rangle = \langle f, x \rangle \:\: {\bf in } \:\: g y : {\bf K}(A \to B) \to {\bf K} A \to {\bf K} B$}
+  \end{prooftree}
+
+\begin{prooftree}
+  \AxiomC{$f : A \to B \vdash f : A \to B$}
+  \UnaryInfC{$f : A \to B \vdash {\bf pure \:\:} f : {\bf K }(A \to B)$}
+  \AxiomC{$x : {\bf K} A \vdash x : {\bf K} A $}
+  \AxiomC{$g : A \to B$}
+  \AxiomC{$y : A$}
+  \BinaryInfC{$g : A \to B, y : A \vdash g y : B$}
+  \TrinaryInfC{$f : A \to B, x : {\bf K} A \vdash {\bf let \:\: pure} \:\: \langle g, y \rangle = \langle {\bf pure \:\:} f, x \rangle \:\: {\bf in } \:\: g y : {\bf K} B$}
+  \UnaryInfC{$f : A \to B \vdash \lambda x. {\bf let \:\: pure} \:\: \langle g, y \rangle = \langle {\bf pure \:\:} f, x \rangle \:\: {\bf in } \:\: g y : {\bf K} A \to {\bf K} B$}
+  \UnaryInfC{$\lambda f. \lambda x. {\bf let \:\: pure} \:\: \langle g, y \rangle = \langle {\bf pure \:\:} f, x \rangle \:\: {\bf in } \:\: g y : (A \to B) \to {\bf K} A \to {\bf K} B$}
 \end{prooftree}

### 2018-01-13 article.tex
added details

@@
   \begin{prooftree}
-    \AxiomC{$\Gamma \vdash \vec{x} : {\bf K} \vec{A}$}
-    \AxiomC{$\vec{y} : \vec{A} \vdash M : B$}
-    \BinaryInfC{$\Gamma \vdash {\bf let \:\: pure} \:\: \vec{y} = \vec{x} \:\: {\bf in} \:\: M : {\bf K} B$}
+    \AxiomC{$\Gamma \vdash \vec{M} : {\bf K} \vec{A}$}
+    \AxiomC{$\vec{N} : \vec{A} \vdash M : B$}
+    \BinaryInfC{$\Gamma \vdash {\bf let \:\: pure} \:\: \vec{N} = \vec{M} \:\: {\bf in} \:\: M : {\bf K} B$}
   \end{prooftree}
@@
 
-  7) $FV(M \star N) = FV(M) \cup FV(N)$.
+  7) $FV({\bf let \:\: pure} \:\: \vec{N} = \vec{M} \:\: {\bf in} \:\: M) = \bigcup \limits_{i = 1}^n FV(M), \text{where $n = |\vec{M}|$}$.
   \end{defin}
@@
 
-  7) $(M \star N) [x := P] = (M [x := P]) \star (N [x := P])$.
+  7) $({\bf let \:\: pure} \:\: \vec{N} = \vec{M} \:\: {\bf in} \:\: M) [x := P] = {\bf let \:\: pure} \:\: \vec{N} = (\vec{M} [x := P]) \:\: {\bf in} \:\: M$.
   \end{defin}
@@
 
-  1) ${\bf pure \: } (\lambda x. x) \star M \rightarrow_{\beta} M$;
-
-  2) ${\bf pure \: } (\lambda f g x. f (g x)) \star M \star N \star P \rightarrow_{\beta} M \star (N \star
-  P)$;
-
-  3) $({\bf pure \: } M) \star (\text{pure } N) \rightarrow_{\beta} \text{pure } (M N)$;
-
-  4) $M \star {\bf pure \: } N \rightarrow_{\beta} (\lambda f. f N) \star M$;
-
-  5) $M_1 \rightarrow_{\beta} M_2 \Leftrightarrow {\bf pure \: } \: M_1 \rightarrow_{\beta} {\bf pure \: } \: M_2$;
-
-  6) $M_1 \rightarrow_{\beta} M_2 \Rightarrow M_1 \star N \rightarrow_{\beta} M_2 \star N$;
-
-  7) $N_1 \rightarrow_{\beta} N_2 \Rightarrow M \star N_1 \rightarrow_{\beta} M \star N_2$;
-  \end{defin}
-
-  In Haskell any implementation of the Applicative instance should preserve the laws proposed by McBride and
-  Paterson \cite{McP}. We preserve these laws by the introduction of corresponding reduction rules (rules
-  (1)-(4)). These rules will be called respectively identity rule, composition rule, homomorphism rule and
-  interchange rule.
-
-  Rules (5)-(7) are would be useful for strong normalization proof.
+  \end{defin}
 
@@
 
-  2) $\Gamma \vdash M \star N : \textbf{K}\beta$ implies that $\Gamma \vdash M : \textbf{K}(\alpha \to \beta)$ and
-  $\Gamma \vdash N : \textbf{K}\alpha$.
-  \end{lemma}
-
-  \begin{proof}
-
-  $ $
-
-  Simple induction on the derivation of $\Gamma \vdash {\bf pure \: } M : \textbf{K}\alpha$ and $\Gamma \vdash M \star N : \textbf{K}(\alpha \to \beta)$ respectively.
+  2) $\Gamma \vdash {\bf let \: pure} \: \vec{N} = \vec{M} \: {\bf in} \: M : {\bf K} B$ implies that $\Gamma \vdash \vec{M} : {\bf K} \vec{A}$ and
+$\vec{N} : \vec{A} \vdash M : B$.
+  \end{lemma}
+
+  \begin{proof}
+
+  $ $
+
+Induction on the derivation of $\Gamma \vdash {\bf pure \: } M : \textbf{K}\alpha$ and $\Gamma \vdash {\bf let \: pure} \: \vec{N} = \vec{M} \: {\bf in} \: M : {\bf K} B$ respectively.
   \end{proof}

### 2018-01-13 article.tex
added some fixes and rewrite

@@
 
+  \begin{defin} Intuitionistic normal modal logic {\bf IK}
+
+    1) IPC axioms;
+
+    2) $\Box (A \to B) \to (\Box A \to \Box B)$;
+
+    3) Rules: MP and necessiation.
+
+  \end{defin}
+
+  \begin{defin} Translation from {\bf IK} into {\bf IEL}$^{-}$
+
+    1) $\bot^{\circ} = \bot$;
+
+    2) $p^{\circ} = p$;
+
+    3) $(A \alpha B)^{\circ} = A^{\circ} \alpha B^{\circ}$, $\alpha \in \{ \to, \wedge, \vee \}$.
+
+    4) $(\Box A)^{\circ} = {\bf K} A^{\circ}$
+  \end{defin}
+
+  \begin{lemma}
+    $ $
+    ${\bf IK} \vdash A \Rightarrow {\bf IEL}^{-} \vdash A^{\circ}$
+  \end{lemma}
+
+  \begin{proof}
+    Straightforward induction on the structure of $M$.
+  \end{proof}
+
+  It is clearly that we may prove similar fact for ${\bf IEL}$ by the same way.
+
   At first we define the natural deduction for IEL$^{-}$ with $\textbf{K}$-modality and binary connectives
@@
   \RightLabel{\scriptsize{ax}}
-  \UnaryInfC{$\Gamma , x : \alpha \vdash x : \alpha$}
+  \UnaryInfC{$\Gamma , x : A \vdash x : A$}
   \end{prooftree}
@@
   \begin{prooftree}
-  \AxiomC{$\Gamma, x : \alpha \vdash M : \beta$}
+  \AxiomC{$\Gamma, x : A \vdash M : B$}
   \RightLabel{$\rightarrow_i$}
-  \UnaryInfC{$\Gamma \vdash \lambda x. M : \alpha \to \beta$}
-  \end{prooftree}
-
-  \begin{prooftree}
-  \AxiomC{ $\Gamma \vdash x : \alpha$ }
-  \AxiomC{ $\Gamma \vdash y : \beta$ }
+  \UnaryInfC{$\Gamma \vdash \lambda x. M : A \to B$}
+  \end{prooftree}
+
+  \begin{prooftree}
+  \AxiomC{ $\Gamma \vdash x : A$ }
+  \AxiomC{ $\Gamma \vdash y : B$ }
   \RightLabel{$\times_i$}
-  \BinaryInfC{$\Gamma \vdash (x, y) : \alpha \times \beta$}
-  \end{prooftree}
-
-  \begin{prooftree}
-    \AxiomC{$\Gamma \vdash x : \alpha$}
+  \BinaryInfC{$\Gamma \vdash \langle x, y \rangle : A \times B$}
+  \end{prooftree}
+
+  \begin{prooftree}
+    \AxiomC{$\Gamma \vdash x : A$}
     \RightLabel{$\textbf{K}_I$}
-    \UnaryInfC{$\Gamma \vdash {\bf pure \: } \: x : \textbf{K}\alpha $}
+    \UnaryInfC{$\Gamma \vdash {\bf pure \: } \: x : \textbf{K}A $}
   \end{prooftree}
@@
   \begin{prooftree}
-  \AxiomC{$\Gamma \vdash f : \alpha \to \beta$}
-  \AxiomC{$\Gamma \vdash x : \alpha$}
+  \AxiomC{$\Gamma \vdash f : A \to B$}
+  \AxiomC{$\Gamma \vdash x : A$}
   \RightLabel{$\rightarrow_e$}
-  \BinaryInfC{$\Gamma \vdash fx : \beta$}
-  \end{prooftree}
-
-  \begin{prooftree}
-  \AxiomC{ $\Gamma \vdash p : \alpha_1 \times \alpha_2$ }
+  \BinaryInfC{$\Gamma \vdash fx : B$}
+  \end{prooftree}
+
+  \begin{prooftree}
+  \AxiomC{ $\Gamma \vdash M : A_1 \times A_2$ }
   \RightLabel{$\times_e$, $i \in \{ 1, 2 \}$}
-  \UnaryInfC{$\Gamma \vdash \pi_i p : \alpha_i$}
+  \UnaryInfC{$\Gamma \vdash \pi_i M : A_i$}
   \end{prooftree}
@@
     \AxiomC{$\Gamma \vdash \vec{M} : {\bf K} \vec{A}$}
-    \AxiomC{$\vec{N} : \vec{A} \vdash M : B$}
-    \BinaryInfC{$\Gamma \vdash {\bf let \:\: pure} \:\: \vec{N} = \vec{M} \:\: {\bf in} \:\: M : {\bf K} B$}
+    \AxiomC{$\vec{x} : \vec{A} \vdash N : B$}
+    \RightLabel{${\bf let}_{\textbf{K}}$}
+    \BinaryInfC{$\Gamma \vdash {\bf let \:\: pure} \:\: \vec{N} = \vec{M} \:\: {\bf in} \:\: N : {\bf K} B$}
   \end{prooftree}
@@
 
-  7) $FV({\bf let \:\: pure} \:\: \vec{N} = \vec{M} \:\: {\bf in} \:\: M) = \bigcup \limits_{i = 1}^n FV(M), \text{where $n = |\vec{M}|$}$.
+  7) $FV({\bf let \:\: pure} \:\: \vec{x} = \vec{M} \:\: {\bf in} \:\: N) = \bigcup \limits_{i = 1}^n FV(M_i), \text{where $n = |\vec{M}|$}$.
   \end{defin}
@@
 
-  In $\lambda \textbf{K}$ we have the following computational rules. We will define them for terms with
-  ${\bf pure \: }$ or $\star$. Reduction rules for $(,)$ and $\pi_i$ are described, for example, in
-  \cite{Pierce}.
+  In $\lambda \textbf{K}$ we have the following computational rules:
 
   \begin{defin} $\beta$-reduction rules for \emph{$\lambda \textbf{K}$}.
+
+$\begin{array}{lll}
+&\text{1) } (\lambda x. M) N \rightarrow_{\beta} M [x := N]& \\
+&\text{2) } \pi_1 \langle M, N \rangle \rightarrow_{\beta} M& \\
+&\text{3) } \pi_2 \langle M, N \rangle \rightarrow_{\beta} N& \\
+&\text{4) } {\bf let \: pure \: } \vec{y} = \vec{M_2} {\bf \: in \: } ({\bf let \: pure \: } \vec{x} = \vec{M_1} {\bf \: in \: }  N) \rightarrow_{\beta}& \\
+&\quad {\bf let \: pure \: } \vec{x} = ({\bf let \: pure \: } \vec{y} = \vec{M_2} \: {\bf in} \: \vec{M_1}) \: {\bf in} \: {\bf let \: pure \: } \vec{x} = \vec{x} {\bf \: in \:} N&
+\end{array}$
 
@@
 
-  Let $\Gamma = \{ x_1 : \alpha_1, ..., x_n : \alpha_n \}$. Then the domain of $\Gamma$, or $dom(\Gamma)$, is a set $\{  x_1, ..., x_n\}$.
+  Let $\Gamma = \{ x_1 : A_1, ..., x_n : A_n \}$. Then the domain of $\Gamma$, or $dom(\Gamma)$, is a set $\{  x_1, ..., x_n\}$.
 
@@
 
-  If $\Gamma \vdash M : \alpha$, then $FV(M) \subseteq dom(\Gamma)$
-
-  \end{lemma}
-
-  \begin{proof}
-
-  Induction on the derivation of $\Gamma \vdash M : \alpha$.
+  If $\Gamma \vdash M : A$, then $FV(M) \subseteq dom(\Gamma)$
+
+  \end{lemma}
+
+  \begin{proof}
+
+  Induction on the derivation of $\Gamma \vdash M : A$.
 
@@
 
-  1) $\Gamma \vdash {\bf pure \: } M : \textbf{K}\alpha$ implies that $\Gamma \vdash M : \alpha$;
+  1) $\Gamma \vdash {\bf pure \: } M : \textbf{K}A$ implies that $\Gamma \vdash M : A$;
 
@@
 
-  Let $\Gamma \vdash M : \alpha$ and $\Gamma \subseteq \Delta$, then $\Delta \vdash M : \alpha$.
+  Let $\Gamma \vdash M : A$ and $\Gamma \subseteq \Delta$, then $\Delta \vdash M : A$.
 
@@
 
-  1) Let $\Gamma \vdash x : \alpha$, such that $\Gamma = \Delta, x : \alpha$ and $\Theta \subseteq \Gamma$.
+  1) Let $\Gamma \vdash x : A$, such that $\Gamma = \Delta, x : A$ and $\Theta \subseteq \Gamma$.
   Let $\Sigma = \Theta \setminus \Gamma$, or, which is the same, $\Sigma = \Theta \setminus \Delta, x :
-  \alpha$, then $\Sigma,
-  \Delta, x : \alpha \vdash x : \alpha$, or, $\Theta \vdash x : \alpha$.
-
-  2) Let $\Gamma \vdash {\bf pure \: } M : \textbf{K}\alpha$ and $\Gamma \subseteq \Theta$.
-
-  If $\Gamma \vdash {\bf pure \: } M : \textbf{K}\alpha$, then $\Gamma \vdash M : \alpha$ by generation and, by
-  hypothesis, $\Theta \vdash M : \alpha$, so $\Theta \vdash {\bf pure \: } M : \textbf{K}\alpha$ by applying
+  A$, then $\Sigma, \Delta, x : A \vdash x : \alpha$, or, $\Theta \vdash x : A$.
+
+  2) Let $\Gamma \vdash {\bf pure \: } M : \textbf{K}A$ and $\Gamma \subseteq \Theta$.
+
+  If $\Gamma \vdash {\bf pure \: } M : \textbf{K}A$, then $\Gamma \vdash M : A$ by generation and, by
+  hypothesis, $\Theta \vdash M : A$, so $\Theta \vdash {\bf pure \: } M : \textbf{K}A$ by applying
   $\textbf{K}_I$-rule.
 
-  3) Let $\Gamma \vdash M \star N : \textbf{K}\beta$. So $\Gamma \vdash M : \textbf{K}(\alpha \to \beta)$ and $\Gamma
-  \vdash N : \textbf{K}\alpha$. By hypothesis $\Delta \vdash M : \textbf{K}(\alpha \to \beta)$ and $\Delta \vdash N :
-  \textbf{K}\alpha$. Then $\Delta \vdash M \star N : \textbf{K}\beta$.
+  3) Let $\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\bf \: in \:} N : \textbf{K}B$ and $\Gamma \subseteq \Delta$.
+
+  By generation $\Gamma \vdash \vec{M} : \textbf{K}\vec{A}$ and $\vec{x} : \vec{A} \vdash N : \textbf{K}B$.
+
+  By hypothesis we have $\Delta \vdash \vec{M} : \textbf{K}\vec{A}$. So $\Delta \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\bf \: in \:} N : \textbf{K}B$ by ${\bf let}_{\textbf{K}}$.
 
@@
 
-  If $\Gamma \vdash M : \alpha$, then $\Gamma \uparrow FV(M) \vdash M : \alpha$, where $\Gamma \uparrow
+  If $\Gamma \vdash M : A$, then $\Gamma \uparrow FV(M) \vdash M : A$, where $\Gamma \uparrow
   FV(M)$ is a subcontext of $\Gamma$, such that $dom(\Gamma \uparrow FV(M)) = dom(\Gamma) \cap FV(M)$.
@@
 
-  Induction by derivation. We consider the base of induction and the case with $\textbf{K}_{app}$. The rest
+  Induction by derivation. We consider the base of induction and the case with  ${\bf let}_{\textbf{K}}$. The rest
   cases are
@@
 
-  1) Let $\Gamma \vdash x : \alpha$, where $\Gamma = \Delta, x : \alpha$, $x \in \mathbb{V}$.
-
-  $FV(x) = \{ x\}$, then $dom(\Gamma) \cap \{ x\} = \{ x\}. $ So $(\Delta, x : \alpha) \uparrow FV(x) = \{ x
-  : \alpha\}$, then $x : \alpha \vdash x : \alpha$ by axiom.
-
-  2) Let $\Gamma \vdash M \star N : \textbf{K}\beta$.
-
-  By generation $\Gamma \vdash M : \textbf{K}(\alpha \to \beta)$ and $\Gamma \vdash N : \textbf{K}\alpha$.
-
-  By induction hypothesis $\Gamma \uparrow FV(M) \vdash M : \textbf{K}(\alpha \to \beta)$ and $\Gamma
-  \uparrow FV(N) \vdash N : \textbf{K}\beta$.
-
-  But $\Gamma \uparrow FV(M) \subseteq \Gamma \uparrow FV(M \star N)$ and $\Gamma
-  \uparrow FV(N) \subseteq \Gamma \uparrow FV(M \star N)$, so $\Gamma \uparrow FV(M \star N) \vdash M :
-  \textbf{K}(\alpha \to \beta)$ and $\Gamma \uparrow FV(M \star N) \vdash N : \textbf{K}\alpha$ by weakening.
-
-  Then $\Gamma \uparrow FV(M \star N) \vdash M \star N : \textbf{K}\beta$ by $\textbf{K}_{app}$.
+  1) Let $\Gamma \vdash x : A$, where $\Gamma = \Delta, x : A$, $x \in \mathbb{V}$.
+
+  $FV(x) = \{ x\}$, then $dom(\Gamma) \cap \{ x\} = \{ x\}. $ So $(\Delta, x : A) \uparrow FV(x) = \{ x
+  : A\}$, then $x : A \vdash x : A$ by axiom.
+
+  2) Let $\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\bf \: in \:} N : \textbf{K}B$.
+
+  By generation $\Gamma \vdash \vec{M} : \textbf{K}\vec{A}$ and $\vec{x} : \vec{A} \vdash N : \textbf{K}B$.
+
+  By hypothesis $\Gamma \uparrow FV(\vec{M}) \vdash \vec{M} : \textbf{K}\vec{A}$.
+
+  So $\Gamma \uparrow FV(\vec{M}) \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\bf \: in \:} N : \textbf{K}B$.
 
@@
 
-  If $\Gamma, x : \alpha \vdash M : \beta$  and $\Gamma \vdash N : \alpha$, then $\Gamma \vdash (M [x := N]) :
-  \beta $
-
-  \end{lemma}
-
-  \begin{proof}
-
-  $ $
-
-  1) Let $\Gamma, x : \alpha \vdash {\bf pure \: } M : \textbf{K}\beta$ and $\Gamma \vdash N : \alpha$.
-
-  If $\Gamma, x : \alpha \vdash {\bf pure \: } M : \textbf{K}\beta$, then, by generation, $\Gamma, x : A \vdash
-  M : \beta$. So, by induction hypothesis, $\Gamma \vdash (M [x := N]) : \beta $, then $\Gamma \vdash
-  {\bf pure \: } (M [x := N]) : \textbf{K}\beta$ by $\textbf{K}_I$, but ${\bf pure \: } (M [x := N]) =
-  ({\bf pure \: } M) (M [x := N])$
-  by substitution definition, so $\Gamma \vdash ({\bf pure \: } M) (M [x := N]) : \textbf{K}\beta$
-
-  2) Let $\Gamma, x : \gamma \vdash M \star N : \textbf{K}\beta$, and $\Gamma \vdash y : \gamma$.
-
-  So, by generation, $\Gamma, x : \gamma \vdash M : \textbf{K}(\alpha \to \beta)$ and $\Gamma, x : \gamma
-  \vdash N : \textbf{K}\alpha$.
-
-  Hence $\Gamma, x : \gamma \vdash (M [x := y]) : \textbf{K}(\alpha \to \beta)$ and
-  $\Gamma \vdash (N [x := y]): \textbf{K}\alpha$ by hypothesis.
-
-  So $\Gamma \vdash (M [x := y]) \star (N [x := y]) : \textbf{K}\beta$, or, $\Gamma \vdash (M \star N) ([x
-  := y]) : \textbf{K}\beta$.
+  If $\Gamma, x : A \vdash M : B$  and $\Gamma \vdash N : A$, then $\Gamma \vdash (M [x := N]) :
+  B $
+
+  \end{lemma}
+
+  \begin{proof}
+
+  $ $
+
+  1) Let $\Gamma, x : A \vdash {\bf pure \: } M : \textbf{K}B$ and $\Gamma \vdash N : A$.
+
+  If $\Gamma, x : \alpha \vdash {\bf pure \: } M : \textbf{K}\beta$.
+
+  By generation, $\Gamma, x : A \vdash M : B$.
+
+  So, by induction hypothesis, $\Gamma \vdash (M [x := N]) : B $.
+  Then $\Gamma \vdash {\bf pure \: } (M [x := N]) : \textbf{K}B$ by $\textbf{K}_I$, but ${\bf pure \: } (M [x := N]) =
+  ({\bf pure \: } M [x := N])$ by substitution definition.
+
+  So $\Gamma \vdash ({\bf pure \: } M [x := N]) : \textbf{K}B$
+
+  2) .
 

### 2018-01-14 article.tex
added rewrite

@@
 
-  \begin{defin} Intuitionistic normal modal logic {\bf IK}
-
-    1) IPC axioms;
-
-    2) $\Box (A \to B) \to (\Box A \to \Box B)$;
-
-    3) Rules: MP and necessiation.
-
-  \end{defin}
-
-  \begin{defin} Translation from {\bf IK} into {\bf IEL}$^{-}$
-
-    1) $\bot^{\circ} = \bot$;
-
-    2) $p^{\circ} = p$;
-
-    3) $(A \alpha B)^{\circ} = A^{\circ} \alpha B^{\circ}$, $\alpha \in \{ \to, \wedge, \vee \}$.
-
-    4) $(\Box A)^{\circ} = {\bf K} A^{\circ}$
-  \end{defin}
-
-  \begin{lemma}
-    $ $
-    ${\bf IK} \vdash A \Rightarrow {\bf IEL}^{-} \vdash A^{\circ}$
-  \end{lemma}
-
-  \begin{proof}
-    Straightforward induction on the structure of $M$.
-  \end{proof}
-
-  It is clearly that we may prove similar fact for ${\bf IEL}$ by the same way.
-
   At first we define the natural deduction for IEL$^{-}$ with $\textbf{K}$-modality and binary connectives
@@
   	\begin{prooftree}
-    	\AxiomC{$\Gamma \vdash A$}
-    	\RightLabel{$\textbf{K}_I$}
-    	\UnaryInfC{$\Gamma \vdash \textbf{K}A$}
+    	\AxiomC{$\Gamma, \Delta \vdash A$}
+    	\RightLabel{$\vdash {\bf K}_I$}
+    	\UnaryInfC{${\bf K}(\Gamma), \Delta \vdash {\bf K}A$}
   \end{prooftree}
@@
   	\begin{prooftree}
-    \AxiomC{$\Gamma \vdash {\bf K} \vec{A}$}
-    \AxiomC{$\vec{A} \vdash B$}
-    \BinaryInfC{$\Gamma \vdash {\bf K} B$}
+    \AxiomC{${\bf K}(\Gamma), \Delta \vdash {\bf K} A$}
+    \AxiomC{${\bf K}(\Gamma), {\bf K} A, \Delta \vdash B$}
+    \BinaryInfC{${\bf K}(\Gamma), \Delta \vdash B$}
   	\end{prooftree}
@@
 
-Where $\Gamma \vdash {\bf K} \vec{A}$ is a syntax sugar for $\Gamma \vdash {\bf K} A_1 ,\dots, \Gamma \vdash {\bf K} A_n$.
-
   \vspace{\baselineskip}
@@
 
-1) If $\Gamma \vdash_{NIEL^{-}_{\land, \to}} A$, then $IEL^{-} \vdash \bigwedge \Gamma \rightarrow \textbf{K}A$.
+1) If $\Gamma, \Delta \vdash_{NIEL^{-}_{\land, \to}} A$, then $IEL^{-} \vdash \bigwedge {\bf K} \Gamma \wedge \bigwedge \Delta \rightarrow {\bf K}A$.
+
+See \cite{Krupski}.
+
+\vspace{\baselineskip}
+
+2) If ${\bf K} \Gamma, \Delta \vdash_{NIEL^{-}_{\land, \to}} {\bf K} A$ and ${\bf K}\Gamma, {\bf K} A, \Delta \vdash B$, then $IEL^{-} \vdash \bigwedge {\bf K} \Gamma \wedge \bigwedge \Delta \rightarrow B$.
 
 $\begin{array}{lll}
-(1) & \bigwedge \Gamma \rightarrow A & \text{assumption}\\
-(2) & A \rightarrow \textbf{K}A &\text{co-reflection}\\
-(3) & (\bigwedge \Gamma \rightarrow A) \rightarrow ((A \rightarrow \textbf{K}A) \rightarrow (\bigwedge \Gamma \rightarrow \textbf{K}A))&\text{IPC theorem}\\
-(4) & (A \rightarrow \textbf{K}A) \rightarrow (\bigwedge \Gamma \rightarrow \textbf{K}A) &\text{from (1), (3) and MP}\\
-(5) & \bigwedge \Gamma \rightarrow \textbf{K}A &\text{from (2), (4) and MP}\\
+&\text{(1) } \bigwedge {\bf K} \Gamma \wedge \bigwedge \Delta \to {\bf K} A& \text{assumption} \\
+&\text{(2) } \bigwedge {\bf K} \Gamma \wedge {\bf K}A \wedge \bigwedge \Delta \rightarrow B& \text{assumption} \\
+&\text{(3) } {\bf K} A \wedge (\bigwedge {\bf K} \Gamma \wedge \bigwedge \Delta) \rightarrow B& \text{conjuction properties} \\
+&\text{(4) } ({\bf K} A \wedge (\bigwedge {\bf K} \Gamma \wedge \bigwedge \Delta) \rightarrow B) \rightarrow ({\bf K } A \rightarrow ((\bigwedge {\bf K} \Gamma \wedge \bigwedge \Delta) \rightarrow B)) & \text{IPC theorem} \\
+&\text{(5) } {\bf K} A \rightarrow ((\bigwedge {\bf K} \Gamma \wedge \bigwedge \Delta) \rightarrow B) & \text{MP } \\
+&\text{(6) } \bigwedge {\bf K} \Gamma \wedge \bigwedge \Delta \to ((\bigwedge {\bf K} \Gamma \wedge \bigwedge \Delta) \rightarrow B) & \text{(1), (5), transitivity} \\
+&\text{(7) } (\bigwedge {\bf K} \Gamma \wedge \bigwedge \Delta) \rightarrow B & \text{(6), contraction}
 \end{array}$
 
-\vspace{\baselineskip}
-
-2) If $\Gamma \vdash_{NIEL^{-}_{\land, \to}} {\bf K} \vec{A}$ and $\vec{A} \vdash B$, then $IEL^{-} \vdash \bigwedge \Gamma \rightarrow \textbf{K}B$.
-
-$\begin{array}{lll}
-(1) &\bigwedge \Gamma \rightarrow \bigwedge \limits_{i = 1}^{n} {\bf K} A_i & \text{assumption} \\
-(2) &\bigwedge \limits_{i = 1}^{n} {\bf K} A_i \rightarrow {\bf K} \bigwedge \limits_{i = 1}^{n} A_i& \text{IEL theorem} \\
-(3) &\bigwedge \Gamma \rightarrow {\bf K} \bigwedge \limits_{i = 1}^{n} A_i & \text{from (1), (2) and transitivity} \\
-(4) &\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B& \text{assumption} \\
-(5) &(\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B) \rightarrow {\bf K} (\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B)& \text{co-reflection}\\
-(6) &{\bf K} (\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B)& \text{from (2), (3) and MP} \\
-(7) &{\bf K} \bigwedge \limits_{i = 1}^{n} A_i \rightarrow {\bf K} B & \text{from (6) and normality} \\
-(8) &\bigwedge \Gamma \rightarrow {\bf K} B & \text{from (3), (7) and transitivity}
-\end{array}$
-
+  \end{proof}
+
+  \begin{lemma}
+    $ $
+
+    If $\text{IEL}^{-} \vdash A$, then $\text{NIEL}^{-} \vdash A$.
+  \end{lemma}
+
+  \begin{proof}
+    $ $
+    Straightforward derivation of modal axioms in NIEL.
   \end{proof}
@@
   Context, domain of context and range of context are defined standardly \cite{Neder}\cite{Morten}.
+
+  We divide into two parts, the first one zone is a modal part (enviroment with applicaitve types), the second one
+  is a intuitionistic part (enviroment with pure types). From a logical point of view, $\Gamma | \Delta \vdash M : A$ denotes that
+  $M$ is a proof of ${\bf K} \Gamma, \Delta \vdash A$. This idea similar to \cite{DePaiva}.
 
@@
 
-  2) $(M N) [x := N] = M[x := N] N [x := N]$;
+  2) $(M N) [x := P] = M[x := P] N [x := P]$;
 
@@
 &\text{3) } \pi_2 \langle M, N \rangle \rightarrow_{\beta} N& \\
-&\text{4) } {\bf let \: pure \: } \vec{y} = \vec{M_2} {\bf \: in \: } ({\bf let \: pure \: } \vec{x} = \vec{M_1} {\bf \: in \: }  N) \rightarrow_{\beta}& \\
-&\quad {\bf let \: pure \: } \vec{x} = ({\bf let \: pure \: } \vec{y} = \vec{M_2} \: {\bf in} \: \vec{M_1}) \: {\bf in} \: {\bf let \: pure \: } \vec{x} = \vec{x} {\bf \: in \:} N&
 \end{array}$
@@
 
-  So $\Gamma \uparrow FV(\vec{M}) \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\bf \: in \:} N : \textbf{K}B$.
+  So $\Gamma \uparrow FV(\vec{M}) \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\bf \: in \:} N : \textbf{K}B$ by ${\bf let}_{\textbf{K}}$.
 
@@
 
-  2) .
-
-  \end{proof}
-
-  \vspace{\baselineskip}
+  2) Let $\Gamma, y : C \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\bf \: in \:} N : \textbf{K}B$ and $\Gamma \vdash P : C$.
+
+  By generation $\Gamma, y : C \vdash \vec{M} : \textbf{K}\vec{A}$ and $\vec{x} : \vec{A} \vdash N : \textbf{K}B$.
+
+  By assumption $\Gamma \vdash \overrightarrow{M [y := P]} : \textbf{K}A$.
+
+  So $\Gamma \vdash {\bf let \: pure \:} \vec{x} = \overrightarrow{M [y := P]} {\bf \: in \:} N : \textbf{K}B$ by ${\bf let}_{\textbf{K}}$.
+
+  \end{proof}
+
+  \vspace{\baselineskip}
+
+  \begin{defin} Type substitution
+
+    The substitution of type $C$ for type variable $B$ in some type $A$ is defined by:
+
+    1) $A [ A := B ] = B$;
+
+    2) $A^{'} [A := C] = A^{'}$;
+
+    3) $(A_1 \alpha A_2) [B := C] = A_1 [B := C] \alpha A_2 [A := B]$, where $\alpha \{ \times, \to \}$;
+
+    4) $(\textbf{K}A) [B := C] = \textbf{K}(A [B := C])$.
+
+  \end{defin}
+
+    \vspace{\baselineskip}
+
+  \begin{lemma} Substitution for types.
+
+    Let $\Gamma \vdash M : A$, then $\Gamma [B := C] \vdash M : A [B := C]$, where $\Gamma [B := C]$ denotes $\{ x_1 : A_1 [B := C],\dots, x_n : A_n [B := C] \}$.
+
+  \end{lemma}
+
+  1) Let $\Gamma, x : A \vdash x : A$, so $\Gamma [B := C], x : A [B := C] \vdash x : A [B := C]$.
+
+  \vspace{\baselineskip}
+
+  2) Let $\Gamma \vdash {\bf pure \: } M : \textbf{K}A$. By generation $\Gamma \vdash M : A$.
+
+  By assumption $\Gamma [B := C] \vdash M : A [B := C]$.
+
+  By $\textbf{K}_I$, $\Gamma [B := C] \vdash {\bf pure \:} M : \textbf{K} (A [B := C])$.
+
+  But, by substitution definition, $\textbf{K} (A [B := C]) = (\textbf{K}A) [B := C]$.
+
+  Then $\Gamma [B := C] \vdash {\bf pure \:} M : (\textbf{K}A) [B := C]$.
+
+  3) $\Gamma \vdash {\bf let \: pure} \: \vec{x} = \vec{M} \: {\bf in} \: M : {\bf K} B$.
+
+  By generation $\Gamma \vdash \vec{M} : \textbf{K}\vec{A}$ and $\vec{x} : \vec{A} \vdash N : \textbf{K}B$.
+
+  $\Gamma [D := C] \vdash \vec{M} : \textbf{K}\overrightarrow{A [D := C]}$ and $\vec{x} : \overrightarrow{A [D := C]} \vdash M : B [D := C]$ by assumption.
+
+  Then $\Gamma [D := C] \vdash {\bf let \: pure } \: \vec{x} = \vec{M} \: {\bf in} \: M : {\bf } (B [D := C])$.
 
@@
 
-  We consider cases with reduction rules which are applicative laws. The general statement for $\twoheadrightarrow_{\beta}$ follows from transitivity of multi-step $\beta$-reduction.
-
-  \begin{proof}
-
-  $ $
-
-\vspace{\baselineskip}
-
-  1) Let $\Gamma \vdash {\bf pure \: } (\lambda x. x) \star M : \textbf{K}\alpha$. Then $\Gamma \vdash
-  {\bf pure \: } (\lambda x.x) : \textbf{K}(\alpha \to \alpha)$ and $\Gamma \vdash M : \textbf{K}\alpha$ by
-  generation. Then $\Gamma \vdash M : \textbf{K}\alpha$ trivially.
-
-\vspace{\baselineskip}
-
-  2) Let $\Gamma \vdash {\bf pure \: }(\lambda f g x. f (g x)) \star M \star N \star P : \textbf{K}\gamma$.
-
-  Then $\Gamma \vdash {\bf pure \: } (\lambda f g x. f (g x)) : \textbf{K}((\beta \to \gamma) \to (\alpha \to \beta)
-  \to \alpha \to \gamma)$, $\Gamma \vdash M : \textbf{K}(\beta \to \gamma)$, $\Gamma \vdash N :
-  \textbf{K}(\alpha \to \beta)$ and $\Gamma \vdash P : \textbf{K}\alpha$ by generation.
-
-  If $\Gamma \vdash N : \textbf{K}(\alpha \to \beta)$ and $\Gamma \vdash P : \textbf{K}\alpha$, then $\Gamma
-  \vdash N \star P : \textbf{K}\beta$ by $\textbf{K}_{app}$.
-
-  Hence, if $\Gamma \vdash M : \textbf{K}(\beta \to \gamma)$, then $\Gamma \vdash M \star (N \star P) :
-  \textbf{K}\gamma$ by $\textbf{K}_{app}$.
-
-\vspace{\baselineskip}
-
-  3) Let $\Gamma \vdash ({\bf pure \: } M) \star ({\bf pure \: } N) : \textbf{K}\beta$. Then $\Gamma \vdash
-  {\bf pure \: } M : \textbf{K}(\alpha \to \beta)$ and $\Gamma \vdash {\bf pure \: } N : \textbf{K}\alpha$ by
-  generation. Moreover, $\Gamma \vdash M : \alpha \to \beta$ and $\Gamma \vdash N : \alpha$.
-
-  Then $\Gamma \vdash M N : \beta$ by application.
-
-  Hence, $\Gamma \vdash {\bf pure \: } (M N) : \textbf{K}\beta$ by $K_I$.
-
-\vspace{\baselineskip}
-
-  4) Let $\Gamma \vdash M \star ({\bf pure \: } N) : \textbf{K}\beta$.
-
-  Then $\Gamma \vdash M : \textbf{K}(\alpha \to \beta)$ and $\Gamma \vdash {\bf pure \: } N : \textbf{K}\alpha$.
-
-  Moreover, $\Gamma \vdash N : \alpha$ by generation.
-
-  Let $\Gamma, f : \alpha \to \beta \vdash f : \alpha \to \beta$ and $\Gamma, f : \alpha \to \beta \vdash N :
-  \alpha$ by weakening.
-
-  So $\Gamma, f : \alpha \to \beta \vdash f N : \beta$ by application, so $\Gamma \vdash \lambda f. f N :
-  (\alpha \to \beta) \to \beta$ by abstraction.
-
-  Then $\Gamma \vdash {\bf pure \: } (\lambda f. f N) : \textbf{K}((\alpha \to \beta) \to \beta)$ by $\textbf{K}_I$.
-
-  Hence, $\Gamma \vdash {\bf pure \: } (\lambda f. f N) \star M : \textbf{K}\beta$.
+  \begin{proof}
 
@@
   the Foundations of Mathematics, vol. 103. Amsterdam: North-Holland, 1985.
+
+  \bibitem{DePaiva} de Paiva V., Ritter E., Bellin, G.
+Extended Curry-Howard correspondence for a basic constructive modal logic.
+Methods for Modalities Conference (M4M-2); 2001 November 29-30; Amsterdam; the Netherlands.
 \end{thebibliography}

### 2018-01-14 article.tex
added
Revert "added rewrite"
This reverts commit 82d65e58f1e2ae7113751d29f684b3a94ca5ce90.

@@
   	\begin{prooftree}
-    	\AxiomC{$\Gamma, \Delta \vdash A$}
-    	\RightLabel{$\vdash {\bf K}_I$}
-    	\UnaryInfC{${\bf K}(\Gamma), \Delta \vdash {\bf K}A$}
+    	\AxiomC{$\Gamma \vdash A$}
+    	\RightLabel{$\textbf{K}_I$}
+    	\UnaryInfC{$\Gamma \vdash \textbf{K}A$}
   \end{prooftree}
@@
   	\begin{prooftree}
-    \AxiomC{${\bf K}(\Gamma), \Delta \vdash {\bf K} A$}
-    \AxiomC{${\bf K}(\Gamma), {\bf K} A, \Delta \vdash B$}
-    \BinaryInfC{${\bf K}(\Gamma), \Delta \vdash B$}
+    \AxiomC{$\Gamma \vdash {\bf K} \vec{A}$}
+    \AxiomC{$\vec{A} \vdash B$}
+    \BinaryInfC{$\Gamma \vdash {\bf K} B$}
   	\end{prooftree}
@@
 
-1) If $\Gamma, \Delta \vdash_{NIEL^{-}_{\land, \to}} A$, then $IEL^{-} \vdash \bigwedge {\bf K} \Gamma \wedge \bigwedge \Delta \rightarrow {\bf K}A$.
-
-See \cite{Krupski}.
-
-\vspace{\baselineskip}
-
-2) If ${\bf K} \Gamma, \Delta \vdash_{NIEL^{-}_{\land, \to}} {\bf K} A$ and ${\bf K}\Gamma, {\bf K} A, \Delta \vdash B$, then $IEL^{-} \vdash \bigwedge {\bf K} \Gamma \wedge \bigwedge \Delta \rightarrow B$.
+1) If $\Gamma \vdash_{NIEL^{-}_{\land, \to}} A$, then $IEL^{-} \vdash \bigwedge \Gamma \rightarrow \textbf{K}A$.
 
 $\begin{array}{lll}
-&\text{(1) } \bigwedge {\bf K} \Gamma \wedge \bigwedge \Delta \to {\bf K} A& \text{assumption} \\
-&\text{(2) } \bigwedge {\bf K} \Gamma \wedge {\bf K}A \wedge \bigwedge \Delta \rightarrow B& \text{assumption} \\
-&\text{(3) } {\bf K} A \wedge (\bigwedge {\bf K} \Gamma \wedge \bigwedge \Delta) \rightarrow B& \text{conjuction properties} \\
-&\text{(4) } ({\bf K} A \wedge (\bigwedge {\bf K} \Gamma \wedge \bigwedge \Delta) \rightarrow B) \rightarrow ({\bf K } A \rightarrow ((\bigwedge {\bf K} \Gamma \wedge \bigwedge \Delta) \rightarrow B)) & \text{IPC theorem} \\
-&\text{(5) } {\bf K} A \rightarrow ((\bigwedge {\bf K} \Gamma \wedge \bigwedge \Delta) \rightarrow B) & \text{MP } \\
-&\text{(6) } \bigwedge {\bf K} \Gamma \wedge \bigwedge \Delta \to ((\bigwedge {\bf K} \Gamma \wedge \bigwedge \Delta) \rightarrow B) & \text{(1), (5), transitivity} \\
-&\text{(7) } (\bigwedge {\bf K} \Gamma \wedge \bigwedge \Delta) \rightarrow B & \text{(6), contraction}
+(1) & \bigwedge \Gamma \rightarrow A & \text{assumption}\\
+(2) & A \rightarrow \textbf{K}A &\text{co-reflection}\\
+(3) & (\bigwedge \Gamma \rightarrow A) \rightarrow ((A \rightarrow \textbf{K}A) \rightarrow (\bigwedge \Gamma \rightarrow \textbf{K}A))&\text{IPC theorem}\\
+(4) & (A \rightarrow \textbf{K}A) \rightarrow (\bigwedge \Gamma \rightarrow \textbf{K}A) &\text{from (1), (3) and MP}\\
+(5) & \bigwedge \Gamma \rightarrow \textbf{K}A &\text{from (2), (4) and MP}\\
 \end{array}$
 
-  \end{proof}
-
-  \begin{lemma}
-    $ $
-
-    If $\text{IEL}^{-} \vdash A$, then $\text{NIEL}^{-} \vdash A$.
-  \end{lemma}
-
-  \begin{proof}
-    $ $
-    Straightforward derivation of modal axioms in NIEL.
+\vspace{\baselineskip}
+
+2) If $\Gamma \vdash_{NIEL^{-}_{\land, \to}} {\bf K} \vec{A}$ and $\vec{A} \vdash B$, then $IEL^{-} \vdash \bigwedge \Gamma \rightarrow \textbf{K}B$.
+
+$\begin{array}{lll}
+(1) &\bigwedge \Gamma \rightarrow \bigwedge \limits_{i = 1}^{n} {\bf K} A_i & \text{assumption} \\
+(2) &\bigwedge \limits_{i = 1}^{n} {\bf K} A_i \rightarrow {\bf K} \bigwedge \limits_{i = 1}^{n} A_i& \text{IEL theorem} \\
+(3) &\bigwedge \Gamma \rightarrow {\bf K} \bigwedge \limits_{i = 1}^{n} A_i & \text{from (1), (2) and transitivity} \\
+(4) &\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B& \text{assumption} \\
+(5) &(\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B) \rightarrow {\bf K} (\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B)& \text{co-reflection}\\
+(6) &{\bf K} (\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B)& \text{from (2), (3) and MP} \\
+(7) &{\bf K} \bigwedge \limits_{i = 1}^{n} A_i \rightarrow {\bf K} B & \text{from (6) and normality} \\
+(8) &\bigwedge \Gamma \rightarrow {\bf K} B & \text{from (3), (7) and transitivity}
+\end{array}$
+
   \end{proof}
@@
   Context, domain of context and range of context are defined standardly \cite{Neder}\cite{Morten}.
-
-  We divide into two parts, the first one zone is a modal part (enviroment with applicaitve types), the second one
-  is a intuitionistic part (enviroment with pure types). From a logical point of view, $\Gamma | \Delta \vdash M : A$ denotes that
-  $M$ is a proof of ${\bf K} \Gamma, \Delta \vdash A$. This idea similar to \cite{DePaiva}.
 
@@
 &\text{3) } \pi_2 \langle M, N \rangle \rightarrow_{\beta} N& \\
+&\text{4) } {\bf let \: pure \: } \vec{y} = \vec{M_2} {\bf \: in \: } ({\bf let \: pure \: } \vec{x} = \vec{M_1} {\bf \: in \: }  N) \rightarrow_{\beta}& \\
+&\quad {\bf let \: pure \: } \vec{x} = ({\bf let \: pure \: } \vec{y} = \vec{M_2} \: {\bf in} \: \vec{M_1}) \: {\bf in} \: {\bf let \: pure \: } \vec{x} = \vec{x} {\bf \: in \:} N&
 \end{array}$
@@
   the Foundations of Mathematics, vol. 103. Amsterdam: North-Holland, 1985.
-
-  \bibitem{DePaiva} de Paiva V., Ritter E., Bellin, G.
-Extended Curry-Howard correspondence for a basic constructive modal logic.
-Methods for Modalities Conference (M4M-2); 2001 November 29-30; Amsterdam; the Netherlands.
   \end{thebibliography}

### 2018-01-14 article.tex
Revert "added some fixes and rewrite"
This reverts commit 3a317452fa7ea77dedaa76a6123d357676ea22dd.

--- 2018-01-13 article.tex
@@
   \RightLabel{\scriptsize{ax}}
-  \UnaryInfC{$\Gamma , x : A \vdash x : A$}
+  \UnaryInfC{$\Gamma , x : \alpha \vdash x : \alpha$}
   \end{prooftree}
@@
   \begin{prooftree}
-  \AxiomC{$\Gamma, x : A \vdash M : B$}
+  \AxiomC{$\Gamma, x : \alpha \vdash M : \beta$}
   \RightLabel{$\rightarrow_i$}
-  \UnaryInfC{$\Gamma \vdash \lambda x. M : A \to B$}
-  \end{prooftree}
-
-  \begin{prooftree}
-  \AxiomC{ $\Gamma \vdash x : A$ }
-  \AxiomC{ $\Gamma \vdash y : B$ }
+  \UnaryInfC{$\Gamma \vdash \lambda x. M : \alpha \to \beta$}
+  \end{prooftree}
+
+  \begin{prooftree}
+  \AxiomC{ $\Gamma \vdash x : \alpha$ }
+  \AxiomC{ $\Gamma \vdash y : \beta$ }
   \RightLabel{$\times_i$}
-  \BinaryInfC{$\Gamma \vdash \langle x, y \rangle : A \times B$}
-  \end{prooftree}
-
-  \begin{prooftree}
-    \AxiomC{$\Gamma \vdash x : A$}
+  \BinaryInfC{$\Gamma \vdash (x, y) : \alpha \times \beta$}
+  \end{prooftree}
+
+  \begin{prooftree}
+    \AxiomC{$\Gamma \vdash x : \alpha$}
     \RightLabel{$\textbf{K}_I$}
-    \UnaryInfC{$\Gamma \vdash {\bf pure \: } \: x : \textbf{K}A $}
+    \UnaryInfC{$\Gamma \vdash {\bf pure \: } \: x : \textbf{K}\alpha $}
   \end{prooftree}
@@
   \begin{prooftree}
-  \AxiomC{$\Gamma \vdash f : A \to B$}
-  \AxiomC{$\Gamma \vdash x : A$}
+  \AxiomC{$\Gamma \vdash f : \alpha \to \beta$}
+  \AxiomC{$\Gamma \vdash x : \alpha$}
   \RightLabel{$\rightarrow_e$}
-  \BinaryInfC{$\Gamma \vdash fx : B$}
-  \end{prooftree}
-
-  \begin{prooftree}
-  \AxiomC{ $\Gamma \vdash M : A_1 \times A_2$ }
+  \BinaryInfC{$\Gamma \vdash fx : \beta$}
+  \end{prooftree}
+
+  \begin{prooftree}
+  \AxiomC{ $\Gamma \vdash p : \alpha_1 \times \alpha_2$ }
   \RightLabel{$\times_e$, $i \in \{ 1, 2 \}$}
-  \UnaryInfC{$\Gamma \vdash \pi_i M : A_i$}
+  \UnaryInfC{$\Gamma \vdash \pi_i p : \alpha_i$}
   \end{prooftree}
@@
     \AxiomC{$\Gamma \vdash \vec{M} : {\bf K} \vec{A}$}
-    \AxiomC{$\vec{x} : \vec{A} \vdash N : B$}
-    \RightLabel{${\bf let}_{\textbf{K}}$}
-    \BinaryInfC{$\Gamma \vdash {\bf let \:\: pure} \:\: \vec{N} = \vec{M} \:\: {\bf in} \:\: N : {\bf K} B$}
+    \AxiomC{$\vec{N} : \vec{A} \vdash M : B$}
+    \BinaryInfC{$\Gamma \vdash {\bf let \:\: pure} \:\: \vec{N} = \vec{M} \:\: {\bf in} \:\: M : {\bf K} B$}
   \end{prooftree}
@@
 
-  7) $FV({\bf let \:\: pure} \:\: \vec{x} = \vec{M} \:\: {\bf in} \:\: N) = \bigcup \limits_{i = 1}^n FV(M_i), \text{where $n = |\vec{M}|$}$.
+  7) $FV({\bf let \:\: pure} \:\: \vec{N} = \vec{M} \:\: {\bf in} \:\: M) = \bigcup \limits_{i = 1}^n FV(M), \text{where $n = |\vec{M}|$}$.
   \end{defin}
@@
 
-  In $\lambda \textbf{K}$ we have the following computational rules:
+  In $\lambda \textbf{K}$ we have the following computational rules. We will define them for terms with
+  ${\bf pure \: }$ or $\star$. Reduction rules for $(,)$ and $\pi_i$ are described, for example, in
+  \cite{Pierce}.
 
@@
 
-$\begin{array}{lll}
-&\text{1) } (\lambda x. M) N \rightarrow_{\beta} M [x := N]& \\
-&\text{2) } \pi_1 \langle M, N \rangle \rightarrow_{\beta} M& \\
-&\text{3) } \pi_2 \langle M, N \rangle \rightarrow_{\beta} N& \\
-&\text{4) } {\bf let \: pure \: } \vec{y} = \vec{M_2} {\bf \: in \: } ({\bf let \: pure \: } \vec{x} = \vec{M_1} {\bf \: in \: }  N) \rightarrow_{\beta}& \\
-&\quad {\bf let \: pure \: } \vec{x} = ({\bf let \: pure \: } \vec{y} = \vec{M_2} \: {\bf in} \: \vec{M_1}) \: {\bf in} \: {\bf let \: pure \: } \vec{x} = \vec{x} {\bf \: in \:} N&
-\end{array}$
+  1) $(\lambda x. M) N \rightarrow_{\beta} M [x := N]$;
+
+  2) $\pi_1 \langle M, N \rangle \rightarrow_{\beta} M$;
+
+  3) $\pi_2 \langle M, N \rangle \rightarrow_{\beta} N$;
+
+  4) ${\bf let \: pure \: } \vec{N_1} = \vec{M_1} \: {\bf in} \: {\bf let \: pure \:} \vec{N_2} = \vec{M_2} \: {\bf in} \: M \rightarrow_{\beta} $
 
@@
 
-  Let $\Gamma = \{ x_1 : A_1, ..., x_n : A_n \}$. Then the domain of $\Gamma$, or $dom(\Gamma)$, is a set $\{  x_1, ..., x_n\}$.
+  Let $\Gamma = \{ x_1 : \alpha_1, ..., x_n : \alpha_n \}$. Then the domain of $\Gamma$, or $dom(\Gamma)$, is a set $\{  x_1, ..., x_n\}$.
 
@@
 
-  If $\Gamma \vdash M : A$, then $FV(M) \subseteq dom(\Gamma)$
-
-  \end{lemma}
-
-  \begin{proof}
-
-  Induction on the derivation of $\Gamma \vdash M : A$.
+  If $\Gamma \vdash M : \alpha$, then $FV(M) \subseteq dom(\Gamma)$
+
+  \end{lemma}
+
+  \begin{proof}
+
+  Induction on the derivation of $\Gamma \vdash M : \alpha$.
 
@@
 
-  1) $\Gamma \vdash {\bf pure \: } M : \textbf{K}A$ implies that $\Gamma \vdash M : A$;
+  1) $\Gamma \vdash {\bf pure \: } M : \textbf{K}\alpha$ implies that $\Gamma \vdash M : \alpha$;
 
@@
 
-  Let $\Gamma \vdash M : A$ and $\Gamma \subseteq \Delta$, then $\Delta \vdash M : A$.
+  Let $\Gamma \vdash M : \alpha$ and $\Gamma \subseteq \Delta$, then $\Delta \vdash M : \alpha$.
 
@@
 
-  1) Let $\Gamma \vdash x : A$, such that $\Gamma = \Delta, x : A$ and $\Theta \subseteq \Gamma$.
+  1) Let $\Gamma \vdash x : \alpha$, such that $\Gamma = \Delta, x : \alpha$ and $\Theta \subseteq \Gamma$.
   Let $\Sigma = \Theta \setminus \Gamma$, or, which is the same, $\Sigma = \Theta \setminus \Delta, x :
-  A$, then $\Sigma, \Delta, x : A \vdash x : \alpha$, or, $\Theta \vdash x : A$.
-
-  2) Let $\Gamma \vdash {\bf pure \: } M : \textbf{K}A$ and $\Gamma \subseteq \Theta$.
-
-  If $\Gamma \vdash {\bf pure \: } M : \textbf{K}A$, then $\Gamma \vdash M : A$ by generation and, by
-  hypothesis, $\Theta \vdash M : A$, so $\Theta \vdash {\bf pure \: } M : \textbf{K}A$ by applying
+  \alpha$, then $\Sigma,
+  \Delta, x : \alpha \vdash x : \alpha$, or, $\Theta \vdash x : \alpha$.
+
+  2) Let $\Gamma \vdash {\bf pure \: } M : \textbf{K}\alpha$ and $\Gamma \subseteq \Theta$.
+
+  If $\Gamma \vdash {\bf pure \: } M : \textbf{K}\alpha$, then $\Gamma \vdash M : \alpha$ by generation and, by
+  hypothesis, $\Theta \vdash M : \alpha$, so $\Theta \vdash {\bf pure \: } M : \textbf{K}\alpha$ by applying
   $\textbf{K}_I$-rule.
 
-  3) Let $\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\bf \: in \:} N : \textbf{K}B$ and $\Gamma \subseteq \Delta$.
-
-  By generation $\Gamma \vdash \vec{M} : \textbf{K}\vec{A}$ and $\vec{x} : \vec{A} \vdash N : \textbf{K}B$.
-
-  By hypothesis we have $\Delta \vdash \vec{M} : \textbf{K}\vec{A}$. So $\Delta \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\bf \: in \:} N : \textbf{K}B$ by ${\bf let}_{\textbf{K}}$.
+  3) Let $\Gamma \vdash M \star N : \textbf{K}\beta$. So $\Gamma \vdash M : \textbf{K}(\alpha \to \beta)$ and $\Gamma
+  \vdash N : \textbf{K}\alpha$. By hypothesis $\Delta \vdash M : \textbf{K}(\alpha \to \beta)$ and $\Delta \vdash N :
+  \textbf{K}\alpha$. Then $\Delta \vdash M \star N : \textbf{K}\beta$.
 
@@
 
-  If $\Gamma \vdash M : A$, then $\Gamma \uparrow FV(M) \vdash M : A$, where $\Gamma \uparrow
+  If $\Gamma \vdash M : \alpha$, then $\Gamma \uparrow FV(M) \vdash M : \alpha$, where $\Gamma \uparrow
   FV(M)$ is a subcontext of $\Gamma$, such that $dom(\Gamma \uparrow FV(M)) = dom(\Gamma) \cap FV(M)$.
@@
 
-  Induction by derivation. We consider the base of induction and the case with  ${\bf let}_{\textbf{K}}$. The rest
+  Induction by derivation. We consider the base of induction and the case with $\textbf{K}_{app}$. The rest
   cases are
@@
 
-  1) Let $\Gamma \vdash x : A$, where $\Gamma = \Delta, x : A$, $x \in \mathbb{V}$.
-
-  $FV(x) = \{ x\}$, then $dom(\Gamma) \cap \{ x\} = \{ x\}. $ So $(\Delta, x : A) \uparrow FV(x) = \{ x
-  : A\}$, then $x : A \vdash x : A$ by axiom.
-
-  2) Let $\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\bf \: in \:} N : \textbf{K}B$.
-
-  By generation $\Gamma \vdash \vec{M} : \textbf{K}\vec{A}$ and $\vec{x} : \vec{A} \vdash N : \textbf{K}B$.
-
-  By hypothesis $\Gamma \uparrow FV(\vec{M}) \vdash \vec{M} : \textbf{K}\vec{A}$.
-
-  So $\Gamma \uparrow FV(\vec{M}) \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\bf \: in \:} N : \textbf{K}B$.
+  1) Let $\Gamma \vdash x : \alpha$, where $\Gamma = \Delta, x : \alpha$, $x \in \mathbb{V}$.
+
+  $FV(x) = \{ x\}$, then $dom(\Gamma) \cap \{ x\} = \{ x\}. $ So $(\Delta, x : \alpha) \uparrow FV(x) = \{ x
+  : \alpha\}$, then $x : \alpha \vdash x : \alpha$ by axiom.
+
+  2) Let $\Gamma \vdash M \star N : \textbf{K}\beta$.
+
+  By generation $\Gamma \vdash M : \textbf{K}(\alpha \to \beta)$ and $\Gamma \vdash N : \textbf{K}\alpha$.
+
+  By induction hypothesis $\Gamma \uparrow FV(M) \vdash M : \textbf{K}(\alpha \to \beta)$ and $\Gamma
+  \uparrow FV(N) \vdash N : \textbf{K}\beta$.
+
+  But $\Gamma \uparrow FV(M) \subseteq \Gamma \uparrow FV(M \star N)$ and $\Gamma
+  \uparrow FV(N) \subseteq \Gamma \uparrow FV(M \star N)$, so $\Gamma \uparrow FV(M \star N) \vdash M :
+  \textbf{K}(\alpha \to \beta)$ and $\Gamma \uparrow FV(M \star N) \vdash N : \textbf{K}\alpha$ by weakening.
+
+  Then $\Gamma \uparrow FV(M \star N) \vdash M \star N : \textbf{K}\beta$ by $\textbf{K}_{app}$.
 
@@
 
-  If $\Gamma, x : A \vdash M : B$  and $\Gamma \vdash N : A$, then $\Gamma \vdash (M [x := N]) :
-  B $
-
-  \end{lemma}
-
-  \begin{proof}
-
-  $ $
-
-  1) Let $\Gamma, x : A \vdash {\bf pure \: } M : \textbf{K}B$ and $\Gamma \vdash N : A$.
-
-  If $\Gamma, x : \alpha \vdash {\bf pure \: } M : \textbf{K}\beta$.
-
-  By generation, $\Gamma, x : A \vdash M : B$.
-
-  So, by induction hypothesis, $\Gamma \vdash (M [x := N]) : B $.
-  Then $\Gamma \vdash {\bf pure \: } (M [x := N]) : \textbf{K}B$ by $\textbf{K}_I$, but ${\bf pure \: } (M [x := N]) =
-  ({\bf pure \: } M [x := N])$ by substitution definition.
-
-  So $\Gamma \vdash ({\bf pure \: } M [x := N]) : \textbf{K}B$
-
-  2) .
+  If $\Gamma, x : \alpha \vdash M : \beta$  and $\Gamma \vdash N : \alpha$, then $\Gamma \vdash (M [x := N]) :
+  \beta $
+
+  \end{lemma}
+
+  \begin{proof}
+
+  $ $
+
+  1) Let $\Gamma, x : \alpha \vdash {\bf pure \: } M : \textbf{K}\beta$ and $\Gamma \vdash N : \alpha$.
+
+  If $\Gamma, x : \alpha \vdash {\bf pure \: } M : \textbf{K}\beta$, then, by generation, $\Gamma, x : A \vdash
+  M : \beta$. So, by induction hypothesis, $\Gamma \vdash (M [x := N]) : \beta $, then $\Gamma \vdash
+  {\bf pure \: } (M [x := N]) : \textbf{K}\beta$ by $\textbf{K}_I$, but ${\bf pure \: } (M [x := N]) =
+  ({\bf pure \: } M) (M [x := N])$
+  by substitution definition, so $\Gamma \vdash ({\bf pure \: } M) (M [x := N]) : \textbf{K}\beta$
+
+  2) Let $\Gamma, x : \gamma \vdash M \star N : \textbf{K}\beta$, and $\Gamma \vdash y : \gamma$.
+
+  So, by generation, $\Gamma, x : \gamma \vdash M : \textbf{K}(\alpha \to \beta)$ and $\Gamma, x : \gamma
+  \vdash N : \textbf{K}\alpha$.
+
+  Hence $\Gamma, x : \gamma \vdash (M [x := y]) : \textbf{K}(\alpha \to \beta)$ and
+  $\Gamma \vdash (N [x := y]): \textbf{K}\alpha$ by hypothesis.
+
+  So $\Gamma \vdash (M [x := y]) \star (N [x := y]) : \textbf{K}\beta$, or, $\Gamma \vdash (M \star N) ([x
+  := y]) : \textbf{K}\beta$.
 

### 2018-01-14 article.tex
rewrite

--- 2018-01-14 article.tex
@@
 
+\begin{lemma}
+  $ $
+  If $\text{IEL}^{-} \vdash A$, then $\text{NIEL}^{-} \vdash A$.
+\end{lemma}
+
+\begin{proof}
+  Straightforward derivation of modal axioms in $\text{NIEL}^{-}$. We consider this derivation below using terms.
+\end{proof}
+
   \vspace{\baselineskip}
@@
     \AxiomC{$\Gamma \vdash \vec{M} : {\bf K} \vec{A}$}
-    \AxiomC{$\vec{x} : \vec{A} \vdash N : B$}
-    \RightLabel{${\bf let}_{\textbf{K}}$}
-    \BinaryInfC{$\Gamma \vdash {\bf let \:\: pure} \:\: \vec{N} = \vec{M} \:\: {\bf in} \:\: N : {\bf K} B$}
+    \AxiomC{$\vec{x} : \vec{A} \vdash M : B$}
+    \RightLabel{$\text{let}_{\bf K}$}
+    \BinaryInfC{$\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } M : {\bf K} B$}
   \end{prooftree}
@@
   Haskell method {\bf pure} for Applicative class. It plays the same role as the {\bf return} method in Monad class.
+
+  $\text{let}_{\bf K}$ is similar to $\Box_I$-rule in typed lambda calculus for intuitionistic normal modal logic {\bf IK}, which is described in \cite{ModalK}.
 
@@
 
-  2) $(M N) [x := P] = M[x := P] N [x := P]$;
+  2) $(M N) [x := N] = M[x := N] N [x := N]$;
 
@@
 
-  In $\lambda \textbf{K}$ we have the following computational rules:
-
-  \begin{defin} $\beta$-reduction rules for \emph{$\lambda \textbf{K}$}.
-
-$\begin{array}{lll}
-&\text{1) } (\lambda x. M) N \rightarrow_{\beta} M [x := N]& \\
-&\text{2) } \pi_1 \langle M, N \rangle \rightarrow_{\beta} M& \\
-&\text{3) } \pi_2 \langle M, N \rangle \rightarrow_{\beta} N& \\
-&\text{4) } {\bf let \: pure \: } \vec{y} = \vec{M_2} {\bf \: in \: } ({\bf let \: pure \: } \vec{x} = \vec{M_1} {\bf \: in \: }  N) \rightarrow_{\beta}& \\
-&\quad {\bf let \: pure \: } \vec{x} = ({\bf let \: pure \: } \vec{y} = \vec{M_2} \: {\bf in} \: \vec{M_1}) \: {\bf in} \: {\bf let \: pure \: } \vec{x} = \vec{x} {\bf \: in \:} N&
+  \begin{defin} $\beta$-reduction and $\eta$-reduction rules for \emph{$\lambda \textbf{K}$}.
+
+  1) $(\lambda x. M) N \rightarrow_{\beta} M [x := N]$;
+
+  2) $\pi_1 \langle M, N \rangle \rightarrow_{\beta} M$;
+
+  3) $\pi_2 \langle M, N \rangle \rightarrow_{\beta} N$;
+
+  4) $\begin{array}{llll}
+  & {\bf let \: pure \:} \langle \vec{x}, y, \vec{z} \rangle = \langle \vec{M}, {\bf let \: pure \: } \vec{w} = \vec{N} {\bf \: in \: } Q, \vec{P} \rangle {\: in \:} R \rightarrow_{\beta} & \\
+  & {\bf let \: pure \:} \langle \vec{x}, \vec{w}, \vec{z} \rangle = \langle \vec{M}, \vec{N}, \vec{P} \rangle \: { \bf in } \: R [y := Q]
   \end{array}$
 
+  5) ${\bf pure \:} ((\lambda x. M) N) \rightarrow_{\beta} {\bf pure \: } (M [x := N])$;
+
+  6) ${\bf pure \:} (\pi_i \langle M_1, M_2 \rangle) \rightarrow_{\beta} {\bf pure \:} M_i$, where $i \in \{ 1, 2 \}$.
+
+  7) $\begin{array}{llll}
+  & {\bf pure \:} ({\bf let \: pure \:} \langle \vec{x}, y, \vec{z} \rangle = \langle \vec{M}, {\bf let \: pure \: } \vec{w} = \vec{N} {\bf \: in \: } Q, \vec{P} \rangle {\: in \:} R) \rightarrow_{\beta} & \\
+  & {\bf pure \:} ({\bf let \: pure \:} \langle \vec{x}, \vec{w}, \vec{z} \rangle = \langle \vec{M}, \vec{N}, \vec{P} \rangle \: { \bf in } \: R [y := Q])
+  \end{array}$
+
+  8) $\lambda x. f x \rightarrow_{\eta} f$;
+
+  9) $\langle \pi_1 P, \pi_2 P \rangle \rightarrow_{\eta} P$;
+
+  10) ${\bf let \: pure \:} x = M {\: \bf in \: } x \rightarrow_{\eta} M$;
+
+  11) ${\bf pure \:} (\lambda x. f x) \rightarrow_{\eta} {\bf pure \:} f$;
+
+  12) ${\bf pure \:} (\langle \pi_1 P, \pi_2 P \rangle) \rightarrow_{\eta} {\bf pure \:} P$;
+
+  13) ${\bf pure \:} ({\bf let \: pure \:} x = M {\: \bf in \: } x) \rightarrow_{\eta} {\bf pure \:} M$.
   \end{defin}
@@
 
-  Induction on derivation of $\Gamma \vdash M : \alpha$. Let us assume $\Gamma \subseteq \Delta$.
+  Induction on derivation of $\Gamma \vdash M : A$. Let us assume $\Gamma \subseteq \Delta$.
 
@@
   Let $\Sigma = \Theta \setminus \Gamma$, or, which is the same, $\Sigma = \Theta \setminus \Delta, x :
-  A$, then $\Sigma, \Delta, x : A \vdash x : \alpha$, or, $\Theta \vdash x : A$.
+  A$, then $\Sigma, \Delta, x : A \vdash x : A$, or, $\Theta \vdash x : A$.
+
+  \vspace{\baselineskip}
 
@@
 
-  If $\Gamma \vdash {\bf pure \: } M : \textbf{K}A$, then $\Gamma \vdash M : A$ by generation and, by
-  hypothesis, $\Theta \vdash M : A$, so $\Theta \vdash {\bf pure \: } M : \textbf{K}A$ by applying
+  By generation $\Gamma \vdash M : A$
+
+  By hypothesis, $\Theta \vdash M : A$, so $\Theta \vdash {\bf pure \: } M : \textbf{K}A$ by applying
   $\textbf{K}_I$-rule.
 
-  3) Let $\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\bf \: in \:} N : \textbf{K}B$ and $\Gamma \subseteq \Delta$.
-
-  By generation $\Gamma \vdash \vec{M} : \textbf{K}\vec{A}$ and $\vec{x} : \vec{A} \vdash N : \textbf{K}B$.
-
-  By hypothesis we have $\Delta \vdash \vec{M} : \textbf{K}\vec{A}$. So $\Delta \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\bf \: in \:} N : \textbf{K}B$ by ${\bf let}_{\textbf{K}}$.
+    \vspace{\baselineskip}
+
+  3) Let $\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N : {\bf K} B$ and $\Gamma \subseteq \Theta$.
+
+  By generation $\Gamma \vdash \vec{M} : {\bf K} \vec{A}$ and $\vec{x} : \vec{A} \vdash N : B$.
+
+  By assumption $\Theta \vdash \vec{M} : {\bf K} \vec{A}$.
+
+  Hence $\Theta \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N : {\bf K} B$.
 
@@
 
-  Induction by derivation. We consider the base of induction and the case with  ${\bf let}_{\textbf{K}}$. The rest
-  cases are
-  proven by the same way.
-
-  $ $
-
-  1) Let $\Gamma \vdash x : A$, where $\Gamma = \Delta, x : A$, $x \in \mathbb{V}$.
+  $ $
+
+  1) Let $\Gamma \vdash x : A$, where $\Gamma = \Delta, x : \alpha$, $x \in \mathbb{V}$.
 
@@
 
-  2) Let $\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\bf \: in \:} N : \textbf{K}B$.
-
-  By generation $\Gamma \vdash \vec{M} : \textbf{K}\vec{A}$ and $\vec{x} : \vec{A} \vdash N : \textbf{K}B$.
-
-  By hypothesis $\Gamma \uparrow FV(\vec{M}) \vdash \vec{M} : \textbf{K}\vec{A}$.
-
-  So $\Gamma \uparrow FV(\vec{M}) \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\bf \: in \:} N : \textbf{K}B$ by ${\bf let}_{\textbf{K}}$.
+  \vspace{\baselineskip}
+
+  2) Let $\Gamma \vdash {\bf pure \: } M : {\bf K} A$.
+
+  By generation, $\Gamma \vdash M : A$ and $\Gamma \uparrow FV(M) \vdash M : A$ by hypothesis.
+
+  So $\Gamma \uparrow FV(M) \vdash {\bf pure \: } M : {\bf K} A$ by ${\bf K}_I$.
+
+  \vspace{\baselineskip}
+
+  3) Let $\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N : {\bf K} B$.
+
+By generation, $\Gamma \vdash \vec{M} : {\bf K} \vec{A}$ and $\vec{x} : \vec{A} \vdash N : B$.
+
+By assumption, $\Gamma \uparrow FV(\vec{M}) \vdash \vec{M} : {\bf } \vec{A}$.
+
+By $\text{let}_{\bf K}$, $\Gamma \uparrow FV(\vec{M}) \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N : {\bf K} B$
 
@@
 
-  If $\Gamma, x : A \vdash M : B$  and $\Gamma \vdash N : A$, then $\Gamma \vdash (M [x := N]) :
-  B $
+  If $\Gamma, x : A \vdash M : B$  and $\Gamma \vdash N : A$, then $\Gamma \vdash (M [x := N]) : B$
 
@@
 
-  If $\Gamma, x : \alpha \vdash {\bf pure \: } M : \textbf{K}\beta$.
-
   By generation, $\Gamma, x : A \vdash M : B$.
 
-  So, by induction hypothesis, $\Gamma \vdash (M [x := N]) : B $.
-  Then $\Gamma \vdash {\bf pure \: } (M [x := N]) : \textbf{K}B$ by $\textbf{K}_I$, but ${\bf pure \: } (M [x := N]) =
-  ({\bf pure \: } M [x := N])$ by substitution definition.
-
-  So $\Gamma \vdash ({\bf pure \: } M [x := N]) : \textbf{K}B$
-
-  2) Let $\Gamma, y : C \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\bf \: in \:} N : \textbf{K}B$ and $\Gamma \vdash P : C$.
-
-  By generation $\Gamma, y : C \vdash \vec{M} : \textbf{K}\vec{A}$ and $\vec{x} : \vec{A} \vdash N : \textbf{K}B$.
-
-  By assumption $\Gamma \vdash \overrightarrow{M [y := P]} : \textbf{K}A$.
-
-  So $\Gamma \vdash {\bf let \: pure \:} \vec{x} = \overrightarrow{M [y := P]} {\bf \: in \:} N : \textbf{K}B$ by ${\bf let}_{\textbf{K}}$.
-
-  \end{proof}
-
-  \vspace{\baselineskip}
-
-  \begin{defin} Type substitution
-
-    The substitution of type $C$ for type variable $B$ in some type $A$ is defined by:
-
-    1) $A [ A := B ] = B$;
-
-    2) $A^{'} [A := C] = A^{'}$;
-
-    3) $(A_1 \alpha A_2) [B := C] = A_1 [B := C] \alpha A_2 [A := B]$, where $\alpha \{ \times, \to \}$;
-
-    4) $(\textbf{K}A) [B := C] = \textbf{K}(A [B := C])$.
-
-  \end{defin}
-
-    \vspace{\baselineskip}
-
-  \begin{lemma} Substitution for types.
-
-    Let $\Gamma \vdash M : A$, then $\Gamma [B := C] \vdash M : A [B := C]$, where $\Gamma [B := C]$ denotes $\{ x_1 : A_1 [B := C],\dots, x_n : A_n [B := C] \}$.
-
-  \end{lemma}
-
-  1) Let $\Gamma, x : A \vdash x : A$, so $\Gamma [B := C], x : A [B := C] \vdash x : A [B := C]$.
-
-  \vspace{\baselineskip}
-
-  2) Let $\Gamma \vdash {\bf pure \: } M : \textbf{K}A$. By generation $\Gamma \vdash M : A$.
-
-  By assumption $\Gamma [B := C] \vdash M : A [B := C]$.
-
-  By $\textbf{K}_I$, $\Gamma [B := C] \vdash {\bf pure \:} M : \textbf{K} (A [B := C])$.
-
-  But, by substitution definition, $\textbf{K} (A [B := C]) = (\textbf{K}A) [B := C]$.
-
-  Then $\Gamma [B := C] \vdash {\bf pure \:} M : (\textbf{K}A) [B := C]$.
-
-  3) $\Gamma \vdash {\bf let \: pure} \: \vec{x} = \vec{M} \: {\bf in} \: M : {\bf K} B$.
-
-  By generation $\Gamma \vdash \vec{M} : \textbf{K}\vec{A}$ and $\vec{x} : \vec{A} \vdash N : \textbf{K}B$.
-
-  $\Gamma [D := C] \vdash \vec{M} : \textbf{K}\overrightarrow{A [D := C]}$ and $\vec{x} : \overrightarrow{A [D := C]} \vdash M : B [D := C]$ by assumption.
-
-  Then $\Gamma [D := C] \vdash {\bf let \: pure } \: \vec{x} = \vec{M} \: {\bf in} \: M : {\bf } (B [D := C])$.
+  By assumption, $\Gamma \vdash (M [x := N]) : B$
+
+  Then $\Gamma \vdash {\bf pure \: } (M [x := N]) : \textbf{K}B$ by $\textbf{K}_I$.
+
+  But ${\bf pure \: } (M [x := N]) = ({\bf pure \: } M [x := N]) $ by substitution definition, so
+  $\Gamma \vdash ({\bf pure \: } M [x := N]) : \textbf{K}B$
+
+  \vspace{\baselineskip}
+
+  2) Let $\Gamma, y : A \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N : {\bf K} B$ and $\Gamma \vdash N : A$.
+
+  By generation, $\Gamma, y : A \vdash \vec{M} : {\bf K} \vec{A}$ and $\vec{x} : \vec{A} \vdash N : B$.
+
+  By hypothesis, $\Gamma \vdash \vec{M}[x := N] : {\bf K} \vec{A}$.
+
+  Hence $\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M}[x := N] {\: \bf in \:} N : {\bf K} B$.
+
+  \end{proof}
+
+  \vspace{\baselineskip}
 
@@
 
-  \begin{proof}
-
+We consider only modal $\beta$-reduction rules. The general statement for $\twoheadrightarrow_{\beta}$ follows from transitivity of multi-step $\beta$-reduction.
+
+  \begin{proof}
+
+  $ $
+
+1) Let $\Gamma \vdash {\bf let \: pure \:} \langle \vec{x}, y, \vec{z} \rangle = \langle \vec{M}, {\bf let \: pure \: } \vec{w} = \vec{N} {\bf \: in \: } Q, \vec{P} \rangle {\: in \:} R : {\bf K }B$
+
+By generation we have $\Gamma \vdash \vec{M} : {\bf K}\vec{A_1}$, $\Gamma \vdash {\bf let \: pure \: } \vec{w} = \vec{N} {\bf \: in \: } Q : {\bf K}\vec{A_2}$, $\Gamma \vdash \vec{P} : {\bf K}\vec{A_3}$
+and $\vec{x} : \vec{A_1}, y : A_2, \vec{z} : \vec{A_3} \vdash R : B$.
+
+If $\Gamma \vdash {\bf let \: pure \: } \vec{w} = \vec{N} {\bf \: in \: } Q : {\bf K}\vec{A_2}$, then
+
+$\Gamma \vdash \vec{N} : {\bf K}\vec{A_4}$ and $\vec{w} : \vec{A_4} \vdash Q : A_2$.
+
+Then $\vec{x} : \vec{A_1}, \vec{w} : \vec{A_4}, \vec{z} : \vec{A_3} \vdash R [y := Q] : B$ by substitution lemma and weakening.
+
+Hence $\Gamma \vdash {\bf let \: pure \:} \langle \vec{x}, \vec{w}, \vec{z} \rangle = \langle \vec{M}, \vec{N}, \vec{P} \rangle \: { \bf in } \: R [y := Q] : {\bf K} B$
+by $let_{{\bf K}}$.
+
+  \vspace{\baselineskip}
+
+2) Let $\Gamma \vdash {\bf pure } ((\lambda x. M) N) : {\bf K} B$.
+
+By generation $\Gamma \vdash (\lambda x. M) N : B$, but $\Gamma \vdash M [x := N] : B$, then, by ${\bf K}_I$, $\Gamma \vdash {\bf pure \:} (M [x := N]) : {\bf K} B$.
+
+  \vspace{\baselineskip}
+
+3) Let $\Gamma \vdash {\bf pure } (\pi_i \langle M_1, M_2 \rangle) : {\bf K} A_i$, where $i \in \{ 1,2 \}$.
+
+By generation $\Gamma \vdash \pi_i \langle M_1, M_2 \rangle : A_i$ and $\Gamma \vdash M_i : A_i$.
+
+Hence $\Gamma \vdash {\bf pure \: } M_i : { \bf K } A_i$ by ${\bf K}_I$
   \end{proof}
@@
   the Foundations of Mathematics, vol. 103. Amsterdam: North-Holland, 1985.
+
+  \bibitem{ModalK} Yoshihiko KAKUTANI, A Curry-Howard Correspondence for Intuitionistic Normal Modal Logic, Computer Software, Released February 29, 2008, Online ISSN , Print ISSN 0289-6540.
   \end{thebibliography}

### 2018-01-16 article.tex
finished subject reduction
added some beauty
added eta rules
extended proof
added interpretation for typing rules

@@
     \RightLabel{\scriptsize{ax}}
-  \UnaryInfC{$\Gamma , x : \alpha \vdash x : \alpha$}
+    \UnaryInfC{$\Gamma , x : A \vdash x : A$}
     \end{prooftree}
@@
 
-  \begin{minipage}{0.5\textwidth}
-    \begin{flushleft}
-  \begin{prooftree}
-  \AxiomC{$\Gamma, x : \alpha \vdash M : \beta$}
+    \begin{minipage}{0.45\textwidth}
+      \begin{prooftree}
+      \AxiomC{$\Gamma, x : A \vdash M : B$}
       \RightLabel{$\rightarrow_i$}
-  \UnaryInfC{$\Gamma \vdash \lambda x. M : \alpha \to \beta$}
-  \end{prooftree}
-
-  \begin{prooftree}
-  \AxiomC{ $\Gamma \vdash x : \alpha$ }
-  \AxiomC{ $\Gamma \vdash y : \beta$ }
+      \UnaryInfC{$\Gamma \vdash \lambda x. M : A \to B$}
+      \end{prooftree}
+
+      \begin{prooftree}
+      \AxiomC{ $\Gamma \vdash M : A$ }
+      \AxiomC{ $\Gamma \vdash N : B$ }
       \RightLabel{$\times_i$}
-  \BinaryInfC{$\Gamma \vdash (x, y) : \alpha \times \beta$}
-  \end{prooftree}
-
-  \begin{prooftree}
-    \AxiomC{$\Gamma \vdash x : \alpha$}
+      \BinaryInfC{$\Gamma \vdash \langle x, y \rangle : A \times B$}
+      \end{prooftree}
+
+      \begin{prooftree}
+        \AxiomC{$\Gamma \vdash x : A$}
         \RightLabel{$\textbf{K}_I$}
-    \UnaryInfC{$\Gamma \vdash {\bf pure \: } \: x : \textbf{K}\alpha $}
-  \end{prooftree}
-    \end{flushleft}
-  \end{minipage}
-
-  \begin{minipage}{0.5\textwidth}
-    \begin{flushright}
-  \begin{prooftree}
-  \AxiomC{$\Gamma \vdash f : \alpha \to \beta$}
-  \AxiomC{$\Gamma \vdash x : \alpha$}
+        \UnaryInfC{$\Gamma \vdash {\bf pure \: } \: x : {\bf K}A $}
+      \end{prooftree}
+  \end{minipage}%
+  \hfill
+  \begin{minipage}{0.45\textwidth}
+  \begin{tabular}{p{\textwidth}}
+    \begin{prooftree}
+    \AxiomC{$\Gamma \vdash f : A \to B$}
+    \AxiomC{$\Gamma \vdash x : A$}
     \RightLabel{$\rightarrow_e$}
-  \BinaryInfC{$\Gamma \vdash fx : \beta$}
-  \end{prooftree}
-
-  \begin{prooftree}
-  \AxiomC{ $\Gamma \vdash p : \alpha_1 \times \alpha_2$ }
+    \BinaryInfC{$\Gamma \vdash fx : B$}
+    \end{prooftree}
+
+    \begin{prooftree}
+    \AxiomC{ $\Gamma \vdash M : A_1 \times A_2$ }
     \RightLabel{$\times_e$, $i \in \{ 1, 2 \}$}
-  \UnaryInfC{$\Gamma \vdash \pi_i p : \alpha_i$}
+    \UnaryInfC{$\Gamma \vdash \pi_i M : A_i$}
     \end{prooftree}
@@
     \end{prooftree}
-    \end{flushright}
-  \end{minipage}
+  \end{tabular}
+  \end{minipage}%
+
   \end{defin}
@@
 
-  10) ${\bf let \: pure \:} x = M {\: \bf in \: } x \rightarrow_{\eta} M$;
-
-  11) ${\bf pure \:} (\lambda x. f x) \rightarrow_{\eta} {\bf pure \:} f$;
-
-  12) ${\bf pure \:} (\langle \pi_1 P, \pi_2 P \rangle) \rightarrow_{\eta} {\bf pure \:} P$;
-
-  13) ${\bf pure \:} ({\bf let \: pure \:} x = M {\: \bf in \: } x) \rightarrow_{\eta} {\bf pure \:} M$.
+  10) ${\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} N \rightarrow_{\eta} {\bf pure \:} N$;
+
+  11) ${\bf let \: pure \:} x = M {\: \bf in \: } x \rightarrow_{\eta} M$;
+
+  12) ${\bf pure \:} (\lambda x. f x) \rightarrow_{\eta} {\bf pure \:} f$;
+
+  13) ${\bf pure \:} (\langle \pi_1 P, \pi_2 P \rangle) \rightarrow_{\eta} {\bf pure \:} P$;
+
+  14) ${\bf pure \:} ({\bf let \: pure \:} x = M {\: \bf in \: } x) \rightarrow_{\eta} {\bf pure \:} M$;
+
+  15) ${\bf pure \:} ({\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} N) \rightarrow_{\eta} { \bf pure \:} ({\bf pure \:} N)$.
   \end{defin}
@@
 
-  1) Let $\Gamma \vdash x : A$, where $\Gamma = \Delta, x : \alpha$, $x \in \mathbb{V}$.
+  1) Let $\Gamma \vdash x : A$, where $\Gamma = \Delta, x : A$, $x \in \mathbb{V}$.
 
@@
 
-  Let $\Gamma \vdash M : \alpha$ and $M \twoheadrightarrow_{\beta} N$, then $\Gamma \vdash N : \alpha$
-
+  i) Let $\Gamma \vdash M : A$ and $M \twoheadrightarrow_{\beta} N$, then $\Gamma \vdash N : A$
+
+  ii) Let $\Gamma \vdash M : A$ and $M \twoheadrightarrow_{\eta} N$, then $\Gamma \vdash N : A$
   \end{theorem}
@@
 
-1) Let $\Gamma \vdash {\bf let \: pure \:} \langle \vec{x}, y, \vec{z} \rangle = \langle \vec{M}, {\bf let \: pure \: } \vec{w} = \vec{N} {\bf \: in \: } Q, \vec{P} \rangle {\: in \:} R : {\bf K }B$
-
-By generation we have $\Gamma \vdash \vec{M} : {\bf K}\vec{A_1}$, $\Gamma \vdash {\bf let \: pure \: } \vec{w} = \vec{N} {\bf \: in \: } Q : {\bf K}\vec{A_2}$, $\Gamma \vdash \vec{P} : {\bf K}\vec{A_3}$
+  i) For multistep $\beta$-reduction:
+
+  \vspace{\baselineskip}
+
+1) Let $\Gamma \vdash {\bf let \: pure \:} \langle \vec{x}, y, \vec{z} \rangle = \langle \vec{M}, {\bf let \: pure \: } \vec{w} = \vec{N} { \: \bf in \: } Q, \vec{P} \rangle { \: \bf in \: } R : {\bf K }B$
+
+By generation we have $\Gamma \vdash \vec{M} : {\bf K}\vec{A_1}$, $\Gamma \vdash {\bf let \: pure \: } \vec{w} = \vec{N} {\: \bf in \:} Q : {\bf K}\vec{A_2}$, $\Gamma \vdash \vec{P} : {\bf K}\vec{A_3}$
 and $\vec{x} : \vec{A_1}, y : A_2, \vec{z} : \vec{A_3} \vdash R : B$.
@@
 
-Hence $\Gamma \vdash {\bf pure \: } M_i : { \bf K } A_i$ by ${\bf K}_I$
+Hence $\Gamma \vdash {\bf pure \: } M_i : { \bf K } A_i$ by ${\bf K}_I$.
+
+\vspace{\baselineskip}
+
+4) Let $\Gamma \vdash {\bf let \: pure \:} (\langle \vec{x}, y, \vec{z} \rangle = \langle \vec{M}, {\bf let \: pure \: } \vec{w} = \vec{N} {\bf \: in \: } Q, \vec{P} \rangle { \: \bf in \: } R) : {\bf K }^2 B$.
+
+By generation $\Gamma \vdash {\bf let \: pure \:} \langle \vec{x}, y, \vec{z} \rangle = \langle \vec{M}, {\bf let \: pure \: } \vec{w} = \vec{N} {\bf \: in \: } Q, \vec{P} \rangle { \: \bf in \: } R : {\bf K }B$,
+
+hence $\Gamma \vdash {\bf let \: pure \:} \langle \vec{x}, \vec{w}, \vec{z} \rangle = \langle \vec{M}, \vec{N}, \vec{P} \rangle \: { \bf in } \: R [y := Q] : {\bf K} B$ by the first case.
+
+So $\Gamma \vdash {\bf pure \:} ({\bf let \: pure \:} \langle \vec{x}, \vec{w}, \vec{z} \rangle = \langle \vec{M}, \vec{N}, \vec{P} \rangle \: { \bf in } \: R [y := Q]) : {\bf K}^2 B$
+by ${\bf K}_I$.
+
+\vspace{\baselineskip}
+
+ii) For multistep $\eta$-reduction:
+
+\vspace{\baselineskip}
+
+  1) Let $\vdash {\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} N : {\bf K} A$.
+
+  Then by generation $\vdash N : A$, so $\vdash {\bf pure \:} N : {\bf K} A$ by ${\bf K}_I$.
+
+  \vspace{\baselineskip}
+
+  2) Let $\Gamma \vdash {\bf let \: pure \:} x = M {\: \bf in \: } x : {\bf K} A$.
+
+  By generation $\Gamma \vdash M : {\bf K} A$ and $x : A \vdash x : A$, hence $\Gamma \vdash M : {\bf K} A$.
+
+  \vspace{\baselineskip}
+
+  3) Let $\Gamma \vdash {\bf pure \:} (\lambda x. f x) : {\bf K}(A \to B)$.
+
+  By generation $\Gamma \vdash \lambda x. f x : A \to B$, so $\Gamma \vdash f : A \to B$, then $\Gamma \vdash {\bf pure \:} f : {\bf K}(A \to B)$ by ${\bf K}_I$.
+
+  \vspace{\baselineskip}
+
+  4) Let $\Gamma \vdash {\bf pure \:} (\langle \pi_1 P, \pi_2 P \rangle) : {\bf K}(A \times B)$.
+
+  By generation $\Gamma \vdash \langle \pi_1 P, \pi_2 P \rangle : A \times B$, then $\Gamma \vdash P : A \times B$.
+
+  By ${\bf K}_I$, $\Gamma \vdash {\bf pure \:} P : {\bf K}(A \times B)$.
+
+  \vspace{\baselineskip}
+
+  5) $\Gamma \vdash {\bf pure \:} ({\bf let \: pure \:} x = M {\: \bf in \: } x) : {\bf K}^2 A$.
+
+  Then $\Gamma \vdash M : {\bf K}^2 A$ and $x : {\bf K} A \vdash x : {\bf K} A$, so $\Gamma \vdash M : {\bf K}^2 A$.
+
+  \vspace{\baselineskip}
+
+  6) Let $\vdash {\bf pure \:} ({\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} N) : {\bf K}^2 A$.
+
+  By generation ${\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} N : {\bf K}A$, so $\vdash N : A$, then $\vdash {\bf pure \:} N : {\bf K}$.
+
   \end{proof}
@@
 
+\section{Categorical semantics}
+
+\begin{defin} Lax monoidal functor
+
+  Let $\langle \mathcal{C}, \oplus_1, \mathds{1} \rangle$ and $\langle \mathcal{D}, \oplus_2, \mathds{1}' \rangle$ are monoidal categories.
+
+  A lax monoidal functor $\mathcal{F} : \langle \mathcal{C}, \oplus_1, \mathds{1} \rangle \to \langle \mathcal{D}, \oplus_2, \mathds{1}' \rangle$ is a functor
+  $\mathcal{F} : \mathcal{C} \to \mathcal{D}$ with additional natural transformations:
+
+  1) $u : \mathds{1}' \to \mathcal{F}\mathds{1}$;
+
+  2) $\ast_{A, B} \mathcal{F}A \otimes_2 \mathcal{F}B \to \mathcal{F}(A \otimes_1 B)$
+
+\end{defin}
+
+\begin{defin} Applicative functor
+
+  An applicative functor is a triple $\langle \mathcal{C}, \mathcal{K}, \eta \rangle$,
+where $\mathcal{C}$ is a symmetric monoidal category, $\mathcal{K}$ is a lax monoidal endofunctor and $\eta$ is a natural transformation, such that:
+
+1) $u = \eta_{\mathds{1}}$;
+
+2) $\ast_{A,B} \circ (\eta_A \otimes \eta_B) = \eta_{A \otimes B}$;
+
+3) Weak commutativity condition holds:
+
+\xymatrix{
+A \otimes \mathcal{K}B & \mathcal{K}A \otimes \mathcal{K}B & \mathcal{K}(A \otimes B) \\
+\mathcal{K}B \otimes A & \mathcal{K}B \otimes \mathcal{K}A & \mathcal{K}(B \otimes A)
+}
+\end{defin}
+
+By default we will consider an arbitrary closed functor on some cartersian closed category, which is the special case of an applicative functor.
+
+\subsection{Soundness}
+
+\begin{defin} Semantical translation from $\lambda_{\textbf{K}}$ to CCC with applicative functor $\mathcal{K}$:
+
+1) Interpretation for types:
+
+$[\![A]\!] := \hat{A}, A \in \mathbb{T}$;
+
+$[\![A \to B]\!] := [\![A]\!] \to [\![B]\!]$;
+
+$[\![A \times B]\!] := [\![A]\!] \times [\![B]\!]$.
+
+2) Interpretation for modal types: $[\![\textbf{K}A]\!] = \mathcal{K}[\![A]\!]$;
+
+3) Interpretaion for contexts:
+
+$[\![\Gamma = \{ x_1 : A_1, ..., x_n : A_n\}]\!] := [\![\Gamma]\!] = [\![A_1]\!] \times ... \times [\![A_n]\!]$;
+
+4) Interpretation for typing assignment: $[\![\Gamma \vdash M : A]\!] := [\![M]\!] : [\![\Gamma]\!] \to
+[\![A]\!]$.
+
+5) Interpretation for typing rules:
+
+\begin{prooftree}
+\AxiomC{$ $}
+\UnaryInfC{$[\![\Gamma, x : A \vdash x : A]\!] = \pi_2 : [\![\Gamma]\!] \times [\![A]\!] \rightarrow
+[\![A]\!]$}
+\end{prooftree}
+
+\begin{prooftree}
+\AxiomC{$[\![\Gamma, x : A \vdash M : B]\!] = f : [\![\Gamma]\!] \times [\![A]\!] \rightarrow [\![B]\!]$}
+\UnaryInfC{$[\![\Gamma \vdash (\lambda x. M) : A \to B]\!] = \Lambda(f) : [\![\Gamma]\!]
+\rightarrow[\![B]\!]^{[\![A]\!]}$}
+\end{prooftree}
+
+\begin{prooftree}
+\AxiomC{$[\![\Gamma \vdash M : A \to B]\!] = [\![M]\!] : [\![\Gamma]\!] \rightarrow [\![B]\!]^{[\![A]\!]}$}
+\AxiomC{$[\![\Gamma \vdash N : A]\!] = [\![N]\!] : [\![\Gamma]\!] \rightarrow [\![A]\!]$}
+\BinaryInfC{$[\![\Gamma \vdash (M N) : B]\!] = [\![\Gamma]\!] \xrightarrow{\langle [\![M]\!], [\![N]\!]
+\rangle} [\![B]\!]^{[\![A]\!]} \times [\![A]\!] \xrightarrow{\epsilon} [\![B]\!] $}
+\end{prooftree}
+
+\begin{prooftree}
+\AxiomC{$[\![\Gamma \vdash M : A ]\!] = f : [\![\Gamma]\!] \rightarrow [\![A]\!]$}
+\AxiomC{$[\![\Gamma \vdash N : B ]\!] = g : [\![\Gamma]\!] \rightarrow [\![B]\!]$}
+\BinaryInfC{$[\![\Gamma \vdash (M, N) : A \times B]\!] = \langle f, g \rangle : [\![\Gamma]\!] \rightarrow
+[\![A]\!] \times [\![B]\!]$}
+\end{prooftree}
+
+\begin{prooftree}
+\AxiomC{$[\![\Gamma \vdash p : A_1 \times A_2]\!] = f : [\![\Gamma]\!] \rightarrow [\![A_1]\!] \times
+[\![A_2]\!]$}
+\RightLabel{$i \in \{1,2\}$}
+\UnaryInfC{$[\![\Gamma \vdash \pi_i p : A_i]\!] = [\![\Gamma]\!] \xrightarrow{f} [\![A_1]\!] \times
+[\![A_2]\!] \xrightarrow{\pi_i} [\![A_i]\!]$}
+\end{prooftree}
+
+\begin{prooftree}
+\AxiomC{$[\![\Gamma \vdash M : A]\!] = [\![M]\!] : [\![\Gamma]\!] \rightarrow [\![A]\!]$}
+\UnaryInfC{$[\![\Gamma \vdash {\bf pure \:} M : \textbf{K}A]\!] := [\![\Gamma]\!] \xrightarrow{[\![M]\!]}
+[\![A]\!] \xrightarrow{\eta_{[\![A]\!]}} \mathcal{K}[\![A]\!]$}
+\end{prooftree}
+
+  \begin{prooftree}
+    \AxiomC{$[\![\Gamma \vdash \vec{M} : {\bf K} \vec{A}]\!] = \langle [\![M_1]\!],\dots, [\![M_n]\!] \rangle : [\![\Gamma]\!] \rightarrow \prod \limits_{i=i}^n \mathcal{K}[\![A_i]\!]$}
+    \AxiomC{$[\![\vec{x} : \vec{A} \vdash N : B]\!] = [\![N]\!] : \prod \limits_{i=i}^n [\![A_i]\!] \rightarrow [\![B]\!]$}
+    \BinaryInfC{$[\![\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } M : {\bf K} B]\!] = \mathcal{K}([\![N]\!]) \circ \ast_{[\![A_1]\!],\dots,[\![A_n]\!]} \circ \langle [\![M_1]\!],\dots, [\![M_n]\!] \rangle : [\![\Gamma]\!] \rightarrow \mathcal{K}[\![B]\!]$}
+  \end{prooftree}
+
+\end{defin}
+
+\begin{defin} Simultaneous substitution
+
+Let $\Gamma = \{ x_1 : A_1, ..., x_n : A_n \}$, $\Gamma \vdash M : A$ and for all $i \in \{ 1,..., n \}$,
+$\Gamma \vdash M_i : A_i$.
+
+We define simultaneous substitution $M [ \vec{x} := \vec{M}]$ recursively by:
+
+1) $x_i [ \vec{x} := \vec{M}] := M_i $;
+
+2) $(\lambda x. M) [ \vec{x} := \vec{M}] := \lambda x. (M [ \vec{x} := \vec{M}])$;
+
+3) $(M N) [ \vec{x} := \vec{M}] = (M [ \vec{x} := \vec{M}]) (N [ \vec{x} := \vec{M}])$;
+
+4) $\langle M, N \rangle = \langle (M [ \vec{x} := \vec{M}]), (N [ \vec{x} := \vec{M}])\rangle$;
+
+5) $(\pi_i P) [ \vec{x} := \vec{M}] = \pi_i (P [ \vec{x} := \vec{M}])$;
+
+6) $(\text{pure } M) [ \vec{x} := \vec{M}] = \text{pure } (M [ \vec{x} := \vec{M}])$;
+
+\end{defin}
+
+\begin{lemma}
+$ $
+
+$[\![M [x_1 := M_1,\dots, x_n := M_n]]\!] = [\![M]\!] \circ \langle [\![M_1]\!], \dots, [\![M_n]\!] \rangle$.
+
+\end{lemma}
+
+\begin{proof}
+
+$ $
+
+1) $[\![(\text{pure } M) [ \vec{x} := \vec{M}]]\!] = [\![\text{pure } M]\!] \circ \langle [\![M_1]\!], \dots,
+[\![M_n]\!] \rangle$.
+
+\vspace{\baselineskip}
+
+$\begin{array}{lll}
+& [\![(\text{pure } M) [ \vec{x} := \vec{M}]]\!] = [\![\text{pure } (M [ \vec{x} := \vec{M}])]\!] &
+\text{Substitution definition} \\
+& \quad\quad\quad\quad\quad\quad\quad = p \circ [\![(M [ \vec{x} := \vec{M}])]\!]&
+\text{Translation for pure}\\
+& \quad\quad\quad\quad\quad\quad\quad = p \circ [\![M]\!] \circ \langle [\![M_1]\!], \dots,
+[\![M_n]\!] \rangle & \text{Induction hypothesis} \\
+& \quad\quad\quad\quad\quad\quad\quad = (p \circ [\![M]\!]) \circ \langle [\![M_1]\!], \dots,
+[\![M_n]\!] \rangle & \text{Associativity of composition}\\
+& \quad\quad\quad\quad\quad\quad\quad = [\![\text{pure } M]\!] \circ \langle [\![M_1]\!], \dots,
+[\![M_n]\!] \rangle& \text{Translation for pure}\\
+\end{array}$
+
+\vspace{\baselineskip}
+
+\end{proof}
+
+\begin{theorem} Soundness
+
+  Let $\Gamma \vdash M : A$ and $M =_{\beta\eta} N$, then $[\![\Gamma \vdash M : A]\!] = [\![\Gamma \vdash N : A]\!]$
+\end{theorem}
+
   \section{Acknowledgement.}

### 2018-01-17 article.tex
added lemma proof

@@
   \RightLabel{\scriptsize{ax}}
-  \UnaryInfC{$\Gamma , \alpha \vdash A$}
+  \UnaryInfC{$\Gamma , A \vdash A$}
   \end{prooftree}
@@
 
-  4) $FV((M,N)) = FV(M) \cup FV(N)$;
-
-  5) $FV(\pi_i p) \subseteq FV(p)$, $i \in \{ 1, 2\}$;
+  4) $FV(\langle M,N \rangle) = FV(M) \cup FV(N)$;
+
+  5) $FV(\pi_i M) \subseteq FV(M)$, $i \in \{ 1, 2\}$;
 
@@
 
-  7) $({\bf let \:\: pure} \:\: \vec{N} = \vec{M} \:\: {\bf in} \:\: M) [x := P] = {\bf let \:\: pure} \:\: \vec{N} = (\vec{M} [x := P]) \:\: {\bf in} \:\: M$.
+  7) $({\bf let \:\: pure} \:\: \vec{x} = \vec{M} \:\: {\bf in} \:\: M) [y := P] = {\bf let \:\: pure} \:\: \vec{x} = (\vec{M} [y := P]) \:\: {\bf in} \:\: M$.
   \end{defin}
@@
   \begin{prooftree}
-    \AxiomC{$[\![\Gamma \vdash \vec{M} : {\bf K} \vec{A}]\!] = \langle [\![M_1]\!],\dots, [\![M_n]\!] \rangle : [\![\Gamma]\!] \rightarrow \prod \limits_{i=i}^n \mathcal{K}[\![A_i]\!]$}
-    \AxiomC{$[\![\vec{x} : \vec{A} \vdash N : B]\!] = [\![N]\!] : \prod \limits_{i=i}^n [\![A_i]\!] \rightarrow [\![B]\!]$}
+    \AxiomC{$[\![\Gamma \vdash \vec{M} : {\bf K} \vec{A}]\!] = \langle [\![M_1]\!],\dots, [\![M_n]\!] \rangle : [\![\Gamma]\!] \rightarrow \prod \limits_{i=1}^n \mathcal{K}[\![A_i]\!]$}
+    \AxiomC{$[\![\vec{x} : \vec{A} \vdash N : B]\!] = [\![N]\!] : \prod \limits_{i=1}^n [\![A_i]\!] \rightarrow [\![B]\!]$}
     \BinaryInfC{$[\![\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } M : {\bf K} B]\!] = \mathcal{K}([\![N]\!]) \circ \ast_{[\![A_1]\!],\dots,[\![A_n]\!]} \circ \langle [\![M_1]\!],\dots, [\![M_n]\!] \rangle : [\![\Gamma]\!] \rightarrow \mathcal{K}[\![B]\!]$}
@@
 
-1) $x_i [ \vec{x} := \vec{M}] := M_i $;
-
-2) $(\lambda x. M) [ \vec{x} := \vec{M}] := \lambda x. (M [ \vec{x} := \vec{M}])$;
-
-3) $(M N) [ \vec{x} := \vec{M}] = (M [ \vec{x} := \vec{M}]) (N [ \vec{x} := \vec{M}])$;
-
-4) $\langle M, N \rangle = \langle (M [ \vec{x} := \vec{M}]), (N [ \vec{x} := \vec{M}])\rangle$;
-
-5) $(\pi_i P) [ \vec{x} := \vec{M}] = \pi_i (P [ \vec{x} := \vec{M}])$;
-
-6) $(\text{pure } M) [ \vec{x} := \vec{M}] = \text{pure } (M [ \vec{x} := \vec{M}])$;
+1) $x_i [ \vec{x} := \vec{M}] = M_i $;
+
+2) $(\lambda x. M) [ \vec{x} := \vec{M}] = \lambda x. (M [ \vec{x} := \vec{M}])$;
+
+3) $(M N) [ \vec{x} := \vec{M}] = (M [ \vec{x} = \vec{M}]) (N [ \vec{x} := \vec{M}])$;
+
+4) $\langle M, N \rangle = \langle (M [ \vec{x} = \vec{M}]), (N [ \vec{x} := \vec{M}])\rangle$;
+
+5) $(\pi_i P) [ \vec{x} := \vec{M}] = \pi_i (P [ \vec{x} = \vec{M}])$;
+
+6) $({\bf pure \:} M) [ \vec{x} := \vec{M}] = {\bf pure \:} (M [ \vec{x} = \vec{M}])$;
+
+7) $({\bf let \: pure} \: \vec{x} = \vec{M} {\: \bf in \:} N) [\vec{y} := \vec{P}] =
+{\bf let \: pure} \: \vec{x} = (\vec{M} [\vec{y} := \vec{P}]) {\: \bf in \:} N$
 
@@
 
-1) $[\![(\text{pure } M) [ \vec{x} := \vec{M}]]\!] = [\![\text{pure } M]\!] \circ \langle [\![M_1]\!], \dots,
+1) $[\![\Gamma \vdash ({\bf pure \:} M) [ \vec{x} := \vec{M}] : {\bf K}A ]\!] = [\![ \Gamma \vdash {\bf pure \:} M : {\bf K}A]\!] \circ \langle [\![M_1]\!], \dots,
 [\![M_n]\!] \rangle$.
@@
 $\begin{array}{lll}
-& [\![(\text{pure } M) [ \vec{x} := \vec{M}]]\!] = [\![\text{pure } (M [ \vec{x} := \vec{M}])]\!] &
+& [\![\Gamma \vdash ({\bf pure \:} M) [ \vec{x} := \vec{M}] : {\bf K}A]\!] = [\![ \Gamma \vdash {\bf pure \:} (M [ \vec{x} := \vec{M}]) : {\bf K}A]\!] &
 \text{Substitution definition} \\
-& \quad\quad\quad\quad\quad\quad\quad = p \circ [\![(M [ \vec{x} := \vec{M}])]\!]&
+& \quad\quad\quad\quad\quad\quad\quad = \eta_{[\![A]\!]} \circ [\![(M [ \vec{x} := \vec{M}])]\!]&
 \text{Translation for pure}\\
-& \quad\quad\quad\quad\quad\quad\quad = p \circ [\![M]\!] \circ \langle [\![M_1]\!], \dots,
-[\![M_n]\!] \rangle & \text{Induction hypothesis} \\
-& \quad\quad\quad\quad\quad\quad\quad = (p \circ [\![M]\!]) \circ \langle [\![M_1]\!], \dots,
+& \quad\quad\quad\quad\quad\quad\quad = \eta_{[\![A]\!]} \circ ([\![M]\!] \circ \langle [\![M_1]\!], \dots,
+[\![M_n]\!] \rangle) & \text{Induction hypothesis} \\
+& \quad\quad\quad\quad\quad\quad\quad = (\eta_{[\![A]\!]} \circ [\![M]\!]) \circ \langle [\![M_1]\!], \dots,
 [\![M_n]\!] \rangle & \text{Associativity of composition}\\
-& \quad\quad\quad\quad\quad\quad\quad = [\![\text{pure } M]\!] \circ \langle [\![M_1]\!], \dots,
+& \quad\quad\quad\quad\quad\quad\quad = [\![ \Gamma \vdash {\bf pure \:} M : {\bf K}A]\!] \circ \langle [\![M_1]\!], \dots,
 [\![M_n]\!] \rangle& \text{Translation for pure}\\
+\end{array}$
+
+\vspace{\baselineskip}
+
+2) $\begin{array}{lll}&[\![\Gamma \vdash ({\bf let \: pure} \: \vec{x} = \vec{M} {\: \bf in \:} N) [\vec{y} := \vec{P}] : {\bf K} B]\!]
+=[\![\Gamma \vdash {\bf let \: pure} \: \vec{x} = \vec{M} {\: \bf in \:} N : {\bf K}B]\!] \circ \langle [\![P_1]\!],\dots,[\![P_n]\!]\rangle& \end{array}$
+
+\vspace{\baselineskip}
+
+$\begin{array}{lll}
+&[\![\Gamma \vdash ({\bf let \: pure} \: \vec{x} = \vec{M} {\: \bf in \:} N) [\vec{y} := \vec{P}] : {\bf K} B]\!] =& \\
+&\text{Substitution definition}& \\
+&[\![\Gamma \vdash {\bf let \: pure} \: \vec{x} = (\vec{M} [\vec{y} := \vec{P}]) {\: \bf in \:} N : {\bf K}B]\!] =& \\
+&\text{Interpretaion for $let_{\bf K}$}& \\
+&\mathcal{K}([\![N]\!]) \circ \ast_{[\![A_1]\!],\dots,[\![A_n]\!]} \circ [\![\Gamma \vdash (\vec{M} [\vec{y} := \vec{P}]) \vdash : {\bf K}\vec{A}]\!] =& \\
+&\text{Induction hypothesis}& \\
+&\mathcal{K}([\![N]\!]) \circ \ast_{[\![A_1]\!],\dots,[\![A_n]\!]} \circ ([\![\vec{M}]\!] \circ \langle [\![P_1]\!],\dots,[\![P_n]\!]\rangle) = & \\
+&\text{Associativity of composition}& \\
+&(\mathcal{K}([\![N]\!]) \circ \ast_{[\![A_1]\!],\dots,[\![A_n]\!]} \circ [\![\vec{M}]\!]) \circ \langle [\![P_1]\!],\dots,[\![P_n]\!]\rangle = & \\
+&\text{By interpretation}&\\
+&[\![\Gamma \vdash ({\bf let \: pure} \: \vec{x} = \vec{M} {\: \bf in \:} N]\!] \circ \langle [\![P_1]\!],\dots,[\![P_n]\!]\rangle&
 \end{array}$

### 2018-01-17 article.tex
added proof and erased some rubbish

@@
 
-  \begin{proof}
-
-  We build the smallest of subset of strongly normalazing terms of modal types and show that an arbitrary
-  term belongs to this subset.
-
-  \begin{defin} The set of strongly computable terms of type $\phi \in \mathbb{T}_{\textbf{K}}$, $SC_{\phi}$:
-
-  \begin{itemize}
-  \item Let $\phi = \textbf{K}\alpha$ and $\alpha \in \mathbb{T}$ , then:
-  \begin{equation}
-  SC_{\textbf{K}\alpha} = \{ M : \textbf{K}\alpha \: | \: \text{$M$ is strongly normalizing} \}
-  \end{equation}
-
-  \item Let $\phi = \textbf{K}(\tau \to \psi)$ and $\tau, \psi \in \mathbb{T}_{\textbf{K}}$, then:
-  \begin{equation}
-  SC_{\textbf{K}(\tau \to \psi)} = \{ M : \textbf{K}(\tau \to \psi) \: | \: \forall N \in SC_{\textbf{K}\tau}, M \star N \in SC_{\textbf{K}\psi}\}
-  \end{equation}
-
-  \item Let $\phi = \textbf{K}(\tau_1 \times \tau_2)$ and $\tau_1, \tau_2 \in \mathbb{T}_{\textbf{K}}$, then:
-  \begin{equation}
-  SC_{\textbf{K}(\tau_1 \times \tau_2)} = \{ P : \textbf{K}(\tau_1 \times \tau_2) \: |
-  \: {\bf pure \:} (\lambda x. \pi_i x) \star P \in SC_{\textbf{K}\tau_i}, i \in \{1,2\} \}
-  \end{equation}
-  \end{itemize}
-  \end{defin}
-
-  \begin{lemma}
-
-  $ $
-
-  If $M \in SC_{\alpha}$, then $M$ is strongly normalizing.
-
-  \end{lemma}
-
-  \begin{proof}
-
-$ $
-
-  1) If $M \in SC_{{\bf K} \alpha}$ and $\alpha \in \mathbb{T}$, then $M$ is strongly normalizing by the
-  definition of $SC_{{\bf K} \alpha}$.
-
-  2) Let $M \in SC_{\textbf{K}(\tau \to \psi)}$, so by every $N \in SC_{\textbf{K}\tau}$,
-  $M \star N \in SC_{\textbf{K}\psi}$, which is strongly normalizing by hypothesis. So $M$ is
-  strongly normalizing.
-
-  3) Let $M \in SC_{\textbf{K}(\tau_1 \times \tau_2)}$, so ${\bf pure \:} (\lambda x. \pi_i x) \star M \in
-  SC_{\textbf{K}\tau_i}$, $i \in \{ 1,2 \}$, which are strongly normalizing. So $M$ is
-  strongly normalizing.
-
-  \end{proof}
-
-  \begin{lemma}
-
-  $ $
-
-  Let $M \rightarrow_{\beta} M'$ and $M \in SC_{\alpha}$, then $M' \in SC_{\alpha}$.
-
-  \end{lemma}
-
-  \begin{proof}
-
-$ $
-
-  1) Let $M \rightarrow_{\beta} M'$ and $M \in SC_{\textbf{K}\alpha}$, where $\alpha \in \mathbb{T}$.
-
-  $M$ has the longest reduction path (which we denote as $p(M)$). So $p(M') < p(M)$, then $M' \in SC_{\textbf{K}\alpha}$.
-
-  2) Let $M \in SC_{\textbf{K}(\alpha \to \beta)}$ and $M \rightarrow_{\beta} M'$. Let $N \in
-  SC_{\textbf{K}\alpha}$. So $M \star N \in SC_{\textbf{K}\beta}$.
-
-  If $M \rightarrow_{\beta} M'$, then $M \star N \rightarrow_{\beta} M' \star N$ by reduction
-  rule, so $M' \star N \in SC_{\textbf{K}\beta}$ and $M' \in SC_{\textbf{K}(\alpha \to \beta)}$ by
-  hypothesis.
-
-  3) Let $M \in SC_{\textbf{K}(\tau_1 \times \tau_2)}$ and $M \rightarrow_{\beta} M'$.
-
-  So ${\bf pure \: } (\lambda x. \pi_i x) \star M \rightarrow_{\beta} {\bf pure \: } (\lambda x. \pi_i x)
-  \star M'$, $i \in \{ 1, 2 \}$ by reduction rule. So ${\bf pure \: } (\lambda x. \pi_i x) \star M' \in
-  SC_{\textbf{K} \tau_i }$ and $M' \in SC_{\textbf{K}(\tau_1 \times \tau_2)}$.
-
-  \end{proof}
-
-  \begin{defin} Neutral term:
-
-  We define a term $M$ to be neutral if it has of the next forms:
-
-  1) $M = x$, where $x \in \mathbb{V}$;
-
-  2) $M = (P Q)$;
-
-  3) $M = \pi_i M$, $i \in \{ 1,2 \}$;
-
-  4) $M = P \star Q$;
-
-  5) If $M$ is a neutral, then ${\bf pure \: } M$ is a neutral.
-
-  \end{defin}
-
-  \begin{lemma}
-
-  Let $M \rightarrow_{\beta} M'$ and $M' \in SC_{\alpha}$ for every one-step reduction. So if $M'$ is a
-  neutral, then $M \in SC_{\alpha}$.
-
-  \end{lemma}
-
-  \begin{proof}
-
-  $ $
-
-  Simple induction on the structure of $M'$.
-
-  \end{proof}
-
-  \begin{lemma}
-
-  $ $
-
-  Let $x_1 : \phi_1, \dots, x_n : \phi_n \vdash M : \phi$ and for all $i \in \{ 1, \dots, n\}$, $N_i
-  \in SC_{\phi_i}$, then $(M [x_1 := N_1, \dots, x_n := N_n]) \in SC_{\phi}$.
-
-  \end{lemma}
-
-  \begin{proof}
-
-$ $
-
-  1) If $\phi$ is an atomic and $M$ is a variable, then this condition holds straightforwardly.
-
-  2) Let $\Gamma = \{ x_1 : \phi_1, \dots, x_n : \phi_n \}$, $\Gamma \vdash {\bf pure \: } M :
-  \textbf{K}\alpha$ and for all $i \in \{1, \dots, n\}, N_i \in SC_{\phi_i}$.
-
-  Then by $\Gamma \vdash M : \alpha$ by generation
-  and $(M [x_1 := N_1, \dots, x_n := N_n]) \in SC_{\alpha}$ by induction hypothesis.
-
-  Hence, $\Gamma \vdash {\bf pure \: } M : \textbf{K}\alpha$ and $({\bf pure \: } M ([x_1 := N_1, \dots, x_n := N_n])) \in SC_{\textbf{K}\alpha}$ by definition of $SC_{\textbf{K}\alpha}$.
-
-  3) Let $\Gamma = \{ x_1 : \phi_1, \dots, x_n : \phi_n \}$, $\Gamma : \phi_n \vdash M \star P : \textbf{K}\beta$ and forall $i \in \{ 1, \dots, n\}, N_i \in SC_{\phi_i}$.
-
-  Then $\Gamma \vdash M : \textbf{K}(\alpha \to \beta)$, $\Gamma \vdash P : \textbf{K}\alpha$ by generation.
-
-  But by induction hypothesis $M [x_1 := N_1, \dots, x_n := N_n] \in SC_{\textbf{K}(\alpha \to
-  \beta)}$ and $P [x_1 := N_1, \dots, x_n := N_n] \in SC_{\textbf{K}\alpha}$.
-
-  Then, by definition of $SC_{\textbf{K}\beta}$, $((M [x_1 := N_1, \dots, x_n := N_n]) \star (P [x_1 := N_1, \dots, x_n := N_n])) \in SC_{\textbf{K}\beta}$, i.e.
-$(M \star N ([x_1 := N_1, \dots, x_n := N_n])) \in
-  SC_{\textbf{K}\beta}$.
-  \end{proof}
-
-  \begin{col}
-
-  $ $
-
-  If $\vdash M : \alpha$, then $M$ is strongly normalizing.
-
-  \end{col}
-
-  \begin{proof}
-
-  $M \in SC_{\alpha}$ by Lemma 10, so $M$ is strongly normalizing.
-
-  \end{proof}
-
-  \end{proof}
-
 \section{Confluence}
-
-In the confluence proof (below) we treat the cases with ${\bf pure}$ and $\star$ similar to \cite{Baren} \cite{Baren2}.
-
-\begin{defin} Alphabet for the labelled terms:
-
-variables: $x, y, z, x_1, y_1, z_1, ...$;
-
-lambdas: $\lambda, \lambda_0, \lambda_1, \lambda_2, ...$;
-
-constructors for an applicative functor: ${\bf pure \: }, \star$;
-
-parentheses (,).
-
-\end{defin}
-
-\begin{defin} The set of labelled terms $\Lambda^{'}_{\textbf{K}}$ inductively defined as a set of words on the alphabet
-described above:
-
-1) $x \in \Lambda^{'}$;
-
-2) If $M \in \Lambda^{'}_{\textbf{K}}$, then $(\lambda x. M) \in \Lambda^{'}_{\textbf{K}}$;
-
-3) If $M, N \in \Lambda^{'}_{\textbf{K}}$, then $(MN) \in \Lambda^{'}_{\textbf{K}}$;
-
-4) If $M \in \Lambda^{'}_{\textbf{K}}$, then ${\bf pure \: } M \in \Lambda^{'}_{\textbf{K}}$;
-
-5) If $M, N \in \Lambda^{'}_{\textbf{K}}$, then $M \star N \in \Lambda^{'}_{\textbf{K}}$;
-
-6) If $M, N \in \Lambda^{'}_{\textbf{K}}$, then for all $i \in \mathbb{N}$, $((\lambda_i x. M) N) \in
-\Lambda^{'}_{\textbf{K}}$.
-
-\end{defin}
-
-\begin{defin} Erasing map
-
-Erasing map is a map $|.| : \Lambda^{'}_{\textbf{K}} \to \Lambda_{\textbf{K}}$, such that:
-
-1) $|x| = x$;
-
-2) $|(\lambda x. M)| = \lambda x. |M|$;
-
-3) $|(M N)| = |M| |N|$;
-
-4) $|({\bf pure \: } M)| = {\bf pure \: } |M|$;
-
-5) $|M \star N| = |M| \star |N|$;
-
-6) $|((\lambda_i x. M) N)| = (\lambda x. |M|) |N|$
-
-\end{defin}
-
-\begin{ex}
-
-$ $
-
-$|{\bf pure \: } ((\lambda_i x. M) N) \star P| = {\bf pure \: } (\lambda x. |M|) |N|) \star |P|$
-
-\end{ex}
-
-\begin{defin} Substitution for $\Lambda^{'}_{\textbf{K}}$:
-
-1) $x [x := N] = N$, $x [y := N] = x$;
-
-2) $(M N) [x := N] = M[x := N] N [x := N]$;
-
-3) $(\lambda x. M) [x := N] = \lambda x. M [x := N]$;
-
-4) $({\bf pure \: } M) [x := P] = {\bf pure \: } \: (M [x := P])$;
-
-5) $(M \star N) [x := P] = (M [x := P]) \star (N [x := P])$;
-
-6) $(\lambda_i x. M) N [y := P] = (\lambda_i x. M [y := P]) (N [y := P])$.
-
-\end{defin}
-
-\begin{defin} One-step reduction $\rightarrow_{\beta^{'}}$ for $\Lambda^{'}_{\textbf{K}}$:
-
-1) $(\lambda x. M) N \rightarrow_{\beta^{'}} M [x := N]$;
-
-2) ${\bf pure \: } (\lambda x. x) \star M \rightarrow_{\beta^{'}} M$;
-
-3) ${\bf pure \: } (\lambda f g x. f (g x)) \star M \star N \star P \rightarrow_{\beta^{'}} M \star (N \star
-P)$;
-
-4) $({\bf pure \: } M) \star ({\bf pure \: } N) \rightarrow_{\beta^{'}} {\bf pure \: } (M N)$;
-
-5) $M \star ({\bf pure \: } N) \rightarrow_{\beta^{'}} {\bf pure \: } (\lambda f. f N) \star M$;
-
-6) $(\lambda_i x. M) N \rightarrow_{\beta^{'}} M [x := N]$.
-
-Multi-step reduction $\twoheadrightarrow_{\beta^{'}}$ is a reflexive-transitive closure of
-$\rightarrow_{\beta^{'}}$.
-
-\end{defin}
-
-\begin{defin} Let us define a map $\phi : \Lambda^{'}_{\textbf{K}} \to \Lambda_{\textbf{K}}$ inductively as
-follows:
-
-1) $\phi(x) = x$;
-
-2) $\phi(M N) = \phi(M)\phi(N)$;
-
-3) $\phi(\lambda x. M) = \lambda x. \phi(M)$;
-
-4) $\phi({\bf pure \: } M) = {\bf pure \: } (\phi(M))$;
-
-5) $\phi(M \star N) = \phi(M) \star \phi(N)$;
-
-6) $\phi((\lambda_i x. M) N) = \phi(M) [x := \phi(N)]$.
-
-\end{defin}
-
-\begin{ex}
-
-$ $
-
-$\phi({\bf pure \: } ((\lambda_i x. M) N) \star P) = {\bf pure \: } (\phi(M) [x := \phi(N)]) \star \phi(P)$
-
-\end{ex}
-
-\begin{lemma}
-
-$ $
-
-1) Let $M, N \in \Lambda^{'}_{\textbf{K}}$ and $|M| \twoheadrightarrow_{\beta} |N|$, then $M \twoheadrightarrow_{\beta^{'}} N$.
-
-2) Let $M, N \in \Lambda^{'}_{\textbf{K}}$ and $M \twoheadrightarrow_{\beta^{'}} N$, then $|M| \twoheadrightarrow_{\beta} |N|$.
-
-\end{lemma}
-
-\begin{proof}
-
-$ $
-
-Induction on the generation of $\twoheadrightarrow_{\beta}$ ($\twoheadrightarrow_{\beta^{'}}$).
-
-1) Let us consider homomorphism rule. The rest applicative reduction rules are considered similary.
-
-Let $({\bf pure \: } M') \star ({\bf pure \: } N'), {\bf pure \: } (M' N') \in \Lambda^{'}_{\textbf{K}}$.
-
-So $|({\bf pure \: } M') \star ({\bf pure \: } N')| = ({\bf pure \: } |M'|) \star ({\bf pure \: } |N'|)$ and
-$|{\bf pure \: } (M' N')| = {\bf pure \: } (|M'| |N'|)$.
-
-By reduction rule, $({\bf pure \: } |M'|) \star ({\bf pure \: } |N'|) \rightarrow_{\beta} {\bf pure \: }
-(|M'||N'|)$.
-
-But $({\bf pure \: } M') \star ({\bf pure \: } N') \rightarrow_{\beta^{'}} {\bf pure \: } (M' N')$ by
-reduction rule for $\rightarrow_{\beta^{'}}$.
-
-  \vspace{\baselineskip}
-
-2) Let us consider interchange rule.
-
-Let $M \star ({\bf pure \: } N), {\bf pure \: } (\lambda f. f N) \star M \in \Lambda^{'}_{\textbf{K}}$ and
-$M \star ({\bf pure \: } N) \rightarrow_{\beta^{'}} {\bf pure \: } (\lambda f. f N) \star M$.
-
-But $|M \star ({\bf pure \: } N)| = |M| \star ({\bf pure \: } |N|)$ and $|{\bf pure \: } (\lambda f. f N)
-\star M| = {\bf pure \: } (\lambda f. f |N|) \star |M|$.
-
-So $|M| \star ({\bf pure \: } |N|) \rightarrow_{\beta} {\bf pure \: } (\lambda f. f |N|) \star |M|$ by
-$\beta$-reduction rule.
-
-\vspace{\baselineskip}
-
-It is easy to see, that the statement for $\twoheadrightarrow_{\beta^{'}}$ and $\twoheadrightarrow_{\beta}$
-immedeatly follows from transitivity of multi-step rediction for labelled terms and for usual terms
-respectively.
-
-\end{proof}
-
-\begin{lemma}
-
-$ $
-
-$\phi(M [x := N]) = \phi(M) [x := \phi(N)]$.
-
-\end{lemma}
-
-\begin{proof} Induction on $M$.
-
-1) Let $M = x$. Then $\phi(x [x := N]) = \phi(N)$.
-
-On the other hand, $\phi(x) [x := \phi(N)] = x [x := \phi(N)] = \phi(N)$.
-
-So $\phi(x [x := N]) = \phi(x) [x := \phi(N)]$.
-
-  \vspace{\baselineskip}
-
-2) Let $M = y$ and $y \neq x$. Then $\phi(y [x := N]) = \phi(y) = y$.
-
-But $\phi(y) [x := \phi(N)] = y [x := \phi(N)] = y$.
-
-Therefore $\phi(y [x := N]) = \phi(y) [x := \phi(N)]$.
-
-  \vspace{\baselineskip}
-
-3) Let $M = {\bf pure \: } M'$. Then $\phi({\bf pure \: } M' [x := N]) = {\bf pure \: } \phi(M' [x := N])$.
-
-By hypothesis, ${\bf pure \: } (\phi(M' [x := N])) = {\bf pure \: } (\phi(M') [x := \phi(N)])$, which is
-$({\bf pure \: } \phi(M')) [x := \phi(N)]$ by substitution definition.
-
-\vspace{\baselineskip}
-
-4) Let $M = M' \star N'$. So $\phi((M' \star N') [x := N])) = \phi(M' [x := N] \star N' [x := N])$.
-
-By definition of $\phi$,
-
-\quad $\phi(M' [x := N] \star N' [x := N]) = \phi(M' [x := N]) \star \phi(N' [x := N])$.
-
-But by induction hypothesis,
-
-\quad $\phi(M' [x := N]) = \phi(M') [x := \phi(N)]$ and
-
-\quad $\phi(N' [x := N]) = \phi(N') [x := \phi(N)]$.
-
-Hence,
-
-\quad $\phi(M' [x := N]) \star \phi(N' [x := N]) = \phi(M') [x := \phi(N)] \star \phi(N') [x := \phi(N)]$.
-
-So,
-
-\quad $\phi(M') [x := \phi(N)] \star \phi(N')[x := \phi(N)] = (\phi(M') \star \phi(N')) [x := \phi(N)]$.
-
-And by definition of $\phi$, $(\phi(M') \star \phi(N')) [x := \phi(N)] = \phi(M' \star N') [x := \phi(N)]$.
-
-\end{proof}
-
-\begin{lemma}
-
-$ $
-
-  Let $M, N \in \Lambda^{'}_{\textbf{K}}$ and $M \twoheadrightarrow_{\beta^{'}} N$, then $\phi(M)
-  \twoheadrightarrow_{\beta} \phi(N)$.
-
-\end{lemma}
-
-\begin{proof}
-
-$ $
-
-1) Let ${\bf pure \: } (\lambda x. x) \star M, M \in \Lambda^{'}_{\textbf{K}}$ and ${\bf pure \: } (\lambda
-x. x) \star M \rightarrow_{\beta^{'}} M$.
-
-But $\phi ({\bf pure \: } (\lambda x. x) \star M) = {\bf pure \: } (\lambda x. x) \star \phi(M)$.\
-
-So ${\bf pure \: } (\lambda x. x) \star \phi(M) \rightarrow_{\beta} \phi(M)$ by $\beta$-reduction rule.
-
-\vspace{\baselineskip}
-
-2) Let ${\bf pure \: } (\lambda f g x. f (g x)) \star M \star N \star P, M \star (N \star P) \in
-\Lambda^{'}_{\textbf{K}}$ and ${\bf pure \: } (\lambda f g x. f (g x)) \star M \star N \star P
-\rightarrow_{\beta^{'}} M \star (N \star P)$.
-
-By the definition of $\phi$:
-
-$\phi({\bf pure \: } (\lambda f g x. f (g x)) \star M \star N \star P) = {\bf pure \: } (\lambda f g x. f (g
-x)) \star \phi(M) \star \phi(N) \star \phi(P)$;
-
-$M \star (N \star P) = \phi(M) \star (\phi(N) \star \phi(P))$.
-
-Hence, ${\bf pure \: } (\lambda f g x. f (g x))
-\star \phi(M) \star \phi(N) \star \phi(P) \rightarrow_{\beta} \phi(M) \star (\phi(N) \star \phi(P))$ by
-$\beta$-reduction rule.
-
-\vspace{\baselineskip}
-
-3) Let $({\bf pure \: } M) \star ({\bf pure \: } N), {\bf pure \: } (M N) \in \Lambda^{'}_{\textbf{K}}$ and
-$({\bf pure \: } M) \star ({\bf pure \: } N) \rightarrow_{\beta} {\bf pure \: } (M N)$.
-
-By the definition of $\phi$:
-
-$\phi(({\bf pure \: } M) \star ({\bf pure \: } N)) = ({\bf pure \: } \phi(M)) \star ({\bf pure \: }
-\phi(N))$;
-
-$\phi({\bf pure \: } (M N)) = {\bf pure \: } (\phi(M) \phi(N))$.
-
-So, by reduction rule, $({\bf pure \: } \phi(M)) \star ({\bf pure \: } \phi(N)) \rightarrow_{\beta} {\bf
-pure \: } (\phi(M) \phi(N))$.
-
-\vspace{\baselineskip}
-
-4) Let $M \star ({\bf pure \: }), {\bf pure \: } (\lambda f. f N) \star M$ and $M \star ({\bf pure \: } N)
-\rightarrow_{\beta^{'}} (\lambda f. f N) \star M$.
-
-$\phi(M \star ({\bf pure \: } N)) = \phi(M) \star ({\bf pure \: } \phi(N))$
-
-$\phi((\lambda f. f N) \star M) = (\lambda f. f \phi(N)) \star \phi(M)$.
-
-So, $\phi(M) \star ({\bf pure \: } \phi(N)) \rightarrow_{\beta} {\bf pure \: } (\lambda f. f \phi(N)) \star
-\phi(M)$.
-\end{proof}
-
-\begin{lemma}
-
-$ $
-
-Let $M \in \Lambda^{'}_{\textbf{K}}$. Then $|M| \twoheadrightarrow_{\beta} \phi(M)$.
-
-\end{lemma}
-
-\begin{proof}
-
-Induction on the structure of $M$.
-
-\end{proof}
-
-\begin{lemma} Strip lemma.
-
-$ $
-
-If $M \rightarrow_{\beta} N$ and $M \twoheadrightarrow_{\beta} P$. Then there exists some term $Q$, such that
-$N \twoheadrightarrow_{\beta} Q$ and $P \twoheadrightarrow_{\beta} Q$.
-
-\end{lemma}
-
-\begin{proof}
-
-  Proof is similar to \cite{Baren} \cite{Baren2}. We build the following diagram
-
-\vspace{\baselineskip}
-
-\xymatrix{
-&&& M \ar[dd]_{\beta} \ar@{->>}[rrr] &&& P \ar@{-->>}[dd]^{\beta} \\
-&&&& N' \ar[ul]^{|.|} \ar[dl]_{\phi} \ar@{-->>}[rrr]^{\beta^{'}} &&& P' \ar[ul]^{|.|} \ar[dl]_{\phi} \\
-&&& N \ar@{-->>}[rrr]_{\beta} &&& Q
-}
-
-\end{proof}
-
-which is commutes by lemmas 11 -- 14.
-
-\begin{theorem} Confluence.
-
-$ $
-
-If $M \twoheadrightarrow_{\beta} N$ and $M \twoheadrightarrow_{\beta} P$. Then there exists some term $Q$,
-such that $N \twoheadrightarrow_{\beta} Q$ and $P \twoheadrightarrow_{\beta} Q$.
-
-\end{theorem}
-
-\begin{proof}
-
-$ $
-
-  By unfolding $M \twoheadrightarrow_{\beta} N$ as the sequence of one-step reductions $M \rightarrow_{\beta}
-  M_1 \rightarrow_{\beta} ... \rightarrow_{\beta} M_n \rightarrow_{\beta} N$ and applying strip lemma on every step.
-
-\end{proof}
 
@@
 
-\vspace{\baselineskip}
+\end{proof}
+
+\begin{lemma}
+  $ $
+
+  i) Let $\Gamma \vdash M : A$ and $M \twoheadrightarrow_{\beta} N$, then $[\![\Gamma \vdash M : A]\!] = [\![\Gamma \vdash N : A]\!]$;
+
+  ii) Let $\Gamma \vdash M : A$ and $M \twoheadrightarrow_{\eta} N$, then $[\![\Gamma \vdash M : A]\!] = [\![\Gamma \vdash N : A]\!]$;
+\end{lemma}
+
+\begin{proof}
+  i) For $\beta$-reduction
+
+\vspace{\baselineskip}
+
+Cases with $\beta$-reductions for $let_{\bf K}$ are shown in \cite{ModalK1}. Let us consider cases with ${\bf pure}$.
+
+\vspace{\baselineskip}
+
+1) $[\![\Gamma \vdash {\bf pure \:} ((\lambda x. M) N) : {\bf K}B]\!] = [\![\Gamma \vdash {\bf pure \:} (M [x := N]]) : {\bf K}B]\!]$
+
+$\begin{array}{lll}
+&[\![\Gamma \vdash {\bf pure \:} (\lambda x. M) N : {\bf K}B]\!] = & \text{By interpretation} \\
+&\eta_{[\![B]\!]} \circ (\epsilon \circ \langle \Lambda([\![M]\!]), [\![N]\!]\rangle) =& \text{Property of $\times$} \\
+&\eta_{[\![B]\!]} \circ (\epsilon \circ (\Lambda([\![M]\!]) \times id_{[\![A]\!]}) \circ \langle id_{[\![\Gamma]\!]}, [\![N]\!] \rangle) = & \text{Associativity of composition} \\
+&\eta_{[\![B]\!]} \circ ((\epsilon \circ (\Lambda([\![M]\!]) \times id_{[\![A]\!]})) \circ \langle id_{[\![\Gamma]\!]}, [\![N]\!] \rangle) = & \text{Exponentiation property} \\
+&\eta_{[\![B]\!]} \circ ([\![M]\!] \circ \langle id_{[\![\Gamma]\!]}, [\![N]\!] \rangle) = & \text{Substitution lemma} \\
+&\eta_{[\![B]\!]} \circ [\![M [\vec{x}, x := \vec{x}, N]]\!] =& \text{By interpretation} \\
+&[\![\Gamma \vdash {\bf pure \:} (M [x := N]]) : {\bf K}B]\!]&
+\end{array}$
+
+\vspace{\baselineskip}
+
+2) $[\![\Gamma \vdash {\bf pure \:} (\pi_i \langle [\![M_1]\!], [\![M_2]\!] \rangle) : {\bf K}A_i ]\!] = [\![\Gamma \vdash {\bf \: pure} M_i : {\bf K}A_i ]\!]$
+
+\vspace{\baselineskip}
+
+$\begin{array}{lll}
+&[\![\Gamma \vdash {\bf pure \:} (\pi_i \langle M_1, M_2] \rangle) : {\bf K}A_i ]\!] = & \text{By interpretation} \\
+&\eta_{[\![A_i]\!]} \circ \pi_i \circ \langle [\![M_1]\!], [\![M_2]\!] \rangle = & \text{Property of $\times$} \\
+&\eta_{[\![A_i]\!]} \circ [\![M_i]\!] =& \text{By interpetation} \\
+&[\![\Gamma \vdash {\bf pure \:} M_i : {\bf K}A_i]\!]&
+\end{array}$
 
@@
   \bibitem{ModalK} Yoshihiko KAKUTANI, A Curry-Howard Correspondence for Intuitionistic Normal Modal Logic, Computer Software, Released February 29, 2008, Online ISSN , Print ISSN 0289-6540.
+
+    \bibitem{ModalK1} Kakutani Y.
+(2007) Call-by-Name and Call-by-Value in Normal Modal Logic.
+In: Shao Z.
+(eds) Programming Languages and Systems.
+APLAS 2007.
+Lecture Notes in Computer Science, vol 4807.
+Springer, Berlin, Heidelberg
   \end{thebibliography}

### 2018-01-17 article.tex
added SN proof

@@
 \newtheorem{col}{Corollary}
-\newtheorem{con}{Convention}
+\newtheorem{con}{Consequence}
 \usepackage{listings} 		% for source code
@@
 
+  Let us show the next simple observation, which immeadelty follows from the previous definition.
+
+  \begin{lemma}
+    $ $
+
+    If $M \rightarrow_{\beta \eta} N$, then ${\bf pure \:} M \rightarrow_{\beta \eta} {\bf pure \:} N$.
+  \end{lemma}
+
   \section{Basic lemmas}
@@
 
+  Strong normalization for ${\bf IK}$ is proved in \cite{Abe} \cite{ModalK}.
+So we consider simply typed lambda calculus with ${\bf K}_I$ rule and show that $\lambda_{\to, \times} + {\bf K}_I$ is strongly normalizable.
+
   \begin{theorem}
@@
   \end{theorem}
+
+  \begin{proof}
+    $ $
+
+    We build the subset of strongly normalazing terms and show that an arbitrary term belongs to this subset.
+
+    \begin{defin} The set of strongly computable terms for every type $T \in \mathbb{T_{\bf K}}$.
+      \begin{itemize}
+        \item Let $A \in \mathbb{T}$, then $SC_A = \{ M : A \: | \: \text{$M$ is strongly normalizing}\}$;
+        \item $SC_{A \to B} = \{ M : A \to B \: | \: \forall A \in SC_A, M N \in SC_B \}$;
+        \item $SC_{A_1 \times A_2} = \{ M : A \times B \: | \: \pi_i M \in SC_{A_i}, i \in \{ 1, 2 \} \}$;
+        \item $SC_{{\bf K}A} = \{ {\bf pure \:} M : {\bf K}A \: | \: M \in SC_A \}$
+      \end{itemize}
+    \end{defin}
+
+    Strong normalization proof reduces to the proof of the next lemma:
+
+    \begin{lemma}
+      $ $
+
+      i) If $M \in SC_A$, then $M$ is stronly normalizing;
+
+      ii) If $M \rightarrow_{\beta} M^{'}$ and $M \in SC_A$, then $M^{'}$;
+
+      iii) Let $M \rightarrow_{\beta} M^{'}$ and $M^{'} \in SC_A$, then, if $M$ is a neutral term, then $M \in SC_A$.
+
+      iv) Let $x_1 : A_1,\dots,x_n : A_n \vdash M : B$ and $\forall i \in \{ 1,\dots,n\}, N_i \in SC_{A_i}$, then $M [\vec{x} := \vec{N}] \in SC_B$.
+    \end{lemma}
+
+    \begin{proof}
+      $ $
+
+      i)
+
+      The base case follows from the definition.
+
+      Let us consider case with $SC_{{\bf K}A}$.
+      If ${\bf pure \:} M \in SC_{{\bf K}A}$, then $M \in SC_A$ and $M$ is strongly normalizible.
+      So ${\bf pure \:} M$ is strongly normalizible, otherwise there would be an infinite reduction path in ${\bf pure \:} M$.
+
+      \vspace{\baselineskip}
+
+      ii)
+
+      The base case is trivial.
+
+      Let ${\bf pure \:} M \rightarrow_{\beta} {\bf pure \:} M^{'}$ and ${\bf pure \:} M \in SN_{{\bf K}A}$.
+      By assumption, $M \in SN_A$ and $M \rightarrow_{\beta} M^{'}$, so $M^{'} \in SN_A$. Hence ${\bf pure \:} M^{'} \in SC_{{\bf K}A}$ by the first statement of the lemma.
+
+      \vspace{\baselineskip}
+
+      iii)
+
+      The base case is trivial.
+
+      Let ${\bf pure \:} M \rightarrow_{\beta} {\bf pure \:} M^{'}$ and ${\bf pure \:} M^{'} \in SN_{{\bf K}A}$.
+
+      ${\bf pure \:} M^{'}$ is a neutral by the definition. By assumption $M$ is a strongly normalizing. So ${\bf pure\:} M$ is a strongly normalizing by the first part of the current lemma.
+
+      \vspace{\baselineskip}
+
+      iv)
+
+      Let $x_1 : A_1,\dots,x_n : A_n \vdash {\bf pure \:} M : {\bf K}A$ and $\forall i \in \{ 1,\dots,n\}, N_i \in SC_{A_i}$.
+
+      By generation $x_1 : A_1,\dots,x_n : A_n \vdash M : A$ and by assumption $M [\vec{x} := \vec{N}] \in SC_B$.
+
+      Hence, by the first part of lemma, ${\bf pure \:} (M [\vec{x} := \vec{N}]) \in SC_{{\bf K}B}$.
+    \end{proof}
+
+    \begin{col}
+      $ $
+
+      Let $\vdash N : A$, then $N$ is strongly normalizing.
+    \end{col}
+
+    \begin{proof}
+      $ $
+
+    If $\vdash N : A$, then $N \in SC_A$, hence $N$ is strongly normalizing.
+    \end{proof}
+
+  \end{proof}
 
@@
 
+\begin{proof}
+$ $
+Straightforwardly follows from two previous lemmas.
+\end{proof}
+
   \section{Acknowledgement.}
@@
 Springer, Berlin, Heidelberg
+
+  \bibitem{Abe} T. Abe. Completeness of modal proofs in first-order predicate logic. Computer Software, JSSST Journal, 24:165 -- 177, 2007.
   \end{thebibliography}

### 2018-01-17 article.tex
added some stuff

@@
 
+We identify terminal objects. So $\mathcal{K}(\mathds{1}) = \mathds{1}$ and $\eta_{\mathds{1}} = id_{\mathds{1}}$ since $\mathcal{K}$ is an endofunctor.
+
 \subsection{Soundness}
@@
 \begin{proof}
+  $ $
+
   i) For $\beta$-reduction
@@
 
+\vspace{\baselineskip}
+
+ii) For $\eta$-reduction.
+
+\vspace{\baselineskip}
+
+1) $[\![\Gamma \vdash {\bf pure \:} (\lambda x. M x) : {\bf K}(A \to B)]\!] = [\![\Gamma \vdash {\bf pure \:} M : {\bf K}(A \to B)]\!]$.
+
+\vspace{\baselineskip}
+
+$\begin{array}{lll}
+&[\![\Gamma \vdash {\bf pure \:} (\lambda x. M x) : {\bf K}(A \to B)]\!] = & \text{By interpetation} \\
+&\eta_{[\![B]\!]^{[\![A]\!]}} \circ \Lambda(\epsilon \circ [\![M]\!] \times id_{[\![A]\!])}& \text{Exponentiation property} \\
+&\eta_{[\![B]\!]^{[\![A]\!]}} \circ [\![M]\!] = & \text{By interpretation} \\
+&[\![\Gamma \vdash {\bf pure \:} M : {\bf K}(A \to B)]\!]&
+\end{array}$
+
+\vspace{\baselineskip}
+
+2) $[\![\Gamma \vdash {\bf pure \:} \langle \pi_1 M, \pi_2 M \rangle : {\bf K}(A \times B)]\!] = [\![\Gamma \vdash {\bf pure \:} M : {\bf K}(A \times B)]\!]$
+
+\vspace{\baselineskip}
+
+$\begin{array}{lll}
+&[\![\Gamma \vdash {\bf pure \:} \langle \pi_1 M, \pi_2 M \rangle : {\bf K}(A \times B)]\!] = & \text{By interpetation} \\
+&\eta_{[\![A]\!] \times [\![B]\!]} \circ \langle \pi_1 \circ [\![M]\!], \pi_2 \circ [\![M]\!] \rangle  = & \text{By the property of a product of morphisms} \\
+&\eta_{[\![A]\!] \times [\![B]\!]} \circ [\![M]\!] = & \text{By interpetation} \\
+&[\![\Gamma \vdash {\bf pure \:} M : {\bf K}(A \times B)]\!]&
+\end{array}$
+
+\vspace{\baselineskip}
+
+3) $[\![\vdash {\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} N : {\bf }K A ]\!] = [\![\vdash {\bf pure \:} N : {\bf K A}]\!]$.
+
+\vspace{\baselineskip}
+
+$\begin{array}{lll}
+&[\![\vdash {\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} N : {\bf }K A ]\!] = & \text{By interpetation}\\
+&\mathcal{K}([\![N]\!]) \circ \eta_{\mathds{1}} =& \text{Naturality for $\eta$} \\
+&\eta_{[\![A]\!]} \circ [\![N]\!] =& \text{By interpretation} \\
+&[\![\vdash {\bf pure \:} N : {\bf K} A]\!]& 
+\end{array}$
+
 \end{proof}

### 2018-01-18 article.tex
erased rubbish

@@
   provides the epistimology and the theory of knowledge as based on BHK-semantics of intuitionistic logic.
-  $IEL^{-}$  is a variant of IEL, that corresponds to intuitionistic belief.
-  Informally, $\textbf{K}A$ denotes that $A$ is verified intuitionistically.
+  IEL$^{-}$  is a variant of IEL, that corresponds to intuitionistic belief.
+  Informally, ${\bf K}A$ denotes that $A$ is verified intuitionistically.
 
@@
 
-    2) $\textbf{K}(A \to B) \to (\textbf{K}A \to \textbf{K}B)$ (normality);
-
-    3) $A \to \textbf{K}A$ (co-reflection);
+    2) ${\bf K}(A \to B) \to ({\bf K}A \to {\bf K}B)$ (normality);
+
+    3) $A \to {\bf K}A$ (co-reflection);
 
@@
 
-  $\lambda K$ consists of the rules for simply typed lambda-calculus and special typing rules for
+  $\lambda_{{\bf K}}$ consists of the rules for simply typed lambda-calculus and special typing rules for
   lifting types into the applicative functor $\textbf{K}$. We assume that our type system will
@@
 
-  Let $\mathbb{V}$ be the set of variables. The set $\Lambda_{\textbf{K}}$ of terms is defined by the grammar:
-
-$\begin{array}{lll}
-& \Lambda_{\textbf{K}} ::= \mathbb{V} \: | \:  (\lambda \Lambda.\Lambda_{\textbf{K}}) \: | \: (\Lambda_{\textbf{K}}\Lambda_{\textbf{K}}) \: | \: (\Lambda_{\textbf{K}} , \Lambda_{\textbf{K}}) \: | \: (\pi_1 \Lambda_{\textbf{K}}) \: | \: (\pi_2 \Lambda_{\textbf{K}}) \: | & \\
-& \quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad ({\bf pure \: } \: \Lambda_{\textbf{K}}) \: | \: ({\bf let \:\: pure} \:\: \Lambda_{\textbf{K}} = \Lambda_{\textbf{K}} \:\: {\bf in} \:\: \Lambda_{\textbf{K}})
+  Let $\mathbb{V}$ be the set of variables. The set $\Lambda_{{\bf K}}$ of terms is defined by the grammar:
+
+$\begin{array}{lll}
+& \Lambda_{{\bf K}} ::= \mathbb{V} \: | \:  (\lambda \Lambda.\Lambda_{{\bf K}}) \: | \: (\Lambda_{{\bf K}}\Lambda_{{\bf K}}) \: | \: (\Lambda_{{\bf K}} , \Lambda_{{\bf K}}) \: | \: (\pi_1 \Lambda_{{\bf K}}) \: | \: (\pi_2 \Lambda_{{\bf K}}) \: | & \\
+& \quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad ({\bf pure \: } \: \Lambda_{{\bf K}}) \: | \: ({\bf let \:\: pure} \:\: \Lambda_{{\bf K}} = \Lambda_{{\bf K}} \:\: {\bf in} \:\: \Lambda_{{\bf K}})
 \end{array}$
@@
 
-  Let $\mathbb{T}$ be the set of atomic types. The set $\mathbb{T}_{\textbf{K}}$ of types with applicative
-  functor
-  $\textbf{K}$ is generated by the grammar:
+  Let $\mathbb{T}$ be the set of atomic types. The set $\mathbb{T}_{{\bf K}}$ of types with applicative
+  functor ${\bf K}$ is generated by the grammar:
   \begin{equation}
-    \mathbb{T}_{\textbf{K}} ::= \mathbb{T} \: | \: (\mathbb{T}_{\textbf{K}} \to \mathbb{T}_{\textbf{K}}) \: |
-    \:
-    (\mathbb{T}_{\textbf{K}} \times \mathbb{T}_{\textbf{K}}) \: | \: (\textbf{K}\mathbb{T}_{\textbf{K}})
+    \mathbb{T}_{{\bf K}} ::= \mathbb{T} \: | \: (\mathbb{T}_{{\bf K}} \to \mathbb{T}_{{\bf K}}) \: |
+    \: (\mathbb{T}_{{\bf K}} \times \mathbb{T}_{{\bf K}}) \: | \: ({\bf K}\mathbb{T}_{{\bf K}})
   \end{equation}
@@
 
-  \begin{defin} Modal typed lambda calculus $\lambda \textbf{K}$ based on NIEL$^{-}_{\land, \to}$:
+  \begin{defin} Modal typed lambda calculus $\lambda_{{\bf K}}$ based on NIEL$^{-}_{\land, \to}$:
 
@@
       \AxiomC{$\vec{x} : \vec{A} \vdash M : B$}
-      \RightLabel{$\text{let}_{\bf K}$}
+      \RightLabel{$\text{let}_{{\bf K}}$}
       \BinaryInfC{$\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } M : {\bf K} B$}
@@
 
-  $\textbf{K}_I$-typing rule is the same as $\bigcirc$-introduction in lax logic (also known as monadic
+  ${\bf K}_I$-typing rule is the same as $\bigcirc$-introduction in lax logic (also known as monadic
   metalanguage \cite{Lax}) and in typed lambda-calculus which is derived by proof-assignment for lax-logic
   proofs.
-  $\textbf{K}_I$ allows to inject an object of type $\alpha$ into the functor. $\textbf{K}_I$  reflects the
+  ${\bf K}_I$ allows to inject an object of type $\alpha$ into the functor. ${\bf K}_I$  reflects the
   Haskell method {\bf pure} for Applicative class. It plays the same role as the {\bf return} method in Monad class.
 
-  $\text{let}_{\bf K}$ is similar to $\Box_I$-rule in typed lambda calculus for intuitionistic normal modal logic {\bf IK}, which is described in \cite{ModalK}.
+  $\text{let}_{{\bf K}}$ is the same as the $\Box$-rule in typed lambda calculus for intuitionistic normal modal logic {\bf IK}, which is described in \cite{ModalK}.
+
+  In fact, our calculus is the extention of typed lambda calculus for {\bf IK} with typing rule appropriate to co-reflection.
 
@@
 
-  5) ${\bf pure \:} ((\lambda x. M) N) \rightarrow_{\beta} {\bf pure \: } (M [x := N])$;
-
-  6) ${\bf pure \:} (\pi_i \langle M_1, M_2 \rangle) \rightarrow_{\beta} {\bf pure \:} M_i$, where $i \in \{ 1, 2 \}$.
-
-  7) $\begin{array}{llll}
-  & {\bf pure \:} ({\bf let \: pure \:} \langle \vec{x}, y, \vec{z} \rangle = \langle \vec{M}, {\bf let \: pure \: } \vec{w} = \vec{N} {\bf \: in \: } Q, \vec{P} \rangle {\: in \:} R) \rightarrow_{\beta} & \\
-  & {\bf pure \:} ({\bf let \: pure \:} \langle \vec{x}, \vec{w}, \vec{z} \rangle = \langle \vec{M}, \vec{N}, \vec{P} \rangle \: { \bf in } \: R [y := Q])
-  \end{array}$
-
-  8) $\lambda x. f x \rightarrow_{\eta} f$;
-
-  9) $\langle \pi_1 P, \pi_2 P \rangle \rightarrow_{\eta} P$;
+  5) $M \rightarrow_{\beta} N \Rightarrow {\bf pure \: M} \rightarrow_{\beta} {\bf pure \: N}$
+
+  6) $\lambda x. f x \rightarrow_{\eta} f$;
+
+  7) $\langle \pi_1 P, \pi_2 P \rangle \rightarrow_{\eta} P$;
 
@@
 
-  12) ${\bf pure \:} (\lambda x. f x) \rightarrow_{\eta} {\bf pure \:} f$;
-
-  13) ${\bf pure \:} (\langle \pi_1 P, \pi_2 P \rangle) \rightarrow_{\eta} {\bf pure \:} P$;
-
-  14) ${\bf pure \:} ({\bf let \: pure \:} x = M {\: \bf in \: } x) \rightarrow_{\eta} {\bf pure \:} M$;
-
-  15) ${\bf pure \:} ({\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} N) \rightarrow_{\eta} { \bf pure \:} ({\bf pure \:} N)$.
+  12) $M \rightarrow_{\beta} N \Rightarrow {\bf pure \: M} \rightarrow_{\eta} {\bf pure \: N}$
   \end{defin}
@@
   example see \cite{Neder}\cite{Morten}\cite{Girard}. We will consider only modal cases}:
-
-  \begin{defin} The domain of a context $\Gamma$:
-
-  Let $\Gamma = \{ x_1 : A_1, ..., x_n : A_n \}$. Then the domain of $\Gamma$, or $dom(\Gamma)$, is a set $\{  x_1, ..., x_n\}$.
-
-  \end{defin}
-
-  \begin{lemma}
-
-  If $\Gamma \vdash M : A$, then $FV(M) \subseteq dom(\Gamma)$
-
-  \end{lemma}
-
-  \begin{proof}
-
-  Induction on the derivation of $\Gamma \vdash M : A$.
-
-  \end{proof}
-
-  \vspace{\baselineskip}
-
-  \begin{lemma} Generation for $\lambda \textbf{K}$.
-
-  1) $\Gamma \vdash {\bf pure \: } M : \textbf{K}\alpha$ implies that $\Gamma \vdash M : \alpha$;
-
-  2) $\Gamma \vdash {\bf let \: pure} \: \vec{N} = \vec{M} \: {\bf in} \: M : {\bf K} B$ implies that $\Gamma \vdash \vec{M} : {\bf K} \vec{A}$ and
-$\vec{N} : \vec{A} \vdash M : B$.
-  \end{lemma}
-
-  \begin{proof}
-
-  $ $
-
-Induction on the derivation of $\Gamma \vdash {\bf pure \: } M : \textbf{K}\alpha$ and $\Gamma \vdash {\bf let \: pure} \: \vec{N} = \vec{M} \: {\bf in} \: M : {\bf K} B$ respectively.
-  \end{proof}
-
-  \vspace{\baselineskip}
-
-  The next one lemma allows that weakening structural rule is admissable.
-
-  \begin{lemma} Weakening for $\lambda \textbf{K}$.
-
-  Let $\Gamma \vdash M : A$ and $\Gamma \subseteq \Delta$, then $\Delta \vdash M : A$.
-
-  \end{lemma}
-
-  \begin{proof}
-
-  $ $
-
-  Induction on derivation of $\Gamma \vdash M : A$. Let us assume $\Gamma \subseteq \Delta$.
-
-  1) Let $\Gamma \vdash x : A$, such that $\Gamma = \Delta, x : A$ and $\Theta \subseteq \Gamma$.
-  Let $\Sigma = \Theta \setminus \Gamma$, or, which is the same, $\Sigma = \Theta \setminus \Delta, x :
-  A$, then $\Sigma, \Delta, x : A \vdash x : A$, or, $\Theta \vdash x : A$.
-
-  \vspace{\baselineskip}
-
-  2) Let $\Gamma \vdash {\bf pure \: } M : \textbf{K}A$ and $\Gamma \subseteq \Theta$.
-
-  By generation $\Gamma \vdash M : A$
-
-  By hypothesis, $\Theta \vdash M : A$, so $\Theta \vdash {\bf pure \: } M : \textbf{K}A$ by applying
-  $\textbf{K}_I$-rule.
-
-    \vspace{\baselineskip}
-
-  3) Let $\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N : {\bf K} B$ and $\Gamma \subseteq \Theta$.
-
-  By generation $\Gamma \vdash \vec{M} : {\bf K} \vec{A}$ and $\vec{x} : \vec{A} \vdash N : B$.
-
-  By assumption $\Theta \vdash \vec{M} : {\bf K} \vec{A}$.
-
-  Hence $\Theta \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N : {\bf K} B$.
-
-  \end{proof}
-
-  \begin{lemma} Considering for $\lambda \textbf{K}$.
-
-  If $\Gamma \vdash M : \alpha$, then $\Gamma \uparrow FV(M) \vdash M : \alpha$, where $\Gamma \uparrow
-  FV(M)$ is a subcontext of $\Gamma$, such that $dom(\Gamma \uparrow FV(M)) = dom(\Gamma) \cap FV(M)$.
-
-  \end{lemma}
-
-  \begin{proof}
-
-  $ $
-
-  1) Let $\Gamma \vdash x : A$, where $\Gamma = \Delta, x : A$, $x \in \mathbb{V}$.
-
-  $FV(x) = \{ x\}$, then $dom(\Gamma) \cap \{ x\} = \{ x\}. $ So $(\Delta, x : A) \uparrow FV(x) = \{ x
-  : A\}$, then $x : A \vdash x : A$ by axiom.
-
-  \vspace{\baselineskip}
-
-  2) Let $\Gamma \vdash {\bf pure \: } M : {\bf K} A$.
-
-  By generation, $\Gamma \vdash M : A$ and $\Gamma \uparrow FV(M) \vdash M : A$ by hypothesis.
-
-  So $\Gamma \uparrow FV(M) \vdash {\bf pure \: } M : {\bf K} A$ by ${\bf K}_I$.
-
-  \vspace{\baselineskip}
-
-  3) Let $\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N : {\bf K} B$.
-
-By generation, $\Gamma \vdash \vec{M} : {\bf K} \vec{A}$ and $\vec{x} : \vec{A} \vdash N : B$.
-
-By assumption, $\Gamma \uparrow FV(\vec{M}) \vdash \vec{M} : {\bf } \vec{A}$.
-
-By $\text{let}_{\bf K}$, $\Gamma \uparrow FV(\vec{M}) \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N : {\bf K} B$
-
-  \end{proof}
-
-  \vspace{\baselineskip}
-
-  \begin{lemma}
-
-  If $\Gamma, x : A \vdash M : B$  and $\Gamma \vdash N : A$, then $\Gamma \vdash (M [x := N]) : B$
-
-  \end{lemma}
-
-  \begin{proof}
-
-  $ $
-
-  1) Let $\Gamma, x : A \vdash {\bf pure \: } M : \textbf{K}B$ and $\Gamma \vdash N : A$.
-
-  By generation, $\Gamma, x : A \vdash M : B$.
-
-  By assumption, $\Gamma \vdash (M [x := N]) : B$
-
-  Then $\Gamma \vdash {\bf pure \: } (M [x := N]) : \textbf{K}B$ by $\textbf{K}_I$.
-
-  But ${\bf pure \: } (M [x := N]) = ({\bf pure \: } M [x := N]) $ by substitution definition, so
-  $\Gamma \vdash ({\bf pure \: } M [x := N]) : \textbf{K}B$
-
-  \vspace{\baselineskip}
-
-  2) Let $\Gamma, y : A \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N : {\bf K} B$ and $\Gamma \vdash N : A$.
-
-  By generation, $\Gamma, y : A \vdash \vec{M} : {\bf K} \vec{A}$ and $\vec{x} : \vec{A} \vdash N : B$.
-
-  By hypothesis, $\Gamma \vdash \vec{M}[x := N] : {\bf K} \vec{A}$.
-
-  Hence $\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M}[x := N] {\: \bf in \:} N : {\bf K} B$.
-
-  \end{proof}
-
-  \vspace{\baselineskip}
-
-  \begin{theorem} Subject reduction
-
-  $ $
-
-  i) Let $\Gamma \vdash M : A$ and $M \twoheadrightarrow_{\beta} N$, then $\Gamma \vdash N : A$
-
-  ii) Let $\Gamma \vdash M : A$ and $M \twoheadrightarrow_{\eta} N$, then $\Gamma \vdash N : A$
-  \end{theorem}
-
-We consider only modal $\beta$-reduction rules. The general statement for $\twoheadrightarrow_{\beta}$ follows from transitivity of multi-step $\beta$-reduction.
-
-  \begin{proof}
-
-  $ $
-
-  i) For multistep $\beta$-reduction:
-
-  \vspace{\baselineskip}
-
-1) Let $\Gamma \vdash {\bf let \: pure \:} \langle \vec{x}, y, \vec{z} \rangle = \langle \vec{M}, {\bf let \: pure \: } \vec{w} = \vec{N} { \: \bf in \: } Q, \vec{P} \rangle { \: \bf in \: } R : {\bf K }B$
-
-By generation we have $\Gamma \vdash \vec{M} : {\bf K}\vec{A_1}$, $\Gamma \vdash {\bf let \: pure \: } \vec{w} = \vec{N} {\: \bf in \:} Q : {\bf K}\vec{A_2}$, $\Gamma \vdash \vec{P} : {\bf K}\vec{A_3}$
-and $\vec{x} : \vec{A_1}, y : A_2, \vec{z} : \vec{A_3} \vdash R : B$.
-
-If $\Gamma \vdash {\bf let \: pure \: } \vec{w} = \vec{N} {\bf \: in \: } Q : {\bf K}\vec{A_2}$, then
-
-$\Gamma \vdash \vec{N} : {\bf K}\vec{A_4}$ and $\vec{w} : \vec{A_4} \vdash Q : A_2$.
-
-Then $\vec{x} : \vec{A_1}, \vec{w} : \vec{A_4}, \vec{z} : \vec{A_3} \vdash R [y := Q] : B$ by substitution lemma and weakening.
-
-Hence $\Gamma \vdash {\bf let \: pure \:} \langle \vec{x}, \vec{w}, \vec{z} \rangle = \langle \vec{M}, \vec{N}, \vec{P} \rangle \: { \bf in } \: R [y := Q] : {\bf K} B$
-by $let_{{\bf K}}$.
-
-  \vspace{\baselineskip}
-
-2) Let $\Gamma \vdash {\bf pure } ((\lambda x. M) N) : {\bf K} B$.
-
-By generation $\Gamma \vdash (\lambda x. M) N : B$, but $\Gamma \vdash M [x := N] : B$, then, by ${\bf K}_I$, $\Gamma \vdash {\bf pure \:} (M [x := N]) : {\bf K} B$.
-
-  \vspace{\baselineskip}
-
-3) Let $\Gamma \vdash {\bf pure } (\pi_i \langle M_1, M_2 \rangle) : {\bf K} A_i$, where $i \in \{ 1,2 \}$.
-
-By generation $\Gamma \vdash \pi_i \langle M_1, M_2 \rangle : A_i$ and $\Gamma \vdash M_i : A_i$.
-
-Hence $\Gamma \vdash {\bf pure \: } M_i : { \bf K } A_i$ by ${\bf K}_I$.
-
-\vspace{\baselineskip}
-
-4) Let $\Gamma \vdash {\bf let \: pure \:} (\langle \vec{x}, y, \vec{z} \rangle = \langle \vec{M}, {\bf let \: pure \: } \vec{w} = \vec{N} {\bf \: in \: } Q, \vec{P} \rangle { \: \bf in \: } R) : {\bf K }^2 B$.
-
-By generation $\Gamma \vdash {\bf let \: pure \:} \langle \vec{x}, y, \vec{z} \rangle = \langle \vec{M}, {\bf let \: pure \: } \vec{w} = \vec{N} {\bf \: in \: } Q, \vec{P} \rangle { \: \bf in \: } R : {\bf K }B$,
-
-hence $\Gamma \vdash {\bf let \: pure \:} \langle \vec{x}, \vec{w}, \vec{z} \rangle = \langle \vec{M}, \vec{N}, \vec{P} \rangle \: { \bf in } \: R [y := Q] : {\bf K} B$ by the first case.
-
-So $\Gamma \vdash {\bf pure \:} ({\bf let \: pure \:} \langle \vec{x}, \vec{w}, \vec{z} \rangle = \langle \vec{M}, \vec{N}, \vec{P} \rangle \: { \bf in } \: R [y := Q]) : {\bf K}^2 B$
-by ${\bf K}_I$.
-
-\vspace{\baselineskip}
-
-ii) For multistep $\eta$-reduction:
-
-\vspace{\baselineskip}
-
-  1) Let $\vdash {\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} N : {\bf K} A$.
-
-  Then by generation $\vdash N : A$, so $\vdash {\bf pure \:} N : {\bf K} A$ by ${\bf K}_I$.
-
-  \vspace{\baselineskip}
-
-  2) Let $\Gamma \vdash {\bf let \: pure \:} x = M {\: \bf in \: } x : {\bf K} A$.
-
-  By generation $\Gamma \vdash M : {\bf K} A$ and $x : A \vdash x : A$, hence $\Gamma \vdash M : {\bf K} A$.
-
-  \vspace{\baselineskip}
-
-  3) Let $\Gamma \vdash {\bf pure \:} (\lambda x. f x) : {\bf K}(A \to B)$.
-
-  By generation $\Gamma \vdash \lambda x. f x : A \to B$, so $\Gamma \vdash f : A \to B$, then $\Gamma \vdash {\bf pure \:} f : {\bf K}(A \to B)$ by ${\bf K}_I$.
-
-  \vspace{\baselineskip}
-
-  4) Let $\Gamma \vdash {\bf pure \:} (\langle \pi_1 P, \pi_2 P \rangle) : {\bf K}(A \times B)$.
-
-  By generation $\Gamma \vdash \langle \pi_1 P, \pi_2 P \rangle : A \times B$, then $\Gamma \vdash P : A \times B$.
-
-  By ${\bf K}_I$, $\Gamma \vdash {\bf pure \:} P : {\bf K}(A \times B)$.
-
-  \vspace{\baselineskip}
-
-  5) $\Gamma \vdash {\bf pure \:} ({\bf let \: pure \:} x = M {\: \bf in \: } x) : {\bf K}^2 A$.
-
-  Then $\Gamma \vdash M : {\bf K}^2 A$ and $x : {\bf K} A \vdash x : {\bf K} A$, so $\Gamma \vdash M : {\bf K}^2 A$.
-
-  \vspace{\baselineskip}
-
-  6) Let $\vdash {\bf pure \:} ({\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} N) : {\bf K}^2 A$.
-
-  By generation ${\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} N : {\bf K}A$, so $\vdash N : A$, then $\vdash {\bf pure \:} N : {\bf K}$.
-
-  \end{proof}
-
-  \vspace{\baselineskip}
 
@@
 
-  2) $\ast_{A, B} \mathcal{F}A \otimes_2 \mathcal{F}B \to \mathcal{F}(A \otimes_1 B)$
+  2) $\ast_{A, B} : \mathcal{F}A \otimes_2 \mathcal{F}B \to \mathcal{F}(A \otimes_1 B)$
 

### 2018-01-18 article.tex
erased non needed text

@@
 
-  We modify and apply Tait's technique of logical relation for modalities. Strong normalization proof with Tait's method for simply typed lambda calculus is described here \cite{Pierce}.
-
-  Strong normalization for ${\bf IK}$ is proved in \cite{Abe} \cite{ModalK}.
-So we consider simply typed lambda calculus with ${\bf K}_I$ rule and show that $\lambda_{\to, \times} + {\bf K}_I$ is strongly normalizable.
-
-  \begin{theorem}
-
-    Let $M \in \Lambda_{\textbf{K}}$, then any sequence of reduction $M \rightarrow_{\beta} M_1 \dots$ terminates.
-
-  \end{theorem}
-
-  \begin{proof}
-    $ $
-
-    We build the subset of strongly normalazing terms and show that an arbitrary term belongs to this subset.
-
-    \begin{defin} The set of strongly computable terms for every type $T \in \mathbb{T_{\bf K}}$.
-      \begin{itemize}
-        \item Let $A \in \mathbb{T}$, then $SC_A = \{ M : A \: | \: \text{$M$ is strongly normalizing}\}$;
-        \item $SC_{A \to B} = \{ M : A \to B \: | \: \forall A \in SC_A, M N \in SC_B \}$;
-        \item $SC_{A_1 \times A_2} = \{ M : A \times B \: | \: \pi_i M \in SC_{A_i}, i \in \{ 1, 2 \} \}$;
-        \item $SC_{{\bf K}A} = \{ {\bf pure \:} M : {\bf K}A \: | \: M \in SC_A \}$
-      \end{itemize}
-    \end{defin}
-
-    Strong normalization proof reduces to the proof of the next lemma:
-
-    \begin{lemma}
-      $ $
-
-      i) If $M \in SC_A$, then $M$ is stronly normalizing;
-
-      ii) If $M \rightarrow_{\beta} M^{'}$ and $M \in SC_A$, then $M^{'}$;
-
-      iii) Let $M \rightarrow_{\beta} M^{'}$ and $M^{'} \in SC_A$, then, if $M$ is a neutral term, then $M \in SC_A$.
-
-      iv) Let $x_1 : A_1,\dots,x_n : A_n \vdash M : B$ and $\forall i \in \{ 1,\dots,n\}, N_i \in SC_{A_i}$, then $M [\vec{x} := \vec{N}] \in SC_B$.
-    \end{lemma}
-
-    \begin{proof}
-      $ $
-
-      i)
-
-      The base case follows from the definition.
-
-      Let us consider case with $SC_{{\bf K}A}$.
-      If ${\bf pure \:} M \in SC_{{\bf K}A}$, then $M \in SC_A$ and $M$ is strongly normalizible.
-      So ${\bf pure \:} M$ is strongly normalizible, otherwise there would be an infinite reduction path in ${\bf pure \:} M$.
-
-      \vspace{\baselineskip}
-
-      ii)
-
-      The base case is trivial.
-
-      Let ${\bf pure \:} M \rightarrow_{\beta} {\bf pure \:} M^{'}$ and ${\bf pure \:} M \in SN_{{\bf K}A}$.
-      By assumption, $M \in SN_A$ and $M \rightarrow_{\beta} M^{'}$, so $M^{'} \in SN_A$. Hence ${\bf pure \:} M^{'} \in SC_{{\bf K}A}$ by the first statement of the lemma.
-
-      \vspace{\baselineskip}
-
-      iii)
-
-      The base case is trivial.
-
-      Let ${\bf pure \:} M \rightarrow_{\beta} {\bf pure \:} M^{'}$ and ${\bf pure \:} M^{'} \in SN_{{\bf K}A}$.
-
-      ${\bf pure \:} M^{'}$ is a neutral by the definition. By assumption $M$ is a strongly normalizing. So ${\bf pure\:} M$ is a strongly normalizing by the first part of the current lemma.
-
-      \vspace{\baselineskip}
-
-      iv)
-
-      Let $x_1 : A_1,\dots,x_n : A_n \vdash {\bf pure \:} M : {\bf K}A$ and $\forall i \in \{ 1,\dots,n\}, N_i \in SC_{A_i}$.
-
-      By generation $x_1 : A_1,\dots,x_n : A_n \vdash M : A$ and by assumption $M [\vec{x} := \vec{N}] \in SC_B$.
-
-      Hence, by the first part of lemma, ${\bf pure \:} (M [\vec{x} := \vec{N}]) \in SC_{{\bf K}B}$.
-    \end{proof}
-
-    \begin{col}
-      $ $
-
-      Let $\vdash N : A$, then $N$ is strongly normalizing.
-    \end{col}
-
-    \begin{proof}
-      $ $
-
-    If $\vdash N : A$, then $N \in SC_A$, hence $N$ is strongly normalizing.
-    \end{proof}
-
-  \end{proof}
-
 \section{Confluence}
@@
 
-  \section{Acknowledgement.}
-
-  Author would like to thank his supervisior
-  V.L.Vasukov, V.N. Krupski for general idea and wise
-  advice, V. de Paiva, V.I. Shalack, A.V. Rodin and M.
-  Taldykin for discussing, critics and consulting.
-
   \addcontentsline{toc}{section}{References}

### 2018-01-18 article.tex
added

@@
 
-  5) $M \rightarrow_{\beta} N \Rightarrow {\bf pure \: M} \rightarrow_{\beta} {\bf pure \: N}$
+  5) ${\bf let \: pure \:} \vec{x} = {\bf pure \:} \vec{M} {\: \bf in \:} N \rightarrow_{\beta} {\bf pure \:} N [\vec{x} := \vec{M}]$
 
-  6) $\lambda x. f x \rightarrow_{\eta} f$;
+  6) $M \rightarrow_{\beta} N \Rightarrow {\bf pure \: M} \rightarrow_{\beta} {\bf pure \: N}$
 
-  7) $\langle \pi_1 P, \pi_2 P \rangle \rightarrow_{\eta} P$;
+  7) $\lambda x. f x \rightarrow_{\eta} f$;
 
-  10) ${\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} N \rightarrow_{\eta} {\bf pure \:} N$;
+  8) $\langle \pi_1 P, \pi_2 P \rangle \rightarrow_{\eta} P$;
 
-  11) ${\bf let \: pure \:} x = M {\: \bf in \: } x \rightarrow_{\eta} M$;
+  9) ${\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} N \rightarrow_{\eta} {\bf pure \:} N$;
 
-  12) $M \rightarrow_{\beta} N \Rightarrow {\bf pure \: M} \rightarrow_{\eta} {\bf pure \: N}$
+  10) ${\bf let \: pure \:} x = M {\: \bf in \: } x \rightarrow_{\eta} M$;
+
+  11) $M \rightarrow_{\eta} N \Rightarrow {\bf pure \: M} \rightarrow_{\eta} {\bf pure \: N}$
   \end{defin}
-
-  Let us show the next simple observation, which immeadelty follows from the previous definition.
-
-  \begin{lemma}
-    $ $
-
-    If $M \rightarrow_{\beta \eta} N$, then ${\bf pure \:} M \rightarrow_{\beta \eta} {\bf pure \:} N$.
-  \end{lemma}
 
@@
 \subsection{Soundness}
+
+\begin{theorem} Soundness
+
+  Let $\Gamma \vdash M : A$ and $M =_{\beta\eta} N$, then $[\![\Gamma \vdash M : A]\!] = [\![\Gamma \vdash N : A]\!]$
+\end{theorem}
+
+\begin{proof}
 
@@
 
-1) $[\![\Gamma \vdash {\bf pure \:} ((\lambda x. M) N) : {\bf K}B]\!] = [\![\Gamma \vdash {\bf pure \:} (M [x := N]]) : {\bf K}B]\!]$
-
-$\begin{array}{lll}
-&[\![\Gamma \vdash {\bf pure \:} (\lambda x. M) N : {\bf K}B]\!] = & \text{By interpretation} \\
-&\eta_{[\![B]\!]} \circ (\epsilon \circ \langle \Lambda([\![M]\!]), [\![N]\!]\rangle) =& \text{Property of $\times$} \\
-&\eta_{[\![B]\!]} \circ (\epsilon \circ (\Lambda([\![M]\!]) \times id_{[\![A]\!]}) \circ \langle id_{[\![\Gamma]\!]}, [\![N]\!] \rangle) = & \text{Associativity of composition} \\
-&\eta_{[\![B]\!]} \circ ((\epsilon \circ (\Lambda([\![M]\!]) \times id_{[\![A]\!]})) \circ \langle id_{[\![\Gamma]\!]}, [\![N]\!] \rangle) = & \text{Exponentiation property} \\
-&\eta_{[\![B]\!]} \circ ([\![M]\!] \circ \langle id_{[\![\Gamma]\!]}, [\![N]\!] \rangle) = & \text{Substitution lemma} \\
-&\eta_{[\![B]\!]} \circ [\![M [\vec{x}, x := \vec{x}, N]]\!] =& \text{By interpretation} \\
-&[\![\Gamma \vdash {\bf pure \:} (M [x := N]]) : {\bf K}B]\!]&
-\end{array}$
-
-\vspace{\baselineskip}
-
-2) $[\![\Gamma \vdash {\bf pure \:} (\pi_i \langle [\![M_1]\!], [\![M_2]\!] \rangle) : {\bf K}A_i ]\!] = [\![\Gamma \vdash {\bf \: pure} M_i : {\bf K}A_i ]\!]$
+1) $[\![\Gamma \vdash {\bf let \: pure \:} \vec{x} = {\bf pure \:} \vec{M} {\: \bf in \:} N : {\bf K}B]\!] = [\![\Gamma \vdash {\bf pure \:} N [\vec{x} := \vec{M}] : {\bf K}B]\!]$
 
@@
 $\begin{array}{lll}
-&[\![\Gamma \vdash {\bf pure \:} (\pi_i \langle M_1, M_2] \rangle) : {\bf K}A_i ]\!] = & \text{By interpretation} \\
-&\eta_{[\![A_i]\!]} \circ \pi_i \circ \langle [\![M_1]\!], [\![M_2]\!] \rangle = & \text{Property of $\times$} \\
-&\eta_{[\![A_i]\!]} \circ [\![M_i]\!] =& \text{By interpetation} \\
-&[\![\Gamma \vdash {\bf pure \:} M_i : {\bf K}A_i]\!]&
+&[\![\Gamma \vdash {\bf let \: pure \:} \vec{x} = {\bf pure \:} \vec{M} {\: \bf in \:} N : {\bf K}B]\!] = & \\
+&\text{\quad\quad\quad\quad\quad\quad\quad\quad\quad By interpretation}& \\
+&\mathcal{K}([\![N]\!]) \circ \ast_{[\![A_1]\!],\dots,[\![A_n]\!]} \circ \langle \eta_{[\![A_1]\!]} \circ [\![M_1]\!],\dots,\eta_{[\![A_n]\!]} \circ [\![M_n]\!] \rangle = \\
+&\text{\quad\quad\quad\quad\quad\quad\quad\quad\quad By the property of a pair of morphisms}& \\
+&\mathcal{K}([\![N]\!]) \circ \ast_{[\![A_1]\!],\dots,[\![A_n]\!]} \circ (\eta_{[\![A_1]\!]} \times \dots \times \eta_{[\![A_n]\!]}) \circ \langle [\![M_1]\!], \dots, [\![M_n]\!]\rangle =& \\
+&\text{\quad\quad\quad\quad\quad\quad\quad\quad\quad Associativity of composition}& \\
+&\mathcal{K}([\![N]\!]) \circ (\ast_{[\![A_1]\!],\dots,[\![A_n]\!]} \circ (\eta_{[\![A_1]\!]} \times \dots \eta_{[\![A_n]\!]})) \circ \langle [\![M_1]\!], \dots, [\![M_n]\!] \rangle =& \\
+&\text{\quad\quad\quad\quad\quad\quad\quad\quad\quad By the definition of an applicative functor}& \\
+&\mathcal{K}([\![N]\!]) \circ \eta_{[\![A_1]\!] \times \dots \times [\![A_n]\!]} \circ \langle [\![M_1]\!], \dots, [\![M_n]\!] \rangle =& \\
+&\text{\quad\quad\quad\quad\quad\quad\quad\quad\quad Naturality of $\eta$}& \\
+&\eta_{[\![B]\!]} \circ [\![N]\!] \circ \langle [\![M_1]\!], \dots, [\![M_n]\!] \rangle =& \\
+&\text{\quad\quad\quad\quad\quad\quad\quad\quad\quad Associativity of composition}& \\
+&\eta_{[\![B]\!]} \circ ([\![N]\!] \circ \langle [\![M_1]\!], \dots, [\![M_n]\!]) \rangle =& \\
+&\text{\quad\quad\quad\quad\quad\quad\quad\quad\quad Simultaneous substitution lemma}& \\
+&\eta_{[\![B]\!]} \circ [\![N [\vec{x} := \vec{M}]]\!]& \\
+&\text{\quad\quad\quad\quad\quad\quad\quad\quad\quad By interpetation}& \\
+&[\![\Gamma \vdash {\bf pure \:} (N [\vec{x} := \vec{M}]) : {\bf K}B]\!]&
 \end{array}$
+
+2)
+
+If $\Gamma \vdash M : A$ and $M \rightarrow_{\beta} N$, then $[\![\Gamma \vdash {\bf pure \:} M : {\bf K}A]\!] = [\![\Gamma \vdash {\bf pure \:} N : {\bf K}A]\!]$.
+
+\vspace{\baselineskip}
+
+If $\Gamma \vdash M : A$ and $M \rightarrow_{\beta} N$, then $\Gamma \vdash N : A$ by subject reduction.
+
+By assumption $[\![\Gamma \vdash M : A]\!] = [\![\Gamma \vdash N : A]\!]$.
+
+So $\eta_{[\![A]\!]} \circ [\![\Gamma \vdash M : A]\!] = \eta_{[\![A]\!]} \circ [\![\Gamma \vdash N : A]\!]$.
+
+Hence $[\![\Gamma \vdash {\bf pure \:} M : {\bf K}A]\!] = [\![\Gamma \vdash {\bf pure \:} N : {\bf K}A]\!]$.
 
@@
 
-\vspace{\baselineskip}
-
-1) $[\![\Gamma \vdash {\bf pure \:} (\lambda x. M x) : {\bf K}(A \to B)]\!] = [\![\Gamma \vdash {\bf pure \:} M : {\bf K}(A \to B)]\!]$.
-
-\vspace{\baselineskip}
-
-$\begin{array}{lll}
-&[\![\Gamma \vdash {\bf pure \:} (\lambda x. M x) : {\bf K}(A \to B)]\!] = & \text{By interpetation} \\
-&\eta_{[\![B]\!]^{[\![A]\!]}} \circ \Lambda(\epsilon \circ [\![M]\!] \times id_{[\![A]\!])}& \text{Exponentiation property} \\
-&\eta_{[\![B]\!]^{[\![A]\!]}} \circ [\![M]\!] = & \text{By interpretation} \\
-&[\![\Gamma \vdash {\bf pure \:} M : {\bf K}(A \to B)]\!]&
-\end{array}$
-
-\vspace{\baselineskip}
-
-2) $[\![\Gamma \vdash {\bf pure \:} \langle \pi_1 M, \pi_2 M \rangle : {\bf K}(A \times B)]\!] = [\![\Gamma \vdash {\bf pure \:} M : {\bf K}(A \times B)]\!]$
-
-\vspace{\baselineskip}
-
-$\begin{array}{lll}
-&[\![\Gamma \vdash {\bf pure \:} \langle \pi_1 M, \pi_2 M \rangle : {\bf K}(A \times B)]\!] = & \text{By interpetation} \\
-&\eta_{[\![A]\!] \times [\![B]\!]} \circ \langle \pi_1 \circ [\![M]\!], \pi_2 \circ [\![M]\!] \rangle  = & \text{By the property of a product of morphisms} \\
-&\eta_{[\![A]\!] \times [\![B]\!]} \circ [\![M]\!] = & \text{By interpetation} \\
-&[\![\Gamma \vdash {\bf pure \:} M : {\bf K}(A \times B)]\!]&
-\end{array}$
-
-\vspace{\baselineskip}
-
-3) $[\![\vdash {\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} N : {\bf }K A ]\!] = [\![\vdash {\bf pure \:} N : {\bf K A}]\!]$.
+1) $[\![\vdash {\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} N : {\bf }K A ]\!] = [\![\vdash {\bf pure \:} N : {\bf K A}]\!]$.
 
@@
 
+\vspace{\baselineskip}
+
+2)
+
+If $\Gamma \vdash M : A$ and $M \rightarrow_{\eta} N$, then $[\![\Gamma \vdash {\bf pure \:} M : {\bf K}A]\!] = [\![\Gamma \vdash {\bf pure \:} N : {\bf K}A]\!]$.
+
+\vspace{\baselineskip}
+
+Similar to case with $\beta$-reduction.
+
 \end{proof}
 
-\begin{theorem} Soundness
+\end{proof}
 
-  Let $\Gamma \vdash M : A$ and $M =_{\beta\eta} N$, then $[\![\Gamma \vdash M : A]\!] = [\![\Gamma \vdash N : A]\!]$
-\end{theorem}
-
-\begin{proof}
-$ $
-Straightforwardly follows from two previous lemmas.
-\end{proof}
+\subsection{Completeness}
 

### 2018-01-19 article.tex
added rewritten lemma

@@
 
+It is clearly that these lemmas could be extended for IEL$^{-}$ with $\vee$ and $\neg$ similary.
+
   \vspace{\baselineskip}
@@
   example see \cite{Neder}\cite{Morten}\cite{Girard}. We will consider only modal cases}:
+
+\begin{lemma} Generation lemma.
+
+  i) Let $\Gamma \vdash {\bf pure \:} M : {\bf K}A$, then $\Gamma \vdash M : A$;
+
+  ii) Let $\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N : {\bf K}B$, there are some $A_1,\dots,A_n \in \mathbb{T}_{{\bf K}}$,
+  such that $\Gamma \vdash \vec{M} : {\bf K} \vec{A}$ and $\vec{x} : \vec{A} \vdash N : B$.
+\end{lemma}
+
+\begin{proof}
+  $ $
+
+  Induction on $\Gamma \vdash {\bf pure \:} M : {\bf K}A$ and $\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{N} {\: \bf in \:} N : {\bf K}B$ correspondently.
+\end{proof}
+
+\begin{lemma} Weakening.
+
+  Let $\Gamma \vdash M : A$ and $\Gamma \subseteq \Delta$, then $\Delta \vdash M : A$.
+
+\end{lemma}
+
+\begin{proof}
+  $ $
+
+  1) Let $\Gamma, x : A \vdash x : A$ and $\Gamma \subseteq \Delta$, then $\Delta, x : A \vdash x : A$ trivially.
+
+  2) Let $\Gamma \vdash {\bf pure \:} M : {\bf K}A$. Then $\Gamma \vdash M : A$ by generation and $\Delta \vdash M : A$ by assumption. So $\Delta \vdash {\bf pure \:} M : {\bf K} A$ by ${\bf K}_I$.
+
+  3) Let $\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N : {\bf K}B$ and $\Gamma \subseteq \Delta$.
+Then $\Gamma \vdash \vec{M} : {\bf K}\vec{A}$ and $\vec{x} : \vec{A} \vdash N : B$.
+
+  By assumption $\Delta \vdash \vec{M} : {\bf K}\vec{A}$. So $\Delta \vdash {\bf let \: pure \:} \vec{x} = \vec{N} {\: \bf in \:} N : {\bf K}B$ by $\text{let}_{{\bf K}}$.
+\end{proof}
+
+\begin{defin} Type substituition
+
+  The substituition of type $C$ for type variable $B$ in type $A$ inductively defined as follows:
+
+  1) $B [B := C] = B$ and $D [B := C] = D$, if $B \neq D$;
+
+  2) $(A_1 \alpha A_2)[B := C] = (A_1[B := C]) \alpha (A_2[B := C])$, where $\alpha \in \{ \to, \times \}$;
+
+  3) $({\bf K}A)[B := C] = {\bf K}(A [B := C])$.
+
+  4) Let $\Gamma$ be the context, then $\Gamma [B := C] = \{ x : (A[B := C]) \: | \: x : A \in \Gamma \}$
+\end{defin}
+
+\begin{lemma} Substituition lemma.
+
+    i) Let $\Gamma, x : A \vdash M : B$ and $\Gamma \vdash N : A$, then $\Gamma \vdash M [x := N] : B$.
+
+    ii) Let $\Gamma \vdash M : A$, then $\Gamma [B := C] \vdash M : (A [B := C])$.
+
+\end{lemma}
+
+\begin{proof}
+
+$ $
+
+i) For term substitution:
+
+\vspace{\baselineskip}
+
+1) Let $\Gamma, x : A \vdash x : A$ and $\Gamma \vdash N : A$, but $x [x := N] = N$, so $\Gamma \vdash N : A$.
+
+2) Let $\Gamma, x : A \vdash {\bf pure \:} M : {\bf K}B$ and $\Gamma \vdash N : A$.
+
+By generation $\Gamma, x : A \vdash M : B$ and by assumption $\Gamma \vdash M [x := N] : B$.
+
+By $K_I$, $\Gamma \vdash {\bf pure \:} (M [x := N]) : {\bf K}B$.
+
+3) Let $\Gamma, y : A \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N : {\bf K} B$ and $\Gamma \vdash N : A$.
+
+By generation, $\Gamma, y : A \vdash \vec{M} : {\bf K} \vec{A}$ and $\vec{x} : \vec{A} \vdash N : B$.
+
+By hypothesis, $\Gamma \vdash \vec{M}[x := N] : {\bf K} \vec{A}$.
+
+Hence $\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M}[x := N] {\: \bf in \:} N : {\bf K} B$.
+
+\vspace{\baselineskip}
+
+ii) For type substitution
+
+\vspace{\baselineskip}
+
+1) Let $\Gamma, x : A \vdash x : A$, so $\Gamma [A := C], x : (A[A := C]) \vdash x : (A[A := C])$, or
+
+$\Gamma [A := C], x : C \vdash x : C$.
+
+2) Let $\Gamma \vdash {\bf pure \:} M : {\bf K}A$. By generation $\Gamma \vdash M : A$ and by assumption $\Gamma [B := C] \vdash M : A[B := C]$.
+
+By $K_I$ $\Gamma \vdash {\bf pure \:L} M : {\bf K}(A[B := C])$.
+
+3) $\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N : {\bf K}B$. By generation $\Gamma \vdash \vec{M} : {\bf K}\vec{A}$ and $\vec{x} : \vec{A} \vdash N : B$.
+
+By assumption $\Gamma [B_1 := C] \vdash \vec{M} : {\bf }{K}\vec{A}[B_1 := C]$ and $\vec{x} : \vec{A}[B_1 := C] \vdash N : B[B_1 := C]$.
+
+So by $\text{let}_{{\bf K}}$, $\Gamma [B_1 := C] \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N : {\bf K}(B[B_1 := C])$.
+
+\end{proof}
+
+\begin{theorem} Subject reduction
+\end{theorem}
 

### 2018-01-20 article.tex
added subject reduction
small fixes

@@
 
-  6) $M \rightarrow_{\beta} N \Rightarrow {\bf pure \: M} \rightarrow_{\beta} {\bf pure \: N}$
+  6) $\lambda x. f x \rightarrow_{\eta} f$;
 
-  7) $\lambda x. f x \rightarrow_{\eta} f$;
+  7) $\langle \pi_1 P, \pi_2 P \rangle \rightarrow_{\eta} P$;
 
-  8) $\langle \pi_1 P, \pi_2 P \rangle \rightarrow_{\eta} P$;
+  8) ${\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} N \rightarrow_{\eta} {\bf pure \:} N$;
 
-  9) ${\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} N \rightarrow_{\eta} {\bf pure \:} N$;
+  9) ${\bf let \: pure \:} x = M {\: \bf in \: } x \rightarrow_{\eta} M$;
 
-  10) ${\bf let \: pure \:} x = M {\: \bf in \: } x \rightarrow_{\eta} M$;
-
-  11) $M \rightarrow_{\eta} N \Rightarrow {\bf pure \: M} \rightarrow_{\eta} {\bf pure \: N}$
+  10) $M \rightarrow_{\beta \eta} N \Rightarrow {\bf pure \: M} \rightarrow_{\beta \eta} {\bf pure \: N}$
   \end{defin}
@@
 \begin{theorem} Subject reduction
+
+  Let $\Gamma \vdash M : A$ and $M \twoheadrightarrow_{\beta \eta} N$, then $\Gamma \vdash N : A$
+
 \end{theorem}
+
+For cases with application, abstraction and pairs see \cite{Morten} \cite{Pierce}.
+
+1)
+  Let $\Gamma \vdash {\bf let \: pure \:} \langle \vec{x}, y, \vec{z} \rangle = \langle \vec{M}, {\bf let \: pure \: } \vec{w} = \vec{N} {\bf \: in \: } Q, \vec{P} \rangle {\: in \:} R : {\bf K}B$,
+  then $\Gamma {\bf let \: pure \:} \langle \vec{x}, \vec{w}, \vec{z} \rangle = \langle \vec{M}, \vec{N}, \vec{P} \rangle \: { \bf in } \: R [y := Q] : {\bf K}B$
+
+2) Let $\Gamma \vdash {\bf let \: pure \:} x = M {\: \bf in \: } x : {\bf K}A$, then $\Gamma \vdash M : {\bf K}A$.
+
+See \cite{ModalK}.
+
+3) Let $\Gamma \vdash {\bf let \: pure \:} \vec{x} = {\bf pure \:} \vec{M} {\: \bf in \:} N : {\bf K}B$.
+
+By generation $\Gamma \vdash {\bf pure \:} \vec{M} : {\bf K}\vec{A}$ and $\vec{x} : \vec{A} \vdash N : B$.
+\\ Moreover, $\Gamma \vdash \vec{M} : \vec{A}$. By weakening and substitution lemma $\Gamma \vdash N [\vec{x} = \vec{M}] : B$.
+\\ By ${\bf K}_I$, $\Gamma \vdash {\bf pure \:} N [\vec{x} := \vec{M}] : {\bf K}B$.
+
+4) Let $\vdash {\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} N : {\bf K}A$
+
+By generation $\vdash N : A$. \\ So $\vdash {\bf pure \:} N : {\bf K}A$ by ${\bf K}_I$.
+
+5) Let $\Gamma \vdash {\bf pure \:} M : A$ and $M \twoheadrightarrow_{\beta \eta} N$.
+
+By generation $\Gamma \vdash M : A$ and $\Gamma \vdash N : A$ by assumption.
+\\ So $\Gamma \vdash {\bf pure \:} N : {\bf K}A$.
 
@@
 
-  Let $\langle \mathcal{C}, \oplus_1, \mathds{1} \rangle$ and $\langle \mathcal{D}, \oplus_2, \mathds{1}' \rangle$ are monoidal categories.
+  Let $\langle \mathcal{C}, \otimes_1, \mathds{1} \rangle$ and $\langle \mathcal{D}, \otimes_2, \mathds{1}' \rangle$ are monoidal categories.
 
-  A lax monoidal functor $\mathcal{F} : \langle \mathcal{C}, \oplus_1, \mathds{1} \rangle \to \langle \mathcal{D}, \oplus_2, \mathds{1}' \rangle$ is a functor
+  A lax monoidal functor $\mathcal{F} : \langle \mathcal{C}, \otimes_1, \mathds{1} \rangle \to \langle \mathcal{D}, \otimes_2, \mathds{1}' \rangle$ is a functor
   $\mathcal{F} : \mathcal{C} \to \mathcal{D}$ with additional natural transformations:

### 2018-01-20 article.tex
added
added more draft

@@
 
+We will consider term model for simply typed lambda calculus $\times$ and $\to$ standardly described in \cite{LambekScott} \cite{Abramsky}.
+
+\begin{defin}
+  Let us define an endofunctor $\mathcal{K} : \mathcal{C}(\lambda) \to \mathcal{C}(\lambda)$, such that:
+  \begin{itemize}
+    \item ${\bf K} : A \mapsto {\bf K}A$;
+    \item ${\bf K} : [x,M] \in Hom_{\mathcal{C}(\lambda)}(A,B) \mapsto \text{fmap } f = [y, {\bf let \: pure \:} x = y \in M] \in Hom_{\mathcal{C}(\lambda)}({\bf K} A, {\bf K} B)$.
+  \end{itemize}
+\end{defin}
+
+\begin{lemma} Functoriality
+
+  i) ${\bf K}(g \circ f) = {\bf K}(g) \circ {\bf K}(f)$;
+
+  ii) ${\bf K}(id_A) = id_{{\bf K}A}$.
+\end{lemma}
+
+\begin{proof}
+
+Easy checking.
+
+\end{proof}
+
+\begin{defin}
+
+  Let us define natural transformations:
+  \begin{itemize}
+  \item $\eta : Id_{\mathcal{C}(\lambda)} \to \mathcal{K}$, such that forall $A \in Ob_{\mathcal{C}(\lambda)}$, $\eta_A = [x, {\bf pure \:} x] \in Hom_{\mathcal{C}(\lambda)}(A, {\bf K}A)$;
+
+  \item $\ast_{A,B} : {\bf K}A \times {\bf K}B \to {\bf K}(A \times B)$
+  \end{itemize}
+\end{defin}
+
+\begin{lemma} Naturality for $\eta$ and for $\ast$
+
+  i) $\text{fmap } f \circ \eta_A = \eta_B \circ f$;
+
+  ii) $\text{fmap } (f \times g) \circ \ast_{A, B} = \ast_{C,D} \circ (\text{fmap } f) \times (\text{fmap } g)$.
+
+  iii) $\ast_{A,B} \circ (\eta_{A} \times \eta_{B}) = \eta_{A \times B}$;
+\end{lemma}
+
+\begin{proof}
+  $ $
+
+  i) $\text{fmap } f \circ \eta_A = \eta_B \circ f$
+
+\vspace{\baselineskip}
+
+$\begin{array}{lll}
+&\eta_B \circ f = & \\
+&[y, {\bf pure \:} y] \circ [x, M] = & \\
+&[x, {\bf pure \:} y [y := M]] =& \\
+&[x, {\bf pure \:} M]& \\
+&\text{On the other hand}& \\
+&\text{fmap } f \circ \eta_A = &\\
+&[z, {\bf let \: pure \:} x = z {\: \bf in \:} M] \circ [x, {\bf pure \: x}] = & \\
+&[x, {\bf let \: pure \:} x = z {\: \bf in \:} M [z := {\bf pure \:} x]]& \\
+&[x, {\bf let \: pure \:} x = {\bf pure \: x} {\: \bf in \:} M]& \\
+&[x, {\bf pure \:} M [x := x]]& \\
+&[x, {\bf pure \:} M]&
+\end{array}$
+
+\vspace{\baselineskip}
+
+ii)  $\text{fmap } (f \times g) \circ \ast_{A, B} = \ast_{C,D} \circ (\text{fmap } f) \times (\text{fmap } g)$
+
+\end{proof}
+
+\begin{lemma} Tensorial strength
+\end{lemma}
+
+\begin{lemma}
+
+  ${\bf K}$ is an applicative functor.
+\end{lemma}
+
+\begin{theorem} Completeness
+
+  Let $[\![\Gamma \vdash M : A]\!] = [\![\Gamma \vdash N : A]\!]$, then $M =_{\beta \eta} N$.
+\end{theorem}
+
   \addcontentsline{toc}{section}{References}
@@
   \bibitem{Abe} T. Abe. Completeness of modal proofs in first-order predicate logic. Computer Software, JSSST Journal, 24:165 -- 177, 2007.
+
+  \bibitem{LambekScott}
+
+  \bibitem{Abramsky}
+
+  \bibitem{Gabbay}
+
 \end{thebibliography}

### 2018-01-20 article.tex
small fixes

@@
 
-Here are some examples of derivation trees.
+Here are some examples of derivable terms:
 
-  \begin{prooftree}
-  \AxiomC{$x : A \vdash x : A$}
-  \RightLabel{$\textbf{K}_I$}
-  \UnaryInfC{$x : A \vdash {\bf pure \: } x : \textbf{K} A$}
-  \RightLabel{$\rightarrow_i$}
-  \UnaryInfC{$\vdash (\lambda x. {\bf pure \: } x) : A \to \textbf{K}A$}
-  \end{prooftree}
+1) $\vdash (\lambda x. {\bf pure \: } x) : A \to \textbf{K}A$;
 
-  \begin{prooftree}
-    \AxiomC{$f : {\bf K} (A \to B) \vdash f : {\bf K} (A \to B)$}
-    \AxiomC{$x : {\bf K} A \vdash x : {\bf K} A $}
-    \AxiomC{$g : A \to B$}
-    \AxiomC{$y : A$}
-    \BinaryInfC{$g : A \to B, y : A \vdash g y : B$}
-    \TrinaryInfC{$f : {\bf K} (A \to B), x : {\bf K} A \vdash {\bf let \:\: pure} \:\: \langle g, y \rangle = \langle f, x \rangle \:\: {\bf in } \:\: g y : {\bf K} B$}
-    \UnaryInfC{$f : {\bf K} (A \to B) \vdash \lambda x. {\bf let \:\: pure} \:\: \langle g, y \rangle = \langle f, x \rangle \:\: {\bf in } \:\: g y : {\bf K} A \to {\bf K} B$}
-    \UnaryInfC{$\vdash \lambda f. \lambda x. {\bf let \:\: pure} \:\: \langle g, y \rangle = \langle f, x \rangle \:\: {\bf in } \:\: g y : {\bf K}(A \to B) \to {\bf K} A \to {\bf K} B$}
-  \end{prooftree}
+2) $\vdash \lambda f. \lambda x. {\bf let \: pure} \: \langle g, y \rangle = \langle f, x \rangle {\: \bf in \:} g y : {\bf K}(A \to B) \to {\bf K} A \to {\bf K} B$
 
-\begin{prooftree}
-  \AxiomC{$f : A \to B \vdash f : A \to B$}
-  \UnaryInfC{$f : A \to B \vdash {\bf pure \:\:} f : {\bf K }(A \to B)$}
-  \AxiomC{$x : {\bf K} A \vdash x : {\bf K} A $}
-  \AxiomC{$g : A \to B$}
-  \AxiomC{$y : A$}
-  \BinaryInfC{$g : A \to B, y : A \vdash g y : B$}
-  \TrinaryInfC{$f : A \to B, x : {\bf K} A \vdash {\bf let \:\: pure} \:\: \langle g, y \rangle = \langle {\bf pure \:\:} f, x \rangle \:\: {\bf in } \:\: g y : {\bf K} B$}
-  \UnaryInfC{$f : A \to B \vdash \lambda x. {\bf let \:\: pure} \:\: \langle g, y \rangle = \langle {\bf pure \:\:} f, x \rangle \:\: {\bf in } \:\: g y : {\bf K} A \to {\bf K} B$}
-  \UnaryInfC{$\lambda f. \lambda x. {\bf let \:\: pure} \:\: \langle g, y \rangle = \langle {\bf pure \:\:} f, x \rangle \:\: {\bf in } \:\: g y : (A \to B) \to {\bf K} A \to {\bf K} B$}
-\end{prooftree}
+3) $\lambda f. \lambda x. {\bf let \:pure} \: \langle g, y \rangle = \langle {\bf pure \:} f, x \rangle {\: \bf in \: } g y : (A \to B) \to {\bf K} A \to {\bf K} B$
 
@@
 
-2) $\ast_{A,B} \circ (\eta_A \otimes \eta_B) = \eta_{A \otimes B}$;
-
-3) Weak commutativity condition holds:
-
-\xymatrix{
-A \otimes \mathcal{K}B & \mathcal{K}A \otimes \mathcal{K}B & \mathcal{K}(A \otimes B) \\
-\mathcal{K}B \otimes A & \mathcal{K}B \otimes \mathcal{K}A & \mathcal{K}(B \otimes A)
-}
+2) $\ast_{A,B} \circ (\eta_A \otimes \eta_B) = \eta_{A \otimes B}$.
 \end{defin}
@@
 
-If $\Gamma \vdash M : A$ and $M \rightarrow_{\beta} N$, then $[\![\Gamma \vdash {\bf pure \:} M : {\bf K}A]\!] = [\![\Gamma \vdash {\bf pure \:} N : {\bf K}A]\!]$.
+If $\Gamma \vdash M : A$ and $M \rightarrow_{\beta \eta} N$, then $[\![\Gamma \vdash {\bf pure \:} M : {\bf K}A]\!] = [\![\Gamma \vdash {\bf pure \:} N : {\bf K}A]\!]$.
 
@@
 
-If $\Gamma \vdash M : A$ and $M \rightarrow_{\beta} N$, then $\Gamma \vdash N : A$ by subject reduction.
+If $\Gamma \vdash M : A$ and $M \rightarrow_{\beta \eta} N$, then $\Gamma \vdash N : A$ by subject reduction.
 
@@
 
-\vspace{\baselineskip}
-
-2)
-
-If $\Gamma \vdash M : A$ and $M \rightarrow_{\eta} N$, then $[\![\Gamma \vdash {\bf pure \:} M : {\bf K}A]\!] = [\![\Gamma \vdash {\bf pure \:} N : {\bf K}A]\!]$.
-
-\vspace{\baselineskip}
-
-Similar to case with $\beta$-reduction.
-
 \end{proof}
@@
   Let us define an endofunctor $\mathcal{K} : \mathcal{C}(\lambda) \to \mathcal{C}(\lambda)$, such that:
-  \begin{itemize}
-    \item ${\bf K} : A \mapsto {\bf K}A$;
-    \item ${\bf K} : [x,M] \in Hom_{\mathcal{C}(\lambda)}(A,B) \mapsto \text{fmap } f = [y, {\bf let \: pure \:} x = y \in M] \in Hom_{\mathcal{C}(\lambda)}({\bf K} A, {\bf K} B)$.
-  \end{itemize}
+
+1) ${\bf K} : A \mapsto {\bf K}A$;
+
+2) ${\bf K} : [x,M] \in Hom_{\mathcal{C}(\lambda)}(A,B) \mapsto \text{fmap } f = [y, {\bf let \: pure \:} x = y {\: \bf in \:} M] \in Hom_{\mathcal{C}(\lambda)}({\bf K} A, {\bf K} B)$.
+
 \end{defin}
@@
 
-Easy checking.
+Easy checking using reduction rules.
 
@@
 &[z, {\bf let \: pure \:} x = z {\: \bf in \:} M] \circ [x, {\bf pure \: x}] = & \\
-&[x, {\bf let \: pure \:} x = z {\: \bf in \:} M [z := {\bf pure \:} x]]& \\
-&[x, {\bf let \: pure \:} x = {\bf pure \: x} {\: \bf in \:} M]& \\
-&[x, {\bf pure \:} M [x := x]]& \\
+&[x, {\bf let \: pure \:} x = z {\: \bf in \:} M [z := {\bf pure \:} x]] = & \\
+&[x, {\bf let \: pure \:} x = {\bf pure \: x} {\: \bf in \:} M] = & \\
+&[x, {\bf pure \:} M [x := x]] = & \\
 &[x, {\bf pure \:} M]&
@@
 
+\vspace{\baselineskip}
+
+iii) $\ast_{A,B} \circ (\eta_{A} \times \eta_{B}) = \eta_{A \times B}$
+
+Follows from ii).
+
 \end{proof}
@@
 
+\begin{lemma} Weak commutativity.
+\end{lemma}
+
+\begin{proof}
+  Follows from tensorial strength and commutativity of product.
+\end{proof}
+
 \begin{lemma}
@@
 \end{lemma}
+
+\begin{proof}
+  Follows from previous lemmas.
+\end{proof}
 

### 2018-01-20 article.tex
some text

@@
   Let us define natural transformations:
-  \begin{itemize}
-  \item $\eta : Id_{\mathcal{C}(\lambda)} \to \mathcal{K}$, such that forall $A \in Ob_{\mathcal{C}(\lambda)}$, $\eta_A = [x, {\bf pure \:} x] \in Hom_{\mathcal{C}(\lambda)}(A, {\bf K}A)$;
 
-  \item $\ast_{A,B} : {\bf K}A \times {\bf K}B \to {\bf K}(A \times B)$
-  \end{itemize}
+  1) $\eta:Id \Rightarrow \mathcal{K}$, s. t. $\forall A \in Ob_{\mathcal{C}(\lambda)}$, $\eta_A = [x, {\bf pure \:} x] \in Hom_{\mathcal{C}(\lambda)}(A, {\bf K}A)$;
+
+  2) $\ast_{A,B}:{\bf K}A \times {\bf K}B \to {\bf K}(A \times B)$.
 \end{defin}
@@
 $\begin{array}{lll}
-&\eta_B \circ f = & \\
-&[y, {\bf pure \:} y] \circ [x, M] = & \\
-&[x, {\bf pure \:} y [y := M]] =& \\
+&\eta_B \circ f = & \text{By the definition} \\
+&[y, {\bf pure \:} y] \circ [x, M] = & \text{By the definition of composition}\\
+&[x, {\bf pure \:} y [y := M]] = & \text{By substitution}\\
 &[x, {\bf pure \:} M]& \\
-&\text{On the other hand}& \\
-&\text{fmap } f \circ \eta_A = &\\
-&[z, {\bf let \: pure \:} x = z {\: \bf in \:} M] \circ [x, {\bf pure \: x}] = & \\
-&[x, {\bf let \: pure \:} x = z {\: \bf in \:} M [z := {\bf pure \:} x]] = & \\
-&[x, {\bf let \: pure \:} x = {\bf pure \: x} {\: \bf in \:} M] = & \\
-&[x, {\bf pure \:} M [x := x]] = & \\
+& &\\
+&\text{On the other hand:}& \\
+&\text{fmap } f \circ \eta_A = & \text{By the definiton} \\
+&[z, {\bf let \: pure \:} x = z {\: \bf in \:} M] \circ [x, {\bf pure \: x}] = & \text{By the definition of composition}\\
+&[x, {\bf let \: pure \:} x = z {\: \bf in \:} M [z := {\bf pure \:} x]] = & \text{By substitution} \\
+&[x, {\bf let \: pure \:} x = {\bf pure \: x} {\: \bf in \:} M] = & \text{$\beta$-reduction rule}\\
+&[x, {\bf pure \:} M [x := x]] = & \text{By substitution}\\
 &[x, {\bf pure \:} M]&

### 2018-01-21 article.tex
added some trees
added tensorial strength

@@
 
-  2) $\ast_{A,B}:{\bf K}A \times {\bf K}B \to {\bf K}(A \times B)$.
+  2) $\ast_{A,B}:{\bf K}A \times {\bf K}B \to {\bf K}(A \times B)$, s.
+t.
+$\forall A, B \in Ob_{\mathcal{C}(\lambda)}, \ast_{A,B} = [p, {\bf let \: pure \:} x,y = \pi_1 p, \pi_2 p {\: \bf in \:} \langle x, y \rangle] \in Hom_{\mathcal{C}(\lambda)}({\bf K}A \times {\bf K}B, {\bf K}(A \times B))$.
 \end{defin}
+
+Implementation for $\ast$ in our term model is a modification of ${\text{let}_{\bf K}}$-rule:
+
+\begin{prooftree}
+\AxiomC{$p : {\bf K}A \times {\bf K}B \vdash p : {\bf K}A \times {\bf K}B$}
+\UnaryInfC{$p : {\bf K}A \times {\bf K}B \vdash \pi_1 p : {\bf K} A$}
+\AxiomC{$p : {\bf K}A \times {\bf K}B \vdash p : {\bf K}A \times {\bf K}B$}
+\UnaryInfC{$p : {\bf K}A \times {\bf K}B \vdash \pi_2 p : {\bf K} B$}
+\AxiomC{$x : A \vdash x : A$}
+\AxiomC{$y : B \vdash y : B$}
+\BinaryInfC{$x : A, y : B \vdash \langle x, y \rangle : A \times B$}
+\TrinaryInfC{$p : {\bf K}A \times {\bf K}B \vdash {\bf let \: pure \:} \langle x,y \rangle = \langle \pi_1 p, \pi_2 p \rangle {\: \bf in \:} \langle x, y \rangle : {\bf K}(A \times B)$}
+\end{prooftree}
 
@@
 
+See \cite{ModalK}
+
+\vspace{\baselineskip}
+
 iii) $\ast_{A,B} \circ (\eta_{A} \times \eta_{B}) = \eta_{A \times B}$
 
-Follows from ii).
+Follows from i) and ii).
 
@@
 
-\begin{lemma} Tensorial strength
-\end{lemma}
+  Tensorial strength is defined as follows:
+
+\begin{defin} Tensorial strength
+  $ $
+
+  Let $[p, \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle] \in Hom_{\mathcal{C}(\lambda)}(A \times {\bf K}B, {\bf K}A \times {\bf K}B)$.
+
+  So tensorial strength is defined as $\tau_{A, B} = \ast_{A, B} \circ [p, \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle]$.
+\end{defin}
+
+It is clearly that tensorial strength defined above can be simplified as follows:
+
+$\begin{array}{lll}
+&\ast_{A, B} \circ [p, \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle] = & \text{By definition} \\
+&[p^{'}, {\bf let \: pure \:} x, y = \pi_1 p^{'}, \pi_2 p^{'} {\: \bf in \:} \langle x, y \rangle] \circ [p, \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle] = & \text{By composition} \\
+&[p, {\bf let \: pure \:} x, y = \pi_1 p^{'}, \pi_2 p^{'} {\: \bf in \:} \langle x, y \rangle [p^{'} := \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle]] = & \text{By substitution} \\
+&[p, {\bf let \: pure \:} x, y = \pi_1 (\langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle), \pi_2 (\langle \pi_1 p, {\bf pure \:} (\pi_2 p) \rangle) {\: \bf in \:} \langle x, y \rangle] = & \text{By $\beta$-reduction rules} \\
+&[p, {\bf let \: pure \:} x, y = {\bf pure \:} (\pi_1 p), \pi_2 p {\: \bf in \:} \langle x, y \rangle]&
+\end{array}$
 
 \begin{lemma} Weak commutativity.
+  $ $
+
+  $\begin{array}{lll}
+  & \text{fmap } ([p, \langle \pi_2 p, \pi_1 p \rangle]) \circ \ast_{A, B} \circ [p, \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle] = & \\
+  & \ast_{B, A} \circ [p^{'}, \langle \pi_1 p^{'}, {\bf pure \:} (\pi_2 p) \rangle] \circ [p, \langle \pi_2 p, \pi_1 p \rangle]&
+  \end{array}$
 \end{lemma}

### 2018-01-22 article.tex
finished completeness

@@
 
-See \cite{ModalK}
+See \cite{ModalK}.
 
@@
   $\begin{array}{lll}
-  & \text{fmap } ([p, \langle \pi_2 p, \pi_1 p \rangle]) \circ \ast_{A, B} \circ [p, \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle] = & \\
-  & \ast_{B, A} \circ [p^{'}, \langle \pi_1 p^{'}, {\bf pure \:} (\pi_2 p) \rangle] \circ [p, \langle \pi_2 p, \pi_1 p \rangle]&
+  & \text{fmap } ([p, \langle \pi_2 p, \pi_1 p \rangle]) \circ \tau_{A,B} = & \\
+  & \ast_{B, A} \circ [q, \langle \pi_1 q, {\bf pure \:} (\pi_2 q) \rangle] \circ [p, \langle \pi_2 p, \pi_1 p \rangle]&
   \end{array}$
@@
 \begin{proof}
-  Follows from tensorial strength and commutativity of product.
-\end{proof}
-
-\begin{lemma}
-
-  ${\bf K}$ is an applicative functor.
-\end{lemma}
-
-\begin{proof}
-  Follows from previous lemmas.
+  $ $
+
+  $\begin{array}{lll}
+  & \text{fmap } ([r, \langle \pi_2 r, \pi_1 r \rangle]) \circ \tau_{A,B} = & \\
+  &\text{By the definition of $\tau$}& \\
+  & \text{fmap } ([r, \langle \pi_2 r, \pi_1 r \rangle]) \circ [p, {\bf let \: pure \:} x, y = {\bf pure \:} (\pi_1 p), \pi_2 p {\: \bf in \:} \langle x, y \rangle] = & \\
+  &\text{By the definition of fmap}& \\
+  & [q, {\bf let \: pure \:} r = q {\: \bf in \:} \langle \pi_2 r, \pi_1 r \rangle] \circ [p, {\bf let \: pure \:} x, y = {\bf pure \:} (\pi_1 p), \pi_2 p {\: \bf in \:} \langle x, y \rangle] = & \\
+  &\text{Composition}& \\
+  & [p, {\bf let \: pure \:} r = q {\: \bf in \:} \langle \pi_2 r, \pi_1 r \rangle [q := {\bf let \: pure \:} x, y = {\bf pure \:} (\pi_1 p), \pi_2 p {\: \bf in \:} \langle x, y \rangle]] = &\\
+  &\text{By $\beta$-reduction rules}& \\
+  & [p, {\bf let \: pure \:} r = ({\bf let \: pure \:} x, y = {\bf pure \:} (\pi_1 p), \pi_2 p {\: \bf in \:} \langle x, y \rangle) {\: \bf in \:} \langle \pi_2 r, \pi_1 r \rangle] = & \\
+  &\text{By $\beta$-reduction rules}& \\
+  & [p, {\bf let \: pure \:} x, y = {\bf pure \:} (\pi_1 p), \pi_2 p {\: \bf in \:} \langle \pi_2 r, \pi_1 r \rangle [r := \langle x, y \rangle]] = & \\
+  &\text{By substitution}& \\
+  & [p, {\bf let \: pure \:} x, y = {\bf pure \:} (\pi_1 p), \pi_2 p {\: \bf in \:} \langle \pi_2 \langle x, y \rangle, \pi_1 \langle x, y \rangle \rangle]= & \\
+  &\text{By $\beta$-reduction rules}& \\
+  & [p, {\bf let \: pure \:} x, y = {\bf pure \:} (\pi_1 p), \pi_2 p {\: \bf in \:} \langle y, x \rangle ] = & \\
+  & & \\
+  &\text{On the other hand}&\\
+  &\ast_{B, A} \circ [q, \langle \pi_1 q, {\bf pure \:} (\pi_2 q) \rangle] \circ [p, \langle \pi_2 p, \pi_1 p \rangle] = & \\
+  &\text{By the definition of $\ast$}& \\
+  &[r, {\bf let \: pure \:} y,x = \pi_1 r, \pi_2 r {\: \bf in \:} \langle y, x \rangle] \circ [q, \langle \pi_1 q, {\bf pure \:} (\pi_2 q) \rangle] \circ [p, \langle \pi_2 p, \pi_1 p \rangle] = & \\
+  &\text{Composition}& \\
+  &[r, {\bf let \: pure \:} y,x = \pi_1 r, \pi_2 r {\: \bf in \:} \langle y, x \rangle] \circ [p, \langle \pi_1 q, {\bf pure \:} (\pi_2 q) \rangle [q := \langle \pi_2 p, \pi_1 p  \rangle]] = & \\
+  &\text{By subsitution and by  $\beta$-reduction rules}& \\
+  &[r, {\bf let \: pure \:} y,x = \pi_1 r, \pi_2 r {\: \bf in \:} \langle y, x \rangle] \circ [p, \langle \pi_2 p, {\bf pure \:} (\pi_1 p) \rangle ]] = & \\
+  &\text{Composition}& \\
+  &[p, {\bf let \: pure \:} y,x = \pi_1 r, \pi_2 r {\: \bf in \:} \langle y, x \rangle [r := \langle \pi_2 p, {\bf pure \:} (\pi_1 p) \rangle]] = & \\
+  &\text{By subsitution and by  $\beta$-reduction rules}& \\
+  &[p, {\bf let \: pure \:} y,x = \pi_2 p, {\bf pure \:} (\pi_1 p) {\: \bf in \:} \langle y, x \rangle] = & \\
+  &\text{By symmetricity of assingment}&\\
+  &[p, {\bf let \: pure \:} x,y = {\bf pure \:} (\pi_1 p), \pi_2 p {\: \bf in \:} \langle y, x \rangle]&
+  \end{array}$
 \end{proof}
@@
 \begin{theorem} Completeness
-
-  Let $[\![\Gamma \vdash M : A]\!] = [\![\Gamma \vdash N : A]\!]$, then $M =_{\beta \eta} N$.
+  $ $
+
+  i) {\bf K} is an applicative functor;
+
+  ii) Let $[\![\Gamma \vdash M : A]\!] = [\![\Gamma \vdash N : A]\!]$, then $M =_{\beta \eta} N$.
 \end{theorem}
+
+\begin{proof}
+  Immediately follows from previous lemmas in the section.
+\end{proof}
 

### 2018-01-24 article.tex
small updates

@@
 
-  At first we define the natural deduction for IEL$^{-}$ with $\textbf{K}$-modality and binary connectives
-  $\to$ and $\land$ (we call that calculus NIEL$^{-}_{\land, \to}$):
-
-  \begin{defin} Natural deduction NIEL$^{-}_{\land, \to}$ for IEL$^{-}$ with $\to$ and $\wedge$:
-
-  \begin{center}
-  \begin{prooftree}
-  \AxiomC{}
-  \RightLabel{\scriptsize{ax}}
-  \UnaryInfC{$\Gamma , A \vdash A$}
-  \end{prooftree}
-  \end{center}
+  At first we define the natural deduction for IEL$^{-}$ :
+
+  \begin{defin} Natural deduction NIEL for IEL$^{-}$ is an extenstion of intuitionistic natural deduction with additional derivation rules for modality:
 
@@
     \begin{flushleft}
-  	\begin{prooftree}
-  	\AxiomC{$\Gamma, A \vdash B$}
-  	\RightLabel{$\rightarrow_i$}
-  	\UnaryInfC{$\Gamma \vdash A \to B$}
-    	\end{prooftree}
-
-  	\begin{prooftree}
-  	\AxiomC{ $\Gamma \vdash A$ }
-  	\AxiomC{ $\Gamma \vdash B$ }
-  	\RightLabel{$\land_i$}
-  	\BinaryInfC{$\Gamma \vdash A \land B$}
-  	\end{prooftree}
-
   	\begin{prooftree}
@@
     \begin{flushright}
-    	\begin{prooftree}
-  	\AxiomC{$\Gamma \vdash A \to B$}
-  	\AxiomC{$\Gamma \vdash A$}
-  	\RightLabel{$\rightarrow_i$}
-  	\BinaryInfC{$\Gamma \vdash B$}
-    	\end{prooftree}
-
-  	\begin{prooftree}
-  	\AxiomC{ $\Gamma \vdash A_1 \land A_2$ }
-  	\RightLabel{$\land_e, i \in \{ 1, 2 \}$}
-  	\UnaryInfC{$\Gamma \vdash A_i$}
-  	\end{prooftree}
-
   	\begin{prooftree}
@@
 
-It is clearly that these lemmas could be extended for IEL$^{-}$ with $\vee$ and $\neg$ similary.
-
   \vspace{\baselineskip}
@@
       \begin{prooftree}
-        \AxiomC{$\Gamma \vdash x : A$}
+        \AxiomC{$\Gamma \vdash M : A$}
         \RightLabel{$\textbf{K}_I$}
-        \UnaryInfC{$\Gamma \vdash {\bf pure \: } \: x : {\bf K}A $}
+        \UnaryInfC{$\Gamma \vdash {\bf pure \: } \: M : {\bf K}A $}
       \end{prooftree}
@@
     \begin{prooftree}
-    \AxiomC{$\Gamma \vdash f : A \to B$}
-    \AxiomC{$\Gamma \vdash x : A$}
+    \AxiomC{$\Gamma \vdash M : A \to B$}
+    \AxiomC{$\Gamma \vdash N : A$}
     \RightLabel{$\rightarrow_e$}
-    \BinaryInfC{$\Gamma \vdash fx : B$}
+    \BinaryInfC{$\Gamma \vdash MN : B$}
     \end{prooftree}
@@
 
-  $\text{let}_{{\bf K}}$ is the same as the $\Box$-rule in typed lambda calculus for intuitionistic normal modal logic {\bf IK}, which is described in \cite{ModalK}.
+  $\text{let}_{{\bf K}}$ is similar to the $\Box$-rule in typed lambda calculus for intuitionistic normal modal logic {\bf IK}, which is described in \cite{ModalK}.
 
@@
 
-Here are some examples of derivable terms:
-
-1) $\vdash (\lambda x. {\bf pure \: } x) : A \to \textbf{K}A$;
-
-2) $\vdash \lambda f. \lambda x. {\bf let \: pure} \: \langle g, y \rangle = \langle f, x \rangle {\: \bf in \:} g y : {\bf K}(A \to B) \to {\bf K} A \to {\bf K} B$
-
-3) $\lambda f. \lambda x. {\bf let \:pure} \: \langle g, y \rangle = \langle {\bf pure \:} f, x \rangle {\: \bf in \: } g y : (A \to B) \to {\bf K} A \to {\bf K} B$
+Here are some examples of closed terms:
+
+\begin{itemize}
+\item $(\lambda x. {\bf pure \: } x) : A \to \textbf{K}A$;
+\item $\lambda f. \lambda x. {\bf let \: pure} \: \langle g, y \rangle = \langle f, x \rangle {\: \bf in \:} g y : {\bf K}(A \to B) \to {\bf K} A \to {\bf K} B$
+\item $\lambda f. \lambda x. {\bf let \:pure} \: \langle g, y \rangle = \langle {\bf pure \:} f, x \rangle {\: \bf in \: } g y : (A \to B) \to {\bf K} A \to {\bf K} B$
+\end{itemize}
 
@@
 
+\begin{proof}
 For cases with application, abstraction and pairs see \cite{Morten} \cite{Pierce}.
@@
 \\ So $\Gamma \vdash {\bf pure \:} N : {\bf K}A$.
-
-\section{Strong normalization}
-
-\section{Confluence}
+\end{proof}
+
+Strong normalization and colfuence for ${\bf IK}$ was proved by Kakutani for call-by-value and for call-by name \cite{ModalK} \cite{ModalK1}.
 
@@
 
-  A lax monoidal functor $\mathcal{F} : \langle \mathcal{C}, \otimes_1, \mathds{1} \rangle \to \langle \mathcal{D}, \otimes_2, \mathds{1}' \rangle$ is a functor
-  $\mathcal{F} : \mathcal{C} \to \mathcal{D}$ with additional natural transformations:
+  A monoidal functor $\mathcal{F} : \langle \mathcal{C}, \otimes_1, \mathds{1} \rangle \to \langle \mathcal{D}, \otimes_2, \mathds{1}' \rangle$ is a functor
+  $\mathcal{F} : \mathcal{C} \to \mathcal{D}$ with additional natural transformations, which satisfy the well-known conditions described in \cite{ElKelly}:
 
@@
 
-  2) $\ast_{A, B} : \mathcal{F}A \otimes_2 \mathcal{F}B \to \mathcal{F}(A \otimes_1 B)$
+  2) $\ast_{A, B} : \mathcal{F}A \otimes_2 \mathcal{F}B \to \mathcal{F}(A \otimes_1 B)$.
 
@@
   An applicative functor is a triple $\langle \mathcal{C}, \mathcal{K}, \eta \rangle$,
-where $\mathcal{C}$ is a symmetric monoidal category, $\mathcal{K}$ is a lax monoidal endofunctor and $\eta$ is a natural transformation, such that:
+where $\mathcal{C}$ is a symmetric monoidal category, $\mathcal{K}$ is a monoidal and $\eta : Id_{\mathcal{C}} \Rightarrow \mathcal{K}$ is a natural transformation (similar to unit in monad), such that:
 
@@
 
-2) $\ast_{A,B} \circ (\eta_A \otimes \eta_B) = \eta_{A \otimes B}$.
+2) $\ast_{A,B} \circ (\eta_A \otimes \eta_B) = \eta_{A \otimes B}$;
+
+3) Weak commutativity condition:
+
+\xymatrix{
+&& A \otimes \mathcal{K}B \ar[d]_{\sigma_{A, \mathcal{K}B}} \ar[rr]^{\eta_{A} \otimes id_{\mathcal{K}B}} && \mathcal{K}A \otimes \mathcal{K}B \ar[rr]^{\ast_{A,B}} && \mathcal{K}(A \otimes B) \ar[d]^{\mathcal{K}(\sigma_{A,B})}\\
+&& \mathcal{K}B \otimes A \ar[rr]_{id_{\mathcal{K}B} \otimes \eta_{A}} && \mathcal{K}B \otimes \mathcal{K}A \ar[rr]_{\ast_{B,A}} && \mathcal{K}(B \otimes A)
+}
 \end{defin}
@@
 
-\subsection{Soundness}
+\subsection{Soundness and completeness}
 
@@
 
-\subsection{Completeness}
+\begin{theorem} Completeness
+
+Let $[\![\Gamma \vdash M : A]\!] = [\![\Gamma \vdash N : A]\!]$, then $M =_{\beta \eta} N$.
+\end{theorem}
+
+\begin{proof}
+
+$ $
 
@@
 
-\begin{theorem} Completeness
-  $ $
-
-  i) {\bf K} is an applicative functor;
-
-  ii) Let $[\![\Gamma \vdash M : A]\!] = [\![\Gamma \vdash N : A]\!]$, then $M =_{\beta \eta} N$.
-\end{theorem}
+\begin{lemma}
+
+  {\bf K} is an applicative functor
+\end{lemma}
 
@@
   Immediately follows from previous lemmas in the section.
+\end{proof}
+
 \end{proof}

### 2018-01-24 article.tex
fixes and erases

@@
 which is Curry-Howard isomorphic to IEL$^{-}$ as formal theory of calculations with applicative functors in
-functional programming languages like Haskell or Idris. We prove that this typed lambda-calculus has the
-strong normalization and Church-Rosser properties.
+functional programming languages like Haskell or Idris.
 \end{abstract}
@@
   Haskell see \cite{Morten}.
+
+  The motivation for using an applicative functor is quite natural. Applicative functor allows to generalize the action of a functor for functions with arbitrary number of arguments, for instance:
+  \begin{lstlisting}[language=Haskell]
+  liftA2 :: Applicative f => (a -> b -> c) -> f a -> f b -> f c
+  liftA2 f x y = pure f <*> x <*> y
+  \end{lstlisting}
 
@@
       \RightLabel{$\times_i$}
-      \BinaryInfC{$\Gamma \vdash \langle x, y \rangle : A \times B$}
+      \BinaryInfC{$\Gamma \vdash \langle M, N \rangle : A \times B$}
       \end{prooftree}
@@
 \item $(\lambda x. {\bf pure \: } x) : A \to \textbf{K}A$;
-\item $\lambda f. \lambda x. {\bf let \: pure} \: \langle g, y \rangle = \langle f, x \rangle {\: \bf in \:} g y : {\bf K}(A \to B) \to {\bf K} A \to {\bf K} B$
-\item $\lambda f. \lambda x. {\bf let \:pure} \: \langle g, y \rangle = \langle {\bf pure \:} f, x \rangle {\: \bf in \: } g y : (A \to B) \to {\bf K} A \to {\bf K} B$
+\item $\lambda f. \lambda x. {\bf let \: pure} \: g, y  = f, x {\: \bf in \:} g y : {\bf K}(A \to B) \to {\bf K} A \to {\bf K} B$
+\item $\lambda f. \lambda x. {\bf let \:pure} \: g, y = {\bf pure \:} f, x  {\: \bf in \: } g y : (A \to B) \to {\bf K} A \to {\bf K} B$
 \end{itemize}
@@
 
-  6) $\lambda x. f x \rightarrow_{\eta} f$;
+  6) ${\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} N \rightarrow_{\beta} {\bf pure \:} N$;
 
-  7) $\langle \pi_1 P, \pi_2 P \rangle \rightarrow_{\eta} P$;
+  7) $\lambda x. f x \rightarrow_{\eta} f$;
 
-  8) ${\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} N \rightarrow_{\eta} {\bf pure \:} N$;
+  8) $\langle \pi_1 P, \pi_2 P \rangle \rightarrow_{\eta} P$;
 
@@
   example see \cite{Neder}\cite{Morten}\cite{Girard}. We will consider only modal cases}:
-
-\begin{lemma} Generation lemma.
-
-  i) Let $\Gamma \vdash {\bf pure \:} M : {\bf K}A$, then $\Gamma \vdash M : A$;
-
-  ii) Let $\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N : {\bf K}B$, there are some $A_1,\dots,A_n \in \mathbb{T}_{{\bf K}}$,
-  such that $\Gamma \vdash \vec{M} : {\bf K} \vec{A}$ and $\vec{x} : \vec{A} \vdash N : B$.
-\end{lemma}
-
-\begin{proof}
-  $ $
-
-  Induction on $\Gamma \vdash {\bf pure \:} M : {\bf K}A$ and $\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{N} {\: \bf in \:} N : {\bf K}B$ correspondently.
-\end{proof}
-
-\begin{lemma} Weakening.
-
-  Let $\Gamma \vdash M : A$ and $\Gamma \subseteq \Delta$, then $\Delta \vdash M : A$.
-
-\end{lemma}
-
-\begin{proof}
-  $ $
-
-  1) Let $\Gamma, x : A \vdash x : A$ and $\Gamma \subseteq \Delta$, then $\Delta, x : A \vdash x : A$ trivially.
-
-  2) Let $\Gamma \vdash {\bf pure \:} M : {\bf K}A$. Then $\Gamma \vdash M : A$ by generation and $\Delta \vdash M : A$ by assumption. So $\Delta \vdash {\bf pure \:} M : {\bf K} A$ by ${\bf K}_I$.
-
-  3) Let $\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N : {\bf K}B$ and $\Gamma \subseteq \Delta$.
-Then $\Gamma \vdash \vec{M} : {\bf K}\vec{A}$ and $\vec{x} : \vec{A} \vdash N : B$.
-
-  By assumption $\Delta \vdash \vec{M} : {\bf K}\vec{A}$. So $\Delta \vdash {\bf let \: pure \:} \vec{x} = \vec{N} {\: \bf in \:} N : {\bf K}B$ by $\text{let}_{{\bf K}}$.
-\end{proof}
 
@@
 
-\begin{lemma} Substituition lemma.
+\begin{lemma} Basic lemmas for ${\bf K}_I$.
 
-    i) Let $\Gamma, x : A \vdash M : B$ and $\Gamma \vdash N : A$, then $\Gamma \vdash M [x := N] : B$.
+  i) Let $\Gamma \vdash {\bf pure \:} M : {\bf K}A$, then $\Gamma \vdash M : A$;
 
-    ii) Let $\Gamma \vdash M : A$, then $\Gamma [B := C] \vdash M : (A [B := C])$.
+  ii) Let $\Gamma \vdash M : A$ and $\Gamma \subseteq \Delta$, then $\Delta \vdash M : A$;
 
+  iii) Let $\Gamma, x : A \vdash {\bf pure \:} M : {\bf K}B$ and $\Gamma \vdash N : A$, then $\Gamma \vdash {\bf pure \:} M [x := N] : {\bf K}B$.
+
+  iv) Let $\Gamma \vdash {\bf pure \:} M : {\bf K}A$, then $\Gamma [B := C] \vdash {\bf pure \:} M : {\bf K} (A [B := C])$.
 \end{lemma}
@@
 \begin{proof}
-
   $ $
 
-i) For term substitution:
+  i) Induction on $\Gamma \vdash {\bf pure \:} M : {\bf K}A$ and $\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{N} {\: \bf in \:} N : {\bf K}B$ correspondently.
 
-\vspace{\baselineskip}
+  ii) Let $\Gamma \vdash {\bf pure \:} M : {\bf K}A$. Then $\Gamma \vdash M : A$ by generation and $\Delta \vdash M : A$ by assumption. So $\Delta \vdash {\bf pure \:} M : {\bf K} A$ by ${\bf K}_I$.
 
-1) Let $\Gamma, x : A \vdash x : A$ and $\Gamma \vdash N : A$, but $x [x := N] = N$, so $\Gamma \vdash N : A$.
-
-2) Let $\Gamma, x : A \vdash {\bf pure \:} M : {\bf K}B$ and $\Gamma \vdash N : A$.
+  iii) Let $\Gamma, x : A \vdash {\bf pure \:} M : {\bf K}B$ and $\Gamma \vdash N : A$.
 
@@
 
-3) Let $\Gamma, y : A \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N : {\bf K} B$ and $\Gamma \vdash N : A$.
-
-By generation, $\Gamma, y : A \vdash \vec{M} : {\bf K} \vec{A}$ and $\vec{x} : \vec{A} \vdash N : B$.
-
-By hypothesis, $\Gamma \vdash \vec{M}[x := N] : {\bf K} \vec{A}$.
-
-Hence $\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M}[x := N] {\: \bf in \:} N : {\bf K} B$.
-
-\vspace{\baselineskip}
-
-ii) For type substitution
-
-\vspace{\baselineskip}
-
-1) Let $\Gamma, x : A \vdash x : A$, so $\Gamma [A := C], x : (A[A := C]) \vdash x : (A[A := C])$, or
-
-$\Gamma [A := C], x : C \vdash x : C$.
-
-2) Let $\Gamma \vdash {\bf pure \:} M : {\bf K}A$. By generation $\Gamma \vdash M : A$ and by assumption $\Gamma [B := C] \vdash M : A[B := C]$.
+  iv) Let $\Gamma \vdash {\bf pure \:} M : {\bf K}A$. By generation $\Gamma \vdash M : A$ and by assumption $\Gamma [B := C] \vdash M : A[B := C]$.
 
   By $K_I$ $\Gamma \vdash {\bf pure \:L} M : {\bf K}(A[B := C])$.
-
-3) $\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N : {\bf K}B$. By generation $\Gamma \vdash \vec{M} : {\bf K}\vec{A}$ and $\vec{x} : \vec{A} \vdash N : B$.
-
-By assumption $\Gamma [B_1 := C] \vdash \vec{M} : {\bf }{K}\vec{A}[B_1 := C]$ and $\vec{x} : \vec{A}[B_1 := C] \vdash N : B[B_1 := C]$.
-
-So by $\text{let}_{{\bf K}}$, $\Gamma [B_1 := C] \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N : {\bf K}(B[B_1 := C])$.
-
 \end{proof}
@@
 
-3) Let $\Gamma \vdash {\bf let \: pure \:} \vec{x} = {\bf pure \:} \vec{M} {\: \bf in \:} N : {\bf K}B$.
-
-By generation $\Gamma \vdash {\bf pure \:} \vec{M} : {\bf K}\vec{A}$ and $\vec{x} : \vec{A} \vdash N : B$.
-\\ Moreover, $\Gamma \vdash \vec{M} : \vec{A}$. By weakening and substitution lemma $\Gamma \vdash N [\vec{x} = \vec{M}] : B$.
-\\ By ${\bf K}_I$, $\Gamma \vdash {\bf pure \:} N [\vec{x} := \vec{M}] : {\bf K}B$.
-
-4) Let $\vdash {\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} N : {\bf K}A$
-
-By generation $\vdash N : A$. \\ So $\vdash {\bf pure \:} N : {\bf K}A$ by ${\bf K}_I$.
-
-5) Let $\Gamma \vdash {\bf pure \:} M : A$ and $M \twoheadrightarrow_{\beta \eta} N$.
-
-By generation $\Gamma \vdash M : A$ and $\Gamma \vdash N : A$ by assumption.
-\\ So $\Gamma \vdash {\bf pure \:} N : {\bf K}A$.
 \end{proof}
@@
 
-We will consider term model for simply typed lambda calculus $\times$ and $\to$ standardly described in \cite{LambekScott} \cite{Abramsky}.
+We will consider term model for simply typed lambda calculus $\times$ and $\to$ standardly described in \cite{LambekScott}.
 
@@
 
-  i) ${\bf K}(g \circ f) = {\bf K}(g) \circ {\bf K}(f)$;
+  i) $\text{fmap }(g \circ f) = \text{fmap }(g) \circ \text{fmap }(f)$;
 
-  ii) ${\bf K}(id_A) = id_{{\bf K}A}$.
+  ii) $\text{fmap }(id_A) = id_{{\bf K}A}$.
 \end{lemma}
@@
 
-  \bibitem{LambekScott}
+  \bibitem{LambekScott} Lambek, J. and Scott P.J. (1986) Introduction to Higher Order Categorical Logic, Cambridge Studies in Advanced Mathematics 7, Cambridge: Cambridge University Press.
 
-  \bibitem{Abramsky}
-
-  \bibitem{Gabbay}
+  \bibitem{ElKelly} Samuel Eilenberg and Max Kelly, Closed categories. Proc. Conf. Categorical Algebra (La Jolla, Calif., 1965).
 

### 2018-01-24 article.tex
added some stuff

@@
 \newtheorem{theorem}{Theorem}
+\newtheorem{prop}{Proposition}
 \newtheorem{lemma}{Lemma}
@@
   Functional programming languages such as Haskell \cite{Haskell}, Idris \cite{Idris}, Purescript \cite{Purs}
-  or Elm \cite{Elm} have special type classes\footnote{Type class in Haskell is a general interface for
+  Elm \cite{Elm} or Scala \cite{Scala} have special type classes\footnote{Type class in Haskell is a general interface for
   special group of datatypes.} for calculations with container types like \verb"Functor" and
@@
 
-Strong normalization and colfuence for ${\bf IK}$ was proved by Kakutani for call-by-value and for call-by name \cite{ModalK} \cite{ModalK1}.
+\begin{theorem}
+  $ $
+
+  i) $\twoheadrightarrow_{\beta}$ is strongly normalizing;
+
+  ii) $\twoheadrightarrow_{\beta}$ is confluent.
+\end{theorem}
+
+\begin{proof}
+  $ $
+
+i) Strong normalization for ${\bf IK}$ was proved by Kakutani for call-by-value and for call-by name \cite{ModalK} \cite{ModalK1}.
+
+It is easy to extend Kakutani's result for ${\bf pure}$-rules.
+
+ii) Confluence can be proved by extending Barendregt's technique with term underlying for untyped or simply typed lambda calculus \cite{Baren}.
+
+Hence, it is sufficient to prove strip lemma.
+\end{proof}
 
@@
 
-\begin{defin} Lax monoidal functor
+\begin{defin} Monoidal functor
 

### 2018-01-25 article.tex
added

@@
 
-  It's not difficult to see that modal axioms in $IEL^{-}$ and types of the methods of Applicative class in
-  Haskell-like languages (which is described below) are syntactically similar and we are going to show that
-  this coincidence has a
-  non-trivial computational meaning.
-
   Functional programming languages such as Haskell \cite{Haskell}, Idris \cite{Idris}, Purescript \cite{Purs}
@@
   \end{lstlisting}
+
+  It's not difficult to see that modal axioms in $IEL^{-}$ and types of the methods of Applicative class in
+  Haskell-like languages (which is described below) are syntactically similar and we are going to show that
+  this coincidence has a non-trivial computational meaning.
 
@@
       \AxiomC{$\Gamma \vdash \vec{M} : {\bf K} \vec{A}$}
-      \AxiomC{$\vec{x} : \vec{A} \vdash M : B$}
+      \AxiomC{$\vec{x} : \vec{A} \vdash N : B$}
       \RightLabel{$\text{let}_{{\bf K}}$}
-      \BinaryInfC{$\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } M : {\bf K} B$}
+      \BinaryInfC{$\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } N : {\bf K} B$}
     \end{prooftree}
@@
 
-  7) $({\bf let \:\: pure} \:\: \vec{x} = \vec{M} \:\: {\bf in} \:\: M) [y := P] = {\bf let \:\: pure} \:\: \vec{x} = (\vec{M} [y := P]) \:\: {\bf in} \:\: M$.
+  7) $({\bf let \: pure \:}\vec{x} = \vec{M} {\: \bf in \:} N) [y := P] = {\bf let \: pure \:} \vec{x} = (\vec{M} [y := P]) {\: \bf in \:} M$.
+  \end{defin}
+
+  \begin{defin} Type substituition
+
+    The substituition of type $C$ for type variable $B$ in type $A$ inductively defined as follows:
+
+    1) $B [B := C] = B$ and $D [B := C] = D$, if $B \neq D$;
+
+    2) $(A_1 \alpha A_2)[B := C] = (A_1[B := C]) \alpha (A_2[B := C])$, where $\alpha \in \{ \to, \times \}$;
+
+    3) $({\bf K}A)[B := C] = {\bf K}(A [B := C])$.
+
+    4) Let $\Gamma$ be the context, then $\Gamma [B := C] = \{ x : (A[B := C]) \: | \: x : A \in \Gamma \}$
   \end{defin}
@@
   $\to$-constructor, they are proved standardly in the same lemmas for simply typed lambda calculus, for
-  example see \cite{Neder}\cite{Morten}\cite{Girard}. We will consider only modal cases}:
-
-  \begin{defin} Type substituition
-
-    The substituition of type $C$ for type variable $B$ in type $A$ inductively defined as follows:
-
-    1) $B [B := C] = B$ and $D [B := C] = D$, if $B \neq D$;
-
-    2) $(A_1 \alpha A_2)[B := C] = (A_1[B := C]) \alpha (A_2[B := C])$, where $\alpha \in \{ \to, \times \}$;
-
-    3) $({\bf K}A)[B := C] = {\bf K}(A [B := C])$.
-
-    4) Let $\Gamma$ be the context, then $\Gamma [B := C] = \{ x : (A[B := C]) \: | \: x : A \in \Gamma \}$
-  \end{defin}
+  example see \cite{Neder} \cite{Morten} \cite{Girard}. We will consider only modal cases}:
 
@@
 
-  By $K_I$ $\Gamma \vdash {\bf pure \:L} M : {\bf K}(A[B := C])$.
+  By $K_I$ $\Gamma \vdash {\bf pure \:} M : {\bf K}(A[B := C])$.
 \end{proof}
@@
 
+\begin{theorem}
+  $ $
+
+  Normal form in call-by-name $\lambda_{{\bf K}}$ has the subformula property.
+
+\end{theorem}
+
+\begin{proof}
+  By induction on the structure of term. Case with ${\bf let \: pure\:} \vec{x} = \vec{M} {\: \bf in \:} N$ was considered by Kakutani \cite{ModalK} \cite{ModalK1}.
+  Similary, if ${\bf pure \:} M$ is a normal form, so $M$ is a normal form too by hypothesis.
+\end{proof}
+
 \section{Categorical semantics}
@@
 \end{defin}
-
-By default we will consider an arbitrary closed functor on some cartersian closed category, which is the special case of an applicative functor.
-
-We identify terminal objects. So $\mathcal{K}(\mathds{1}) = \mathds{1}$ and $\eta_{\mathds{1}} = id_{\mathds{1}}$ since $\mathcal{K}$ is an endofunctor.
 
@@
 
-  i) Let $\Gamma \vdash M : A$ and $M \twoheadrightarrow_{\beta} N$, then $[\![\Gamma \vdash M : A]\!] = [\![\Gamma \vdash N : A]\!]$;
-
-  ii) Let $\Gamma \vdash M : A$ and $M \twoheadrightarrow_{\eta} N$, then $[\![\Gamma \vdash M : A]\!] = [\![\Gamma \vdash N : A]\!]$;
+  Let $\Gamma \vdash M : A$ and $M \twoheadrightarrow_{\beta \eta} N$, then $[\![\Gamma \vdash M : A]\!] = [\![\Gamma \vdash N : A]\!]$;
 \end{lemma}

### 2018-01-25 article.tex
added some fixes

@@
 
-  i) For $\beta$-reduction
-
-\vspace{\baselineskip}
-
 Cases with $\beta$-reductions for $let_{\bf K}$ are shown in \cite{ModalK1}. Let us consider cases with ${\bf pure}$.
@@
 
-2)
+2) $[\![\vdash {\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} N : {\bf }K A ]\!] = [\![\vdash {\bf pure \:} N : {\bf K A}]\!]$.
+
+\vspace{\baselineskip}
+
+$\begin{array}{lll}
+&[\![\vdash {\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} N : {\bf }K A ]\!] = & \text{By interpetation}\\
+&\mathcal{K}([\![N]\!]) \circ \eta_{\mathds{1}} =& \text{Naturality for $\eta$} \\
+&\eta_{[\![A]\!]} \circ [\![N]\!] =& \text{By interpretation} \\
+&[\![\vdash {\bf pure \:} N : {\bf K} A]\!]&
+\end{array}$
+
+\vspace{\baselineskip}
+
+3)
 
@@
 Hence $[\![\Gamma \vdash {\bf pure \:} M : {\bf K}A]\!] = [\![\Gamma \vdash {\bf pure \:} N : {\bf K}A]\!]$.
-
-\vspace{\baselineskip}
-
-ii) For $\eta$-reduction.
-
-1) $[\![\vdash {\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} N : {\bf }K A ]\!] = [\![\vdash {\bf pure \:} N : {\bf K A}]\!]$.
-
-\vspace{\baselineskip}
-
-$\begin{array}{lll}
-&[\![\vdash {\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} N : {\bf }K A ]\!] = & \text{By interpetation}\\
-&\mathcal{K}([\![N]\!]) \circ \eta_{\mathds{1}} =& \text{Naturality for $\eta$} \\
-&\eta_{[\![A]\!]} \circ [\![N]\!] =& \text{By interpretation} \\
-&[\![\vdash {\bf pure \:} N : {\bf K} A]\!]&
-\end{array}$
 
@@
 \begin{defin}
-  Let us define an endofunctor $\mathcal{K} : \mathcal{C}(\lambda) \to \mathcal{C}(\lambda)$, such that:
-
-1) ${\bf K} : A \mapsto {\bf K}A$;
-
-2) ${\bf K} : [x,M] \in Hom_{\mathcal{C}(\lambda)}(A,B) \mapsto \text{fmap } f = [y, {\bf let \: pure \:} x = y {\: \bf in \:} M] \in Hom_{\mathcal{C}(\lambda)}({\bf K} A, {\bf K} B)$.
+  Let us define an endofunctor $\mathcal{K} : \mathcal{C}(\lambda) \to \mathcal{C}(\lambda)$, such that
+forall $[x,M] \in Hom_{\mathcal{C}(\lambda)}(A,B), {\bf K}([x,M]) = [y, {\bf let \: pure \:} x = y {\: \bf in \:} M] \in Hom_{\mathcal{C}(\lambda)}({\bf K} A, {\bf K} B)$ (denotation: $\text{fmap } f$ for an arbitrary arrow $f$).
 
@@
 
-  Let $[p, \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle] \in Hom_{\mathcal{C}(\lambda)}(A \times {\bf K}B, {\bf K}A \times {\bf K}B)$.
+  Let $[p, \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle] \in Hom_{\mathcal{C}(\lambda)}(A \times {\bf K}B, {\bf K}(A \times B))$.
 

### 2018-01-28 article.tex
added proof
added confluence proof

@@
 
-  6) ${\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} N \rightarrow_{\beta} {\bf pure \:} N$;
-
-  7) $\lambda x. f x \rightarrow_{\eta} f$;
-
-  8) $\langle \pi_1 P, \pi_2 P \rangle \rightarrow_{\eta} P$;
-
-  9) ${\bf let \: pure \:} x = M {\: \bf in \: } x \rightarrow_{\eta} M$;
-
-  10) $M \rightarrow_{\beta \eta} N \Rightarrow {\bf pure \: M} \rightarrow_{\beta \eta} {\bf pure \: N}$
-  \end{defin}
+  6) $\lambda x. f x \rightarrow_{\eta} f$;
+
+  7) $\langle \pi_1 P, \pi_2 P \rangle \rightarrow_{\eta} P$;
+
+  8) ${\bf let \: pure \:} x = M {\: \bf in \: } x \rightarrow_{\eta} M$;
+  \end{defin}
+
+By default we use call-by-name evaluation strategy.
 
@@
 
-\begin{lemma} Basic lemmas for ${\bf K}_I$.
-
-  i) Let $\Gamma \vdash {\bf pure \:} M : {\bf K}A$, then $\Gamma \vdash M : A$;
-
-  ii) Let $\Gamma \vdash M : A$ and $\Gamma \subseteq \Delta$, then $\Delta \vdash M : A$;
-
-  iii) Let $\Gamma, x : A \vdash {\bf pure \:} M : {\bf K}B$ and $\Gamma \vdash N : A$, then $\Gamma \vdash {\bf pure \:} M [x := N] : {\bf K}B$.
-
-  iv) Let $\Gamma \vdash {\bf pure \:} M : {\bf K}A$, then $\Gamma [B := C] \vdash {\bf pure \:} M : {\bf K} (A [B := C])$.
-\end{lemma}
-
-\begin{proof}
-  $ $
-
-  i) Induction on $\Gamma \vdash {\bf pure \:} M : {\bf K}A$ and $\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{N} {\: \bf in \:} N : {\bf K}B$ correspondently.
-
-  ii) Let $\Gamma \vdash {\bf pure \:} M : {\bf K}A$. Then $\Gamma \vdash M : A$ by generation and $\Delta \vdash M : A$ by assumption. So $\Delta \vdash {\bf pure \:} M : {\bf K} A$ by ${\bf K}_I$.
-
-  iii) Let $\Gamma, x : A \vdash {\bf pure \:} M : {\bf K}B$ and $\Gamma \vdash N : A$.
-
-  By generation $\Gamma, x : A \vdash M : B$ and by assumption $\Gamma \vdash M [x := N] : B$.
-
-  By $K_I$, $\Gamma \vdash {\bf pure \:} (M [x := N]) : {\bf K}B$.
-
-  iv) Let $\Gamma \vdash {\bf pure \:} M : {\bf K}A$. By generation $\Gamma \vdash M : A$ and by assumption $\Gamma [B := C] \vdash M : A[B := C]$.
-
-  By $K_I$ $\Gamma \vdash {\bf pure \:} M : {\bf K}(A[B := C])$.
+\begin{lemma} Generation for ${\bf K}_I$.
+
+  Let $\Gamma \vdash {\bf pure \:} M : {\bf K}A$, then $\Gamma \vdash M : A$;
+\end{lemma}
+
+\begin{proof}
+  Induction on the structure of ${\bf pure \:} M$.
+\end{proof}
+
+\begin{lemma} Basic lemmas .
+
+  i) Let $\Gamma \vdash M : A$ and $\Gamma \subseteq \Delta$, then $\Delta \vdash M : A$;
+
+  ii) Let $\Gamma, x : A \vdash M : B$ and $\Gamma \vdash N : A$, then $\Gamma \vdash M [x := N] : B$.
+
+  iii) Let $\Gamma \vdash M : A$, then $\Gamma [B := C] \vdash M : (A [B := C])$.
+\end{lemma}
+
+\begin{proof}
+  $ $
+
+  i-ii-iii) Induction on $\Gamma \vdash M : A$.
+
 \end{proof}
@@
 
-  i) $\twoheadrightarrow_{\beta}$ is strongly normalizing;
-
-  ii) $\twoheadrightarrow_{\beta}$ is confluent.
+  $\twoheadrightarrow_{\beta}$ is strongly normalizing;
 \end{theorem}
@@
 
-It is easy to extend Kakutani's result for ${\bf pure}$-rules.
-
-ii) Confluence can be proved by extending Barendregt's technique with term underlying for untyped or simply typed lambda calculus \cite{Baren}.
-
-Hence, it is sufficient to prove strip lemma.
 \end{proof}
@@
 
-  Normal form in call-by-name $\lambda_{{\bf K}}$ has the subformula property.
+  $\twoheadrightarrow_{\beta}$ is confluent.
+\end{theorem}
+
+\begin{proof}
+
+  We modify and apply Barendregt's technique with term underlying. We will consider the fragment of the grammar for terms without
+  constructors for pairs for simplicity.
+
+\begin{defin} The set of underlined terms.
+
+  \begin{itemize}
+    \item $x \in \mathbb{V} \Rightarrow x \in \underline{\Lambda}$;
+    \item $M \in \underline{\Lambda} \Rightarrow (\lambda x. M) \in \underline{\Lambda}$;
+    \item $M, N \in \underline{\Lambda} \Rightarrow (M N) \in \underline{\Lambda}$;
+    \item $M \in \underline{\Lambda} \Rightarrow ({\bf pure \:} M) \in \underline{\Lambda}$;
+    \item $\vec{x} \in \mathbb{V}, \vec{M}, N \in \underline{\Lambda} \Rightarrow {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N \in \underline{\Lambda}$;
+    \item $M, N \in \underline{\Lambda} \Rightarrow (\lambda_i x. M) N \in \underline{\Lambda}$, for all $i \in \mathbb{N}$.
+  \end{itemize}
+\end{defin}
+
+\begin{defin} Subsitution for term with labelled lambda:
+
+  $((\lambda_i x. M) N) [y := Z] = (\lambda_i x. M [y := Z]) (N [y := Z])$
+\end{defin}
+
+\begin{defin} Index erasing
+
+  Let us define map $|.| : \underline{\Lambda} \to \Lambda$ as follows:
+
+  \begin{itemize}
+    \item $|x| = x$;
+    \item $|\lambda x. M| = \lambda x. |M|$;
+    \item $|M N| = |M| |N|$;
+    \item $|{\bf pure \:} M| = {\bf pure \:} |M|$;
+    \item $|{\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N| = {\bf let \: pure \:} \vec{x} = \vec{|M|} {\: \bf in \:} |N|$;
+    \item $|(\lambda_i x. M) N| = (\lambda x. M) N$
+  \end{itemize}
+\end{defin}
+
+\begin{defin} Reduction rules:
+
+\begin{itemize}
+  \item $(\lambda x. M) N \rightarrow_{\underline{\beta}} M [x := N]$;
+  \item $\begin{array}{llll}
+  & {\bf let \: pure \:} \langle \vec{x}, y, \vec{z} \rangle = \langle \vec{M}, {\bf let \: pure \: } \vec{w} = \vec{N} {\bf \: in \: } Q, \vec{P} \rangle {\: in \:} R \rightarrow_{\underline{\beta}} & \\
+  & {\bf let \: pure \:} \langle \vec{x}, \vec{w}, \vec{z} \rangle = \langle \vec{M}, \vec{N}, \vec{P} \rangle \: { \bf in } \: R [y := Q]
+  \end{array}$;
+  \item ${\bf let \: pure \:} \vec{x} = {\bf pure \:} \vec{M} {\: \bf in \:} N \rightarrow_{\underline{\beta}} {\bf pure \:} N [\vec{x} := \vec{M}]$;
+  \item $(\lambda x_i. M) N \rightarrow_{\underline{\beta}} M [x := N]$
+\end{itemize}
+\end{defin}
+
+$\twoheadrightarrow_{\underline{\beta}}$ is a reflexive-transitive closure of $\rightarrow_{\underline{\beta}}$.
+
+\begin{defin} Indexed redex erasing:
+
+  Let us define the next map $\phi : \underline{\Lambda} \to \Lambda$:
+
+  \begin{itemize}
+    \item $\phi(x) = x$;
+    \item $\phi(\lambda x. M) = \lambda x. \phi(M)$;
+    \item $\phi(M N) = \phi(M) \phi(N)$;
+    \item $\phi({\bf pure \:} M) = {\bf pure \:} \phi(M)$;
+    \item $\phi({\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N) = {\bf let \: pure \:} \vec{x} = \vec{\phi(M)} {\: \bf in \:} \phi(N)$;
+    \item $\phi((\lambda_i x. M) N) = M [x := N]$
+  \end{itemize}
+\end{defin}
+
+\begin{lemma}
+
+  $\forall \underline{M}, \underline{N} \in \underline{\Lambda} \: \forall M, N \in \Lambda, \text{if } |\underline{M}| = M, |\underline{N}| = N, \text{then}$
+  \begin{itemize}
+    \item If $M \twoheadrightarrow_{\beta} N$, then $\underline{M} \twoheadrightarrow_{\underline{\beta}} \underline{N}$
+    \item Vice versa
+  \end{itemize}
+\end{lemma}
+
+\begin{proof}
+
+  Induction on the generation $\rightarrow_{\beta}$ and $\rightarrow_{\underline{\beta}}$ correspondently. The general statement follows from transitivity
+  of multi-step reductions of both types.
+\end{proof}
+
+\begin{lemma}
+  $\phi(M [x := N]) = \phi(M) [x := \phi(N)]$.
+\end{lemma}
+
+\begin{proof}
+  We treat only cases with {\bf pure} and with {\bf let}. For the rest cases see \cite{Baren}.
+
+  1)
+  $ $
+
+  $\begin{array}{lll}
+  & \phi({\bf pure \:} (M [x := N])) = & \text{By the definition of $\phi$}\\
+  & {\bf pure \:} (\phi(M [x := N])) = & \text{Induction hypothesis} \\
+  & {\bf pure \:} (\phi(M) [x := \phi(N)]) = & \text{Substitution definition}\\
+  & ({\bf pure \:} \phi (M)) [x := \phi(N)]&
+  \end{array}$
+
+  2)
+  $ $
+
+  $\begin{array}{lll}
+  &\phi(({\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N) [y := P]) = & \text{Substitution definition}\\
+  &\phi({\bf let \: pure \:} \vec{x} = (\vec{M}[y := P]) {\: \bf in \:} N) = & \text{By the definition of $\phi$}\\
+  &{\bf let \: pure \:} \vec{x} = \phi(\vec{M}[y := P]) {\: \bf in \:} \phi(N) = & \text{Induction hypothesis} \\
+  &{\bf let \: pure \:} \vec{x} = (\phi(\vec{M}) [y := \phi(P)]) {\: \bf in \:} \phi(N) =& \text{Substitution definition} \\
+  &({\bf let \: pure \:} \vec{x} = \phi(\vec{M}) {\: \bf in \:} \phi(N))[y := \phi(P)]
+  \end{array}$
+
+\end{proof}
+
+\begin{lemma}
+  $ $
+
+  \begin{itemize}
+  \item If $M \twoheadrightarrow_{\underline{\beta}} N$, then $\phi(M) \twoheadrightarrow_{\beta} \phi(N)$
+  \item If $|M| = N$ and $\phi(M) = P$, then $N \twoheadrightarrow_{\beta} P$.
+  \end{itemize}
+\end{lemma}
+
+\begin{proof}
+  $ $
+
+  i) Induction on the generation of $\twoheadrightarrow_{\underline{\beta}}$ using previous lemma.
+
+  ii) Induction on the structure of $M$.
+\end{proof}
+
+\begin{lemma} Strip lemma.
+
+$ $
+
+If $M \rightarrow_{\beta} N$ and $M \twoheadrightarrow_{\beta} P$. Then there exists some term $Q$, such that
+$N \twoheadrightarrow_{\beta} Q$ and $P \twoheadrightarrow_{\beta} Q$.
+\end{lemma}
+
+\begin{proof}
+
+Proof is similar to \cite{Baren} \cite{Baren2}. We build the following diagram, which commutes by lemmas 5 and 7.
+
+\vspace{\baselineskip}
+
+\xymatrix{
+&&& M \ar[dd]_{\beta} \ar@{->>}[rrr] &&& P \ar@{-->>}[dd]^{\beta} \\
+&&&& N' \ar[ul]^{|.|} \ar[dl]_{\phi} \ar@{-->>}[rrr]^{\underline{\beta}} &&& P' \ar[ul]^{|.|} \ar[dl]_{\phi} \\
+&&& N \ar@{-->>}[rrr]_{\beta} &&& Q
+}
+\end{proof}
+
+\begin{col}
+$ $
+If $M \twoheadrightarrow_{\beta} N$ and $M \twoheadrightarrow_{\beta} P$. Then there exists some term $Q$, such that
+$N \twoheadrightarrow_{\beta} Q$ and $P \twoheadrightarrow_{\beta} Q$.
+\end{col}
+
+\begin{proof}
+
+  Unfold $M \twoheadrightarrow_{\beta} N$ as the sequence of one-step reduction and apply strip lemma on the every step.
+\end{proof}
+
+\end{proof}
+
+\begin{theorem}
+  $ $
+
+  Normal form in $\lambda_{{\bf K}}$ has the subformula property.
 
@@
 
-2) $[\![\vdash {\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} N : {\bf }K A ]\!] = [\![\vdash {\bf pure \:} N : {\bf K A}]\!]$.
-
-\vspace{\baselineskip}
-
-$\begin{array}{lll}
-&[\![\vdash {\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} N : {\bf }K A ]\!] = & \text{By interpetation}\\
-&\mathcal{K}([\![N]\!]) \circ \eta_{\mathds{1}} =& \text{Naturality for $\eta$} \\
-&\eta_{[\![A]\!]} \circ [\![N]\!] =& \text{By interpretation} \\
-&[\![\vdash {\bf pure \:} N : {\bf K} A]\!]&
-\end{array}$
-
-\vspace{\baselineskip}
-
-3)
-
-If $\Gamma \vdash M : A$ and $M \rightarrow_{\beta \eta} N$, then $[\![\Gamma \vdash {\bf pure \:} M : {\bf K}A]\!] = [\![\Gamma \vdash {\bf pure \:} N : {\bf K}A]\!]$.
-
-\vspace{\baselineskip}
-
-If $\Gamma \vdash M : A$ and $M \rightarrow_{\beta \eta} N$, then $\Gamma \vdash N : A$ by subject reduction.
-
-By assumption $[\![\Gamma \vdash M : A]\!] = [\![\Gamma \vdash N : A]\!]$.
-
-So $\eta_{[\![A]\!]} \circ [\![\Gamma \vdash M : A]\!] = \eta_{[\![A]\!]} \circ [\![\Gamma \vdash N : A]\!]$.
-
-Hence $[\![\Gamma \vdash {\bf pure \:} M : {\bf K}A]\!] = [\![\Gamma \vdash {\bf pure \:} N : {\bf K}A]\!]$.
-
 \end{proof}
@@
 
-Follows from i) and ii).
+\vspace{\baselineskip}
+
+$\begin{array}{lll}
+& \ast_{A,B} \circ (\eta_{A} \times \eta_{B}) = & \\
+& \text{By unfolding} & \\
+& [q, {\bf let \: pure \:} x, y = \pi_1 q, \pi_2 q {\: \bf in \:} \langle x, y \rangle] \circ [p, \langle {\bf pure \:} (\pi_1 p), {\bf pure \:} (\pi_2 p) \rangle] = & \\
+& \text{Composition}& \\
+& [p, {\bf let \: pure \:} x, y = \pi_1 q, \pi_2 q {\: \bf in \:} \langle x, y \rangle [q := \langle {\bf pure \:} (\pi_1 p), {\bf pure \:} (\pi_2 p) \rangle]] = & \\
+& \text{By substitution}& \\
+& [p, {\bf let \: pure \:} x, y = \pi_1 (\langle {\bf pure \:} (\pi_1 p), {\bf pure \:} (\pi_2 p) \rangle), \pi_2 (\langle {\bf pure \:} (\pi_1 p), {\bf pure \:} (\pi_2 p) \rangle) {\: \bf in \:} \langle x, y \rangle] = & \\
+& \text{Reduction rules}& \\
+& [p, {\bf let \: pure \:} x, y = {\bf pure \:} (\pi_1 p), {\bf pure \:} (\pi_2 p) {\: \bf in \:} \langle x, y \rangle] = & \\
+& \text{Reduction rule}& \\
+& [p, {\bf pure \:} (\langle x,y \rangle [x := \pi_1 p, y := \pi_2 p])] = & \\
+& \text{Substitution}& \\
+& [p, {\bf pure \:} \langle \pi_1 p, \pi_2 p \rangle] = & \\
+& \text{$\eta$-reduction}& \\
+& [p, {\bf pure \:} p] =& \\
+& \text{By definition}& \\
+& \eta_{A \times B}&
+\end{array}$
 
@@
 
+\section{Monadic extention}
+
+In modern Haskell \verb"Monad" class is an extention of the \verb"Applicative" class,
+where monadic binding (\verb"(>>=)") is the single basic class method:
+
+\begin{lstlisting}[language=Haskell]
+  class Applicative m => Monad (m :: * -> *) where
+    (>>=) :: m a -> (a -> m b) -> m b
+    (>>) :: m a -> m b -> m b
+    return :: a -> m a
+    fail :: String -> m a
+\end{lstlisting}
+From a logical point of view we may consider this extention as the extention of IEL$^{-}$ with the imdepotency axiom.
+
+\begin{defin} ``Monadic'' intuitionistic epistemic logic IEL$^{-}$
+
+  1) IPC axioms;
+
+  2) ${\bf K}(A \to B) \to ({\bf K}A \to {\bf K}B)$ (normality);
+
+  3) $A \to {\bf K}A$ (co-reflection);
+
+  4) ${\bf KK} A \to {\bf K}A$
+
+  Rule: MP.
+\end{defin}
+
+Categorically \verb"Monad" class is a Kleisli triple on monoidal category, or, which is the same, strong monoidal monad:
+
+\begin{defin} Strong monoidal monad
+
+\end{defin}
+
+\begin{defin} Monadic $\lambda_{{\bf K}}$
+
+  Monadic $\lambda_{{\bf K}}$ is a $\lambda_{{\bf K}}$ with the next additional typing rule:
+
+  \begin{prooftree}
+    \AxiomC{$\Gamma \vdash M : {\bf KK}A$}
+    \RightLabel{${\bf K}^2$--e}
+    \UnaryInfC{$\Gamma \vdash {\bf join \:} M : {\bf K}A$}
+  \end{prooftree}
+\end{defin}
+
+\begin{defin}
+\end{defin}
+
   \addcontentsline{toc}{section}{References}
 
-  %далее сам список используевой литературы
   \begin{thebibliography}{}

### 2018-01-29 article.tex
added proofs

@@
 
+3) If the derivation ends in
+
+\begin{prooftree}
+\AxiomC{$\Gamma \vdash {\bf pure \:} \vec{M} : {\bf K}\vec{A}$}
+\AxiomC{$\vec{x} : \vec{A} \vdash N : B$}
+\BinaryInfC{$\Gamma \vdash {\bf let \: pure \:} \vec{x} = {\bf pure \:} \vec{M} {\: \bf in \:} N : {\bf K}B$}
+\end{prooftree}
+
+So $\Gamma \vdash \vec{M} : \vec{A}$ by generation and $\Gamma \vdash N [\vec{x} := \vec{M}] : B$ by weakening and substitution.
+
+Then we can transform this into the next derivation:
+
+\begin{prooftree}
+\AxiomC{$\Gamma \vdash N [\vec{x} := \vec{M}] : B$}
+\RightLabel{${\bf K}_I$}
+\UnaryInfC{$\Gamma \vdash {\bf pure \:} N [\vec{x} := \vec{M}] : {\bf K}B$}
+\end{prooftree}
+
 \end{proof}
@@
 
-i) Strong normalization for ${\bf IK}$ was proved by Kakutani for call-by-value and for call-by name \cite{ModalK} \cite{ModalK1}.
+We modify and apply Tait's technique of logical relation for modalities.
+For strong normalization proof with Tait's method for simply typed lambda calculus see \cite{Pierce}.
+
+\begin{defin} The set of strongly computable terms:
+  \begin{itemize}
+    \item $SC_A = \{ M : A \: | \: M \text{ is strongly normalizing} \}$ for $A \in \mathbb{T}$;
+    \item $SC_{A \to N} = \{ M : A \to B \: | \: \forall N \in SC_A, M N \in SC_B \}$, for $A,B \in \mathbb{T}_{\bf K}$
+    \item $SC_{{\bf K}A} = \{ M : {\bf K}A \: | \: M \text{ is strongly normalizing} \}$ for $A \in \mathbb{T}$;
+    \item $SC_{{\bf K}(A \to B)} = \{ M : {\bf K}(A \to B) \: | \: \forall f \in SC_{A \to B}, \forall x \in SC_A, \forall N \in SC_{{\bf K}A}, {\bf let \: pure \:} f, x = M, N {\: \bf in \:} f x \}$ for $A,B \in \mathbb{T}_{\bf K}$.
+  \end{itemize}
+\end{defin}
+
+\begin{lemma}
+  $ $
+
+\begin{itemize}
+\item If $M \in SC_A$, then $M$ is strongly normalizing;
+\item Let $M \in SC_A$ and $M \rightarrow_{\beta} N$, then $N \in SC_A$;
+\item Let $N$ is non-introduced, $N \in SC_A$. Then, if $M \rightarrow_{\beta} N$, then $M \in SC_A$;
+\end{itemize}
+\end{lemma}
+
+\begin{proof}
+  By induction on the structure of $A$.
+
+  1) $A \equiv {\bf K}A$, where $A \in \mathbb{T}$.
+
+  i) Follows from the definition;
+
+  ii) Immediately;
+
+  iii) Let $N$ is non-introduced and $N \in SC_A$, such that $M \rightarrow_{\beta} N$.
+  Any reduction path $M \rightarrow_{\beta} \dots$ passes through $M \rightarrow_{\beta} N$.
+
+  $N$ is strongly normalizing, so $M$ too.
+
+    \vspace{\baselineskip}
+
+  2) $A \equiv {\bf K}(B \to C)$
+
+  i) Suppose $M \in SC_{{\bf K}(B \to C)}$. Let $N \in SC_{{\bf K}B}$. So ${\bf let \: pure \:} f, x = M, N {\: \bf in \: } f x \in SC_{{\bf K}C}$.
+
+  So $M$ is strongly normalizing, since ${\bf let \: pure \:} f, x = M, N {\: \bf in \: } f x$ is strongly normalizing.
+
+  ii) Let $M_1 \in SC_{{\bf K}(B \to C)}$ and $M_1 \rightarrow_{\beta} M_2$. Fix $N \in SC_{{\bf K}B}$.
+
+  Then ${\bf let \: pure \:} f, x = M_1, N {\: \bf in \: } f x \in SC_{{\bf K}C}$.
+
+  Hence, ${\bf let \: pure \:} f, x = M_1, N {\: \bf in \: } f x \in SC_{{\bf K}C} \rightarrow_{\beta} {\bf let \: pure \:} f, x = M_2, N {\: \bf in \: } f x$.
+
+  So ${\bf let \: pure \:} f, x = M_2, N {\: \bf in \: } f x \in SC_{{\bf K}C}$. Then $M_2 \in SC_{{\bf K}(B \to C)}$.
+
+  iii) Let $M_2$ be non-introduced, $M_2 \in SC_{{\bf K}(B \to C)}$ and $M_1 \rightarrow_{\beta} M_2$.
+
+  Let $N \in SC_{{\bf K}B}$.
+\end{proof}
 
@@
 
-  Unfold $M \twoheadrightarrow_{\beta} N$ as the sequence of one-step reduction and apply strip lemma on the every step.
+  Unfold $M \twoheadrightarrow_{\beta} N$ as the sequence of one-step reductions and apply strip lemma on the every step.
 \end{proof}
@@
 
-\section{Monadic extention}
-
-In modern Haskell \verb"Monad" class is an extention of the \verb"Applicative" class,
-where monadic binding (\verb"(>>=)") is the single basic class method:
-
-\begin{lstlisting}[language=Haskell]
-  class Applicative m => Monad (m :: * -> *) where
-    (>>=) :: m a -> (a -> m b) -> m b
-    (>>) :: m a -> m b -> m b
-    return :: a -> m a
-    fail :: String -> m a
-\end{lstlisting}
-From a logical point of view we may consider this extention as the extention of IEL$^{-}$ with the imdepotency axiom.
-
-\begin{defin} ``Monadic'' intuitionistic epistemic logic IEL$^{-}$
-
-  1) IPC axioms;
-
-  2) ${\bf K}(A \to B) \to ({\bf K}A \to {\bf K}B)$ (normality);
-
-  3) $A \to {\bf K}A$ (co-reflection);
-
-  4) ${\bf KK} A \to {\bf K}A$
-
-  Rule: MP.
-\end{defin}
-
-Categorically \verb"Monad" class is a Kleisli triple on monoidal category, or, which is the same, strong monoidal monad:
-
-\begin{defin} Strong monoidal monad
-
-\end{defin}
-
-\begin{defin} Monadic $\lambda_{{\bf K}}$
-
-  Monadic $\lambda_{{\bf K}}$ is a $\lambda_{{\bf K}}$ with the next additional typing rule:
-
-  \begin{prooftree}
-    \AxiomC{$\Gamma \vdash M : {\bf KK}A$}
-    \RightLabel{${\bf K}^2$--e}
-    \UnaryInfC{$\Gamma \vdash {\bf join \:} M : {\bf K}A$}
-  \end{prooftree}
-\end{defin}
-
-\begin{defin}
-\end{defin}
-
   \addcontentsline{toc}{section}{References}

### 2018-01-29 article.tex
added more proofs
added some features

@@
 1)
-  Let $\Gamma \vdash {\bf let \: pure \:} \langle \vec{x}, y, \vec{z} \rangle = \langle \vec{M}, {\bf let \: pure \: } \vec{w} = \vec{N} {\bf \: in \: } Q, \vec{P} \rangle {\: in \:} R : {\bf K}B$,
-  then $\Gamma {\bf let \: pure \:} \langle \vec{x}, \vec{w}, \vec{z} \rangle = \langle \vec{M}, \vec{N}, \vec{P} \rangle \: { \bf in } \: R [y := Q] : {\bf K}B$
+  Let $\Gamma \vdash {\bf let \: pure \:} \vec{x}, y, \vec{z}  = \langle \vec{M}, {\bf let \: pure \: } \vec{w} = \vec{N} {\bf \: in \: } Q, \vec{P} \rangle {\: in \:} R : {\bf K}B$,
+  then $\Gamma \vdash {\bf let \: pure \:} \vec{x}, \vec{w}, \vec{z} = \langle \vec{M}, \vec{N}, \vec{P} \rangle \: { \bf in } \: R [y := Q] : {\bf K}B$
 
@@
     \item $SC_A = \{ M : A \: | \: M \text{ is strongly normalizing} \}$ for $A \in \mathbb{T}$;
-    \item $SC_{A \to N} = \{ M : A \to B \: | \: \forall N \in SC_A, M N \in SC_B \}$, for $A,B \in \mathbb{T}_{\bf K}$
+    \item $SC_{A \to B} = \{ M : A \to B \: | \: \forall N \in SC_A, M N \in SC_B \}$, for $A,B \in \mathbb{T}_{\bf K}$
     \item $SC_{{\bf K}A} = \{ M : {\bf K}A \: | \: M \text{ is strongly normalizing} \}$ for $A \in \mathbb{T}$;
@@
 \begin{proof}
+  $ $
+
   By induction on the structure of $A$.
@@
 
-  Let $N \in SC_{{\bf K}B}$.
-\end{proof}
+  Let $N \in SC_{{\bf K}B}$. So ${\bf let \: pure \:} f, x = M_2, N {\: \bf in \:} f x \in SC_{{\bf K}C}$.
+
+  So ${\bf let \: pure \:} f, x = M_1, N {\: \bf in \:} f x \rightarrow_{\beta} {\bf let \: pure \:} f, x = M_2, N {\: \bf in \:} f x \in SC_{{\bf K}C}$.
+
+  Thus ${\bf let \: pure \:} f, x = M_1, N {\: \bf in \:} f x \in SC_{{\bf K}C}$ by IH, so $M_1 \in SC_{{\bf K}(B \to C)}$.
+\end{proof}
+
+\begin{lemma}
+  $ $
+
+  Let $x_1 : A_1,\dots, x_n : A_n \vdash M : A$, then forall $i, M_i \in SC_{A_i}$. Then $M [x_1 := M_1,\dots,x_n := M_n] \in SC_A$.
+\end{lemma}
+
+\begin{proof}
+  $ $
+
+  1) Let the derivation ends in:
+
+\begin{prooftree}
+\AxiomC{$x_1 : A_1,\dots, x_n : A_n \vdash M : A$}
+\UnaryInfC{$x_1 : A_1,\dots, x_n : A_n \vdash {\bf pure \:} M : {\bf K}A$}
+\end{prooftree}
+
+  By assumption $M [x_1 := M_1,\dots,x_n := M_n] \in SC_A$, so ${\bf pure \:} M [x_1 := M_1,\dots,x_n := M_n] \in SC_{{\bf K}A}$.
+
+\vspace{\baselineskip}
+
+  2) Let the derivation ends in:
+\begin{prooftree}
+  \AxiomC{$x_1 : A_1,\dots, x_n : A_n \vdash \vec{M^{'}} : {\bf K}\vec{A}$}
+  \AxiomC{$\vec{x} : \vec{A} \vdash N : B$}
+  \BinaryInfC{$x_1 : A_1,\dots, x_n : A_n \vdash {\bf let \: pure \:} \vec{x} = \vec{M^{'}} {\: \bf in \:} N : {\bf K}B$}
+\end{prooftree}
+
+By IH forall $i \in \{ 1,\dots,\text{length}(\vec{M^{'}}) \}$, $M_i^{'} [x_1 := M_1,\dots,x_n := M_n] \in SC_{{\bf K}A_i}$.
+
+So ${\bf let \: pure \:} \vec{x} = \vec{M^{'}} [x_1 := M_1,\dots,x_n := M_n] {\: \bf in \:} N \in SC_{{\bf K} B}$.
+\end{proof}
+
+\begin{col}
+
+  All terms are strongly computable, therefore are strongly normalizing.
+\end{col}
 
@@
   An applicative functor is a triple $\langle \mathcal{C}, \mathcal{K}, \eta \rangle$,
-where $\mathcal{C}$ is a symmetric monoidal category, $\mathcal{K}$ is a monoidal and $\eta : Id_{\mathcal{C}} \Rightarrow \mathcal{K}$ is a natural transformation (similar to unit in monad), such that:
+where $\mathcal{C}$ is a symmetric monoidal category, $\mathcal{K}$ is a monoidal endofunctor and $\eta : Id_{\mathcal{C}} \Rightarrow \mathcal{K}$ is a natural transformation (similar to unit in monad), such that:
 
@@
 
-\begin{defin} Semantical translation from $\lambda_{\textbf{K}}$ to CCC with applicative functor $\mathcal{K}$:
+\begin{defin} Semantical translation from $\lambda_{\textbf{K}}$ to some cartesian closed category $\mathcal{C}$ with applicative functor $\mathcal{K}$:
 
@@
 
-$[\![A]\!] := \hat{A}, A \in \mathbb{T}$;
+$[\![A]\!] := \hat{A}, A \in \mathbb{T}$, where $\hat{A}$ is an object of $\mathcal{C}$ obtained by some given assignment;
 
@@
 
-We will consider term model for simply typed lambda calculus $\times$ and $\to$ standardly described in \cite{LambekScott}.
+We will consider term model for simply typed lambda calculus $\times$ and $\to$ standardly described in \cite{LambekScott}:
+
+\begin{defin} Equivalence on term pairs:
+  $ $
+
+  1) $(x, M) \sim_{A, B} (y, N) \Leftrightarrow x : A \vdash M : B \:\: \& \:\: y : A \vdash N : A \:\: \& \:\: M =_{\beta \eta} N [y := x]$;
+
+  2) $[x, M]_{A, B} = \{ (y, N) \: | \: (x, M) \sim_{A, B} (y, N) \}$.
+\end{defin}
+
+We will drop indeces below.
+
+\begin{defin} Category $\mathcal{C}(\lambda)$:
+\begin{itemize}
+  \item $Ob_{\mathcal{C}} = \{ \hat{A} \: | \: A \in \mathbb{T} \} \cup \{ \mathds{1} \}$;
+  \item $Hom_{\mathcal{C}(\lambda)}(\hat{A},\hat{B}) = \{ [x,M] \: | \: x : A \vdash M : B \}$;
+  \item Let $[x, M] \in Hom_{\mathcal{C}(\lambda)}(\hat{A},\hat{B})$ and $[y,N] \in Hom_{\mathcal{C}(\lambda)}(\hat{B},\hat{C})$, then $[y,M] \circ [x, M] = [x, N [y := M]]$;
+  \item Identity morphism $id_{\hat{A}} = [x,x] \in Hom_{\mathcal{C}(\lambda)(\hat{A})}$;
+  \item $\mathds{1}$ is a terminal object;
+  \item $\widehat{A \times B} = \hat{A} \times \hat{B}$;
+  \item Canonical projection is defined as $[x, \pi_i x] \in Hom_{\mathcal{C}(\lambda)}(\hat{A_1} \times \hat{A_2},\hat{A_i})$ for $i \in \{ 1, 2 \}$;
+  \item $\widehat{A \to B} = \hat{B}^{\hat{A}}$;
+  \item Evaluation arrow $\epsilon = [x, (\pi_2 x) (\pi_1 x)] \in Hom_{\mathcal{C}(\lambda)(\hat{B}^{\hat{A}} \times \hat{A}, \hat{B})}$.
+\end{itemize}
+\end{defin}
+
+It is sufficient to show ${\bf K}$ is an applicative functor on $\mathcal{C}(\lambda)$.
 
@@
   Let us define an endofunctor $\mathcal{K} : \mathcal{C}(\lambda) \to \mathcal{C}(\lambda)$, such that
-forall $[x,M] \in Hom_{\mathcal{C}(\lambda)}(A,B), {\bf K}([x,M]) = [y, {\bf let \: pure \:} x = y {\: \bf in \:} M] \in Hom_{\mathcal{C}(\lambda)}({\bf K} A, {\bf K} B)$ (denotation: $\text{fmap } f$ for an arbitrary arrow $f$).
+forall $[x,M] \in Hom_{\mathcal{C}(\lambda)}(\hat{A},\hat{B}), {\bf K}([x,M]) = [y, {\bf let \: pure \:} x = y {\: \bf in \:} M] \in Hom_{\mathcal{C}(\lambda)}({\bf K} \hat{A}, {\bf K} \hat{B})$ (denotation: $\text{fmap } f$ for an arbitrary arrow $f$).
 
@@
 
-  ii) $\text{fmap }(id_A) = id_{{\bf K}A}$.
+  ii) $\text{fmap }(id_{\hat{A}}) = id_{{\bf K}\hat{A}}$.
 \end{lemma}
@@
 
-  1) $\eta:Id \Rightarrow \mathcal{K}$, s. t. $\forall A \in Ob_{\mathcal{C}(\lambda)}$, $\eta_A = [x, {\bf pure \:} x] \in Hom_{\mathcal{C}(\lambda)}(A, {\bf K}A)$;
-
-  2) $\ast_{A,B}:{\bf K}A \times {\bf K}B \to {\bf K}(A \times B)$, s.
+  1) $\eta:Id \Rightarrow \mathcal{K}$, s. t. $\forall \hat{A} \in Ob_{\mathcal{C}(\lambda)}$, $\eta_{\hat{A}} = [x, {\bf pure \:} x] \in Hom_{\mathcal{C}(\lambda)}(\hat{A}, {\bf K}\hat{A})$;
+
+  2) $\ast_{A,B}:{\bf K}\hat{A} \times {\bf K}\hat{B} \to {\bf K}(\hat{A} \times \hat{B})$, s.
 t.
-$\forall A, B \in Ob_{\mathcal{C}(\lambda)}, \ast_{A,B} = [p, {\bf let \: pure \:} x,y = \pi_1 p, \pi_2 p {\: \bf in \:} \langle x, y \rangle] \in Hom_{\mathcal{C}(\lambda)}({\bf K}A \times {\bf K}B, {\bf K}(A \times B))$.
+$\forall \hat{A}, \hat{B} \in Ob_{\mathcal{C}(\lambda)}, \ast_{\hat{A},\hat{B}} = [p, {\bf let \: pure \:} x,y = \pi_1 p, \pi_2 p {\: \bf in \:} \langle x, y \rangle] \in Hom_{\mathcal{C}(\lambda)}({\bf K}A \times {\bf K}B, {\bf K}(A \times B))$.
 \end{defin}
@@
 
-  ii) $\text{fmap } (f \times g) \circ \ast_{A, B} = \ast_{C,D} \circ (\text{fmap } f) \times (\text{fmap } g)$.
-
-  iii) $\ast_{A,B} \circ (\eta_{A} \times \eta_{B}) = \eta_{A \times B}$;
-\end{lemma}
-
-\begin{proof}
-  $ $
-
-  i) $\text{fmap } f \circ \eta_A = \eta_B \circ f$
+  ii) $\text{fmap } (f \times g) \circ \ast_{\hat{A}, \hat{B}} = \ast_{\hat{C},\hat{D}} \circ (\text{fmap } f) \times (\text{fmap } g)$.
+
+  iii) $\ast_{\hat{A},\hat{B}} \circ (\eta_{A} \times \eta_{B}) = \eta_{\hat{A} \times \hat{B}}$;
+\end{lemma}
+
+\begin{proof}
+  $ $
+
+  i) $\text{fmap } f \circ \eta_{\hat{A}} = \eta_{\hat{B}} \circ f$
 
@@
 $\begin{array}{lll}
-&\eta_B \circ f = & \text{By the definition} \\
+&\eta_{\hat{B}} \circ f = & \text{By the definition} \\
 &[y, {\bf pure \:} y] \circ [x, M] = & \text{By the definition of composition}\\
@@
 &\text{On the other hand:}& \\
-&\text{fmap } f \circ \eta_A = & \text{By the definiton} \\
+&\text{fmap } f \circ \eta_{\hat{A}} = & \text{By the definiton} \\
 &[z, {\bf let \: pure \:} x = z {\: \bf in \:} M] \circ [x, {\bf pure \: x}] = & \text{By the definition of composition}\\
 &[x, {\bf let \: pure \:} x = z {\: \bf in \:} M [z := {\bf pure \:} x]] = & \text{By substitution} \\
-&[x, {\bf let \: pure \:} x = {\bf pure \: x} {\: \bf in \:} M] = & \text{$\beta$-reduction rule}\\
+&[x, {\bf let \: pure \:} x = {\bf pure \:} x {\: \bf in \:} M] = & \text{$\beta$-reduction rule}\\
 &[x, {\bf pure \:} M [x := x]] = & \text{By substitution}\\
@@
 
-ii)  $\text{fmap } (f \times g) \circ \ast_{A, B} = \ast_{C,D} \circ (\text{fmap } f) \times (\text{fmap } g)$
+ii)  $\text{fmap } (f \times g) \circ \ast_{\hat{A}, \hat{B}} = \ast_{\hat{C},\hat{D}} \circ (\text{fmap } f) \times (\text{fmap } g)$
 
@@
 
-iii) $\ast_{A,B} \circ (\eta_{A} \times \eta_{B}) = \eta_{A \times B}$
+iii) $\ast_{\hat{A},\hat{B}} \circ (\eta_{\hat{A}} \times \eta_{\hat{B}}) = \eta_{\hat{A} \times \hat{B}}$
 
@@
 $\begin{array}{lll}
-& \ast_{A,B} \circ (\eta_{A} \times \eta_{B}) = & \\
+& \ast_{\hat{A},\hat{B}} \circ (\eta_{\hat{A}} \times \eta_{\hat{B}}) = & \\
 & \text{By unfolding} & \\
@@
 & \text{By definition}& \\
-& \eta_{A \times B}&
+& \eta_{\hat{A} \times \hat{B}}&
 \end{array}$
@@
 
-  Let $[p, \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle] \in Hom_{\mathcal{C}(\lambda)}(A \times {\bf K}B, {\bf K}(A \times B))$.
-
-  So tensorial strength is defined as $\tau_{A, B} = \ast_{A, B} \circ [p, \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle]$.
+  Let $[p, \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle] \in Hom_{\mathcal{C}(\lambda)}(\hat{A} \times {\bf K}\hat{B}, {\bf K}\hat{A} \times {\bf K}\hat{B})$.
+
+  So tensorial strength is defined as $\tau_{\hat{A}, \hat{B}} = \ast_{\hat{A}, \hat{B}} \circ [p, \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle]$.
 \end{defin}
@@
 $\begin{array}{lll}
-&\ast_{A, B} \circ [p, \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle] = & \text{By definition} \\
+&\ast_{\hat{A}, \hat{B}} \circ [p, \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle] = & \text{By definition} \\
 &[p^{'}, {\bf let \: pure \:} x, y = \pi_1 p^{'}, \pi_2 p^{'} {\: \bf in \:} \langle x, y \rangle] \circ [p, \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle] = & \text{By composition} \\
@@
   $\begin{array}{lll}
-  & \text{fmap } ([p, \langle \pi_2 p, \pi_1 p \rangle]) \circ \tau_{A,B} = & \\
-  & \ast_{B, A} \circ [q, \langle \pi_1 q, {\bf pure \:} (\pi_2 q) \rangle] \circ [p, \langle \pi_2 p, \pi_1 p \rangle]&
+  & \text{fmap } ([p, \langle \pi_2 p, \pi_1 p \rangle]) \circ \tau_{\hat{A},\hat{B}} = & \\
+  & \ast_{\hat{B}, \hat{A}} \circ [q, \langle \pi_1 q, {\bf pure \:} (\pi_2 q) \rangle] \circ [p, \langle \pi_2 p, \pi_1 p \rangle]&
   \end{array}$
@@
   $\begin{array}{lll}
-  & \text{fmap } ([r, \langle \pi_2 r, \pi_1 r \rangle]) \circ \tau_{A,B} = & \\
+  & \text{fmap } ([r, \langle \pi_2 r, \pi_1 r \rangle]) \circ \tau_{\hat{A},\hat{B}} = & \\
   &\text{By the definition of $\tau$}& \\
@@
   &\text{On the other hand}&\\
-  &\ast_{B, A} \circ [q, \langle \pi_1 q, {\bf pure \:} (\pi_2 q) \rangle] \circ [p, \langle \pi_2 p, \pi_1 p \rangle] = & \\
+  &\ast_{\hat{B}, \hat{A}} \circ [q, \langle \pi_1 q, {\bf pure \:} (\pi_2 q) \rangle] \circ [p, \langle \pi_2 p, \pi_1 p \rangle] = & \\
   &\text{By the definition of $\ast$}& \\
@@
 
+Similar to \cite{Abramsky}, we apply the translation from $\lambda_{{\bf K}}$ to some cartesian closed category with an abritraty applicative functor $\mathcal{K}$, then
+we have $[\![\Gamma \vdash M : A]\!] = [x, M [x_i := \pi_i x]]$, so $M =_{\beta \eta} N \Leftrightarrow [\![\Gamma \vdash M : A]\!] = [\![\Gamma \vdash N : A]\!]$
+
 \begin{proof}
@@
 
+  \bibitem{Abramsky} Samson Abramsky and Nikos Tzevelekos, Introduction to Categories and Categorical Logic
 \end{thebibliography}

### 2018-01-29 article.tex
added

@@
 
+  If $M \in SC_A$, where $A$ is a modal-free type, then ${\bf pure \:} M \in SC_{{\bf K}A}$
+\end{lemma}
+
+\begin{proof}
+
+  Induction on the structure of M.
+\end{proof}
+
+\begin{lemma}
+  $ $
+
   Let $x_1 : A_1,\dots, x_n : A_n \vdash M : A$, then forall $i, M_i \in SC_{A_i}$. Then $M [x_1 := M_1,\dots,x_n := M_n] \in SC_A$.
@@
 
-Proof is similar to \cite{Baren} \cite{Baren2}. We build the following diagram, which commutes by lemmas 5 and 7.
+Proof is similar to \cite{Baren} \cite{Baren2}. We build the following diagram, which commutes by lemmas 8 and 10.
 
@@
 
+\begin{proof}
+  Immediately follows from previous lemmas in the section.
+\end{proof}
+
 Similar to \cite{Abramsky}, we apply the translation from $\lambda_{{\bf K}}$ to some cartesian closed category with an abritraty applicative functor $\mathcal{K}$, then
-we have $[\![\Gamma \vdash M : A]\!] = [x, M [x_i := \pi_i x]]$, so $M =_{\beta \eta} N \Leftrightarrow [\![\Gamma \vdash M : A]\!] = [\![\Gamma \vdash N : A]\!]$
-
-\begin{proof}
-  Immediately follows from previous lemmas in the section.
-\end{proof}
+we have $[\![\Gamma \vdash M : A]\!] = [x, M [x_i := \pi_i x]]$, so $M =_{\beta \eta} N \Leftrightarrow [\![\Gamma \vdash M : A]\!] = [\![\Gamma \vdash N : A]\!]$.
 

### 2018-01-29 article.tex
added fixes
fixes

@@
 
-  If $M \in SC_A$, where $A$ is a modal-free type, then ${\bf pure \:} M \in SC_{{\bf K}A}$
+  If $M \in SC_A$, then ${\bf pure \:} M \in SC_{{\bf K}A}$
 \end{lemma}
@@
 
-So ${\bf let \: pure \:} \vec{x} = \vec{M^{'}} [x_1 := M_1,\dots,x_n := M_n] {\: \bf in \:} N \in SC_{{\bf K} B}$.
+So ${\bf let \: pure \:} \vec{x} = \vec{M^{'}} [x_1 := M_1,\dots,x_n := M_n] {\: \bf in \:} N \in SC_{{\bf K} B}$,
+otherwise we can build infinite reduction path in $\vec{M^{'}} [x_1 := M_1,\dots,x_n := M_n]$.
 \end{proof}
@@
 
-  1) $(x, M) \sim_{A, B} (y, N) \Leftrightarrow x : A \vdash M : B \:\: \& \:\: y : A \vdash N : A \:\: \& \:\: M =_{\beta \eta} N [y := x]$;
-
-  2) $[x, M]_{A, B} = \{ (y, N) \: | \: (x, M) \sim_{A, B} (y, N) \}$.
-\end{defin}
-
-We will drop indeces below.
+  Let us define relation $\sim_{A, B} \subseteq \mathbb{V} \times \Lambda_{{\bf K}}$, such that:
+
+  $(x, M) \sim_{A, B} (y, N) \Leftrightarrow x : A \vdash M : B \:\: \& \:\: y : A \vdash N : A \:\: \& \:\: M =_{\beta \eta} N [y := x]$;
+\end{defin}
+
+We will denote equivalence class as $[x, M]_{A, B} = \{ (y, N) \: | \: (x, M) \sim_{A, B} (y, N) \}$ (we will drop indeces below).
 
@@
   \item $Ob_{\mathcal{C}} = \{ \hat{A} \: | \: A \in \mathbb{T} \} \cup \{ \mathds{1} \}$;
-  \item $Hom_{\mathcal{C}(\lambda)}(\hat{A},\hat{B}) = \{ [x,M] \: | \: x : A \vdash M : B \}$;
+  \item $Hom_{\mathcal{C}(\lambda)}(\hat{A},\hat{B}) = (\mathbb{V} \times \Lambda_{{\bf K}})/_{\sim_{A, B}}$;
   \item Let $[x, M] \in Hom_{\mathcal{C}(\lambda)}(\hat{A},\hat{B})$ and $[y,N] \in Hom_{\mathcal{C}(\lambda)}(\hat{B},\hat{C})$, then $[y,M] \circ [x, M] = [x, N [y := M]]$;
@@
 
-  i) $\text{fmap }(g \circ f) = \text{fmap }(g) \circ \text{fmap }(f)$;
-
-  ii) $\text{fmap }(id_{\hat{A}}) = id_{{\bf K}\hat{A}}$.
+\begin{itemize}
+  \item $\text{fmap }(g \circ f) = \text{fmap }(g) \circ \text{fmap }(f)$;
+  \item $\text{fmap }(id_{\hat{A}}) = id_{{\bf K}\hat{A}}$.
+\end{itemize}
 \end{lemma}
@@
 
+\begin{lemma}
+  $ $
+
+  {\bf K} is a monoidal endofunctor
+\end{lemma}
+
+\begin{proof}
+$ $
+
+See \cite{ModalK}
+\end{proof}
+
 \begin{lemma} Naturality for $\eta$ and for $\ast$
 
-  i) $\text{fmap } f \circ \eta_A = \eta_B \circ f$;
-
-  ii) $\text{fmap } (f \times g) \circ \ast_{\hat{A}, \hat{B}} = \ast_{\hat{C},\hat{D}} \circ (\text{fmap } f) \times (\text{fmap } g)$.
-
-  iii) $\ast_{\hat{A},\hat{B}} \circ (\eta_{A} \times \eta_{B}) = \eta_{\hat{A} \times \hat{B}}$;
+\begin{itemize}
+  \item $\text{fmap } f \circ \eta_A = \eta_B \circ f$;
+  \item $\ast_{\hat{A},\hat{B}} \circ (\eta_{A} \times \eta_{B}) = \eta_{\hat{A} \times \hat{B}}$;
+\end{itemize}
 \end{lemma}
@@
 
-ii)  $\text{fmap } (f \times g) \circ \ast_{\hat{A}, \hat{B}} = \ast_{\hat{C},\hat{D}} \circ (\text{fmap } f) \times (\text{fmap } g)$
-
-\vspace{\baselineskip}
-
-See \cite{ModalK}.
-
-\vspace{\baselineskip}
-
-iii) $\ast_{\hat{A},\hat{B}} \circ (\eta_{\hat{A}} \times \eta_{\hat{B}}) = \eta_{\hat{A} \times \hat{B}}$
+ii) $\ast_{\hat{A},\hat{B}} \circ (\eta_{\hat{A}} \times \eta_{\hat{B}}) = \eta_{\hat{A} \times \hat{B}}$
 

### 2018-01-30 article.tex
small fixing

@@
   functional programming with effects.
-We show that relationship by building the type system (which is called $\lambda_{\textbf{K}}$) which is Curry-Howard isomorphic to $IEL^{-}$.
-So we will consider $\textbf{K}$-modality as an arbitrary applicative functor.
+We show that relationship by building the type system (which is called $\lambda_{{\bf K}}$) which is Curry-Howard isomorphic to $IEL^{-}$.
+So we will consider ${\bf K}$-modality as an arbitrary applicative functor.
 
   $\lambda_{{\bf K}}$ consists of the rules for simply typed lambda-calculus and special typing rules for
-  lifting types into the applicative functor $\textbf{K}$. We assume that our type system will
+  lifting types into the applicative functor ${\bf K}$. We assume that our type system will
   axiomatize the simplest case of computation with effects with one container. We provide proof-theoretical
@@
     	\AxiomC{$\Gamma \vdash A$}
-    	\RightLabel{$\textbf{K}_I$}
-    	\UnaryInfC{$\Gamma \vdash \textbf{K}A$}
+    	\RightLabel{${\bf K}_I$}
+    	\UnaryInfC{$\Gamma \vdash {\bf K}A$}
   \end{prooftree}
@@
 
-1) If $\Gamma \vdash_{NIEL^{-}_{\land, \to}} A$, then $IEL^{-} \vdash \bigwedge \Gamma \rightarrow \textbf{K}A$.
+1) If $\Gamma \vdash_{NIEL^{-}_{\land, \to}} A$, then $IEL^{-} \vdash \bigwedge \Gamma \rightarrow {\bf K}A$.
 
@@
 (1) & \bigwedge \Gamma \rightarrow A & \text{assumption}\\
-(2) & A \rightarrow \textbf{K}A &\text{co-reflection}\\
-(3) & (\bigwedge \Gamma \rightarrow A) \rightarrow ((A \rightarrow \textbf{K}A) \rightarrow (\bigwedge \Gamma \rightarrow \textbf{K}A))&\text{IPC theorem}\\
-(4) & (A \rightarrow \textbf{K}A) \rightarrow (\bigwedge \Gamma \rightarrow \textbf{K}A) &\text{from (1), (3) and MP}\\
-(5) & \bigwedge \Gamma \rightarrow \textbf{K}A &\text{from (2), (4) and MP}\\
-\end{array}$
-
-\vspace{\baselineskip}
-
-2) If $\Gamma \vdash_{NIEL^{-}_{\land, \to}} {\bf K} \vec{A}$ and $\vec{A} \vdash B$, then $IEL^{-} \vdash \bigwedge \Gamma \rightarrow \textbf{K}B$.
+(2) & A \rightarrow {\bf K}A &\text{co-reflection}\\
+(3) & (\bigwedge \Gamma \rightarrow A) \rightarrow ((A \rightarrow {\bf K}A) \rightarrow (\bigwedge \Gamma \rightarrow {\bf K}A))&\text{IPC theorem}\\
+(4) & (A \rightarrow {\bf K}A) \rightarrow (\bigwedge \Gamma \rightarrow {\bf K}A) &\text{from (1), (3) and MP}\\
+(5) & \bigwedge \Gamma \rightarrow {\bf K}A &\text{from (2), (4) and MP}\\
+\end{array}$
+
+\vspace{\baselineskip}
+
+2) If $\Gamma \vdash_{NIEL^{-}_{\land, \to}} {\bf K} \vec{A}$ and $\vec{A} \vdash B$, then $IEL^{-} \vdash \bigwedge \Gamma \rightarrow {\bf K}B$.
 
@@
 (5) &(\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B) \rightarrow {\bf K} (\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B)& \text{co-reflection}\\
-(6) &{\bf K} (\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B)& \text{from (2), (3) and MP} \\
+(6) &{\bf K} (\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B)& \text{from (4), (5) and MP} \\
 (7) &{\bf K} \bigwedge \limits_{i = 1}^{n} A_i \rightarrow {\bf K} B & \text{from (6) and normality} \\
@@
         \AxiomC{$\Gamma \vdash M : A$}
-        \RightLabel{$\textbf{K}_I$}
+        \RightLabel{${\bf K}_I$}
         \UnaryInfC{$\Gamma \vdash {\bf pure \: } \: M : {\bf K}A $}
@@
 \begin{itemize}
-\item $(\lambda x. {\bf pure \: } x) : A \to \textbf{K}A$;
+\item $(\lambda x. {\bf pure \: } x) : A \to {\bf K}A$;
 \item $\lambda f. \lambda x. {\bf let \: pure} \: g, y  = f, x {\: \bf in \:} g y : {\bf K}(A \to B) \to {\bf K} A \to {\bf K} B$
@@
 
-  \begin{defin} $\beta$-reduction and $\eta$-reduction rules for \emph{$\lambda \textbf{K}$}.
+  \begin{defin} $\beta$-reduction and $\eta$-reduction rules for $\lambda_{{\bf K}}$.
 
@@
   4) $\begin{array}{llll}
-  & {\bf let \: pure \:} \langle \vec{x}, y, \vec{z} \rangle = \langle \vec{M}, {\bf let \: pure \: } \vec{w} = \vec{N} {\bf \: in \: } Q, \vec{P} \rangle {\: in \:} R \rightarrow_{\beta} & \\
-  & {\bf let \: pure \:} \langle \vec{x}, \vec{w}, \vec{z} \rangle = \langle \vec{M}, \vec{N}, \vec{P} \rangle \: { \bf in } \: R [y := Q]
+  & {\bf let \: pure \:} \vec{x}, y, \vec{z} = \vec{M}, {\bf let \: pure \: } \vec{w} = \vec{N} {\: \bf in \: } Q, \vec{P} {\: in \:} R \rightarrow_{\beta} & \\
+  & {\bf let \: pure \:} \vec{x}, \vec{w}, \vec{z} = \vec{M}, \vec{N}, \vec{P} {\: \bf in \: } R [y := Q]
   \end{array}$
@@
 1)
-  Let $\Gamma \vdash {\bf let \: pure \:} \vec{x}, y, \vec{z}  = \langle \vec{M}, {\bf let \: pure \: } \vec{w} = \vec{N} {\bf \: in \: } Q, \vec{P} \rangle {\: in \:} R : {\bf K}B$,
-  then $\Gamma \vdash {\bf let \: pure \:} \vec{x}, \vec{w}, \vec{z} = \langle \vec{M}, \vec{N}, \vec{P} \rangle \: { \bf in } \: R [y := Q] : {\bf K}B$
+  Let $\Gamma \vdash {\bf let \: pure \:} \vec{x}, y, \vec{z} = \vec{M}, {\bf let \: pure \: } \vec{w} = \vec{N} {\bf \: in \: } Q, \vec{P}  {\: \bf in \:} R : {\bf K}B$,
+  then $\Gamma \vdash {\bf let \: pure \:} \vec{x}, \vec{w}, \vec{z} = \vec{M}, \vec{N}, \vec{P} \: { \bf in } \: R [y := Q] : {\bf K}B$
 
@@
   \item $\begin{array}{llll}
-  & {\bf let \: pure \:} \langle \vec{x}, y, \vec{z} \rangle = \langle \vec{M}, {\bf let \: pure \: } \vec{w} = \vec{N} {\bf \: in \: } Q, \vec{P} \rangle {\: in \:} R \rightarrow_{\underline{\beta}} & \\
-  & {\bf let \: pure \:} \langle \vec{x}, \vec{w}, \vec{z} \rangle = \langle \vec{M}, \vec{N}, \vec{P} \rangle \: { \bf in } \: R [y := Q]
+  & {\bf let \: pure \:} \vec{x}, y, \vec{z} = \vec{M}, {\bf let \: pure \: } \vec{w} = \vec{N} {\bf \: in \: } Q, \vec{P} {\: in \:} R \rightarrow_{\underline{\beta}} & \\
+  & {\bf let \: pure \:} \vec{x}, \vec{w}, \vec{z} = \vec{M}, \vec{N}, \vec{P} \: { \bf in } \: R [y := Q]
   \end{array}$;
@@
 
-  Normal form in $\lambda_{{\bf K}}$ has the subformula property.
+  Normal form in call-by-name $\lambda_{{\bf K}}$ has the subformula property: if $M$ is in normal formal, then its all subterms are in normal form too.
 
@@
 \begin{proof}
-  By induction on the structure of term. Case with ${\bf let \: pure\:} \vec{x} = \vec{M} {\: \bf in \:} N$ was considered by Kakutani \cite{ModalK} \cite{ModalK1}.
-  Similary, if ${\bf pure \:} M$ is a normal form, so $M$ is a normal form too by hypothesis.
+  By induction on the structure of $M$.
+  Case with ${\bf let \: pure\:} \vec{x} = \vec{M} {\: \bf in \:} N$ was considered by Kakutani \cite{ModalK} \cite{ModalK1}.
+  Similary, if ${\bf pure \:} M$ is in normal form, so $M$ is in normal form to and has subformula property by hypothesis.
+  So ${\bf pure \:} M$ has subformula property.
 \end{proof}
@@
 
-\begin{defin} Semantical translation from $\lambda_{\textbf{K}}$ to some cartesian closed category $\mathcal{C}$ with applicative functor $\mathcal{K}$:
+\begin{defin} Semantical translation from $\lambda_{{\bf K}}$ to some cartesian closed category $\mathcal{C}$ with applicative functor $\mathcal{K}$:
 
@@
 
-2) Interpretation for modal types: $[\![\textbf{K}A]\!] = \mathcal{K}[\![A]\!]$;
+2) Interpretation for modal types: $[\![{\bf K}A]\!] = \mathcal{K}[\![A]\!]$;
 
@@
 \AxiomC{$[\![\Gamma \vdash M : A]\!] = [\![M]\!] : [\![\Gamma]\!] \rightarrow [\![A]\!]$}
-\UnaryInfC{$[\![\Gamma \vdash {\bf pure \:} M : \textbf{K}A]\!] := [\![\Gamma]\!] \xrightarrow{[\![M]\!]}
+\UnaryInfC{$[\![\Gamma \vdash {\bf pure \:} M : {\bf K}A]\!] := [\![\Gamma]\!] \xrightarrow{[\![M]\!]}
 [\![A]\!] \xrightarrow{\eta_{[\![A]\!]}} \mathcal{K}[\![A]\!]$}
@@
 $\begin{array}{lll}
-&[\![\Gamma \vdash ({\bf let \: pure} \: \vec{x} = \vec{M} {\: \bf in \:} N) [\vec{y} := \vec{P}] : {\bf K} B]\!] =& \\
-&\text{Substitution definition}& \\
-&[\![\Gamma \vdash {\bf let \: pure} \: \vec{x} = (\vec{M} [\vec{y} := \vec{P}]) {\: \bf in \:} N : {\bf K}B]\!] =& \\
-&\text{Interpretaion for $let_{\bf K}$}& \\
-&\mathcal{K}([\![N]\!]) \circ \ast_{[\![A_1]\!],\dots,[\![A_n]\!]} \circ [\![\Gamma \vdash (\vec{M} [\vec{y} := \vec{P}]) \vdash : {\bf K}\vec{A}]\!] =& \\
-&\text{Induction hypothesis}& \\
-&\mathcal{K}([\![N]\!]) \circ \ast_{[\![A_1]\!],\dots,[\![A_n]\!]} \circ ([\![\vec{M}]\!] \circ \langle [\![P_1]\!],\dots,[\![P_n]\!]\rangle) = & \\
-&\text{Associativity of composition}& \\
-&(\mathcal{K}([\![N]\!]) \circ \ast_{[\![A_1]\!],\dots,[\![A_n]\!]} \circ [\![\vec{M}]\!]) \circ \langle [\![P_1]\!],\dots,[\![P_n]\!]\rangle = & \\
-&\text{By interpretation}&\\
-&[\![\Gamma \vdash ({\bf let \: pure} \: \vec{x} = \vec{M} {\: \bf in \:} N]\!] \circ \langle [\![P_1]\!],\dots,[\![P_n]\!]\rangle&
+&[\![\Gamma \vdash ({\bf let \: pure} \: \vec{x} = \vec{M} {\: \bf in \:} N) [\vec{y} := \vec{P}] : {\bf K} B]\!] =& \text{Substitution definition}\\
+&[\![\Gamma \vdash {\bf let \: pure} \: \vec{x} = (\vec{M} [\vec{y} := \vec{P}]) {\: \bf in \:} N : {\bf K}B]\!] =& \text{Interpretaion for $let_{\bf K}$} \\
+&\mathcal{K}([\![N]\!]) \circ \ast_{[\![A_1]\!],\dots,[\![A_n]\!]} \circ [\![\Gamma \vdash (\vec{M} [\vec{y} := \vec{P}]) : {\bf K}\vec{A}]\!] =& \text{Induction hypothesis} \\
+&\mathcal{K}([\![N]\!]) \circ \ast_{[\![A_1]\!],\dots,[\![A_n]\!]} \circ ([\![\vec{M}]\!] \circ \langle [\![P_1]\!],\dots,[\![P_n]\!]\rangle) = & \text{Associativity of composition} \\
+&(\mathcal{K}([\![N]\!]) \circ \ast_{[\![A_1]\!],\dots,[\![A_n]\!]} \circ [\![\vec{M}]\!]) \circ \langle [\![P_1]\!],\dots,[\![P_n]\!]\rangle = & \text{By interpretation} \\
+&[\![\Gamma \vdash {\bf let \: pure} \: \vec{x} = \vec{M} {\: \bf in \:} N : {\bf K}B]\!] \circ \langle [\![P_1]\!],\dots,[\![P_n]\!]\rangle&
 \end{array}$
@@
 \BinaryInfC{$x : A, y : B \vdash \langle x, y \rangle : A \times B$}
-\TrinaryInfC{$p : {\bf K}A \times {\bf K}B \vdash {\bf let \: pure \:} \langle x,y \rangle = \langle \pi_1 p, \pi_2 p \rangle {\: \bf in \:} \langle x, y \rangle : {\bf K}(A \times B)$}
+\TrinaryInfC{$p : {\bf K}A \times {\bf K}B \vdash {\bf let \: pure \:} x,y = \pi_1 p, \pi_2 p {\: \bf in \:} \langle x, y \rangle : {\bf K}(A \times B)$}
 \end{prooftree}

### 2018-01-31 article.tex
update

@@
 \usepackage{dsfont}
+\usepackage{mathabx}
 \usepackage[all, 2cell]{xy}
@@
 
-  6) $\lambda x. f x \rightarrow_{\eta} f$;
-
-  7) $\langle \pi_1 P, \pi_2 P \rangle \rightarrow_{\eta} P$;
-
-  8) ${\bf let \: pure \:} x = M {\: \bf in \: } x \rightarrow_{\eta} M$;
+  6) ${\bf let \: pure \:} \underline{\quad} = \underline {\quad} {\: \bf in \:} M \rightarrow_{\beta} {\bf pure \:} M$
+
+  7) $\lambda x. f x \rightarrow_{\eta} f$;
+
+  8) $\langle \pi_1 P, \pi_2 P \rangle \rightarrow_{\eta} P$;
+
+  9) ${\bf let \: pure \:} x = M {\: \bf in \: } x \rightarrow_{\eta} M$;
+
   \end{defin}
@@
 By default we use call-by-name evaluation strategy.
-
-  \section{Basic lemmas}
 
@@
 $\to$-constructor, they are proved standardly in the same lemmas for simply typed lambda calculus, for
-  example see \cite{Neder} \cite{Morten} \cite{Girard}. We will consider only modal cases}:
+example see \cite{Neder} \cite{Morten}. We will consider only modal cases}:
 
@@
 
+4) If the derivation ends in
+
+\begin{prooftree}
+\AxiomC{$\vdash M : A$}
+\UnaryInfC{$\vdash {\bf let \: pure \:} \underline{\quad} = \underline {\quad} {\: \bf in \:} M : {\bf K} A$}
+\end{prooftree}
+
+So, if $\vdash M : A$, then $\vdash {\bf pure \:} M : {\bf K}A$.
+
+Note that this part of the lemma works conversly too.
 \end{proof}
@@
     \item $SC_{{\bf K}A} = \{ M : {\bf K}A \: | \: M \text{ is strongly normalizing} \}$ for $A \in \mathbb{T}$;
-    \item $SC_{{\bf K}(A \to B)} = \{ M : {\bf K}(A \to B) \: | \: \forall f \in SC_{A \to B}, \forall x \in SC_A, \forall N \in SC_{{\bf K}A}, {\bf let \: pure \:} f, x = M, N {\: \bf in \:} f x \}$ for $A,B \in \mathbb{T}_{\bf K}$.
+    \item $SC_{{\bf K}(A \to B)} = \{ M : {\bf K}(A \to B) \: | \: \forall f \in SC_{A \to B}, \forall x \in SC_A, \forall N \in SC_{{\bf K}A}, {\bf let \: pure \:} f, x = M, N {\: \bf in \:} f x \in SC_{{\bf K}B} \}$ for $A,B \in \mathbb{T}_{\bf K}$.
   \end{itemize}
@@
   \item ${\bf let \: pure \:} \vec{x} = {\bf pure \:} \vec{M} {\: \bf in \:} N \rightarrow_{\underline{\beta}} {\bf pure \:} N [\vec{x} := \vec{M}]$;
+  \item ${\bf let \: pure \:} \underline{\quad} = \underline {\quad} {\: \bf in \:} M \rightarrow_{\underline{\beta}} {\bf pure \:} M$
   \item $(\lambda x_i. M) N \rightarrow_{\underline{\beta}} M [x := N]$
@@
     \item $\phi({\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N) = {\bf let \: pure \:} \vec{x} = \vec{\phi(M)} {\: \bf in \:} \phi(N)$;
-    \item $\phi((\lambda_i x. M) N) = M [x := N]$
+    \item $\phi((\lambda_i x. M) N) = \phi(M) [x := \phi(N)]$
   \end{itemize}
 \end{defin}
+
+\begin{ex}
+  $ $
+
+  $\phi((\lambda_i x. {\bf let \: pure \:} y = x {\: \bf in \:} x) N) = {\bf let \: pure \:} y = \phi(N) {\: \bf in \:} \phi(N)$
+\end{ex}
 
@@
 
+\vspace{\baselineskip}
+
+2) $[\![\vdash {\bf let \: pure \:} \underline{\quad} = \underline {\quad} {\: \bf in \:} M : {\bf K} A]\!] = [\![\vdash {\bf pure \:} M : {\bf K} A]\!]$
+
+$\begin{array}{lll}
+&[\![\vdash {\bf let \: pure \:} \underline{\quad} = \underline {\quad} {\: \bf in \:} M : {\bf K} A]\!] = & \text{By interpretation} \\
+&\mathcal{K}([\![M]\!]) \circ u_{\mathds{1}} = & \text{By the definition of an applicative functor} \\
+&\mathcal{K}([\![M]\!]) \circ \eta_{\mathds{1}} = & \text{By naturality of $\eta$} \\
+&\eta_{[\![A]\!]} \circ [\![M]\!] = & \text{By interpretation} \\
+&[\![\vdash {\bf pure \:} M : {\bf K} A]\!]&
+\end{array}$
 \end{proof}
@@
 
-\begin{lemma} Naturality for $\eta$ and for $\ast$
+\begin{lemma} Properties of $\eta$:
 
@@
 
+\end{proof}
+
+\begin{defin}
+  $ $
+
+  $u_{\mathds{1}} = [\sqbullet, {\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} \sqbullet] \in Hom_{\mathcal{C}(\lambda)}(\mathds{1}, {\bf K}\mathds{1})$.
+\end{defin}
+
+\begin{lemma}
+  $ $
+
+  $u_{\mathds{1}} = \eta_{\mathds{1}}$
+
+\end{lemma}
+
+\begin{proof}
+
+  Immediately.
 \end{proof}

### 2018-02-01 article.tex
added

@@
 
-  Intuitionistic epistemic logic IEL$^{-}$ is defined with by following axioms and derivation rules:
+  Intuitionistic epistemic logic IEL$^{-}$ is defined by following axioms and derivation rules:
 
@@
   \end{defin}
-
-  We have the deduction theorem and necessitation rule which is derivable.
 
@@
   By \emph{container} (or \emph{computational context}) type we mean some type-operator $f$, where $f$ is a
-  ``function'' from $*$ to $*$: type operator takes a simple type (which has kind $*$) and returns another
-  simple type type with kind \verb"*". For more detailed description of the type system with kinds used in
+  ``function'' from $*$ to $*$: type operator takes a simple type (ща kind $*$) and returns another
+  simple type of kind \verb"*". For more detailed description of the type system with kinds used in
   Haskell see \cite{Morten}.
 
-  The motivation for using an applicative functor is quite natural. Applicative functor allows to generalize the action of a functor for functions with arbitrary number of arguments, for instance:
+Applicative functor allows to generalize the action of a functor for functions with arbitrary number of arguments, for instance:
   \begin{lstlisting}[language=Haskell]
   liftA2 :: Applicative f => (a -> b -> c) -> f a -> f b -> f c
-  liftA2 f x y = pure f <*> x <*> y
+  liftA2 f x y = ((pure f) <*> x) <*> y
   \end{lstlisting}
@@
 
-  The main goal of our research is a relationship between intuitionistic epistemic logic $IEL^{-}$ and
-  functional programming with effects.
-We show that relationship by building the type system (which is called $\lambda_{{\bf K}}$) which is Curry-Howard isomorphic to $IEL^{-}$.
-So we will consider ${\bf K}$-modality as an arbitrary applicative functor.
+  We investigate the relationship between intuitionistic epistemic logic IEL$^{-}$ and applicative programming with side-effects
+  by constructing the type system (which is called $\lambda_{{\bf K}}$) which is Curry-Howard isomorphic to $IEL^{-}$.
+  So we will consider ${\bf K}$-modality as an arbitrary applicative functor and we prove that obtained type system
+  is sound and complete for applicative functor on cartesian closed category (using the categorical definition proposed by Paterson \cite{Cons}).
 
@@
   lifting types into the applicative functor ${\bf K}$. We assume that our type system will
-  axiomatize the simplest case of computation with effects with one container. We provide proof-theoretical
-  view on
-  this kind of computations in functional programming and prove strong normalization and confluence.
+  axiomatize the simplest case of computation with effects with one container. We provide a proof-theoretical
+  view at this kind of computations in functional programming and prove strong normalization and confluence.
 
@@
 
-  At first we define the natural deduction for IEL$^{-}$ :
-
-  \begin{defin} Natural deduction NIEL for IEL$^{-}$ is an extenstion of intuitionistic natural deduction with additional derivation rules for modality:
+  The first is to define the natural deduction calculus for NIEL$^{-}$ :
+
+  \begin{defin} Natural deduction NIEL$^{-}$ for IEL$^{-}$ is an extenstion of intuitionistic natural deduction calculus
+    with additional inference rules for modality:
 
@@
   	\begin{prooftree}
-    \AxiomC{$\Gamma \vdash {\bf K} \vec{A}$}
-    \AxiomC{$\vec{A} \vdash B$}
+    \AxiomC{$\Gamma \vdash {\bf K} A_1, \dots, \Gamma \vdash {\bf K} A_n $}
+    \AxiomC{$A_1,\dots,A_n \vdash B$}
     \BinaryInfC{$\Gamma \vdash {\bf K} B$}
@@
 
-Where $\Gamma \vdash {\bf K} \vec{A}$ is a syntax sugar for $\Gamma \vdash {\bf K} A_1 ,\dots, \Gamma \vdash {\bf K} A_n$.
+The first rule allows to derive co-reflextion. The second modal rule is a counterpart of $\Box_I$
+rule in natural deduction calculus for constructive K (see \cite{ModalLa}).
+
+We will denote $\Gamma \vdash {\bf K} A_1, \dots, \Gamma \vdash {\bf K} A_n$ and $\Gamma \vdash {\bf K} B$ as $\Gamma \vdash {\bf K} \vec{A}$ and $\vec{A} \vdash B$ for brevity.
 
@@
   \begin{lemma}
-    $\Gamma \vdash_{\text{NIEL}^{-}_{\land, \to}} A \Rightarrow$ IEL$^{-} \vdash \bigwedge \Gamma \rightarrow A$.
+    $\Gamma \vdash_{\text{NIEL}^{-}} A \Rightarrow$ IEL$^{-} \vdash \bigwedge \Gamma \rightarrow A$.
   \end{lemma}
@@
 
-1) If $\Gamma \vdash_{NIEL^{-}_{\land, \to}} A$, then $IEL^{-} \vdash \bigwedge \Gamma \rightarrow {\bf K}A$.
+1) If $\Gamma \vdash_{\text{NIEL}^{-}} A$, then $\text{IEL}^{-} \vdash \bigwedge \Gamma \rightarrow {\bf K}A$.
 
@@
 
-2) If $\Gamma \vdash_{NIEL^{-}_{\land, \to}} {\bf K} \vec{A}$ and $\vec{A} \vdash B$, then $IEL^{-} \vdash \bigwedge \Gamma \rightarrow {\bf K}B$.
+2) If $\Gamma \vdash_{\text{NIEL}^{-}} {\bf K} \vec{A}$ and $\vec{A} \vdash B$, then $\text{IEL}^{-} \vdash \bigwedge \Gamma \rightarrow {\bf K}B$.
 
@@
 (1) &\bigwedge \Gamma \rightarrow \bigwedge \limits_{i = 1}^{n} {\bf K} A_i & \text{assumption} \\
-(2) &\bigwedge \limits_{i = 1}^{n} {\bf K} A_i \rightarrow {\bf K} \bigwedge \limits_{i = 1}^{n} A_i& \text{IEL theorem} \\
+(2) &\bigwedge \limits_{i = 1}^{n} {\bf K} A_i \rightarrow {\bf K} \bigwedge \limits_{i = 1}^{n} A_i& \text{IEL$^{-}$ theorem} \\
 (3) &\bigwedge \Gamma \rightarrow {\bf K} \bigwedge \limits_{i = 1}^{n} A_i & \text{from (1), (2) and transitivity} \\
@@
 
+  $\Gamma \vdash \vec{M} : {\bf K} \vec{A}$ is a syntax sugar for the sequence $\Gamma \vdash M_1 : {\bf K}A_1,\dots,\Gamma \vdash M_n : {\bf K}A_n$ and $\vec{x} : \vec{A} \vdash N : B$ is an abbrevation for $x_1 : A_1, \dots, x_n : A_n \vdash N : B$.
+${\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } N$ is a simultaneous local binding in $N$.
+We use abbreviated notation instead of ${\bf let \: pure\:} x_1,\dots,x_n = M_1,\dots,M_n {\: \bf in \:} N$.
+
   In fact, our calculus is the extention of typed lambda calculus for {\bf IK} with typing rule appropriate to co-reflection.
@@
 
-\begin{itemize}
-\item $(\lambda x. {\bf pure \: } x) : A \to {\bf K}A$;
-\item $\lambda f. \lambda x. {\bf let \: pure} \: g, y  = f, x {\: \bf in \:} g y : {\bf K}(A \to B) \to {\bf K} A \to {\bf K} B$
-\item $\lambda f. \lambda x. {\bf let \:pure} \: g, y = {\bf pure \:} f, x  {\: \bf in \: } g y : (A \to B) \to {\bf K} A \to {\bf K} B$
-\end{itemize}
+\begin{prooftree}
+\AxiomC{$x : A \vdash x : A$}
+\UnaryInfC{$x : A \vdash {\bf pure \:} x : {\bf K} A$}
+\UnaryInfC{$\vdash (\lambda x. {\bf pure \: } x) : A \to {\bf K}A$}
+\end{prooftree}
+
+\begin{prooftree}
+\AxiomC{$f : {\bf K} (A \to B) \vdash f : {\bf K} (A \to B)$}
+\AxiomC{$x : {\bf K} A \vdash x : {\bf K} A $}
+\AxiomC{$g : A \to B \vdash g : A \to B$}
+\AxiomC{$y : A \vdash y : A$}
+\RightLabel{$\to_e$}
+\BinaryInfC{$g : A \to B, y : A \vdash g y : B$}
+\RightLabel{$\text{let}_{{\bf K}}$}
+\TrinaryInfC{$f : {\bf K} (A \to B), x : {\bf K} A \vdash {\bf let \: pure \:} g, y  = f, x {\: \bf in \:} g y : {\bf K} B$}
+\UnaryInfC{$f : {\bf K} (A \to B) \vdash \lambda x. {\bf let \: pure\:} g, y = f, x {\: \bf in \:} g y : {\bf K} A \to {\bf K} B$}
+\UnaryInfC{$\vdash \lambda f. \lambda x. {\bf let \:pure \:} g, y = f, x {\: \bf in \:} g y : {\bf K}(A \to B) \to {\bf K} A \to {\bf K} B$}
+\end{prooftree}
 
@@
 
-  7) $FV({\bf let \:\: pure} \:\: \vec{N} = \vec{M} \:\: {\bf in} \:\: M) = \bigcup \limits_{i = 1}^n FV(M), \text{where $n = |\vec{M}|$}$.
+  7) $FV({\bf let \:\: pure} \:\: \vec{x} = \vec{M} \:\: {\bf in} \:\: N) = \bigcup \limits_{i = 1}^n FV(M), \text{where $n = |\vec{M}|$}$.
   \end{defin}
@@
 
-  7) $({\bf let \: pure \:}\vec{x} = \vec{M} {\: \bf in \:} N) [y := P] = {\bf let \: pure \:} \vec{x} = (\vec{M} [y := P]) {\: \bf in \:} M$.
+  7) $({\bf let \: pure \:}\vec{x} = \vec{M} {\: \bf in \:} N) [y := P] = {\bf let \: pure \:} \vec{x} = (\vec{M} [y := P]) {\: \bf in \:} N$.
   \end{defin}
@@
 
-  i) Let $\Gamma \vdash M : A$ and $\Gamma \subseteq \Delta$, then $\Delta \vdash M : A$;
-
-  ii) Let $\Gamma, x : A \vdash M : B$ and $\Gamma \vdash N : A$, then $\Gamma \vdash M [x := N] : B$.
-
-  iii) Let $\Gamma \vdash M : A$, then $\Gamma [B := C] \vdash M : (A [B := C])$.
-\end{lemma}
-
-\begin{proof}
-  $ $
-
-  i-ii-iii) Induction on $\Gamma \vdash M : A$.
+\begin{itemize}
+  \item Let $\Gamma \vdash M : A$ and $\Gamma \subseteq \Delta$, then $\Delta \vdash M : A$;
+  \item Let $\Gamma \vdash M : A$, then $\Delta \vdash M : A$, where $\Delta = \{ x_i : A_i \: | \: (x_i : A_i) \in \Gamma \: \& \: x_i \in FV(M) \}$
+  \item Let $\Gamma, x : A \vdash M : B$ and $\Gamma \vdash N : A$, then $\Gamma \vdash M [x := N] : B$.
+  \item Let $\Gamma \vdash M : A$, then $\Gamma [B := C] \vdash M : (A [B := C])$
+\end{itemize}
+\end{lemma}
+
+\begin{proof}
+  $ $
+
+  i--iv) Induction on $\Gamma \vdash M : A$.
 
@@
     \item $SC_{{\bf K}A} = \{ M : {\bf K}A \: | \: M \text{ is strongly normalizing} \}$ for $A \in \mathbb{T}$;
-    \item $SC_{{\bf K}(A \to B)} = \{ M : {\bf K}(A \to B) \: | \: \forall f \in SC_{A \to B}, \forall x \in SC_A, \forall N \in SC_{{\bf K}A}, {\bf let \: pure \:} f, x = M, N {\: \bf in \:} f x \in SC_{{\bf K}B} \}$ for $A,B \in \mathbb{T}_{\bf K}$.
+    \item $\forall i \in \{ 1,\dots,n\}, \prod \limits_{i = 1}^{n} SC_{{\bf K}A_i} = \{ \vec{M} = (M_1,\dots,M_n) \:\: | \:\: \forall N \in SC_{B}, FV(N) = \{ x_1,\dots,x_n\} \: \& \: \forall i, x_i \in SC_{A_i} \Rightarrow {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N \in SC_{{\bf K} B}\}$
   \end{itemize}
@@
 
-  2) $A \equiv {\bf K}(B \to C)$
-
-  i) Suppose $M \in SC_{{\bf K}(B \to C)}$. Let $N \in SC_{{\bf K}B}$. So ${\bf let \: pure \:} f, x = M, N {\: \bf in \: } f x \in SC_{{\bf K}C}$.
-
-  So $M$ is strongly normalizing, since ${\bf let \: pure \:} f, x = M, N {\: \bf in \: } f x$ is strongly normalizing.
-
-  ii) Let $M_1 \in SC_{{\bf K}(B \to C)}$ and $M_1 \rightarrow_{\beta} M_2$. Fix $N \in SC_{{\bf K}B}$.
-
-  Then ${\bf let \: pure \:} f, x = M_1, N {\: \bf in \: } f x \in SC_{{\bf K}C}$.
-
-  Hence, ${\bf let \: pure \:} f, x = M_1, N {\: \bf in \: } f x \in SC_{{\bf K}C} \rightarrow_{\beta} {\bf let \: pure \:} f, x = M_2, N {\: \bf in \: } f x$.
-
-  So ${\bf let \: pure \:} f, x = M_2, N {\: \bf in \: } f x \in SC_{{\bf K}C}$. Then $M_2 \in SC_{{\bf K}(B \to C)}$.
-
-  iii) Let $M_2$ be non-introduced, $M_2 \in SC_{{\bf K}(B \to C)}$ and $M_1 \rightarrow_{\beta} M_2$.
-
-  Let $N \in SC_{{\bf K}B}$. So ${\bf let \: pure \:} f, x = M_2, N {\: \bf in \:} f x \in SC_{{\bf K}C}$.
-
-  So ${\bf let \: pure \:} f, x = M_1, N {\: \bf in \:} f x \rightarrow_{\beta} {\bf let \: pure \:} f, x = M_2, N {\: \bf in \:} f x \in SC_{{\bf K}C}$.
-
-  Thus ${\bf let \: pure \:} f, x = M_1, N {\: \bf in \:} f x \in SC_{{\bf K}C}$ by IH, so $M_1 \in SC_{{\bf K}(B \to C)}$.
+2)
+
+i) Suppose $(M_1,\dots,M_n) \in \prod \limits_{i = 1}^{n} SC_{{\bf K}A_i}$.
+
+Let $N \in SC_{B}$, such that $FV(N) = \{ x_1,\dots,x_n \}$ and $\forall i, x_i \in SC_{A_i}$.
+
+So ${\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } N \in SC_{{\bf K}B}$ by IH.
+
+So $\vec{M}$ are strongly normalizing, since ${\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } N$ is strongly normalizing by IH.
+
+ii) Let $\vec{M_1} \in \prod \limits_{i = 1}^{n} SC_{{\bf K}A_i}$ and $\vec{M_1} \rightarrow_{\beta} \vec{M_2}$.
+
+Let $N \in SC_{B}$, such that $FV(N) = \{ x_1,\dots,x_n \}$ and $\forall i, x_i \in SC_{A_i}$.
+
+So ${\bf let \: pure \:} \vec{x} = \vec{M_1} {\: \bf in \:} N \rightarrow_{\beta} {\bf let \: pure \:} \vec{x} = \vec{M_2} {\: \bf in \:} N$
+
+and ${\bf let \: pure \:} \vec{x} = \vec{M_2} {\: \bf in \:} N \in SC_{{\bf K}B}$ by assumption.
+
+So $\vec{M_2} \in \prod \limits_{i = 1}^{n} SC_{{\bf K}A_i}$.
+
+iii) Let $M_2$ be non-introduced, $M_2 \in \prod \limits_{i = 1}^{n} SC_{{\bf K}A_i}$ and $M_1 \rightarrow_{\beta} M_2$.
+
+Let $N \in SC_{B}$. So ${\bf let \: pure \:} \vec{x} = \vec{M_2} {\: \bf in \:} N \in SC_{{\bf K}B}$.
+
+So ${\bf let \: pure \:} \vec{x} = \vec{M_1} {\: \bf in \:} N \rightarrow_{\beta} {\bf let \: pure \:} \vec{x} = \vec{M_2} {\: \bf in \:} \in N$.
+
+Thus ${\bf let \: pure \:} \vec{x} = \vec{M_1} {\: \bf in \:} N\in SC_{{\bf K}B}$ by IH, so $\vec{M_1} \in \prod \limits_{i = 1}^{n} SC_{{\bf K}A_i}$.
 \end{proof}
@@
 
-  We modify and apply Barendregt's technique with term underlying. We will consider the fragment of the grammar for terms without
-  constructors for pairs for simplicity.
+  We modify and apply Barendregt's technique with term underlying. We will consider the grammar without constructors for pairs.
 
@@
   Case with ${\bf let \: pure\:} \vec{x} = \vec{M} {\: \bf in \:} N$ was considered by Kakutani \cite{ModalK} \cite{ModalK1}.
-  Similary, if ${\bf pure \:} M$ is in normal form, so $M$ is in normal form to and has subformula property by hypothesis.
-  So ${\bf pure \:} M$ has subformula property.
+  Similary, if ${\bf pure \:} M$ is in normal form, so $M$ is in normal form to and its subterms are in normal form too by hypothesis.
 \end{proof}
@@
   \bibitem{Abramsky} Samson Abramsky and Nikos Tzevelekos, Introduction to Categories and Categorical Logic
+
+  \bibitem{ModalLa} G. A. Kavvos. The Many Worlds of Modal $\Lambda$--calculi: I. Curry-Howard for Necessity, Possibility and Time
+
+  \bibitem{Cons} Ross Paterson. in Mathematics of Program Construction, Madrid, 2012, Lecture Notes in Computer Science, vol. 7342, pp. 300--323, Springer, 2012.
 \end{thebibliography}

### 2018-02-06 article.tex
added

@@
 
-We will denote $\Gamma \vdash {\bf K} A_1, \dots, \Gamma \vdash {\bf K} A_n$ and $\Gamma \vdash {\bf K} B$ as $\Gamma \vdash {\bf K} \vec{A}$ and $\vec{A} \vdash B$ for brevity.
+We will denote $\Gamma \vdash {\bf K} A_1, \dots, \Gamma \vdash {\bf K} A_n$ and $A_1,\dots,A_n \vdash B$ as $\Gamma \vdash {\bf K} \vec{A}$ and $\vec{A} \vdash B$ for brevity.
 
@@
 
-  At the next step we build the typed lambda-calculus based on NIEL$^{-}_{\land, \to}$ by proof-assingment in
-  rules.
+  At the next step we build the typed lambda-calculus based on the NIEL$^{-}$ with implication and dijunction
+by proof-assingment in rules. Obtained fragment is equivalent to IEL$^{-}$ without axioms for negation and disjunction.
 
@@
 $\begin{array}{lll}
-& \Lambda_{{\bf K}} ::= \mathbb{V} \: | \:  (\lambda \Lambda.\Lambda_{{\bf K}}) \: | \: (\Lambda_{{\bf K}}\Lambda_{{\bf K}}) \: | \: (\Lambda_{{\bf K}} , \Lambda_{{\bf K}}) \: | \: (\pi_1 \Lambda_{{\bf K}}) \: | \: (\pi_2 \Lambda_{{\bf K}}) \: | & \\
-& \quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad ({\bf pure \: } \: \Lambda_{{\bf K}}) \: | \: ({\bf let \:\: pure} \:\: \Lambda_{{\bf K}} = \Lambda_{{\bf K}} \:\: {\bf in} \:\: \Lambda_{{\bf K}})
+& \Lambda_{{\bf K}} ::= \mathbb{V} \: | \:  (\lambda \mathbb{V}.\Lambda_{{\bf K}}) \: | \: (\Lambda_{{\bf K}}\Lambda_{{\bf K}}) \: | \: (\Lambda_{{\bf K}} , \Lambda_{{\bf K}}) \: | \: (\pi_1 \Lambda_{{\bf K}}) \: | \: (\pi_2 \Lambda_{{\bf K}}) \: | & \\
+& \quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad ({\bf pure \: } \: \Lambda_{{\bf K}}) \: | \: ({\bf let \:\: pure} \:\: \mathbb{V}^{*} = \Lambda_{{\bf K}}^{*} \:\: {\bf in} \:\: \Lambda_{{\bf K}})
 \end{array}$
@@
   \end{defin}
+
+Where $\mathbb{V}^{*}$ and $\Lambda_{{\bf K}}^{*}$ denote the set of finite sequences of variables $\mathbb{V}^0 \times \mathbb{V}^1 \times \dots$
+and the set of finite sequences of terms $\Lambda_{{\bf K}}^0 \times \Lambda_{{\bf K}}^1 \times \dots$.
+Note that the sequence of variables $\vec{x}$ and the sequence of terms $\vec{M}$ should have the same length.
+Otherwise, term is not well-formed.
 
@@
 
-  $\Gamma \vdash \vec{M} : {\bf K} \vec{A}$ is a syntax sugar for the sequence $\Gamma \vdash M_1 : {\bf K}A_1,\dots,\Gamma \vdash M_n : {\bf K}A_n$ and $\vec{x} : \vec{A} \vdash N : B$ is an abbrevation for $x_1 : A_1, \dots, x_n : A_n \vdash N : B$.
+$\Gamma \vdash \vec{M} : {\bf K} \vec{A}$ is a syntax sugar for the sequence $\Gamma \vdash M_1 : {\bf K}A_1,\dots,\Gamma \vdash M_n : {\bf K}A_n$ and $\vec{x} : \vec{A} \vdash N : B$ is a short form for $x_1 : A_1, \dots, x_n : A_n \vdash N : B$.
 ${\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } N$ is a simultaneous local binding in $N$.
-We use abbreviated notation instead of ${\bf let \: pure\:} x_1,\dots,x_n = M_1,\dots,M_n {\: \bf in \:} N$.
-
-  In fact, our calculus is the extention of typed lambda calculus for {\bf IK} with typing rule appropriate to co-reflection.
+We use this short form instead of ${\bf let \: pure\:} x_1,\dots,x_n = M_1,\dots,M_n {\: \bf in \:} N$.
+
+In fact, our calculus is the extention of typed lambda calculus for {\bf IK} by ${\bf K}_I$-rule that is appropriate to co-reflection.
 
@@
 
-  \begin{defin} Set $FV(M)$ of free variables for arbitrary term $M$:
+  \begin{defin} The set $FV(M)$ of free variables for a term $M$:
 
@@
 
-  7) $FV({\bf let \:\: pure} \:\: \vec{x} = \vec{M} \:\: {\bf in} \:\: N) = \bigcup \limits_{i = 1}^n FV(M), \text{where $n = |\vec{M}|$}$.
+  7) $FV({\bf let \: pure} \: \vec{x} = \vec{M} \:\: {\bf in} \:\: N) = \bigcup \limits_{i = 1}^n FV(M), \text{where $n = |\vec{M}|$}$.
   \end{defin}
@@
 
-  3) $(\lambda x. M) [x := N] = \lambda x. M [x := N]$;
+  3) $(\lambda x. M) [x := N] = \lambda x. M [y := N]$, $y \in FV(M)$;
 
@@
 
+Substitution and free variable for terms of kind ${\bf let \: pure}$ are defined similary to \cite{ModalK}.
+
+Because the local binding
+
   \begin{defin} Type substituition
 
-    The substituition of type $C$ for type variable $B$ in type $A$ inductively defined as follows:
+    The substituition of type $C$ for type variable $B$ in type $A$ is defined inductively:
 
@@
 
-    3) $({\bf K}A)[B := C] = {\bf K}(A [B := C])$.
-
-    4) Let $\Gamma$ be the context, then $\Gamma [B := C] = \{ x : (A[B := C]) \: | \: x : A \in \Gamma \}$
+    3) $({\bf K}A)[B := C] = {\bf K}(A [B := C])$;
+
+    4) Let $\Gamma$ be the context, then $\Gamma [B := C] = \{ x : (A[B := C]) \: | \: x : A \in \Gamma \}$.
   \end{defin}
@@
 
-  6) ${\bf let \: pure \:} \underline{\quad} = \underline {\quad} {\: \bf in \:} M \rightarrow_{\beta} {\bf pure \:} M$
+  6) ${\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} M \rightarrow_{\beta} {\bf pure \:} M$, where \underline{\quad} is an empty sequence of terms.
 
@@
 \begin{proof}
-  Induction on the structure of ${\bf pure \:} M$.
-\end{proof}
-
-\begin{lemma} Basic lemmas .
+  Straightforwardly.
+\end{proof}
+
+\begin{lemma} Basic lemmas.
 
 \begin{itemize}
-  \item Let $\Gamma \vdash M : A$ and $\Gamma \subseteq \Delta$, then $\Delta \vdash M : A$;
-  \item Let $\Gamma \vdash M : A$, then $\Delta \vdash M : A$, where $\Delta = \{ x_i : A_i \: | \: (x_i : A_i) \in \Gamma \: \& \: x_i \in FV(M) \}$
-  \item Let $\Gamma, x : A \vdash M : B$ and $\Gamma \vdash N : A$, then $\Gamma \vdash M [x := N] : B$.
-  \item Let $\Gamma \vdash M : A$, then $\Gamma [B := C] \vdash M : (A [B := C])$
+  \item If $\Gamma \vdash M : A$ and $\Gamma \subseteq \Delta$, then $\Delta \vdash M : A$;
+  \item If $\Gamma \vdash M : A$, then $\Delta \vdash M : A$, where $\Delta = \{ x_i : A_i \: | \: (x_i : A_i) \in \Gamma \: \& \: x_i \in FV(M) \}$
+  \item If $\Gamma, x : A \vdash M : B$ and $\Gamma \vdash N : A$, then $\Gamma \vdash M [x := N] : B$.
+  \item If $\Gamma \vdash M : A$, then $\Gamma [B := C] \vdash M : (A [B := C])$;
+  \item If $\Gamma \vdash M : A$ and $\Delta$ is a permutation of $\Gamma$, then $\Delta \vdash M : A$.
 \end{itemize}
@@
 
-  i--iv) Induction on $\Gamma \vdash M : A$.
-
+1) The derivation ends in
+
+\begin{prooftree}
+  \AxiomC{$\Gamma \vdash \vec{M} : {\bf K} \vec{A}$}
+  \AxiomC{$\vec{x} : \vec{A} \vdash N : B$}
+  \RightLabel{$\text{let}_{{\bf K}}$}
+  \BinaryInfC{$\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } N : {\bf K}B$}
+\end{prooftree}
+
+By IH $\Delta \vdash \vec{M} : {\bf K} \vec{A}$, so $\Delta \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } N : {\bf K}B$.
+
+  \vspace{\baselineskip}
+
+2) Similary.
+
+  \vspace{\baselineskip}
+
+3) The derivation ends in
+
+\begin{prooftree}
+  \AxiomC{$\Gamma, z : C \vdash \vec{M} : {\bf K} \vec{A}$}
+  \AxiomC{$\vec{x} : \vec{A} \vdash N : B$}
+  \RightLabel{$\text{let}_{{\bf K}}$}
+  \BinaryInfC{$\Gamma, z : C \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } N : {\bf K}B$}
+\end{prooftree}
+
+Let $\Gamma \vdash P : C$.
+
+By IH, $\Gamma \vdash \vec{M} [z := P] : {\bf K} \vec{A}$. So $\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} [z := P] {\: \bf in \: } N : {\bf K}B$
+
+  \vspace{\baselineskip}
+
+4-5) Similary.
+
+\end{proof}
+
+\begin{prop} Simmetricity of local binding
+
+  If $\Gamma \vdash {\bf let \: pure \:} \vec{x}, \vec{y} = \vec{M}, \vec{N} {\: \bf in \:} P$, then
+  $\Gamma \vdash {\bf let \: pure \:} \vec{y}, \vec{x} = \vec{N}, \vec{M} {\: \bf in \:} P$
+\end{prop}
+
+\begin{proof}
+
+  Follows from the part 5 of the previous lemma.
 \end{proof}
@@
 
-  Let $\Gamma \vdash M : A$ and $M \twoheadrightarrow_{\beta \eta} N$, then $\Gamma \vdash N : A$
+  If $\Gamma \vdash M : A$ and $M \twoheadrightarrow_{\beta \eta} N$, then $\Gamma \vdash N : A$
 
@@
 \begin{proof}
+
+Induction on the derivation $\Gamma \vdash M : A$ and on the generation of $\rightarrow_{\beta \eta}$.
+
 For cases with application, abstraction and pairs see \cite{Morten} \cite{Pierce}.
 
-1)
-  Let $\Gamma \vdash {\bf let \: pure \:} \vec{x}, y, \vec{z} = \vec{M}, {\bf let \: pure \: } \vec{w} = \vec{N} {\bf \: in \: } Q, \vec{P}  {\: \bf in \:} R : {\bf K}B$,
-  then $\Gamma \vdash {\bf let \: pure \:} \vec{x}, \vec{w}, \vec{z} = \vec{M}, \vec{N}, \vec{P} \: { \bf in } \: R [y := Q] : {\bf K}B$
-
-2) Let $\Gamma \vdash {\bf let \: pure \:} x = M {\: \bf in \: } x : {\bf K}A$, then $\Gamma \vdash M : {\bf K}A$.
+1) If $\Gamma \vdash {\bf let \: pure \:} \vec{x}, y, \vec{z} = \vec{M}, {\bf let \: pure \: } \vec{w} = \vec{N} {\bf \: in \: } Q, \vec{P}  {\: \bf in \:} R : {\bf K}B$,
+  then $\Gamma \vdash {\bf let \: pure \:} \vec{x}, \vec{w}, \vec{z} = \vec{M}, \vec{N}, \vec{P} \: { \bf in } \: R [y := Q] : {\bf K}B$ by rule 4).
+
+2) Let $\Gamma \vdash {\bf let \: pure \:} x = M {\: \bf in \: } x : {\bf K}A$, then $\Gamma \vdash M : {\bf K}A$ by rule 9).
 
@@
 
-3) If the derivation ends in
+3) The derivation ends in
 
@@
 
-So $\Gamma \vdash \vec{M} : \vec{A}$ by generation and $\Gamma \vdash N [\vec{x} := \vec{M}] : B$ by weakening and substitution.
-
-Then we can transform this into the next derivation:
+So $\Gamma \vdash \vec{M} : \vec{A}$ by Lemma 4 and $\Gamma \vdash N [\vec{x} := \vec{M}] : B$ by Lemma 4, part 3.
+
+Then we can transform this into the following derivation:
 
@@
 
-4) If the derivation ends in
+4) The derivation ends in
 
@@
 
-We modify and apply Tait's technique of logical relation for modalities.
-For strong normalization proof with Tait's method for simply typed lambda calculus see \cite{Pierce}.
+We modify and apply Tait's technique of logical relation for modalities. See \cite{Pierce} \cite{Troelstra}.
+
+We treat only modal cases below.
 
@@
     \item $SC_A = \{ M : A \: | \: M \text{ is strongly normalizing} \}$ for $A \in \mathbb{T}$;
-    \item $SC_{A \to B} = \{ M : A \to B \: | \: \forall N \in SC_A, M N \in SC_B \}$, for $A,B \in \mathbb{T}_{\bf K}$
+    \item $SC_{A \to B} = \{ M : A \to B \: | \: \forall N \in SC_A, M N \in SC_B \}$, for $A,B \in \mathbb{T}_{\bf K}$ for $A, B \in \mathbb{T}_{{\bf K}}$;
     \item $SC_{{\bf K}A} = \{ M : {\bf K}A \: | \: M \text{ is strongly normalizing} \}$ for $A \in \mathbb{T}$;
@@
 
+\begin{defin}
+  A term $M$ is neutral, if it has one of the following forms:
+  \begin{itemize}
+    \item $M N$;
+    \item If $M$ is neutral, then ${\bf pure \:} M$ is neutral;
+    \item If $\vec{M}$ and $N$ are neutral, then ${\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: N}$ is neutral.
+    $\vec{x}$ is a sequence of free variables of a term $N$.
+  \end{itemize}
+\end{defin}
+
 \begin{lemma}
@@
 \begin{itemize}
-\item If $M \in SC_A$, then $M$ is strongly normalizing;
-\item Let $M \in SC_A$ and $M \rightarrow_{\beta} N$, then $N \in SC_A$;
-\item Let $N$ is non-introduced, $N \in SC_A$. Then, if $M \rightarrow_{\beta} N$, then $M \in SC_A$;
+\item If $M \in SC_A$ and $A \in \mathbb{T}_{{\bf K}}$, then $M$ is strongly normalizing;
+\item If $M \in SC_A$, $A \in \mathbb{T}_{{\bf K}}$ and $M \rightarrow_{\beta} N$, then $N \in SC_A$;
+\item If $N$ is neutral, $N \in SC_A$. Then, if $M \rightarrow_{\beta} N$, then $M \in SC_A$;
 \end{itemize}
@@
 
-  i) Follows from the definition;
-
-  ii) Immediately;
-
-  iii) Let $N$ is non-introduced and $N \in SC_A$, such that $M \rightarrow_{\beta} N$.
-  Any reduction path $M \rightarrow_{\beta} \dots$ passes through $M \rightarrow_{\beta} N$.
-
-  $N$ is strongly normalizing, so $M$ too.
+i-ii-iii) Immediately.
 
@@
 
-i) Suppose $(M_1,\dots,M_n) \in \prod \limits_{i = 1}^{n} SC_{{\bf K}A_i}$.
+i) Suppose $\vec{M} = (M_1,\dots,M_n) \in \prod \limits_{i = 1}^{n} SC_{{\bf K}A_i}$.
+
+Let $N \in SC_{B}$, such that $FV(N) = \{ x_1,\dots, x_n \}$ and $\forall i, x_i \in SC_{A_i}$.
+
+So ${\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } N \in SC_{{\bf K}B}$ by IH.
+
+So $\vec{M}$ are strongly normalizing, since ${\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } N$ is strongly normalizing by IH.
+
+\vspace{\baselineskip}
+
+ii) Let $\vec{M_1} \in \prod \limits_{i = 1}^{n} SC_{{\bf K}A_i}$ and $\vec{M_1} \rightarrow_{\beta} \vec{M_2}$.
 
@@
 
-So ${\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } N \in SC_{{\bf K}B}$ by IH.
-
-So $\vec{M}$ are strongly normalizing, since ${\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } N$ is strongly normalizing by IH.
-
-ii) Let $\vec{M_1} \in \prod \limits_{i = 1}^{n} SC_{{\bf K}A_i}$ and $\vec{M_1} \rightarrow_{\beta} \vec{M_2}$.
-
-Let $N \in SC_{B}$, such that $FV(N) = \{ x_1,\dots,x_n \}$ and $\forall i, x_i \in SC_{A_i}$.
-
 So ${\bf let \: pure \:} \vec{x} = \vec{M_1} {\: \bf in \:} N \rightarrow_{\beta} {\bf let \: pure \:} \vec{x} = \vec{M_2} {\: \bf in \:} N$
@@
 
-iii) Let $M_2$ be non-introduced, $M_2 \in \prod \limits_{i = 1}^{n} SC_{{\bf K}A_i}$ and $M_1 \rightarrow_{\beta} M_2$.
-
-Let $N \in SC_{B}$. So ${\bf let \: pure \:} \vec{x} = \vec{M_2} {\: \bf in \:} N \in SC_{{\bf K}B}$.
-
-So ${\bf let \: pure \:} \vec{x} = \vec{M_1} {\: \bf in \:} N \rightarrow_{\beta} {\bf let \: pure \:} \vec{x} = \vec{M_2} {\: \bf in \:} \in N$.
-
-Thus ${\bf let \: pure \:} \vec{x} = \vec{M_1} {\: \bf in \:} N\in SC_{{\bf K}B}$ by IH, so $\vec{M_1} \in \prod \limits_{i = 1}^{n} SC_{{\bf K}A_i}$.
+iii) Let $M_2$ be neutral, $M_2 \in \prod \limits_{i = 1}^{n} SC_{{\bf K}A_i}$ and $M_1 \rightarrow_{\beta} M_2$.
+
+Let $N \in SC_{B}$, , such that $FV(N) = \{ x_1,\dots,x_n \}$ and $\forall i, x_i \in SC_{A_i}$.
+
+So ${\bf let \: pure \:} \vec{x} = \vec{M_2} {\: \bf in \:} N \in SC_{{\bf K}B}$.
+
+Thus ${\bf let \: pure \:} \vec{x} = \vec{M_1} {\: \bf in \:} N \rightarrow_{\beta} {\bf let \: pure \:} \vec{x} = \vec{M_2} {\: \bf in \:} \in N$.
+
+Hence ${\bf let \: pure \:} \vec{x} = \vec{M_1} {\: \bf in \:} N\in SC_{{\bf K}B}$ by IH, so $\vec{M_1} \in \prod \limits_{i = 1}^{n} SC_{{\bf K}A_i}$.
 \end{proof}
@@
 
-  1) Let the derivation ends in:
+Induction on the derivation of $x_1 : A_1,\dots, x_n : A_n \vdash M : A$.
+
+  1) The derivation ends in:
 
@@
 
-  2) Let the derivation ends in:
+  2) The derivation ends in:
 \begin{prooftree}
@@
     \item $|{\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N| = {\bf let \: pure \:} \vec{x} = \vec{|M|} {\: \bf in \:} |N|$;
-    \item $|(\lambda_i x. M) N| = (\lambda x. M) N$
+    \item $|(\lambda_i x. M) N| = (\lambda x. |M|) |N|$
   \end{itemize}
@@
   Case with ${\bf let \: pure\:} \vec{x} = \vec{M} {\: \bf in \:} N$ was considered by Kakutani \cite{ModalK} \cite{ModalK1}.
-  Similary, if ${\bf pure \:} M$ is in normal form, so $M$ is in normal form to and its subterms are in normal form too by hypothesis.
+
+  If ${\bf pure \:} M$ is in normal form, so $M$ is in normal form and its subterms are in normal form too by hypothesis.
+
+  Thus if ${\bf pure \:} M$ is in normal form, then all its subterms are in normal form too.
 \end{proof}
@@
 
-2) $\ast_{A,B} \circ (\eta_A \otimes \eta_B) = \eta_{A \otimes B}$;
+2) $\ast_{A,B} \circ (\eta_A \otimes \eta_B) = \eta_{A \otimes B}$, i.e. the following diagram commutes:
+
+\xymatrix
+{
+&&& A \otimes B \ar[rr]^{\eta_A \otimes \eta_B} \ar[drr]_{\eta_{A \otimes B}} && \mathcal{K}A \otimes \mathcal{K}B \ar[d]^{\ast_{A,B}} \\
+&&&&& \mathcal{K}(A \otimes B)
+}
 

### 2018-02-07 article.tex
added

@@
 
-Where $\mathbb{V}^{*}$ and $\Lambda_{{\bf K}}^{*}$ denote the set of finite sequences of variables $\mathbb{V}^0 \times \mathbb{V}^1 \times \dots$
-and the set of finite sequences of terms $\Lambda_{{\bf K}}^0 \times \Lambda_{{\bf K}}^1 \times \dots$.
+Where $\mathbb{V}^{*}$ and $\Lambda_{{\bf K}}^{*}$ denote the set of finite sequences of variables $\bigcup \limits_{i=0}^{\infty} \mathbb{V}_i$
+and the set of finite sequences of terms $\bigcup \limits_{i = 0}^{\infty} {\Lambda_{{\bf K}}}_i $.
 Note that the sequence of variables $\vec{x}$ and the sequence of terms $\vec{M}$ should have the same length.
@@
 
-Because the local binding
-
   \begin{defin} Type substituition
@@
 By default we use call-by-name evaluation strategy.
+
+Also we accept the following equality, so called symmetricity of local binding:
+
+\begin{defin} Symmetricity of local binding
+
+  ${\bf let \: pure \:} \vec{x}, \vec{y} = \vec{M}, \vec{N} {\: \bf in \:} P \equiv {\bf let \: pure \:} \vec{y}, \vec{x} = \vec{N}, \vec{M} {\: \bf in \:} P$
+
+\end{defin}
+
+This commuting relation denotes that we identify terms of kind {\bf let pure} if they are differ in the order of binding.
 
@@
 
-\begin{prop} Simmetricity of local binding
-
-  If $\Gamma \vdash {\bf let \: pure \:} \vec{x}, \vec{y} = \vec{M}, \vec{N} {\: \bf in \:} P$, then
-  $\Gamma \vdash {\bf let \: pure \:} \vec{y}, \vec{x} = \vec{N}, \vec{M} {\: \bf in \:} P$
+\begin{prop}
+
+  If $\Gamma \vdash {\bf let \: pure \:} \vec{x}, \vec{y} = \vec{M}, \vec{N} {\: \bf in \:} P : {\bf K}A$, then
+  $\Gamma \vdash {\bf let \: pure \:} \vec{y}, \vec{x} = \vec{N}, \vec{M} {\: \bf in \:} P : {\bf K}A$.
 \end{prop}
@@
 
-\begin{defin} Monoidal functor
+\begin{defin} Lax monoidal functor
 
@@
 
-  A monoidal functor $\mathcal{F} : \langle \mathcal{C}, \otimes_1, \mathds{1} \rangle \to \langle \mathcal{D}, \otimes_2, \mathds{1}' \rangle$ is a functor
+  A lax monoidal functor $\mathcal{F} : \langle \mathcal{C}, \otimes_1, \mathds{1} \rangle \to \langle \mathcal{D}, \otimes_2, \mathds{1}' \rangle$ is a functor
   $\mathcal{F} : \mathcal{C} \to \mathcal{D}$ with additional natural transformations, which satisfy the well-known conditions described in \cite{ElKelly}:
@@
   An applicative functor is a triple $\langle \mathcal{C}, \mathcal{K}, \eta \rangle$,
-where $\mathcal{C}$ is a symmetric monoidal category, $\mathcal{K}$ is a monoidal endofunctor and $\eta : Id_{\mathcal{C}} \Rightarrow \mathcal{K}$ is a natural transformation (similar to unit in monad), such that:
+where $\mathcal{C}$ is a symmetric monoidal category, $\mathcal{K}$ is a lax monoidal endofunctor and $\eta : Id_{\mathcal{C}} \Rightarrow \mathcal{K}$ is a natural transformation (similar to unit in monad), such that:
 
@@
 
+By default we will consider a strong functor on cartesian closed category with $\eta$ \cite{Cons}.
+
 \subsection{Soundness and completeness}
@@
 
-$[\![A \to B]\!] := [\![A]\!] \to [\![B]\!]$;
+$[\![A \to B]\!] := [\![B]\!]^{[\![A]\!]}$;
 
@@
 
+\begin{prop}
+$ $
+
+  $[\![\Gamma \vdash {\bf let \: pure \:} \vec{x}, \vec{y} = \vec{M}, \vec{N} {\: \bf in \:} P : {\bf K} C]\!] = [\![\Gamma \vdash {\bf let \: pure \:} \vec{y}, \vec{x} = \vec{N}, \vec{M} {\: \bf in \:} P : {\bf K} C]\!]$
+\end{prop}
+
+\begin{proof}
+  $ $
+
+  Follows from the commutativity of the following diagram:
+
+\xymatrix{
+[\![\Gamma]\!] \ar[rr]^{\langle g, f \rangle} \ar[d]_{\langle f, g \rangle} && \mathcal{K}[\![B]\!] \times \mathcal{K}[\![A]\!] \ar[rr]^{\ast_{[\![B]\!], [\![A]\!]}}
+\ar[d]_{\ast_{[\![A]\!], [\![B]\!]} \circ \langle \pi_2, \pi_1 \rangle} && \mathcal{K}([\![B]\!] \times [\![A\!]) \ar[d]^{\mathcal{K}{[\![N]\!]} \circ \mathcal{K}(\langle \pi_2, \pi_1 \rangle)} \\
+\mathcal{K}[\![A\!] \times \mathcal{K}[\![B]\!] \ar[rr]_{\ast_{[\![A]\!], [\![B]\!]}} && \mathcal{K}([\![A]\!] \times [\![B]\!]) \ar[rr]_{[\![N]\!]}&& \mathcal{K}[\![C]\!]
+}
+
+\end{proof}
+
 \end{proof}
@@
 
-We will denote equivalence class as $[x, M]_{A, B} = \{ (y, N) \: | \: (x, M) \sim_{A, B} (y, N) \}$ (we will drop indeces below).
+We will denote equivalence class as $[x, M]_{A, B} = \{ (y, N) \: | \: (x, M) \sim_{A, B} (y, N) \}$ (we will drop indices below).
 

### 2018-02-09 article.tex
modified

@@
 
-Where $\mathbb{V}^{*}$ and $\Lambda_{{\bf K}}^{*}$ denote the set of finite sequences of variables $\bigcup \limits_{i=0}^{\infty} \mathbb{V}_i$
-and the set of finite sequences of terms $\bigcup \limits_{i = 0}^{\infty} {\Lambda_{{\bf K}}}_i $.
+Where $\mathbb{V}^{*}$ and $\Lambda_{{\bf K}}^{*}$ denote the set of finite sequences of variables $\bigcup \limits_{i=0}^{\infty} \mathbb{V}^i$
+and the set of finite sequences of terms $\bigcup \limits_{i = 0}^{\infty} {\Lambda_{{\bf K}}}^i $.
 Note that the sequence of variables $\vec{x}$ and the sequence of terms $\vec{M}$ should have the same length.

### 2018-02-10 article.tex
updated definitions and lemmas

@@
 
-Also we accept the following equality, so called symmetricity of local binding:
-
-\begin{defin} Symmetricity of local binding
-
-  ${\bf let \: pure \:} \vec{x}, \vec{y} = \vec{M}, \vec{N} {\: \bf in \:} P \equiv {\bf let \: pure \:} \vec{y}, \vec{x} = \vec{N}, \vec{M} {\: \bf in \:} P$
-
-\end{defin}
-
 This commuting relation denotes that we identify terms of kind {\bf let pure} if they are differ in the order of binding.
@@
 
-\end{proof}
-
-\begin{prop}
-
-  If $\Gamma \vdash {\bf let \: pure \:} \vec{x}, \vec{y} = \vec{M}, \vec{N} {\: \bf in \:} P : {\bf K}A$, then
-  $\Gamma \vdash {\bf let \: pure \:} \vec{y}, \vec{x} = \vec{N}, \vec{M} {\: \bf in \:} P : {\bf K}A$.
-\end{prop}
-
-\begin{proof}
-
-  Follows from the part 5 of the previous lemma.
 \end{proof}
@@
 
+\begin{defin}
+
+Strong endofunctor on a monoidal category is a with a natural transformation so called tensorial strength
+
+\begin{equation}
+  \tau_{A, B} : A \otimes \mathcal{K}B \rightarrow \mathcal{K}(A \otimes B)
+\end{equation}
+
+making two diagrams commute:
+
+\xymatrix{
+(A \otimes B) \otimes \mathcal{K}C \ar[d]_{\alpha_{A, B, \mathcal{K}C}}\ar[rrrr]^{\tau_{A \otimes B, C}} &&&& \mathcal{K}((A \otimes B) \otimes C) \ar[d]^{\mathcal{K}(\alpha_{A,B,C})}\\
+A \otimes (B \otimes \mathcal{K}C) \ar[rr]_{id_A \otimes \tau_{B,C}} && A \otimes \mathcal{K}(B \otimes C) \ar[rr]_{\quad \tau_{A, (B \otimes C)}} && \mathcal{K}(A \otimes (B \otimes C)) \\
+}
+
+\xymatrix{
+&&& \mathds{1} \otimes \mathcal{K}A \ar[drr]_{\rho_{\mathcal{K}A}} \ar[rr]^{\mu_{\mathds{1}, A}} && \mathcal{K}(\mathds{1} \otimes A) \ar[d]^{\mathcal{K}(\rho_A)}\\
+&&&&& \mathcal{K}A
+}
+\end{defin}
+
 \begin{defin} Applicative functor
@@
 
-3) Weak commutativity condition:
-
-\xymatrix{
-&& A \otimes \mathcal{K}B \ar[d]_{\sigma_{A, \mathcal{K}B}} \ar[rr]^{\eta_{A} \otimes id_{\mathcal{K}B}} && \mathcal{K}A \otimes \mathcal{K}B \ar[rr]^{\ast_{A,B}} && \mathcal{K}(A \otimes B) \ar[d]^{\mathcal{K}(\sigma_{A,B})}\\
-&& \mathcal{K}B \otimes A \ar[rr]_{id_{\mathcal{K}B} \otimes \eta_{A}} && \mathcal{K}B \otimes \mathcal{K}A \ar[rr]_{\ast_{B,A}} && \mathcal{K}(B \otimes A)
-}
+3) $\tau_{A, B} = \ast_{A, B} \circ \eta_{A} \otimes id_{\mathcal{K}B}$.
 \end{defin}
@@
 
-\begin{prop}
-$ $
-
-  $[\![\Gamma \vdash {\bf let \: pure \:} \vec{x}, \vec{y} = \vec{M}, \vec{N} {\: \bf in \:} P : {\bf K} C]\!] = [\![\Gamma \vdash {\bf let \: pure \:} \vec{y}, \vec{x} = \vec{N}, \vec{M} {\: \bf in \:} P : {\bf K} C]\!]$
-\end{prop}
-
-\begin{proof}
-  $ $
-
-  Follows from the commutativity of the following diagram:
-
-\xymatrix{
-[\![\Gamma]\!] \ar[rr]^{\langle g, f \rangle} \ar[d]_{\langle f, g \rangle} && \mathcal{K}[\![B]\!] \times \mathcal{K}[\![A]\!] \ar[rr]^{\ast_{[\![B]\!], [\![A]\!]}}
-\ar[d]_{\ast_{[\![A]\!], [\![B]\!]} \circ \langle \pi_2, \pi_1 \rangle} && \mathcal{K}([\![B]\!] \times [\![A\!]) \ar[d]^{\mathcal{K}{[\![N]\!]} \circ \mathcal{K}(\langle \pi_2, \pi_1 \rangle)} \\
-\mathcal{K}[\![A\!] \times \mathcal{K}[\![B]\!] \ar[rr]_{\ast_{[\![A]\!], [\![B]\!]}} && \mathcal{K}([\![A]\!] \times [\![B]\!]) \ar[rr]_{[\![N]\!]}&& \mathcal{K}[\![C]\!]
-}
-
-\end{proof}
-
 \end{proof}
@@
 
-\begin{lemma} Weak commutativity.
-  $ $
-
-  $\begin{array}{lll}
-  & \text{fmap } ([p, \langle \pi_2 p, \pi_1 p \rangle]) \circ \tau_{\hat{A},\hat{B}} = & \\
-  & \ast_{\hat{B}, \hat{A}} \circ [q, \langle \pi_1 q, {\bf pure \:} (\pi_2 q) \rangle] \circ [p, \langle \pi_2 p, \pi_1 p \rangle]&
-  \end{array}$
-\end{lemma}
-
-\begin{proof}
-  $ $
-
-  $\begin{array}{lll}
-  & \text{fmap } ([r, \langle \pi_2 r, \pi_1 r \rangle]) \circ \tau_{\hat{A},\hat{B}} = & \\
-  &\text{By the definition of $\tau$}& \\
-  & \text{fmap } ([r, \langle \pi_2 r, \pi_1 r \rangle]) \circ [p, {\bf let \: pure \:} x, y = {\bf pure \:} (\pi_1 p), \pi_2 p {\: \bf in \:} \langle x, y \rangle] = & \\
-  &\text{By the definition of fmap}& \\
-  & [q, {\bf let \: pure \:} r = q {\: \bf in \:} \langle \pi_2 r, \pi_1 r \rangle] \circ [p, {\bf let \: pure \:} x, y = {\bf pure \:} (\pi_1 p), \pi_2 p {\: \bf in \:} \langle x, y \rangle] = & \\
-  &\text{Composition}& \\
-  & [p, {\bf let \: pure \:} r = q {\: \bf in \:} \langle \pi_2 r, \pi_1 r \rangle [q := {\bf let \: pure \:} x, y = {\bf pure \:} (\pi_1 p), \pi_2 p {\: \bf in \:} \langle x, y \rangle]] = &\\
-  &\text{By $\beta$-reduction rules}& \\
-  & [p, {\bf let \: pure \:} r = ({\bf let \: pure \:} x, y = {\bf pure \:} (\pi_1 p), \pi_2 p {\: \bf in \:} \langle x, y \rangle) {\: \bf in \:} \langle \pi_2 r, \pi_1 r \rangle] = & \\
-  &\text{By $\beta$-reduction rules}& \\
-  & [p, {\bf let \: pure \:} x, y = {\bf pure \:} (\pi_1 p), \pi_2 p {\: \bf in \:} \langle \pi_2 r, \pi_1 r \rangle [r := \langle x, y \rangle]] = & \\
-  &\text{By substitution}& \\
-  & [p, {\bf let \: pure \:} x, y = {\bf pure \:} (\pi_1 p), \pi_2 p {\: \bf in \:} \langle \pi_2 \langle x, y \rangle, \pi_1 \langle x, y \rangle \rangle]= & \\
-  &\text{By $\beta$-reduction rules}& \\
-  & [p, {\bf let \: pure \:} x, y = {\bf pure \:} (\pi_1 p), \pi_2 p {\: \bf in \:} \langle y, x \rangle ] = & \\
-  & & \\
-  &\text{On the other hand}&\\
-  &\ast_{\hat{B}, \hat{A}} \circ [q, \langle \pi_1 q, {\bf pure \:} (\pi_2 q) \rangle] \circ [p, \langle \pi_2 p, \pi_1 p \rangle] = & \\
-  &\text{By the definition of $\ast$}& \\
-  &[r, {\bf let \: pure \:} y,x = \pi_1 r, \pi_2 r {\: \bf in \:} \langle y, x \rangle] \circ [q, \langle \pi_1 q, {\bf pure \:} (\pi_2 q) \rangle] \circ [p, \langle \pi_2 p, \pi_1 p \rangle] = & \\
-  &\text{Composition}& \\
-  &[r, {\bf let \: pure \:} y,x = \pi_1 r, \pi_2 r {\: \bf in \:} \langle y, x \rangle] \circ [p, \langle \pi_1 q, {\bf pure \:} (\pi_2 q) \rangle [q := \langle \pi_2 p, \pi_1 p  \rangle]] = & \\
-  &\text{By subsitution and by  $\beta$-reduction rules}& \\
-  &[r, {\bf let \: pure \:} y,x = \pi_1 r, \pi_2 r {\: \bf in \:} \langle y, x \rangle] \circ [p, \langle \pi_2 p, {\bf pure \:} (\pi_1 p) \rangle ]] = & \\
-  &\text{Composition}& \\
-  &[p, {\bf let \: pure \:} y,x = \pi_1 r, \pi_2 r {\: \bf in \:} \langle y, x \rangle [r := \langle \pi_2 p, {\bf pure \:} (\pi_1 p) \rangle]] = & \\
-  &\text{By subsitution and by  $\beta$-reduction rules}& \\
-  &[p, {\bf let \: pure \:} y,x = \pi_2 p, {\bf pure \:} (\pi_1 p) {\: \bf in \:} \langle y, x \rangle] = & \\
-  &\text{By symmetricity of assingment}&\\
-  &[p, {\bf let \: pure \:} x,y = {\bf pure \:} (\pi_1 p), \pi_2 p {\: \bf in \:} \langle y, x \rangle]&
-  \end{array}$
+\begin{lemma} Coherence for tensorial strength
+
+  \begin{itemize}
+    \item $\text{fmap } \alpha_{\hat{A},\hat{B},\hat{C}} \circ \tau_{\hat{A} \times \hat{B}, \hat{C}} = \tau_{\hat{A}, \hat{B} \times \hat{C}} \circ (id_{\hat{A}} \times \tau_{\hat{B}, \hat{C}}) \circ \alpha_{\hat{A}, \hat{B}, {\bf K}\hat{C}}$;
+    \item ${\bf K}(\rho_{\hat{A}}) \circ \tau_{\mathds{1}, \hat{A}} = \rho_{{\bf K}\hat{A}}$.
+  \end{itemize}
+  where $\alpha_{\hat{A},\hat{B},\hat{C}} = [p, \langle \pi_1 (\pi_1 p), \langle \pi_1 (\pi_2 p), \pi_2 p \rangle \rangle]$ and
+  $\rho = \pi_2$.
+\end{lemma}
+
+\begin{proof}
+  $ $
+
+Routine checking.
 \end{proof}

### 2018-02-10 article.tex
update

@@
   An applicative functor is a triple $\langle \mathcal{C}, \mathcal{K}, \eta \rangle$,
-where $\mathcal{C}$ is a symmetric monoidal category, $\mathcal{K}$ is a lax monoidal endofunctor and $\eta : Id_{\mathcal{C}} \Rightarrow \mathcal{K}$ is a natural transformation (similar to unit in monad), such that:
+where $\mathcal{C}$ is a monoidal category, $\mathcal{K}$ is a strong lax monoidal endofunctor and $\eta : Id_{\mathcal{C}} \Rightarrow \mathcal{K}$ is a natural transformation (similar to unit in monad), such that:
 
@@
 
-Routine checking.
+Simple, but routine checking.
 \end{proof}

### 2018-02-10 article.tex
fixes

@@
 
-By default we will consider a strong functor on cartesian closed category with $\eta$ \cite{Cons}.
+By default we will consider some strong lax monoidal functor on cartesian closed category.
+
+Note that lax monoidal functor on a cartesian closed category is a special case of an applicative functor \cite{Cons}, but
+we will define $\eta$ explicitly for simplicity.
 

### 2018-02-10 article.tex
added fixes

@@
   \end{itemize}
-  where $\alpha_{\hat{A},\hat{B},\hat{C}} = [p, \langle \pi_1 (\pi_1 p), \langle \pi_1 (\pi_2 p), \pi_2 p \rangle \rangle]$ and
+  where $\alpha_{\hat{A},\hat{B},\hat{C}} = [p, \langle \pi_1 (\pi_1 p), \langle \pi_2 (\pi_1 p), \pi_2 p \rangle \rangle]$ and
   $\rho = \pi_2$.
@@
 
-Simple, but routine checking.
+1)
+
+$\begin{array}{lll}
+& \text{fmap } \alpha_{\hat{A},\hat{B},\hat{C}} \circ \tau_{\hat{A} \times \hat{B}, \hat{C}} = & \\
+& [q, {\bf let \: pure \:} z = q {\: \bf in \:} \langle \pi_1 (\pi_1 z), \langle \pi_2 (\pi_1 z), \pi_2 z \rangle \rangle] \circ [p, {\bf let \: pure \:} x, y = {\bf pure \:}(\pi_1 p), \pi_2 {\: \bf in \:} \langle x, y \rangle ] = & \\
+& [p, {\bf let \: pure \:} z = q {\: \bf in \:} \langle \pi_1 (\pi_1 z), \langle \pi_2 (\pi_1 z), \pi_2 z \rangle \rangle [q := {\bf let \: pure \:} x, y = {\bf pure \:}(\pi_1 p), \pi_2 {\: \bf in \:} \langle x, y \rangle]] = & \\
+& [p, {\bf let \: pure \:} z = ({\bf let \: pure \:} x, y = {\bf pure \:}(\pi_1 p), \pi_2 {\: \bf in \:} \langle x, y \rangle) {\: \bf in \:} \langle \pi_1 (\pi_1 z), \langle \pi_2 (\pi_1 z), \pi_2 z \rangle \rangle] = & \\
+& [p, {\bf let \: pure \:} x, y = {\bf pure \:}(\pi_1 p), \pi_2 {\: \bf in \:} \langle \pi_1 (\pi_1 z), \langle \pi_2 (\pi_1 z), \pi_2 z \rangle \rangle [z := x, y]] = & \\
+& [p, {\bf let \: pure \:} x, y = {\bf pure \:}(\pi_1 p), \pi_2 {\: \bf in \:} \langle \pi_1 x, \langle \pi_2 x, y \rangle \rangle]&
+\end{array}$
 \end{proof}

### 2018-02-11 article.tex
modifications

--- 2018-02-10 article.tex
@@
 
-  {\bf K} is a monoidal endofunctor
+  {\bf K} is a lax monoidal endofunctor
 \end{lemma}
@@
 
-Simple, but routine checking.
+1) Let us define $\tau_{\hat{A} \times \hat{B}, \hat{C}}$ as follows:
+
+$\tau_{\hat{A} \times \hat{B}, \hat{C}} = [p, {\bf \: let \: pure \:} x,y,z = {\bf pure\:} (\pi_1 (\pi_1 p)), {\bf pure \:} (\pi_2 (\pi_1 p)), \pi_2 p {\: \bf in \:} \langle \langle x,y \rangle, z \rangle]$
+
+\vspace{\baselineskip}
+
+Then:
+
+\vspace{\baselineskip}
+
+$\begin{array}{lll}
+&\text{fmap } \alpha_{\hat{A},\hat{B},\hat{C}} \circ \tau_{\hat{A} \times \hat{B}, \hat{C}} = & \\
+&[q, {\bf let \: pure \:} r = q {\: \bf in \:}  \langle \pi_1 (\pi_1 r), \langle \pi_2 (\pi_1 r), \pi_2 r \rangle \rangle] \: \circ & \\
+&\quad\quad\quad\quad \circ \: [p, {\bf \: let \: pure \:} x,y,z = {\bf pure\:} (\pi_1 (\pi_1 p)), {\bf pure \:} (\pi_2 (\pi_1 p)), \pi_2 p {\: \bf in \:} \langle \langle x,y \rangle, z \rangle] = & \\
+&[p, {\bf let \: pure \:} r = q {\: \bf in \:}  \langle \pi_1 (\pi_1 r), \langle \pi_2 (\pi_1 r), \pi_2 r \rangle \rangle & \\
+&\quad\quad\quad\quad [q := {\bf \: let \: pure \:} x,y,z = {\bf pure\:} (\pi_1 (\pi_1 p)), {\bf pure \:} (\pi_2 (\pi_1 p)), \pi_2 p {\: \bf in \:} \langle \langle x,y \rangle, z \rangle]] = & \\
+&[p, {\bf let \: pure \:} r = ({\bf let \: pure \:} x,y,z = {\bf pure\:} (\pi_1 (\pi_1 p)), {\bf pure \:} (\pi_2 (\pi_1 p)), \pi_2 p {\: \bf in \:} \langle \langle x,y \rangle, z \rangle) & \\
+&\quad\quad\quad\quad {\: \bf in \:} \langle \pi_1 (\pi_1 r), \langle \pi_2 (\pi_1 r), \pi_2 r \rangle \rangle] = & \\
+&[p, {\bf let \: pure \:} x,y,z = {\bf pure\:} (\pi_1 (\pi_1 p)), {\bf pure \:} (\pi_2 (\pi_1 p)), \pi_2 p {\: \bf in \:}  \langle \pi_1 (\pi_1 r), \langle \pi_2 (\pi_1 r), \pi_2 r \rangle \rangle & \\
+&\quad\quad\quad\quad [r := \langle \langle x,y \rangle, z \rangle]] = & \\
+&[p, {\bf let \: pure \:} x,y,z = {\bf pure\:} (\pi_1 (\pi_1 p)), {\bf pure \:} (\pi_2 (\pi_1 p)), \pi_2 p {\: \bf in \:} \langle x, \langle y, z \rangle \rangle]&
+\end{array}$
+
+\vspace{\baselineskip}
+
+On the other hand:
+
+\vspace{\baselineskip}
+
+Let us define as follows $\tau_{\hat{A}, \hat{B} \times \hat{C}}$:
+
+$\begin{array}{lll}
+&\tau_{\hat{A}, \hat{B} \times \hat{C}} = [r, {\bf let \: pure \:} x, y, z = ({\bf pure \:} \pi_1 r, {\bf let \: pure \:} q^{'} = \pi_2 r {\: \bf in \:} \pi_1 q^{'}, {\bf let \: pure \:} q^{'} = \pi_2 r {\: \bf in \:} \pi_2 q^{'}) & \\
+&\quad\quad\quad\quad {\: \bf in \:} \langle x, \langle y, z \rangle\rangle]& \\
+\end{array}$
+
+We may simplify the current instance of tensorial strength:
+
 \end{proof}

### 2018-02-11 article.tex
added proof of coherence for tensorial strength in the term model

@@
 
+\vspace{\baselineskip}
+
 We may simplify the current instance of tensorial strength:
+
+$\begin{array}{lll}
+&[p, {\bf let \: pure \:} x, y, z = ({\bf pure \:} \pi_1 p, {\bf let \: pure \:} p^{'} = \pi_2 p {\: \bf in \:} \pi_1 p^{'}, {\bf let \: pure \:} p^{'} = \pi_2 p {\: \bf in \:} \pi_2 p^{'}) & \\
+&\quad\quad\quad\quad {\: \bf in \:} \langle x, \langle y, z \rangle\rangle] = & \\
+&[p, {\bf let \: pure \:} x, p^{'}, z = ({\bf pure \:} \pi_1 p, \pi_2 p, {\bf let \: pure \:} p^{'} = \pi_2 p {\: \bf in \:} \pi_2 p^{'}) {\: \bf in \:} \langle x, \langle \pi_1 p^{'}, z \rangle\rangle] = & \\
+&[p, {\bf let \: pure \:} x, p^{'} = {\bf pure \:} \pi_1 p, \pi_2 p {\: \bf in \:} \langle x, \langle \pi_1 p^{'}, \pi_2 p^{'} \rangle\rangle]&
+\end{array}$
+
+\vspace{\baselineskip}
+
+Then:
+
+\vspace{\baselineskip}
+
+$\begin{array}{lll}
+&\tau_{\hat{A}, \hat{B} \times \hat{C}} \circ (id_{\hat{A}} \times \tau_{\hat{B}, \hat{C}}) \circ \alpha_{\hat{A}, \hat{B}, {\bf K}\hat{C}} = & \\
+&\tau_{\hat{A}, \hat{B} \times \hat{C}} \circ [q, \langle \pi_1 q, {\bf let \: pure \:} y, z = {\bf pure \:} (\pi_1 (\pi_2 q)), \pi_2 (\pi_2 q) {\: \bf in \:} \langle y, z \rangle \rangle] \: \circ & \\
+&\quad\quad\quad\quad \circ \: [p, \langle \pi_1 (\pi_1 p), \langle \pi_2 (\pi_1 p), \pi_2 p \rangle \rangle] = & \\
+&\tau_{\hat{A}, \hat{B} \times \hat{C}} \circ [p, \langle \pi_1 q, {\bf let \: pure \:} y, z = {\bf pure \:} (\pi_1 (\pi_2 q)), \pi_2 (\pi_2 q) {\: \bf in \:} \langle y, z \rangle \rangle & \\
+&\quad\quad\quad\quad [q := \langle \pi_1 (\pi_1 p), \langle \pi_2 (\pi_1 p), \pi_2 p \rangle \rangle]] = & \\
+&\tau_{\hat{A}, \hat{B} \times \hat{C}} \circ [p, \langle \pi_1 (\pi_1 p), {\bf let \: pure \:} y, z = {\bf pure \:} \pi_2 (\pi_1 p), \pi_2 p {\: \bf in \:} \langle y, z \rangle \rangle] = & \\
+&[r, {\bf let \: pure \:} x, p^{'} = {\bf pure \:} \pi_1 r, \pi_2 r {\: \bf in \:} \langle x, \langle \pi_1 p^{'}, \pi_2 p^{'} \rangle\rangle] \: \circ & \\
+&\quad\quad\quad\quad \circ \: [p, \langle \pi_1 (\pi_1 p), {\bf let \: pure \:} y, z = {\bf pure \:} \pi_2 (\pi_1 p), \pi_2 p {\: \bf in \:} \langle y, z \rangle \rangle] = & \\
+&[p, {\bf let \: pure \:} x, p^{'} = ({\bf pure \:} (\pi_1 (\pi_1 p)), {\bf let \: pure \:} y, z = ({\bf pure \:} \pi_2 (\pi_1 p), \pi_2 p {\: \bf in \:} \langle y, z \rangle \rangle) & \\
+&\quad\quad\quad\quad {\: \bf in \:} \langle x, \langle \pi_1 p^{'}, \pi_2 p^{'} \rangle\rangle ] = & \\
+&[p, {\bf let \: pure \:} x,y,z = {\bf pure \:} (\pi_1 (\pi_1 p)), {\bf pure \:} \pi_2 (\pi_1 p), \pi_2 p {\: \bf in \:} \langle x, \langle \pi_1 p^{'}, \pi_2 p^{'} \rangle\rangle [p^{'} := \langle y, z \rangle]] = & \\
+&[p, {\bf let \: pure \:} x,y,z = {\bf pure \:} (\pi_1 (\pi_1 p)), {\bf pure \:} \pi_2 (\pi_1 p), \pi_2 p {\: \bf in \:} \langle x, \langle y, z \rangle\rangle]&
+\end{array}$
 

### 2018-02-11 article.tex
added simple mod
added comments for proof of coherence in the term model

--- 2018-02-11 article.tex
@@
 &\quad\quad\quad\quad \circ \: [p, {\bf \: let \: pure \:} x,y,z = {\bf pure\:} (\pi_1 (\pi_1 p)), {\bf pure \:} (\pi_2 (\pi_1 p)), \pi_2 p {\: \bf in \:} \langle \langle x,y \rangle, z \rangle] = & \\
+&\text{Composition}& \\
 &[p, {\bf let \: pure \:} r = q {\: \bf in \:}  \langle \pi_1 (\pi_1 r), \langle \pi_2 (\pi_1 r), \pi_2 r \rangle \rangle & \\
 &\quad\quad\quad\quad [q := {\bf \: let \: pure \:} x,y,z = {\bf pure\:} (\pi_1 (\pi_1 p)), {\bf pure \:} (\pi_2 (\pi_1 p)), \pi_2 p {\: \bf in \:} \langle \langle x,y \rangle, z \rangle]] = & \\
+&\text{Substitution and reduction}& \\
 &[p, {\bf let \: pure \:} r = ({\bf let \: pure \:} x,y,z = {\bf pure\:} (\pi_1 (\pi_1 p)), {\bf pure \:} (\pi_2 (\pi_1 p)), \pi_2 p {\: \bf in \:} \langle \langle x,y \rangle, z \rangle) & \\
 &\quad\quad\quad\quad {\: \bf in \:} \langle \pi_1 (\pi_1 r), \langle \pi_2 (\pi_1 r), \pi_2 r \rangle \rangle] = & \\
+&\text{$\beta$-reduction rule}& \\
 &[p, {\bf let \: pure \:} x,y,z = {\bf pure\:} (\pi_1 (\pi_1 p)), {\bf pure \:} (\pi_2 (\pi_1 p)), \pi_2 p {\: \bf in \:}  \langle \pi_1 (\pi_1 r), \langle \pi_2 (\pi_1 r), \pi_2 r \rangle \rangle & \\
 &\quad\quad\quad\quad [r := \langle \langle x,y \rangle, z \rangle]] = & \\
+&\text{$\beta$-reduction rule}& \\
 &[p, {\bf let \: pure \:} x,y,z = {\bf pure\:} (\pi_1 (\pi_1 p)), {\bf pure \:} (\pi_2 (\pi_1 p)), \pi_2 p {\: \bf in \:} \langle x, \langle y, z \rangle \rangle]&
@@
 
+\vspace{\baselineskip}
+
 We may simplify the current instance of tensorial strength:
+
+$\begin{array}{lll}
+&[p, {\bf let \: pure \:} x, y, z = ({\bf pure \:} \pi_1 p, {\bf let \: pure \:} p^{'} = \pi_2 p {\: \bf in \:} \pi_1 p^{'}, {\bf let \: pure \:} p^{'} = \pi_2 p {\: \bf in \:} \pi_2 p^{'}) & \\
+&\quad\quad\quad\quad {\: \bf in \:} \langle x, \langle y, z \rangle\rangle] = & \\
+&\text{Substitution and reduction}& \\
+&[p, {\bf let \: pure \:} x, p^{'}, z = ({\bf pure \:} \pi_1 p, \pi_2 p, {\bf let \: pure \:} p^{'} = \pi_2 p {\: \bf in \:} \pi_2 p^{'}) {\: \bf in \:} \langle x, \langle \pi_1 p^{'}, z \rangle\rangle] = & \\
+&\text{Substitution and reduction}& \\
+&[p, {\bf let \: pure \:} x, p^{'} = {\bf pure \:} \pi_1 p, \pi_2 p {\: \bf in \:} \langle x, \langle \pi_1 p^{'}, \pi_2 p^{'} \rangle\rangle]&
+\end{array}$
+
+\vspace{\baselineskip}
+
+Then:
+
+\vspace{\baselineskip}
+
+$\begin{array}{lll}
+&\tau_{\hat{A}, \hat{B} \times \hat{C}} \circ (id_{\hat{A}} \times \tau_{\hat{B}, \hat{C}}) \circ \alpha_{\hat{A}, \hat{B}, {\bf K}\hat{C}} = & \\
+&\text{Unfolding}& \\
+&\tau_{\hat{A}, \hat{B} \times \hat{C}} \circ [q, \langle \pi_1 q, {\bf let \: pure \:} y, z = {\bf pure \:} (\pi_1 (\pi_2 q)), \pi_2 (\pi_2 q) {\: \bf in \:} \langle y, z \rangle \rangle] \: \circ & \\
+&\quad\quad\quad\quad \circ \: [p, \langle \pi_1 (\pi_1 p), \langle \pi_2 (\pi_1 p), \pi_2 p \rangle \rangle] = & \\
+&\text{Composition}& \\
+&\tau_{\hat{A}, \hat{B} \times \hat{C}} \circ [p, \langle \pi_1 q, {\bf let \: pure \:} y, z = {\bf pure \:} (\pi_1 (\pi_2 q)), \pi_2 (\pi_2 q) {\: \bf in \:} \langle y, z \rangle \rangle & \\
+&\quad\quad\quad\quad [q := \langle \pi_1 (\pi_1 p), \langle \pi_2 (\pi_1 p), \pi_2 p \rangle \rangle]] = & \\
+&\text{Substitution and reduction}& \\
+&\tau_{\hat{A}, \hat{B} \times \hat{C}} \circ [p, \langle \pi_1 (\pi_1 p), {\bf let \: pure \:} y, z = {\bf pure \:} \pi_2 (\pi_1 p), \pi_2 p {\: \bf in \:} \langle y, z \rangle \rangle] = & \\
+&\text{Composition}& \\
+&[r, {\bf let \: pure \:} x, p^{'} = {\bf pure \:} \pi_1 r, \pi_2 r {\: \bf in \:} \langle x, \langle \pi_1 p^{'}, \pi_2 p^{'} \rangle\rangle] \: \circ & \\
+&\quad\quad\quad\quad \circ \: [p, \langle \pi_1 (\pi_1 p), {\bf let \: pure \:} y, z = {\bf pure \:} \pi_2 (\pi_1 p), \pi_2 p {\: \bf in \:} \langle y, z \rangle \rangle] = & \\
+&\text{Substitution and reduction}& \\
+&[p, {\bf let \: pure \:} x, p^{'} = ({\bf pure \:} (\pi_1 (\pi_1 p)), {\bf let \: pure \:} y, z = ({\bf pure \:} \pi_2 (\pi_1 p), \pi_2 p {\: \bf in \:} \langle y, z \rangle \rangle) & \\
+&\quad\quad\quad\quad {\: \bf in \:} \langle x, \langle \pi_1 p^{'}, \pi_2 p^{'} \rangle\rangle ] = & \\
+&\text{Reduction}& \\
+&[p, {\bf let \: pure \:} x,y,z = {\bf pure \:} (\pi_1 (\pi_1 p)), {\bf pure \:} \pi_2 (\pi_1 p), \pi_2 p {\: \bf in \:} \langle x, \langle \pi_1 p^{'}, \pi_2 p^{'} \rangle\rangle [p^{'} := \langle y, z \rangle]] = & \\
+&\text{Substitution and reduction}& \\
+&[p, {\bf let \: pure \:} x,y,z = {\bf pure \:} (\pi_1 (\pi_1 p)), {\bf pure \:} \pi_2 (\pi_1 p), \pi_2 p {\: \bf in \:} \langle x, \langle y, z \rangle\rangle]&
+\end{array}$
+
+\vspace{\baselineskip}
+
+2) Easy
 

### 2018-02-13 article.tex
modified comments in proofs

--- 2018-02-11 article.tex
@@
 
-\begin{lemma} Properties of $\eta$:
+\begin{lemma} Naturality and coherence for $\eta$:
 
@@
 & \ast_{\hat{A},\hat{B}} \circ (\eta_{\hat{A}} \times \eta_{\hat{B}}) = & \\
-& \text{By unfolding} & \\
+& \quad\quad\quad\quad\quad\quad \text{By unfolding} & \\
 & [q, {\bf let \: pure \:} x, y = \pi_1 q, \pi_2 q {\: \bf in \:} \langle x, y \rangle] \circ [p, \langle {\bf pure \:} (\pi_1 p), {\bf pure \:} (\pi_2 p) \rangle] = & \\
-& \text{Composition}& \\
+& \quad\quad\quad\quad\quad\quad \text{Composition}& \\
 & [p, {\bf let \: pure \:} x, y = \pi_1 q, \pi_2 q {\: \bf in \:} \langle x, y \rangle [q := \langle {\bf pure \:} (\pi_1 p), {\bf pure \:} (\pi_2 p) \rangle]] = & \\
-& \text{By substitution}& \\
+& \quad\quad\quad\quad\quad\quad \text{By substitution}& \\
 & [p, {\bf let \: pure \:} x, y = \pi_1 (\langle {\bf pure \:} (\pi_1 p), {\bf pure \:} (\pi_2 p) \rangle), \pi_2 (\langle {\bf pure \:} (\pi_1 p), {\bf pure \:} (\pi_2 p) \rangle) {\: \bf in \:} \langle x, y \rangle] = & \\
-& \text{Reduction rules}& \\
+& \quad\quad\quad\quad\quad\quad \text{Reduction rules}& \\
 & [p, {\bf let \: pure \:} x, y = {\bf pure \:} (\pi_1 p), {\bf pure \:} (\pi_2 p) {\: \bf in \:} \langle x, y \rangle] = & \\
-& \text{Reduction rule}& \\
+& \quad\quad\quad\quad\quad\quad \text{Reduction rule}& \\
 & [p, {\bf pure \:} (\langle x,y \rangle [x := \pi_1 p, y := \pi_2 p])] = & \\
-& \text{Substitution}& \\
+& \quad\quad\quad\quad\quad\quad \text{Substitution}& \\
 & [p, {\bf pure \:} \langle \pi_1 p, \pi_2 p \rangle] = & \\
-& \text{$\eta$-reduction}& \\
+& \quad\quad\quad\quad\quad\quad \text{$\eta$-reduction}& \\
 & [p, {\bf pure \:} p] =& \\
-& \text{By definition}& \\
+& \quad\quad\quad\quad\quad\quad \text{By definition}& \\
 & \eta_{\hat{A} \times \hat{B}}&
@@
 &\quad\quad\quad\quad \circ \: [p, {\bf \: let \: pure \:} x,y,z = {\bf pure\:} (\pi_1 (\pi_1 p)), {\bf pure \:} (\pi_2 (\pi_1 p)), \pi_2 p {\: \bf in \:} \langle \langle x,y \rangle, z \rangle] = & \\
+&\quad\quad\quad\quad\quad\quad \text{Composition}& \\
 &[p, {\bf let \: pure \:} r = q {\: \bf in \:}  \langle \pi_1 (\pi_1 r), \langle \pi_2 (\pi_1 r), \pi_2 r \rangle \rangle & \\
 &\quad\quad\quad\quad [q := {\bf \: let \: pure \:} x,y,z = {\bf pure\:} (\pi_1 (\pi_1 p)), {\bf pure \:} (\pi_2 (\pi_1 p)), \pi_2 p {\: \bf in \:} \langle \langle x,y \rangle, z \rangle]] = & \\
+&\quad\quad\quad\quad\quad\quad \text{Substitution and reduction}& \\
 &[p, {\bf let \: pure \:} r = ({\bf let \: pure \:} x,y,z = {\bf pure\:} (\pi_1 (\pi_1 p)), {\bf pure \:} (\pi_2 (\pi_1 p)), \pi_2 p {\: \bf in \:} \langle \langle x,y \rangle, z \rangle) & \\
 &\quad\quad\quad\quad {\: \bf in \:} \langle \pi_1 (\pi_1 r), \langle \pi_2 (\pi_1 r), \pi_2 r \rangle \rangle] = & \\
+&\quad\quad\quad\quad\quad\quad \text{$\beta$-reduction rule}& \\
 &[p, {\bf let \: pure \:} x,y,z = {\bf pure\:} (\pi_1 (\pi_1 p)), {\bf pure \:} (\pi_2 (\pi_1 p)), \pi_2 p {\: \bf in \:}  \langle \pi_1 (\pi_1 r), \langle \pi_2 (\pi_1 r), \pi_2 r \rangle \rangle & \\
 &\quad\quad\quad\quad [r := \langle \langle x,y \rangle, z \rangle]] = & \\
+&\quad\quad\quad\quad\quad\quad \text{$\beta$-reduction rule}& \\
 &[p, {\bf let \: pure \:} x,y,z = {\bf pure\:} (\pi_1 (\pi_1 p)), {\bf pure \:} (\pi_2 (\pi_1 p)), \pi_2 p {\: \bf in \:} \langle x, \langle y, z \rangle \rangle]&
@@
 
+\vspace{\baselineskip}
+
 We may simplify the current instance of tensorial strength:
+
+$\begin{array}{lll}
+&[p, {\bf let \: pure \:} x, y, z = ({\bf pure \:} \pi_1 p, {\bf let \: pure \:} p^{'} = \pi_2 p {\: \bf in \:} \pi_1 p^{'}, {\bf let \: pure \:} p^{'} = \pi_2 p {\: \bf in \:} \pi_2 p^{'}) & \\
+&\quad\quad\quad\quad {\: \bf in \:} \langle x, \langle y, z \rangle\rangle] = & \\
+&\quad\quad\quad\quad\quad\quad \text{Substitution and reduction}& \\
+&[p, {\bf let \: pure \:} x, p^{'}, z = ({\bf pure \:} \pi_1 p, \pi_2 p, {\bf let \: pure \:} p^{'} = \pi_2 p {\: \bf in \:} \pi_2 p^{'}) {\: \bf in \:} \langle x, \langle \pi_1 p^{'}, z \rangle\rangle] = & \\
+&\quad\quad\quad\quad\quad\quad \text{Substitution and reduction}& \\
+&[p, {\bf let \: pure \:} x, p^{'} = {\bf pure \:} \pi_1 p, \pi_2 p {\: \bf in \:} \langle x, \langle \pi_1 p^{'}, \pi_2 p^{'} \rangle\rangle]&
+\end{array}$
+
+\vspace{\baselineskip}
+
+Then:
+
+\vspace{\baselineskip}
+
+$\begin{array}{lll}
+&\tau_{\hat{A}, \hat{B} \times \hat{C}} \circ (id_{\hat{A}} \times \tau_{\hat{B}, \hat{C}}) \circ \alpha_{\hat{A}, \hat{B}, {\bf K}\hat{C}} = & \\
+&\quad\quad\quad\quad\quad\quad \text{Unfolding}& \\
+&\tau_{\hat{A}, \hat{B} \times \hat{C}} \circ [q, \langle \pi_1 q, {\bf let \: pure \:} y, z = {\bf pure \:} (\pi_1 (\pi_2 q)), \pi_2 (\pi_2 q) {\: \bf in \:} \langle y, z \rangle \rangle] \: \circ & \\
+&\quad\quad\quad\quad \circ \: [p, \langle \pi_1 (\pi_1 p), \langle \pi_2 (\pi_1 p), \pi_2 p \rangle \rangle] = & \\
+&\quad\quad\quad\quad\quad\quad \text{Composition}& \\
+&\tau_{\hat{A}, \hat{B} \times \hat{C}} \circ [p, \langle \pi_1 q, {\bf let \: pure \:} y, z = {\bf pure \:} (\pi_1 (\pi_2 q)), \pi_2 (\pi_2 q) {\: \bf in \:} \langle y, z \rangle \rangle & \\
+&\quad\quad\quad\quad [q := \langle \pi_1 (\pi_1 p), \langle \pi_2 (\pi_1 p), \pi_2 p \rangle \rangle]] = & \\
+&\quad\quad\quad\quad\quad\quad \text{Substitution and reduction}& \\
+&\tau_{\hat{A}, \hat{B} \times \hat{C}} \circ [p, \langle \pi_1 (\pi_1 p), {\bf let \: pure \:} y, z = {\bf pure \:} \pi_2 (\pi_1 p), \pi_2 p {\: \bf in \:} \langle y, z \rangle \rangle] = & \\
+&\quad\quad\quad\quad\quad\quad \text{Composition}& \\
+&[r, {\bf let \: pure \:} x, p^{'} = {\bf pure \:} \pi_1 r, \pi_2 r {\: \bf in \:} \langle x, \langle \pi_1 p^{'}, \pi_2 p^{'} \rangle\rangle] \: \circ & \\
+&\quad\quad\quad\quad \circ \: [p, \langle \pi_1 (\pi_1 p), {\bf let \: pure \:} y, z = {\bf pure \:} \pi_2 (\pi_1 p), \pi_2 p {\: \bf in \:} \langle y, z \rangle \rangle] = & \\
+&\quad\quad\quad\quad\quad\quad \text{Substitution and reduction}& \\
+&[p, {\bf let \: pure \:} x, p^{'} = ({\bf pure \:} (\pi_1 (\pi_1 p)), {\bf let \: pure \:} y, z = ({\bf pure \:} \pi_2 (\pi_1 p), \pi_2 p {\: \bf in \:} \langle y, z \rangle \rangle) & \\
+&\quad\quad\quad\quad {\: \bf in \:} \langle x, \langle \pi_1 p^{'}, \pi_2 p^{'} \rangle\rangle ] = & \\
+&\quad\quad\quad\quad\quad\quad \text{Reduction}& \\
+&[p, {\bf let \: pure \:} x,y,z = {\bf pure \:} (\pi_1 (\pi_1 p)), {\bf pure \:} \pi_2 (\pi_1 p), \pi_2 p {\: \bf in \:} \langle x, \langle \pi_1 p^{'}, \pi_2 p^{'} \rangle\rangle [p^{'} := \langle y, z \rangle]] = & \\
+&\quad\quad\quad\quad\quad\quad \text{Substitution and reduction}& \\
+&[p, {\bf let \: pure \:} x,y,z = {\bf pure \:} (\pi_1 (\pi_1 p)), {\bf pure \:} \pi_2 (\pi_1 p), \pi_2 p {\: \bf in \:} \langle x, \langle y, z \rangle\rangle]&
+\end{array}$
+
+\vspace{\baselineskip}
+
+2) Easy
 

### 2018-02-14 article.tex
made some beauty

@@
 
+\begin{defin} Monoidal category
+
+  Monoidal category is a category $\mathcal{C}$ with additional monoidal structure:
+  \begin{itemize}
+    \item A bifunctor $\otimes : \mathcal{C} \times \mathcal{C} \to C$ called tensor product;
+    \item Identity object $\mathds{1}$;
+    \item A natural isomorphism called associator: $\alpha_{A,B,C}: (A \otimes B) \otimes C \cong A \otimes (B \otimes C)$;
+    \item A natural isomorphism called left unitor: $L_A : \mathds{1} \otimes A \cong A$;
+    \item A natural isomorphism called right unitor $R_A : A \otimes \mathds{1} \cong A$;
+    \item A coherence condition called MacLane pentagon, i.e. the following diagram commutes:
+
+    \xymatrix{
+      & (A \otimes (B \otimes C)) \otimes D \ar[dr]^{\alpha_{A,B \otimes C,D}}\\
+      ((A \otimes B) \otimes C) \otimes D \ar[d]_{\alpha_{A \otimes B, C, D}} \ar[ur]^{\alpha_{A,B,C} \otimes id_D \quad} && A \otimes ((B \otimes C) \otimes D) \ar[d]^{id_A \otimes \alpha_{B,C,D}}\\
+      (A \otimes B) \otimes (C \otimes D) \ar[rr]_{\alpha_{A,B,C \otimes D}}&& A \otimes (B \otimes (C \otimes D))
+    }
+    \item A coherence condition called triangle identity:
+
+    \xymatrix{
+    && (A \otimes \mathds{1}) \otimes B \ar[rr]^{\alpha_{A, \mathds{1}, B}} \ar[dr]_{R_A \otimes id_B} && A \otimes (\mathds{1} \otimes B) \ar[dl]^{id_A \otimes L_B} \\
+    &&& A \otimes B
+    }
+
+    \item A monoidal category $\mathcal{C}$ is called symmetric if forall $A,B \in Ob_{\mathcal{C}}$, there is an isomorphism $\sigma_{A,B} : A \otimes B \cong B \otimes A$.
+  \end{itemize}
+\end{defin}
+
+\begin{defin} Cartesian closed category
+
+  Cartesian closed category is a category with a terminal object, finite products and exponentiation.
+\end{defin}
+
+Note that, any cartesian closed category is the special case of a (symmetric) monoidal category, where tensor is a product and identity object is a terminal object.
+
 \begin{defin} Lax monoidal functor
 
-  Let $\langle \mathcal{C}, \otimes_1, \mathds{1} \rangle$ and $\langle \mathcal{D}, \otimes_2, \mathds{1}' \rangle$ are monoidal categories.
+  Let $\langle \mathcal{C}, \otimes_1, \mathds{1}_{\mathcal{C}} \rangle$ and $\langle \mathcal{D}, \otimes_2, \mathds{1}_{\mathcal{D}} \rangle$ are monoidal categories.
 
   A lax monoidal functor $\mathcal{F} : \langle \mathcal{C}, \otimes_1, \mathds{1} \rangle \to \langle \mathcal{D}, \otimes_2, \mathds{1}' \rangle$ is a functor
-  $\mathcal{F} : \mathcal{C} \to \mathcal{D}$ with additional natural transformations, which satisfy the well-known conditions described in \cite{ElKelly}:
-
-  1) $u : \mathds{1}' \to \mathcal{F}\mathds{1}$;
-
-  2) $\ast_{A, B} : \mathcal{F}A \otimes_2 \mathcal{F}B \to \mathcal{F}(A \otimes_1 B)$.
-
+  $\mathcal{F} : \mathcal{C} \to \mathcal{D}$ with additional natural transformations:
+
+  \begin{itemize}
+  \item $u : \mathds{1}_{\mathcal{D}} \to \mathcal{F}\mathds{1}_{\mathcal{C}}$;
+  \item $\ast_{A, B} : \mathcal{F}A \otimes_{\mathcal{D}} \mathcal{F}B \to \mathcal{F}(A \otimes_{\mathcal{C}} B)$.
+  \end{itemize}
+
+  and coherence maps:
+
+  \begin{itemize}
+    \item Associativity:
+
+  \xymatrix{
+    && (\mathcal{F}A \otimes_{\mathcal{D}} \mathcal{F}B) \otimes_{\mathcal{D}} \mathcal{F}C \ar[d]_{\ast_{A,B} \otimes_{\mathcal{D}} id_{\mathcal{F}B}}
+    \ar[rr]^{\alpha^{\mathcal{D}}_{{\mathcal{F}A, \mathcal{F}B, \mathcal{F}C}}} && \mathcal{F}A \otimes_{\mathcal{D}} (\mathcal{F}B \otimes_{\mathcal{D}} \mathcal{F}C) \ar[d]^{id_{\mathcal{F}A} \otimes_{\mathcal{D}} \ast_{B,C}}\\
+    && \mathcal{F}(A \otimes_{\mathcal{C}} B) \otimes_{\mathcal{D}} \mathcal{C} \ar[d]_{\ast_{A \otimes_{\mathcal{C}} B, C}} && \mathcal{F}A \otimes_{\mathcal{D}} \mathcal{F}(B \otimes_{\mathcal{C}} C) \ar[d]^{\ast_{A, B \otimes_{\mathcal{C}} C}}\\
+    && \mathcal{F}((A \otimes_{\mathcal{C}} B) \otimes_{\mathcal{C}} C) \ar[rr]_{\mathcal{F}(\alpha^{\mathcal{C}}_{A,B,C})}&& \mathcal{F}(A \otimes_{\mathcal{C}} (B \otimes_{\mathcal{C}} C))
+  }
+
+    \item Left unitality:
+
+    \xymatrix{
+    &&  \mathds{1}_{\mathcal{D}} \otimes_{\mathcal{D}} \mathcal{F}A \ar[d]_{L^{\mathcal{D}}_{\mathcal{F}A}}\ar[rr]^{u \otimes_{\mathcal{D}} id_{\mathcal{F}A}} && \mathcal{F}\mathds{1}_{\mathcal{C}} \otimes_{\mathcal{D}} \mathcal{F}A \ar[d]^{\ast_{\mathds{1}_{\mathcal{C}}, A}} \\
+    &&  \mathcal{F}A && \mathcal{F}(\mathds{1}_{\mathcal{C}} \otimes_{\mathcal{C}} A) \ar[ll]_{\mathcal{F}(L^{\mathcal{C}}_A)}
+    }
+
+    \item Right unitality:
+
+    \xymatrix{
+    &&  \mathcal{F}A \otimes_{\mathcal{D}} \mathds{1}_{\mathcal{D}} \ar[d]_{R^{\mathcal{D}}_{\mathcal{F}A}}\ar[rr]^{id_{\mathcal{F}A} \otimes_{\mathcal{D}} u} && \mathcal{F}A \otimes_{\mathcal{D}} \mathcal{F}\mathds{1}_{\mathcal{C}} \ar[d]^{\ast_{A, \mathds{1}_{\mathcal{C}}}} \\
+    &&  \mathcal{F}A && \mathcal{F}(A \otimes_{\mathcal{C}} \mathds{1}_{\mathcal{C}} ) \ar[ll]^{\mathcal{F}(R^{\mathcal{C}}_A)}
+    }
+  \end{itemize}
 \end{defin}
@@
 
-Strong endofunctor on a monoidal category is a with a natural transformation so called tensorial strength
+Strong endofunctor on a monoidal category is with a natural transformation so called tensorial strength
 
@@
 
-1) $u = \eta_{\mathds{1}}$;
-
-2) $\ast_{A,B} \circ (\eta_A \otimes \eta_B) = \eta_{A \otimes B}$, i.e. the following diagram commutes:
+\begin{itemize}
+\item $u = \eta_{\mathds{1}}$;
+\item $\ast_{A,B} \circ (\eta_A \otimes \eta_B) = \eta_{A \otimes B}$, i.e. the following diagram commutes:
 
@@
 }
-
-3) $\tau_{A, B} = \ast_{A, B} \circ \eta_{A} \otimes id_{\mathcal{K}B}$.
+\item $\tau_{A, B} = \ast_{A, B} \circ \eta_{A} \otimes id_{\mathcal{K}B}$.
+\end{itemize}
 \end{defin}
@@
 
-\begin{defin} Semantical translation from $\lambda_{{\bf K}}$ to some cartesian closed category $\mathcal{C}$ with applicative functor $\mathcal{K}$:
-
-1) Interpretation for types:
-
-$[\![A]\!] := \hat{A}, A \in \mathbb{T}$, where $\hat{A}$ is an object of $\mathcal{C}$ obtained by some given assignment;
-
-$[\![A \to B]\!] := [\![B]\!]^{[\![A]\!]}$;
-
-$[\![A \times B]\!] := [\![A]\!] \times [\![B]\!]$.
-
-2) Interpretation for modal types: $[\![{\bf K}A]\!] = \mathcal{K}[\![A]\!]$;
-
-3) Interpretaion for contexts:
-
-$[\![\Gamma = \{ x_1 : A_1, ..., x_n : A_n\}]\!] := [\![\Gamma]\!] = [\![A_1]\!] \times ... \times [\![A_n]\!]$;
-
-4) Interpretation for typing assignment: $[\![\Gamma \vdash M : A]\!] := [\![M]\!] : [\![\Gamma]\!] \to
-[\![A]\!]$.
-
-5) Interpretation for typing rules:
+\begin{defin} Semantical translation from $\lambda_{{\bf K}}$ to some cartesian closed category $\mathcal{C}$ with an applicative functor $\mathcal{K}$:
+
+\begin{itemize}
+\item Interpretation for types:
+  \begin{itemize}
+    \item $[\![A]\!] := \hat{A}, A \in \mathbb{T}$, where $\hat{A}$ is an object of $\mathcal{C}$ obtained by some given assignment;
+    \item $[\![A \to B]\!] := [\![B]\!]^{[\![A]\!]}$;
+    \item $[\![A \times B]\!] := [\![A]\!] \times [\![B]\!]$.
+  \end{itemize}
+\item Interpretation for modal types:
+  \begin{itemize}
+    \item $[\![{\bf K}A]\!] = \mathcal{K}[\![A]\!]$;
+  \end{itemize}
+\item Interpretaion for contexts:
+  \begin{itemize}
+    \item $[\![ \quad ]\!] = \mathds{1}$, where $\mathds{1}$ is a terminal object of a given ccc;
+    \item $[\![\Gamma, x : A]\!] = [\![\Gamma]\!] \times [\![A]\!]$
+  \end{itemize}
+\item Interpretation for typing assignment:
+  \begin{itemize}
+    \item $[\![\Gamma \vdash M : A]\!] := [\![M]\!] : [\![\Gamma]\!] \to [\![A]\!]$.
+  \end{itemize}
+\item Interpretation for typing rules:
 
@@
 
+\begin{small}
   \begin{prooftree}
@@
   \end{prooftree}
-
+\end{small}
+\end{itemize}
 \end{defin}
@@
 
-1) $x_i [ \vec{x} := \vec{M}] = M_i $;
-
-2) $(\lambda x. M) [ \vec{x} := \vec{M}] = \lambda x. (M [ \vec{x} := \vec{M}])$;
-
-3) $(M N) [ \vec{x} := \vec{M}] = (M [ \vec{x} = \vec{M}]) (N [ \vec{x} := \vec{M}])$;
-
-4) $\langle M, N \rangle = \langle (M [ \vec{x} = \vec{M}]), (N [ \vec{x} := \vec{M}])\rangle$;
-
-5) $(\pi_i P) [ \vec{x} := \vec{M}] = \pi_i (P [ \vec{x} = \vec{M}])$;
-
-6) $({\bf pure \:} M) [ \vec{x} := \vec{M}] = {\bf pure \:} (M [ \vec{x} = \vec{M}])$;
-
-7) $({\bf let \: pure} \: \vec{x} = \vec{M} {\: \bf in \:} N) [\vec{y} := \vec{P}] =
+\begin{itemize}
+\item $x_i [ \vec{x} := \vec{M}] = M_i $;
+\item $(\lambda x. M) [ \vec{x} := \vec{M}] = \lambda x. (M [ \vec{x} := \vec{M}])$;
+\item $(M N) [ \vec{x} := \vec{M}] = (M [ \vec{x} = \vec{M}]) (N [ \vec{x} := \vec{M}])$;
+\item $\langle M, N \rangle = \langle (M [ \vec{x} = \vec{M}]), (N [ \vec{x} := \vec{M}])\rangle$;
+\item $(\pi_i P) [ \vec{x} := \vec{M}] = \pi_i (P [ \vec{x} = \vec{M}])$;
+\item $({\bf pure \:} M) [ \vec{x} := \vec{M}] = {\bf pure \:} (M [ \vec{x} = \vec{M}])$;
+\item $({\bf let \: pure} \: \vec{x} = \vec{M} {\: \bf in \:} N) [\vec{y} := \vec{P}] =
 {\bf let \: pure} \: \vec{x} = (\vec{M} [\vec{y} := \vec{P}]) {\: \bf in \:} N$
-
+\end{itemize}
 \end{defin}
@@
 
-1) $[\![\Gamma \vdash ({\bf pure \:} M) [ \vec{x} := \vec{M}] : {\bf K}A ]\!] = [\![ \Gamma \vdash {\bf pure \:} M : {\bf K}A]\!] \circ \langle [\![M_1]\!], \dots,
-[\![M_n]\!] \rangle$.
-
-\vspace{\baselineskip}
+1)
 
@@
 
-2) $\begin{array}{lll}&[\![\Gamma \vdash ({\bf let \: pure} \: \vec{x} = \vec{M} {\: \bf in \:} N) [\vec{y} := \vec{P}] : {\bf K} B]\!]
-=[\![\Gamma \vdash {\bf let \: pure} \: \vec{x} = \vec{M} {\: \bf in \:} N : {\bf K}B]\!] \circ \langle [\![P_1]\!],\dots,[\![P_n]\!]\rangle& \end{array}$
+2)
 
@@
 
-  1) $\eta:Id \Rightarrow \mathcal{K}$, s. t. $\forall \hat{A} \in Ob_{\mathcal{C}(\lambda)}$, $\eta_{\hat{A}} = [x, {\bf pure \:} x] \in Hom_{\mathcal{C}(\lambda)}(\hat{A}, {\bf K}\hat{A})$;
-
-  2) $\ast_{A,B}:{\bf K}\hat{A} \times {\bf K}\hat{B} \to {\bf K}(\hat{A} \times \hat{B})$, s.
+\begin{itemize}
+  \item $\eta:Id \Rightarrow \mathcal{K}$, s. t. $\forall \hat{A} \in Ob_{\mathcal{C}(\lambda)}$, $\eta_{\hat{A}} = [x, {\bf pure \:} x] \in Hom_{\mathcal{C}(\lambda)}(\hat{A}, {\bf K}\hat{A})$;
+  \item $\ast_{A,B}:{\bf K}\hat{A} \times {\bf K}\hat{B} \to {\bf K}(\hat{A} \times \hat{B})$, s.
 t.
 $\forall \hat{A}, \hat{B} \in Ob_{\mathcal{C}(\lambda)}, \ast_{\hat{A},\hat{B}} = [p, {\bf let \: pure \:} x,y = \pi_1 p, \pi_2 p {\: \bf in \:} \langle x, y \rangle] \in Hom_{\mathcal{C}(\lambda)}({\bf K}A \times {\bf K}B, {\bf K}(A \times B))$.
+\end{itemize}
 \end{defin}

### 2018-02-15 article.tex
added acknowledgment
added some beauty again

@@
   \item If $\Gamma, x : A \vdash M : B$ and $\Gamma \vdash N : A$, then $\Gamma \vdash M [x := N] : B$.
-  \item If $\Gamma \vdash M : A$, then $\Gamma [B := C] \vdash M : (A [B := C])$;
-  \item If $\Gamma \vdash M : A$ and $\Delta$ is a permutation of $\Gamma$, then $\Delta \vdash M : A$.
+  \item If $\Gamma \vdash M : A$, then $\Gamma [B := C] \vdash M : (A [B := C])$.
 \end{itemize}
@@
 
-\begin{ex}
-  $ $
-
-  $\phi((\lambda_i x. {\bf let \: pure \:} y = x {\: \bf in \:} x) N) = {\bf let \: pure \:} y = \phi(N) {\: \bf in \:} \phi(N)$
-\end{ex}
-
 \begin{lemma}
@@
   $\begin{array}{lll}
-  & \phi({\bf pure \:} (M [x := N])) = & \text{By the definition of $\phi$}\\
-  & {\bf pure \:} (\phi(M [x := N])) = & \text{Induction hypothesis} \\
-  & {\bf pure \:} (\phi(M) [x := \phi(N)]) = & \text{Substitution definition}\\
+  & \phi({\bf pure \:} (M [x := N])) = & \\
+  &\quad\quad\quad\quad\quad\quad \text{By the definition of $\phi$} & \\
+  & {\bf pure \:} (\phi(M [x := N])) = & \\
+  &\quad\quad\quad\quad\quad\quad \text{Induction hypothesis}& \\
+  & {\bf pure \:} (\phi(M) [x := \phi(N)]) = &\\
+  &\quad\quad\quad\quad\quad\quad \text{Substitution definition}& \\
   & ({\bf pure \:} \phi (M)) [x := \phi(N)]&
@@
   $\begin{array}{lll}
-  &\phi(({\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N) [y := P]) = & \text{Substitution definition}\\
-  &\phi({\bf let \: pure \:} \vec{x} = (\vec{M}[y := P]) {\: \bf in \:} N) = & \text{By the definition of $\phi$}\\
-  &{\bf let \: pure \:} \vec{x} = \phi(\vec{M}[y := P]) {\: \bf in \:} \phi(N) = & \text{Induction hypothesis} \\
-  &{\bf let \: pure \:} \vec{x} = (\phi(\vec{M}) [y := \phi(P)]) {\: \bf in \:} \phi(N) =& \text{Substitution definition} \\
+  &\phi(({\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N) [y := P]) = &\\
+  &\quad\quad\quad\quad\quad\quad \text{Substitution definition}& \\
+  &\phi({\bf let \: pure \:} \vec{x} = (\vec{M}[y := P]) {\: \bf in \:} N) = &\\
+  &\quad\quad\quad\quad\quad\quad \text{By the definition of $\phi$}& \\
+  &{\bf let \: pure \:} \vec{x} = \phi(\vec{M}[y := P]) {\: \bf in \:} \phi(N) = &\\
+  &\quad\quad\quad\quad\quad\quad \text{Induction hypothesis}& \\
+  &{\bf let \: pure \:} \vec{x} = (\phi(\vec{M}) [y := \phi(P)]) {\: \bf in \:} \phi(N) =&\\
+  &\quad\quad\quad\quad\quad\quad \text{Substitution definition} & \\
   &({\bf let \: pure \:} \vec{x} = \phi(\vec{M}) {\: \bf in \:} \phi(N))[y := \phi(P)]
@@
 
-Strong endofunctor on a monoidal category is with a natural transformation so called tensorial strength
-
-\begin{equation}
+Strong functor on a monoidal category is an endofunctor equipped with a natural transformation so called tensorial strength:
+
+\begin{center}
+$\begin{array}{lll}
   \tau_{A, B} : A \otimes \mathcal{K}B \rightarrow \mathcal{K}(A \otimes B)
-\end{equation}
+\end{array}$
+\end{center}
 
@@
 \xymatrix{
-&&& \mathds{1} \otimes \mathcal{K}A \ar[drr]_{\rho_{\mathcal{K}A}} \ar[rr]^{\mu_{\mathds{1}, A}} && \mathcal{K}(\mathds{1} \otimes A) \ar[d]^{\mathcal{K}(\rho_A)}\\
+&&& \mathds{1} \otimes \mathcal{K}A \ar[drr]_{R_{\mathcal{K}A}} \ar[rr]^{\mu_{\mathds{1}, A}} && \mathcal{K}(\mathds{1} \otimes A) \ar[d]^{\mathcal{K}(R_A)}\\
 &&&&& \mathcal{K}A
@@
 \begin{prooftree}
-\AxiomC{$[\![\Gamma, x : A \vdash M : B]\!] = f : [\![\Gamma]\!] \times [\![A]\!] \rightarrow [\![B]\!]$}
-\UnaryInfC{$[\![\Gamma \vdash (\lambda x. M) : A \to B]\!] = \Lambda(f) : [\![\Gamma]\!]
+\AxiomC{$[\![\Gamma, x : A \vdash M : B]\!] = [\![M]\!] : [\![\Gamma]\!] \times [\![A]\!] \rightarrow [\![B]\!]$}
+\UnaryInfC{$[\![\Gamma \vdash (\lambda x. M) : A \to B]\!] = \Lambda([\![M]\!]) : [\![\Gamma]\!]
 \rightarrow[\![B]\!]^{[\![A]\!]}$}
@@
 \begin{prooftree}
-\AxiomC{$[\![\Gamma \vdash M : A ]\!] = f : [\![\Gamma]\!] \rightarrow [\![A]\!]$}
-\AxiomC{$[\![\Gamma \vdash N : B ]\!] = g : [\![\Gamma]\!] \rightarrow [\![B]\!]$}
-\BinaryInfC{$[\![\Gamma \vdash (M, N) : A \times B]\!] = \langle f, g \rangle : [\![\Gamma]\!] \rightarrow
+\AxiomC{$[\![\Gamma \vdash M : A ]\!] = [\![M]\!] : [\![\Gamma]\!] \rightarrow [\![A]\!]$}
+\AxiomC{$[\![\Gamma \vdash N : B ]\!] = [\![N]\!] : [\![\Gamma]\!] \rightarrow [\![B]\!]$}
+\BinaryInfC{$[\![\Gamma \vdash \langle M, N \rangle : A \times B]\!] = \langle [\![M]\!], [\![N]\!] \rangle : [\![\Gamma]\!] \rightarrow
 [\![A]\!] \times [\![B]\!]$}
@@
 \begin{prooftree}
-\AxiomC{$[\![\Gamma \vdash p : A_1 \times A_2]\!] = f : [\![\Gamma]\!] \rightarrow [\![A_1]\!] \times
+\AxiomC{$[\![\Gamma \vdash M : A_1 \times A_2]\!] = [\![M]\!] : [\![\Gamma]\!] \rightarrow [\![A_1]\!] \times
 [\![A_2]\!]$}
 \RightLabel{$i \in \{1,2\}$}
-\UnaryInfC{$[\![\Gamma \vdash \pi_i p : A_i]\!] = [\![\Gamma]\!] \xrightarrow{f} [\![A_1]\!] \times
+\UnaryInfC{$[\![\Gamma \vdash \pi_i M : A_i]\!] = [\![\Gamma]\!] \xrightarrow{[\![M]\!]} [\![A_1]\!] \times
 [\![A_2]\!] \xrightarrow{\pi_i} [\![A_i]\!]$}
@@
 $\begin{array}{lll}
-&[\![\vdash {\bf let \: pure \:} \underline{\quad} = \underline {\quad} {\: \bf in \:} M : {\bf K} A]\!] = & \text{By interpretation} \\
-&\mathcal{K}([\![M]\!]) \circ u_{\mathds{1}} = & \text{By the definition of an applicative functor} \\
-&\mathcal{K}([\![M]\!]) \circ \eta_{\mathds{1}} = & \text{By naturality of $\eta$} \\
-&\eta_{[\![A]\!]} \circ [\![M]\!] = & \text{By interpretation} \\
+&[\![\vdash {\bf let \: pure \:} \underline{\quad} = \underline {\quad} {\: \bf in \:} M : {\bf K} A]\!] = & \\
+&\quad\quad\quad\quad\quad\quad \text{By interpretation}& \\
+&\mathcal{K}([\![M]\!]) \circ u_{\mathds{1}} = & \\
+&\quad\quad\quad\quad\quad\quad \text{By the definition of an applicative functor}& \\
+&\mathcal{K}([\![M]\!]) \circ \eta_{\mathds{1}} = & \\
+&\quad\quad\quad\quad\quad\quad \text{By naturality of $\eta$} & \\
+&\eta_{[\![A]\!]} \circ [\![M]\!] = &\\
+&\quad\quad\quad\quad\quad\quad \text{By interpretation} & \\
 &[\![\vdash {\bf pure \:} M : {\bf K} A]\!]&
@@
 $\begin{array}{lll}
-&\eta_{\hat{B}} \circ f = & \text{By the definition} \\
-&[y, {\bf pure \:} y] \circ [x, M] = & \text{By the definition of composition}\\
-&[x, {\bf pure \:} y [y := M]] = & \text{By substitution}\\
+&\eta_{\hat{B}} \circ f = & \\
+&\quad\quad\quad\quad\quad\quad \text{By the definition}&\\
+&[y, {\bf pure \:} y] \circ [x, M] = &\\
+&\quad\quad\quad\quad\quad\quad \text{By the definition of composition} & \\
+&[x, {\bf pure \:} y [y := M]] = &\\
+&\quad\quad\quad\quad\quad\quad \text{By substitution}& \\
 &[x, {\bf pure \:} M]& \\
@@
 &\text{On the other hand:}& \\
-&\text{fmap } f \circ \eta_{\hat{A}} = & \text{By the definiton} \\
-&[z, {\bf let \: pure \:} x = z {\: \bf in \:} M] \circ [x, {\bf pure \: x}] = & \text{By the definition of composition}\\
-&[x, {\bf let \: pure \:} x = z {\: \bf in \:} M [z := {\bf pure \:} x]] = & \text{By substitution} \\
-&[x, {\bf let \: pure \:} x = {\bf pure \:} x {\: \bf in \:} M] = & \text{$\beta$-reduction rule}\\
-&[x, {\bf pure \:} M [x := x]] = & \text{By substitution}\\
+&\text{fmap } f \circ \eta_{\hat{A}} = & \\
+&\quad\quad\quad\quad\quad\quad \text{By the definiton}& \\
+&[z, {\bf let \: pure \:} x = z {\: \bf in \:} M] \circ [x, {\bf pure \: x}] = &\\
+&\quad\quad\quad\quad\quad\quad \text{By the definition of composition}& \\
+&[x, {\bf let \: pure \:} x = z {\: \bf in \:} M [z := {\bf pure \:} x]] = & \\
+&\quad\quad\quad\quad\quad\quad \text{By substitution}& \\
+&[x, {\bf let \: pure \:} x = {\bf pure \:} x {\: \bf in \:} M] = & \\
+&\quad\quad\quad\quad\quad\quad \text{$\beta$-reduction rule}& \\
+&[x, {\bf pure \:} M [x := x]] = &\\
+&\quad\quad\quad\quad\quad\quad \text{By substitution} & \\
 &[x, {\bf pure \:} M]&
@@
 $\begin{array}{lll}
-&\ast_{\hat{A}, \hat{B}} \circ [p, \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle] = & \text{By definition} \\
-&[p^{'}, {\bf let \: pure \:} x, y = \pi_1 p^{'}, \pi_2 p^{'} {\: \bf in \:} \langle x, y \rangle] \circ [p, \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle] = & \text{By composition} \\
-&[p, {\bf let \: pure \:} x, y = \pi_1 p^{'}, \pi_2 p^{'} {\: \bf in \:} \langle x, y \rangle [p^{'} := \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle]] = & \text{By substitution} \\
-&[p, {\bf let \: pure \:} x, y = \pi_1 (\langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle), \pi_2 (\langle \pi_1 p, {\bf pure \:} (\pi_2 p) \rangle) {\: \bf in \:} \langle x, y \rangle] = & \text{By $\beta$-reduction rules} \\
+&\ast_{\hat{A}, \hat{B}} \circ [p, \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle] = & \\
+&\quad\quad\quad\quad\quad\quad \text{By definition}&\\
+&[p^{'}, {\bf let \: pure \:} x, y = \pi_1 p^{'}, \pi_2 p^{'} {\: \bf in \:} \langle x, y \rangle] \circ [p, \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle] = & \\
+&\quad\quad\quad\quad\quad\quad \text{By composition} & \\
+&[p, {\bf let \: pure \:} x, y = \pi_1 p^{'}, \pi_2 p^{'} {\: \bf in \:} \langle x, y \rangle [p^{'} := \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle]] = & \\
+&\quad\quad\quad\quad\quad\quad \text{By substitution} & \\
+&[p, {\bf let \: pure \:} x, y = \pi_1 (\langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle), \pi_2 (\langle \pi_1 p, {\bf pure \:} (\pi_2 p) \rangle) {\: \bf in \:} \langle x, y \rangle] = & \\
+&\quad\quad\quad\quad\quad\quad \text{By $\beta$-reduction rules} & \\
 &[p, {\bf let \: pure \:} x, y = {\bf pure \:} (\pi_1 p), \pi_2 p {\: \bf in \:} \langle x, y \rangle]&
@@
     \item $\text{fmap } \alpha_{\hat{A},\hat{B},\hat{C}} \circ \tau_{\hat{A} \times \hat{B}, \hat{C}} = \tau_{\hat{A}, \hat{B} \times \hat{C}} \circ (id_{\hat{A}} \times \tau_{\hat{B}, \hat{C}}) \circ \alpha_{\hat{A}, \hat{B}, {\bf K}\hat{C}}$;
-    \item ${\bf K}(\rho_{\hat{A}}) \circ \tau_{\mathds{1}, \hat{A}} = \rho_{{\bf K}\hat{A}}$.
+    \item ${\bf K}(R_{\hat{A}}) \circ \tau_{\mathds{1}, \hat{A}} = R_{{\bf K}\hat{A}}$.
   \end{itemize}
   where $\alpha_{\hat{A},\hat{B},\hat{C}} = [p, \langle \pi_1 (\pi_1 p), \langle \pi_2 (\pi_1 p), \pi_2 p \rangle \rangle]$ and
-  $\rho = \pi_2$.
+  $R = \pi_2$.
 \end{lemma}
@@
 
+\section{Acknowledgment}
+
+  We are grateful to Valeria de Paiva, Neel Krishnaswami and Vladimir Vasyukov for consulting and advice.
+
+  Special thanks to Vladimir Krupski for general idea, critics and support.
+
   \addcontentsline{toc}{section}{References}

### 2018-02-24 MT/CategoryFAQ.tex
started russian version

--- 2018-02-14 article.tex
@@
-\documentclass[a4paper]{article}
-\usepackage{amsmath}
-\usepackage{amsthm}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{bussproofs}
-\usepackage{mathtools}
-\usepackage{verbatim}
-\usepackage{dsfont}
-\usepackage{mathabx}
-\usepackage[all, 2cell]{xy}
-\usepackage[all]{xy}
-\newtheorem{theorem}{Theorem}
-\newtheorem{prop}{Proposition}
-\newtheorem{lemma}{Lemma}
-\newtheorem{defin}{Definition}
-\newtheorem{ex}{Example}
-\newtheorem{col}{Corollary}
-\newtheorem{con}{Consequence}
-\usepackage{listings} 		% for source code
-\date{}
-\title{Modal type theory based on the intuitionistic epistemic logic}
+\section{Приложение А. Глоссарий по теории категорий.}
 
-\begin{document}
-\maketitle
+\begin{defin}
 
-\begin{abstract}
-Modal intuitionistic epistemic logic IEL$^{-}$ was proposed by S.Artemov and T. Protopopescu as the formal
-foundation for the intuitionistic theory of knowledge. We construct a modal simply typed lambda-calculus
-which is Curry-Howard isomorphic to IEL$^{-}$ as formal theory of calculations with applicative functors in
-functional programming languages like Haskell or Idris.
-\end{abstract}
-
-  % This is a comment, not shown in final output.
-  % The following shows typesetting  power of LaTeX:
-
-  \section{Introduction}
-
-  Modal intutionistic epistemic logic IEL was proposed by S. Artemov and T. Proropopescu \cite{Artemov}. IEL
-  provides the epistimology and the theory of knowledge as based on BHK-semantics of intuitionistic logic.
-  IEL$^{-}$  is a variant of IEL, that corresponds to intuitionistic belief.
-  Informally, ${\bf K}A$ denotes that $A$ is verified intuitionistically.
-
-  Intuitionistic epistemic logic IEL$^{-}$ is defined by following axioms and derivation rules:
-
-  \begin{defin} Intuitionistic epistemic logic IEL:
-
-    1) IPC axioms;
-
-    2) ${\bf K}(A \to B) \to ({\bf K}A \to {\bf K}B)$ (normality);
-
-    3) $A \to {\bf K}A$ (co-reflection);
-
-    Rule: MP.
-
-  \end{defin}
-
-  V. Krupski and  A. Yatmanov provided the sequential calculus for IEL and proved that this calculus is
-  PSPACE-complete \cite{Krupski}.
-
-  Functional programming languages such as Haskell \cite{Haskell}, Idris \cite{Idris}, Purescript \cite{Purs}
-  Elm \cite{Elm} or Scala \cite{Scala} have special type classes\footnote{Type class in Haskell is a general interface for
-  special group of datatypes.} for calculations with container types like \verb"Functor" and
-  \verb"Applicative" \footnote{Reader may read more about container types in the Haskell standard library documentation\cite{Base} or in the next one textbook
-  \cite{Miran}}:
-
-  \begin{lstlisting}[language=Haskell]
-    class Functor f where
-      fmap :: (a -> b) -> f a -> f b
-
-    class Functor f => Applicative f where
-      pure :: a -> f a
-      (<*>) :: f (a -> b) -> f a -> f b
-  \end{lstlisting}
-
-  By \emph{container} (or \emph{computational context}) type we mean some type-operator $f$, where $f$ is a
-  ``function'' from $*$ to $*$: type operator takes a simple type (ща kind $*$) and returns another
-  simple type of kind \verb"*". For more detailed description of the type system with kinds used in
-  Haskell see \cite{Morten}.
-
-Applicative functor allows to generalize the action of a functor for functions with arbitrary number of arguments, for instance:
-  \begin{lstlisting}[language=Haskell]
-  liftA2 :: Applicative f => (a -> b -> c) -> f a -> f b -> f c
-  liftA2 f x y = ((pure f) <*> x) <*> y
-  \end{lstlisting}
-
-  It's not difficult to see that modal axioms in $IEL^{-}$ and types of the methods of Applicative class in
-  Haskell-like languages (which is described below) are syntactically similar and we are going to show that
-  this coincidence has a non-trivial computational meaning.
-
-  We investigate the relationship between intuitionistic epistemic logic IEL$^{-}$ and applicative programming with side-effects
-  by constructing the type system (which is called $\lambda_{{\bf K}}$) which is Curry-Howard isomorphic to $IEL^{-}$.
-  So we will consider ${\bf K}$-modality as an arbitrary applicative functor and we prove that obtained type system
-  is sound and complete for applicative functor on cartesian closed category (using the categorical definition proposed by Paterson \cite{Cons}).
-
-  $\lambda_{{\bf K}}$ consists of the rules for simply typed lambda-calculus and special typing rules for
-  lifting types into the applicative functor ${\bf K}$. We assume that our type system will
-  axiomatize the simplest case of computation with effects with one container. We provide a proof-theoretical
-  view at this kind of computations in functional programming and prove strong normalization and confluence.
-
-  \section{Typed lambda-calculus based on IEL$^{-}$}
-
-  The first is to define the natural deduction calculus for NIEL$^{-}$ :
-
-  \begin{defin} Natural deduction NIEL$^{-}$ for IEL$^{-}$ is an extenstion of intuitionistic natural deduction calculus
-    with additional inference rules for modality:
-
-  \begin{minipage}{0.5\textwidth}
-    \begin{flushleft}
-  	\begin{prooftree}
-    	\AxiomC{$\Gamma \vdash A$}
-    	\RightLabel{${\bf K}_I$}
-    	\UnaryInfC{$\Gamma \vdash {\bf K}A$}
-  \end{prooftree}
-    \end{flushleft}
-  \end{minipage}
-  \begin{minipage}{0.5\textwidth}
-    \begin{flushright}
-  	\begin{prooftree}
-    \AxiomC{$\Gamma \vdash {\bf K} A_1, \dots, \Gamma \vdash {\bf K} A_n $}
-    \AxiomC{$A_1,\dots,A_n \vdash B$}
-    \BinaryInfC{$\Gamma \vdash {\bf K} B$}
-  	\end{prooftree}
-    \end{flushright}
-  \end{minipage}
-  \end{defin}
-
-The first rule allows to derive co-reflextion. The second modal rule is a counterpart of $\Box_I$
-rule in natural deduction calculus for constructive K (see \cite{ModalLa}).
-
-We will denote $\Gamma \vdash {\bf K} A_1, \dots, \Gamma \vdash {\bf K} A_n$ and $A_1,\dots,A_n \vdash B$ as $\Gamma \vdash {\bf K} \vec{A}$ and $\vec{A} \vdash B$ for brevity.
-
-  \vspace{\baselineskip}
-
-  \begin{lemma}
-    $\Gamma \vdash_{\text{NIEL}^{-}} A \Rightarrow$ IEL$^{-} \vdash \bigwedge \Gamma \rightarrow A$.
-  \end{lemma}
-
-  \begin{proof}
-Induction on the derivation.
-
-  \vspace{\baselineskip}
-
-Let us consider cases with modality.
-
-\vspace{\baselineskip}
-
-1) If $\Gamma \vdash_{\text{NIEL}^{-}} A$, then $\text{IEL}^{-} \vdash \bigwedge \Gamma \rightarrow {\bf K}A$.
-
-$\begin{array}{lll}
-(1) & \bigwedge \Gamma \rightarrow A & \text{assumption}\\
-(2) & A \rightarrow {\bf K}A &\text{co-reflection}\\
-(3) & (\bigwedge \Gamma \rightarrow A) \rightarrow ((A \rightarrow {\bf K}A) \rightarrow (\bigwedge \Gamma \rightarrow {\bf K}A))&\text{IPC theorem}\\
-(4) & (A \rightarrow {\bf K}A) \rightarrow (\bigwedge \Gamma \rightarrow {\bf K}A) &\text{from (1), (3) and MP}\\
-(5) & \bigwedge \Gamma \rightarrow {\bf K}A &\text{from (2), (4) and MP}\\
-\end{array}$
-
-\vspace{\baselineskip}
-
-2) If $\Gamma \vdash_{\text{NIEL}^{-}} {\bf K} \vec{A}$ and $\vec{A} \vdash B$, then $\text{IEL}^{-} \vdash \bigwedge \Gamma \rightarrow {\bf K}B$.
-
-$\begin{array}{lll}
-(1) &\bigwedge \Gamma \rightarrow \bigwedge \limits_{i = 1}^{n} {\bf K} A_i & \text{assumption} \\
-(2) &\bigwedge \limits_{i = 1}^{n} {\bf K} A_i \rightarrow {\bf K} \bigwedge \limits_{i = 1}^{n} A_i& \text{IEL$^{-}$ theorem} \\
-(3) &\bigwedge \Gamma \rightarrow {\bf K} \bigwedge \limits_{i = 1}^{n} A_i & \text{from (1), (2) and transitivity} \\
-(4) &\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B& \text{assumption} \\
-(5) &(\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B) \rightarrow {\bf K} (\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B)& \text{co-reflection}\\
-(6) &{\bf K} (\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B)& \text{from (4), (5) and MP} \\
-(7) &{\bf K} \bigwedge \limits_{i = 1}^{n} A_i \rightarrow {\bf K} B & \text{from (6) and normality} \\
-(8) &\bigwedge \Gamma \rightarrow {\bf K} B & \text{from (3), (7) and transitivity}
-\end{array}$
-
-  \end{proof}
-
-\begin{lemma}
-  $ $
-  If $\text{IEL}^{-} \vdash A$, then $\text{NIEL}^{-} \vdash A$.
-\end{lemma}
-
-\begin{proof}
-  Straightforward derivation of modal axioms in $\text{NIEL}^{-}$. We consider this derivation below using terms.
-\end{proof}
-
-  \vspace{\baselineskip}
-
-  At the next step we build the typed lambda-calculus based on the NIEL$^{-}$ with implication and dijunction
-by proof-assingment in rules. Obtained fragment is equivalent to IEL$^{-}$ without axioms for negation and disjunction.
-
-  At first, we define lambda-terms and types for this lambda-calculus.
-
-  \vspace{\baselineskip}
-
-  \begin{defin} The set of terms:
-
-  Let $\mathbb{V}$ be the set of variables. The set $\Lambda_{{\bf K}}$ of terms is defined by the grammar:
-
-$\begin{array}{lll}
-& \Lambda_{{\bf K}} ::= \mathbb{V} \: | \:  (\lambda \mathbb{V}.\Lambda_{{\bf K}}) \: | \: (\Lambda_{{\bf K}}\Lambda_{{\bf K}}) \: | \: (\Lambda_{{\bf K}} , \Lambda_{{\bf K}}) \: | \: (\pi_1 \Lambda_{{\bf K}}) \: | \: (\pi_2 \Lambda_{{\bf K}}) \: | & \\
-& \quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad ({\bf pure \: } \: \Lambda_{{\bf K}}) \: | \: ({\bf let \:\: pure} \:\: \mathbb{V}^{*} = \Lambda_{{\bf K}}^{*} \:\: {\bf in} \:\: \Lambda_{{\bf K}})
-\end{array}$
-
-  \end{defin}
-
-Where $\mathbb{V}^{*}$ and $\Lambda_{{\bf K}}^{*}$ denote the set of finite sequences of variables $\bigcup \limits_{i=0}^{\infty} \mathbb{V}^i$
-and the set of finite sequences of terms $\bigcup \limits_{i = 0}^{\infty} {\Lambda_{{\bf K}}}^i $.
-Note that the sequence of variables $\vec{x}$ and the sequence of terms $\vec{M}$ should have the same length.
-Otherwise, term is not well-formed.
-
-  \begin{defin} The set of types:
-
-  Let $\mathbb{T}$ be the set of atomic types. The set $\mathbb{T}_{{\bf K}}$ of types with applicative
-  functor ${\bf K}$ is generated by the grammar:
-  \begin{equation}
-    \mathbb{T}_{{\bf K}} ::= \mathbb{T} \: | \: (\mathbb{T}_{{\bf K}} \to \mathbb{T}_{{\bf K}}) \: |
-    \: (\mathbb{T}_{{\bf K}} \times \mathbb{T}_{{\bf K}}) \: | \: ({\bf K}\mathbb{T}_{{\bf K}})
-  \end{equation}
-  \end{defin}
-
-  Context, domain of context and range of context are defined standardly \cite{Neder}\cite{Morten}.
-
-  Our type system is based on the Curry-style typing rules:
-
-  \begin{defin} Modal typed lambda calculus $\lambda_{{\bf K}}$ based on NIEL$^{-}_{\land, \to}$:
-
-    \begin{center}
-    \begin{prooftree}
-    \AxiomC{$ $}
-    \RightLabel{\scriptsize{ax}}
-    \UnaryInfC{$\Gamma , x : A \vdash x : A$}
-    \end{prooftree}
-    \end{center}
-
-    \begin{minipage}{0.45\textwidth}
-      \begin{prooftree}
-      \AxiomC{$\Gamma, x : A \vdash M : B$}
-      \RightLabel{$\rightarrow_i$}
-      \UnaryInfC{$\Gamma \vdash \lambda x. M : A \to B$}
-      \end{prooftree}
-
-      \begin{prooftree}
-      \AxiomC{ $\Gamma \vdash M : A$ }
-      \AxiomC{ $\Gamma \vdash N : B$ }
-      \RightLabel{$\times_i$}
-      \BinaryInfC{$\Gamma \vdash \langle M, N \rangle : A \times B$}
-      \end{prooftree}
-
-      \begin{prooftree}
-        \AxiomC{$\Gamma \vdash M : A$}
-        \RightLabel{${\bf K}_I$}
-        \UnaryInfC{$\Gamma \vdash {\bf pure \: } \: M : {\bf K}A $}
-      \end{prooftree}
-  \end{minipage}%
-  \hfill
-  \begin{minipage}{0.45\textwidth}
-  \begin{tabular}{p{\textwidth}}
-    \begin{prooftree}
-    \AxiomC{$\Gamma \vdash M : A \to B$}
-    \AxiomC{$\Gamma \vdash N : A$}
-    \RightLabel{$\rightarrow_e$}
-    \BinaryInfC{$\Gamma \vdash MN : B$}
-    \end{prooftree}
-
-    \begin{prooftree}
-    \AxiomC{ $\Gamma \vdash M : A_1 \times A_2$ }
-    \RightLabel{$\times_e$, $i \in \{ 1, 2 \}$}
-    \UnaryInfC{$\Gamma \vdash \pi_i M : A_i$}
-    \end{prooftree}
-
-    \begin{prooftree}
-      \AxiomC{$\Gamma \vdash \vec{M} : {\bf K} \vec{A}$}
-      \AxiomC{$\vec{x} : \vec{A} \vdash N : B$}
-      \RightLabel{$\text{let}_{{\bf K}}$}
-      \BinaryInfC{$\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } N : {\bf K} B$}
-    \end{prooftree}
-  \end{tabular}
-  \end{minipage}%
-
-  \end{defin}
-
-  ${\bf K}_I$-typing rule is the same as $\bigcirc$-introduction in lax logic (also known as monadic
-  metalanguage \cite{Lax}) and in typed lambda-calculus which is derived by proof-assignment for lax-logic
-  proofs.
-  ${\bf K}_I$ allows to inject an object of type $\alpha$ into the functor. ${\bf K}_I$  reflects the
-  Haskell method {\bf pure} for Applicative class. It plays the same role as the {\bf return} method in Monad class.
-
-  $\text{let}_{{\bf K}}$ is similar to the $\Box$-rule in typed lambda calculus for intuitionistic normal modal logic {\bf IK}, which is described in \cite{ModalK}.
-
-$\Gamma \vdash \vec{M} : {\bf K} \vec{A}$ is a syntax sugar for the sequence $\Gamma \vdash M_1 : {\bf K}A_1,\dots,\Gamma \vdash M_n : {\bf K}A_n$ and $\vec{x} : \vec{A} \vdash N : B$ is a short form for $x_1 : A_1, \dots, x_n : A_n \vdash N : B$.
-${\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } N$ is a simultaneous local binding in $N$.
-We use this short form instead of ${\bf let \: pure\:} x_1,\dots,x_n = M_1,\dots,M_n {\: \bf in \:} N$.
-
-In fact, our calculus is the extention of typed lambda calculus for {\bf IK} by ${\bf K}_I$-rule that is appropriate to co-reflection.
-
-  \vspace{\baselineskip}
-
-Here are some examples of closed terms:
-
-\begin{prooftree}
-\AxiomC{$x : A \vdash x : A$}
-\UnaryInfC{$x : A \vdash {\bf pure \:} x : {\bf K} A$}
-\UnaryInfC{$\vdash (\lambda x. {\bf pure \: } x) : A \to {\bf K}A$}
-\end{prooftree}
-
-\begin{prooftree}
-\AxiomC{$f : {\bf K} (A \to B) \vdash f : {\bf K} (A \to B)$}
-\AxiomC{$x : {\bf K} A \vdash x : {\bf K} A $}
-\AxiomC{$g : A \to B \vdash g : A \to B$}
-\AxiomC{$y : A \vdash y : A$}
-\RightLabel{$\to_e$}
-\BinaryInfC{$g : A \to B, y : A \vdash g y : B$}
-\RightLabel{$\text{let}_{{\bf K}}$}
-\TrinaryInfC{$f : {\bf K} (A \to B), x : {\bf K} A \vdash {\bf let \: pure \:} g, y  = f, x {\: \bf in \:} g y : {\bf K} B$}
-\UnaryInfC{$f : {\bf K} (A \to B) \vdash \lambda x. {\bf let \: pure\:} g, y = f, x {\: \bf in \:} g y : {\bf K} A \to {\bf K} B$}
-\UnaryInfC{$\vdash \lambda f. \lambda x. {\bf let \:pure \:} g, y = f, x {\: \bf in \:} g y : {\bf K}(A \to B) \to {\bf K} A \to {\bf K} B$}
-\end{prooftree}
-
-  \vspace{\baselineskip}
-
-  Now we define free variables and substitutions. $\beta$-reduction, multi-step $\beta$-reduction and $\beta$-equality are defined standardly:
-
-  \begin{defin} The set $FV(M)$ of free variables for a term $M$:
-
-  1) $FV(x) = \{ x \}$;
-
-  2) $FV(\lambda x. M) = FV(M) \setminus \{ x\}$;
-
-  3) $FV(M N) = FV(M) \cup FV(N);$
-
-  4) $FV(\langle M,N \rangle) = FV(M) \cup FV(N)$;
-
-  5) $FV(\pi_i M) \subseteq FV(M)$, $i \in \{ 1, 2\}$;
-
-  6) $FV(\text{pure } M) = FV(M)$;
-
-  7) $FV({\bf let \: pure} \: \vec{x} = \vec{M} \:\: {\bf in} \:\: N) = \bigcup \limits_{i = 1}^n FV(M), \text{where $n = |\vec{M}|$}$.
-  \end{defin}
-
-  \begin{defin} Substitution:
-
-  1) $x [x := N] = N$, $x [y := N] = x$;
-
-  2) $(M N) [x := N] = M[x := N] N [x := N]$;
-
-  3) $(\lambda x. M) [x := N] = \lambda x. M [y := N]$, $y \in FV(M)$;
-
-  4) $(M, N)[x := P] = (M[x := P], N [x := P])$;
-
-  5) $(\pi_i M) [x := P] = \pi_i (M[x := P])$, $i \in \{ 1, 2\}$;
-
-  6) $({\bf pure \: } M) [x := P] = {\bf pure \: } (M [x := P])$;
-
-  7) $({\bf let \: pure \:}\vec{x} = \vec{M} {\: \bf in \:} N) [y := P] = {\bf let \: pure \:} \vec{x} = (\vec{M} [y := P]) {\: \bf in \:} N$.
-  \end{defin}
-
-Substitution and free variable for terms of kind ${\bf let \: pure}$ are defined similary to \cite{ModalK}.
-
-  \begin{defin} Type substituition
-
-    The substituition of type $C$ for type variable $B$ in type $A$ is defined inductively:
-
-    1) $B [B := C] = B$ and $D [B := C] = D$, if $B \neq D$;
-
-    2) $(A_1 \alpha A_2)[B := C] = (A_1[B := C]) \alpha (A_2[B := C])$, where $\alpha \in \{ \to, \times \}$;
-
-    3) $({\bf K}A)[B := C] = {\bf K}(A [B := C])$;
-
-    4) Let $\Gamma$ be the context, then $\Gamma [B := C] = \{ x : (A[B := C]) \: | \: x : A \in \Gamma \}$.
-  \end{defin}
-
-  \begin{defin} $\beta$-reduction and $\eta$-reduction rules for $\lambda_{{\bf K}}$.
-
-  1) $(\lambda x. M) N \rightarrow_{\beta} M [x := N]$;
-
-  2) $\pi_1 \langle M, N \rangle \rightarrow_{\beta} M$;
-
-  3) $\pi_2 \langle M, N \rangle \rightarrow_{\beta} N$;
-
-  4) $\begin{array}{llll}
-  & {\bf let \: pure \:} \vec{x}, y, \vec{z} = \vec{M}, {\bf let \: pure \: } \vec{w} = \vec{N} {\: \bf in \: } Q, \vec{P} {\: in \:} R \rightarrow_{\beta} & \\
-  & {\bf let \: pure \:} \vec{x}, \vec{w}, \vec{z} = \vec{M}, \vec{N}, \vec{P} {\: \bf in \: } R [y := Q]
-  \end{array}$
-
-  5) ${\bf let \: pure \:} \vec{x} = {\bf pure \:} \vec{M} {\: \bf in \:} N \rightarrow_{\beta} {\bf pure \:} N [\vec{x} := \vec{M}]$
-
-  6) ${\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} M \rightarrow_{\beta} {\bf pure \:} M$, where \underline{\quad} is an empty sequence of terms.
-
-  7) $\lambda x. f x \rightarrow_{\eta} f$;
-
-  8) $\langle \pi_1 P, \pi_2 P \rangle \rightarrow_{\eta} P$;
-
-  9) ${\bf let \: pure \:} x = M {\: \bf in \: } x \rightarrow_{\eta} M$;
-
-  \end{defin}
-
-By default we use call-by-name evaluation strategy.
-
-This commuting relation denotes that we identify terms of kind {\bf let pure} if they are differ in the order of binding.
-
-Now we will prove standard lemmas for contexts in type systems\footnote{We will not prove cases with
-$\to$-constructor, they are proved standardly in the same lemmas for simply typed lambda calculus, for
-example see \cite{Neder} \cite{Morten}. We will consider only modal cases}:
-
-\begin{lemma} Generation for ${\bf K}_I$.
-
-  Let $\Gamma \vdash {\bf pure \:} M : {\bf K}A$, then $\Gamma \vdash M : A$;
-\end{lemma}
-
-\begin{proof}
-  Straightforwardly.
-\end{proof}
-
-\begin{lemma} Basic lemmas.
-
+  Категория $\mathcal{C}$ состоит из:
   \begin{itemize}
-  \item If $\Gamma \vdash M : A$ and $\Gamma \subseteq \Delta$, then $\Delta \vdash M : A$;
-  \item If $\Gamma \vdash M : A$, then $\Delta \vdash M : A$, where $\Delta = \{ x_i : A_i \: | \: (x_i : A_i) \in \Gamma \: \& \: x_i \in FV(M) \}$
-  \item If $\Gamma, x : A \vdash M : B$ and $\Gamma \vdash N : A$, then $\Gamma \vdash M [x := N] : B$.
-  \item If $\Gamma \vdash M : A$, then $\Gamma [B := C] \vdash M : (A [B := C])$;
-  \item If $\Gamma \vdash M : A$ and $\Delta$ is a permutation of $\Gamma$, then $\Delta \vdash M : A$.
-\end{itemize}
-\end{lemma}
-
-\begin{proof}
-  $ $
-
-1) The derivation ends in
-
-\begin{prooftree}
-  \AxiomC{$\Gamma \vdash \vec{M} : {\bf K} \vec{A}$}
-  \AxiomC{$\vec{x} : \vec{A} \vdash N : B$}
-  \RightLabel{$\text{let}_{{\bf K}}$}
-  \BinaryInfC{$\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } N : {\bf K}B$}
-\end{prooftree}
-
-By IH $\Delta \vdash \vec{M} : {\bf K} \vec{A}$, so $\Delta \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } N : {\bf K}B$.
-
-  \vspace{\baselineskip}
-
-2) Similary.
-
-  \vspace{\baselineskip}
-
-3) The derivation ends in
-
-\begin{prooftree}
-  \AxiomC{$\Gamma, z : C \vdash \vec{M} : {\bf K} \vec{A}$}
-  \AxiomC{$\vec{x} : \vec{A} \vdash N : B$}
-  \RightLabel{$\text{let}_{{\bf K}}$}
-  \BinaryInfC{$\Gamma, z : C \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } N : {\bf K}B$}
-\end{prooftree}
-
-Let $\Gamma \vdash P : C$.
-
-By IH, $\Gamma \vdash \vec{M} [z := P] : {\bf K} \vec{A}$. So $\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} [z := P] {\: \bf in \: } N : {\bf K}B$
-
-  \vspace{\baselineskip}
-
-4-5) Similary.
-
-\end{proof}
-
-\begin{theorem} Subject reduction
-
-  If $\Gamma \vdash M : A$ and $M \twoheadrightarrow_{\beta \eta} N$, then $\Gamma \vdash N : A$
-
-\end{theorem}
-
-\begin{proof}
-
-Induction on the derivation $\Gamma \vdash M : A$ and on the generation of $\rightarrow_{\beta \eta}$.
-
-For cases with application, abstraction and pairs see \cite{Morten} \cite{Pierce}.
-
-1) If $\Gamma \vdash {\bf let \: pure \:} \vec{x}, y, \vec{z} = \vec{M}, {\bf let \: pure \: } \vec{w} = \vec{N} {\bf \: in \: } Q, \vec{P}  {\: \bf in \:} R : {\bf K}B$,
-  then $\Gamma \vdash {\bf let \: pure \:} \vec{x}, \vec{w}, \vec{z} = \vec{M}, \vec{N}, \vec{P} \: { \bf in } \: R [y := Q] : {\bf K}B$ by rule 4).
-
-2) Let $\Gamma \vdash {\bf let \: pure \:} x = M {\: \bf in \: } x : {\bf K}A$, then $\Gamma \vdash M : {\bf K}A$ by rule 9).
-
-See \cite{ModalK}.
-
-3) The derivation ends in
-
-\begin{prooftree}
-\AxiomC{$\Gamma \vdash {\bf pure \:} \vec{M} : {\bf K}\vec{A}$}
-\AxiomC{$\vec{x} : \vec{A} \vdash N : B$}
-\BinaryInfC{$\Gamma \vdash {\bf let \: pure \:} \vec{x} = {\bf pure \:} \vec{M} {\: \bf in \:} N : {\bf K}B$}
-\end{prooftree}
-
-So $\Gamma \vdash \vec{M} : \vec{A}$ by Lemma 4 and $\Gamma \vdash N [\vec{x} := \vec{M}] : B$ by Lemma 4, part 3.
-
-Then we can transform this into the following derivation:
-
-\begin{prooftree}
-\AxiomC{$\Gamma \vdash N [\vec{x} := \vec{M}] : B$}
-\RightLabel{${\bf K}_I$}
-\UnaryInfC{$\Gamma \vdash {\bf pure \:} N [\vec{x} := \vec{M}] : {\bf K}B$}
-\end{prooftree}
-
-4) The derivation ends in
-
-\begin{prooftree}
-\AxiomC{$\vdash M : A$}
-\UnaryInfC{$\vdash {\bf let \: pure \:} \underline{\quad} = \underline {\quad} {\: \bf in \:} M : {\bf K} A$}
-\end{prooftree}
-
-So, if $\vdash M : A$, then $\vdash {\bf pure \:} M : {\bf K}A$.
-
-Note that this part of the lemma works conversly too.
-\end{proof}
-
-\begin{theorem}
-  $ $
-
-  $\twoheadrightarrow_{\beta}$ is strongly normalizing;
-\end{theorem}
-
-\begin{proof}
-  $ $
-
-We modify and apply Tait's technique of logical relation for modalities. See \cite{Pierce} \cite{Troelstra}.
-
-We treat only modal cases below.
-
-\begin{defin} The set of strongly computable terms:
-  \begin{itemize}
-    \item $SC_A = \{ M : A \: | \: M \text{ is strongly normalizing} \}$ for $A \in \mathbb{T}$;
-    \item $SC_{A \to B} = \{ M : A \to B \: | \: \forall N \in SC_A, M N \in SC_B \}$, for $A,B \in \mathbb{T}_{\bf K}$ for $A, B \in \mathbb{T}_{{\bf K}}$;
-    \item $SC_{{\bf K}A} = \{ M : {\bf K}A \: | \: M \text{ is strongly normalizing} \}$ for $A \in \mathbb{T}$;
-    \item $\forall i \in \{ 1,\dots,n\}, \prod \limits_{i = 1}^{n} SC_{{\bf K}A_i} = \{ \vec{M} = (M_1,\dots,M_n) \:\: | \:\: \forall N \in SC_{B}, FV(N) = \{ x_1,\dots,x_n\} \: \& \: \forall i, x_i \in SC_{A_i} \Rightarrow {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N \in SC_{{\bf K} B}\}$
+    \item Класса объектов $Ob_{\mathcal{C}}$;
+    \item Для любых объекта $A, B \in Ob_{\mathcal{C}}$ определено множество стрелок (или морфизмов) из $A$ в $B$ $Hom_{\mathcal{C}}(A, B)$;
+    \item Если $f \in Hom_{\mathcal{C}}(A, B)$ и $g \in Hom_{\mathcal{C}}(B,C)$, то $g \circ f \in Hom_{\mathcal{C}}(A, C)$;
+    \item Для любого объекта $A \in Ob_{\mathcal{C}}$, определен тождественный морфизм $id_A \in Hom_{\mathcal{C}}(A, A)$;
+    \item Для любой стрелки $f \in Hom_{\mathcal{C}}(A, B)$, для любой стрелки $g \in Hom_{\mathcal{C}}(B,C)$ и для любой стрелки $h \in Hom_{\mathcal{C}}(C,D)$, $h \circ (g \circ f) = (h \circ g) \circ f$.
+    \item Для любой стрелки $f \in Hom_{\mathcal{C}}(A, B)$, $f \circ id_A = f$ и $id_B \circ f = f$.
   \end{itemize}
@@
 
-\begin{defin}
-  A term $M$ is neutral, if it has one of the following forms:
+\begin{defin} Моноидальная категория
+
+  Моноидальная категория -- это категория $\mathcal{C}$ с дополнительной структурой:
   \begin{itemize}
-    \item $M N$;
-    \item If $M$ is neutral, then ${\bf pure \:} M$ is neutral;
-    \item If $\vec{M}$ and $N$ are neutral, then ${\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: N}$ is neutral.
-    $\vec{x}$ is a sequence of free variables of a term $N$.
-  \end{itemize}
-\end{defin}
-
-\begin{lemma}
-  $ $
-
-\begin{itemize}
-\item If $M \in SC_A$ and $A \in \mathbb{T}_{{\bf K}}$, then $M$ is strongly normalizing;
-\item If $M \in SC_A$, $A \in \mathbb{T}_{{\bf K}}$ and $M \rightarrow_{\beta} N$, then $N \in SC_A$;
-\item If $N$ is neutral, $N \in SC_A$. Then, if $M \rightarrow_{\beta} N$, then $M \in SC_A$;
-\end{itemize}
-\end{lemma}
-
-\begin{proof}
-  $ $
-
-  By induction on the structure of $A$.
-
-1) $A \equiv {\bf K}A$, where $A \in \mathbb{T}$.
-
-i-ii-iii) Immediately.
-
-    \vspace{\baselineskip}
-
-2)
-
-i) Suppose $\vec{M} = (M_1,\dots,M_n) \in \prod \limits_{i = 1}^{n} SC_{{\bf K}A_i}$.
-
-Let $N \in SC_{B}$, such that $FV(N) = \{ x_1,\dots, x_n \}$ and $\forall i, x_i \in SC_{A_i}$.
-
-So ${\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } N \in SC_{{\bf K}B}$ by IH.
-
-So $\vec{M}$ are strongly normalizing, since ${\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } N$ is strongly normalizing by IH.
-
-\vspace{\baselineskip}
-
-ii) Let $\vec{M_1} \in \prod \limits_{i = 1}^{n} SC_{{\bf K}A_i}$ and $\vec{M_1} \rightarrow_{\beta} \vec{M_2}$.
-
-Let $N \in SC_{B}$, such that $FV(N) = \{ x_1,\dots,x_n \}$ and $\forall i, x_i \in SC_{A_i}$.
-
-So ${\bf let \: pure \:} \vec{x} = \vec{M_1} {\: \bf in \:} N \rightarrow_{\beta} {\bf let \: pure \:} \vec{x} = \vec{M_2} {\: \bf in \:} N$
-
-and ${\bf let \: pure \:} \vec{x} = \vec{M_2} {\: \bf in \:} N \in SC_{{\bf K}B}$ by assumption.
-
-So $\vec{M_2} \in \prod \limits_{i = 1}^{n} SC_{{\bf K}A_i}$.
-
-iii) Let $M_2$ be neutral, $M_2 \in \prod \limits_{i = 1}^{n} SC_{{\bf K}A_i}$ and $M_1 \rightarrow_{\beta} M_2$.
-
-Let $N \in SC_{B}$, , such that $FV(N) = \{ x_1,\dots,x_n \}$ and $\forall i, x_i \in SC_{A_i}$.
-
-So ${\bf let \: pure \:} \vec{x} = \vec{M_2} {\: \bf in \:} N \in SC_{{\bf K}B}$.
-
-Thus ${\bf let \: pure \:} \vec{x} = \vec{M_1} {\: \bf in \:} N \rightarrow_{\beta} {\bf let \: pure \:} \vec{x} = \vec{M_2} {\: \bf in \:} \in N$.
-
-Hence ${\bf let \: pure \:} \vec{x} = \vec{M_1} {\: \bf in \:} N\in SC_{{\bf K}B}$ by IH, so $\vec{M_1} \in \prod \limits_{i = 1}^{n} SC_{{\bf K}A_i}$.
-\end{proof}
-
-\begin{lemma}
-  $ $
-
-  If $M \in SC_A$, then ${\bf pure \:} M \in SC_{{\bf K}A}$
-\end{lemma}
-
-\begin{proof}
-
-  Induction on the structure of M.
-\end{proof}
-
-\begin{lemma}
-  $ $
-
-  Let $x_1 : A_1,\dots, x_n : A_n \vdash M : A$, then forall $i, M_i \in SC_{A_i}$. Then $M [x_1 := M_1,\dots,x_n := M_n] \in SC_A$.
-\end{lemma}
-
-\begin{proof}
-  $ $
-
-Induction on the derivation of $x_1 : A_1,\dots, x_n : A_n \vdash M : A$.
-
-  1) The derivation ends in:
-
-\begin{prooftree}
-\AxiomC{$x_1 : A_1,\dots, x_n : A_n \vdash M : A$}
-\UnaryInfC{$x_1 : A_1,\dots, x_n : A_n \vdash {\bf pure \:} M : {\bf K}A$}
-\end{prooftree}
-
-  By assumption $M [x_1 := M_1,\dots,x_n := M_n] \in SC_A$, so ${\bf pure \:} M [x_1 := M_1,\dots,x_n := M_n] \in SC_{{\bf K}A}$.
-
-\vspace{\baselineskip}
-
-  2) The derivation ends in:
-\begin{prooftree}
-  \AxiomC{$x_1 : A_1,\dots, x_n : A_n \vdash \vec{M^{'}} : {\bf K}\vec{A}$}
-  \AxiomC{$\vec{x} : \vec{A} \vdash N : B$}
-  \BinaryInfC{$x_1 : A_1,\dots, x_n : A_n \vdash {\bf let \: pure \:} \vec{x} = \vec{M^{'}} {\: \bf in \:} N : {\bf K}B$}
-\end{prooftree}
-
-By IH forall $i \in \{ 1,\dots,\text{length}(\vec{M^{'}}) \}$, $M_i^{'} [x_1 := M_1,\dots,x_n := M_n] \in SC_{{\bf K}A_i}$.
-
-So ${\bf let \: pure \:} \vec{x} = \vec{M^{'}} [x_1 := M_1,\dots,x_n := M_n] {\: \bf in \:} N \in SC_{{\bf K} B}$,
-otherwise we can build infinite reduction path in $\vec{M^{'}} [x_1 := M_1,\dots,x_n := M_n]$.
-\end{proof}
-
-\begin{col}
-
-  All terms are strongly computable, therefore are strongly normalizing.
-\end{col}
-
-\end{proof}
-
-\begin{theorem}
-  $ $
-
-  $\twoheadrightarrow_{\beta}$ is confluent.
-\end{theorem}
-
-\begin{proof}
-
-  We modify and apply Barendregt's technique with term underlying. We will consider the grammar without constructors for pairs.
-
-\begin{defin} The set of underlined terms.
-
-  \begin{itemize}
-    \item $x \in \mathbb{V} \Rightarrow x \in \underline{\Lambda}$;
-    \item $M \in \underline{\Lambda} \Rightarrow (\lambda x. M) \in \underline{\Lambda}$;
-    \item $M, N \in \underline{\Lambda} \Rightarrow (M N) \in \underline{\Lambda}$;
-    \item $M \in \underline{\Lambda} \Rightarrow ({\bf pure \:} M) \in \underline{\Lambda}$;
-    \item $\vec{x} \in \mathbb{V}, \vec{M}, N \in \underline{\Lambda} \Rightarrow {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N \in \underline{\Lambda}$;
-    \item $M, N \in \underline{\Lambda} \Rightarrow (\lambda_i x. M) N \in \underline{\Lambda}$, for all $i \in \mathbb{N}$.
-  \end{itemize}
-\end{defin}
-
-\begin{defin} Subsitution for term with labelled lambda:
-
-  $((\lambda_i x. M) N) [y := Z] = (\lambda_i x. M [y := Z]) (N [y := Z])$
-\end{defin}
-
-\begin{defin} Index erasing
-
-  Let us define map $|.| : \underline{\Lambda} \to \Lambda$ as follows:
-
-  \begin{itemize}
-    \item $|x| = x$;
-    \item $|\lambda x. M| = \lambda x. |M|$;
-    \item $|M N| = |M| |N|$;
-    \item $|{\bf pure \:} M| = {\bf pure \:} |M|$;
-    \item $|{\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N| = {\bf let \: pure \:} \vec{x} = \vec{|M|} {\: \bf in \:} |N|$;
-    \item $|(\lambda_i x. M) N| = (\lambda x. |M|) |N|$
-  \end{itemize}
-\end{defin}
-
-\begin{defin} Reduction rules:
-
-\begin{itemize}
-  \item $(\lambda x. M) N \rightarrow_{\underline{\beta}} M [x := N]$;
-  \item $\begin{array}{llll}
-  & {\bf let \: pure \:} \vec{x}, y, \vec{z} = \vec{M}, {\bf let \: pure \: } \vec{w} = \vec{N} {\bf \: in \: } Q, \vec{P} {\: in \:} R \rightarrow_{\underline{\beta}} & \\
-  & {\bf let \: pure \:} \vec{x}, \vec{w}, \vec{z} = \vec{M}, \vec{N}, \vec{P} \: { \bf in } \: R [y := Q]
-  \end{array}$;
-  \item ${\bf let \: pure \:} \vec{x} = {\bf pure \:} \vec{M} {\: \bf in \:} N \rightarrow_{\underline{\beta}} {\bf pure \:} N [\vec{x} := \vec{M}]$;
-  \item ${\bf let \: pure \:} \underline{\quad} = \underline {\quad} {\: \bf in \:} M \rightarrow_{\underline{\beta}} {\bf pure \:} M$
-  \item $(\lambda x_i. M) N \rightarrow_{\underline{\beta}} M [x := N]$
-\end{itemize}
-\end{defin}
-
-$\twoheadrightarrow_{\underline{\beta}}$ is a reflexive-transitive closure of $\rightarrow_{\underline{\beta}}$.
-
-\begin{defin} Indexed redex erasing:
-
-  Let us define the next map $\phi : \underline{\Lambda} \to \Lambda$:
-
-  \begin{itemize}
-    \item $\phi(x) = x$;
-    \item $\phi(\lambda x. M) = \lambda x. \phi(M)$;
-    \item $\phi(M N) = \phi(M) \phi(N)$;
-    \item $\phi({\bf pure \:} M) = {\bf pure \:} \phi(M)$;
-    \item $\phi({\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N) = {\bf let \: pure \:} \vec{x} = \vec{\phi(M)} {\: \bf in \:} \phi(N)$;
-    \item $\phi((\lambda_i x. M) N) = \phi(M) [x := \phi(N)]$
-  \end{itemize}
-\end{defin}
-
-\begin{ex}
-  $ $
-
-  $\phi((\lambda_i x. {\bf let \: pure \:} y = x {\: \bf in \:} x) N) = {\bf let \: pure \:} y = \phi(N) {\: \bf in \:} \phi(N)$
-\end{ex}
-
-\begin{lemma}
-
-  $\forall \underline{M}, \underline{N} \in \underline{\Lambda} \: \forall M, N \in \Lambda, \text{if } |\underline{M}| = M, |\underline{N}| = N, \text{then}$
-  \begin{itemize}
-    \item If $M \twoheadrightarrow_{\beta} N$, then $\underline{M} \twoheadrightarrow_{\underline{\beta}} \underline{N}$
-    \item Vice versa
-  \end{itemize}
-\end{lemma}
-
-\begin{proof}
-
-  Induction on the generation $\rightarrow_{\beta}$ and $\rightarrow_{\underline{\beta}}$ correspondently. The general statement follows from transitivity
-  of multi-step reductions of both types.
-\end{proof}
-
-\begin{lemma}
-  $\phi(M [x := N]) = \phi(M) [x := \phi(N)]$.
-\end{lemma}
-
-\begin{proof}
-  We treat only cases with {\bf pure} and with {\bf let}. For the rest cases see \cite{Baren}.
-
-  1)
-  $ $
-
-  $\begin{array}{lll}
-  & \phi({\bf pure \:} (M [x := N])) = & \text{By the definition of $\phi$}\\
-  & {\bf pure \:} (\phi(M [x := N])) = & \text{Induction hypothesis} \\
-  & {\bf pure \:} (\phi(M) [x := \phi(N)]) = & \text{Substitution definition}\\
-  & ({\bf pure \:} \phi (M)) [x := \phi(N)]&
-  \end{array}$
-
-  2)
-  $ $
-
-  $\begin{array}{lll}
-  &\phi(({\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N) [y := P]) = & \text{Substitution definition}\\
-  &\phi({\bf let \: pure \:} \vec{x} = (\vec{M}[y := P]) {\: \bf in \:} N) = & \text{By the definition of $\phi$}\\
-  &{\bf let \: pure \:} \vec{x} = \phi(\vec{M}[y := P]) {\: \bf in \:} \phi(N) = & \text{Induction hypothesis} \\
-  &{\bf let \: pure \:} \vec{x} = (\phi(\vec{M}) [y := \phi(P)]) {\: \bf in \:} \phi(N) =& \text{Substitution definition} \\
-  &({\bf let \: pure \:} \vec{x} = \phi(\vec{M}) {\: \bf in \:} \phi(N))[y := \phi(P)]
-  \end{array}$
-
-\end{proof}
-
-\begin{lemma}
-  $ $
-
-  \begin{itemize}
-  \item If $M \twoheadrightarrow_{\underline{\beta}} N$, then $\phi(M) \twoheadrightarrow_{\beta} \phi(N)$
-  \item If $|M| = N$ and $\phi(M) = P$, then $N \twoheadrightarrow_{\beta} P$.
-  \end{itemize}
-\end{lemma}
-
-\begin{proof}
-  $ $
-
-  i) Induction on the generation of $\twoheadrightarrow_{\underline{\beta}}$ using previous lemma.
-
-  ii) Induction on the structure of $M$.
-\end{proof}
-
-\begin{lemma} Strip lemma.
-
-$ $
-
-If $M \rightarrow_{\beta} N$ and $M \twoheadrightarrow_{\beta} P$. Then there exists some term $Q$, such that
-$N \twoheadrightarrow_{\beta} Q$ and $P \twoheadrightarrow_{\beta} Q$.
-\end{lemma}
-
-\begin{proof}
-
-Proof is similar to \cite{Baren} \cite{Baren2}. We build the following diagram, which commutes by lemmas 8 and 10.
-
-\vspace{\baselineskip}
-
-\xymatrix{
-&&& M \ar[dd]_{\beta} \ar@{->>}[rrr] &&& P \ar@{-->>}[dd]^{\beta} \\
-&&&& N' \ar[ul]^{|.|} \ar[dl]_{\phi} \ar@{-->>}[rrr]^{\underline{\beta}} &&& P' \ar[ul]^{|.|} \ar[dl]_{\phi} \\
-&&& N \ar@{-->>}[rrr]_{\beta} &&& Q
-}
-\end{proof}
-
-\begin{col}
-$ $
-If $M \twoheadrightarrow_{\beta} N$ and $M \twoheadrightarrow_{\beta} P$. Then there exists some term $Q$, such that
-$N \twoheadrightarrow_{\beta} Q$ and $P \twoheadrightarrow_{\beta} Q$.
-\end{col}
-
-\begin{proof}
-
-  Unfold $M \twoheadrightarrow_{\beta} N$ as the sequence of one-step reductions and apply strip lemma on the every step.
-\end{proof}
-
-\end{proof}
-
-\begin{theorem}
-  $ $
-
-  Normal form in call-by-name $\lambda_{{\bf K}}$ has the subformula property: if $M$ is in normal formal, then its all subterms are in normal form too.
-
-\end{theorem}
-
-\begin{proof}
-  By induction on the structure of $M$.
-  Case with ${\bf let \: pure\:} \vec{x} = \vec{M} {\: \bf in \:} N$ was considered by Kakutani \cite{ModalK} \cite{ModalK1}.
-
-  If ${\bf pure \:} M$ is in normal form, so $M$ is in normal form and its subterms are in normal form too by hypothesis.
-
-  Thus if ${\bf pure \:} M$ is in normal form, then all its subterms are in normal form too.
-\end{proof}
-
-\section{Categorical semantics}
-
-\begin{defin} Monoidal category
-
-  Monoidal category is a category $\mathcal{C}$ with additional monoidal structure:
-  \begin{itemize}
-    \item A bifunctor $\otimes : \mathcal{C} \times \mathcal{C} \to C$ called tensor product;
-    \item Identity object $\mathds{1}$;
-    \item A natural isomorphism called associator: $\alpha_{A,B,C}: (A \otimes B) \otimes C \cong A \otimes (B \otimes C)$;
-    \item A natural isomorphism called left unitor: $L_A : \mathds{1} \otimes A \cong A$;
-    \item A natural isomorphism called right unitor $R_A : A \otimes \mathds{1} \cong A$;
-    \item A coherence condition called MacLane pentagon, i.e. the following diagram commutes:
+    \item Бифунктор $\otimes : \mathcal{C} \times \mathcal{C} \to C$, который мы будем называть тензором;
+    \item Единица $\mathds{1}$;
+    \item Изоморфизм, который мы будем называть ассоциатором: $\alpha_{A,B,C}: (A \otimes B) \otimes C \cong A \otimes (B \otimes C)$;
+    \item Изоморфизм  $L_A : \mathds{1} \otimes A \cong A$;
+    \item Изоморфизм $R_A : A \otimes \mathds{1} \cong A$;
+    \item Первое условие когерентности (пятиугольник Маклейна) (данная диаграмма коммутирует):
 
@@
     }
-    \item A coherence condition called triangle identity:
+    \item Второе условие когерентности (тождество треугольника):
 
@@
 
-    \item A monoidal category $\mathcal{C}$ is called symmetric if forall $A,B \in Ob_{\mathcal{C}}$, there is an isomorphism $\sigma_{A,B} : A \otimes B \cong B \otimes A$.
+    \item Моноидальная категория $\mathcal{C}$ называется симметрической, если для любых $A,B \in Ob_{\mathcal{C}}$, имеет место изоморфизм $\sigma_{A,B} : A \otimes B \cong B \otimes A$.
   \end{itemize}
@@
 
-\begin{defin} Cartesian closed category
+\begin{defin} Декартово замкнутная категория
 
-  Cartesian closed category is a category with a terminal object, finite products and exponentiation.
+  Декартово замкнутная категория -- это категория с терминальным объектом, конечными произведениями и экспоненцированием.
 \end{defin}
 
-Note that, any cartesian closed category is the special case of a (symmetric) monoidal category, where tensor is a product and identity object is a terminal object.
+Заметим, что декартово замкнутая категория -- это частный случай (симметрической) моноидальной категории, в котором тензор -- это произведения, а единица -- это терминальный объект.
 
-\begin{defin} Lax monoidal functor
+\begin{defin} Нестрогий моноидальный функтор
 
-  Let $\langle \mathcal{C}, \otimes_1, \mathds{1}_{\mathcal{C}} \rangle$ and $\langle \mathcal{D}, \otimes_2, \mathds{1}_{\mathcal{D}} \rangle$ are monoidal categories.
+  Пусть $\langle \mathcal{C}, \otimes_1, \mathds{1}_{\mathcal{C}} \rangle$ и $\langle \mathcal{D}, \otimes_2, \mathds{1}_{\mathcal{D}} \rangle$ моноидальные категории.
 
-  A lax monoidal functor $\mathcal{F} : \langle \mathcal{C}, \otimes_1, \mathds{1} \rangle \to \langle \mathcal{D}, \otimes_2, \mathds{1}' \rangle$ is a functor
-  $\mathcal{F} : \mathcal{C} \to \mathcal{D}$ with additional natural transformations:
+  Нестрогий моноидальный функтор $\mathcal{F} : \langle \mathcal{C}, \otimes_1, \mathds{1} \rangle \to \langle \mathcal{D}, \otimes_2, \mathds{1}' \rangle$ это функтор
+  $\mathcal{F} : \mathcal{C} \to \mathcal{D}$ с дополнительными естественными преобразованиями:
 
@@
 
-  and coherence maps:
+  и условиями когерентности:
 
   \begin{itemize}
-    \item Associativity:
+    \item Ассоциативность:
 
@@
 
-    \item Left unitality:
+    \item Свойство левой единицы:
 
@@
 
-    \item Right unitality:
+    \item Свойство правой единицы:
 
@@
 
-Strong endofunctor on a monoidal category is with a natural transformation so called tensorial strength
+Тензорно-сильный функтор -- это эндофунктор над моноидальной категорией с дополнительным естественным преобразованием и условиями когерентности для него (ниже соответствующие коммутирующие диаграмы):
 
-\begin{equation}
+\begin{center}
+$\begin{array}{lll}
   \tau_{A, B} : A \otimes \mathcal{K}B \rightarrow \mathcal{K}(A \otimes B)
-\end{equation}
-
-making two diagrams commute:
+\end{array}$
+\end{center}
 
@@
 
-\begin{defin} Applicative functor
+\begin{defin} Аппликативный функтор
 
-  An applicative functor is a triple $\langle \mathcal{C}, \mathcal{K}, \eta \rangle$,
-where $\mathcal{C}$ is a monoidal category, $\mathcal{K}$ is a strong lax monoidal endofunctor and $\eta : Id_{\mathcal{C}} \Rightarrow \mathcal{K}$ is a natural transformation (similar to unit in monad), such that:
+  Аппликативный функтор -- это тройка $\langle \mathcal{C}, \mathcal{K}, \eta \rangle$,
+где $\mathcal{C}$ -- это моноидальная категория, $\mathcal{K}$ - это тензорно-сильный нестрогий моноидальный эндофунктор и
+$\eta : Id_{\mathcal{C}} \Rightarrow \mathcal{K}$ -- это естественное преобразование, такое, что:
 
@@
 \item $u = \eta_{\mathds{1}}$;
-\item $\ast_{A,B} \circ (\eta_A \otimes \eta_B) = \eta_{A \otimes B}$, i.e. the following diagram commutes:
+\item $\ast_{A,B} \circ (\eta_A \otimes \eta_B) = \eta_{A \otimes B}$, то есть диаграмма коммутирует:
 
@@
 
-By default we will consider some strong lax monoidal functor on cartesian closed category.
-
-Note that lax monoidal functor on a cartesian closed category is a special case of an applicative functor \cite{Cons}, but
-we will define $\eta$ explicitly for simplicity.
-
-\subsection{Soundness and completeness}
-
-\begin{theorem} Soundness
-
-  Let $\Gamma \vdash M : A$ and $M =_{\beta\eta} N$, then $[\![\Gamma \vdash M : A]\!] = [\![\Gamma \vdash N : A]\!]$
-\end{theorem}
-
-\begin{proof}
-
-\begin{defin} Semantical translation from $\lambda_{{\bf K}}$ to some cartesian closed category $\mathcal{C}$ with an applicative functor $\mathcal{K}$:
-
-\begin{itemize}
-\item Interpretation for types:
-  \begin{itemize}
-    \item $[\![A]\!] := \hat{A}, A \in \mathbb{T}$, where $\hat{A}$ is an object of $\mathcal{C}$ obtained by some given assignment;
-    \item $[\![A \to B]\!] := [\![B]\!]^{[\![A]\!]}$;
-    \item $[\![A \times B]\!] := [\![A]\!] \times [\![B]\!]$.
-  \end{itemize}
-\item Interpretation for modal types:
-  \begin{itemize}
-    \item $[\![{\bf K}A]\!] = \mathcal{K}[\![A]\!]$;
-  \end{itemize}
-\item Interpretaion for contexts:
-  \begin{itemize}
-    \item $[\![ \quad ]\!] = \mathds{1}$, where $\mathds{1}$ is a terminal object of a given ccc;
-    \item $[\![\Gamma, x : A]\!] = [\![\Gamma]\!] \times [\![A]\!]$
-  \end{itemize}
-\item Interpretation for typing assignment:
-  \begin{itemize}
-    \item $[\![\Gamma \vdash M : A]\!] := [\![M]\!] : [\![\Gamma]\!] \to [\![A]\!]$.
-  \end{itemize}
-\item Interpretation for typing rules:
-
-\begin{prooftree}
-\AxiomC{$ $}
-\UnaryInfC{$[\![\Gamma, x : A \vdash x : A]\!] = \pi_2 : [\![\Gamma]\!] \times [\![A]\!] \rightarrow
-[\![A]\!]$}
-\end{prooftree}
-
-\begin{prooftree}
-\AxiomC{$[\![\Gamma, x : A \vdash M : B]\!] = f : [\![\Gamma]\!] \times [\![A]\!] \rightarrow [\![B]\!]$}
-\UnaryInfC{$[\![\Gamma \vdash (\lambda x. M) : A \to B]\!] = \Lambda(f) : [\![\Gamma]\!]
-\rightarrow[\![B]\!]^{[\![A]\!]}$}
-\end{prooftree}
-
-\begin{prooftree}
-\AxiomC{$[\![\Gamma \vdash M : A \to B]\!] = [\![M]\!] : [\![\Gamma]\!] \rightarrow [\![B]\!]^{[\![A]\!]}$}
-\AxiomC{$[\![\Gamma \vdash N : A]\!] = [\![N]\!] : [\![\Gamma]\!] \rightarrow [\![A]\!]$}
-\BinaryInfC{$[\![\Gamma \vdash (M N) : B]\!] = [\![\Gamma]\!] \xrightarrow{\langle [\![M]\!], [\![N]\!]
-\rangle} [\![B]\!]^{[\![A]\!]} \times [\![A]\!] \xrightarrow{\epsilon} [\![B]\!] $}
-\end{prooftree}
-
-\begin{prooftree}
-\AxiomC{$[\![\Gamma \vdash M : A ]\!] = f : [\![\Gamma]\!] \rightarrow [\![A]\!]$}
-\AxiomC{$[\![\Gamma \vdash N : B ]\!] = g : [\![\Gamma]\!] \rightarrow [\![B]\!]$}
-\BinaryInfC{$[\![\Gamma \vdash (M, N) : A \times B]\!] = \langle f, g \rangle : [\![\Gamma]\!] \rightarrow
-[\![A]\!] \times [\![B]\!]$}
-\end{prooftree}
-
-\begin{prooftree}
-\AxiomC{$[\![\Gamma \vdash p : A_1 \times A_2]\!] = f : [\![\Gamma]\!] \rightarrow [\![A_1]\!] \times
-[\![A_2]\!]$}
-\RightLabel{$i \in \{1,2\}$}
-\UnaryInfC{$[\![\Gamma \vdash \pi_i p : A_i]\!] = [\![\Gamma]\!] \xrightarrow{f} [\![A_1]\!] \times
-[\![A_2]\!] \xrightarrow{\pi_i} [\![A_i]\!]$}
-\end{prooftree}
-
-\begin{prooftree}
-\AxiomC{$[\![\Gamma \vdash M : A]\!] = [\![M]\!] : [\![\Gamma]\!] \rightarrow [\![A]\!]$}
-\UnaryInfC{$[\![\Gamma \vdash {\bf pure \:} M : {\bf K}A]\!] := [\![\Gamma]\!] \xrightarrow{[\![M]\!]}
-[\![A]\!] \xrightarrow{\eta_{[\![A]\!]}} \mathcal{K}[\![A]\!]$}
-\end{prooftree}
-
-\begin{small}
-  \begin{prooftree}
-    \AxiomC{$[\![\Gamma \vdash \vec{M} : {\bf K} \vec{A}]\!] = \langle [\![M_1]\!],\dots, [\![M_n]\!] \rangle : [\![\Gamma]\!] \rightarrow \prod \limits_{i=1}^n \mathcal{K}[\![A_i]\!]$}
-    \AxiomC{$[\![\vec{x} : \vec{A} \vdash N : B]\!] = [\![N]\!] : \prod \limits_{i=1}^n [\![A_i]\!] \rightarrow [\![B]\!]$}
-    \BinaryInfC{$[\![\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } M : {\bf K} B]\!] = \mathcal{K}([\![N]\!]) \circ \ast_{[\![A_1]\!],\dots,[\![A_n]\!]} \circ \langle [\![M_1]\!],\dots, [\![M_n]\!] \rangle : [\![\Gamma]\!] \rightarrow \mathcal{K}[\![B]\!]$}
-  \end{prooftree}
-\end{small}
-\end{itemize}
-\end{defin}
-
-\begin{defin} Simultaneous substitution
-
-Let $\Gamma = \{ x_1 : A_1, ..., x_n : A_n \}$, $\Gamma \vdash M : A$ and for all $i \in \{ 1,..., n \}$,
-$\Gamma \vdash M_i : A_i$.
-
-We define simultaneous substitution $M [ \vec{x} := \vec{M}]$ recursively by:
-
-\begin{itemize}
-\item $x_i [ \vec{x} := \vec{M}] = M_i $;
-\item $(\lambda x. M) [ \vec{x} := \vec{M}] = \lambda x. (M [ \vec{x} := \vec{M}])$;
-\item $(M N) [ \vec{x} := \vec{M}] = (M [ \vec{x} = \vec{M}]) (N [ \vec{x} := \vec{M}])$;
-\item $\langle M, N \rangle = \langle (M [ \vec{x} = \vec{M}]), (N [ \vec{x} := \vec{M}])\rangle$;
-\item $(\pi_i P) [ \vec{x} := \vec{M}] = \pi_i (P [ \vec{x} = \vec{M}])$;
-\item $({\bf pure \:} M) [ \vec{x} := \vec{M}] = {\bf pure \:} (M [ \vec{x} = \vec{M}])$;
-\item $({\bf let \: pure} \: \vec{x} = \vec{M} {\: \bf in \:} N) [\vec{y} := \vec{P}] =
-{\bf let \: pure} \: \vec{x} = (\vec{M} [\vec{y} := \vec{P}]) {\: \bf in \:} N$
-\end{itemize}
-\end{defin}
-
-\begin{lemma}
-$ $
-
-$[\![M [x_1 := M_1,\dots, x_n := M_n]]\!] = [\![M]\!] \circ \langle [\![M_1]\!], \dots, [\![M_n]\!] \rangle$.
-
-\end{lemma}
-
-\begin{proof}
-
-$ $
-
-1)
-
-$\begin{array}{lll}
-& [\![\Gamma \vdash ({\bf pure \:} M) [ \vec{x} := \vec{M}] : {\bf K}A]\!] = [\![ \Gamma \vdash {\bf pure \:} (M [ \vec{x} := \vec{M}]) : {\bf K}A]\!] &
-\text{Substitution definition} \\
-& \quad\quad\quad\quad\quad\quad\quad = \eta_{[\![A]\!]} \circ [\![(M [ \vec{x} := \vec{M}])]\!]&
-\text{Translation for pure}\\
-& \quad\quad\quad\quad\quad\quad\quad = \eta_{[\![A]\!]} \circ ([\![M]\!] \circ \langle [\![M_1]\!], \dots,
-[\![M_n]\!] \rangle) & \text{Induction hypothesis} \\
-& \quad\quad\quad\quad\quad\quad\quad = (\eta_{[\![A]\!]} \circ [\![M]\!]) \circ \langle [\![M_1]\!], \dots,
-[\![M_n]\!] \rangle & \text{Associativity of composition}\\
-& \quad\quad\quad\quad\quad\quad\quad = [\![ \Gamma \vdash {\bf pure \:} M : {\bf K}A]\!] \circ \langle [\![M_1]\!], \dots,
-[\![M_n]\!] \rangle& \text{Translation for pure}\\
-\end{array}$
-
-\vspace{\baselineskip}
-
-2)
-
-\vspace{\baselineskip}
-
-$\begin{array}{lll}
-&[\![\Gamma \vdash ({\bf let \: pure} \: \vec{x} = \vec{M} {\: \bf in \:} N) [\vec{y} := \vec{P}] : {\bf K} B]\!] =& \text{Substitution definition}\\
-&[\![\Gamma \vdash {\bf let \: pure} \: \vec{x} = (\vec{M} [\vec{y} := \vec{P}]) {\: \bf in \:} N : {\bf K}B]\!] =& \text{Interpretaion for $let_{\bf K}$} \\
-&\mathcal{K}([\![N]\!]) \circ \ast_{[\![A_1]\!],\dots,[\![A_n]\!]} \circ [\![\Gamma \vdash (\vec{M} [\vec{y} := \vec{P}]) : {\bf K}\vec{A}]\!] =& \text{Induction hypothesis} \\
-&\mathcal{K}([\![N]\!]) \circ \ast_{[\![A_1]\!],\dots,[\![A_n]\!]} \circ ([\![\vec{M}]\!] \circ \langle [\![P_1]\!],\dots,[\![P_n]\!]\rangle) = & \text{Associativity of composition} \\
-&(\mathcal{K}([\![N]\!]) \circ \ast_{[\![A_1]\!],\dots,[\![A_n]\!]} \circ [\![\vec{M}]\!]) \circ \langle [\![P_1]\!],\dots,[\![P_n]\!]\rangle = & \text{By interpretation} \\
-&[\![\Gamma \vdash {\bf let \: pure} \: \vec{x} = \vec{M} {\: \bf in \:} N : {\bf K}B]\!] \circ \langle [\![P_1]\!],\dots,[\![P_n]\!]\rangle&
-\end{array}$
-
-\end{proof}
-
-\begin{lemma}
-  $ $
-
-  Let $\Gamma \vdash M : A$ and $M \twoheadrightarrow_{\beta \eta} N$, then $[\![\Gamma \vdash M : A]\!] = [\![\Gamma \vdash N : A]\!]$;
-\end{lemma}
-
-\begin{proof}
-  $ $
-
-Cases with $\beta$-reductions for $let_{\bf K}$ are shown in \cite{ModalK1}. Let us consider cases with ${\bf pure}$.
-
-\vspace{\baselineskip}
-
-1) $[\![\Gamma \vdash {\bf let \: pure \:} \vec{x} = {\bf pure \:} \vec{M} {\: \bf in \:} N : {\bf K}B]\!] = [\![\Gamma \vdash {\bf pure \:} N [\vec{x} := \vec{M}] : {\bf K}B]\!]$
-
-\vspace{\baselineskip}
-
-$\begin{array}{lll}
-&[\![\Gamma \vdash {\bf let \: pure \:} \vec{x} = {\bf pure \:} \vec{M} {\: \bf in \:} N : {\bf K}B]\!] = & \\
-&\text{\quad\quad\quad\quad\quad\quad\quad\quad\quad By interpretation}& \\
-&\mathcal{K}([\![N]\!]) \circ \ast_{[\![A_1]\!],\dots,[\![A_n]\!]} \circ \langle \eta_{[\![A_1]\!]} \circ [\![M_1]\!],\dots,\eta_{[\![A_n]\!]} \circ [\![M_n]\!] \rangle = \\
-&\text{\quad\quad\quad\quad\quad\quad\quad\quad\quad By the property of a pair of morphisms}& \\
-&\mathcal{K}([\![N]\!]) \circ \ast_{[\![A_1]\!],\dots,[\![A_n]\!]} \circ (\eta_{[\![A_1]\!]} \times \dots \times \eta_{[\![A_n]\!]}) \circ \langle [\![M_1]\!], \dots, [\![M_n]\!]\rangle =& \\
-&\text{\quad\quad\quad\quad\quad\quad\quad\quad\quad Associativity of composition}& \\
-&\mathcal{K}([\![N]\!]) \circ (\ast_{[\![A_1]\!],\dots,[\![A_n]\!]} \circ (\eta_{[\![A_1]\!]} \times \dots \eta_{[\![A_n]\!]})) \circ \langle [\![M_1]\!], \dots, [\![M_n]\!] \rangle =& \\
-&\text{\quad\quad\quad\quad\quad\quad\quad\quad\quad By the definition of an applicative functor}& \\
-&\mathcal{K}([\![N]\!]) \circ \eta_{[\![A_1]\!] \times \dots \times [\![A_n]\!]} \circ \langle [\![M_1]\!], \dots, [\![M_n]\!] \rangle =& \\
-&\text{\quad\quad\quad\quad\quad\quad\quad\quad\quad Naturality of $\eta$}& \\
-&\eta_{[\![B]\!]} \circ [\![N]\!] \circ \langle [\![M_1]\!], \dots, [\![M_n]\!] \rangle =& \\
-&\text{\quad\quad\quad\quad\quad\quad\quad\quad\quad Associativity of composition}& \\
-&\eta_{[\![B]\!]} \circ ([\![N]\!] \circ \langle [\![M_1]\!], \dots, [\![M_n]\!]) \rangle =& \\
-&\text{\quad\quad\quad\quad\quad\quad\quad\quad\quad Simultaneous substitution lemma}& \\
-&\eta_{[\![B]\!]} \circ [\![N [\vec{x} := \vec{M}]]\!]& \\
-&\text{\quad\quad\quad\quad\quad\quad\quad\quad\quad By interpetation}& \\
-&[\![\Gamma \vdash {\bf pure \:} (N [\vec{x} := \vec{M}]) : {\bf K}B]\!]&
-\end{array}$
-
-\vspace{\baselineskip}
-
-2) $[\![\vdash {\bf let \: pure \:} \underline{\quad} = \underline {\quad} {\: \bf in \:} M : {\bf K} A]\!] = [\![\vdash {\bf pure \:} M : {\bf K} A]\!]$
-
-$\begin{array}{lll}
-&[\![\vdash {\bf let \: pure \:} \underline{\quad} = \underline {\quad} {\: \bf in \:} M : {\bf K} A]\!] = & \text{By interpretation} \\
-&\mathcal{K}([\![M]\!]) \circ u_{\mathds{1}} = & \text{By the definition of an applicative functor} \\
-&\mathcal{K}([\![M]\!]) \circ \eta_{\mathds{1}} = & \text{By naturality of $\eta$} \\
-&\eta_{[\![A]\!]} \circ [\![M]\!] = & \text{By interpretation} \\
-&[\![\vdash {\bf pure \:} M : {\bf K} A]\!]&
-\end{array}$
-\end{proof}
-
-\end{proof}
-
-\begin{theorem} Completeness
-
-Let $[\![\Gamma \vdash M : A]\!] = [\![\Gamma \vdash N : A]\!]$, then $M =_{\beta \eta} N$.
-\end{theorem}
-
-\begin{proof}
-
-$ $
-
-We will consider term model for simply typed lambda calculus $\times$ and $\to$ standardly described in \cite{LambekScott}:
-
-\begin{defin} Equivalence on term pairs:
-  $ $
-
-  Let us define relation $\sim_{A, B} \subseteq \mathbb{V} \times \Lambda_{{\bf K}}$, such that:
-
-  $(x, M) \sim_{A, B} (y, N) \Leftrightarrow x : A \vdash M : B \:\: \& \:\: y : A \vdash N : A \:\: \& \:\: M =_{\beta \eta} N [y := x]$;
-\end{defin}
-
-We will denote equivalence class as $[x, M]_{A, B} = \{ (y, N) \: | \: (x, M) \sim_{A, B} (y, N) \}$ (we will drop indices below).
-
-\begin{defin} Category $\mathcal{C}(\lambda)$:
-\begin{itemize}
-  \item $Ob_{\mathcal{C}} = \{ \hat{A} \: | \: A \in \mathbb{T} \} \cup \{ \mathds{1} \}$;
-  \item $Hom_{\mathcal{C}(\lambda)}(\hat{A},\hat{B}) = (\mathbb{V} \times \Lambda_{{\bf K}})/_{\sim_{A, B}}$;
-  \item Let $[x, M] \in Hom_{\mathcal{C}(\lambda)}(\hat{A},\hat{B})$ and $[y,N] \in Hom_{\mathcal{C}(\lambda)}(\hat{B},\hat{C})$, then $[y,M] \circ [x, M] = [x, N [y := M]]$;
-  \item Identity morphism $id_{\hat{A}} = [x,x] \in Hom_{\mathcal{C}(\lambda)(\hat{A})}$;
-  \item $\mathds{1}$ is a terminal object;
-  \item $\widehat{A \times B} = \hat{A} \times \hat{B}$;
-  \item Canonical projection is defined as $[x, \pi_i x] \in Hom_{\mathcal{C}(\lambda)}(\hat{A_1} \times \hat{A_2},\hat{A_i})$ for $i \in \{ 1, 2 \}$;
-  \item $\widehat{A \to B} = \hat{B}^{\hat{A}}$;
-  \item Evaluation arrow $\epsilon = [x, (\pi_2 x) (\pi_1 x)] \in Hom_{\mathcal{C}(\lambda)(\hat{B}^{\hat{A}} \times \hat{A}, \hat{B})}$.
-\end{itemize}
-\end{defin}
-
-It is sufficient to show ${\bf K}$ is an applicative functor on $\mathcal{C}(\lambda)$.
-
-\begin{defin}
-  Let us define an endofunctor $\mathcal{K} : \mathcal{C}(\lambda) \to \mathcal{C}(\lambda)$, such that
-forall $[x,M] \in Hom_{\mathcal{C}(\lambda)}(\hat{A},\hat{B}), {\bf K}([x,M]) = [y, {\bf let \: pure \:} x = y {\: \bf in \:} M] \in Hom_{\mathcal{C}(\lambda)}({\bf K} \hat{A}, {\bf K} \hat{B})$ (denotation: $\text{fmap } f$ for an arbitrary arrow $f$).
-
-\end{defin}
-
-\begin{lemma} Functoriality
-
-\begin{itemize}
-  \item $\text{fmap }(g \circ f) = \text{fmap }(g) \circ \text{fmap }(f)$;
-  \item $\text{fmap }(id_{\hat{A}}) = id_{{\bf K}\hat{A}}$.
-\end{itemize}
-\end{lemma}
-
-\begin{proof}
-
-Easy checking using reduction rules.
-
-\end{proof}
-
-\begin{defin}
-
-  Let us define natural transformations:
-
-\begin{itemize}
-  \item $\eta:Id \Rightarrow \mathcal{K}$, s. t. $\forall \hat{A} \in Ob_{\mathcal{C}(\lambda)}$, $\eta_{\hat{A}} = [x, {\bf pure \:} x] \in Hom_{\mathcal{C}(\lambda)}(\hat{A}, {\bf K}\hat{A})$;
-  \item $\ast_{A,B}:{\bf K}\hat{A} \times {\bf K}\hat{B} \to {\bf K}(\hat{A} \times \hat{B})$, s.
-t.
-$\forall \hat{A}, \hat{B} \in Ob_{\mathcal{C}(\lambda)}, \ast_{\hat{A},\hat{B}} = [p, {\bf let \: pure \:} x,y = \pi_1 p, \pi_2 p {\: \bf in \:} \langle x, y \rangle] \in Hom_{\mathcal{C}(\lambda)}({\bf K}A \times {\bf K}B, {\bf K}(A \times B))$.
-\end{itemize}
-\end{defin}
-
-Implementation for $\ast$ in our term model is a modification of ${\text{let}_{\bf K}}$-rule:
-
-\begin{prooftree}
-\AxiomC{$p : {\bf K}A \times {\bf K}B \vdash p : {\bf K}A \times {\bf K}B$}
-\UnaryInfC{$p : {\bf K}A \times {\bf K}B \vdash \pi_1 p : {\bf K} A$}
-\AxiomC{$p : {\bf K}A \times {\bf K}B \vdash p : {\bf K}A \times {\bf K}B$}
-\UnaryInfC{$p : {\bf K}A \times {\bf K}B \vdash \pi_2 p : {\bf K} B$}
-\AxiomC{$x : A \vdash x : A$}
-\AxiomC{$y : B \vdash y : B$}
-\BinaryInfC{$x : A, y : B \vdash \langle x, y \rangle : A \times B$}
-\TrinaryInfC{$p : {\bf K}A \times {\bf K}B \vdash {\bf let \: pure \:} x,y = \pi_1 p, \pi_2 p {\: \bf in \:} \langle x, y \rangle : {\bf K}(A \times B)$}
-\end{prooftree}
-
-\begin{lemma}
-  $ $
-
-  {\bf K} is a lax monoidal endofunctor
-\end{lemma}
-
-\begin{proof}
-$ $
-
-See \cite{ModalK}
-\end{proof}
-
-\begin{lemma} Naturality and coherence for $\eta$:
-
-\begin{itemize}
-  \item $\text{fmap } f \circ \eta_A = \eta_B \circ f$;
-  \item $\ast_{\hat{A},\hat{B}} \circ (\eta_{A} \times \eta_{B}) = \eta_{\hat{A} \times \hat{B}}$;
-\end{itemize}
-\end{lemma}
-
-\begin{proof}
-  $ $
-
-  i) $\text{fmap } f \circ \eta_{\hat{A}} = \eta_{\hat{B}} \circ f$
-
-\vspace{\baselineskip}
-
-$\begin{array}{lll}
-&\eta_{\hat{B}} \circ f = & \text{By the definition} \\
-&[y, {\bf pure \:} y] \circ [x, M] = & \text{By the definition of composition}\\
-&[x, {\bf pure \:} y [y := M]] = & \text{By substitution}\\
-&[x, {\bf pure \:} M]& \\
-& &\\
-&\text{On the other hand:}& \\
-&\text{fmap } f \circ \eta_{\hat{A}} = & \text{By the definiton} \\
-&[z, {\bf let \: pure \:} x = z {\: \bf in \:} M] \circ [x, {\bf pure \: x}] = & \text{By the definition of composition}\\
-&[x, {\bf let \: pure \:} x = z {\: \bf in \:} M [z := {\bf pure \:} x]] = & \text{By substitution} \\
-&[x, {\bf let \: pure \:} x = {\bf pure \:} x {\: \bf in \:} M] = & \text{$\beta$-reduction rule}\\
-&[x, {\bf pure \:} M [x := x]] = & \text{By substitution}\\
-&[x, {\bf pure \:} M]&
-\end{array}$
-
-\vspace{\baselineskip}
-
-ii) $\ast_{\hat{A},\hat{B}} \circ (\eta_{\hat{A}} \times \eta_{\hat{B}}) = \eta_{\hat{A} \times \hat{B}}$
-
-\vspace{\baselineskip}
-
-$\begin{array}{lll}
-& \ast_{\hat{A},\hat{B}} \circ (\eta_{\hat{A}} \times \eta_{\hat{B}}) = & \\
-& \quad\quad\quad\quad\quad\quad \text{By unfolding} & \\
-& [q, {\bf let \: pure \:} x, y = \pi_1 q, \pi_2 q {\: \bf in \:} \langle x, y \rangle] \circ [p, \langle {\bf pure \:} (\pi_1 p), {\bf pure \:} (\pi_2 p) \rangle] = & \\
-& \quad\quad\quad\quad\quad\quad \text{Composition}& \\
-& [p, {\bf let \: pure \:} x, y = \pi_1 q, \pi_2 q {\: \bf in \:} \langle x, y \rangle [q := \langle {\bf pure \:} (\pi_1 p), {\bf pure \:} (\pi_2 p) \rangle]] = & \\
-& \quad\quad\quad\quad\quad\quad \text{By substitution}& \\
-& [p, {\bf let \: pure \:} x, y = \pi_1 (\langle {\bf pure \:} (\pi_1 p), {\bf pure \:} (\pi_2 p) \rangle), \pi_2 (\langle {\bf pure \:} (\pi_1 p), {\bf pure \:} (\pi_2 p) \rangle) {\: \bf in \:} \langle x, y \rangle] = & \\
-& \quad\quad\quad\quad\quad\quad \text{Reduction rules}& \\
-& [p, {\bf let \: pure \:} x, y = {\bf pure \:} (\pi_1 p), {\bf pure \:} (\pi_2 p) {\: \bf in \:} \langle x, y \rangle] = & \\
-& \quad\quad\quad\quad\quad\quad \text{Reduction rule}& \\
-& [p, {\bf pure \:} (\langle x,y \rangle [x := \pi_1 p, y := \pi_2 p])] = & \\
-& \quad\quad\quad\quad\quad\quad \text{Substitution}& \\
-& [p, {\bf pure \:} \langle \pi_1 p, \pi_2 p \rangle] = & \\
-& \quad\quad\quad\quad\quad\quad \text{$\eta$-reduction}& \\
-& [p, {\bf pure \:} p] =& \\
-& \quad\quad\quad\quad\quad\quad \text{By definition}& \\
-& \eta_{\hat{A} \times \hat{B}}&
-\end{array}$
-
-\end{proof}
-
-\begin{defin}
-  $ $
-
-  $u_{\mathds{1}} = [\sqbullet, {\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} \sqbullet] \in Hom_{\mathcal{C}(\lambda)}(\mathds{1}, {\bf K}\mathds{1})$.
-\end{defin}
-
-\begin{lemma}
-  $ $
-
-  $u_{\mathds{1}} = \eta_{\mathds{1}}$
-
-\end{lemma}
-
-\begin{proof}
-
-  Immediately.
-\end{proof}
-
-  Tensorial strength is defined as follows:
-
-\begin{defin} Tensorial strength
-  $ $
-
-  Let $[p, \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle] \in Hom_{\mathcal{C}(\lambda)}(\hat{A} \times {\bf K}\hat{B}, {\bf K}\hat{A} \times {\bf K}\hat{B})$.
-
-  So tensorial strength is defined as $\tau_{\hat{A}, \hat{B}} = \ast_{\hat{A}, \hat{B}} \circ [p, \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle]$.
-\end{defin}
-
-It is clearly that tensorial strength defined above can be simplified as follows:
-
-$\begin{array}{lll}
-&\ast_{\hat{A}, \hat{B}} \circ [p, \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle] = & \text{By definition} \\
-&[p^{'}, {\bf let \: pure \:} x, y = \pi_1 p^{'}, \pi_2 p^{'} {\: \bf in \:} \langle x, y \rangle] \circ [p, \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle] = & \text{By composition} \\
-&[p, {\bf let \: pure \:} x, y = \pi_1 p^{'}, \pi_2 p^{'} {\: \bf in \:} \langle x, y \rangle [p^{'} := \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle]] = & \text{By substitution} \\
-&[p, {\bf let \: pure \:} x, y = \pi_1 (\langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle), \pi_2 (\langle \pi_1 p, {\bf pure \:} (\pi_2 p) \rangle) {\: \bf in \:} \langle x, y \rangle] = & \text{By $\beta$-reduction rules} \\
-&[p, {\bf let \: pure \:} x, y = {\bf pure \:} (\pi_1 p), \pi_2 p {\: \bf in \:} \langle x, y \rangle]&
-\end{array}$
-
-\begin{lemma} Coherence for tensorial strength
-
-  \begin{itemize}
-    \item $\text{fmap } \alpha_{\hat{A},\hat{B},\hat{C}} \circ \tau_{\hat{A} \times \hat{B}, \hat{C}} = \tau_{\hat{A}, \hat{B} \times \hat{C}} \circ (id_{\hat{A}} \times \tau_{\hat{B}, \hat{C}}) \circ \alpha_{\hat{A}, \hat{B}, {\bf K}\hat{C}}$;
-    \item ${\bf K}(\rho_{\hat{A}}) \circ \tau_{\mathds{1}, \hat{A}} = \rho_{{\bf K}\hat{A}}$.
-  \end{itemize}
-  where $\alpha_{\hat{A},\hat{B},\hat{C}} = [p, \langle \pi_1 (\pi_1 p), \langle \pi_2 (\pi_1 p), \pi_2 p \rangle \rangle]$ and
-  $\rho = \pi_2$.
-\end{lemma}
-
-\begin{proof}
-  $ $
-
-1) Let us define $\tau_{\hat{A} \times \hat{B}, \hat{C}}$ as follows:
-
-$\tau_{\hat{A} \times \hat{B}, \hat{C}} = [p, {\bf \: let \: pure \:} x,y,z = {\bf pure\:} (\pi_1 (\pi_1 p)), {\bf pure \:} (\pi_2 (\pi_1 p)), \pi_2 p {\: \bf in \:} \langle \langle x,y \rangle, z \rangle]$
-
-\vspace{\baselineskip}
-
-Then:
-
-\vspace{\baselineskip}
-
-$\begin{array}{lll}
-&\text{fmap } \alpha_{\hat{A},\hat{B},\hat{C}} \circ \tau_{\hat{A} \times \hat{B}, \hat{C}} = & \\
-&[q, {\bf let \: pure \:} r = q {\: \bf in \:}  \langle \pi_1 (\pi_1 r), \langle \pi_2 (\pi_1 r), \pi_2 r \rangle \rangle] \: \circ & \\
-&\quad\quad\quad\quad \circ \: [p, {\bf \: let \: pure \:} x,y,z = {\bf pure\:} (\pi_1 (\pi_1 p)), {\bf pure \:} (\pi_2 (\pi_1 p)), \pi_2 p {\: \bf in \:} \langle \langle x,y \rangle, z \rangle] = & \\
-&\quad\quad\quad\quad\quad\quad \text{Composition}& \\
-&[p, {\bf let \: pure \:} r = q {\: \bf in \:}  \langle \pi_1 (\pi_1 r), \langle \pi_2 (\pi_1 r), \pi_2 r \rangle \rangle & \\
-&\quad\quad\quad\quad [q := {\bf \: let \: pure \:} x,y,z = {\bf pure\:} (\pi_1 (\pi_1 p)), {\bf pure \:} (\pi_2 (\pi_1 p)), \pi_2 p {\: \bf in \:} \langle \langle x,y \rangle, z \rangle]] = & \\
-&\quad\quad\quad\quad\quad\quad \text{Substitution and reduction}& \\
-&[p, {\bf let \: pure \:} r = ({\bf let \: pure \:} x,y,z = {\bf pure\:} (\pi_1 (\pi_1 p)), {\bf pure \:} (\pi_2 (\pi_1 p)), \pi_2 p {\: \bf in \:} \langle \langle x,y \rangle, z \rangle) & \\
-&\quad\quad\quad\quad {\: \bf in \:} \langle \pi_1 (\pi_1 r), \langle \pi_2 (\pi_1 r), \pi_2 r \rangle \rangle] = & \\
-&\quad\quad\quad\quad\quad\quad \text{$\beta$-reduction rule}& \\
-&[p, {\bf let \: pure \:} x,y,z = {\bf pure\:} (\pi_1 (\pi_1 p)), {\bf pure \:} (\pi_2 (\pi_1 p)), \pi_2 p {\: \bf in \:}  \langle \pi_1 (\pi_1 r), \langle \pi_2 (\pi_1 r), \pi_2 r \rangle \rangle & \\
-&\quad\quad\quad\quad [r := \langle \langle x,y \rangle, z \rangle]] = & \\
-&\quad\quad\quad\quad\quad\quad \text{$\beta$-reduction rule}& \\
-&[p, {\bf let \: pure \:} x,y,z = {\bf pure\:} (\pi_1 (\pi_1 p)), {\bf pure \:} (\pi_2 (\pi_1 p)), \pi_2 p {\: \bf in \:} \langle x, \langle y, z \rangle \rangle]&
-\end{array}$
-
-\vspace{\baselineskip}
-
-On the other hand:
-
-\vspace{\baselineskip}
-
-Let us define as follows $\tau_{\hat{A}, \hat{B} \times \hat{C}}$:
-
-$\begin{array}{lll}
-&\tau_{\hat{A}, \hat{B} \times \hat{C}} = [r, {\bf let \: pure \:} x, y, z = ({\bf pure \:} \pi_1 r, {\bf let \: pure \:} q^{'} = \pi_2 r {\: \bf in \:} \pi_1 q^{'}, {\bf let \: pure \:} q^{'} = \pi_2 r {\: \bf in \:} \pi_2 q^{'}) & \\
-&\quad\quad\quad\quad {\: \bf in \:} \langle x, \langle y, z \rangle\rangle]& \\
-\end{array}$
-
-\vspace{\baselineskip}
-
-We may simplify the current instance of tensorial strength:
-
-$\begin{array}{lll}
-&[p, {\bf let \: pure \:} x, y, z = ({\bf pure \:} \pi_1 p, {\bf let \: pure \:} p^{'} = \pi_2 p {\: \bf in \:} \pi_1 p^{'}, {\bf let \: pure \:} p^{'} = \pi_2 p {\: \bf in \:} \pi_2 p^{'}) & \\
-&\quad\quad\quad\quad {\: \bf in \:} \langle x, \langle y, z \rangle\rangle] = & \\
-&\quad\quad\quad\quad\quad\quad \text{Substitution and reduction}& \\
-&[p, {\bf let \: pure \:} x, p^{'}, z = ({\bf pure \:} \pi_1 p, \pi_2 p, {\bf let \: pure \:} p^{'} = \pi_2 p {\: \bf in \:} \pi_2 p^{'}) {\: \bf in \:} \langle x, \langle \pi_1 p^{'}, z \rangle\rangle] = & \\
-&\quad\quad\quad\quad\quad\quad \text{Substitution and reduction}& \\
-&[p, {\bf let \: pure \:} x, p^{'} = {\bf pure \:} \pi_1 p, \pi_2 p {\: \bf in \:} \langle x, \langle \pi_1 p^{'}, \pi_2 p^{'} \rangle\rangle]&
-\end{array}$
-
-\vspace{\baselineskip}
-
-Then:
-
-\vspace{\baselineskip}
-
-$\begin{array}{lll}
-&\tau_{\hat{A}, \hat{B} \times \hat{C}} \circ (id_{\hat{A}} \times \tau_{\hat{B}, \hat{C}}) \circ \alpha_{\hat{A}, \hat{B}, {\bf K}\hat{C}} = & \\
-&\quad\quad\quad\quad\quad\quad \text{Unfolding}& \\
-&\tau_{\hat{A}, \hat{B} \times \hat{C}} \circ [q, \langle \pi_1 q, {\bf let \: pure \:} y, z = {\bf pure \:} (\pi_1 (\pi_2 q)), \pi_2 (\pi_2 q) {\: \bf in \:} \langle y, z \rangle \rangle] \: \circ & \\
-&\quad\quad\quad\quad \circ \: [p, \langle \pi_1 (\pi_1 p), \langle \pi_2 (\pi_1 p), \pi_2 p \rangle \rangle] = & \\
-&\quad\quad\quad\quad\quad\quad \text{Composition}& \\
-&\tau_{\hat{A}, \hat{B} \times \hat{C}} \circ [p, \langle \pi_1 q, {\bf let \: pure \:} y, z = {\bf pure \:} (\pi_1 (\pi_2 q)), \pi_2 (\pi_2 q) {\: \bf in \:} \langle y, z \rangle \rangle & \\
-&\quad\quad\quad\quad [q := \langle \pi_1 (\pi_1 p), \langle \pi_2 (\pi_1 p), \pi_2 p \rangle \rangle]] = & \\
-&\quad\quad\quad\quad\quad\quad \text{Substitution and reduction}& \\
-&\tau_{\hat{A}, \hat{B} \times \hat{C}} \circ [p, \langle \pi_1 (\pi_1 p), {\bf let \: pure \:} y, z = {\bf pure \:} \pi_2 (\pi_1 p), \pi_2 p {\: \bf in \:} \langle y, z \rangle \rangle] = & \\
-&\quad\quad\quad\quad\quad\quad \text{Composition}& \\
-&[r, {\bf let \: pure \:} x, p^{'} = {\bf pure \:} \pi_1 r, \pi_2 r {\: \bf in \:} \langle x, \langle \pi_1 p^{'}, \pi_2 p^{'} \rangle\rangle] \: \circ & \\
-&\quad\quad\quad\quad \circ \: [p, \langle \pi_1 (\pi_1 p), {\bf let \: pure \:} y, z = {\bf pure \:} \pi_2 (\pi_1 p), \pi_2 p {\: \bf in \:} \langle y, z \rangle \rangle] = & \\
-&\quad\quad\quad\quad\quad\quad \text{Substitution and reduction}& \\
-&[p, {\bf let \: pure \:} x, p^{'} = ({\bf pure \:} (\pi_1 (\pi_1 p)), {\bf let \: pure \:} y, z = ({\bf pure \:} \pi_2 (\pi_1 p), \pi_2 p {\: \bf in \:} \langle y, z \rangle \rangle) & \\
-&\quad\quad\quad\quad {\: \bf in \:} \langle x, \langle \pi_1 p^{'}, \pi_2 p^{'} \rangle\rangle ] = & \\
-&\quad\quad\quad\quad\quad\quad \text{Reduction}& \\
-&[p, {\bf let \: pure \:} x,y,z = {\bf pure \:} (\pi_1 (\pi_1 p)), {\bf pure \:} \pi_2 (\pi_1 p), \pi_2 p {\: \bf in \:} \langle x, \langle \pi_1 p^{'}, \pi_2 p^{'} \rangle\rangle [p^{'} := \langle y, z \rangle]] = & \\
-&\quad\quad\quad\quad\quad\quad \text{Substitution and reduction}& \\
-&[p, {\bf let \: pure \:} x,y,z = {\bf pure \:} (\pi_1 (\pi_1 p)), {\bf pure \:} \pi_2 (\pi_1 p), \pi_2 p {\: \bf in \:} \langle x, \langle y, z \rangle\rangle]&
-\end{array}$
-
-\vspace{\baselineskip}
-
-2) Easy
-
-\end{proof}
-
-\begin{lemma}
-
-  {\bf K} is an applicative functor
-\end{lemma}
-
-\begin{proof}
-  Immediately follows from previous lemmas in the section.
-\end{proof}
-
-Similar to \cite{Abramsky}, we apply the translation from $\lambda_{{\bf K}}$ to some cartesian closed category with an abritraty applicative functor $\mathcal{K}$, then
-we have $[\![\Gamma \vdash M : A]\!] = [x, M [x_i := \pi_i x]]$, so $M =_{\beta \eta} N \Leftrightarrow [\![\Gamma \vdash M : A]\!] = [\![\Gamma \vdash N : A]\!]$.
-
-\end{proof}
-
-  \addcontentsline{toc}{section}{References}
-
-  \begin{thebibliography}{}
-
-  \bibitem{Artemov} Artemov S. and Protopopescu T., \/ ``Intuitionistic Epistemic Logic'', \textit{The
-  Review of Symbolic Logic}, 2016, vol. 9, no 2. pp. 266-298.\parskip=1mm
-
-  \bibitem{Krupski} Krupski V. N. and Yatmanov A., \/ ``Sequent Calculus for Intuitionistic Epistemic Logic
-  IEL'', \textit{Logical Foundations of Computer Science: International Symposium, LFCS 2016, Deerfield
-  Beach, FL, USA, January 4-7, 2016. Proceedings}, 2016, pp. 187-201.\parskip=1mm
-
-  \bibitem{Haskell} Haskell Language. // URL: https://www.haskell.org. (Date: 1.08.2017) \parskip=1mm
-
-  \bibitem{Idris} Idris. A Language with Dependent Types.// URL:https://www.idris-lang.org. (Date:
-  1.08.2017) \parskip=1mm
-
-  \bibitem{Purs} Purescript. A strongly-typed functional programming language that compiles to JavaScript.
-  URL: http://www.purescript.org. (Date: 1.08.2017) \parskip=1mm
-
-  \bibitem{Elm} Elm. A delightful language for reliable webapps. // URL: http://elm-lang.org. (Date:
-  1.08.2017) \parskip=1mm
-
-  \bibitem{Base} Hackage, \/ ``The base package'' // URL: https://hackage.haskell.org/package/base-4.10.0.0
-  (Date: 1.08.2017) \parskip=1mm
-
-  \bibitem{Miran} Lipovaca M, \/ ``Learn you a Haskell for Great Good!''. //URL:
-  http://learnyouahaskell.com/chapters (Date: 1.08.2017) \parskip=1mm
-
-  \bibitem{McP} McBride C. and Paterson R., ``Applicative programming with effects", \textit{Journal of
-  Functional Programming}, 2008, vol. 18, no 01. pp 1-13. \parskip=1mm
-
-  \bibitem{McP2} McBride C. and Paterson R, ``Functional Pearl. Idioms: applicative programming with
-  effects'', \textit{Journal of Functional Programming}, 2005. vol. 18, no 01. pp 1-20. \parskip=1mm
-
-  \bibitem{Neder} R. Nederpelt and H. Geuvers, ``Type Theory and Formal Proof: An Introduction''.
-  \textit{Cambridge University Press}, New York, NY, USA, 2014. pp. 436. \parskip=1mm
-
-  \bibitem{Morten} Sorensen M. H. and Urzyczyn P, ``Lectures on the Curry-Howard isomorphism'',
-  \textit{Studies in Logic and the Foundations of Mathematics}, vol. 149, \textit{Elsevier Science}, 1998.
-  pp 261. \parskip=1mm
-
-  \bibitem{Pierce} Pierce B. C., ``Types and Programming Languages''. \textit{Cambridge, Mass: The MIT
-  Press}, 2002. pp. 605. \parskip=1mm
-
-  \bibitem{Girard} Girard J.-Y., Taylor P. and  Lafont Y, ``Proofs and Types'', \textit{Cambridge University
-  Press}, New York, NY, USA, 1989. pp. 175. \parskip=1mm
-
-  \bibitem{Baren} Barendregt. H. P., ``Lambda calculi with types" // Abramsky S., Gabbay Dov M., and S. E.
-  Maibaum, ``Handbook of logic in computer science (vol. 2), Osborne Handbooks Of Logic In Computer
-  Science'', Vol. 2. \textit{Oxford University Press, Inc.}, New York, NY, USA, 1993. pp 117-309.
-  \parskip=1mm
-
-  \bibitem{Hindley} Hindley J. Roger, ``Basic Simple Type Theory''. \textit{Cambridge University Press}, New
-  York, NY, USA, 1997. pp. 185. \parskip=1mm
-
-  \bibitem{Lax} Pfenning F. and Davies R., ``A judgmental reconstruction of modal logic'',
-  \textit{Mathematical Structures in Computer Science}, vol. 11, no 4, 2001, pp. 511-540. \parskip=1mm
-
-  \bibitem{Baren2} H.P. Barendregt. The Lambda Calculus --- Its Syntax and Semantics. Studies in Logic and
-  the Foundations of Mathematics, vol. 103. Amsterdam: North-Holland, 1985.
-
-  \bibitem{ModalK} Yoshihiko KAKUTANI, A Curry-Howard Correspondence for Intuitionistic Normal Modal Logic, Computer Software, Released February 29, 2008, Online ISSN , Print ISSN 0289-6540.
-
-  \bibitem{ModalK1} Kakutani Y.
-(2007) Call-by-Name and Call-by-Value in Normal Modal Logic.
-In: Shao Z.
-(eds) Programming Languages and Systems.
-APLAS 2007.
-Lecture Notes in Computer Science, vol 4807.
-Springer, Berlin, Heidelberg
-
-  \bibitem{Abe} T. Abe. Completeness of modal proofs in first-order predicate logic. Computer Software, JSSST Journal, 24:165 -- 177, 2007.
-
-  \bibitem{LambekScott} Lambek, J. and Scott P.J. (1986) Introduction to Higher Order Categorical Logic, Cambridge Studies in Advanced Mathematics 7, Cambridge: Cambridge University Press.
-
-  \bibitem{ElKelly} Samuel Eilenberg and Max Kelly, Closed categories. Proc. Conf. Categorical Algebra (La Jolla, Calif., 1965).
-
-  \bibitem{Abramsky} Samson Abramsky and Nikos Tzevelekos, Introduction to Categories and Categorical Logic
-
-  \bibitem{ModalLa} G. A. Kavvos. The Many Worlds of Modal $\Lambda$--calculi: I. Curry-Howard for Necessity, Possibility and Time
-
-  \bibitem{Cons} Ross Paterson. in Mathematics of Program Construction, Madrid, 2012, Lecture Notes in Computer Science, vol. 7342, pp. 300--323, Springer, 2012.
-\end{thebibliography}
-
-\end{document}
+По умолчанию мы будем рассматривать ниже аппликативный функтор над декартово замкнутой категорией.

### 2018-02-25 MT/CategoryFAQ.tex
added more definitions

@@
   \end{itemize}
+\end{defin}
+
+\begin{defin} Функтор
+
+  Пусть $\mathcal{C}, \mathcal{D}$ -- категории. Функтором называется отображение $F : \mathcal{C} \to \mathcal{D}$, такое, что:
+  \begin{itemize}
+    \item $F : A \mapsto F A$, где $A \in Ob_{\mathcal{C}}$;
+    \item $F (g \circ f) = F(g) \circ F(f)$;
+    \item $F (id_A) = id_{F A}$.
+  \end{itemize}
+\end{defin}
+
+\begin{defin} Естественное преобразование
+  Пусть $\mathcal{F}, \mathcal{G} : \mathcal{C} \to \mathcal{D}$ -- функторы.
+  Естественным преобразованием $\alpha : \mathcal{F} \Rightarrow \mathcal{G}$ называется такое индексированное семейство стрелок $(\alpha_{X})_{X \in Ob_{\mathcal{C}}}$,
+что для любых $A, B \in Ob_{\mathcal{C}}$, для любой стрелки $f \in Hom_{\mathcal{C}}(A, B)$, диаграмма коммутирует:
+
+\xymatrix{
+&&&& \mathcal{F} A \ar[dd]_{\alpha_A} \ar[rr]^{\mathcal{F}(f)} && \mathcal{F} B \ar[dd]^{\alpha_B}\\
+\\
+&&&& \mathcal{G}A \ar[rr]_{\mathcal{G}(f)} && \mathcal{G} B
+}
 \end{defin}
@@
 
-Заметим, что декартово замкнутая категория -- это частный случай (симметрической) моноидальной категории, в котором тензор -- это произведения, а единица -- это терминальный объект.
+Легко видеть, что декартово замкнутая категория -- это частный случай (симметрической) моноидальной категории, в котором тензор -- это произведения, а единица -- это терминальный объект.
 

### 2018-02-25 MT/HaskellFAQ.tex
added more definitions

\section{Глоссарий по основным конструкциям функционального языка программирования Haskell: функторы, монады, аппликативные функторы}

\begin{defin} Класс типов

  Классом типов в языке Haskell -- это реализация некоторого общего интерфейса для совокупности типов.

  Представителем (или наследников) класса типов называется реализация данного класса для конкретного типа.
\end{defin}

\begin{defin} Функтор

  Функтор -- это однопараметрический класс типов, позволяющий пронести действие функции через значения,
  полученные в результате применения к их типу одноместного типового оператора.
\end{defin}

Определение в стандартной
библиотеке выглядит следующим образом:

\begin{lstlisting}[language=Haskell]
  class Functor f where
    fmap :: (a -> b) -> f a -> f b
\end{lstlisting}

Рассмотрим примеры:

\begin{itemize}
  \item Список (неограниченная в длине последовательность) является функтором:
  Определение в стандартной
  библиотеке выглядит следующим образом:

  \begin{lstlisting}[language=Haskell]
    instance Functor [] where
      fmap :: (a -> b) -> [a] -> [b]
      fmap f [] = []
      fmap f (x:xs) = (f x) : (fmap f xs)
  \end{lstlisting}

  Данный пример достаточно прост: реализация функтора для списка -- это функция высшего порядка, которая,
  принимая на входе одноместную функцию из типа $a$ в тип $b$ и список элементов типа $a$, возвращает список элементов типа $b$, который
  получен применением функции к каждому элементу списка, полученного на вход.
  \item Пара (тип декартова произведения типов) также функтор:
  \begin{lstlisting}[language=Haskell]
  instance Functor (b,) where
    fmap :: (a -> c) -> (b,a) -> (b,c)
    fmap f (x,y) = (x, f y)
  \end{lstlisting}

  Конструктор пары является двухпараметрическим типовым оператором, но мы сделали из него однопараметрический оператор фиксацией первого параметра.

  Данная реализация также довольно проста: на вход принимается функция из типа $a$ в тип $c$ и кортеж, в котором первая координата имеет тип $b$, а вторая -- тип $a$.
  На выходе мы получаем кортеж типа $(b,c)$, применяя полученную на вход функцию ко второй координате пары.

  \item Тип $Maybe$ -- это однопараметрический типовой оператор, для обработки неопределенных значений:
  \begin{lstlisting}[language=Haskell]
    data Maybe a = Nothing | Just a
  \end{lstlisting}

  Реализация функтора для типа $Maybe$:
  \begin{lstlisting}[language=Haskell]
    instance Functor Maybe where
      fmap :: (a -> b) -> Maybe a -> Maybe b
      fmap f Nothing = Nothing
      fmap f (Just x) = Just (f x)
  \end{lstlisting}

  Если второй аргумент является неопределенным значением (на вход передан $Nothing$), то и возвращается $Nothing$. Если же значение определено, то есть оно имеет вид $Just \: x$, тогда
  мы применяем функцию функцию к $x$, а результат вычисления оборачиваем в конструктор $Just$.
\end{itemize}

### 2018-03-25 MT/CategoryFAQ.tex
update

@@
-\section{Приложение А. Глоссарий по теории категорий.}
+\subsection{Приложение А. Глоссарий по теории категорий.}
 
@@
 
+Примеры категорий:
+\begin{itemize}
+  \item $Set$ -- категория всех множеств, объекты -- класс всех множеств, морфизмы -- теоретико-множественные функции;
+  \item $Group$ -- категория всех групп, морфизмы которой -- гомоморфизмы групп;
+  \item Пусть $\langle A, \preceq \rangle$ -- частично-упорядоченное множество;
+Тогда $\langle A, \preceq \rangle$ -- это категория, объекты которой -- элементы множества $A$, а морфизмы определяются порядком: $a \rightarrow b$ тогда и только тогда, когда $a \preceq b$, для $a, b \in A$;
+  \item $Top$ -- категория топологических пространств и непрерывных отображений между ними.
+  \item Каждый моноид $\langle M, \cdot, e \rangle$ является категорией: моноид можно определять как категорию с одним объектом.
+\end{itemize}
+
 \begin{defin} Функтор
@@
 \end{defin}
+
+Примеры функторов:
+\begin{itemize}
+  \item Булеан множества является эндофунктором $\mathcal{P}() : Set \Rightarrow Set$.
+  \item Пусть $AssocRing$ -- категория ассоциативных колец, $Semi$ -- категория полугрупп.
+$F : AssocRing \Rightarrow Semi$ -- забывающий функтор, который берет кольцо $\langle A, +, \cdot \rangle$ ``стирает'' аддитивную структуру кольца, оставляя при этом мультипликативную, и возвращает
+полугруппу $\langle A, \cdot \rangle$;
+  \item Пусть $\langle G, \cdot, {}^{-1}, e \rangle$ -- группа. Коммутантом группы $G$ называется называется группа $[G,G] = \{ x \cdot y \cdot x^{-1} \cdot y^{-1} | x, y \in G \}$.
+Легко проверить, что коммутант функториален: гомоморфизм групп переводится в гомоморфизм между порожденными им коммутантами, сохраняя при этом композицию гоморфзимов и тождественный гомоморфизм;
+  \item Гомоморфизм моноидов является функтором (поскольку каждый моноид является категорией).
+\end{itemize}
 
@@
     \item Единица $\mathds{1}$;
-    \item Изоморфизм, который мы будем называть ассоциатором: $\alpha_{A,B,C}: (A \otimes B) \otimes C \cong A \otimes (B \otimes C)$;
+    \item Изоморфизм, который мы будем называть ассоциатором: для любых $A, B, C \in Ob_{\mathcal{C}}, \alpha_{A,B,C}: (A \otimes B) \otimes C \cong A \otimes (B \otimes C)$;
     \item Изоморфизм  $L_A : \mathds{1} \otimes A \cong A$;
@@
     }
-
-    \item Моноидальная категория $\mathcal{C}$ называется симметрической, если для любых $A,B \in Ob_{\mathcal{C}}$, имеет место изоморфизм $\sigma_{A,B} : A \otimes B \cong B \otimes A$.
   \end{itemize}
@@
 
-  Декартово замкнутная категория -- это категория с терминальным объектом, конечными произведениями и экспоненцированием.
+  Декартово замкнутная категория -- это категория с терминальным объектом, произведениями и экспоненцированием:
+
+1) Объект $\mathds{1}$ в категории $\mathcal{C}$ называется терминальных, если для любого объекта $A \in Ob_{\mathcal{C}}$ и для любых морфизмов $f, g \in Hom_{\mathcal{C}}(A, 1)$, $f = g$.
+
+2) Пусть $A, B \in Ob_{\mathcal{C}}$, тогда произведением объектов $A$ и $B$ называется объект $A \times B$, такой, для любого $C \in Ob_{\mathcal{C}}$ и для любых морфизмов $f \in Hom_{\mathcal{C}}(C, A)$ и $g \in Hom_{\mathcal{C}}(C, B)$,
+что диаграмма коммутирует:
+
+\xymatrix{
+&&&&&&& C \ar[dr]^{f} \ar[dl]_{g} \ar[d]^{\langle f,g \rangle} \\
+&&&&&& A & \ar[l]^{\pi_1} A \times B \ar[r]_{\pi_2} & B
+}
+
+Морфизм $\langle f,g \rangle$ называется парой морфизмов $f$ и $g$, а морфизмы вида $\pi_1$ и $\pi_2$ -- каноническими проекциями.
+
+3) Пусть $A, B \in Ob_{\mathcal{C}}$, тогда экспонентой объектов $A$ и $B$ называется объект $B^A$, такой, что диаграмма коммутирует для любого объекта $C \in Ob_{\mathcal{C}}$ и для любого морфизма $f \in Hom_{\mathcal{C}}(C \times A, B)$:
+
+\xymatrix{
+&&&&&& B^A \times A \ar[r]^{\epsilon_{A,B}} & B \\
+&&&&&& C \times A \ar[u]^{\Lambda(f) \times id_A} \ar[ur]_{f}
+}
+
+Морфизмы вида $\epsilon_{A,B}$ называются вычисляющими стрелками, а морфизмы вида $\Lambda(f)$ -- каррированием стрелки $f$.
 \end{defin}
 
-Легко видеть, что декартово замкнутая категория -- это частный случай (симметрической) моноидальной категории, в котором тензор -- это произведения, а единица -- это терминальный объект.
+Примеры декартово замкнутных категорий:
+\begin{itemize}
+  \item Категория всех множеств $Set$: терминальный объект -- любой синглетон, произведение -- декартово произведение, экспоненциальный объект -- множество $B^A = \{ f \: | \: f : A \to B \}$;
+  \item Пусть $A$ -- алгебра Гейтинга, тогда терминальным объектом будет верхний элемент множества $A$, произведение -- точная нижняя грань, экспоненциальный объект -- импликация;
+\end{itemize}
 
-\begin{defin} Нестрогий моноидальный функтор
+\begin{defin} Моноидальный функтор
 
@@
 
-  Нестрогий моноидальный функтор $\mathcal{F} : \langle \mathcal{C}, \otimes_1, \mathds{1} \rangle \to \langle \mathcal{D}, \otimes_2, \mathds{1}' \rangle$ это функтор
+  Моноидальный функтор $\mathcal{F} : \langle \mathcal{C}, \otimes_1, \mathds{1} \rangle \to \langle \mathcal{D}, \otimes_2, \mathds{1}' \rangle$ это функтор
   $\mathcal{F} : \mathcal{C} \to \mathcal{D}$ с дополнительными естественными преобразованиями:
@@
 
-\begin{defin}
-
-Тензорно-сильный функтор -- это эндофунктор над моноидальной категорией с дополнительным естественным преобразованием и условиями когерентности для него (ниже соответствующие коммутирующие диаграмы):
-
-\begin{center}
-$\begin{array}{lll}
-  \tau_{A, B} : A \otimes \mathcal{K}B \rightarrow \mathcal{K}(A \otimes B)
-\end{array}$
-\end{center}
-
-\xymatrix{
-(A \otimes B) \otimes \mathcal{K}C \ar[d]_{\alpha_{A, B, \mathcal{K}C}}\ar[rrrr]^{\tau_{A \otimes B, C}} &&&& \mathcal{K}((A \otimes B) \otimes C) \ar[d]^{\mathcal{K}(\alpha_{A,B,C})}\\
-A \otimes (B \otimes \mathcal{K}C) \ar[rr]_{id_A \otimes \tau_{B,C}} && A \otimes \mathcal{K}(B \otimes C) \ar[rr]_{\quad \tau_{A, (B \otimes C)}} && \mathcal{K}(A \otimes (B \otimes C)) \\
-}
-
-\xymatrix{
-&&& \mathds{1} \otimes \mathcal{K}A \ar[drr]_{R_{\mathcal{K}A}} \ar[rr]^{\mu_{\mathds{1}, A}} && \mathcal{K}(\mathds{1} \otimes A) \ar[d]^{\mathcal{K}(R_A)}\\
-&&&&& \mathcal{K}A
-}
-\end{defin}
-
 \begin{defin} Аппликативный функтор
@@
   Аппликативный функтор -- это тройка $\langle \mathcal{C}, \mathcal{K}, \eta \rangle$,
-где $\mathcal{C}$ -- это моноидальная категория, $\mathcal{K}$ - это тензорно-сильный нестрогий моноидальный эндофунктор и
+где $\mathcal{C}$ -- это моноидальная категория, $\mathcal{K}$ - это моноидальный эндофунктор и
 $\eta : Id_{\mathcal{C}} \Rightarrow \mathcal{K}$ -- это естественное преобразование, такое, что:
@@
 }
-\item $\tau_{A, B} = \ast_{A, B} \circ \eta_{A} \otimes id_{\mathcal{K}B}$.
 \end{itemize}

### 2018-03-25 MT/CategoryFAQ.tex
modify

@@
 
-Примеры категорий:
-\begin{itemize}
-  \item $Set$ -- категория всех множеств, объекты -- класс всех множеств, морфизмы -- теоретико-множественные функции;
-  \item $Group$ -- категория всех групп, морфизмы которой -- гомоморфизмы групп;
-  \item Пусть $\langle A, \preceq \rangle$ -- частично-упорядоченное множество;
-Тогда $\langle A, \preceq \rangle$ -- это категория, объекты которой -- элементы множества $A$, а морфизмы определяются порядком: $a \rightarrow b$ тогда и только тогда, когда $a \preceq b$, для $a, b \in A$;
-  \item $Top$ -- категория топологических пространств и непрерывных отображений между ними.
-  \item Каждый моноид $\langle M, \cdot, e \rangle$ является категорией: моноид можно определять как категорию с одним объектом.
-\end{itemize}
-
 \begin{defin} Функтор
@@
 \end{defin}
-
-Примеры функторов:
-\begin{itemize}
-  \item Булеан множества является эндофунктором $\mathcal{P}() : Set \Rightarrow Set$.
-  \item Пусть $AssocRing$ -- категория ассоциативных колец, $Semi$ -- категория полугрупп.
-$F : AssocRing \Rightarrow Semi$ -- забывающий функтор, который берет кольцо $\langle A, +, \cdot \rangle$ ``стирает'' аддитивную структуру кольца, оставляя при этом мультипликативную, и возвращает
-полугруппу $\langle A, \cdot \rangle$;
-  \item Пусть $\langle G, \cdot, {}^{-1}, e \rangle$ -- группа. Коммутантом группы $G$ называется называется группа $[G,G] = \{ x \cdot y \cdot x^{-1} \cdot y^{-1} | x, y \in G \}$.
-Легко проверить, что коммутант функториален: гомоморфизм групп переводится в гомоморфизм между порожденными им коммутантами, сохраняя при этом композицию гоморфзимов и тождественный гомоморфизм;
-  \item Гомоморфизм моноидов является функтором (поскольку каждый моноид является категорией).
-\end{itemize}
 
@@
 }
+где $\Lambda(f) \times id_A = \langle \Lambda(f) \circ \pi_1, id_A \circ \pi_2 \rangle$
 
@@
 \end{defin}
-
-Примеры декартово замкнутных категорий:
-\begin{itemize}
-  \item Категория всех множеств $Set$: терминальный объект -- любой синглетон, произведение -- декартово произведение, экспоненциальный объект -- множество $B^A = \{ f \: | \: f : A \to B \}$;
-  \item Пусть $A$ -- алгебра Гейтинга, тогда терминальным объектом будет верхний элемент множества $A$, произведение -- точная нижняя грань, экспоненциальный объект -- импликация;
-\end{itemize}
 

### 2018-03-25 MT/References.tex
modify

--- 2018-02-13 article.tex
@@
-\documentclass[a4paper]{article}
-\usepackage{amsmath}
-\usepackage{amsthm}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{bussproofs}
-\usepackage{mathtools}
-\usepackage{verbatim}
-\usepackage{dsfont}
-\usepackage{mathabx}
-\usepackage[all, 2cell]{xy}
-\usepackage[all]{xy}
-\newtheorem{theorem}{Theorem}
-\newtheorem{prop}{Proposition}
-\newtheorem{lemma}{Lemma}
-\newtheorem{defin}{Definition}
-\newtheorem{ex}{Example}
-\newtheorem{col}{Corollary}
-\newtheorem{con}{Consequence}
-\usepackage{listings} 		% for source code
-\date{}
-\title{Modal type theory based on the intuitionistic epistemic logic}
-
-\begin{document}
-\maketitle
-
-\begin{abstract}
-Modal intuitionistic epistemic logic IEL$^{-}$ was proposed by S.Artemov and T. Protopopescu as the formal
-foundation for the intuitionistic theory of knowledge. We construct a modal simply typed lambda-calculus
-which is Curry-Howard isomorphic to IEL$^{-}$ as formal theory of calculations with applicative functors in
-functional programming languages like Haskell or Idris.
-\end{abstract}
-
-  % This is a comment, not shown in final output.
-  % The following shows typesetting  power of LaTeX:
-
-  \section{Introduction}
-
-  Modal intutionistic epistemic logic IEL was proposed by S. Artemov and T. Proropopescu \cite{Artemov}. IEL
-  provides the epistimology and the theory of knowledge as based on BHK-semantics of intuitionistic logic.
-  IEL$^{-}$  is a variant of IEL, that corresponds to intuitionistic belief.
-  Informally, ${\bf K}A$ denotes that $A$ is verified intuitionistically.
-
-  Intuitionistic epistemic logic IEL$^{-}$ is defined by following axioms and derivation rules:
-
-  \begin{defin} Intuitionistic epistemic logic IEL:
-
-    1) IPC axioms;
-
-    2) ${\bf K}(A \to B) \to ({\bf K}A \to {\bf K}B)$ (normality);
-
-    3) $A \to {\bf K}A$ (co-reflection);
-
-    Rule: MP.
-
-  \end{defin}
-
-  V. Krupski and  A. Yatmanov provided the sequential calculus for IEL and proved that this calculus is
-  PSPACE-complete \cite{Krupski}.
-
-  Functional programming languages such as Haskell \cite{Haskell}, Idris \cite{Idris}, Purescript \cite{Purs}
-  Elm \cite{Elm} or Scala \cite{Scala} have special type classes\footnote{Type class in Haskell is a general interface for
-  special group of datatypes.} for calculations with container types like \verb"Functor" and
-  \verb"Applicative" \footnote{Reader may read more about container types in the Haskell standard library documentation\cite{Base} or in the next one textbook
-  \cite{Miran}}:
-
-  \begin{lstlisting}[language=Haskell]
-    class Functor f where
-      fmap :: (a -> b) -> f a -> f b
-
-    class Functor f => Applicative f where
-      pure :: a -> f a
-      (<*>) :: f (a -> b) -> f a -> f b
-  \end{lstlisting}
-
-  By \emph{container} (or \emph{computational context}) type we mean some type-operator $f$, where $f$ is a
-  ``function'' from $*$ to $*$: type operator takes a simple type (ща kind $*$) and returns another
-  simple type of kind \verb"*". For more detailed description of the type system with kinds used in
-  Haskell see \cite{Morten}.
-
-Applicative functor allows to generalize the action of a functor for functions with arbitrary number of arguments, for instance:
-  \begin{lstlisting}[language=Haskell]
-  liftA2 :: Applicative f => (a -> b -> c) -> f a -> f b -> f c
-  liftA2 f x y = ((pure f) <*> x) <*> y
-  \end{lstlisting}
-
-  It's not difficult to see that modal axioms in $IEL^{-}$ and types of the methods of Applicative class in
-  Haskell-like languages (which is described below) are syntactically similar and we are going to show that
-  this coincidence has a non-trivial computational meaning.
-
-  We investigate the relationship between intuitionistic epistemic logic IEL$^{-}$ and applicative programming with side-effects
-  by constructing the type system (which is called $\lambda_{{\bf K}}$) which is Curry-Howard isomorphic to $IEL^{-}$.
-  So we will consider ${\bf K}$-modality as an arbitrary applicative functor and we prove that obtained type system
-  is sound and complete for applicative functor on cartesian closed category (using the categorical definition proposed by Paterson \cite{Cons}).
-
-  $\lambda_{{\bf K}}$ consists of the rules for simply typed lambda-calculus and special typing rules for
-  lifting types into the applicative functor ${\bf K}$. We assume that our type system will
-  axiomatize the simplest case of computation with effects with one container. We provide a proof-theoretical
-  view at this kind of computations in functional programming and prove strong normalization and confluence.
-
-  \section{Typed lambda-calculus based on IEL$^{-}$}
-
-  The first is to define the natural deduction calculus for NIEL$^{-}$ :
-
-  \begin{defin} Natural deduction NIEL$^{-}$ for IEL$^{-}$ is an extenstion of intuitionistic natural deduction calculus
-    with additional inference rules for modality:
-
-  \begin{minipage}{0.5\textwidth}
-    \begin{flushleft}
-  	\begin{prooftree}
-    	\AxiomC{$\Gamma \vdash A$}
-    	\RightLabel{${\bf K}_I$}
-    	\UnaryInfC{$\Gamma \vdash {\bf K}A$}
-  \end{prooftree}
-    \end{flushleft}
-  \end{minipage}
-  \begin{minipage}{0.5\textwidth}
-    \begin{flushright}
-  	\begin{prooftree}
-    \AxiomC{$\Gamma \vdash {\bf K} A_1, \dots, \Gamma \vdash {\bf K} A_n $}
-    \AxiomC{$A_1,\dots,A_n \vdash B$}
-    \BinaryInfC{$\Gamma \vdash {\bf K} B$}
-  	\end{prooftree}
-    \end{flushright}
-  \end{minipage}
-  \end{defin}
-
-The first rule allows to derive co-reflextion. The second modal rule is a counterpart of $\Box_I$
-rule in natural deduction calculus for constructive K (see \cite{ModalLa}).
-
-We will denote $\Gamma \vdash {\bf K} A_1, \dots, \Gamma \vdash {\bf K} A_n$ and $A_1,\dots,A_n \vdash B$ as $\Gamma \vdash {\bf K} \vec{A}$ and $\vec{A} \vdash B$ for brevity.
-
-  \vspace{\baselineskip}
-
-  \begin{lemma}
-    $\Gamma \vdash_{\text{NIEL}^{-}} A \Rightarrow$ IEL$^{-} \vdash \bigwedge \Gamma \rightarrow A$.
-  \end{lemma}
-
-  \begin{proof}
-Induction on the derivation.
-
-  \vspace{\baselineskip}
-
-Let us consider cases with modality.
-
-\vspace{\baselineskip}
-
-1) If $\Gamma \vdash_{\text{NIEL}^{-}} A$, then $\text{IEL}^{-} \vdash \bigwedge \Gamma \rightarrow {\bf K}A$.
-
-$\begin{array}{lll}
-(1) & \bigwedge \Gamma \rightarrow A & \text{assumption}\\
-(2) & A \rightarrow {\bf K}A &\text{co-reflection}\\
-(3) & (\bigwedge \Gamma \rightarrow A) \rightarrow ((A \rightarrow {\bf K}A) \rightarrow (\bigwedge \Gamma \rightarrow {\bf K}A))&\text{IPC theorem}\\
-(4) & (A \rightarrow {\bf K}A) \rightarrow (\bigwedge \Gamma \rightarrow {\bf K}A) &\text{from (1), (3) and MP}\\
-(5) & \bigwedge \Gamma \rightarrow {\bf K}A &\text{from (2), (4) and MP}\\
-\end{array}$
-
-\vspace{\baselineskip}
-
-2) If $\Gamma \vdash_{\text{NIEL}^{-}} {\bf K} \vec{A}$ and $\vec{A} \vdash B$, then $\text{IEL}^{-} \vdash \bigwedge \Gamma \rightarrow {\bf K}B$.
-
-$\begin{array}{lll}
-(1) &\bigwedge \Gamma \rightarrow \bigwedge \limits_{i = 1}^{n} {\bf K} A_i & \text{assumption} \\
-(2) &\bigwedge \limits_{i = 1}^{n} {\bf K} A_i \rightarrow {\bf K} \bigwedge \limits_{i = 1}^{n} A_i& \text{IEL$^{-}$ theorem} \\
-(3) &\bigwedge \Gamma \rightarrow {\bf K} \bigwedge \limits_{i = 1}^{n} A_i & \text{from (1), (2) and transitivity} \\
-(4) &\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B& \text{assumption} \\
-(5) &(\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B) \rightarrow {\bf K} (\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B)& \text{co-reflection}\\
-(6) &{\bf K} (\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B)& \text{from (4), (5) and MP} \\
-(7) &{\bf K} \bigwedge \limits_{i = 1}^{n} A_i \rightarrow {\bf K} B & \text{from (6) and normality} \\
-(8) &\bigwedge \Gamma \rightarrow {\bf K} B & \text{from (3), (7) and transitivity}
-\end{array}$
-
-  \end{proof}
-
-\begin{lemma}
-  $ $
-  If $\text{IEL}^{-} \vdash A$, then $\text{NIEL}^{-} \vdash A$.
-\end{lemma}
-
-\begin{proof}
-  Straightforward derivation of modal axioms in $\text{NIEL}^{-}$. We consider this derivation below using terms.
-\end{proof}
-
-  \vspace{\baselineskip}
-
-  At the next step we build the typed lambda-calculus based on the NIEL$^{-}$ with implication and dijunction
-by proof-assingment in rules. Obtained fragment is equivalent to IEL$^{-}$ without axioms for negation and disjunction.
-
-  At first, we define lambda-terms and types for this lambda-calculus.
-
-  \vspace{\baselineskip}
-
-  \begin{defin} The set of terms:
-
-  Let $\mathbb{V}$ be the set of variables. The set $\Lambda_{{\bf K}}$ of terms is defined by the grammar:
-
-$\begin{array}{lll}
-& \Lambda_{{\bf K}} ::= \mathbb{V} \: | \:  (\lambda \mathbb{V}.\Lambda_{{\bf K}}) \: | \: (\Lambda_{{\bf K}}\Lambda_{{\bf K}}) \: | \: (\Lambda_{{\bf K}} , \Lambda_{{\bf K}}) \: | \: (\pi_1 \Lambda_{{\bf K}}) \: | \: (\pi_2 \Lambda_{{\bf K}}) \: | & \\
-& \quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad ({\bf pure \: } \: \Lambda_{{\bf K}}) \: | \: ({\bf let \:\: pure} \:\: \mathbb{V}^{*} = \Lambda_{{\bf K}}^{*} \:\: {\bf in} \:\: \Lambda_{{\bf K}})
-\end{array}$
-
-  \end{defin}
-
-Where $\mathbb{V}^{*}$ and $\Lambda_{{\bf K}}^{*}$ denote the set of finite sequences of variables $\bigcup \limits_{i=0}^{\infty} \mathbb{V}^i$
-and the set of finite sequences of terms $\bigcup \limits_{i = 0}^{\infty} {\Lambda_{{\bf K}}}^i $.
-Note that the sequence of variables $\vec{x}$ and the sequence of terms $\vec{M}$ should have the same length.
-Otherwise, term is not well-formed.
-
-  \begin{defin} The set of types:
-
-  Let $\mathbb{T}$ be the set of atomic types. The set $\mathbb{T}_{{\bf K}}$ of types with applicative
-  functor ${\bf K}$ is generated by the grammar:
-  \begin{equation}
-    \mathbb{T}_{{\bf K}} ::= \mathbb{T} \: | \: (\mathbb{T}_{{\bf K}} \to \mathbb{T}_{{\bf K}}) \: |
-    \: (\mathbb{T}_{{\bf K}} \times \mathbb{T}_{{\bf K}}) \: | \: ({\bf K}\mathbb{T}_{{\bf K}})
-  \end{equation}
-  \end{defin}
-
-  Context, domain of context and range of context are defined standardly \cite{Neder}\cite{Morten}.
-
-  Our type system is based on the Curry-style typing rules:
-
-  \begin{defin} Modal typed lambda calculus $\lambda_{{\bf K}}$ based on NIEL$^{-}_{\land, \to}$:
-
-    \begin{center}
-    \begin{prooftree}
-    \AxiomC{$ $}
-    \RightLabel{\scriptsize{ax}}
-    \UnaryInfC{$\Gamma , x : A \vdash x : A$}
-    \end{prooftree}
-    \end{center}
-
-    \begin{minipage}{0.45\textwidth}
-      \begin{prooftree}
-      \AxiomC{$\Gamma, x : A \vdash M : B$}
-      \RightLabel{$\rightarrow_i$}
-      \UnaryInfC{$\Gamma \vdash \lambda x. M : A \to B$}
-      \end{prooftree}
-
-      \begin{prooftree}
-      \AxiomC{ $\Gamma \vdash M : A$ }
-      \AxiomC{ $\Gamma \vdash N : B$ }
-      \RightLabel{$\times_i$}
-      \BinaryInfC{$\Gamma \vdash \langle M, N \rangle : A \times B$}
-      \end{prooftree}
-
-      \begin{prooftree}
-        \AxiomC{$\Gamma \vdash M : A$}
-        \RightLabel{${\bf K}_I$}
-        \UnaryInfC{$\Gamma \vdash {\bf pure \: } \: M : {\bf K}A $}
-      \end{prooftree}
-  \end{minipage}%
-  \hfill
-  \begin{minipage}{0.45\textwidth}
-  \begin{tabular}{p{\textwidth}}
-    \begin{prooftree}
-    \AxiomC{$\Gamma \vdash M : A \to B$}
-    \AxiomC{$\Gamma \vdash N : A$}
-    \RightLabel{$\rightarrow_e$}
-    \BinaryInfC{$\Gamma \vdash MN : B$}
-    \end{prooftree}
-
-    \begin{prooftree}
-    \AxiomC{ $\Gamma \vdash M : A_1 \times A_2$ }
-    \RightLabel{$\times_e$, $i \in \{ 1, 2 \}$}
-    \UnaryInfC{$\Gamma \vdash \pi_i M : A_i$}
-    \end{prooftree}
-
-    \begin{prooftree}
-      \AxiomC{$\Gamma \vdash \vec{M} : {\bf K} \vec{A}$}
-      \AxiomC{$\vec{x} : \vec{A} \vdash N : B$}
-      \RightLabel{$\text{let}_{{\bf K}}$}
-      \BinaryInfC{$\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } N : {\bf K} B$}
-    \end{prooftree}
-  \end{tabular}
-  \end{minipage}%
-
-  \end{defin}
-
-  ${\bf K}_I$-typing rule is the same as $\bigcirc$-introduction in lax logic (also known as monadic
-  metalanguage \cite{Lax}) and in typed lambda-calculus which is derived by proof-assignment for lax-logic
-  proofs.
-  ${\bf K}_I$ allows to inject an object of type $\alpha$ into the functor. ${\bf K}_I$  reflects the
-  Haskell method {\bf pure} for Applicative class. It plays the same role as the {\bf return} method in Monad class.
-
-  $\text{let}_{{\bf K}}$ is similar to the $\Box$-rule in typed lambda calculus for intuitionistic normal modal logic {\bf IK}, which is described in \cite{ModalK}.
-
-$\Gamma \vdash \vec{M} : {\bf K} \vec{A}$ is a syntax sugar for the sequence $\Gamma \vdash M_1 : {\bf K}A_1,\dots,\Gamma \vdash M_n : {\bf K}A_n$ and $\vec{x} : \vec{A} \vdash N : B$ is a short form for $x_1 : A_1, \dots, x_n : A_n \vdash N : B$.
-${\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } N$ is a simultaneous local binding in $N$.
-We use this short form instead of ${\bf let \: pure\:} x_1,\dots,x_n = M_1,\dots,M_n {\: \bf in \:} N$.
-
-In fact, our calculus is the extention of typed lambda calculus for {\bf IK} by ${\bf K}_I$-rule that is appropriate to co-reflection.
-
-  \vspace{\baselineskip}
-
-Here are some examples of closed terms:
-
-\begin{prooftree}
-\AxiomC{$x : A \vdash x : A$}
-\UnaryInfC{$x : A \vdash {\bf pure \:} x : {\bf K} A$}
-\UnaryInfC{$\vdash (\lambda x. {\bf pure \: } x) : A \to {\bf K}A$}
-\end{prooftree}
-
-\begin{prooftree}
-\AxiomC{$f : {\bf K} (A \to B) \vdash f : {\bf K} (A \to B)$}
-\AxiomC{$x : {\bf K} A \vdash x : {\bf K} A $}
-\AxiomC{$g : A \to B \vdash g : A \to B$}
-\AxiomC{$y : A \vdash y : A$}
-\RightLabel{$\to_e$}
-\BinaryInfC{$g : A \to B, y : A \vdash g y : B$}
-\RightLabel{$\text{let}_{{\bf K}}$}
-\TrinaryInfC{$f : {\bf K} (A \to B), x : {\bf K} A \vdash {\bf let \: pure \:} g, y  = f, x {\: \bf in \:} g y : {\bf K} B$}
-\UnaryInfC{$f : {\bf K} (A \to B) \vdash \lambda x. {\bf let \: pure\:} g, y = f, x {\: \bf in \:} g y : {\bf K} A \to {\bf K} B$}
-\UnaryInfC{$\vdash \lambda f. \lambda x. {\bf let \:pure \:} g, y = f, x {\: \bf in \:} g y : {\bf K}(A \to B) \to {\bf K} A \to {\bf K} B$}
-\end{prooftree}
-
-  \vspace{\baselineskip}
-
-  Now we define free variables and substitutions. $\beta$-reduction, multi-step $\beta$-reduction and $\beta$-equality are defined standardly:
-
-  \begin{defin} The set $FV(M)$ of free variables for a term $M$:
-
-  1) $FV(x) = \{ x \}$;
-
-  2) $FV(\lambda x. M) = FV(M) \setminus \{ x\}$;
-
-  3) $FV(M N) = FV(M) \cup FV(N);$
-
-  4) $FV(\langle M,N \rangle) = FV(M) \cup FV(N)$;
-
-  5) $FV(\pi_i M) \subseteq FV(M)$, $i \in \{ 1, 2\}$;
-
-  6) $FV(\text{pure } M) = FV(M)$;
-
-  7) $FV({\bf let \: pure} \: \vec{x} = \vec{M} \:\: {\bf in} \:\: N) = \bigcup \limits_{i = 1}^n FV(M), \text{where $n = |\vec{M}|$}$.
-  \end{defin}
-
-  \begin{defin} Substitution:
-
-  1) $x [x := N] = N$, $x [y := N] = x$;
-
-  2) $(M N) [x := N] = M[x := N] N [x := N]$;
-
-  3) $(\lambda x. M) [x := N] = \lambda x. M [y := N]$, $y \in FV(M)$;
-
-  4) $(M, N)[x := P] = (M[x := P], N [x := P])$;
-
-  5) $(\pi_i M) [x := P] = \pi_i (M[x := P])$, $i \in \{ 1, 2\}$;
-
-  6) $({\bf pure \: } M) [x := P] = {\bf pure \: } (M [x := P])$;
-
-  7) $({\bf let \: pure \:}\vec{x} = \vec{M} {\: \bf in \:} N) [y := P] = {\bf let \: pure \:} \vec{x} = (\vec{M} [y := P]) {\: \bf in \:} N$.
-  \end{defin}
-
-Substitution and free variable for terms of kind ${\bf let \: pure}$ are defined similary to \cite{ModalK}.
-
-  \begin{defin} Type substituition
-
-    The substituition of type $C$ for type variable $B$ in type $A$ is defined inductively:
-
-    1) $B [B := C] = B$ and $D [B := C] = D$, if $B \neq D$;
-
-    2) $(A_1 \alpha A_2)[B := C] = (A_1[B := C]) \alpha (A_2[B := C])$, where $\alpha \in \{ \to, \times \}$;
-
-    3) $({\bf K}A)[B := C] = {\bf K}(A [B := C])$;
-
-    4) Let $\Gamma$ be the context, then $\Gamma [B := C] = \{ x : (A[B := C]) \: | \: x : A \in \Gamma \}$.
-  \end{defin}
-
-  \begin{defin} $\beta$-reduction and $\eta$-reduction rules for $\lambda_{{\bf K}}$.
-
-  1) $(\lambda x. M) N \rightarrow_{\beta} M [x := N]$;
-
-  2) $\pi_1 \langle M, N \rangle \rightarrow_{\beta} M$;
-
-  3) $\pi_2 \langle M, N \rangle \rightarrow_{\beta} N$;
-
-  4) $\begin{array}{llll}
-  & {\bf let \: pure \:} \vec{x}, y, \vec{z} = \vec{M}, {\bf let \: pure \: } \vec{w} = \vec{N} {\: \bf in \: } Q, \vec{P} {\: in \:} R \rightarrow_{\beta} & \\
-  & {\bf let \: pure \:} \vec{x}, \vec{w}, \vec{z} = \vec{M}, \vec{N}, \vec{P} {\: \bf in \: } R [y := Q]
-  \end{array}$
-
-  5) ${\bf let \: pure \:} \vec{x} = {\bf pure \:} \vec{M} {\: \bf in \:} N \rightarrow_{\beta} {\bf pure \:} N [\vec{x} := \vec{M}]$
-
-  6) ${\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} M \rightarrow_{\beta} {\bf pure \:} M$, where \underline{\quad} is an empty sequence of terms.
-
-  7) $\lambda x. f x \rightarrow_{\eta} f$;
-
-  8) $\langle \pi_1 P, \pi_2 P \rangle \rightarrow_{\eta} P$;
-
-  9) ${\bf let \: pure \:} x = M {\: \bf in \: } x \rightarrow_{\eta} M$;
-
-  \end{defin}
-
-By default we use call-by-name evaluation strategy.
-
-This commuting relation denotes that we identify terms of kind {\bf let pure} if they are differ in the order of binding.
-
-Now we will prove standard lemmas for contexts in type systems\footnote{We will not prove cases with
-$\to$-constructor, they are proved standardly in the same lemmas for simply typed lambda calculus, for
-example see \cite{Neder} \cite{Morten}. We will consider only modal cases}:
-
-\begin{lemma} Generation for ${\bf K}_I$.
-
-  Let $\Gamma \vdash {\bf pure \:} M : {\bf K}A$, then $\Gamma \vdash M : A$;
-\end{lemma}
-
-\begin{proof}
-  Straightforwardly.
-\end{proof}
-
-\begin{lemma} Basic lemmas.
-
-\begin{itemize}
-  \item If $\Gamma \vdash M : A$ and $\Gamma \subseteq \Delta$, then $\Delta \vdash M : A$;
-  \item If $\Gamma \vdash M : A$, then $\Delta \vdash M : A$, where $\Delta = \{ x_i : A_i \: | \: (x_i : A_i) \in \Gamma \: \& \: x_i \in FV(M) \}$
-  \item If $\Gamma, x : A \vdash M : B$ and $\Gamma \vdash N : A$, then $\Gamma \vdash M [x := N] : B$.
-  \item If $\Gamma \vdash M : A$, then $\Gamma [B := C] \vdash M : (A [B := C])$;
-  \item If $\Gamma \vdash M : A$ and $\Delta$ is a permutation of $\Gamma$, then $\Delta \vdash M : A$.
-\end{itemize}
-\end{lemma}
-
-\begin{proof}
-  $ $
-
-1) The derivation ends in
-
-\begin{prooftree}
-  \AxiomC{$\Gamma \vdash \vec{M} : {\bf K} \vec{A}$}
-  \AxiomC{$\vec{x} : \vec{A} \vdash N : B$}
-  \RightLabel{$\text{let}_{{\bf K}}$}
-  \BinaryInfC{$\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } N : {\bf K}B$}
-\end{prooftree}
-
-By IH $\Delta \vdash \vec{M} : {\bf K} \vec{A}$, so $\Delta \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } N : {\bf K}B$.
-
-  \vspace{\baselineskip}
-
-2) Similary.
-
-  \vspace{\baselineskip}
-
-3) The derivation ends in
-
-\begin{prooftree}
-  \AxiomC{$\Gamma, z : C \vdash \vec{M} : {\bf K} \vec{A}$}
-  \AxiomC{$\vec{x} : \vec{A} \vdash N : B$}
-  \RightLabel{$\text{let}_{{\bf K}}$}
-  \BinaryInfC{$\Gamma, z : C \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } N : {\bf K}B$}
-\end{prooftree}
-
-Let $\Gamma \vdash P : C$.
-
-By IH, $\Gamma \vdash \vec{M} [z := P] : {\bf K} \vec{A}$. So $\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} [z := P] {\: \bf in \: } N : {\bf K}B$
-
-  \vspace{\baselineskip}
-
-4-5) Similary.
-
-\end{proof}
-
-\begin{theorem} Subject reduction
-
-  If $\Gamma \vdash M : A$ and $M \twoheadrightarrow_{\beta \eta} N$, then $\Gamma \vdash N : A$
-
-\end{theorem}
-
-\begin{proof}
-
-Induction on the derivation $\Gamma \vdash M : A$ and on the generation of $\rightarrow_{\beta \eta}$.
-
-For cases with application, abstraction and pairs see \cite{Morten} \cite{Pierce}.
-
-1) If $\Gamma \vdash {\bf let \: pure \:} \vec{x}, y, \vec{z} = \vec{M}, {\bf let \: pure \: } \vec{w} = \vec{N} {\bf \: in \: } Q, \vec{P}  {\: \bf in \:} R : {\bf K}B$,
-  then $\Gamma \vdash {\bf let \: pure \:} \vec{x}, \vec{w}, \vec{z} = \vec{M}, \vec{N}, \vec{P} \: { \bf in } \: R [y := Q] : {\bf K}B$ by rule 4).
-
-2) Let $\Gamma \vdash {\bf let \: pure \:} x = M {\: \bf in \: } x : {\bf K}A$, then $\Gamma \vdash M : {\bf K}A$ by rule 9).
-
-See \cite{ModalK}.
-
-3) The derivation ends in
-
-\begin{prooftree}
-\AxiomC{$\Gamma \vdash {\bf pure \:} \vec{M} : {\bf K}\vec{A}$}
-\AxiomC{$\vec{x} : \vec{A} \vdash N : B$}
-\BinaryInfC{$\Gamma \vdash {\bf let \: pure \:} \vec{x} = {\bf pure \:} \vec{M} {\: \bf in \:} N : {\bf K}B$}
-\end{prooftree}
-
-So $\Gamma \vdash \vec{M} : \vec{A}$ by Lemma 4 and $\Gamma \vdash N [\vec{x} := \vec{M}] : B$ by Lemma 4, part 3.
-
-Then we can transform this into the following derivation:
-
-\begin{prooftree}
-\AxiomC{$\Gamma \vdash N [\vec{x} := \vec{M}] : B$}
-\RightLabel{${\bf K}_I$}
-\UnaryInfC{$\Gamma \vdash {\bf pure \:} N [\vec{x} := \vec{M}] : {\bf K}B$}
-\end{prooftree}
-
-4) The derivation ends in
-
-\begin{prooftree}
-\AxiomC{$\vdash M : A$}
-\UnaryInfC{$\vdash {\bf let \: pure \:} \underline{\quad} = \underline {\quad} {\: \bf in \:} M : {\bf K} A$}
-\end{prooftree}
-
-So, if $\vdash M : A$, then $\vdash {\bf pure \:} M : {\bf K}A$.
-
-Note that this part of the lemma works conversly too.
-\end{proof}
-
-\begin{theorem}
-  $ $
-
-  $\twoheadrightarrow_{\beta}$ is strongly normalizing;
-\end{theorem}
-
-\begin{proof}
-  $ $
-
-We modify and apply Tait's technique of logical relation for modalities. See \cite{Pierce} \cite{Troelstra}.
-
-We treat only modal cases below.
-
-\begin{defin} The set of strongly computable terms:
-  \begin{itemize}
-    \item $SC_A = \{ M : A \: | \: M \text{ is strongly normalizing} \}$ for $A \in \mathbb{T}$;
-    \item $SC_{A \to B} = \{ M : A \to B \: | \: \forall N \in SC_A, M N \in SC_B \}$, for $A,B \in \mathbb{T}_{\bf K}$ for $A, B \in \mathbb{T}_{{\bf K}}$;
-    \item $SC_{{\bf K}A} = \{ M : {\bf K}A \: | \: M \text{ is strongly normalizing} \}$ for $A \in \mathbb{T}$;
-    \item $\forall i \in \{ 1,\dots,n\}, \prod \limits_{i = 1}^{n} SC_{{\bf K}A_i} = \{ \vec{M} = (M_1,\dots,M_n) \:\: | \:\: \forall N \in SC_{B}, FV(N) = \{ x_1,\dots,x_n\} \: \& \: \forall i, x_i \in SC_{A_i} \Rightarrow {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N \in SC_{{\bf K} B}\}$
-  \end{itemize}
-\end{defin}
-
-\begin{defin}
-  A term $M$ is neutral, if it has one of the following forms:
-  \begin{itemize}
-    \item $M N$;
-    \item If $M$ is neutral, then ${\bf pure \:} M$ is neutral;
-    \item If $\vec{M}$ and $N$ are neutral, then ${\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: N}$ is neutral.
-    $\vec{x}$ is a sequence of free variables of a term $N$.
-  \end{itemize}
-\end{defin}
-
-\begin{lemma}
-  $ $
-
-\begin{itemize}
-\item If $M \in SC_A$ and $A \in \mathbb{T}_{{\bf K}}$, then $M$ is strongly normalizing;
-\item If $M \in SC_A$, $A \in \mathbb{T}_{{\bf K}}$ and $M \rightarrow_{\beta} N$, then $N \in SC_A$;
-\item If $N$ is neutral, $N \in SC_A$. Then, if $M \rightarrow_{\beta} N$, then $M \in SC_A$;
-\end{itemize}
-\end{lemma}
-
-\begin{proof}
-  $ $
-
-  By induction on the structure of $A$.
-
-1) $A \equiv {\bf K}A$, where $A \in \mathbb{T}$.
-
-i-ii-iii) Immediately.
-
-    \vspace{\baselineskip}
-
-2)
-
-i) Suppose $\vec{M} = (M_1,\dots,M_n) \in \prod \limits_{i = 1}^{n} SC_{{\bf K}A_i}$.
-
-Let $N \in SC_{B}$, such that $FV(N) = \{ x_1,\dots, x_n \}$ and $\forall i, x_i \in SC_{A_i}$.
-
-So ${\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } N \in SC_{{\bf K}B}$ by IH.
-
-So $\vec{M}$ are strongly normalizing, since ${\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } N$ is strongly normalizing by IH.
-
-\vspace{\baselineskip}
-
-ii) Let $\vec{M_1} \in \prod \limits_{i = 1}^{n} SC_{{\bf K}A_i}$ and $\vec{M_1} \rightarrow_{\beta} \vec{M_2}$.
-
-Let $N \in SC_{B}$, such that $FV(N) = \{ x_1,\dots,x_n \}$ and $\forall i, x_i \in SC_{A_i}$.
-
-So ${\bf let \: pure \:} \vec{x} = \vec{M_1} {\: \bf in \:} N \rightarrow_{\beta} {\bf let \: pure \:} \vec{x} = \vec{M_2} {\: \bf in \:} N$
-
-and ${\bf let \: pure \:} \vec{x} = \vec{M_2} {\: \bf in \:} N \in SC_{{\bf K}B}$ by assumption.
-
-So $\vec{M_2} \in \prod \limits_{i = 1}^{n} SC_{{\bf K}A_i}$.
-
-iii) Let $M_2$ be neutral, $M_2 \in \prod \limits_{i = 1}^{n} SC_{{\bf K}A_i}$ and $M_1 \rightarrow_{\beta} M_2$.
-
-Let $N \in SC_{B}$, , such that $FV(N) = \{ x_1,\dots,x_n \}$ and $\forall i, x_i \in SC_{A_i}$.
-
-So ${\bf let \: pure \:} \vec{x} = \vec{M_2} {\: \bf in \:} N \in SC_{{\bf K}B}$.
-
-Thus ${\bf let \: pure \:} \vec{x} = \vec{M_1} {\: \bf in \:} N \rightarrow_{\beta} {\bf let \: pure \:} \vec{x} = \vec{M_2} {\: \bf in \:} \in N$.
-
-Hence ${\bf let \: pure \:} \vec{x} = \vec{M_1} {\: \bf in \:} N\in SC_{{\bf K}B}$ by IH, so $\vec{M_1} \in \prod \limits_{i = 1}^{n} SC_{{\bf K}A_i}$.
-\end{proof}
-
-\begin{lemma}
-  $ $
-
-  If $M \in SC_A$, then ${\bf pure \:} M \in SC_{{\bf K}A}$
-\end{lemma}
-
-\begin{proof}
-
-  Induction on the structure of M.
-\end{proof}
-
-\begin{lemma}
-  $ $
-
-  Let $x_1 : A_1,\dots, x_n : A_n \vdash M : A$, then forall $i, M_i \in SC_{A_i}$. Then $M [x_1 := M_1,\dots,x_n := M_n] \in SC_A$.
-\end{lemma}
-
-\begin{proof}
-  $ $
-
-Induction on the derivation of $x_1 : A_1,\dots, x_n : A_n \vdash M : A$.
-
-  1) The derivation ends in:
-
-\begin{prooftree}
-\AxiomC{$x_1 : A_1,\dots, x_n : A_n \vdash M : A$}
-\UnaryInfC{$x_1 : A_1,\dots, x_n : A_n \vdash {\bf pure \:} M : {\bf K}A$}
-\end{prooftree}
-
-  By assumption $M [x_1 := M_1,\dots,x_n := M_n] \in SC_A$, so ${\bf pure \:} M [x_1 := M_1,\dots,x_n := M_n] \in SC_{{\bf K}A}$.
-
-\vspace{\baselineskip}
-
-  2) The derivation ends in:
-\begin{prooftree}
-  \AxiomC{$x_1 : A_1,\dots, x_n : A_n \vdash \vec{M^{'}} : {\bf K}\vec{A}$}
-  \AxiomC{$\vec{x} : \vec{A} \vdash N : B$}
-  \BinaryInfC{$x_1 : A_1,\dots, x_n : A_n \vdash {\bf let \: pure \:} \vec{x} = \vec{M^{'}} {\: \bf in \:} N : {\bf K}B$}
-\end{prooftree}
-
-By IH forall $i \in \{ 1,\dots,\text{length}(\vec{M^{'}}) \}$, $M_i^{'} [x_1 := M_1,\dots,x_n := M_n] \in SC_{{\bf K}A_i}$.
-
-So ${\bf let \: pure \:} \vec{x} = \vec{M^{'}} [x_1 := M_1,\dots,x_n := M_n] {\: \bf in \:} N \in SC_{{\bf K} B}$,
-otherwise we can build infinite reduction path in $\vec{M^{'}} [x_1 := M_1,\dots,x_n := M_n]$.
-\end{proof}
-
-\begin{col}
-
-  All terms are strongly computable, therefore are strongly normalizing.
-\end{col}
-
-\end{proof}
-
-\begin{theorem}
-  $ $
-
-  $\twoheadrightarrow_{\beta}$ is confluent.
-\end{theorem}
-
-\begin{proof}
-
-  We modify and apply Barendregt's technique with term underlying. We will consider the grammar without constructors for pairs.
-
-\begin{defin} The set of underlined terms.
-
-  \begin{itemize}
-    \item $x \in \mathbb{V} \Rightarrow x \in \underline{\Lambda}$;
-    \item $M \in \underline{\Lambda} \Rightarrow (\lambda x. M) \in \underline{\Lambda}$;
-    \item $M, N \in \underline{\Lambda} \Rightarrow (M N) \in \underline{\Lambda}$;
-    \item $M \in \underline{\Lambda} \Rightarrow ({\bf pure \:} M) \in \underline{\Lambda}$;
-    \item $\vec{x} \in \mathbb{V}, \vec{M}, N \in \underline{\Lambda} \Rightarrow {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N \in \underline{\Lambda}$;
-    \item $M, N \in \underline{\Lambda} \Rightarrow (\lambda_i x. M) N \in \underline{\Lambda}$, for all $i \in \mathbb{N}$.
-  \end{itemize}
-\end{defin}
-
-\begin{defin} Subsitution for term with labelled lambda:
-
-  $((\lambda_i x. M) N) [y := Z] = (\lambda_i x. M [y := Z]) (N [y := Z])$
-\end{defin}
-
-\begin{defin} Index erasing
-
-  Let us define map $|.| : \underline{\Lambda} \to \Lambda$ as follows:
-
-  \begin{itemize}
-    \item $|x| = x$;
-    \item $|\lambda x. M| = \lambda x. |M|$;
-    \item $|M N| = |M| |N|$;
-    \item $|{\bf pure \:} M| = {\bf pure \:} |M|$;
-    \item $|{\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N| = {\bf let \: pure \:} \vec{x} = \vec{|M|} {\: \bf in \:} |N|$;
-    \item $|(\lambda_i x. M) N| = (\lambda x. |M|) |N|$
-  \end{itemize}
-\end{defin}
-
-\begin{defin} Reduction rules:
-
-\begin{itemize}
-  \item $(\lambda x. M) N \rightarrow_{\underline{\beta}} M [x := N]$;
-  \item $\begin{array}{llll}
-  & {\bf let \: pure \:} \vec{x}, y, \vec{z} = \vec{M}, {\bf let \: pure \: } \vec{w} = \vec{N} {\bf \: in \: } Q, \vec{P} {\: in \:} R \rightarrow_{\underline{\beta}} & \\
-  & {\bf let \: pure \:} \vec{x}, \vec{w}, \vec{z} = \vec{M}, \vec{N}, \vec{P} \: { \bf in } \: R [y := Q]
-  \end{array}$;
-  \item ${\bf let \: pure \:} \vec{x} = {\bf pure \:} \vec{M} {\: \bf in \:} N \rightarrow_{\underline{\beta}} {\bf pure \:} N [\vec{x} := \vec{M}]$;
-  \item ${\bf let \: pure \:} \underline{\quad} = \underline {\quad} {\: \bf in \:} M \rightarrow_{\underline{\beta}} {\bf pure \:} M$
-  \item $(\lambda x_i. M) N \rightarrow_{\underline{\beta}} M [x := N]$
-\end{itemize}
-\end{defin}
-
-$\twoheadrightarrow_{\underline{\beta}}$ is a reflexive-transitive closure of $\rightarrow_{\underline{\beta}}$.
-
-\begin{defin} Indexed redex erasing:
-
-  Let us define the next map $\phi : \underline{\Lambda} \to \Lambda$:
-
-  \begin{itemize}
-    \item $\phi(x) = x$;
-    \item $\phi(\lambda x. M) = \lambda x. \phi(M)$;
-    \item $\phi(M N) = \phi(M) \phi(N)$;
-    \item $\phi({\bf pure \:} M) = {\bf pure \:} \phi(M)$;
-    \item $\phi({\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N) = {\bf let \: pure \:} \vec{x} = \vec{\phi(M)} {\: \bf in \:} \phi(N)$;
-    \item $\phi((\lambda_i x. M) N) = \phi(M) [x := \phi(N)]$
-  \end{itemize}
-\end{defin}
-
-\begin{ex}
-  $ $
-
-  $\phi((\lambda_i x. {\bf let \: pure \:} y = x {\: \bf in \:} x) N) = {\bf let \: pure \:} y = \phi(N) {\: \bf in \:} \phi(N)$
-\end{ex}
-
-\begin{lemma}
-
-  $\forall \underline{M}, \underline{N} \in \underline{\Lambda} \: \forall M, N \in \Lambda, \text{if } |\underline{M}| = M, |\underline{N}| = N, \text{then}$
-  \begin{itemize}
-    \item If $M \twoheadrightarrow_{\beta} N$, then $\underline{M} \twoheadrightarrow_{\underline{\beta}} \underline{N}$
-    \item Vice versa
-  \end{itemize}
-\end{lemma}
-
-\begin{proof}
-
-  Induction on the generation $\rightarrow_{\beta}$ and $\rightarrow_{\underline{\beta}}$ correspondently. The general statement follows from transitivity
-  of multi-step reductions of both types.
-\end{proof}
-
-\begin{lemma}
-  $\phi(M [x := N]) = \phi(M) [x := \phi(N)]$.
-\end{lemma}
-
-\begin{proof}
-  We treat only cases with {\bf pure} and with {\bf let}. For the rest cases see \cite{Baren}.
-
-  1)
-  $ $
-
-  $\begin{array}{lll}
-  & \phi({\bf pure \:} (M [x := N])) = & \text{By the definition of $\phi$}\\
-  & {\bf pure \:} (\phi(M [x := N])) = & \text{Induction hypothesis} \\
-  & {\bf pure \:} (\phi(M) [x := \phi(N)]) = & \text{Substitution definition}\\
-  & ({\bf pure \:} \phi (M)) [x := \phi(N)]&
-  \end{array}$
-
-  2)
-  $ $
-
-  $\begin{array}{lll}
-  &\phi(({\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N) [y := P]) = & \text{Substitution definition}\\
-  &\phi({\bf let \: pure \:} \vec{x} = (\vec{M}[y := P]) {\: \bf in \:} N) = & \text{By the definition of $\phi$}\\
-  &{\bf let \: pure \:} \vec{x} = \phi(\vec{M}[y := P]) {\: \bf in \:} \phi(N) = & \text{Induction hypothesis} \\
-  &{\bf let \: pure \:} \vec{x} = (\phi(\vec{M}) [y := \phi(P)]) {\: \bf in \:} \phi(N) =& \text{Substitution definition} \\
-  &({\bf let \: pure \:} \vec{x} = \phi(\vec{M}) {\: \bf in \:} \phi(N))[y := \phi(P)]
-  \end{array}$
-
-\end{proof}
-
-\begin{lemma}
-  $ $
-
-  \begin{itemize}
-  \item If $M \twoheadrightarrow_{\underline{\beta}} N$, then $\phi(M) \twoheadrightarrow_{\beta} \phi(N)$
-  \item If $|M| = N$ and $\phi(M) = P$, then $N \twoheadrightarrow_{\beta} P$.
-  \end{itemize}
-\end{lemma}
-
-\begin{proof}
-  $ $
-
-  i) Induction on the generation of $\twoheadrightarrow_{\underline{\beta}}$ using previous lemma.
-
-  ii) Induction on the structure of $M$.
-\end{proof}
-
-\begin{lemma} Strip lemma.
-
-$ $
-
-If $M \rightarrow_{\beta} N$ and $M \twoheadrightarrow_{\beta} P$. Then there exists some term $Q$, such that
-$N \twoheadrightarrow_{\beta} Q$ and $P \twoheadrightarrow_{\beta} Q$.
-\end{lemma}
-
-\begin{proof}
-
-Proof is similar to \cite{Baren} \cite{Baren2}. We build the following diagram, which commutes by lemmas 8 and 10.
-
-\vspace{\baselineskip}
-
-\xymatrix{
-&&& M \ar[dd]_{\beta} \ar@{->>}[rrr] &&& P \ar@{-->>}[dd]^{\beta} \\
-&&&& N' \ar[ul]^{|.|} \ar[dl]_{\phi} \ar@{-->>}[rrr]^{\underline{\beta}} &&& P' \ar[ul]^{|.|} \ar[dl]_{\phi} \\
-&&& N \ar@{-->>}[rrr]_{\beta} &&& Q
-}
-\end{proof}
-
-\begin{col}
-$ $
-If $M \twoheadrightarrow_{\beta} N$ and $M \twoheadrightarrow_{\beta} P$. Then there exists some term $Q$, such that
-$N \twoheadrightarrow_{\beta} Q$ and $P \twoheadrightarrow_{\beta} Q$.
-\end{col}
-
-\begin{proof}
-
-  Unfold $M \twoheadrightarrow_{\beta} N$ as the sequence of one-step reductions and apply strip lemma on the every step.
-\end{proof}
-
-\end{proof}
-
-\begin{theorem}
-  $ $
-
-  Normal form in call-by-name $\lambda_{{\bf K}}$ has the subformula property: if $M$ is in normal formal, then its all subterms are in normal form too.
-
-\end{theorem}
-
-\begin{proof}
-  By induction on the structure of $M$.
-  Case with ${\bf let \: pure\:} \vec{x} = \vec{M} {\: \bf in \:} N$ was considered by Kakutani \cite{ModalK} \cite{ModalK1}.
-
-  If ${\bf pure \:} M$ is in normal form, so $M$ is in normal form and its subterms are in normal form too by hypothesis.
-
-  Thus if ${\bf pure \:} M$ is in normal form, then all its subterms are in normal form too.
-\end{proof}
-
-\section{Categorical semantics}
-
-\begin{defin} Lax monoidal functor
-
-  Let $\langle \mathcal{C}, \otimes_1, \mathds{1} \rangle$ and $\langle \mathcal{D}, \otimes_2, \mathds{1}' \rangle$ are monoidal categories.
-
-  A lax monoidal functor $\mathcal{F} : \langle \mathcal{C}, \otimes_1, \mathds{1} \rangle \to \langle \mathcal{D}, \otimes_2, \mathds{1}' \rangle$ is a functor
-  $\mathcal{F} : \mathcal{C} \to \mathcal{D}$ with additional natural transformations, which satisfy the well-known conditions described in \cite{ElKelly}:
-
-  1) $u : \mathds{1}' \to \mathcal{F}\mathds{1}$;
-
-  2) $\ast_{A, B} : \mathcal{F}A \otimes_2 \mathcal{F}B \to \mathcal{F}(A \otimes_1 B)$.
-
-\end{defin}
-
-\begin{defin}
-
-Strong endofunctor on a monoidal category is a with a natural transformation so called tensorial strength
-
-\begin{equation}
-  \tau_{A, B} : A \otimes \mathcal{K}B \rightarrow \mathcal{K}(A \otimes B)
-\end{equation}
-
-making two diagrams commute:
-
-\xymatrix{
-(A \otimes B) \otimes \mathcal{K}C \ar[d]_{\alpha_{A, B, \mathcal{K}C}}\ar[rrrr]^{\tau_{A \otimes B, C}} &&&& \mathcal{K}((A \otimes B) \otimes C) \ar[d]^{\mathcal{K}(\alpha_{A,B,C})}\\
-A \otimes (B \otimes \mathcal{K}C) \ar[rr]_{id_A \otimes \tau_{B,C}} && A \otimes \mathcal{K}(B \otimes C) \ar[rr]_{\quad \tau_{A, (B \otimes C)}} && \mathcal{K}(A \otimes (B \otimes C)) \\
-}
-
-\xymatrix{
-&&& \mathds{1} \otimes \mathcal{K}A \ar[drr]_{\rho_{\mathcal{K}A}} \ar[rr]^{\mu_{\mathds{1}, A}} && \mathcal{K}(\mathds{1} \otimes A) \ar[d]^{\mathcal{K}(\rho_A)}\\
-&&&&& \mathcal{K}A
-}
-\end{defin}
-
-\begin{defin} Applicative functor
-
-  An applicative functor is a triple $\langle \mathcal{C}, \mathcal{K}, \eta \rangle$,
-where $\mathcal{C}$ is a monoidal category, $\mathcal{K}$ is a strong lax monoidal endofunctor and $\eta : Id_{\mathcal{C}} \Rightarrow \mathcal{K}$ is a natural transformation (similar to unit in monad), such that:
-
-1) $u = \eta_{\mathds{1}}$;
-
-2) $\ast_{A,B} \circ (\eta_A \otimes \eta_B) = \eta_{A \otimes B}$, i.e. the following diagram commutes:
-
-\xymatrix
-{
-&&& A \otimes B \ar[rr]^{\eta_A \otimes \eta_B} \ar[drr]_{\eta_{A \otimes B}} && \mathcal{K}A \otimes \mathcal{K}B \ar[d]^{\ast_{A,B}} \\
-&&&&& \mathcal{K}(A \otimes B)
-}
-
-3) $\tau_{A, B} = \ast_{A, B} \circ \eta_{A} \otimes id_{\mathcal{K}B}$.
-\end{defin}
-
-By default we will consider some strong lax monoidal functor on cartesian closed category.
-
-Note that lax monoidal functor on a cartesian closed category is a special case of an applicative functor \cite{Cons}, but
-we will define $\eta$ explicitly for simplicity.
-
-\subsection{Soundness and completeness}
-
-\begin{theorem} Soundness
-
-  Let $\Gamma \vdash M : A$ and $M =_{\beta\eta} N$, then $[\![\Gamma \vdash M : A]\!] = [\![\Gamma \vdash N : A]\!]$
-\end{theorem}
-
-\begin{proof}
-
-\begin{defin} Semantical translation from $\lambda_{{\bf K}}$ to some cartesian closed category $\mathcal{C}$ with applicative functor $\mathcal{K}$:
-
-1) Interpretation for types:
-
-$[\![A]\!] := \hat{A}, A \in \mathbb{T}$, where $\hat{A}$ is an object of $\mathcal{C}$ obtained by some given assignment;
-
-$[\![A \to B]\!] := [\![B]\!]^{[\![A]\!]}$;
-
-$[\![A \times B]\!] := [\![A]\!] \times [\![B]\!]$.
-
-2) Interpretation for modal types: $[\![{\bf K}A]\!] = \mathcal{K}[\![A]\!]$;
-
-3) Interpretaion for contexts:
-
-$[\![\Gamma = \{ x_1 : A_1, ..., x_n : A_n\}]\!] := [\![\Gamma]\!] = [\![A_1]\!] \times ... \times [\![A_n]\!]$;
-
-4) Interpretation for typing assignment: $[\![\Gamma \vdash M : A]\!] := [\![M]\!] : [\![\Gamma]\!] \to
-[\![A]\!]$.
-
-5) Interpretation for typing rules:
-
-\begin{prooftree}
-\AxiomC{$ $}
-\UnaryInfC{$[\![\Gamma, x : A \vdash x : A]\!] = \pi_2 : [\![\Gamma]\!] \times [\![A]\!] \rightarrow
-[\![A]\!]$}
-\end{prooftree}
-
-\begin{prooftree}
-\AxiomC{$[\![\Gamma, x : A \vdash M : B]\!] = f : [\![\Gamma]\!] \times [\![A]\!] \rightarrow [\![B]\!]$}
-\UnaryInfC{$[\![\Gamma \vdash (\lambda x. M) : A \to B]\!] = \Lambda(f) : [\![\Gamma]\!]
-\rightarrow[\![B]\!]^{[\![A]\!]}$}
-\end{prooftree}
-
-\begin{prooftree}
-\AxiomC{$[\![\Gamma \vdash M : A \to B]\!] = [\![M]\!] : [\![\Gamma]\!] \rightarrow [\![B]\!]^{[\![A]\!]}$}
-\AxiomC{$[\![\Gamma \vdash N : A]\!] = [\![N]\!] : [\![\Gamma]\!] \rightarrow [\![A]\!]$}
-\BinaryInfC{$[\![\Gamma \vdash (M N) : B]\!] = [\![\Gamma]\!] \xrightarrow{\langle [\![M]\!], [\![N]\!]
-\rangle} [\![B]\!]^{[\![A]\!]} \times [\![A]\!] \xrightarrow{\epsilon} [\![B]\!] $}
-\end{prooftree}
-
-\begin{prooftree}
-\AxiomC{$[\![\Gamma \vdash M : A ]\!] = f : [\![\Gamma]\!] \rightarrow [\![A]\!]$}
-\AxiomC{$[\![\Gamma \vdash N : B ]\!] = g : [\![\Gamma]\!] \rightarrow [\![B]\!]$}
-\BinaryInfC{$[\![\Gamma \vdash (M, N) : A \times B]\!] = \langle f, g \rangle : [\![\Gamma]\!] \rightarrow
-[\![A]\!] \times [\![B]\!]$}
-\end{prooftree}
-
-\begin{prooftree}
-\AxiomC{$[\![\Gamma \vdash p : A_1 \times A_2]\!] = f : [\![\Gamma]\!] \rightarrow [\![A_1]\!] \times
-[\![A_2]\!]$}
-\RightLabel{$i \in \{1,2\}$}
-\UnaryInfC{$[\![\Gamma \vdash \pi_i p : A_i]\!] = [\![\Gamma]\!] \xrightarrow{f} [\![A_1]\!] \times
-[\![A_2]\!] \xrightarrow{\pi_i} [\![A_i]\!]$}
-\end{prooftree}
-
-\begin{prooftree}
-\AxiomC{$[\![\Gamma \vdash M : A]\!] = [\![M]\!] : [\![\Gamma]\!] \rightarrow [\![A]\!]$}
-\UnaryInfC{$[\![\Gamma \vdash {\bf pure \:} M : {\bf K}A]\!] := [\![\Gamma]\!] \xrightarrow{[\![M]\!]}
-[\![A]\!] \xrightarrow{\eta_{[\![A]\!]}} \mathcal{K}[\![A]\!]$}
-\end{prooftree}
-
-  \begin{prooftree}
-    \AxiomC{$[\![\Gamma \vdash \vec{M} : {\bf K} \vec{A}]\!] = \langle [\![M_1]\!],\dots, [\![M_n]\!] \rangle : [\![\Gamma]\!] \rightarrow \prod \limits_{i=1}^n \mathcal{K}[\![A_i]\!]$}
-    \AxiomC{$[\![\vec{x} : \vec{A} \vdash N : B]\!] = [\![N]\!] : \prod \limits_{i=1}^n [\![A_i]\!] \rightarrow [\![B]\!]$}
-    \BinaryInfC{$[\![\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } M : {\bf K} B]\!] = \mathcal{K}([\![N]\!]) \circ \ast_{[\![A_1]\!],\dots,[\![A_n]\!]} \circ \langle [\![M_1]\!],\dots, [\![M_n]\!] \rangle : [\![\Gamma]\!] \rightarrow \mathcal{K}[\![B]\!]$}
-  \end{prooftree}
-
-\end{defin}
-
-\begin{defin} Simultaneous substitution
-
-Let $\Gamma = \{ x_1 : A_1, ..., x_n : A_n \}$, $\Gamma \vdash M : A$ and for all $i \in \{ 1,..., n \}$,
-$\Gamma \vdash M_i : A_i$.
-
-We define simultaneous substitution $M [ \vec{x} := \vec{M}]$ recursively by:
-
-1) $x_i [ \vec{x} := \vec{M}] = M_i $;
-
-2) $(\lambda x. M) [ \vec{x} := \vec{M}] = \lambda x. (M [ \vec{x} := \vec{M}])$;
-
-3) $(M N) [ \vec{x} := \vec{M}] = (M [ \vec{x} = \vec{M}]) (N [ \vec{x} := \vec{M}])$;
-
-4) $\langle M, N \rangle = \langle (M [ \vec{x} = \vec{M}]), (N [ \vec{x} := \vec{M}])\rangle$;
-
-5) $(\pi_i P) [ \vec{x} := \vec{M}] = \pi_i (P [ \vec{x} = \vec{M}])$;
-
-6) $({\bf pure \:} M) [ \vec{x} := \vec{M}] = {\bf pure \:} (M [ \vec{x} = \vec{M}])$;
-
-7) $({\bf let \: pure} \: \vec{x} = \vec{M} {\: \bf in \:} N) [\vec{y} := \vec{P}] =
-{\bf let \: pure} \: \vec{x} = (\vec{M} [\vec{y} := \vec{P}]) {\: \bf in \:} N$
-
-\end{defin}
-
-\begin{lemma}
-$ $
-
-$[\![M [x_1 := M_1,\dots, x_n := M_n]]\!] = [\![M]\!] \circ \langle [\![M_1]\!], \dots, [\![M_n]\!] \rangle$.
-
-\end{lemma}
-
-\begin{proof}
-
-$ $
-
-1) $[\![\Gamma \vdash ({\bf pure \:} M) [ \vec{x} := \vec{M}] : {\bf K}A ]\!] = [\![ \Gamma \vdash {\bf pure \:} M : {\bf K}A]\!] \circ \langle [\![M_1]\!], \dots,
-[\![M_n]\!] \rangle$.
-
-\vspace{\baselineskip}
-
-$\begin{array}{lll}
-& [\![\Gamma \vdash ({\bf pure \:} M) [ \vec{x} := \vec{M}] : {\bf K}A]\!] = [\![ \Gamma \vdash {\bf pure \:} (M [ \vec{x} := \vec{M}]) : {\bf K}A]\!] &
-\text{Substitution definition} \\
-& \quad\quad\quad\quad\quad\quad\quad = \eta_{[\![A]\!]} \circ [\![(M [ \vec{x} := \vec{M}])]\!]&
-\text{Translation for pure}\\
-& \quad\quad\quad\quad\quad\quad\quad = \eta_{[\![A]\!]} \circ ([\![M]\!] \circ \langle [\![M_1]\!], \dots,
-[\![M_n]\!] \rangle) & \text{Induction hypothesis} \\
-& \quad\quad\quad\quad\quad\quad\quad = (\eta_{[\![A]\!]} \circ [\![M]\!]) \circ \langle [\![M_1]\!], \dots,
-[\![M_n]\!] \rangle & \text{Associativity of composition}\\
-& \quad\quad\quad\quad\quad\quad\quad = [\![ \Gamma \vdash {\bf pure \:} M : {\bf K}A]\!] \circ \langle [\![M_1]\!], \dots,
-[\![M_n]\!] \rangle& \text{Translation for pure}\\
-\end{array}$
-
-\vspace{\baselineskip}
-
-2) $\begin{array}{lll}&[\![\Gamma \vdash ({\bf let \: pure} \: \vec{x} = \vec{M} {\: \bf in \:} N) [\vec{y} := \vec{P}] : {\bf K} B]\!]
-=[\![\Gamma \vdash {\bf let \: pure} \: \vec{x} = \vec{M} {\: \bf in \:} N : {\bf K}B]\!] \circ \langle [\![P_1]\!],\dots,[\![P_n]\!]\rangle& \end{array}$
-
-\vspace{\baselineskip}
-
-$\begin{array}{lll}
-&[\![\Gamma \vdash ({\bf let \: pure} \: \vec{x} = \vec{M} {\: \bf in \:} N) [\vec{y} := \vec{P}] : {\bf K} B]\!] =& \text{Substitution definition}\\
-&[\![\Gamma \vdash {\bf let \: pure} \: \vec{x} = (\vec{M} [\vec{y} := \vec{P}]) {\: \bf in \:} N : {\bf K}B]\!] =& \text{Interpretaion for $let_{\bf K}$} \\
-&\mathcal{K}([\![N]\!]) \circ \ast_{[\![A_1]\!],\dots,[\![A_n]\!]} \circ [\![\Gamma \vdash (\vec{M} [\vec{y} := \vec{P}]) : {\bf K}\vec{A}]\!] =& \text{Induction hypothesis} \\
-&\mathcal{K}([\![N]\!]) \circ \ast_{[\![A_1]\!],\dots,[\![A_n]\!]} \circ ([\![\vec{M}]\!] \circ \langle [\![P_1]\!],\dots,[\![P_n]\!]\rangle) = & \text{Associativity of composition} \\
-&(\mathcal{K}([\![N]\!]) \circ \ast_{[\![A_1]\!],\dots,[\![A_n]\!]} \circ [\![\vec{M}]\!]) \circ \langle [\![P_1]\!],\dots,[\![P_n]\!]\rangle = & \text{By interpretation} \\
-&[\![\Gamma \vdash {\bf let \: pure} \: \vec{x} = \vec{M} {\: \bf in \:} N : {\bf K}B]\!] \circ \langle [\![P_1]\!],\dots,[\![P_n]\!]\rangle&
-\end{array}$
-
-\end{proof}
-
-\begin{lemma}
-  $ $
-
-  Let $\Gamma \vdash M : A$ and $M \twoheadrightarrow_{\beta \eta} N$, then $[\![\Gamma \vdash M : A]\!] = [\![\Gamma \vdash N : A]\!]$;
-\end{lemma}
-
-\begin{proof}
-  $ $
-
-Cases with $\beta$-reductions for $let_{\bf K}$ are shown in \cite{ModalK1}. Let us consider cases with ${\bf pure}$.
-
-\vspace{\baselineskip}
-
-1) $[\![\Gamma \vdash {\bf let \: pure \:} \vec{x} = {\bf pure \:} \vec{M} {\: \bf in \:} N : {\bf K}B]\!] = [\![\Gamma \vdash {\bf pure \:} N [\vec{x} := \vec{M}] : {\bf K}B]\!]$
-
-\vspace{\baselineskip}
-
-$\begin{array}{lll}
-&[\![\Gamma \vdash {\bf let \: pure \:} \vec{x} = {\bf pure \:} \vec{M} {\: \bf in \:} N : {\bf K}B]\!] = & \\
-&\text{\quad\quad\quad\quad\quad\quad\quad\quad\quad By interpretation}& \\
-&\mathcal{K}([\![N]\!]) \circ \ast_{[\![A_1]\!],\dots,[\![A_n]\!]} \circ \langle \eta_{[\![A_1]\!]} \circ [\![M_1]\!],\dots,\eta_{[\![A_n]\!]} \circ [\![M_n]\!] \rangle = \\
-&\text{\quad\quad\quad\quad\quad\quad\quad\quad\quad By the property of a pair of morphisms}& \\
-&\mathcal{K}([\![N]\!]) \circ \ast_{[\![A_1]\!],\dots,[\![A_n]\!]} \circ (\eta_{[\![A_1]\!]} \times \dots \times \eta_{[\![A_n]\!]}) \circ \langle [\![M_1]\!], \dots, [\![M_n]\!]\rangle =& \\
-&\text{\quad\quad\quad\quad\quad\quad\quad\quad\quad Associativity of composition}& \\
-&\mathcal{K}([\![N]\!]) \circ (\ast_{[\![A_1]\!],\dots,[\![A_n]\!]} \circ (\eta_{[\![A_1]\!]} \times \dots \eta_{[\![A_n]\!]})) \circ \langle [\![M_1]\!], \dots, [\![M_n]\!] \rangle =& \\
-&\text{\quad\quad\quad\quad\quad\quad\quad\quad\quad By the definition of an applicative functor}& \\
-&\mathcal{K}([\![N]\!]) \circ \eta_{[\![A_1]\!] \times \dots \times [\![A_n]\!]} \circ \langle [\![M_1]\!], \dots, [\![M_n]\!] \rangle =& \\
-&\text{\quad\quad\quad\quad\quad\quad\quad\quad\quad Naturality of $\eta$}& \\
-&\eta_{[\![B]\!]} \circ [\![N]\!] \circ \langle [\![M_1]\!], \dots, [\![M_n]\!] \rangle =& \\
-&\text{\quad\quad\quad\quad\quad\quad\quad\quad\quad Associativity of composition}& \\
-&\eta_{[\![B]\!]} \circ ([\![N]\!] \circ \langle [\![M_1]\!], \dots, [\![M_n]\!]) \rangle =& \\
-&\text{\quad\quad\quad\quad\quad\quad\quad\quad\quad Simultaneous substitution lemma}& \\
-&\eta_{[\![B]\!]} \circ [\![N [\vec{x} := \vec{M}]]\!]& \\
-&\text{\quad\quad\quad\quad\quad\quad\quad\quad\quad By interpetation}& \\
-&[\![\Gamma \vdash {\bf pure \:} (N [\vec{x} := \vec{M}]) : {\bf K}B]\!]&
-\end{array}$
-
-\vspace{\baselineskip}
-
-2) $[\![\vdash {\bf let \: pure \:} \underline{\quad} = \underline {\quad} {\: \bf in \:} M : {\bf K} A]\!] = [\![\vdash {\bf pure \:} M : {\bf K} A]\!]$
-
-$\begin{array}{lll}
-&[\![\vdash {\bf let \: pure \:} \underline{\quad} = \underline {\quad} {\: \bf in \:} M : {\bf K} A]\!] = & \text{By interpretation} \\
-&\mathcal{K}([\![M]\!]) \circ u_{\mathds{1}} = & \text{By the definition of an applicative functor} \\
-&\mathcal{K}([\![M]\!]) \circ \eta_{\mathds{1}} = & \text{By naturality of $\eta$} \\
-&\eta_{[\![A]\!]} \circ [\![M]\!] = & \text{By interpretation} \\
-&[\![\vdash {\bf pure \:} M : {\bf K} A]\!]&
-\end{array}$
-\end{proof}
-
-\end{proof}
-
-\begin{theorem} Completeness
-
-Let $[\![\Gamma \vdash M : A]\!] = [\![\Gamma \vdash N : A]\!]$, then $M =_{\beta \eta} N$.
-\end{theorem}
-
-\begin{proof}
-
-$ $
-
-We will consider term model for simply typed lambda calculus $\times$ and $\to$ standardly described in \cite{LambekScott}:
-
-\begin{defin} Equivalence on term pairs:
-  $ $
-
-  Let us define relation $\sim_{A, B} \subseteq \mathbb{V} \times \Lambda_{{\bf K}}$, such that:
-
-  $(x, M) \sim_{A, B} (y, N) \Leftrightarrow x : A \vdash M : B \:\: \& \:\: y : A \vdash N : A \:\: \& \:\: M =_{\beta \eta} N [y := x]$;
-\end{defin}
-
-We will denote equivalence class as $[x, M]_{A, B} = \{ (y, N) \: | \: (x, M) \sim_{A, B} (y, N) \}$ (we will drop indices below).
-
-\begin{defin} Category $\mathcal{C}(\lambda)$:
-\begin{itemize}
-  \item $Ob_{\mathcal{C}} = \{ \hat{A} \: | \: A \in \mathbb{T} \} \cup \{ \mathds{1} \}$;
-  \item $Hom_{\mathcal{C}(\lambda)}(\hat{A},\hat{B}) = (\mathbb{V} \times \Lambda_{{\bf K}})/_{\sim_{A, B}}$;
-  \item Let $[x, M] \in Hom_{\mathcal{C}(\lambda)}(\hat{A},\hat{B})$ and $[y,N] \in Hom_{\mathcal{C}(\lambda)}(\hat{B},\hat{C})$, then $[y,M] \circ [x, M] = [x, N [y := M]]$;
-  \item Identity morphism $id_{\hat{A}} = [x,x] \in Hom_{\mathcal{C}(\lambda)(\hat{A})}$;
-  \item $\mathds{1}$ is a terminal object;
-  \item $\widehat{A \times B} = \hat{A} \times \hat{B}$;
-  \item Canonical projection is defined as $[x, \pi_i x] \in Hom_{\mathcal{C}(\lambda)}(\hat{A_1} \times \hat{A_2},\hat{A_i})$ for $i \in \{ 1, 2 \}$;
-  \item $\widehat{A \to B} = \hat{B}^{\hat{A}}$;
-  \item Evaluation arrow $\epsilon = [x, (\pi_2 x) (\pi_1 x)] \in Hom_{\mathcal{C}(\lambda)(\hat{B}^{\hat{A}} \times \hat{A}, \hat{B})}$.
-\end{itemize}
-\end{defin}
-
-It is sufficient to show ${\bf K}$ is an applicative functor on $\mathcal{C}(\lambda)$.
-
-\begin{defin}
-  Let us define an endofunctor $\mathcal{K} : \mathcal{C}(\lambda) \to \mathcal{C}(\lambda)$, such that
-forall $[x,M] \in Hom_{\mathcal{C}(\lambda)}(\hat{A},\hat{B}), {\bf K}([x,M]) = [y, {\bf let \: pure \:} x = y {\: \bf in \:} M] \in Hom_{\mathcal{C}(\lambda)}({\bf K} \hat{A}, {\bf K} \hat{B})$ (denotation: $\text{fmap } f$ for an arbitrary arrow $f$).
-
-\end{defin}
-
-\begin{lemma} Functoriality
-
-\begin{itemize}
-  \item $\text{fmap }(g \circ f) = \text{fmap }(g) \circ \text{fmap }(f)$;
-  \item $\text{fmap }(id_{\hat{A}}) = id_{{\bf K}\hat{A}}$.
-\end{itemize}
-\end{lemma}
-
-\begin{proof}
-
-Easy checking using reduction rules.
-
-\end{proof}
-
-\begin{defin}
-
-  Let us define natural transformations:
-
-  1) $\eta:Id \Rightarrow \mathcal{K}$, s. t. $\forall \hat{A} \in Ob_{\mathcal{C}(\lambda)}$, $\eta_{\hat{A}} = [x, {\bf pure \:} x] \in Hom_{\mathcal{C}(\lambda)}(\hat{A}, {\bf K}\hat{A})$;
-
-  2) $\ast_{A,B}:{\bf K}\hat{A} \times {\bf K}\hat{B} \to {\bf K}(\hat{A} \times \hat{B})$, s.
-t.
-$\forall \hat{A}, \hat{B} \in Ob_{\mathcal{C}(\lambda)}, \ast_{\hat{A},\hat{B}} = [p, {\bf let \: pure \:} x,y = \pi_1 p, \pi_2 p {\: \bf in \:} \langle x, y \rangle] \in Hom_{\mathcal{C}(\lambda)}({\bf K}A \times {\bf K}B, {\bf K}(A \times B))$.
-\end{defin}
-
-Implementation for $\ast$ in our term model is a modification of ${\text{let}_{\bf K}}$-rule:
-
-\begin{prooftree}
-\AxiomC{$p : {\bf K}A \times {\bf K}B \vdash p : {\bf K}A \times {\bf K}B$}
-\UnaryInfC{$p : {\bf K}A \times {\bf K}B \vdash \pi_1 p : {\bf K} A$}
-\AxiomC{$p : {\bf K}A \times {\bf K}B \vdash p : {\bf K}A \times {\bf K}B$}
-\UnaryInfC{$p : {\bf K}A \times {\bf K}B \vdash \pi_2 p : {\bf K} B$}
-\AxiomC{$x : A \vdash x : A$}
-\AxiomC{$y : B \vdash y : B$}
-\BinaryInfC{$x : A, y : B \vdash \langle x, y \rangle : A \times B$}
-\TrinaryInfC{$p : {\bf K}A \times {\bf K}B \vdash {\bf let \: pure \:} x,y = \pi_1 p, \pi_2 p {\: \bf in \:} \langle x, y \rangle : {\bf K}(A \times B)$}
-\end{prooftree}
-
-\begin{lemma}
-  $ $
-
-  {\bf K} is a lax monoidal endofunctor
-\end{lemma}
-
-\begin{proof}
-$ $
-
-See \cite{ModalK}
-\end{proof}
-
-\begin{lemma} Naturality and coherence for $\eta$:
-
-\begin{itemize}
-  \item $\text{fmap } f \circ \eta_A = \eta_B \circ f$;
-  \item $\ast_{\hat{A},\hat{B}} \circ (\eta_{A} \times \eta_{B}) = \eta_{\hat{A} \times \hat{B}}$;
-\end{itemize}
-\end{lemma}
-
-\begin{proof}
-  $ $
-
-  i) $\text{fmap } f \circ \eta_{\hat{A}} = \eta_{\hat{B}} \circ f$
-
-\vspace{\baselineskip}
-
-$\begin{array}{lll}
-&\eta_{\hat{B}} \circ f = & \text{By the definition} \\
-&[y, {\bf pure \:} y] \circ [x, M] = & \text{By the definition of composition}\\
-&[x, {\bf pure \:} y [y := M]] = & \text{By substitution}\\
-&[x, {\bf pure \:} M]& \\
-& &\\
-&\text{On the other hand:}& \\
-&\text{fmap } f \circ \eta_{\hat{A}} = & \text{By the definiton} \\
-&[z, {\bf let \: pure \:} x = z {\: \bf in \:} M] \circ [x, {\bf pure \: x}] = & \text{By the definition of composition}\\
-&[x, {\bf let \: pure \:} x = z {\: \bf in \:} M [z := {\bf pure \:} x]] = & \text{By substitution} \\
-&[x, {\bf let \: pure \:} x = {\bf pure \:} x {\: \bf in \:} M] = & \text{$\beta$-reduction rule}\\
-&[x, {\bf pure \:} M [x := x]] = & \text{By substitution}\\
-&[x, {\bf pure \:} M]&
-\end{array}$
-
-\vspace{\baselineskip}
-
-ii) $\ast_{\hat{A},\hat{B}} \circ (\eta_{\hat{A}} \times \eta_{\hat{B}}) = \eta_{\hat{A} \times \hat{B}}$
-
-\vspace{\baselineskip}
-
-$\begin{array}{lll}
-& \ast_{\hat{A},\hat{B}} \circ (\eta_{\hat{A}} \times \eta_{\hat{B}}) = & \\
-& \quad\quad\quad\quad\quad\quad \text{By unfolding} & \\
-& [q, {\bf let \: pure \:} x, y = \pi_1 q, \pi_2 q {\: \bf in \:} \langle x, y \rangle] \circ [p, \langle {\bf pure \:} (\pi_1 p), {\bf pure \:} (\pi_2 p) \rangle] = & \\
-& \quad\quad\quad\quad\quad\quad \text{Composition}& \\
-& [p, {\bf let \: pure \:} x, y = \pi_1 q, \pi_2 q {\: \bf in \:} \langle x, y \rangle [q := \langle {\bf pure \:} (\pi_1 p), {\bf pure \:} (\pi_2 p) \rangle]] = & \\
-& \quad\quad\quad\quad\quad\quad \text{By substitution}& \\
-& [p, {\bf let \: pure \:} x, y = \pi_1 (\langle {\bf pure \:} (\pi_1 p), {\bf pure \:} (\pi_2 p) \rangle), \pi_2 (\langle {\bf pure \:} (\pi_1 p), {\bf pure \:} (\pi_2 p) \rangle) {\: \bf in \:} \langle x, y \rangle] = & \\
-& \quad\quad\quad\quad\quad\quad \text{Reduction rules}& \\
-& [p, {\bf let \: pure \:} x, y = {\bf pure \:} (\pi_1 p), {\bf pure \:} (\pi_2 p) {\: \bf in \:} \langle x, y \rangle] = & \\
-& \quad\quad\quad\quad\quad\quad \text{Reduction rule}& \\
-& [p, {\bf pure \:} (\langle x,y \rangle [x := \pi_1 p, y := \pi_2 p])] = & \\
-& \quad\quad\quad\quad\quad\quad \text{Substitution}& \\
-& [p, {\bf pure \:} \langle \pi_1 p, \pi_2 p \rangle] = & \\
-& \quad\quad\quad\quad\quad\quad \text{$\eta$-reduction}& \\
-& [p, {\bf pure \:} p] =& \\
-& \quad\quad\quad\quad\quad\quad \text{By definition}& \\
-& \eta_{\hat{A} \times \hat{B}}&
-\end{array}$
-
-\end{proof}
-
-\begin{defin}
-  $ $
-
-  $u_{\mathds{1}} = [\sqbullet, {\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} \sqbullet] \in Hom_{\mathcal{C}(\lambda)}(\mathds{1}, {\bf K}\mathds{1})$.
-\end{defin}
-
-\begin{lemma}
-  $ $
-
-  $u_{\mathds{1}} = \eta_{\mathds{1}}$
-
-\end{lemma}
-
-\begin{proof}
-
-  Immediately.
-\end{proof}
-
-  Tensorial strength is defined as follows:
-
-\begin{defin} Tensorial strength
-  $ $
-
-  Let $[p, \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle] \in Hom_{\mathcal{C}(\lambda)}(\hat{A} \times {\bf K}\hat{B}, {\bf K}\hat{A} \times {\bf K}\hat{B})$.
-
-  So tensorial strength is defined as $\tau_{\hat{A}, \hat{B}} = \ast_{\hat{A}, \hat{B}} \circ [p, \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle]$.
-\end{defin}
-
-It is clearly that tensorial strength defined above can be simplified as follows:
-
-$\begin{array}{lll}
-&\ast_{\hat{A}, \hat{B}} \circ [p, \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle] = & \text{By definition} \\
-&[p^{'}, {\bf let \: pure \:} x, y = \pi_1 p^{'}, \pi_2 p^{'} {\: \bf in \:} \langle x, y \rangle] \circ [p, \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle] = & \text{By composition} \\
-&[p, {\bf let \: pure \:} x, y = \pi_1 p^{'}, \pi_2 p^{'} {\: \bf in \:} \langle x, y \rangle [p^{'} := \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle]] = & \text{By substitution} \\
-&[p, {\bf let \: pure \:} x, y = \pi_1 (\langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle), \pi_2 (\langle \pi_1 p, {\bf pure \:} (\pi_2 p) \rangle) {\: \bf in \:} \langle x, y \rangle] = & \text{By $\beta$-reduction rules} \\
-&[p, {\bf let \: pure \:} x, y = {\bf pure \:} (\pi_1 p), \pi_2 p {\: \bf in \:} \langle x, y \rangle]&
-\end{array}$
-
-\begin{lemma} Coherence for tensorial strength
-
-  \begin{itemize}
-    \item $\text{fmap } \alpha_{\hat{A},\hat{B},\hat{C}} \circ \tau_{\hat{A} \times \hat{B}, \hat{C}} = \tau_{\hat{A}, \hat{B} \times \hat{C}} \circ (id_{\hat{A}} \times \tau_{\hat{B}, \hat{C}}) \circ \alpha_{\hat{A}, \hat{B}, {\bf K}\hat{C}}$;
-    \item ${\bf K}(\rho_{\hat{A}}) \circ \tau_{\mathds{1}, \hat{A}} = \rho_{{\bf K}\hat{A}}$.
-  \end{itemize}
-  where $\alpha_{\hat{A},\hat{B},\hat{C}} = [p, \langle \pi_1 (\pi_1 p), \langle \pi_2 (\pi_1 p), \pi_2 p \rangle \rangle]$ and
-  $\rho = \pi_2$.
-\end{lemma}
-
-\begin{proof}
-  $ $
-
-1) Let us define $\tau_{\hat{A} \times \hat{B}, \hat{C}}$ as follows:
-
-$\tau_{\hat{A} \times \hat{B}, \hat{C}} = [p, {\bf \: let \: pure \:} x,y,z = {\bf pure\:} (\pi_1 (\pi_1 p)), {\bf pure \:} (\pi_2 (\pi_1 p)), \pi_2 p {\: \bf in \:} \langle \langle x,y \rangle, z \rangle]$
-
-\vspace{\baselineskip}
-
-Then:
-
-\vspace{\baselineskip}
-
-$\begin{array}{lll}
-&\text{fmap } \alpha_{\hat{A},\hat{B},\hat{C}} \circ \tau_{\hat{A} \times \hat{B}, \hat{C}} = & \\
-&[q, {\bf let \: pure \:} r = q {\: \bf in \:}  \langle \pi_1 (\pi_1 r), \langle \pi_2 (\pi_1 r), \pi_2 r \rangle \rangle] \: \circ & \\
-&\quad\quad\quad\quad \circ \: [p, {\bf \: let \: pure \:} x,y,z = {\bf pure\:} (\pi_1 (\pi_1 p)), {\bf pure \:} (\pi_2 (\pi_1 p)), \pi_2 p {\: \bf in \:} \langle \langle x,y \rangle, z \rangle] = & \\
-&\quad\quad\quad\quad\quad\quad \text{Composition}& \\
-&[p, {\bf let \: pure \:} r = q {\: \bf in \:}  \langle \pi_1 (\pi_1 r), \langle \pi_2 (\pi_1 r), \pi_2 r \rangle \rangle & \\
-&\quad\quad\quad\quad [q := {\bf \: let \: pure \:} x,y,z = {\bf pure\:} (\pi_1 (\pi_1 p)), {\bf pure \:} (\pi_2 (\pi_1 p)), \pi_2 p {\: \bf in \:} \langle \langle x,y \rangle, z \rangle]] = & \\
-&\quad\quad\quad\quad\quad\quad \text{Substitution and reduction}& \\
-&[p, {\bf let \: pure \:} r = ({\bf let \: pure \:} x,y,z = {\bf pure\:} (\pi_1 (\pi_1 p)), {\bf pure \:} (\pi_2 (\pi_1 p)), \pi_2 p {\: \bf in \:} \langle \langle x,y \rangle, z \rangle) & \\
-&\quad\quad\quad\quad {\: \bf in \:} \langle \pi_1 (\pi_1 r), \langle \pi_2 (\pi_1 r), \pi_2 r \rangle \rangle] = & \\
-&\quad\quad\quad\quad\quad\quad \text{$\beta$-reduction rule}& \\
-&[p, {\bf let \: pure \:} x,y,z = {\bf pure\:} (\pi_1 (\pi_1 p)), {\bf pure \:} (\pi_2 (\pi_1 p)), \pi_2 p {\: \bf in \:}  \langle \pi_1 (\pi_1 r), \langle \pi_2 (\pi_1 r), \pi_2 r \rangle \rangle & \\
-&\quad\quad\quad\quad [r := \langle \langle x,y \rangle, z \rangle]] = & \\
-&\quad\quad\quad\quad\quad\quad \text{$\beta$-reduction rule}& \\
-&[p, {\bf let \: pure \:} x,y,z = {\bf pure\:} (\pi_1 (\pi_1 p)), {\bf pure \:} (\pi_2 (\pi_1 p)), \pi_2 p {\: \bf in \:} \langle x, \langle y, z \rangle \rangle]&
-\end{array}$
-
-\vspace{\baselineskip}
-
-On the other hand:
-
-\vspace{\baselineskip}
-
-Let us define as follows $\tau_{\hat{A}, \hat{B} \times \hat{C}}$:
-
-$\begin{array}{lll}
-&\tau_{\hat{A}, \hat{B} \times \hat{C}} = [r, {\bf let \: pure \:} x, y, z = ({\bf pure \:} \pi_1 r, {\bf let \: pure \:} q^{'} = \pi_2 r {\: \bf in \:} \pi_1 q^{'}, {\bf let \: pure \:} q^{'} = \pi_2 r {\: \bf in \:} \pi_2 q^{'}) & \\
-&\quad\quad\quad\quad {\: \bf in \:} \langle x, \langle y, z \rangle\rangle]& \\
-\end{array}$
-
-\vspace{\baselineskip}
-
-We may simplify the current instance of tensorial strength:
-
-$\begin{array}{lll}
-&[p, {\bf let \: pure \:} x, y, z = ({\bf pure \:} \pi_1 p, {\bf let \: pure \:} p^{'} = \pi_2 p {\: \bf in \:} \pi_1 p^{'}, {\bf let \: pure \:} p^{'} = \pi_2 p {\: \bf in \:} \pi_2 p^{'}) & \\
-&\quad\quad\quad\quad {\: \bf in \:} \langle x, \langle y, z \rangle\rangle] = & \\
-&\quad\quad\quad\quad\quad\quad \text{Substitution and reduction}& \\
-&[p, {\bf let \: pure \:} x, p^{'}, z = ({\bf pure \:} \pi_1 p, \pi_2 p, {\bf let \: pure \:} p^{'} = \pi_2 p {\: \bf in \:} \pi_2 p^{'}) {\: \bf in \:} \langle x, \langle \pi_1 p^{'}, z \rangle\rangle] = & \\
-&\quad\quad\quad\quad\quad\quad \text{Substitution and reduction}& \\
-&[p, {\bf let \: pure \:} x, p^{'} = {\bf pure \:} \pi_1 p, \pi_2 p {\: \bf in \:} \langle x, \langle \pi_1 p^{'}, \pi_2 p^{'} \rangle\rangle]&
-\end{array}$
-
-\vspace{\baselineskip}
-
-Then:
-
-\vspace{\baselineskip}
-
-$\begin{array}{lll}
-&\tau_{\hat{A}, \hat{B} \times \hat{C}} \circ (id_{\hat{A}} \times \tau_{\hat{B}, \hat{C}}) \circ \alpha_{\hat{A}, \hat{B}, {\bf K}\hat{C}} = & \\
-&\quad\quad\quad\quad\quad\quad \text{Unfolding}& \\
-&\tau_{\hat{A}, \hat{B} \times \hat{C}} \circ [q, \langle \pi_1 q, {\bf let \: pure \:} y, z = {\bf pure \:} (\pi_1 (\pi_2 q)), \pi_2 (\pi_2 q) {\: \bf in \:} \langle y, z \rangle \rangle] \: \circ & \\
-&\quad\quad\quad\quad \circ \: [p, \langle \pi_1 (\pi_1 p), \langle \pi_2 (\pi_1 p), \pi_2 p \rangle \rangle] = & \\
-&\quad\quad\quad\quad\quad\quad \text{Composition}& \\
-&\tau_{\hat{A}, \hat{B} \times \hat{C}} \circ [p, \langle \pi_1 q, {\bf let \: pure \:} y, z = {\bf pure \:} (\pi_1 (\pi_2 q)), \pi_2 (\pi_2 q) {\: \bf in \:} \langle y, z \rangle \rangle & \\
-&\quad\quad\quad\quad [q := \langle \pi_1 (\pi_1 p), \langle \pi_2 (\pi_1 p), \pi_2 p \rangle \rangle]] = & \\
-&\quad\quad\quad\quad\quad\quad \text{Substitution and reduction}& \\
-&\tau_{\hat{A}, \hat{B} \times \hat{C}} \circ [p, \langle \pi_1 (\pi_1 p), {\bf let \: pure \:} y, z = {\bf pure \:} \pi_2 (\pi_1 p), \pi_2 p {\: \bf in \:} \langle y, z \rangle \rangle] = & \\
-&\quad\quad\quad\quad\quad\quad \text{Composition}& \\
-&[r, {\bf let \: pure \:} x, p^{'} = {\bf pure \:} \pi_1 r, \pi_2 r {\: \bf in \:} \langle x, \langle \pi_1 p^{'}, \pi_2 p^{'} \rangle\rangle] \: \circ & \\
-&\quad\quad\quad\quad \circ \: [p, \langle \pi_1 (\pi_1 p), {\bf let \: pure \:} y, z = {\bf pure \:} \pi_2 (\pi_1 p), \pi_2 p {\: \bf in \:} \langle y, z \rangle \rangle] = & \\
-&\quad\quad\quad\quad\quad\quad \text{Substitution and reduction}& \\
-&[p, {\bf let \: pure \:} x, p^{'} = ({\bf pure \:} (\pi_1 (\pi_1 p)), {\bf let \: pure \:} y, z = ({\bf pure \:} \pi_2 (\pi_1 p), \pi_2 p {\: \bf in \:} \langle y, z \rangle \rangle) & \\
-&\quad\quad\quad\quad {\: \bf in \:} \langle x, \langle \pi_1 p^{'}, \pi_2 p^{'} \rangle\rangle ] = & \\
-&\quad\quad\quad\quad\quad\quad \text{Reduction}& \\
-&[p, {\bf let \: pure \:} x,y,z = {\bf pure \:} (\pi_1 (\pi_1 p)), {\bf pure \:} \pi_2 (\pi_1 p), \pi_2 p {\: \bf in \:} \langle x, \langle \pi_1 p^{'}, \pi_2 p^{'} \rangle\rangle [p^{'} := \langle y, z \rangle]] = & \\
-&\quad\quad\quad\quad\quad\quad \text{Substitution and reduction}& \\
-&[p, {\bf let \: pure \:} x,y,z = {\bf pure \:} (\pi_1 (\pi_1 p)), {\bf pure \:} \pi_2 (\pi_1 p), \pi_2 p {\: \bf in \:} \langle x, \langle y, z \rangle\rangle]&
-\end{array}$
-
-\vspace{\baselineskip}
-
-2) Easy
-
-\end{proof}
-
-\begin{lemma}
-
-  {\bf K} is an applicative functor
-\end{lemma}
-
-\begin{proof}
-  Immediately follows from previous lemmas in the section.
-\end{proof}
-
-Similar to \cite{Abramsky}, we apply the translation from $\lambda_{{\bf K}}$ to some cartesian closed category with an abritraty applicative functor $\mathcal{K}$, then
-we have $[\![\Gamma \vdash M : A]\!] = [x, M [x_i := \pi_i x]]$, so $M =_{\beta \eta} N \Leftrightarrow [\![\Gamma \vdash M : A]\!] = [\![\Gamma \vdash N : A]\!]$.
-
-\end{proof}
-
-  \addcontentsline{toc}{section}{References}
+\addcontentsline{toc}{section}{Список использованной литературы}
 
@@
 \end{thebibliography}
-
-\end{document}

### 2018-03-31 MT/HaskellFAQ.tex
update

@@
-\section{Глоссарий по основным конструкциям функционального языка программирования Haskell: функторы, монады, аппликативные функторы}
+\subsection{Глоссарий по основным конструкциям функционального языка программирования Haskell: функторы и аппликативные функторы}
 
@@
 
-  Классом типов в языке Haskell -- это реализация некоторого общего интерфейса для совокупности типов.
+  Классом типов в языке Haskell -- это реализация некоторого общего интерфейса для некоторой совокупности типов.
 
-  Представителем (или наследников) класса типов называется реализация данного класса для конкретного типа.
+  Представителем класса типов называется реализация данного класса для конкретного типа.
 \end{defin}
@@
 
-  Функтор -- это однопараметрический класс типов, позволяющий пронести действие функции через значения,
+  Функтор -- это однопараметрический класс типов, позволяющий пронести действие одноместной функции через значения,
   полученные в результате применения к их типу одноместного типового оператора.
@@
 
-Определение в стандартной
-библиотеке выглядит следующим образом:
+Определение в стандартной библиотеке выглядит следующим образом:
 
@@
   \item Список (неограниченная в длине последовательность) является функтором:
-  Определение в стандартной
-  библиотеке выглядит следующим образом:
 
@@
   получен применением функции к каждому элементу списка, полученного на вход.
+
+  Пример использования:
+
+\begin{lstlisting}[language=Haskell]
+  > fmap succ ''bg`lo`fmd\USrtodqmnu`''
+  ''champagne supernova''
+\end{lstlisting}
+
+В данном примере мы отобразили функцию ``следующий за''
+\footnote{В языке Haskell функция ``следующий за'' определения для
+любого типа, между элементами которого есть линейный порядок,
+в частности, лексиграфический порядок на символах, вводимых с клавиатуры компьютера}
+на строчку, указанную выше. Строка в языка Haskell -- это частный случай списка, списка символов.
+Тогда \verb"fmap" применяет функцию \verb"succ" к каждой букве строки, возвращая в результате
+строку со словосочетанием ``champagne supernova''.
+
   \item Пара (тип декартова произведения типов) также функтор:
@@
 \end{itemize}
+
+\begin{defin} Аппликативный функтор
+
+  Аппликативным функтором называется класс типов, обобщающий функтор для функций произвольной арности.
+\end{defin}
+
+Определение класса в языке Haskell:
+
+\begin{lstlisting}[language=Haskell]
+  class Functor f => Applicative f where
+    pure :: a -> f a
+    (<*>) :: f (a -> b) -> f a -> f b
+\end{lstlisting}
+
+Обобщение действия функтора с помощью методой класса \verb"Applicative" для произвольной функции:
+\begin{lstlisting}[language=Haskell]
+  liftA2 :: (a -> b -> c) -> f a -> f b -> f c
+  liftA2 g x y = pure g <*> x <*> y
+\end{lstlisting}
+
+Данный комбинатор берет функцию, которая по объектам из типов $a$ и $b$ сопоставляет объект типа $c$ и,
+по аргументам $x$ и $y$ типов соответственно $f \: a$ и $f \: b$, полученных в результате применения к данным типов функтора $f$,
+возвращает объект типа $f \: c$, тип которого также получен в результате применения функтора к типу $f \: c$.
+
+Рассмотрим реализацию данной функции детальнее. Имея функцию $g$ типа \verb"a -> b -> c", мы
+применяет к ней \verb"pure", получая в результате объект \verb"pure g" типа \verb"f (a -> b -> c)",
+иными словами, мы подняли функцию $g$ на уровень функтора $f$.
+
+Далее мы применяем поднятую функцию $f$ к аргументу $x$ типа $f \: a$ с использованием \verb"(<*>)"
+и получаем объект \verb"pure g <*> x" типа \verb"f (b -> c)", получив одноместную функцию из \verb"b" в \verb"c".
+Затем мы \verb"pure g <*> x" применяем к аргументу $y$ типа $f \: b$ опять же с использованием
+\verb"(<*>)" и получаем объект типа $f \: c$.
+
+Рассмотрим примеры представителей класса типов \verb"Applicative":
+
+\begin{itemize}
+  \item Списки
+  \begin{lstlisting}[language=Haskell]
+    instance Applicative [] where
+      pure x = [x]
+      fs <*> xs = [ f x | f <- fs , x <- xs]
+  \end{lstlisting}
+
+  Метод \verb"pure" для списков переводит объект $x$ произвольного типа $a$ в
+  одноэлементный список $[x]$. Метод \verb"(<*>)" в качестве левого операдна
+  принимает список функций и список аргументов -- в качестве правого и возващает
+  список всевозможных применений элементов первого списка к элементам второго списка.
+  Нотация здесь является калькой с теоретико-множественной нотации, которая могла бы иметь
+  следующий вид $\{ f(x) \: | \: f \in B^A \land x \in A \}$.
+  \item Пары
+  \begin{lstlisting}[language=Haskell]
+  instance Monoid a => Applicative ((,) a) where
+      pure x = (mempty, x)
+      (u, f) <*> (v, x) = (u <> v, f x)
+  \end{lstlisting}
+
+  Чтобы пару сделать аппликативным функтором, необходимо соблюсти следующее ограничение:
+  тип первой координаты должен быть моноидом \footnote{То есть тип должен содержать нейтральный и бинарную ассоциативную операцию.
+  Моноид в языке Haskell также является классом типов, который, как видно из названия, является калькой с одноименной структуры в алгебре.}.
+
+  Метод \verb"pure" переводит объект $x$ произвольного типа $b$ в упорядоченную пару, первый элемент которой единица моноида $a$,
+  а второй -- $x$, полученный при входе. Метод \verb"(<*>)" принимает две пары: первая пара -- это пара элемента $u$ моноида и некоторой функции $f$,
+  вторая пара состоит из другого элемента $v$ моноида и аргумента $x$. Возвращаемый результат:
+  соединение элементво $u$ и $v$ с помощью моноидной операции в первой координате, а во второй координате -- применение функции $f$ к аргументу $x$.
+
+  Пример использования:
+  \begin{lstlisting}[language=Haskell]
+  > (''(what's the story) '' , succ) <*> (''morning glory?'' , 1994)
+  (''(what's the story) morning glory?'',1995)
+  \end{lstlisting}
+  где \verb"succ" -- функция ``следующий за''.
+
+  \item Тип Maybe
+  \begin{lstlisting}[language=Haskell]
+    instance Applicative Maybe where
+      pure = Just
+      Just f  <*> m = fmap f m
+      Nothing <*> _m = Nothing
+  \end{lstlisting}
+
+  Пример использования:
+  \begin{lstlisting}[language=Haskell]
+  > liftA2 (++) (Just "Definitely") Nothing
+  Nothing
+
+  > liftA2 (++) (Just "Definitely ") (Just "Maybe")
+  Just ''Definitely Maybe''
+  \end{lstlisting}
+
+Данный пример как раз является примером того, как действие функтора обобщается на функций многих аргументов,
+в данном случае, двух аргументов. Здесь функцией двух аргументов является функция $(++)$, конкатенация строк.
+
+В первом случае определен только первый аргумент, строка \verb"Definitely", обернутая в \verb"Just",
+второй же аргумент является неопределенным (так как в качестве второго аргумента передан \verb"Nothing"),
+поэтому данная попытка конкатенации строк также вернет \verb"Nothing".
+
+Во втором примере, второй аргумент определен, это строка \verb"Maybe", обернутая в \verb"Just".
+Тогда конкатенация пройдет успешно и \verb"liftA2" пронесет комбинатор $(++)$ и совершит конкатенацию строк
+\verb"Definitely" и \verb"Maybe" внутри \verb"Just", и на выходе будет возвращена строка \verb"Definitely Maybe",
+к которой применен \verb"Just".
+\end{itemize}

### 2018-03-31 MT/HaskellFAQ.tex
update

@@
 
-  \item Тип $Maybe$ -- это однопараметрический типовой оператор, для обработки неопределенных значений:
+  \item Тип $Maybe$ -- это однопараметрический типовой оператор, для обработки неопределенных значений, иными словами тип $Maybe$ доопределяет
+  частично-определенную функцию до тотальной:
   \begin{lstlisting}[language=Haskell]
@@
 
+  В данной реализации представителя, \verb"pure" просто оборачивает значение типа произольно типа $a$ в $Just$,
+  создавая таким образом объект типа \verb"Maybe a". Метод \verb"(<*>)" принимает функцию типа \verb"Maybe (a -> b)" и
+  аргумент типа \verb"Maybe a". Если аргументы, переданные \verb"(<*>)", имеют вид $\verb"Just f"$ и $\verb"Just m"$,
+  где $f$ и $m$ -- это объекты типов \verb"a -> b" и \verb"a" соответственно, тогда происходит применение функции $f$
+  к аргументу $m$ внутри \verb"Just" и возвращаемое значение в таком случае имеет вид \verb"Just (f m)".
+  В остальных случаях возвращается \verb"Nothing".
+
   Пример использования:
@@
 
-  > liftA2 (++) (Just "Definitely ") (Just "Maybe")
+  > liftA2 (++) (Just ''Definitely '') (Just "Maybe")
   Just ''Definitely Maybe''

### 2018-04-22 article.tex
added brand new lemma in article
update in article

@@
 \usepackage{mathabx}
+\usepackage{stmaryrd}
 \usepackage[all, 2cell]{xy}
@@
 
-  % This is a comment, not shown in final output.
-  % The following shows typesetting  power of LaTeX:
-
   \section{Introduction}
@@
   IEL$^{-}$  is a variant of IEL, that corresponds to intuitionistic belief.
-  Informally, ${\bf K}A$ denotes that $A$ is verified intuitionistically.
+  Informally, $\Box A$ denotes that $A$ is verified intuitionistically.
 
@@
 
-    2) ${\bf K}(A \to B) \to ({\bf K}A \to {\bf K}B)$ (normality);
-
-    3) $A \to {\bf K}A$ (co-reflection);
+    2) $\Box (A \to B) \to (\Box A \to \Box B)$ (normality);
+
+    3) $A \to \Box A$ (co-reflection);
 
@@
 
-  Functional programming languages such as Haskell \cite{Haskell}, Idris \cite{Idris}, Purescript \cite{Purs}
-  Elm \cite{Elm} or Scala \cite{Scala} have special type classes\footnote{Type class in Haskell is a general interface for
+  Functional programming languages such as Haskell \cite{Haskell}, Idris \cite{Idris} or Purescript \cite{Purs}
+  have special type classes\footnote{Type class in Haskell is a general interface for
   special group of datatypes.} for calculations with container types like \verb"Functor" and
@@
   We investigate the relationship between intuitionistic epistemic logic IEL$^{-}$ and applicative programming with side-effects
-  by constructing the type system (which is called $\lambda_{{\bf K}}$) which is Curry-Howard isomorphic to $IEL^{-}$.
-  So we will consider ${\bf K}$-modality as an arbitrary applicative functor and we prove that obtained type system
+  by constructing the type system (which is called $\lambda_{\text{IEL}^{-}}$) which is Curry-Howard isomorphic to $IEL^{-}$.
+  So we will consider IEL$^{-}$ modality as an arbitrary applicative functor and we prove that obtained type system
   is sound and complete for applicative functor on cartesian closed category (using the categorical definition proposed by Paterson \cite{Cons}).
 
-  $\lambda_{{\bf K}}$ consists of the rules for simply typed lambda-calculus and special typing rules for
-  lifting types into the applicative functor ${\bf K}$. We assume that our type system will
+  $\lambda_{\text{IEL}^{-}}$ consists of the rules for simply typed lambda-calculus and special typing rules for
+  lifting types into the applicative functor $\Box$. We assume that our type system will
   axiomatize the simplest case of computation with effects with one container. We provide a proof-theoretical
@@
     	\AxiomC{$\Gamma \vdash A$}
-    	\RightLabel{${\bf K}_I$}
-    	\UnaryInfC{$\Gamma \vdash {\bf K}A$}
+    	\RightLabel{$\Box_I$}
+    	\UnaryInfC{$\Gamma \vdash \Box A$}
   \end{prooftree}
@@
   	\begin{prooftree}
-    \AxiomC{$\Gamma \vdash {\bf K} A_1, \dots, \Gamma \vdash {\bf K} A_n $}
+    \AxiomC{$\Gamma \vdash \Box A_1, \dots, \Gamma \vdash \Box A_n $}
     \AxiomC{$A_1,\dots,A_n \vdash B$}
-    \BinaryInfC{$\Gamma \vdash {\bf K} B$}
+    \BinaryInfC{$\Gamma \vdash \Box B$}
   	\end{prooftree}
@@
 
-We will denote $\Gamma \vdash {\bf K} A_1, \dots, \Gamma \vdash {\bf K} A_n$ and $A_1,\dots,A_n \vdash B$ as $\Gamma \vdash {\bf K} \vec{A}$ and $\vec{A} \vdash B$ for brevity.
+We will denote $\Gamma \vdash \Box A_1, \dots, \Gamma \vdash \Box A_n$ and $A_1,\dots,A_n \vdash B$ as $\Gamma \vdash \Box \vec{A}$ and $\vec{A} \vdash B$ for brevity.
 
@@
 
-1) If $\Gamma \vdash_{\text{NIEL}^{-}} A$, then $\text{IEL}^{-} \vdash \bigwedge \Gamma \rightarrow {\bf K}A$.
+1) If $\Gamma \vdash_{\text{NIEL}^{-}} A$, then $\text{IEL}^{-} \vdash \bigwedge \Gamma \rightarrow \Box A$.
 
@@
 (1) & \bigwedge \Gamma \rightarrow A & \text{assumption}\\
-(2) & A \rightarrow {\bf K}A &\text{co-reflection}\\
-(3) & (\bigwedge \Gamma \rightarrow A) \rightarrow ((A \rightarrow {\bf K}A) \rightarrow (\bigwedge \Gamma \rightarrow {\bf K}A))&\text{IPC theorem}\\
-(4) & (A \rightarrow {\bf K}A) \rightarrow (\bigwedge \Gamma \rightarrow {\bf K}A) &\text{from (1), (3) and MP}\\
-(5) & \bigwedge \Gamma \rightarrow {\bf K}A &\text{from (2), (4) and MP}\\
-\end{array}$
-
-\vspace{\baselineskip}
-
-2) If $\Gamma \vdash_{\text{NIEL}^{-}} {\bf K} \vec{A}$ and $\vec{A} \vdash B$, then $\text{IEL}^{-} \vdash \bigwedge \Gamma \rightarrow {\bf K}B$.
-
-$\begin{array}{lll}
-(1) &\bigwedge \Gamma \rightarrow \bigwedge \limits_{i = 1}^{n} {\bf K} A_i & \text{assumption} \\
-(2) &\bigwedge \limits_{i = 1}^{n} {\bf K} A_i \rightarrow {\bf K} \bigwedge \limits_{i = 1}^{n} A_i& \text{IEL$^{-}$ theorem} \\
-(3) &\bigwedge \Gamma \rightarrow {\bf K} \bigwedge \limits_{i = 1}^{n} A_i & \text{from (1), (2) and transitivity} \\
-(4) &\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B& \text{assumption} \\
-(5) &(\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B) \rightarrow {\bf K} (\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B)& \text{co-reflection}\\
-(6) &{\bf K} (\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B)& \text{from (4), (5) and MP} \\
-(7) &{\bf K} \bigwedge \limits_{i = 1}^{n} A_i \rightarrow {\bf K} B & \text{from (6) and normality} \\
-(8) &\bigwedge \Gamma \rightarrow {\bf K} B & \text{from (3), (7) and transitivity}
+(2) & A \rightarrow \Box A &\text{co-reflection}\\
+(3) & (\bigwedge \Gamma \rightarrow A) \rightarrow ((A \rightarrow \Box A) \rightarrow (\bigwedge \Gamma \rightarrow \Box A))&\text{IPC theorem}\\
+(4) & (A \rightarrow \Box A) \rightarrow (\bigwedge \Gamma \rightarrow \Box A) &\text{from (1), (3) and MP}\\
+(5) & \bigwedge \Gamma \rightarrow \Box A &\text{from (2), (4) and MP}\\
+\end{array}$
+
+\vspace{\baselineskip}
+
+2) If $\Gamma \vdash_{\text{NIEL}^{-}} \Box \vec{A}$ and $\vec{A} \vdash B$, then $\text{IEL}^{-} \vdash \bigwedge \Gamma \rightarrow \Box B$.
+
+$\begin{array}{lll}
+(1) &\bigwedge \Gamma \rightarrow \Box A_1, \dots, \bigwedge \Gamma \rightarrow \Box A_1 & \text{assumption} \\
+(2) &\bigwedge \Gamma \rightarrow \bigwedge \limits_{i = 1}^{n} \Box A_i & \text{IEL$^{-}$ theorem} \\
+(3) &\bigwedge \limits_{i = 1}^{n} \Box A_i \rightarrow \Box \bigwedge \limits_{i = 1}^{n} A_i & \text{IEL$^{-}$ theorem} \\
+(4) &\bigwedge \Gamma \rightarrow \Box \bigwedge \limits_{i = 1}^{n} A_i & \text{from (1), (2) and transitivity} \\
+(5) &\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B& \text{assumption} \\
+(6) &(\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B) \rightarrow \Box (\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B)& \text{co-reflection}\\
+(7) &\Box (\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B)& \text{from (4), (5) and MP} \\
+(8) &\Box \bigwedge \limits_{i = 1}^{n} A_i \rightarrow \Box B & \text{from (6) and normality} \\
+(9) &\bigwedge \Gamma \rightarrow \Box B & \text{from (3), (7) and transitivity}
 \end{array}$
@@
 
-  Let $\mathbb{V}$ be the set of variables. The set $\Lambda_{{\bf K}}$ of terms is defined by the grammar:
-
-$\begin{array}{lll}
-& \Lambda_{{\bf K}} ::= \mathbb{V} \: | \:  (\lambda \mathbb{V}.\Lambda_{{\bf K}}) \: | \: (\Lambda_{{\bf K}}\Lambda_{{\bf K}}) \: | \: (\Lambda_{{\bf K}} , \Lambda_{{\bf K}}) \: | \: (\pi_1 \Lambda_{{\bf K}}) \: | \: (\pi_2 \Lambda_{{\bf K}}) \: | & \\
-& \quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad ({\bf pure \: } \: \Lambda_{{\bf K}}) \: | \: ({\bf let \:\: pure} \:\: \mathbb{V}^{*} = \Lambda_{{\bf K}}^{*} \:\: {\bf in} \:\: \Lambda_{{\bf K}})
-\end{array}$
-
-  \end{defin}
-
-Where $\mathbb{V}^{*}$ and $\Lambda_{{\bf K}}^{*}$ denote the set of finite sequences of variables $\bigcup \limits_{i=0}^{\infty} \mathbb{V}^i$
-and the set of finite sequences of terms $\bigcup \limits_{i = 0}^{\infty} {\Lambda_{{\bf K}}}^i $.
+  Let $\mathbb{V}$ be the set of variables. The set $\Lambda_{\Box}$ of terms is defined by the grammar:
+
+$\begin{array}{lll}
+& \Lambda_{\Box} ::= \mathbb{V} \: | \:  (\lambda \mathbb{V}.\Lambda_{\Box}) \: | \: (\Lambda_{\Box} \Lambda_{\Box}) \: | \: \langle \Lambda_{\Box} , \Lambda_{\Box} \rangle \: | \: (\pi_1 \Lambda_{\Box}) \: | \: (\pi_2 \Lambda_{\Box}) \: | & \\
+& \quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad ({\bf box \: } \:\Lambda_{\Box}) \: | \: ({\bf let \: box \:} \mathbb{V}^{*} = \Lambda_{\Box}^{*} \:\: {\bf in} \:\: \Lambda_{\Box})
+\end{array}$
+
+  \end{defin}
+
+Where $\mathbb{V}^{*}$ and $\Lambda_{\Box}^{*}$ denote the set of finite sequences of variables $\bigcup \limits_{i=0}^{\infty} \mathbb{V}^i$
+and the set of finite sequences of terms $\bigcup \limits_{i = 0}^{\infty} \Lambda_{\Box}^i $.
 Note that the sequence of variables $\vec{x}$ and the sequence of terms $\vec{M}$ should have the same length.
@@
 
-  Let $\mathbb{T}$ be the set of atomic types. The set $\mathbb{T}_{{\bf K}}$ of types with applicative
-  functor ${\bf K}$ is generated by the grammar:
+  Let $\mathbb{T}$ be the set of atomic types. The set $\mathbb{T}_{\Box}$ of types with applicative
+  functor $\Box$ is generated by the grammar:
   \begin{equation}
-    \mathbb{T}_{{\bf K}} ::= \mathbb{T} \: | \: (\mathbb{T}_{{\bf K}} \to \mathbb{T}_{{\bf K}}) \: |
-    \: (\mathbb{T}_{{\bf K}} \times \mathbb{T}_{{\bf K}}) \: | \: ({\bf K}\mathbb{T}_{{\bf K}})
+    \mathbb{T}_{\Box} ::= \mathbb{T} \: | \: (\mathbb{T}_{\Box} \to \mathbb{T}_{\Box} \: |
+    \: (\mathbb{T}_{\Box} \times \mathbb{T}_{\Box}) \: | \: (\Box \mathbb{T}_{\Box})
   \end{equation}
@@
 
-  \begin{defin} Modal typed lambda calculus $\lambda_{{\bf K}}$ based on NIEL$^{-}_{\land, \to}$:
+  \begin{defin} Modal typed lambda calculus $\lambda_{\text{IEL}^{-}}$ based on NIEL$^{-}_{\land, \to}$:
 
@@
         \AxiomC{$\Gamma \vdash M : A$}
-        \RightLabel{${\bf K}_I$}
-        \UnaryInfC{$\Gamma \vdash {\bf pure \: } \: M : {\bf K}A $}
+        \RightLabel{$\Box_I$}
+        \UnaryInfC{$\Gamma \vdash {\bf box \: } \: M : \Box A $}
       \end{prooftree}
@@
     \begin{prooftree}
-      \AxiomC{$\Gamma \vdash \vec{M} : {\bf K} \vec{A}$}
+      \AxiomC{$\Gamma \vdash \vec{M} : \Box \vec{A}$}
       \AxiomC{$\vec{x} : \vec{A} \vdash N : B$}
-      \RightLabel{$\text{let}_{{\bf K}}$}
-      \BinaryInfC{$\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } N : {\bf K} B$}
+      \RightLabel{$\text{let}_{\Box}$}
+      \BinaryInfC{$\Gamma \vdash {\bf let \: box \:} \vec{x} = \vec{M} {\: \bf in \: } N : \Box B$}
     \end{prooftree}
@@
 
-  ${\bf K}_I$-typing rule is the same as $\bigcirc$-introduction in lax logic (also known as monadic
-  metalanguage \cite{Lax}) and in typed lambda-calculus which is derived by proof-assignment for lax-logic
-  proofs.
-  ${\bf K}_I$ allows to inject an object of type $\alpha$ into the functor. ${\bf K}_I$  reflects the
-  Haskell method {\bf pure} for Applicative class. It plays the same role as the {\bf return} method in Monad class.
-
-  $\text{let}_{{\bf K}}$ is similar to the $\Box$-rule in typed lambda calculus for intuitionistic normal modal logic {\bf IK}, which is described in \cite{ModalK}.
-
-$\Gamma \vdash \vec{M} : {\bf K} \vec{A}$ is a syntax sugar for the sequence $\Gamma \vdash M_1 : {\bf K}A_1,\dots,\Gamma \vdash M_n : {\bf K}A_n$ and $\vec{x} : \vec{A} \vdash N : B$ is a short form for $x_1 : A_1, \dots, x_n : A_n \vdash N : B$.
-${\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } N$ is a simultaneous local binding in $N$.
-We use this short form instead of ${\bf let \: pure\:} x_1,\dots,x_n = M_1,\dots,M_n {\: \bf in \:} N$.
-
-In fact, our calculus is the extention of typed lambda calculus for {\bf IK} by ${\bf K}_I$-rule that is appropriate to co-reflection.
-
-  \vspace{\baselineskip}
-
-Here are some examples of closed terms:
+  $\Box_I$-typing rule is the same as $\bigcirc$-introduction in monadic
+  metalanguage \cite{Lax}. $\Box_I$ allows to inject an object of type $A$ into the functor $\Box$. $\Box_I$ reflects the
+  Haskell method {\bf box} for Applicative class. It plays the same role as the {\bf return} method in Monad class.
+
+  $\text{let}_{\Box}$ is similar to the $\Box$-rule in typed lambda calculus for intuitionistic normal modal logic {\bf IK}, which is described in \cite{ModalK}.
+
+$\Gamma \vdash \vec{M} : \Box \vec{A}$ is a syntax sugar for the sequence $\Gamma \vdash M_1 : \Box A_1,\dots,\Gamma \vdash M_n : \Box A_n$ and $\vec{x} : \vec{A} \vdash N : B$ is a short form for $x_1 : A_1, \dots, x_n : A_n \vdash N : B$.
+${\bf let \: box \:} \vec{x} = \vec{M} {\: \bf in \: } N$ is a simultaneous local binding in $N$.
+We use this short form instead of ${\bf let \: box \:} x_1,\dots,x_n = M_1,\dots,M_n {\: \bf in \:} N$.
+
+In fact, our calculus is the extention of typed lambda calculus for {\bf IK} by $\Box_I$-rule that is appropriate to co-reflection.
+
+  \vspace{\baselineskip}
+
+Here are some examples of derivation trees:
 
@@
 \AxiomC{$x : A \vdash x : A$}
-\UnaryInfC{$x : A \vdash {\bf pure \:} x : {\bf K} A$}
-\UnaryInfC{$\vdash (\lambda x. {\bf pure \: } x) : A \to {\bf K}A$}
-\end{prooftree}
-
-\begin{prooftree}
-\AxiomC{$f : {\bf K} (A \to B) \vdash f : {\bf K} (A \to B)$}
-\AxiomC{$x : {\bf K} A \vdash x : {\bf K} A $}
+\UnaryInfC{$x : A \vdash {\bf box \:} x : \Box A$}
+\UnaryInfC{$\vdash (\lambda x. {\bf box \: } x) : A \to \Box A$}
+\end{prooftree}
+
+\begin{prooftree}
+\AxiomC{$f : \Box (A \to B) \vdash f : \Box (A \to B)$}
+\AxiomC{$x : \Box A \vdash x : \Box A $}
 \AxiomC{$g : A \to B \vdash g : A \to B$}
@@
 \BinaryInfC{$g : A \to B, y : A \vdash g y : B$}
-\RightLabel{$\text{let}_{{\bf K}}$}
-\TrinaryInfC{$f : {\bf K} (A \to B), x : {\bf K} A \vdash {\bf let \: pure \:} g, y  = f, x {\: \bf in \:} g y : {\bf K} B$}
-\UnaryInfC{$f : {\bf K} (A \to B) \vdash \lambda x. {\bf let \: pure\:} g, y = f, x {\: \bf in \:} g y : {\bf K} A \to {\bf K} B$}
-\UnaryInfC{$\vdash \lambda f. \lambda x. {\bf let \:pure \:} g, y = f, x {\: \bf in \:} g y : {\bf K}(A \to B) \to {\bf K} A \to {\bf K} B$}
-\end{prooftree}
-
-  \vspace{\baselineskip}
-
-  Now we define free variables and substitutions. $\beta$-reduction, multi-step $\beta$-reduction and $\beta$-equality are defined standardly:
+\RightLabel{$\text{let}_{\Box}$}
+\TrinaryInfC{$f : \Box (A \to B), x : \Box A \vdash {\bf let \: box \:} g, y  = f, x {\: \bf in \:} g y : \Box B$}
+\UnaryInfC{$f : \Box (A \to B) \vdash \lambda x. {\bf let \: box \:} g, y = f, x {\: \bf in \:} g y : \Box A \to \Box B$}
+\UnaryInfC{$\vdash \lambda f. \lambda x. {\bf let \: box \:} g, y = f, x {\: \bf in \:} g y : \Box (A \to B) \to \Box A \to \Box B$}
+\end{prooftree}
+
+  \vspace{\baselineskip}
+
+  Now we define free variables and substitutions:
 
@@
 
-  1) $FV(x) = \{ x \}$;
-
-  2) $FV(\lambda x. M) = FV(M) \setminus \{ x\}$;
-
-  3) $FV(M N) = FV(M) \cup FV(N);$
-
-  4) $FV(\langle M,N \rangle) = FV(M) \cup FV(N)$;
-
-  5) $FV(\pi_i M) \subseteq FV(M)$, $i \in \{ 1, 2\}$;
-
-  6) $FV(\text{pure } M) = FV(M)$;
-
-  7) $FV({\bf let \: pure} \: \vec{x} = \vec{M} \:\: {\bf in} \:\: N) = \bigcup \limits_{i = 1}^n FV(M), \text{where $n = |\vec{M}|$}$.
+\begin{enumerate}
+  \item $FV(x) = \{ x \}$;
+  \item $FV(\lambda x. M) = FV(M) \setminus \{ x\}$;
+  \item $FV(M N) = FV(M) \cup FV(N);$
+  \item $FV(\langle M,N \rangle) = FV(M) \cup FV(N)$;
+  \item $FV(\pi_i M) \subseteq FV(M)$, $i \in \{ 1, 2\}$;
+  \item $FV(\text{\bf box } M) = FV(M)$;
+  \item $FV({\bf let \: box} \: \vec{x} = \vec{M} {\: \bf in \:} N) = \bigcup \limits_{i = 1}^n FV(M), \text{where $n = |\vec{M}|$}$.
+\end{enumerate}
   \end{defin}
@@
 
-  1) $x [x := N] = N$, $x [y := N] = x$;
-
-  2) $(M N) [x := N] = M[x := N] N [x := N]$;
-
-  3) $(\lambda x. M) [x := N] = \lambda x. M [y := N]$, $y \in FV(M)$;
-
-  4) $(M, N)[x := P] = (M[x := P], N [x := P])$;
-
-  5) $(\pi_i M) [x := P] = \pi_i (M[x := P])$, $i \in \{ 1, 2\}$;
-
-  6) $({\bf pure \: } M) [x := P] = {\bf pure \: } (M [x := P])$;
-
-  7) $({\bf let \: pure \:}\vec{x} = \vec{M} {\: \bf in \:} N) [y := P] = {\bf let \: pure \:} \vec{x} = (\vec{M} [y := P]) {\: \bf in \:} N$.
-  \end{defin}
-
-Substitution and free variable for terms of kind ${\bf let \: pure}$ are defined similary to \cite{ModalK}.
-
-  \begin{defin} Type substituition
-
-    The substituition of type $C$ for type variable $B$ in type $A$ is defined inductively:
-
-    1) $B [B := C] = B$ and $D [B := C] = D$, if $B \neq D$;
-
-    2) $(A_1 \alpha A_2)[B := C] = (A_1[B := C]) \alpha (A_2[B := C])$, where $\alpha \in \{ \to, \times \}$;
-
-    3) $({\bf K}A)[B := C] = {\bf K}(A [B := C])$;
-
-    4) Let $\Gamma$ be the context, then $\Gamma [B := C] = \{ x : (A[B := C]) \: | \: x : A \in \Gamma \}$.
-  \end{defin}
-
-  \begin{defin} $\beta$-reduction and $\eta$-reduction rules for $\lambda_{{\bf K}}$.
-
-  1) $(\lambda x. M) N \rightarrow_{\beta} M [x := N]$;
-
-  2) $\pi_1 \langle M, N \rangle \rightarrow_{\beta} M$;
-
-  3) $\pi_2 \langle M, N \rangle \rightarrow_{\beta} N$;
-
-  4) $\begin{array}{llll}
-  & {\bf let \: pure \:} \vec{x}, y, \vec{z} = \vec{M}, {\bf let \: pure \: } \vec{w} = \vec{N} {\: \bf in \: } Q, \vec{P} {\: in \:} R \rightarrow_{\beta} & \\
-  & {\bf let \: pure \:} \vec{x}, \vec{w}, \vec{z} = \vec{M}, \vec{N}, \vec{P} {\: \bf in \: } R [y := Q]
-  \end{array}$
-
-  5) ${\bf let \: pure \:} \vec{x} = {\bf pure \:} \vec{M} {\: \bf in \:} N \rightarrow_{\beta} {\bf pure \:} N [\vec{x} := \vec{M}]$
-
-  6) ${\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} M \rightarrow_{\beta} {\bf pure \:} M$, where \underline{\quad} is an empty sequence of terms.
-
-  7) $\lambda x. f x \rightarrow_{\eta} f$;
-
-  8) $\langle \pi_1 P, \pi_2 P \rangle \rightarrow_{\eta} P$;
-
-  9) ${\bf let \: pure \:} x = M {\: \bf in \: } x \rightarrow_{\eta} M$;
-
+\begin{enumerate}
+  \item $x [x := N] = N$, $x [y := N] = x$;
+  \item $(M N) [x := N] = M[x := N] N [x := N]$;
+  \item $(\lambda x. M) [x := N] = \lambda x. M [y := N]$, $y \in FV(M)$;
+  \item $(M, N)[x := P] = (M[x := P], N [x := P])$;
+  \item $(\pi_i M) [x := P] = \pi_i (M[x := P])$, $i \in \{ 1, 2\}$;
+  \item $({\bf box \: } M) [x := P] = {\bf box \: } (M [x := P])$;
+  \item $({\bf let \: box \:}\vec{x} = \vec{M} {\: \bf in \:} N) [y := P] = {\bf let \: box \:} \vec{x} = (\vec{M} [y := P]) {\: \bf in \:} N$.
+\end{enumerate}
+  \end{defin}
+
+Substitution and free variable for terms of kind ${\bf let \: box}$ are defined similary to \cite{ModalK}.
+
+  \begin{defin} $\beta$-reduction and $\eta$-reduction rules for $\lambda_{\text{IEL}^{-}}$.
+
+\begin{enumerate}
+  \item $(\lambda x. M) N \rightarrow_{\beta} M [x := N]$;
+  \item $\pi_1 \langle M, N \rangle \rightarrow_{\beta} M$;
+  \item $\pi_2 \langle M, N \rangle \rightarrow_{\beta} N$;
+  \item ${\bf let \: box \:} \vec{x}, y, \vec{z} = \vec{M}, {\bf let \: box \: } \vec{w} = \vec{N} {\: \bf in \: } Q, \vec{P} {\: in \:} R \rightarrow_{\beta} \\
+  {\bf let \: box \:} \vec{x}, \vec{w}, \vec{z} = \vec{M}, \vec{N}, \vec{P} {\: \bf in \: } R [y := Q]$
+  \item ${\bf let \: box \:} \vec{x} = {\bf box \:} \vec{M} {\: \bf in \:} N \rightarrow_{\beta} {\bf box \:} N [\vec{x} := \vec{M}]$
+  \item ${\bf let \: box \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} M \rightarrow_{\beta} {\bf box \:} M$, where \underline{\quad} is an empty sequence of terms.
+  \item $\lambda x. f x \rightarrow_{\eta} f$;
+  \item $\langle \pi_1 P, \pi_2 P \rangle \rightarrow_{\eta} P$;
+  \item ${\bf let \: box \:} x = M {\: \bf in \: } x \rightarrow_{\eta} M$;
+\end{enumerate}
   \end{defin}
@@
 By default we use call-by-name evaluation strategy.
-
-This commuting relation denotes that we identify terms of kind {\bf let pure} if they are differ in the order of binding.
 
@@
 
-\begin{lemma} Generation for ${\bf K}_I$.
-
-  Let $\Gamma \vdash {\bf pure \:} M : {\bf K}A$, then $\Gamma \vdash M : A$;
+\begin{lemma} Generation for $\Box_I$.
+
+  Let $\Gamma \vdash {\bf box \:} M : \Box A$, then $\Gamma \vdash M : A$;
 \end{lemma}
@@
 
-\begin{itemize}
+\begin{enumerate}
   \item If $\Gamma \vdash M : A$ and $\Gamma \subseteq \Delta$, then $\Delta \vdash M : A$;
@@
   \item If $\Gamma, x : A \vdash M : B$ and $\Gamma \vdash N : A$, then $\Gamma \vdash M [x := N] : B$.
-  \item If $\Gamma \vdash M : A$, then $\Gamma [B := C] \vdash M : (A [B := C])$.
-\end{itemize}
+\end{enumerate}
 \end{lemma}
@@
 \begin{prooftree}
-  \AxiomC{$\Gamma \vdash \vec{M} : {\bf K} \vec{A}$}
+  \AxiomC{$\Gamma \vdash \vec{M} : \Box \vec{A}$}
   \AxiomC{$\vec{x} : \vec{A} \vdash N : B$}
-  \RightLabel{$\text{let}_{{\bf K}}$}
-  \BinaryInfC{$\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } N : {\bf K}B$}
-\end{prooftree}
-
-By IH $\Delta \vdash \vec{M} : {\bf K} \vec{A}$, so $\Delta \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } N : {\bf K}B$.
-
-  \vspace{\baselineskip}
-
-2) Similary.
-
-  \vspace{\baselineskip}
+  \RightLabel{$\text{let}_{\Box}$}
+  \BinaryInfC{$\Gamma \vdash {\bf let \: box \:} \vec{x} = \vec{M} {\: \bf in \: } N : \Box B$}
+\end{prooftree}
+
+By IH $\Delta \vdash \vec{M} : \Box \vec{A}$, so $\Delta \vdash {\bf let \: box \:} \vec{x} = \vec{M} {\: \bf in \: } N : \Box B$.
+
+  \vspace{\baselineskip}
+
+2)-3) Similary.
+
+\end{proof}
+
+\begin{theorem} Subject reduction
+
+  If $\Gamma \vdash M : A$ and $M \twoheadrightarrow_r N$, then $\Gamma \vdash N : A$
+
+\end{theorem}
+
+\begin{proof}
+
+Induction on the derivation $\Gamma \vdash M : A$ and on the generation of $\rightarrow_r$.
+
+For cases with axiom, application, abstraction and pairs see \cite{Morten} \cite{Pierce}.
+
+1) If $\Gamma \vdash {\bf let \: box \:} \vec{x}, y, \vec{z} = \vec{M}, {\bf let \: box \: } \vec{w} = \vec{N} {\bf \: in \: } Q, \vec{P}  {\: \bf in \:} R : \Box B$,
+  then $\Gamma \vdash {\bf let \: box \:} \vec{x}, \vec{w}, \vec{z} = \vec{M}, \vec{N}, \vec{P} \: { \bf in } \: R [y := Q] : \Box B$ by rule 4).
+
+2) Let $\Gamma \vdash {\bf let \: box \:} x = M {\: \bf in \: } x : \Box A$, then $\Gamma \vdash M : \Box A$ by rule 9).
+
+See \cite{ModalK}.
 
@@
 \begin{prooftree}
-  \AxiomC{$\Gamma, z : C \vdash \vec{M} : {\bf K} \vec{A}$}
+\AxiomC{$\Gamma \vdash {\bf box \:} \vec{M} : \Box \vec{A}$}
 \AxiomC{$\vec{x} : \vec{A} \vdash N : B$}
-  \RightLabel{$\text{let}_{{\bf K}}$}
-  \BinaryInfC{$\Gamma, z : C \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } N : {\bf K}B$}
-\end{prooftree}
-
-Let $\Gamma \vdash P : C$.
-
-By IH, $\Gamma \vdash \vec{M} [z := P] : {\bf K} \vec{A}$. So $\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} [z := P] {\: \bf in \: } N : {\bf K}B$
-
-  \vspace{\baselineskip}
-
-4-5) Similary.
-
-\end{proof}
-
-\begin{theorem} Subject reduction
-
-  If $\Gamma \vdash M : A$ and $M \twoheadrightarrow_{\beta \eta} N$, then $\Gamma \vdash N : A$
-
+\BinaryInfC{$\Gamma \vdash {\bf let \: box \:} \vec{x} = {\bf box \:} \vec{M} {\: \bf in \:} N : \Box B$}
+\end{prooftree}
+
+So $\Gamma \vdash \vec{M} : \vec{A}$ by Lemma 4 and $\Gamma \vdash N [\vec{x} := \vec{M}] : B$ by Lemma 4, part 3.
+
+Then we can transform this into the following derivation:
+
+\begin{prooftree}
+\AxiomC{$\Gamma \vdash N [\vec{x} := \vec{M}] : B$}
+\RightLabel{$\Box_I$}
+\UnaryInfC{$\Gamma \vdash {\bf box \:} N [\vec{x} := \vec{M}] : \Box B$}
+\end{prooftree}
+
+4) The derivation ends in
+
+\begin{prooftree}
+\AxiomC{$\vdash M : A$}
+\UnaryInfC{$\vdash {\bf let \: box \:} \underline{\quad} = \underline {\quad} {\: \bf in \:} M : \Box A$}
+\end{prooftree}
+
+So, if $\vdash M : A$, then $\vdash {\bf box \:} M : \Box A$.
+
+\end{proof}
+
+\begin{theorem}
+  $ $
+
+  $\twoheadrightarrow_r$ is strongly normalizing;
 \end{theorem}
@@
 \begin{proof}
-
-Induction on the derivation $\Gamma \vdash M : A$ and on the generation of $\rightarrow_{\beta \eta}$.
-
-For cases with application, abstraction and pairs see \cite{Morten} \cite{Pierce}.
-
-1) If $\Gamma \vdash {\bf let \: pure \:} \vec{x}, y, \vec{z} = \vec{M}, {\bf let \: pure \: } \vec{w} = \vec{N} {\bf \: in \: } Q, \vec{P}  {\: \bf in \:} R : {\bf K}B$,
-  then $\Gamma \vdash {\bf let \: pure \:} \vec{x}, \vec{w}, \vec{z} = \vec{M}, \vec{N}, \vec{P} \: { \bf in } \: R [y := Q] : {\bf K}B$ by rule 4).
-
-2) Let $\Gamma \vdash {\bf let \: pure \:} x = M {\: \bf in \: } x : {\bf K}A$, then $\Gamma \vdash M : {\bf K}A$ by rule 9).
-
-See \cite{ModalK}.
-
-3) The derivation ends in
-
-\begin{prooftree}
-\AxiomC{$\Gamma \vdash {\bf pure \:} \vec{M} : {\bf K}\vec{A}$}
-\AxiomC{$\vec{x} : \vec{A} \vdash N : B$}
-\BinaryInfC{$\Gamma \vdash {\bf let \: pure \:} \vec{x} = {\bf pure \:} \vec{M} {\: \bf in \:} N : {\bf K}B$}
-\end{prooftree}
-
-So $\Gamma \vdash \vec{M} : \vec{A}$ by Lemma 4 and $\Gamma \vdash N [\vec{x} := \vec{M}] : B$ by Lemma 4, part 3.
-
-Then we can transform this into the following derivation:
-
-\begin{prooftree}
-\AxiomC{$\Gamma \vdash N [\vec{x} := \vec{M}] : B$}
-\RightLabel{${\bf K}_I$}
-\UnaryInfC{$\Gamma \vdash {\bf pure \:} N [\vec{x} := \vec{M}] : {\bf K}B$}
-\end{prooftree}
-
-4) The derivation ends in
-
-\begin{prooftree}
-\AxiomC{$\vdash M : A$}
-\UnaryInfC{$\vdash {\bf let \: pure \:} \underline{\quad} = \underline {\quad} {\: \bf in \:} M : {\bf K} A$}
-\end{prooftree}
-
-So, if $\vdash M : A$, then $\vdash {\bf pure \:} M : {\bf K}A$.
-
-Note that this part of the lemma works conversly too.
+  $ $
+
+  Let us define the transformation from $\lambda_{\text{IEL}^{-}}$ into the simple type theory with
+  $\to$, $\times$ and natural number type $\mathbb{N}$ \footnote{Strong normalization for stronger system was shown here \cite{Girard}} with additional typing and reduction rules:
+
+  \begin{prooftree}
+  \AxiomC{$ $}
+  \UnaryInfC{$\Gamma \vdash 0 : \mathbb{N}$}
+  \end{prooftree}
+
+  \begin{prooftree}
+  \AxiomC{$\Gamma \vdash n : \mathbb{N}$}
+  \UnaryInfC{$\Gamma \vdash {\bf succ \:} n : \mathbb{N}$}
+  \end{prooftree}
+
+  \begin{prooftree}
+  \AxiomC{$\Gamma \vdash n : \mathbb{N}$}
+  \AxiomC{$\Gamma \vdash m : \mathbb{N}$}
+  \BinaryInfC{$\Gamma \vdash n + m : \mathbb{N}$}
+  \end{prooftree}
+
+  \begin{enumerate}
+    \item $n + 0 \rightarrow_{\beta} n$;
+    \item $(n + {\bf succ \:} m) \rightarrow_{\beta} {\bf succ \:} (n + m)$
+  \end{enumerate}
+
+  The transformation $|.|$ is defined as follows:
+
+  \begin{defin} Interpretaion of types:
+
+    \begin{enumerate}
+      \item $A \in \mathbb{T} \Rightarrow |A| = A$;
+      \item $|A \to B| = |A| \to |B|$;
+      \item $|A \times B| = |A| \times |B|$;
+      \item $|\Box A| = \mathbb{N} \times |A|$.
+    \end{enumerate}
+  \end{defin}
+
+  \begin{defin} Interpretation of terms:
+    \begin{enumerate}
+      \item $x \in \mathbb{V} \Rightarrow |x| = x$;
+      \item $|\lambda x. M| = \lambda x. |M|$;
+      \item $|(M N)| = |M| |N|$;
+      \item $|\langle M, N \rangle| = \langle |M|, |N| \rangle$;
+      \item $|\pi_i M| = \pi_i |M|$, $i \in \{ 1, 2\}$;
+      \item $|{\bf box \:} M| = \langle 0, |M| \rangle$;
+      \item $|{\bf let \: box \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} M| = \langle 0, |M| \rangle$;
+      \item $|{\bf let \: box \:} x = N {\: \bf in \:} M| = \langle \pi_1 |N| , |M| [x := \pi_2 |N|] \rangle$
+      \item $|{\bf let \: box \:} \vec{x} = \vec{N} {\: \bf in \:} M| = \langle \sum \limits_{i = 1}^n \pi_1 |N|, |M| [\vec{x} := \pi_2 \vec{N}] \rangle$
+    \end{enumerate}
+  \end{defin}
+
+  Let us consider the interpretation for $\text{let}_{\Box}$ rule:
+
+  \begin{prooftree}
+    \AxiomC{$|\Gamma \vdash \vec{N} : \Box \vec{A}| = |\Gamma| \vdash \vec{|N|} : \mathbb{N} \times \vec{|A|}$}
+    \AxiomC{$|\vec{x} : \vec{A} \vdash M : B| = \vec{x} : \vec{|A|} \vdash |M| : |B|$}
+    \RightLabel{$\text{let}_{\Box}$}
+    \BinaryInfC{$|\Gamma \vdash {\bf let \: box \:} \vec{x} = \vec{N} {\: \bf in \: } M : \Box B| = |\Gamma| \vdash \langle \sum \limits_{i = 1}^n \pi_1 |N|, |M| [\vec{x} := \pi_2 \vec{N}] \rangle : \mathbb{N} \times |B|$}
+  \end{prooftree}
+
+  \begin{lemma}:
+
+    $|M [x := N]| = |M| [x := |N|]$ for any term $M$.
+  \end{lemma}
+
+  \begin{proof}
+
+    Induction on the structure of $M$.
+  \end{proof}
+
+  \begin{lemma}
+
+    $M \twoheadrightarrow_{r} N \Rightarrow |M| \twoheadrightarrow_{\beta\eta} |N|$
+  \end{lemma}
+
+  \begin{proof}
+
+    Let us consider cases with $\beta \Box$, $\beta \Box {\bf box}$ and $\Box id$.
+
+\begin{enumerate}
+\item $|{\bf let \: box \:} x = ({\bf let \: box \:} y = N {\: \bf in \:} P) {\: \bf in \:} M| = \\
+\quad\quad\quad\quad |{\bf let \: box \:} y = N {\: \bf in \:} M [x := P]|$
+
+  $\begin{array}{lll}
+  &|{\bf let \: box \:} x = ({\bf let \: box \:} y = N {\: \bf in \:} P) {\: \bf in \:} M| = & \\
+  &\quad\quad\quad\quad\quad\quad\quad \text{Interpretation}& \\
+  &\langle \pi_1 |N|, |M| [x := |P| [y := \pi_2 |N|]] \rangle& \\
+  &|{\bf let \: box \:} y = N {\: \bf in \:} M [x := P]| = & \\
+  &\langle \pi_1 |N|, |M| [x := |P|] [y := \pi_2 |N|]\rangle \equiv & \\
+  &\quad\quad\quad\quad\quad\quad\quad \text{So far as $y \notin FV(M)$} & \\
+  &\langle \pi_1 |N|, |M| [x := |P| [y := \pi_2 |N|]]&
+  \end{array}$
+
+\item $|{\bf let \: box \:} \vec{x} = {\bf box \:} \vec{N} {\: \bf in \:} M| = |{\bf box \:} M [\vec{x} := \vec{N}]|$\\
+
+  $\begin{array}{lll}
+  &|{\bf let \: box \:} \vec{x} = {\bf box \:} \vec{N} {\: \bf in \:} M| = & \\
+  &\quad\quad\quad\quad\quad\quad\quad \text{Interpretation}& \\
+  &\langle 0 + \dots + 0, |M| [\vec{x} := \vec{|N|}]\rangle \twoheadrightarrow_{\beta}& \\
+  &\quad\quad\quad\quad\quad\quad\quad \text{Multistep reduction for natural numbers}& \\
+  &\langle 0, |M| [\vec{x} := \vec{|N|}] \rangle = & \\
+  &\quad\quad\quad\quad\quad\quad\quad \text{Interpretation}& \\
+  &|{\bf box \:} M [\vec{x} := \vec{N}]|&
+  \end{array}$
+
+\item $|{\bf let \: box \:} x = M {\: \bf in \:} x| = |M|$\\
+
+  $\begin{array}{lll}
+  &|{\bf let \: box \:} x = M {\: \bf in \:} x| = & \\
+  &\quad\quad\quad\quad\quad\quad\quad \text{Interpretation}& \\
+  &\langle \pi_1 |M|, x [x := \pi_2 |M|] \rangle = & \\
+  &\quad\quad\quad\quad\quad\quad\quad \text{Substitution}& \\
+  &\langle \pi_1 |M|, \pi_2 |M| \rangle \rightarrow_{\eta}& \\
+  &\quad\quad\quad\quad\quad\quad\quad \text{$\eta$-reduction for pairs}& \\
+  &|M|&
+  \end{array}$
+
+\item $\begin{array}{lll}
+  &|{\bf let \: box \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} M| = \langle 0 , |M| \rangle = |{\bf box \:} M|&
+  \end{array}$
+\end{enumerate}
+
+  Hence $\lambda_{\text{IEL}^{-}}$ sounds for $\lambda_{\to, \times, \mathbb{N}}$,
+  then multistep reduction in $\lambda_{\text{IEL}^{-}}$ is strongly normalizing, so far as multistep reduction in $\lambda_{\to, \times, \mathbb{N}}$ is strongly normalizing.
+
+  \end{proof}
 \end{proof}
@@
 
-  $\twoheadrightarrow_{\beta}$ is strongly normalizing;
+  $\twoheadrightarrow_r$ is confluent.
 \end{theorem}
@@
 \begin{proof}
-  $ $
-
-We modify and apply Tait's technique of logical relation for modalities. See \cite{Pierce} \cite{Troelstra}.
-
-We treat only modal cases below.
-
-\begin{defin} The set of strongly computable terms:
-  \begin{itemize}
-    \item $SC_A = \{ M : A \: | \: M \text{ is strongly normalizing} \}$ for $A \in \mathbb{T}$;
-    \item $SC_{A \to B} = \{ M : A \to B \: | \: \forall N \in SC_A, M N \in SC_B \}$, for $A,B \in \mathbb{T}_{\bf K}$ for $A, B \in \mathbb{T}_{{\bf K}}$;
-    \item $SC_{{\bf K}A} = \{ M : {\bf K}A \: | \: M \text{ is strongly normalizing} \}$ for $A \in \mathbb{T}$;
-    \item $\forall i \in \{ 1,\dots,n\}, \prod \limits_{i = 1}^{n} SC_{{\bf K}A_i} = \{ \vec{M} = (M_1,\dots,M_n) \:\: | \:\: \forall N \in SC_{B}, FV(N) = \{ x_1,\dots,x_n\} \: \& \: \forall i, x_i \in SC_{A_i} \Rightarrow {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N \in SC_{{\bf K} B}\}$
-  \end{itemize}
-\end{defin}
-
-\begin{defin}
-  A term $M$ is neutral, if it has one of the following forms:
-  \begin{itemize}
-    \item $M N$;
-    \item If $M$ is neutral, then ${\bf pure \:} M$ is neutral;
-    \item If $\vec{M}$ and $N$ are neutral, then ${\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: N}$ is neutral.
-    $\vec{x}$ is a sequence of free variables of a term $N$.
-  \end{itemize}
-\end{defin}
-
-\begin{lemma}
-  $ $
-
-\begin{itemize}
-\item If $M \in SC_A$ and $A \in \mathbb{T}_{{\bf K}}$, then $M$ is strongly normalizing;
-\item If $M \in SC_A$, $A \in \mathbb{T}_{{\bf K}}$ and $M \rightarrow_{\beta} N$, then $N \in SC_A$;
-\item If $N$ is neutral, $N \in SC_A$. Then, if $M \rightarrow_{\beta} N$, then $M \in SC_A$;
-\end{itemize}
-\end{lemma}
-
-\begin{proof}
-  $ $
-
-  By induction on the structure of $A$.
-
-1) $A \equiv {\bf K}A$, where $A \in \mathbb{T}$.
-
-i-ii-iii) Immediately.
-
-    \vspace{\baselineskip}
-
-2)
-
-i) Suppose $\vec{M} = (M_1,\dots,M_n) \in \prod \limits_{i = 1}^{n} SC_{{\bf K}A_i}$.
-
-Let $N \in SC_{B}$, such that $FV(N) = \{ x_1,\dots, x_n \}$ and $\forall i, x_i \in SC_{A_i}$.
-
-So ${\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } N \in SC_{{\bf K}B}$ by IH.
-
-So $\vec{M}$ are strongly normalizing, since ${\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } N$ is strongly normalizing by IH.
-
-\vspace{\baselineskip}
-
-ii) Let $\vec{M_1} \in \prod \limits_{i = 1}^{n} SC_{{\bf K}A_i}$ and $\vec{M_1} \rightarrow_{\beta} \vec{M_2}$.
-
-Let $N \in SC_{B}$, such that $FV(N) = \{ x_1,\dots,x_n \}$ and $\forall i, x_i \in SC_{A_i}$.
-
-So ${\bf let \: pure \:} \vec{x} = \vec{M_1} {\: \bf in \:} N \rightarrow_{\beta} {\bf let \: pure \:} \vec{x} = \vec{M_2} {\: \bf in \:} N$
-
-and ${\bf let \: pure \:} \vec{x} = \vec{M_2} {\: \bf in \:} N \in SC_{{\bf K}B}$ by assumption.
-
-So $\vec{M_2} \in \prod \limits_{i = 1}^{n} SC_{{\bf K}A_i}$.
-
-iii) Let $M_2$ be neutral, $M_2 \in \prod \limits_{i = 1}^{n} SC_{{\bf K}A_i}$ and $M_1 \rightarrow_{\beta} M_2$.
-
-Let $N \in SC_{B}$, , such that $FV(N) = \{ x_1,\dots,x_n \}$ and $\forall i, x_i \in SC_{A_i}$.
-
-So ${\bf let \: pure \:} \vec{x} = \vec{M_2} {\: \bf in \:} N \in SC_{{\bf K}B}$.
-
-Thus ${\bf let \: pure \:} \vec{x} = \vec{M_1} {\: \bf in \:} N \rightarrow_{\beta} {\bf let \: pure \:} \vec{x} = \vec{M_2} {\: \bf in \:} \in N$.
-
-Hence ${\bf let \: pure \:} \vec{x} = \vec{M_1} {\: \bf in \:} N\in SC_{{\bf K}B}$ by IH, so $\vec{M_1} \in \prod \limits_{i = 1}^{n} SC_{{\bf K}A_i}$.
-\end{proof}
-
-\begin{lemma}
-  $ $
-
-  If $M \in SC_A$, then ${\bf pure \:} M \in SC_{{\bf K}A}$
-\end{lemma}
-
-\begin{proof}
-
-  Induction on the structure of M.
-\end{proof}
-
-\begin{lemma}
-  $ $
-
-  Let $x_1 : A_1,\dots, x_n : A_n \vdash M : A$, then forall $i, M_i \in SC_{A_i}$. Then $M [x_1 := M_1,\dots,x_n := M_n] \in SC_A$.
-\end{lemma}
-
-\begin{proof}
-  $ $
-
-Induction on the derivation of $x_1 : A_1,\dots, x_n : A_n \vdash M : A$.
-
-  1) The derivation ends in:
-
-\begin{prooftree}
-\AxiomC{$x_1 : A_1,\dots, x_n : A_n \vdash M : A$}
-\UnaryInfC{$x_1 : A_1,\dots, x_n : A_n \vdash {\bf pure \:} M : {\bf K}A$}
-\end{prooftree}
-
-  By assumption $M [x_1 := M_1,\dots,x_n := M_n] \in SC_A$, so ${\bf pure \:} M [x_1 := M_1,\dots,x_n := M_n] \in SC_{{\bf K}A}$.
-
-\vspace{\baselineskip}
-
-  2) The derivation ends in:
-\begin{prooftree}
-  \AxiomC{$x_1 : A_1,\dots, x_n : A_n \vdash \vec{M^{'}} : {\bf K}\vec{A}$}
-  \AxiomC{$\vec{x} : \vec{A} \vdash N : B$}
-  \BinaryInfC{$x_1 : A_1,\dots, x_n : A_n \vdash {\bf let \: pure \:} \vec{x} = \vec{M^{'}} {\: \bf in \:} N : {\bf K}B$}
-\end{prooftree}
-
-By IH forall $i \in \{ 1,\dots,\text{length}(\vec{M^{'}}) \}$, $M_i^{'} [x_1 := M_1,\dots,x_n := M_n] \in SC_{{\bf K}A_i}$.
-
-So ${\bf let \: pure \:} \vec{x} = \vec{M^{'}} [x_1 := M_1,\dots,x_n := M_n] {\: \bf in \:} N \in SC_{{\bf K} B}$,
-otherwise we can build infinite reduction path in $\vec{M^{'}} [x_1 := M_1,\dots,x_n := M_n]$.
-\end{proof}
-
-\begin{col}
-
-  All terms are strongly computable, therefore are strongly normalizing.
-\end{col}
+
+  By Newman's lemma \cite{Morten}, if relation is strongly normalizing and locally conluent, then this relation is confluent.
+
+It is sufficient to show that $\twoheadrightarrow_r$ is locally confluent.
+
+\begin{lemma} Local confluence
+
+If $M \rightarrow_{r} N$ and $M \rightarrow_{r} Q$, then there exists some term $P$,
+such that $N \twoheadrightarrow_{r} P$ and $Q \twoheadrightarrow_{r} P$.
+
+\end{lemma}
+
+\begin{proof}
+
+Let us consider the following critical pairs and show that they are joinable:
+
+\begin{enumerate}
+\item $ $ \\
+
+\xymatrix{
+{\bf let \: box \:} x = ({\bf let \: box \:} \vec{y} = {\bf box \:} \vec{N} {\: \bf in \:} P) {\: \bf in \:} M \ar[d]_{\beta \Box} \ar[dr]^{\beta \Box {\bf box}} \\
+{\bf let \: box \:} \vec{y} = {\bf box \:} \vec{N} {\: \bf in \:} M [x := P] & {\bf let \: box \:} x = {\bf box \:} P [\vec{y} := \vec{N}] {\: \bf in \:} M
+}
+
+\vspace{\baselineskip}
+
+$\begin{array}{lll}
+&{\bf let \: box \:} \vec{y} = {\bf box \:} \vec{N} {\: \bf in \:} M [x := P] \rightarrow_{\beta \Box {\bf box}}& \\
+&\quad\quad\quad\quad\quad\quad\quad {\bf box \:} M [x := P] [\vec{y} := \vec{N}]& \\
+&{\bf let \: box \:} x = {\bf box \:} P [\vec{y} := \vec{N}] {\: \bf in \:} M \rightarrow_{\beta \Box {\bf box}}& \\
+&\quad\quad\quad\quad\quad\quad\quad {\bf box \:} M [x := P[\vec{y} := \vec{N}]] \equiv & \\
+&\text{So far as $x \notin \vec{y}$}& \\
+&\quad\quad\quad\quad\quad\quad\quad {\bf box \:} M [x := P] [\vec{y} := \vec{N}] & \\
+\end{array}$
+
+\item $ $ \\
+
+\xymatrix{
+{\bf let \: box \:} x = ({\bf let \: box \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} N) {\: \bf in \:} M \ar[d]^{\beta \Box} \ar[dr]^{\beta \Box \underline{\quad}} \\
+{\bf let \: box \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} M [x := N] & {\bf let \: box \:} x = {\bf box \:} N {\: \bf in \:} M
+}
+
+$\begin{array}{lll}
+&{\bf let \: box \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} M [x := N] \rightarrow_{\beta \Box \underline{\quad}} {\bf box \:} (M [x := N])& \\
+&{\bf let \: box \:} x = {\bf box \:} N {\: \bf in \:} M \rightarrow_{\beta \Box {\bf box}} {\bf box \:} (M [x := N])&
+\end{array}$
+\end{enumerate}
+\end{proof}
+
+Also we may consider four critical pairs which are used in local confluence proof for lambda-calculus based on {\bf IK} and \cite{ModalK}.
 
@@
 
-  $\twoheadrightarrow_{\beta}$ is confluent.
+  Normal form in call-by-name $\lambda_{\text{IEL}^{-}}$ has the subformula property: if $M$ is in normal formal, then its all subterms are in normal form too.
+
 \end{theorem}
@@
 \begin{proof}
-
-  We modify and apply Barendregt's technique with term underlying. We will consider the grammar without constructors for pairs.
-
-\begin{defin} The set of underlined terms.
-
-  \begin{itemize}
-    \item $x \in \mathbb{V} \Rightarrow x \in \underline{\Lambda}$;
-    \item $M \in \underline{\Lambda} \Rightarrow (\lambda x. M) \in \underline{\Lambda}$;
-    \item $M, N \in \underline{\Lambda} \Rightarrow (M N) \in \underline{\Lambda}$;
-    \item $M \in \underline{\Lambda} \Rightarrow ({\bf pure \:} M) \in \underline{\Lambda}$;
-    \item $\vec{x} \in \mathbb{V}, \vec{M}, N \in \underline{\Lambda} \Rightarrow {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N \in \underline{\Lambda}$;
-    \item $M, N \in \underline{\Lambda} \Rightarrow (\lambda_i x. M) N \in \underline{\Lambda}$, for all $i \in \mathbb{N}$.
-  \end{itemize}
-\end{defin}
-
-\begin{defin} Subsitution for term with labelled lambda:
-
-  $((\lambda_i x. M) N) [y := Z] = (\lambda_i x. M [y := Z]) (N [y := Z])$
-\end{defin}
-
-\begin{defin} Index erasing
-
-  Let us define map $|.| : \underline{\Lambda} \to \Lambda$ as follows:
-
-  \begin{itemize}
-    \item $|x| = x$;
-    \item $|\lambda x. M| = \lambda x. |M|$;
-    \item $|M N| = |M| |N|$;
-    \item $|{\bf pure \:} M| = {\bf pure \:} |M|$;
-    \item $|{\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N| = {\bf let \: pure \:} \vec{x} = \vec{|M|} {\: \bf in \:} |N|$;
-    \item $|(\lambda_i x. M) N| = (\lambda x. |M|) |N|$
-  \end{itemize}
-\end{defin}
-
-\begin{defin} Reduction rules:
-
-\begin{itemize}
-  \item $(\lambda x. M) N \rightarrow_{\underline{\beta}} M [x := N]$;
-  \item $\begin{array}{llll}
-  & {\bf let \: pure \:} \vec{x}, y, \vec{z} = \vec{M}, {\bf let \: pure \: } \vec{w} = \vec{N} {\bf \: in \: } Q, \vec{P} {\: in \:} R \rightarrow_{\underline{\beta}} & \\
-  & {\bf let \: pure \:} \vec{x}, \vec{w}, \vec{z} = \vec{M}, \vec{N}, \vec{P} \: { \bf in } \: R [y := Q]
-  \end{array}$;
-  \item ${\bf let \: pure \:} \vec{x} = {\bf pure \:} \vec{M} {\: \bf in \:} N \rightarrow_{\underline{\beta}} {\bf pure \:} N [\vec{x} := \vec{M}]$;
-  \item ${\bf let \: pure \:} \underline{\quad} = \underline {\quad} {\: \bf in \:} M \rightarrow_{\underline{\beta}} {\bf pure \:} M$
-  \item $(\lambda x_i. M) N \rightarrow_{\underline{\beta}} M [x := N]$
-\end{itemize}
-\end{defin}
-
-$\twoheadrightarrow_{\underline{\beta}}$ is a reflexive-transitive closure of $\rightarrow_{\underline{\beta}}$.
-
-\begin{defin} Indexed redex erasing:
-
-  Let us define the next map $\phi : \underline{\Lambda} \to \Lambda$:
-
-  \begin{itemize}
-    \item $\phi(x) = x$;
-    \item $\phi(\lambda x. M) = \lambda x. \phi(M)$;
-    \item $\phi(M N) = \phi(M) \phi(N)$;
-    \item $\phi({\bf pure \:} M) = {\bf pure \:} \phi(M)$;
-    \item $\phi({\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N) = {\bf let \: pure \:} \vec{x} = \vec{\phi(M)} {\: \bf in \:} \phi(N)$;
-    \item $\phi((\lambda_i x. M) N) = \phi(M) [x := \phi(N)]$
-  \end{itemize}
-\end{defin}
-
-\begin{lemma}
-
-  $\forall \underline{M}, \underline{N} \in \underline{\Lambda} \: \forall M, N \in \Lambda, \text{if } |\underline{M}| = M, |\underline{N}| = N, \text{then}$
-  \begin{itemize}
-    \item If $M \twoheadrightarrow_{\beta} N$, then $\underline{M} \twoheadrightarrow_{\underline{\beta}} \underline{N}$
-    \item Vice versa
-  \end{itemize}
-\end{lemma}
-
-\begin{proof}
-
-  Induction on the generation $\rightarrow_{\beta}$ and $\rightarrow_{\underline{\beta}}$ correspondently. The general statement follows from transitivity
-  of multi-step reductions of both types.
-\end{proof}
-
-\begin{lemma}
-  $\phi(M [x := N]) = \phi(M) [x := \phi(N)]$.
-\end{lemma}
-
-\begin{proof}
-  We treat only cases with {\bf pure} and with {\bf let}. For the rest cases see \cite{Baren}.
-
-  1)
-  $ $
-
-  $\begin{array}{lll}
-  & \phi({\bf pure \:} (M [x := N])) = & \\
-  &\quad\quad\quad\quad\quad\quad \text{By the definition of $\phi$} & \\
-  & {\bf pure \:} (\phi(M [x := N])) = & \\
-  &\quad\quad\quad\quad\quad\quad \text{Induction hypothesis}& \\
-  & {\bf pure \:} (\phi(M) [x := \phi(N)]) = &\\
-  &\quad\quad\quad\quad\quad\quad \text{Substitution definition}& \\
-  & ({\bf pure \:} \phi (M)) [x := \phi(N)]&
-  \end{array}$
-
-  2)
-  $ $
-
-  $\begin{array}{lll}
-  &\phi(({\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N) [y := P]) = &\\
-  &\quad\quad\quad\quad\quad\quad \text{Substitution definition}& \\
-  &\phi({\bf let \: pure \:} \vec{x} = (\vec{M}[y := P]) {\: \bf in \:} N) = &\\
-  &\quad\quad\quad\quad\quad\quad \text{By the definition of $\phi$}& \\
-  &{\bf let \: pure \:} \vec{x} = \phi(\vec{M}[y := P]) {\: \bf in \:} \phi(N) = &\\
-  &\quad\quad\quad\quad\quad\quad \text{Induction hypothesis}& \\
-  &{\bf let \: pure \:} \vec{x} = (\phi(\vec{M}) [y := \phi(P)]) {\: \bf in \:} \phi(N) =&\\
-  &\quad\quad\quad\quad\quad\quad \text{Substitution definition} & \\
-  &({\bf let \: pure \:} \vec{x} = \phi(\vec{M}) {\: \bf in \:} \phi(N))[y := \phi(P)]
-  \end{array}$
-
-\end{proof}
-
-\begin{lemma}
-  $ $
-
-  \begin{itemize}
-  \item If $M \twoheadrightarrow_{\underline{\beta}} N$, then $\phi(M) \twoheadrightarrow_{\beta} \phi(N)$
-  \item If $|M| = N$ and $\phi(M) = P$, then $N \twoheadrightarrow_{\beta} P$.
-  \end{itemize}
-\end{lemma}
-
-\begin{proof}
-  $ $
-
-  i) Induction on the generation of $\twoheadrightarrow_{\underline{\beta}}$ using previous lemma.
-
-  ii) Induction on the structure of $M$.
-\end{proof}
-
-\begin{lemma} Strip lemma.
-
-$ $
-
-If $M \rightarrow_{\beta} N$ and $M \twoheadrightarrow_{\beta} P$. Then there exists some term $Q$, such that
-$N \twoheadrightarrow_{\beta} Q$ and $P \twoheadrightarrow_{\beta} Q$.
-\end{lemma}
-
-\begin{proof}
-
-Proof is similar to \cite{Baren} \cite{Baren2}. We build the following diagram, which commutes by lemmas 8 and 10.
-
-\vspace{\baselineskip}
-
-\xymatrix{
-&&& M \ar[dd]_{\beta} \ar@{->>}[rrr] &&& P \ar@{-->>}[dd]^{\beta} \\
-&&&& N' \ar[ul]^{|.|} \ar[dl]_{\phi} \ar@{-->>}[rrr]^{\underline{\beta}} &&& P' \ar[ul]^{|.|} \ar[dl]_{\phi} \\
-&&& N \ar@{-->>}[rrr]_{\beta} &&& Q
-}
-\end{proof}
-
-\begin{col}
-$ $
-If $M \twoheadrightarrow_{\beta} N$ and $M \twoheadrightarrow_{\beta} P$. Then there exists some term $Q$, such that
-$N \twoheadrightarrow_{\beta} Q$ and $P \twoheadrightarrow_{\beta} Q$.
-\end{col}
-
-\begin{proof}
-
-  Unfold $M \twoheadrightarrow_{\beta} N$ as the sequence of one-step reductions and apply strip lemma on the every step.
-\end{proof}
-
-\end{proof}
-
-\begin{theorem}
-  $ $
-
-  Normal form in call-by-name $\lambda_{{\bf K}}$ has the subformula property: if $M$ is in normal formal, then its all subterms are in normal form too.
-
-\end{theorem}
-
-\begin{proof}
   By induction on the structure of $M$.
-  Case with ${\bf let \: pure\:} \vec{x} = \vec{M} {\: \bf in \:} N$ was considered by Kakutani \cite{ModalK} \cite{ModalK1}.
-
-  If ${\bf pure \:} M$ is in normal form, so $M$ is in normal form and its subterms are in normal form too by hypothesis.
-
-  Thus if ${\bf pure \:} M$ is in normal form, then all its subterms are in normal form too.
+  Case with ${\bf let \: box \:} \vec{x} = \vec{M} {\: \bf in \:} N$ was considered by Kakutani \cite{ModalK} \cite{ModalK1}.
+
+  If ${\bf box \:} M$ is in normal form, so $M$ is in normal form and its subterms are in normal form too by hypothesis.
+
+  Thus if ${\bf box \:} M$ is in normal form, then all its subterms are in normal form too.
 \end{proof}
@@
   Monoidal category is a category $\mathcal{C}$ with additional monoidal structure:
-  \begin{itemize}
+  \begin{enumerate}
     \item A bifunctor $\otimes : \mathcal{C} \times \mathcal{C} \to C$ called tensor product;
@@
     }
-
-    \item A monoidal category $\mathcal{C}$ is called symmetric if forall $A,B \in Ob_{\mathcal{C}}$, there is an isomorphism $\sigma_{A,B} : A \otimes B \cong B \otimes A$.
-  \end{itemize}
+  \end{enumerate}
 \end{defin}
@@
 
-Note that, any cartesian closed category is the special case of a (symmetric) monoidal category, where tensor is a product and identity object is a terminal object.
-
-\begin{defin} Lax monoidal functor
+Note that, any cartesian closed category is the special case of a monoidal category, where tensor is a product and identity object is a terminal object.
+
+\begin{defin} Monoidal functor
 
@@
 
-  A lax monoidal functor $\mathcal{F} : \langle \mathcal{C}, \otimes_1, \mathds{1} \rangle \to \langle \mathcal{D}, \otimes_2, \mathds{1}' \rangle$ is a functor
+  A monoidal functor $\mathcal{F} : \langle \mathcal{C}, \otimes_1, \mathds{1} \rangle \to \langle \mathcal{D}, \otimes_2, \mathds{1}' \rangle$ is a functor
   $\mathcal{F} : \mathcal{C} \to \mathcal{D}$ with additional natural transformations:
 
-  \begin{itemize}
+  \begin{enumerate}
   \item $u : \mathds{1}_{\mathcal{D}} \to \mathcal{F}\mathds{1}_{\mathcal{C}}$;
-  \item $\ast_{A, B} : \mathcal{F}A \otimes_{\mathcal{D}} \mathcal{F}B \to \mathcal{F}(A \otimes_{\mathcal{C}} B)$.
-  \end{itemize}
+  \item $A \varoast B : \mathcal{F}A \otimes_{\mathcal{D}} \mathcal{F}B \to \mathcal{F}(A \otimes_{\mathcal{C}} B)$.
+\end{enumerate}
 
@@
   \xymatrix{
-    && (\mathcal{F}A \otimes_{\mathcal{D}} \mathcal{F}B) \otimes_{\mathcal{D}} \mathcal{F}C \ar[d]_{\ast_{A,B} \otimes_{\mathcal{D}} id_{\mathcal{F}B}}
+    && (\mathcal{F}A \otimes_{\mathcal{D}} \mathcal{F}B) \otimes_{\mathcal{D}} \mathcal{F}C \ar[d]_{(A \varoast B) \otimes_{\mathcal{D}} id_{\mathcal{F}B}}
     \ar[rr]^{\alpha^{\mathcal{D}}_{{\mathcal{F}A, \mathcal{F}B, \mathcal{F}C}}} && \mathcal{F}A \otimes_{\mathcal{D}} (\mathcal{F}B \otimes_{\mathcal{D}} \mathcal{F}C) \ar[d]^{id_{\mathcal{F}A} \otimes_{\mathcal{D}} \ast_{B,C}}\\
-    && \mathcal{F}(A \otimes_{\mathcal{C}} B) \otimes_{\mathcal{D}} \mathcal{C} \ar[d]_{\ast_{A \otimes_{\mathcal{C}} B, C}} && \mathcal{F}A \otimes_{\mathcal{D}} \mathcal{F}(B \otimes_{\mathcal{C}} C) \ar[d]^{\ast_{A, B \otimes_{\mathcal{C}} C}}\\
+    && \mathcal{F}(A \otimes_{\mathcal{C}} B) \otimes_{\mathcal{D}} \mathcal{C} \ar[d]_{(A \otimes_{\mathcal{C}} B) \varoast C} && \mathcal{F}A \otimes_{\mathcal{D}} \mathcal{F}(B \otimes_{\mathcal{C}} C) \ar[d]^{A \varoast (B \otimes_{\mathcal{C}} C) }\\
     && \mathcal{F}((A \otimes_{\mathcal{C}} B) \otimes_{\mathcal{C}} C) \ar[rr]_{\mathcal{F}(\alpha^{\mathcal{C}}_{A,B,C})}&& \mathcal{F}(A \otimes_{\mathcal{C}} (B \otimes_{\mathcal{C}} C))
@@
     \xymatrix{
-    &&  \mathds{1}_{\mathcal{D}} \otimes_{\mathcal{D}} \mathcal{F}A \ar[d]_{L^{\mathcal{D}}_{\mathcal{F}A}}\ar[rr]^{u \otimes_{\mathcal{D}} id_{\mathcal{F}A}} && \mathcal{F}\mathds{1}_{\mathcal{C}} \otimes_{\mathcal{D}} \mathcal{F}A \ar[d]^{\ast_{\mathds{1}_{\mathcal{C}}, A}} \\
+    &&  \mathds{1}_{\mathcal{D}} \otimes_{\mathcal{D}} \mathcal{F}A \ar[d]_{L^{\mathcal{D}}_{\mathcal{F}A}}\ar[rr]^{u \otimes_{\mathcal{D}} id_{\mathcal{F}A}} && \mathcal{F}\mathds{1}_{\mathcal{C}} \otimes_{\mathcal{D}} \mathcal{F}A \ar[d]^{\mathds{1}_{\mathcal{C}} \varoast A} \\
     &&  \mathcal{F}A && \mathcal{F}(\mathds{1}_{\mathcal{C}} \otimes_{\mathcal{C}} A) \ar[ll]_{\mathcal{F}(L^{\mathcal{C}}_A)}
@@
     \xymatrix{
-    &&  \mathcal{F}A \otimes_{\mathcal{D}} \mathds{1}_{\mathcal{D}} \ar[d]_{R^{\mathcal{D}}_{\mathcal{F}A}}\ar[rr]^{id_{\mathcal{F}A} \otimes_{\mathcal{D}} u} && \mathcal{F}A \otimes_{\mathcal{D}} \mathcal{F}\mathds{1}_{\mathcal{C}} \ar[d]^{\ast_{A, \mathds{1}_{\mathcal{C}}}} \\
+    &&  \mathcal{F}A \otimes_{\mathcal{D}} \mathds{1}_{\mathcal{D}} \ar[d]_{R^{\mathcal{D}}_{\mathcal{F}A}}\ar[rr]^{id_{\mathcal{F}A} \otimes_{\mathcal{D}} u} && \mathcal{F}A \otimes_{\mathcal{D}} \mathcal{F}\mathds{1}_{\mathcal{C}} \ar[d]^{A \varoast \mathds{1}_{\mathcal{C}}} \\
     &&  \mathcal{F}A && \mathcal{F}(A \otimes_{\mathcal{C}} \mathds{1}_{\mathcal{C}} ) \ar[ll]^{\mathcal{F}(R^{\mathcal{C}}_A)}
@@
 
-\begin{defin}
-
-Strong functor on a monoidal category is an endofunctor equipped with a natural transformation so called tensorial strength:
-
-\begin{center}
-$\begin{array}{lll}
-  \tau_{A, B} : A \otimes \mathcal{K}B \rightarrow \mathcal{K}(A \otimes B)
-\end{array}$
-\end{center}
-
-making two diagrams commute:
-
-\xymatrix{
-(A \otimes B) \otimes \mathcal{K}C \ar[d]_{\alpha_{A, B, \mathcal{K}C}}\ar[rrrr]^{\tau_{A \otimes B, C}} &&&& \mathcal{K}((A \otimes B) \otimes C) \ar[d]^{\mathcal{K}(\alpha_{A,B,C})}\\
-A \otimes (B \otimes \mathcal{K}C) \ar[rr]_{id_A \otimes \tau_{B,C}} && A \otimes \mathcal{K}(B \otimes C) \ar[rr]_{\quad \tau_{A, (B \otimes C)}} && \mathcal{K}(A \otimes (B \otimes C)) \\
-}
-
-\xymatrix{
-&&& \mathds{1} \otimes \mathcal{K}A \ar[drr]_{R_{\mathcal{K}A}} \ar[rr]^{\mu_{\mathds{1}, A}} && \mathcal{K}(\mathds{1} \otimes A) \ar[d]^{\mathcal{K}(R_A)}\\
-&&&&& \mathcal{K}A
-}
-\end{defin}
-
 \begin{defin} Applicative functor
 
-  An applicative functor is a triple $\langle \mathcal{C}, \mathcal{K}, \eta \rangle$,
-where $\mathcal{C}$ is a monoidal category, $\mathcal{K}$ is a strong lax monoidal endofunctor and $\eta : Id_{\mathcal{C}} \Rightarrow \mathcal{K}$ is a natural transformation (similar to unit in monad), such that:
-
-\begin{itemize}
+  An applicative functor is a triple $\langle \mathcal{C}, \mathcal{F}, \eta \rangle$,
+where $\mathcal{C}$ is a monoidal category, $\mathcal{F}$ is a monoidal endofunctor and $\eta : Id_{\mathcal{C}} \Rightarrow \mathcal{F}$ is a natural transformation (similar to unit in monad), such that:
+
+\begin{enumerate}
 \item $u = \eta_{\mathds{1}}$;
-\item $\ast_{A,B} \circ (\eta_A \otimes \eta_B) = \eta_{A \otimes B}$, i.e. the following diagram commutes:
+\item $A \varoast B \circ (\eta_A \otimes \eta_B) = \eta_{A \otimes B}$, i.e. the following diagram commutes:
 
@@
 {
-&&& A \otimes B \ar[rr]^{\eta_A \otimes \eta_B} \ar[drr]_{\eta_{A \otimes B}} && \mathcal{K}A \otimes \mathcal{K}B \ar[d]^{\ast_{A,B}} \\
-&&&&& \mathcal{K}(A \otimes B)
+&&& A \otimes B \ar[rr]^{\eta_A \otimes \eta_B} \ar[drr]_{\eta_{A \otimes B}} && \mathcal{F}A \otimes \mathcal{F}B \ar[d]^{A \varoast B} \\
+&&&&& \mathcal{F}(A \otimes B)
 }
-\item $\tau_{A, B} = \ast_{A, B} \circ \eta_{A} \otimes id_{\mathcal{K}B}$.
-\end{itemize}
-\end{defin}
-
-By default we will consider some strong lax monoidal functor on cartesian closed category.
-
-Note that lax monoidal functor on a cartesian closed category is a special case of an applicative functor \cite{Cons}, but
-we will define $\eta$ explicitly for simplicity.
+\end{enumerate}
+\end{defin}
+
+By default we will consider some monoidal functor on cartesian closed category below.
 
@@
 
-\begin{theorem} Soundness
-
-  Let $\Gamma \vdash M : A$ and $M =_{\beta\eta} N$, then $[\![\Gamma \vdash M : A]\!] = [\![\Gamma \vdash N : A]\!]$
-\end{theorem}
-
-\begin{proof}
-
-\begin{defin} Semantical translation from $\lambda_{{\bf K}}$ to some cartesian closed category $\mathcal{C}$ with an applicative functor $\mathcal{K}$:
-
-\begin{itemize}
+\begin{defin} Semantical translation from $\lambda_{\text{IEL}^{-}}$ to some cartesian closed category $\mathcal{C}$ with an applicative functor $\boxdot$:
+
+\begin{enumerate}
 \item Interpretation for types:
@@
 \item Interpretation for modal types:
-  \begin{itemize}
-    \item $[\![{\bf K}A]\!] = \mathcal{K}[\![A]\!]$;
-  \end{itemize}
+  \begin{enumerate}
+    \item $[\![\Box A]\!] = \boxdot [\![A]\!]$;
+  \end{enumerate}
 \item Interpretaion for contexts:
-  \begin{itemize}
+  \begin{enumerate}
     \item $[\![ \quad ]\!] = \mathds{1}$, where $\mathds{1}$ is a terminal object of a given ccc;
     \item $[\![\Gamma, x : A]\!] = [\![\Gamma]\!] \times [\![A]\!]$
-  \end{itemize}
+  \end{enumerate}
 \item Interpretation for typing assignment:
-  \begin{itemize}
+  \begin{enumerate}
     \item $[\![\Gamma \vdash M : A]\!] := [\![M]\!] : [\![\Gamma]\!] \to [\![A]\!]$.
-  \end{itemize}
+  \end{enumerate}
 \item Interpretation for typing rules:
@@
 \AxiomC{$[\![\Gamma \vdash M : A]\!] = [\![M]\!] : [\![\Gamma]\!] \rightarrow [\![A]\!]$}
-\UnaryInfC{$[\![\Gamma \vdash {\bf pure \:} M : {\bf K}A]\!] := [\![\Gamma]\!] \xrightarrow{[\![M]\!]}
-[\![A]\!] \xrightarrow{\eta_{[\![A]\!]}} \mathcal{K}[\![A]\!]$}
+\UnaryInfC{$[\![\Gamma \vdash {\bf box \:} M : \Box A]\!] := [\![\Gamma]\!] \xrightarrow{[\![M]\!]}
+[\![A]\!] \xrightarrow{\eta_{[\![A]\!]}} \boxdot [\![A]\!]$}
 \end{prooftree}
@@
   \begin{prooftree}
-    \AxiomC{$[\![\Gamma \vdash \vec{M} : {\bf K} \vec{A}]\!] = \langle [\![M_1]\!],\dots, [\![M_n]\!] \rangle : [\![\Gamma]\!] \rightarrow \prod \limits_{i=1}^n \mathcal{K}[\![A_i]\!]$}
+    \AxiomC{$[\![\Gamma \vdash \vec{M} : \Box \vec{A}]\!] = \langle [\![M_1]\!],\dots, [\![M_n]\!] \rangle : [\![\Gamma]\!] \rightarrow \prod \limits_{i=1}^n \boxdot [\![A_i]\!]$}
     \AxiomC{$[\![\vec{x} : \vec{A} \vdash N : B]\!] = [\![N]\!] : \prod \limits_{i=1}^n [\![A_i]\!] \rightarrow [\![B]\!]$}
-    \BinaryInfC{$[\![\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } M : {\bf K} B]\!] = \mathcal{K}([\![N]\!]) \circ \ast_{[\![A_1]\!],\dots,[\![A_n]\!]} \circ \langle [\![M_1]\!],\dots, [\![M_n]\!] \rangle : [\![\Gamma]\!] \rightarrow \mathcal{K}[\![B]\!]$}
+    \BinaryInfC{$[\![\Gamma \vdash {\bf let \: box \:} \vec{x} = \vec{M} {\: \bf in \: } M : \Box B]\!] = \boxdot ([\![N]\!]) \circ \ast_{[\![A_1]\!],\dots,[\![A_n]\!]} \circ \langle [\![M_1]\!],\dots, [\![M_n]\!] \rangle : [\![\Gamma]\!] \rightarrow \boxdot [\![B]\!]$}
   \end{prooftree}
 \end{small}
-\end{itemize}
-\end{defin}
-
-\begin{defin} Simultaneous substitution
-
-Let $\Gamma = \{ x_1 : A_1, ..., x_n : A_n \}$, $\Gamma \vdash M : A$ and for all $i \in \{ 1,..., n \}$,
-$\Gamma \vdash M_i : A_i$.
-
-We define simultaneous substitution $M [ \vec{x} := \vec{M}]$ recursively by:
-
-\begin{itemize}
-\item $x_i [ \vec{x} := \vec{M}] = M_i $;
-\item $(\lambda x. M) [ \vec{x} := \vec{M}] = \lambda x. (M [ \vec{x} := \vec{M}])$;
-\item $(M N) [ \vec{x} := \vec{M}] = (M [ \vec{x} = \vec{M}]) (N [ \vec{x} := \vec{M}])$;
-\item $\langle M, N \rangle = \langle (M [ \vec{x} = \vec{M}]), (N [ \vec{x} := \vec{M}])\rangle$;
-\item $(\pi_i P) [ \vec{x} := \vec{M}] = \pi_i (P [ \vec{x} = \vec{M}])$;
-\item $({\bf pure \:} M) [ \vec{x} := \vec{M}] = {\bf pure \:} (M [ \vec{x} = \vec{M}])$;
-\item $({\bf let \: pure} \: \vec{x} = \vec{M} {\: \bf in \:} N) [\vec{y} := \vec{P}] =
-{\bf let \: pure} \: \vec{x} = (\vec{M} [\vec{y} := \vec{P}]) {\: \bf in \:} N$
-\end{itemize}
-\end{defin}
+\end{enumerate}
+\end{defin}
+
+\begin{theorem} Soundness
+
+  Let $\Gamma \vdash M : A$ and $M =_r N$, then $[\![\Gamma \vdash M : A]\!] = [\![\Gamma \vdash N : A]\!]$
+\end{theorem}
+
+\begin{proof}
 
@@
 $\begin{array}{lll}
-& [\![\Gamma \vdash ({\bf pure \:} M) [ \vec{x} := \vec{M}] : {\bf K}A]\!] = [\![ \Gamma \vdash {\bf pure \:} (M [ \vec{x} := \vec{M}]) : {\bf K}A]\!] &
-\text{Substitution definition} \\
-& \quad\quad\quad\quad\quad\quad\quad = \eta_{[\![A]\!]} \circ [\![(M [ \vec{x} := \vec{M}])]\!]&
-\text{Translation for pure}\\
-& \quad\quad\quad\quad\quad\quad\quad = \eta_{[\![A]\!]} \circ ([\![M]\!] \circ \langle [\![M_1]\!], \dots,
-[\![M_n]\!] \rangle) & \text{Induction hypothesis} \\
-& \quad\quad\quad\quad\quad\quad\quad = (\eta_{[\![A]\!]} \circ [\![M]\!]) \circ \langle [\![M_1]\!], \dots,
-[\![M_n]\!] \rangle & \text{Associativity of composition}\\
-& \quad\quad\quad\quad\quad\quad\quad = [\![ \Gamma \vdash {\bf pure \:} M : {\bf K}A]\!] \circ \langle [\![M_1]\!], \dots,
-[\![M_n]\!] \rangle& \text{Translation for pure}\\
+& [\![\Gamma \vdash ({\bf box \:} M) [ \vec{x} := \vec{M}] : \Box A]\!] = & \\
+&\quad\quad\quad\quad\quad\quad\quad \text{By substitution definition}& \\
+&[\![ \Gamma \vdash {\bf box \:} (M [ \vec{x} := \vec{M}]) : \Box A]\!]& \\
+&\quad\quad\quad\quad\quad\quad\quad \text{Interpretation for {\bf box}}\\
+&\eta_{[\![A]\!]} \circ [\![(M [ \vec{x} := \vec{M}])]\!]& \\
+&\quad\quad\quad\quad\quad\quad\quad \text{Assumption} \\
+&\eta_{[\![A]\!]} \circ ([\![M]\!] \circ \langle [\![M_1]\!], \dots, [\![M_n]\!] \rangle) = & \\
+&\quad\quad\quad\quad\quad\quad\quad \text{Associativity of composition}&\\
+&(\eta_{[\![A]\!]} \circ [\![M]\!]) \circ \langle [\![M_1]\!], \dots, [\![M_n]\!] \rangle = & \\
+&\quad\quad\quad\quad\quad\quad\quad  \text{Interpretation for {\bf box}}& \\
+&[\![ \Gamma \vdash {\bf box \:} M : \Box A]\!] \circ \langle [\![M_1]\!], \dots, [\![M_n]\!] \rangle = & \\
 \end{array}$
@@
 $\begin{array}{lll}
-&[\![\Gamma \vdash ({\bf let \: pure} \: \vec{x} = \vec{M} {\: \bf in \:} N) [\vec{y} := \vec{P}] : {\bf K} B]\!] =& \text{Substitution definition}\\
-&[\![\Gamma \vdash {\bf let \: pure} \: \vec{x} = (\vec{M} [\vec{y} := \vec{P}]) {\: \bf in \:} N : {\bf K}B]\!] =& \text{Interpretaion for $let_{\bf K}$} \\
-&\mathcal{K}([\![N]\!]) \circ \ast_{[\![A_1]\!],\dots,[\![A_n]\!]} \circ [\![\Gamma \vdash (\vec{M} [\vec{y} := \vec{P}]) : {\bf K}\vec{A}]\!] =& \text{Induction hypothesis} \\
-&\mathcal{K}([\![N]\!]) \circ \ast_{[\![A_1]\!],\dots,[\![A_n]\!]} \circ ([\![\vec{M}]\!] \circ \langle [\![P_1]\!],\dots,[\![P_n]\!]\rangle) = & \text{Associativity of composition} \\
-&(\mathcal{K}([\![N]\!]) \circ \ast_{[\![A_1]\!],\dots,[\![A_n]\!]} \circ [\![\vec{M}]\!]) \circ \langle [\![P_1]\!],\dots,[\![P_n]\!]\rangle = & \text{By interpretation} \\
-&[\![\Gamma \vdash {\bf let \: pure} \: \vec{x} = \vec{M} {\: \bf in \:} N : {\bf K}B]\!] \circ \langle [\![P_1]\!],\dots,[\![P_n]\!]\rangle&
+&[\![\Gamma \vdash ({\bf let \: box} \: \vec{x} = \vec{M} {\: \bf in \:} N) [\vec{y} := \vec{P}] : \Box B]\!] =& \\
+&\quad\quad\quad\quad\quad\quad\quad \text{Substitution} &\\
+&[\![\Gamma \vdash {\bf let \: box} \: \vec{x} = (\vec{M} [\vec{y} := \vec{P}]) {\: \bf in \:} N : \Box B]\!] =& \\
+&\quad\quad\quad\quad\quad\quad\quad  \text{Interpretation for $let_{\Box}$} \\
+&\boxdot ([\![N]\!]) \circ ([\![A_1]\!] \ast \dots \ast [\![A_n]\!]) \circ [\![\Gamma \vdash (\vec{M} [\vec{y} := \vec{P}]) : \Box \vec{A}]\!] =& \\
+&\quad\quad\quad\quad\quad\quad\quad \text{Induction hypothesis}& \\
+&\boxdot ([\![N]\!]) \circ ([\![A_1]\!] \ast \dots \ast [\![A_n]\!]) \circ ([\![\vec{M}]\!] \circ \langle [\![P_1]\!],\dots,[\![P_n]\!]\rangle) = & \\
+&\quad\quad\quad\quad\quad\quad\quad \text{Associativity of composition}& \\
+&(\boxdot ([\![N]\!]) \circ ([\![A_1]\!] \ast \dots \ast [\![A_n]\!]) \circ [\![\vec{M}]\!]) \circ \langle [\![P_1]\!],\dots,[\![P_n]\!]\rangle = & \\
+&\quad\quad\quad\quad\quad\quad\quad \text{Interpretaion}& \\
+&[\![\Gamma \vdash {\bf let \: box} \: \vec{x} = \vec{M} {\: \bf in \:} N : \Box B]\!] \circ \langle [\![P_1]\!],\dots,[\![P_n]\!]\rangle&
 \end{array}$
@@
 
-  Let $\Gamma \vdash M : A$ and $M \twoheadrightarrow_{\beta \eta} N$, then $[\![\Gamma \vdash M : A]\!] = [\![\Gamma \vdash N : A]\!]$;
-\end{lemma}
-
-\begin{proof}
-  $ $
-
-Cases with $\beta$-reductions for $let_{\bf K}$ are shown in \cite{ModalK1}. Let us consider cases with ${\bf pure}$.
-
-\vspace{\baselineskip}
-
-1) $[\![\Gamma \vdash {\bf let \: pure \:} \vec{x} = {\bf pure \:} \vec{M} {\: \bf in \:} N : {\bf K}B]\!] = [\![\Gamma \vdash {\bf pure \:} N [\vec{x} := \vec{M}] : {\bf K}B]\!]$
-
-\vspace{\baselineskip}
-
-$\begin{array}{lll}
-&[\![\Gamma \vdash {\bf let \: pure \:} \vec{x} = {\bf pure \:} \vec{M} {\: \bf in \:} N : {\bf K}B]\!] = & \\
-&\text{\quad\quad\quad\quad\quad\quad\quad\quad\quad By interpretation}& \\
-&\mathcal{K}([\![N]\!]) \circ \ast_{[\![A_1]\!],\dots,[\![A_n]\!]} \circ \langle \eta_{[\![A_1]\!]} \circ [\![M_1]\!],\dots,\eta_{[\![A_n]\!]} \circ [\![M_n]\!] \rangle = \\
-&\text{\quad\quad\quad\quad\quad\quad\quad\quad\quad By the property of a pair of morphisms}& \\
-&\mathcal{K}([\![N]\!]) \circ \ast_{[\![A_1]\!],\dots,[\![A_n]\!]} \circ (\eta_{[\![A_1]\!]} \times \dots \times \eta_{[\![A_n]\!]}) \circ \langle [\![M_1]\!], \dots, [\![M_n]\!]\rangle =& \\
-&\text{\quad\quad\quad\quad\quad\quad\quad\quad\quad Associativity of composition}& \\
-&\mathcal{K}([\![N]\!]) \circ (\ast_{[\![A_1]\!],\dots,[\![A_n]\!]} \circ (\eta_{[\![A_1]\!]} \times \dots \eta_{[\![A_n]\!]})) \circ \langle [\![M_1]\!], \dots, [\![M_n]\!] \rangle =& \\
-&\text{\quad\quad\quad\quad\quad\quad\quad\quad\quad By the definition of an applicative functor}& \\
-&\mathcal{K}([\![N]\!]) \circ \eta_{[\![A_1]\!] \times \dots \times [\![A_n]\!]} \circ \langle [\![M_1]\!], \dots, [\![M_n]\!] \rangle =& \\
-&\text{\quad\quad\quad\quad\quad\quad\quad\quad\quad Naturality of $\eta$}& \\
+  Let $\Gamma \vdash M : A$ and $M \to_r N$, then $[\![\Gamma \vdash M : A]\!] = [\![\Gamma \vdash N : A]\!]$;
+\end{lemma}
+
+\begin{proof}
+  $ $
+
+Cases with $\beta$-reductions for $let_{\Box}$ are shown in \cite{ModalK1}. Let us consider cases with ${\bf box}$.
+
+\vspace{\baselineskip}
+
+1) $[\![\Gamma \vdash {\bf let \: box \:} \vec{x} = {\bf box \:} \vec{M} {\: \bf in \:} N : \Box B]\!] = [\![\Gamma \vdash {\bf box \:} N [\vec{x} := \vec{M}] : \Box B]\!]$
+
+\vspace{\baselineskip}
+
+$\begin{array}{lll}
+&[\![\Gamma \vdash {\bf let \: box \:} \vec{x} = {\bf box \:} \vec{M} {\: \bf in \:} N : \Box B]\!] = & \\
+&\quad\quad\quad\quad\quad\quad\quad \text{By interpretation}& \\
+&\boxdot ([\![N]\!]) \circ ([\![A_1]\!] \varoast \dots \varoast [\![A_n]\!]) \circ \langle \eta_{[\![A_1]\!]} \circ [\![M_1]\!],\dots,\eta_{[\![A_n]\!]} \circ [\![M_n]\!] \rangle = \\
+&\quad\quad\quad\quad\quad\quad\quad \text{By the property of a pair of morphisms}& \\
+&\boxdot ([\![N]\!]) \circ ([\![A_1]\!] \varoast \dots \varoast [\![A_n]\!]) \circ (\eta_{[\![A_1]\!]} \times \dots \times \eta_{[\![A_n]\!]}) \circ \langle [\![M_1]\!], \dots, [\![M_n]\!]\rangle =& \\
+&\quad\quad\quad\quad\quad\quad\quad \text{Associativity of composition}& \\
+&\boxdot ([\![N]\!]) \circ (([\![A_1]\!] \varoast \dots \varoast [\![A_n]\!]) \circ (\eta_{[\![A_1]\!]} \times \dots \times \eta_{[\![A_n]\!]})) \circ \langle [\![M_1]\!], \dots, [\![M_n]\!] \rangle =& \\
+&\quad\quad\quad\quad\quad\quad\quad \text{By the definition of an applicative functor}& \\
+&\boxdot ([\![N]\!]) \circ \eta_{[\![A_1]\!] \times \dots \times [\![A_n]\!]} \circ \langle [\![M_1]\!], \dots, [\![M_n]\!] \rangle =& \\
+&\quad\quad\quad\quad\quad\quad\quad \text{Naturality of $\eta$}& \\
 &\eta_{[\![B]\!]} \circ [\![N]\!] \circ \langle [\![M_1]\!], \dots, [\![M_n]\!] \rangle =& \\
-&\text{\quad\quad\quad\quad\quad\quad\quad\quad\quad Associativity of composition}& \\
+&\quad\quad\quad\quad\quad\quad\quad \text{Associativity of composition}& \\
 &\eta_{[\![B]\!]} \circ ([\![N]\!] \circ \langle [\![M_1]\!], \dots, [\![M_n]\!]) \rangle =& \\
-&\text{\quad\quad\quad\quad\quad\quad\quad\quad\quad Simultaneous substitution lemma}& \\
-&\eta_{[\![B]\!]} \circ [\![N [\vec{x} := \vec{M}]]\!]& \\
-&\text{\quad\quad\quad\quad\quad\quad\quad\quad\quad By interpetation}& \\
-&[\![\Gamma \vdash {\bf pure \:} (N [\vec{x} := \vec{M}]) : {\bf K}B]\!]&
-\end{array}$
-
-\vspace{\baselineskip}
-
-2) $[\![\vdash {\bf let \: pure \:} \underline{\quad} = \underline {\quad} {\: \bf in \:} M : {\bf K} A]\!] = [\![\vdash {\bf pure \:} M : {\bf K} A]\!]$
-
-$\begin{array}{lll}
-&[\![\vdash {\bf let \: pure \:} \underline{\quad} = \underline {\quad} {\: \bf in \:} M : {\bf K} A]\!] = & \\
-&\quad\quad\quad\quad\quad\quad \text{By interpretation}& \\
-&\mathcal{K}([\![M]\!]) \circ u_{\mathds{1}} = & \\
-&\quad\quad\quad\quad\quad\quad \text{By the definition of an applicative functor}& \\
-&\mathcal{K}([\![M]\!]) \circ \eta_{\mathds{1}} = & \\
-&\quad\quad\quad\quad\quad\quad \text{By naturality of $\eta$} & \\
+&\quad\quad\quad\quad\quad\quad\quad \text{Substitution lemma}& \\
+&\eta_{[\![B]\!]} \circ [\![\Gamma \vdash N [\vec{x} := \vec{M}] : \Box B]\!]& \\
+&\quad\quad\quad\quad\quad\quad\quad \text{By interpetation}& \\
+&[\![\Gamma \vdash {\bf box \:} (N [\vec{x} := \vec{M}]) : \Box B]\!]&
+\end{array}$
+
+\vspace{\baselineskip}
+
+2) $[\![\vdash {\bf let \: box \:} \underline{\quad} = \underline {\quad} {\: \bf in \:} M : \Box A]\!] = [\![\vdash {\bf box \:} M : \Box A]\!]$
+
+$\begin{array}{lll}
+&[\![\vdash {\bf let \: box \:} \underline{\quad} = \underline {\quad} {\: \bf in \:} M : \Box A]\!] = & \\
+&\quad\quad\quad\quad\quad\quad\quad \text{By interpretation}& \\
+&\boxdot ([\![M]\!]) \circ u_{\mathds{1}} = & \\
+&\quad\quad\quad\quad\quad\quad\quad \text{By the definition of an applicative functor}& \\
+&\boxdot {K}([\![M]\!]) \circ \eta_{\mathds{1}} = & \\
+&\quad\quad\quad\quad\quad\quad\quad \text{By naturality of $\eta$} & \\
 &\eta_{[\![A]\!]} \circ [\![M]\!] = &\\
-&\quad\quad\quad\quad\quad\quad \text{By interpretation} & \\
-&[\![\vdash {\bf pure \:} M : {\bf K} A]\!]&
+&\quad\quad\quad\quad\quad\quad\quad \text{By interpretation} & \\
+&[\![\vdash {\bf box \:} M : \Box A]\!]&
 \end{array}$
@@
 
-Let $[\![\Gamma \vdash M : A]\!] = [\![\Gamma \vdash N : A]\!]$, then $M =_{\beta \eta} N$.
+Let $[\![\Gamma \vdash M : A]\!] = [\![\Gamma \vdash N : A]\!]$, then $M =_r N$.
 \end{theorem}
@@
 
-  Let us define relation $\sim_{A, B} \subseteq \mathbb{V} \times \Lambda_{{\bf K}}$, such that:
-
-  $(x, M) \sim_{A, B} (y, N) \Leftrightarrow x : A \vdash M : B \:\: \& \:\: y : A \vdash N : A \:\: \& \:\: M =_{\beta \eta} N [y := x]$;
+  Let us define relation $\sim_{A, B} \subseteq (\mathbb{V} \times \Lambda_{\Box})^2$, such that:
+
+  $(x, M) \sim_{A, B} (y, N) \Leftrightarrow x : A \vdash M : B \:\: \& \:\: y : A \vdash N : A \:\: \& \:\: M =_r N [y := x]$;
 \end{defin}
@@
 \begin{defin} Category $\mathcal{C}(\lambda)$:
-\begin{itemize}
+\begin{enumerate}
   \item $Ob_{\mathcal{C}} = \{ \hat{A} \: | \: A \in \mathbb{T} \} \cup \{ \mathds{1} \}$;
-  \item $Hom_{\mathcal{C}(\lambda)}(\hat{A},\hat{B}) = (\mathbb{V} \times \Lambda_{{\bf K}})/_{\sim_{A, B}}$;
+  \item $Hom_{\mathcal{C}(\lambda)}(\hat{A},\hat{B}) = \{ [x, M] \: | \: x : A \vdash_{\lambda_{\text{IEL}^{-}}} M : B \}$;
   \item Let $[x, M] \in Hom_{\mathcal{C}(\lambda)}(\hat{A},\hat{B})$ and $[y,N] \in Hom_{\mathcal{C}(\lambda)}(\hat{B},\hat{C})$, then $[y,M] \circ [x, M] = [x, N [y := M]]$;
-  \item Identity morphism $id_{\hat{A}} = [x,x] \in Hom_{\mathcal{C}(\lambda)(\hat{A})}$;
+  \item Identity morphism $id_{\hat{A}} = [x,x] \in Hom_{\mathcal{C}(\lambda)(\hat{A}, \hat{A})}$;
   \item $\mathds{1}$ is a terminal object;
@@
   \item Evaluation arrow $\epsilon = [x, (\pi_2 x) (\pi_1 x)] \in Hom_{\mathcal{C}(\lambda)(\hat{B}^{\hat{A}} \times \hat{A}, \hat{B})}$.
-\end{itemize}
-\end{defin}
-
-It is sufficient to show ${\bf K}$ is an applicative functor on $\mathcal{C}(\lambda)$.
+\end{enumerate}
+\end{defin}
+
+We define endofunctor $\boxdot$ on $\mathcal{C}(\lambda)$ and natural transformation $\eta$ from $id_{\mathcal{C}(\lambda)}$ to this endofunctor.
+It is sufficient to show $\boxdot$ and $\eta$ form an applicative functor on $\mathcal{C}(\lambda)$ for theorem proof.
 
 \begin{defin}
-  Let us define an endofunctor $\mathcal{K} : \mathcal{C}(\lambda) \to \mathcal{C}(\lambda)$, such that
-forall $[x,M] \in Hom_{\mathcal{C}(\lambda)}(\hat{A},\hat{B}), {\bf K}([x,M]) = [y, {\bf let \: pure \:} x = y {\: \bf in \:} M] \in Hom_{\mathcal{C}(\lambda)}({\bf K} \hat{A}, {\bf K} \hat{B})$ (denotation: $\text{fmap } f$ for an arbitrary arrow $f$).
+  Let us define an endofunctor $\boxdot : \mathcal{C}(\lambda) \to \mathcal{C}(\lambda)$, such that
+forall $[x,M] \in Hom_{\mathcal{C}(\lambda)}(\hat{A},\hat{B}), \boxdot ([x,M]) = [y, {\bf let \: box \:} x = y {\: \bf in \:} M] \in Hom_{\mathcal{C}(\lambda)}(\boxdot \hat{A}, \boxdot \hat{B})$
+(denotation: $\text{fmap } f$ for an arbitrary arrow $f$).
 
@@
 
-\begin{itemize}
+\begin{enumerate}
   \item $\text{fmap }(g \circ f) = \text{fmap }(g) \circ \text{fmap }(f)$;
-  \item $\text{fmap }(id_{\hat{A}}) = id_{{\bf K}\hat{A}}$.
-\end{itemize}
+  \item $\text{fmap }(id_{\hat{A}}) = id_{\boxdot}\hat{A}$.
+\end{enumerate}
 \end{lemma}
@@
 
-\begin{itemize}
-  \item $\eta:Id \Rightarrow \mathcal{K}$, s. t. $\forall \hat{A} \in Ob_{\mathcal{C}(\lambda)}$, $\eta_{\hat{A}} = [x, {\bf pure \:} x] \in Hom_{\mathcal{C}(\lambda)}(\hat{A}, {\bf K}\hat{A})$;
-  \item $\ast_{A,B}:{\bf K}\hat{A} \times {\bf K}\hat{B} \to {\bf K}(\hat{A} \times \hat{B})$, s.
+\begin{enumerate}
+  \item $\eta: Id_{\mathcal{C}(\lambda)} \Rightarrow \boxdot$, s.
 t.
-$\forall \hat{A}, \hat{B} \in Ob_{\mathcal{C}(\lambda)}, \ast_{\hat{A},\hat{B}} = [p, {\bf let \: pure \:} x,y = \pi_1 p, \pi_2 p {\: \bf in \:} \langle x, y \rangle] \in Hom_{\mathcal{C}(\lambda)}({\bf K}A \times {\bf K}B, {\bf K}(A \times B))$.
-\end{itemize}
-\end{defin}
-
-Implementation for $\ast$ in our term model is a modification of ${\text{let}_{\bf K}}$-rule:
-
-\begin{prooftree}
-\AxiomC{$p : {\bf K}A \times {\bf K}B \vdash p : {\bf K}A \times {\bf K}B$}
-\UnaryInfC{$p : {\bf K}A \times {\bf K}B \vdash \pi_1 p : {\bf K} A$}
-\AxiomC{$p : {\bf K}A \times {\bf K}B \vdash p : {\bf K}A \times {\bf K}B$}
-\UnaryInfC{$p : {\bf K}A \times {\bf K}B \vdash \pi_2 p : {\bf K} B$}
+$\forall \hat{A} \in Ob_{\mathcal{C}(\lambda)}$, $\eta_{\hat{A}} = [x, {\bf box \:} x] \in Hom_{\mathcal{C}(\lambda)}(\hat{A}, \boxdot \hat{A})$;
+  \item $\hat{A} \varoast \hat{B} : \boxdot \hat{A} \times \boxdot \hat{B} \to \boxdot (\hat{A} \times \hat{B})$, s.
+t.
+$\forall \hat{A}, \hat{B} \in Ob_{\mathcal{C}(\lambda)}, \ast_{\hat{A},\hat{B}} = [p, {\bf let \: box \:} x,y = \pi_1 p, \pi_2 p {\: \bf in \:} \langle x, y \rangle] \in Hom_{\mathcal{C}(\lambda)}(\boxdot \hat{A} \times \boxdot \hat{B}, \boxdot (\hat{A} \times \hat{B}))$.
+\end{enumerate}
+\end{defin}
+
+Implementation for $\ast$ in our term model is the instance of ${\text{let}_{\Box}}$-rule:
+
+\begin{prooftree}
+\AxiomC{$p : \Box A \times \Box B \vdash p : \Box A \times \Box B$}
+\UnaryInfC{$p : \Box A \times \Box B \vdash \pi_1 p : \Box A$}
+\AxiomC{$p : \Box A \times \Box B \vdash p : \Box A \times \Box B$}
+\UnaryInfC{$p : \Box A \times \Box B \vdash \pi_2 p : \Box B$}
 \AxiomC{$x : A \vdash x : A$}
@@
 \BinaryInfC{$x : A, y : B \vdash \langle x, y \rangle : A \times B$}
-\TrinaryInfC{$p : {\bf K}A \times {\bf K}B \vdash {\bf let \: pure \:} x,y = \pi_1 p, \pi_2 p {\: \bf in \:} \langle x, y \rangle : {\bf K}(A \times B)$}
+\TrinaryInfC{$p : \Box A \times \Box B \vdash {\bf let \: box \:} x,y = \pi_1 p, \pi_2 p {\: \bf in \:} \langle x, y \rangle : \Box (A \times B)$}
 \end{prooftree}
@@
 
-  {\bf K} is a lax monoidal endofunctor
+  $\boxdot$ is a monoidal endofunctor
 \end{lemma}
@@
 
-\begin{itemize}
+\begin{enumerate}
   \item $\text{fmap } f \circ \eta_A = \eta_B \circ f$;
-  \item $\ast_{\hat{A},\hat{B}} \circ (\eta_{A} \times \eta_{B}) = \eta_{\hat{A} \times \hat{B}}$;
-\end{itemize}
-\end{lemma}
-
-\begin{proof}
-  $ $
-
-  i) $\text{fmap } f \circ \eta_{\hat{A}} = \eta_{\hat{B}} \circ f$
-
-\vspace{\baselineskip}
+  \item $(\hat{A} \varoast \hat{B}) \circ (\eta_{A} \times \eta_{B}) = \eta_{\hat{A} \times \hat{B}}$;
+\end{enumerate}
+\end{lemma}
+
+\begin{proof}
+  $ $
+
+\begin{enumerate}
+\item $\text{fmap } f \circ \eta_{\hat{A}} = \eta_{\hat{B}} \circ f$ \\
 
@@
 &\eta_{\hat{B}} \circ f = & \\
-&\quad\quad\quad\quad\quad\quad \text{By the definition}&\\
-&[y, {\bf pure \:} y] \circ [x, M] = &\\
-&\quad\quad\quad\quad\quad\quad \text{By the definition of composition} & \\
-&[x, {\bf pure \:} y [y := M]] = &\\
+&[y, {\bf box \:} y] \circ [x, M] = &\\
+&\quad\quad\quad\quad\quad\quad \text{Composition} & \\
+&[x, {\bf box \:} y [y := M]] = &\\
 &\quad\quad\quad\quad\quad\quad \text{By substitution}& \\
-&[x, {\bf pure \:} M]& \\
+&[x, {\bf box \:} M]& \\
 & &\\
@@
 &\text{fmap } f \circ \eta_{\hat{A}} = & \\
-&\quad\quad\quad\quad\quad\quad \text{By the definiton}& \\
-&[z, {\bf let \: pure \:} x = z {\: \bf in \:} M] \circ [x, {\bf pure \: x}] = &\\
+&[z, {\bf let \: box \:} x = z {\: \bf in \:} M] \circ [x, {\bf box \:} x] = &\\
 &\quad\quad\quad\quad\quad\quad \text{By the definition of composition}& \\
-&[x, {\bf let \: pure \:} x = z {\: \bf in \:} M [z := {\bf pure \:} x]] = & \\
+&[x, {\bf let \: box \:} x = z {\: \bf in \:} M [z := {\bf box \:} x]] = & \\
 &\quad\quad\quad\quad\quad\quad \text{By substitution}& \\
-&[x, {\bf let \: pure \:} x = {\bf pure \:} x {\: \bf in \:} M] = & \\
+&[x, {\bf let \: box \:} x = {\bf box \:} x {\: \bf in \:} M] = & \\
 &\quad\quad\quad\quad\quad\quad \text{$\beta$-reduction rule}& \\
-&[x, {\bf pure \:} M [x := x]] = &\\
+&[x, {\bf box \:} M [x := x]] = &\\
 &\quad\quad\quad\quad\quad\quad \text{By substitution} & \\
-&[x, {\bf pure \:} M]&
-\end{array}$
-
-\vspace{\baselineskip}
-
-ii) $\ast_{\hat{A},\hat{B}} \circ (\eta_{\hat{A}} \times \eta_{\hat{B}}) = \eta_{\hat{A} \times \hat{B}}$
-
-\vspace{\baselineskip}
-
-$\begin{array}{lll}
-& \ast_{\hat{A},\hat{B}} \circ (\eta_{\hat{A}} \times \eta_{\hat{B}}) = & \\
-& \quad\quad\quad\quad\quad\quad \text{By unfolding} & \\
-& [q, {\bf let \: pure \:} x, y = \pi_1 q, \pi_2 q {\: \bf in \:} \langle x, y \rangle] \circ [p, \langle {\bf pure \:} (\pi_1 p), {\bf pure \:} (\pi_2 p) \rangle] = & \\
+&[x, {\bf box \:} M]&
+\end{array}$
+
+\item $(\hat{A} \varoast \hat{B}) \circ (\eta_{A} \times \eta_{B}) = \eta_{\hat{A} \times \hat{B}}$ \\
+
+$\begin{array}{lll}
+& (\hat{A} \varoast \hat{B}) \circ (\eta_{A} \times \eta_{B}) = & \\
+& [q, {\bf let \: box \:} x, y = \pi_1 q, \pi_2 q {\: \bf in \:} \langle x, y \rangle] \circ [p, \langle {\bf box \:} (\pi_1 p), {\bf box \:} (\pi_2 p) \rangle] = & \\
 & \quad\quad\quad\quad\quad\quad \text{Composition}& \\
-& [p, {\bf let \: pure \:} x, y = \pi_1 q, \pi_2 q {\: \bf in \:} \langle x, y \rangle [q := \langle {\bf pure \:} (\pi_1 p), {\bf pure \:} (\pi_2 p) \rangle]] = & \\
+& [p, {\bf let \: box \:} x, y = \pi_1 q, \pi_2 q {\: \bf in \:} \langle x, y \rangle [q := \langle {\bf box \:} (\pi_1 p), {\bf box \:} (\pi_2 p) \rangle]] = & \\
 & \quad\quad\quad\quad\quad\quad \text{By substitution}& \\
-& [p, {\bf let \: pure \:} x, y = \pi_1 (\langle {\bf pure \:} (\pi_1 p), {\bf pure \:} (\pi_2 p) \rangle), \pi_2 (\langle {\bf pure \:} (\pi_1 p), {\bf pure \:} (\pi_2 p) \rangle) {\: \bf in \:} \langle x, y \rangle] = & \\
+& [p, {\bf let \: box \:} x, y = \pi_1 (\langle {\bf box \:} (\pi_1 p), {\bf box \:} (\pi_2 p) \rangle), \pi_2 (\langle {\bf box \:} (\pi_1 p), {\bf box \:} (\pi_2 p) \rangle) {\: \bf in \:} \langle x, y \rangle] = & \\
 & \quad\quad\quad\quad\quad\quad \text{Reduction rules}& \\
-& [p, {\bf let \: pure \:} x, y = {\bf pure \:} (\pi_1 p), {\bf pure \:} (\pi_2 p) {\: \bf in \:} \langle x, y \rangle] = & \\
+& [p, {\bf let \: box \:} x, y = {\bf box \:} (\pi_1 p), {\bf box \:} (\pi_2 p) {\: \bf in \:} \langle x, y \rangle] = & \\
 & \quad\quad\quad\quad\quad\quad \text{Reduction rule}& \\
-& [p, {\bf pure \:} (\langle x,y \rangle [x := \pi_1 p, y := \pi_2 p])] = & \\
+& [p, {\bf box \:} (\langle x,y \rangle [x := \pi_1 p, y := \pi_2 p])] = & \\
 & \quad\quad\quad\quad\quad\quad \text{Substitution}& \\
-& [p, {\bf pure \:} \langle \pi_1 p, \pi_2 p \rangle] = & \\
+& [p, {\bf box \:} \langle \pi_1 p, \pi_2 p \rangle] = & \\
 & \quad\quad\quad\quad\quad\quad \text{$\eta$-reduction}& \\
-& [p, {\bf pure \:} p] =& \\
+& [p, {\bf box \:} p] =& \\
 & \quad\quad\quad\quad\quad\quad \text{By definition}& \\
@@
 \end{array}$
-
+\end{enumerate}
 \end{proof}
@@
 
-  $u_{\mathds{1}} = [\sqbullet, {\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} \sqbullet] \in Hom_{\mathcal{C}(\lambda)}(\mathds{1}, {\bf K}\mathds{1})$.
+  $u_{\mathds{1}} = [\sqbullet, {\bf let \: box \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} \sqbullet] \in Hom_{\mathcal{C}(\lambda)}(\mathds{1}, \boxdot \mathds{1})$.
 \end{defin}
@@
 
-  Tensorial strength is defined as follows:
-
-\begin{defin} Tensorial strength
-  $ $
-
-  Let $[p, \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle] \in Hom_{\mathcal{C}(\lambda)}(\hat{A} \times {\bf K}\hat{B}, {\bf K}\hat{A} \times {\bf K}\hat{B})$.
-
-  So tensorial strength is defined as $\tau_{\hat{A}, \hat{B}} = \ast_{\hat{A}, \hat{B}} \circ [p, \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle]$.
-\end{defin}
-
-It is clearly that tensorial strength defined above can be simplified as follows:
-
-$\begin{array}{lll}
-&\ast_{\hat{A}, \hat{B}} \circ [p, \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle] = & \\
-&\quad\quad\quad\quad\quad\quad \text{By definition}&\\
-&[p^{'}, {\bf let \: pure \:} x, y = \pi_1 p^{'}, \pi_2 p^{'} {\: \bf in \:} \langle x, y \rangle] \circ [p, \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle] = & \\
-&\quad\quad\quad\quad\quad\quad \text{By composition} & \\
-&[p, {\bf let \: pure \:} x, y = \pi_1 p^{'}, \pi_2 p^{'} {\: \bf in \:} \langle x, y \rangle [p^{'} := \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle]] = & \\
-&\quad\quad\quad\quad\quad\quad \text{By substitution} & \\
-&[p, {\bf let \: pure \:} x, y = \pi_1 (\langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle), \pi_2 (\langle \pi_1 p, {\bf pure \:} (\pi_2 p) \rangle) {\: \bf in \:} \langle x, y \rangle] = & \\
-&\quad\quad\quad\quad\quad\quad \text{By $\beta$-reduction rules} & \\
-&[p, {\bf let \: pure \:} x, y = {\bf pure \:} (\pi_1 p), \pi_2 p {\: \bf in \:} \langle x, y \rangle]&
-\end{array}$
-
-\begin{lemma} Coherence for tensorial strength
-
-  \begin{itemize}
-    \item $\text{fmap } \alpha_{\hat{A},\hat{B},\hat{C}} \circ \tau_{\hat{A} \times \hat{B}, \hat{C}} = \tau_{\hat{A}, \hat{B} \times \hat{C}} \circ (id_{\hat{A}} \times \tau_{\hat{B}, \hat{C}}) \circ \alpha_{\hat{A}, \hat{B}, {\bf K}\hat{C}}$;
-    \item ${\bf K}(R_{\hat{A}}) \circ \tau_{\mathds{1}, \hat{A}} = R_{{\bf K}\hat{A}}$.
-  \end{itemize}
-  where $\alpha_{\hat{A},\hat{B},\hat{C}} = [p, \langle \pi_1 (\pi_1 p), \langle \pi_2 (\pi_1 p), \pi_2 p \rangle \rangle]$ and
-  $R = \pi_2$.
-\end{lemma}
-
-\begin{proof}
-  $ $
-
-1) Let us define $\tau_{\hat{A} \times \hat{B}, \hat{C}}$ as follows:
-
-$\tau_{\hat{A} \times \hat{B}, \hat{C}} = [p, {\bf \: let \: pure \:} x,y,z = {\bf pure\:} (\pi_1 (\pi_1 p)), {\bf pure \:} (\pi_2 (\pi_1 p)), \pi_2 p {\: \bf in \:} \langle \langle x,y \rangle, z \rangle]$
-
-\vspace{\baselineskip}
-
-Then:
-
-\vspace{\baselineskip}
-
-$\begin{array}{lll}
-&\text{fmap } \alpha_{\hat{A},\hat{B},\hat{C}} \circ \tau_{\hat{A} \times \hat{B}, \hat{C}} = & \\
-&[q, {\bf let \: pure \:} r = q {\: \bf in \:}  \langle \pi_1 (\pi_1 r), \langle \pi_2 (\pi_1 r), \pi_2 r \rangle \rangle] \: \circ & \\
-&\quad\quad\quad\quad \circ \: [p, {\bf \: let \: pure \:} x,y,z = {\bf pure\:} (\pi_1 (\pi_1 p)), {\bf pure \:} (\pi_2 (\pi_1 p)), \pi_2 p {\: \bf in \:} \langle \langle x,y \rangle, z \rangle] = & \\
-&\quad\quad\quad\quad\quad\quad \text{Composition}& \\
-&[p, {\bf let \: pure \:} r = q {\: \bf in \:}  \langle \pi_1 (\pi_1 r), \langle \pi_2 (\pi_1 r), \pi_2 r \rangle \rangle & \\
-&\quad\quad\quad\quad [q := {\bf \: let \: pure \:} x,y,z = {\bf pure\:} (\pi_1 (\pi_1 p)), {\bf pure \:} (\pi_2 (\pi_1 p)), \pi_2 p {\: \bf in \:} \langle \langle x,y \rangle, z \rangle]] = & \\
-&\quad\quad\quad\quad\quad\quad \text{Substitution and reduction}& \\
-&[p, {\bf let \: pure \:} r = ({\bf let \: pure \:} x,y,z = {\bf pure\:} (\pi_1 (\pi_1 p)), {\bf pure \:} (\pi_2 (\pi_1 p)), \pi_2 p {\: \bf in \:} \langle \langle x,y \rangle, z \rangle) & \\
-&\quad\quad\quad\quad {\: \bf in \:} \langle \pi_1 (\pi_1 r), \langle \pi_2 (\pi_1 r), \pi_2 r \rangle \rangle] = & \\
-&\quad\quad\quad\quad\quad\quad \text{$\beta$-reduction rule}& \\
-&[p, {\bf let \: pure \:} x,y,z = {\bf pure\:} (\pi_1 (\pi_1 p)), {\bf pure \:} (\pi_2 (\pi_1 p)), \pi_2 p {\: \bf in \:}  \langle \pi_1 (\pi_1 r), \langle \pi_2 (\pi_1 r), \pi_2 r \rangle \rangle & \\
-&\quad\quad\quad\quad [r := \langle \langle x,y \rangle, z \rangle]] = & \\
-&\quad\quad\quad\quad\quad\quad \text{$\beta$-reduction rule}& \\
-&[p, {\bf let \: pure \:} x,y,z = {\bf pure\:} (\pi_1 (\pi_1 p)), {\bf pure \:} (\pi_2 (\pi_1 p)), \pi_2 p {\: \bf in \:} \langle x, \langle y, z \rangle \rangle]&
-\end{array}$
-
-\vspace{\baselineskip}
-
-On the other hand:
-
-\vspace{\baselineskip}
-
-Let us define as follows $\tau_{\hat{A}, \hat{B} \times \hat{C}}$:
-
-$\begin{array}{lll}
-&\tau_{\hat{A}, \hat{B} \times \hat{C}} = [r, {\bf let \: pure \:} x, y, z = ({\bf pure \:} \pi_1 r, {\bf let \: pure \:} q^{'} = \pi_2 r {\: \bf in \:} \pi_1 q^{'}, {\bf let \: pure \:} q^{'} = \pi_2 r {\: \bf in \:} \pi_2 q^{'}) & \\
-&\quad\quad\quad\quad {\: \bf in \:} \langle x, \langle y, z \rangle\rangle]& \\
-\end{array}$
-
-\vspace{\baselineskip}
-
-We may simplify the current instance of tensorial strength:
-
-$\begin{array}{lll}
-&[p, {\bf let \: pure \:} x, y, z = ({\bf pure \:} \pi_1 p, {\bf let \: pure \:} p^{'} = \pi_2 p {\: \bf in \:} \pi_1 p^{'}, {\bf let \: pure \:} p^{'} = \pi_2 p {\: \bf in \:} \pi_2 p^{'}) & \\
-&\quad\quad\quad\quad {\: \bf in \:} \langle x, \langle y, z \rangle\rangle] = & \\
-&\quad\quad\quad\quad\quad\quad \text{Substitution and reduction}& \\
-&[p, {\bf let \: pure \:} x, p^{'}, z = ({\bf pure \:} \pi_1 p, \pi_2 p, {\bf let \: pure \:} p^{'} = \pi_2 p {\: \bf in \:} \pi_2 p^{'}) {\: \bf in \:} \langle x, \langle \pi_1 p^{'}, z \rangle\rangle] = & \\
-&\quad\quad\quad\quad\quad\quad \text{Substitution and reduction}& \\
-&[p, {\bf let \: pure \:} x, p^{'} = {\bf pure \:} \pi_1 p, \pi_2 p {\: \bf in \:} \langle x, \langle \pi_1 p^{'}, \pi_2 p^{'} \rangle\rangle]&
-\end{array}$
-
-\vspace{\baselineskip}
-
-Then:
-
-\vspace{\baselineskip}
-
-$\begin{array}{lll}
-&\tau_{\hat{A}, \hat{B} \times \hat{C}} \circ (id_{\hat{A}} \times \tau_{\hat{B}, \hat{C}}) \circ \alpha_{\hat{A}, \hat{B}, {\bf K}\hat{C}} = & \\
-&\quad\quad\quad\quad\quad\quad \text{Unfolding}& \\
-&\tau_{\hat{A}, \hat{B} \times \hat{C}} \circ [q, \langle \pi_1 q, {\bf let \: pure \:} y, z = {\bf pure \:} (\pi_1 (\pi_2 q)), \pi_2 (\pi_2 q) {\: \bf in \:} \langle y, z \rangle \rangle] \: \circ & \\
-&\quad\quad\quad\quad \circ \: [p, \langle \pi_1 (\pi_1 p), \langle \pi_2 (\pi_1 p), \pi_2 p \rangle \rangle] = & \\
-&\quad\quad\quad\quad\quad\quad \text{Composition}& \\
-&\tau_{\hat{A}, \hat{B} \times \hat{C}} \circ [p, \langle \pi_1 q, {\bf let \: pure \:} y, z = {\bf pure \:} (\pi_1 (\pi_2 q)), \pi_2 (\pi_2 q) {\: \bf in \:} \langle y, z \rangle \rangle & \\
-&\quad\quad\quad\quad [q := \langle \pi_1 (\pi_1 p), \langle \pi_2 (\pi_1 p), \pi_2 p \rangle \rangle]] = & \\
-&\quad\quad\quad\quad\quad\quad \text{Substitution and reduction}& \\
-&\tau_{\hat{A}, \hat{B} \times \hat{C}} \circ [p, \langle \pi_1 (\pi_1 p), {\bf let \: pure \:} y, z = {\bf pure \:} \pi_2 (\pi_1 p), \pi_2 p {\: \bf in \:} \langle y, z \rangle \rangle] = & \\
-&\quad\quad\quad\quad\quad\quad \text{Composition}& \\
-&[r, {\bf let \: pure \:} x, p^{'} = {\bf pure \:} \pi_1 r, \pi_2 r {\: \bf in \:} \langle x, \langle \pi_1 p^{'}, \pi_2 p^{'} \rangle\rangle] \: \circ & \\
-&\quad\quad\quad\quad \circ \: [p, \langle \pi_1 (\pi_1 p), {\bf let \: pure \:} y, z = {\bf pure \:} \pi_2 (\pi_1 p), \pi_2 p {\: \bf in \:} \langle y, z \rangle \rangle] = & \\
-&\quad\quad\quad\quad\quad\quad \text{Substitution and reduction}& \\
-&[p, {\bf let \: pure \:} x, p^{'} = ({\bf pure \:} (\pi_1 (\pi_1 p)), {\bf let \: pure \:} y, z = ({\bf pure \:} \pi_2 (\pi_1 p), \pi_2 p {\: \bf in \:} \langle y, z \rangle \rangle) & \\
-&\quad\quad\quad\quad {\: \bf in \:} \langle x, \langle \pi_1 p^{'}, \pi_2 p^{'} \rangle\rangle ] = & \\
-&\quad\quad\quad\quad\quad\quad \text{Reduction}& \\
-&[p, {\bf let \: pure \:} x,y,z = {\bf pure \:} (\pi_1 (\pi_1 p)), {\bf pure \:} \pi_2 (\pi_1 p), \pi_2 p {\: \bf in \:} \langle x, \langle \pi_1 p^{'}, \pi_2 p^{'} \rangle\rangle [p^{'} := \langle y, z \rangle]] = & \\
-&\quad\quad\quad\quad\quad\quad \text{Substitution and reduction}& \\
-&[p, {\bf let \: pure \:} x,y,z = {\bf pure \:} (\pi_1 (\pi_1 p)), {\bf pure \:} \pi_2 (\pi_1 p), \pi_2 p {\: \bf in \:} \langle x, \langle y, z \rangle\rangle]&
-\end{array}$
-
-\vspace{\baselineskip}
-
-2) Easy
-
-\end{proof}
-
 \begin{lemma}
 
-  {\bf K} is an applicative functor
+  $\langle \mathcal{C}(\lambda), \boxdot, \eta \rangle$ is an applicative functor
 \end{lemma}
@@
 
-Similar to \cite{Abramsky}, we apply the translation from $\lambda_{{\bf K}}$ to some cartesian closed category with an abritraty applicative functor $\mathcal{K}$, then
-we have $[\![\Gamma \vdash M : A]\!] = [x, M [x_i := \pi_i x]]$, so $M =_{\beta \eta} N \Leftrightarrow [\![\Gamma \vdash M : A]\!] = [\![\Gamma \vdash N : A]\!]$.
-
+\end{proof}
+
+\section{Relation with Moggi's monadic metalanguage}
+
+\begin{defin} Monadic metalanguage
+
+  Monadic metalanguage is a simply typed lambda calculus with additional typing rules:
+
+\begin{minipage}{0.45\textwidth}
+\begin{prooftree}
+  \AxiomC{$\Gamma \vdash M : A$}
+  \RightLabel{$\bigcirc_I$}
+  \UnaryInfC{$\Gamma \vdash {\bf val \:} M : \bigcirc A$}
+\end{prooftree}
+\end{minipage}%
+\hfill
+\begin{minipage}{0.45\textwidth}
+\begin{tabular}{p{\textwidth}}
+\begin{prooftree}
+  \AxiomC{$\Gamma \vdash M : \bigcirc A$}
+  \AxiomC{$\Gamma, x : A \vdash N : \bigcirc B$}
+  \RightLabel{$\bigcirc_E$}
+  \BinaryInfC{$\Gamma \vdash {\bf let \: val \:} x = M {\: \bf in \:} N : \bigcirc B$}
+\end{prooftree}
+\end{tabular}
+\end{minipage}
+
+\end{defin}
+
+\begin{defin} Reduction rules for monadic metalanguage
+  \begin{enumerate}
+    \item ${\bf let \: val \:} x = {\bf val \:} M {\: \bf in \:} N \rightarrow_{\beta} N [x := M]$;
+    \item ${\bf let \: val \:} x = ({\bf let \: val \:} y = N {\: \bf in \:} P) {\: \bf in \:} M \rightarrow_{\beta} {\bf let \: val \:} y = N {\: \bf in \:} ({\bf let \: val \:} x = P {\: \bf in \:} M)$;
+    \item ${\bf let \: val \:} x = M {\: \bf in \:} {\bf val} x \rightarrow_{\eta} M$.
+  \end{enumerate}
+\end{defin}
+
+Let us define translation $\ulcorner . \urcorner$ from $\lambda_{\text{IEL}^{-}}$ into monadic metalanguage.
+Without loss of generality we will consider $\lambda_{\text{IEL}^{-}}$ with $\to$ and $\Box$:
+
+\begin{defin} Translation for types
+  \begin{enumerate}
+    \item $\ulcorner A \urcorner = A$ for an atomic $A$;
+    \item $\ulcorner A \to B \urcorner = \ulcorner A \urcorner \to \ulcorner B \urcorner$;
+    \item $\ulcorner \Box A \urcorner = \bigcirc \ulcorner A \urcorner$
+  \end{enumerate}
+\end{defin}
+
+\begin{defin} Translation for terms
+  \begin{enumerate}
+    \item $\ulcorner x \urcorner = x$ for $x \in \mathbb{V}$;
+    \item $\ulcorner \lambda x. M \urcorner = \lambda x. \ulcorner M \urcorner$;
+    \item $\ulcorner M N \urcorner = \ulcorner M \urcorner \ulcorner N \urcorner$;
+    \item $\ulcorner {\bf box \:} M \urcorner = {\bf val \:} \ulcorner M \urcorner$
+    \item $\ulcorner {\bf let \: box \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} N \urcorner = {\bf val \:} \ulcorner N \urcorner$;
+    \item $\ulcorner {\bf let \: box \:} \vec{x} = \vec{M} {\: \bf in \:} N \urcorner = {\bf let \: val \:} \vec{x} = \ulcorner \vec{M} \urcorner {\: \bf in \:} {\bf val \:} \ulcorner N \urcorner$;
+  \end{enumerate}
+\end{defin}
+
+Where ${\bf let \: val \:} \vec{x} = \ulcorner \vec{M} \urcorner {\: \bf in \:} {\bf val \:} \ulcorner N \urcorner$ denotes
+${\bf let \: val \:} x_n = \ulcorner M_n \urcorner {\: \bf in \:} (\dots ({\bf let \: val \:} x_1 = \ulcorner M_1 \urcorner {\bf \: in \: } \ulcorner N \urcorner) \dots)$
+
+\begin{defin} Interpretation for modal rules
+  \begin{prooftree}
+    \AxiomC{$\ulcorner \Gamma \vdash M : A \urcorner = \ulcorner \Gamma \urcorner \vdash \ulcorner M \urcorner : \ulcorner A \urcorner$}
+    \UnaryInfC{$\ulcorner \Gamma \vdash {\bf box \:} M : \Box A \urcorner = \ulcorner \Gamma \urcorner \vdash {\bf val \:} \ulcorner M \urcorner : \bigcirc \ulcorner A \urcorner$}
+  \end{prooftree}
+
+  \begin{prooftree}
+    \AxiomC{$\ulcorner \Gamma \vdash \vec{M} : \Box \vec{A} \urcorner = \ulcorner \Gamma \urcorner \vdash \ulcorner \vec{M} \urcorner : \bigcirc \ulcorner \vec{A} \urcorner$}
+    \AxiomC{$\ulcorner \vec{x} : \vec{A} \vdash N : B \urcorner = \vec{x} : \ulcorner \vec{A} \urcorner \vdash \ulcorner N \urcorner : \ulcorner B \urcorner$}
+    \UnaryInfC{$\vec{x} : \ulcorner \vec{A} \urcorner \vdash {\bf val \:} \ulcorner N \urcorner : \bigcirc \ulcorner B \urcorner$}
+    \BinaryInfC{$\ulcorner \Gamma \vdash {\bf let \: box \:} \vec{x} = \vec{M} {\: \bf in \:} N : \Box B \urcorner = \ulcorner \Gamma \urcorner \vdash {\bf let \: val \:} \vec{x} = \ulcorner \vec{M} \urcorner {\: \bf in \:} {\bf val \:} \ulcorner N \urcorner : \bigcirc \ulcorner B \urcorner$}
+  \end{prooftree}
+\end{defin}
+
+\begin{lemma}
+  $ $
+
+  $\ulcorner M [x := N] \urcorner = \ulcorner M \urcorner [x := \ulcorner N \urcorner]$
+\end{lemma}
+
+\begin{proof}
+  Induction on the structure of $M$.
+\end{proof}
+
+\begin{lemma}
+  $ $
+
+  If $M =_{r} N$, then $\ulcorner M \urcorner =_{\beta \eta} \ulcorner N \urcorner$.
+\end{lemma}
+
+\begin{proof}
+$ $
+
+1) \\
+$\begin{array}{lll}
+&\ulcorner {\bf let \: box \:} x = ({\bf let \: box \:} \vec{y} = \vec{N} {\: \bf in \:} P) {\: \bf in \:} M \urcorner = & \\
+&\quad {\bf let \: val \:} x = ({\bf let \: val \:} \vec{y} = \ulcorner \vec{N} \urcorner {\: \bf in \:} {\bf val \:} \ulcorner P \urcorner) {\: \bf in \:} {\bf val \:} \ulcorner M \urcorner \rightarrow_{\beta} & \\
+&\quad {\bf let \: val \:} \vec{y} = \ulcorner \vec{N} \urcorner {\: \bf in \:} ({\bf let \: val \:} x = \ulcorner P \urcorner {\: \bf in \:} {\bf val \:} \ulcorner M \urcorner) \rightarrow_{\beta} & \\
+&\quad {\bf let \: val \:} \vec{y} = \ulcorner \vec{N} \urcorner {\: \bf in \:} {\bf val \:} \ulcorner M \urcorner [x := \ulcorner P \urcorner] = & \\
+&\ulcorner {\bf let \: box \:} \vec{y} = \vec{N} {\: \bf in \:} M [x := P] \urcorner&
+\end{array}$
+
+\vspace{\baselineskip}
+
+2) \\
+$\begin{array}{lll}
+&\ulcorner {\bf let \: box \:} \vec{x} = {\bf box \:} \vec{N} {\: \bf in \:} M \urcorner = & \\
+&\quad {\bf let \: val \:} \vec{x} = {\bf val \:} \ulcorner \vec{N} \urcorner {\: \bf in \:} {\bf val} \ulcorner M \urcorner \rightarrow_{\beta} & \\
+&\quad {\bf val \:} \ulcorner M \urcorner [\vec{x} := \ulcorner \vec{N} \urcorner] =& \\
+&\ulcorner {\bf box \:} M [\vec{x} := \vec{N}] \urcorner&
+\end{array}$
+
+\vspace{\baselineskip}
+
+3) \\
+$\begin{array}{lll}
+&\ulcorner {\bf let \: box \:} x = M {\: \bf in \:} x \urcorner = & \\
+&\quad {\bf let \: val \:} x = \ulcorner M \urcorner {\: \bf in \:} {\bf val \:} x \rightarrow_{\eta}& \\
+&\ulcorner M \urcorner&
+\end{array}$
+
+\end{proof}
+
+\begin{theorem}
+  $ $
+
+  If $\Gamma \vdash M : A$, then $\ulcorner \Gamma \urcorner \vdash \ulcorner M \urcorner : \ulcorner A \urcorner$
+\end{theorem}
+
+\begin{proof}
+  Follows from lemmas above.
 \end{proof}
@@
 
-  We are grateful to Valeria de Paiva, Neel Krishnaswami and Vladimir Vasyukov for consulting and advice.
-
-  Special thanks to Vladimir Krupski for general idea, critics and support.
+  We are grateful to Neel Krishnaswami, Vladimir Krupski, Valeria de Paiva, Valerii Plisko and Vladimir Vasyukov for consulting and advice.
 
@@
   URL: http://www.purescript.org. (Date: 1.08.2017) \parskip=1mm
-
-  \bibitem{Elm} Elm. A delightful language for reliable webapps. // URL: http://elm-lang.org. (Date:
-  1.08.2017) \parskip=1mm
 

### 2018-05-08 article.tex
update in article

@@
 \usepackage[all]{xy}
+\usepackage{authblk}
 \newtheorem{theorem}{Theorem}
@@
 \date{}
-\title{Modal type theory based on the intuitionistic epistemic logic}
+\author[1]{Daniel Rogozin}
+\affil[1]{Moscow State University, Moscow, Russia}
+\affil[1]{Serokell O\"{U}, Tallinn, Estonia}
+\title{Modal type theory as based on the intuitionistic modal logic IEL$^{-}$}
 
@@
 \begin{abstract}
-Modal intuitionistic epistemic logic IEL$^{-}$ was proposed by S.Artemov and T. Protopopescu as the formal
+Modal intuitionistic epistemic logic IEL$^{-}$ was proposed by S. Artemov and T. Protopopescu as the formal
 foundation for the intuitionistic theory of knowledge. We construct a modal simply typed lambda-calculus
-which is Curry-Howard isomorphic to IEL$^{-}$ as formal theory of calculations with applicative functors in
-functional programming languages like Haskell or Idris.
+which is Curry-Howard isomorphic to IEL$^{-}$ as the formal theory of calculations with applicative functors in
+functional programming.
 \end{abstract}
@@
   IEL$^{-}$  is a variant of IEL, that corresponds to intuitionistic belief.
-  Informally, $\Box A$ denotes that $A$ is verified intuitionistically.
+  Informally, $\Box A$ denotes that the agent believes in $A$.
 
@@
 
-  Functional programming languages such as Haskell \cite{Haskell}, Idris \cite{Idris} or Purescript \cite{Purs}
-  have special type classes\footnote{Type class in Haskell is a general interface for
-  special group of datatypes.} for calculations with container types like \verb"Functor" and
-  \verb"Applicative" \footnote{Reader may read more about container types in the Haskell standard library documentation\cite{Base} or in the next one textbook
-  \cite{Miran}}:
+  Functional programming languages such as Haskell, Idris or Purescript
+  have special type classes \footnote{Type class in Haskell is a general interface for
+  special group of datatypes.} for calculations with type operators like \verb"Functor" and
+  \verb"Applicative" \footnote{Reader may read more about these types in the Haskell standard library documentation.}:
 
@@
 
-  By \emph{container} (or \emph{computational context}) type we mean some type-operator $f$, where $f$ is a
+  By \emph{container} (or \emph{computational context}) type we mean some type operator $f$, where $f$ is a
   ``function'' from $*$ to $*$: type operator takes a simple type (ща kind $*$) and returns another
@@
   by constructing the type system (which is called $\lambda_{\text{IEL}^{-}}$) which is Curry-Howard isomorphic to $IEL^{-}$.
-  So we will consider IEL$^{-}$ modality as an arbitrary applicative functor and we prove that obtained type system
-  is sound and complete for applicative functor on cartesian closed category (using the categorical definition proposed by Paterson \cite{Cons}).
+  So we will consider IEL$^{-}$ modality as an arbitrary applicative functor and we prove that the obtained type system
+  is sound and complete for an applicative functor on cartesian closed category
+  (using the categorical definition proposed by Paterson \cite{Cons}).
 
@@
 
+  \begin{center}
+    \begin{prooftree}
+      \AxiomC{$ $}
+      \RightLabel{\bf{ax}}
+      \UnaryInfC{$\Gamma, A \vdash A$}
+    \end{prooftree}
+  \end{center}
+
   \begin{minipage}{0.5\textwidth}
@@
     \begin{prooftree}
+      \AxiomC{$\Gamma, A \vdash B$}
+      \RightLabel{$\to_I$}
+      \UnaryInfC{$\Gamma \vdash A \to B$}
+    \end{prooftree}
+    \begin{prooftree}
       \AxiomC{$\Gamma \vdash A$}
-    	\RightLabel{$\Box_I$}
+      \AxiomC{$\Gamma \vdash B$}
+      \RightLabel{$\land_I$}
+      \BinaryInfC{$\Gamma \vdash A \land B$}
+    \end{prooftree}
+  	\begin{prooftree}
+    	\AxiomC{$\Gamma \vdash A$}
+    	\RightLabel{${\Box_I}_1$}
     	\UnaryInfC{$\Gamma \vdash \Box A$}
@@
     \begin{prooftree}
+      \AxiomC{$\Gamma \vdash A \to B$}
+      \AxiomC{$\Gamma \vdash A$}
+      \RightLabel{$\to_E$}
+      \BinaryInfC{$\Gamma \vdash B$}
+    \end{prooftree}
+    \begin{prooftree}
+      \AxiomC{$\Gamma \vdash A_1 \land A_2$}
+      \RightLabel{$\land_E, i \in {1,2}$}
+      \UnaryInfC{$\Gamma \vdash A_i$}
+    \end{prooftree}
+  	\begin{prooftree}
     \AxiomC{$\Gamma \vdash \Box A_1, \dots, \Gamma \vdash \Box A_n $}
     \AxiomC{$A_1,\dots,A_n \vdash B$}
+    \RightLabel{${\Box_I}_2$}
     \BinaryInfC{$\Gamma \vdash \Box B$}
@@
 
-The first rule allows to derive co-reflextion. The second modal rule is a counterpart of $\Box_I$
+Without loss of generality we will consider the calculus without inference rule for disjunction and negation.
+Obtained calculus is equivalent to Hilbert-style IEL$^{-}$ without axioms for negation and disjunction.
+
+The first modal rule allows to derive co-reflextion. The second modal rule is a counterpart of $\Box_I$
 rule in natural deduction calculus for constructive K (see \cite{ModalLa}).
@@
 (3) &\bigwedge \limits_{i = 1}^{n} \Box A_i \rightarrow \Box \bigwedge \limits_{i = 1}^{n} A_i & \text{IEL$^{-}$ theorem} \\
-(4) &\bigwedge \Gamma \rightarrow \Box \bigwedge \limits_{i = 1}^{n} A_i & \text{from (1), (2) and transitivity} \\
+(4) &\bigwedge \Gamma \rightarrow \Box \bigwedge \limits_{i = 1}^{n} A_i & \text{from (2), (3) and transitivity} \\
 (5) &\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B& \text{assumption} \\
 (6) &(\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B) \rightarrow \Box (\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B)& \text{co-reflection}\\
-(7) &\Box (\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B)& \text{from (4), (5) and MP} \\
-(8) &\Box \bigwedge \limits_{i = 1}^{n} A_i \rightarrow \Box B & \text{from (6) and normality} \\
-(9) &\bigwedge \Gamma \rightarrow \Box B & \text{from (3), (7) and transitivity}
+(7) &\Box (\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B)& \text{from (5), (6) and MP} \\
+(8) &\Box \bigwedge \limits_{i = 1}^{n} A_i \rightarrow \Box B & \text{from (7) and normality} \\
+(9) &\bigwedge \Gamma \rightarrow \Box B & \text{from (4), (8) and transitivity}
 \end{array}$
@@
 \begin{proof}
-  Straightforward derivation of modal axioms in $\text{NIEL}^{-}$. We consider this derivation below using terms.
-\end{proof}
-
-  \vspace{\baselineskip}
-
-  At the next step we build the typed lambda-calculus based on the NIEL$^{-}$ with implication and dijunction
-by proof-assingment in rules. Obtained fragment is equivalent to IEL$^{-}$ without axioms for negation and disjunction.
+  Straightforward derivation of modal axioms in $\text{NIEL}^{-}$. We shall consider these derivations using terms below.
+\end{proof}
+
+  \vspace{\baselineskip}
+
+  At the next step we build the typed lambda-calculus based on the NIEL$^{-}$.
+by proof-assingment in rules.
 
@@
   \item $FV(\langle M,N \rangle) = FV(M) \cup FV(N)$;
-  \item $FV(\pi_i M) \subseteq FV(M)$, $i \in \{ 1, 2\}$;
+  \item $FV(\pi_i M) = FV(M)$, $i \in \{ 1, 2\}$;
   \item $FV(\text{\bf box } M) = FV(M)$;
@@
 
-1) The derivation ends in
-
-\begin{prooftree}
-  \AxiomC{$\Gamma \vdash \vec{M} : \Box \vec{A}$}
-  \AxiomC{$\vec{x} : \vec{A} \vdash N : B$}
-  \RightLabel{$\text{let}_{\Box}$}
-  \BinaryInfC{$\Gamma \vdash {\bf let \: box \:} \vec{x} = \vec{M} {\: \bf in \: } N : \Box B$}
-\end{prooftree}
-
-By IH $\Delta \vdash \vec{M} : \Box \vec{A}$, so $\Delta \vdash {\bf let \: box \:} \vec{x} = \vec{M} {\: \bf in \: } N : \Box B$.
-
-  \vspace{\baselineskip}
-
-2)-3) Similary.
+Induction on the derivation of $\Gamma \vdash M : A$.
 
@@
 
-1) If $\Gamma \vdash {\bf let \: box \:} \vec{x}, y, \vec{z} = \vec{M}, {\bf let \: box \: } \vec{w} = \vec{N} {\bf \: in \: } Q, \vec{P}  {\: \bf in \:} R : \Box B$,
+\begin{enumerate}
+
+\item If $\Gamma \vdash {\bf let \: box \:} \vec{x}, y, \vec{z} = \vec{M}, {\bf let \: box \: } \vec{w} = \vec{N} {\bf \: in \: } Q, \vec{P}  {\: \bf in \:} R : \Box B$,
   then $\Gamma \vdash {\bf let \: box \:} \vec{x}, \vec{w}, \vec{z} = \vec{M}, \vec{N}, \vec{P} \: { \bf in } \: R [y := Q] : \Box B$ by rule 4).
-
-2) Let $\Gamma \vdash {\bf let \: box \:} x = M {\: \bf in \: } x : \Box A$, then $\Gamma \vdash M : \Box A$ by rule 9).
-
-See \cite{ModalK}.
-
-3) The derivation ends in
-
+\item Let $\Gamma \vdash {\bf let \: box \:} x = M {\: \bf in \: } x : \Box A$,
+then $\Gamma \vdash M : \Box A$ by rule 9). \\ See \cite{ModalK}.
+\item The derivation ends in
 \begin{prooftree}
@@
 \end{prooftree}
-
 So $\Gamma \vdash \vec{M} : \vec{A}$ by Lemma 4 and $\Gamma \vdash N [\vec{x} := \vec{M}] : B$ by Lemma 4, part 3.
-
 Then we can transform this into the following derivation:
-
 \begin{prooftree}
@@
 \end{prooftree}
-
-4) The derivation ends in
-
+\item The derivation ends in
 \begin{prooftree}
@@
 \end{prooftree}
-
 So, if $\vdash M : A$, then $\vdash {\bf box \:} M : \Box A$.
-
+\end{enumerate}
 \end{proof}
@@
   Let us define the transformation from $\lambda_{\text{IEL}^{-}}$ into the simple type theory with
-  $\to$, $\times$ and natural number type $\mathbb{N}$ \footnote{Strong normalization for stronger system was shown here \cite{Girard}} with additional typing and reduction rules:
+  $\to$, $\times$ and natural number type $\mathbb{N}$ with additional typing and reduction rules
+  \footnote{Strong normalization for stronger system was shown here \cite{Girard}} :
 
@@
       \item $|{\bf let \: box \:} x = N {\: \bf in \:} M| = \langle \pi_1 |N| , |M| [x := \pi_2 |N|] \rangle$
-      \item $|{\bf let \: box \:} \vec{x} = \vec{N} {\: \bf in \:} M| = \langle \sum \limits_{i = 1}^n \pi_1 |N|, |M| [\vec{x} := \pi_2 \vec{N}] \rangle$
+      \item $|{\bf let \: box \:} \vec{x} = \vec{N} {\: \bf in \:} M| = \langle \sum \limits_{i = 1}^n \pi_1 |N|, |M| [\vec{x} := \pi_2 |\vec{N}|] \rangle$
     \end{enumerate}
@@
 
-Also we may consider four critical pairs which are used in local confluence proof for lambda-calculus based on {\bf IK} and \cite{ModalK}.
+Also we may consider four critical pairs which are considered in
+confluence proof for lambda-calculus as based on {\bf IK} \cite{ModalK}.
 
@@
 \begin{proof}
+  $ $
+
   By induction on the structure of $M$.
@@
     && (\mathcal{F}A \otimes_{\mathcal{D}} \mathcal{F}B) \otimes_{\mathcal{D}} \mathcal{F}C \ar[d]_{(A \varoast B) \otimes_{\mathcal{D}} id_{\mathcal{F}B}}
-    \ar[rr]^{\alpha^{\mathcal{D}}_{{\mathcal{F}A, \mathcal{F}B, \mathcal{F}C}}} && \mathcal{F}A \otimes_{\mathcal{D}} (\mathcal{F}B \otimes_{\mathcal{D}} \mathcal{F}C) \ar[d]^{id_{\mathcal{F}A} \otimes_{\mathcal{D}} \ast_{B,C}}\\
+    \ar[rr]^{\alpha^{\mathcal{D}}_{{\mathcal{F}A, \mathcal{F}B, \mathcal{F}C}}} && \mathcal{F}A \otimes_{\mathcal{D}} (\mathcal{F}B \otimes_{\mathcal{D}} \mathcal{F}C) \ar[d]^{id_{\mathcal{F}A} \otimes_{\mathcal{D}} B \varoast C}\\
     && \mathcal{F}(A \otimes_{\mathcal{C}} B) \otimes_{\mathcal{D}} \mathcal{C} \ar[d]_{(A \otimes_{\mathcal{C}} B) \varoast C} && \mathcal{F}A \otimes_{\mathcal{D}} \mathcal{F}(B \otimes_{\mathcal{C}} C) \ar[d]^{A \varoast (B \otimes_{\mathcal{C}} C) }\\
@@
 
-By default we will consider some monoidal functor on cartesian closed category below.
+By default we will consider an arbitrary monoidal functor on cartesian closed category below.
 
@@
 
-\begin{defin} Semantical translation from $\lambda_{\text{IEL}^{-}}$ to some cartesian closed category $\mathcal{C}$ with an applicative functor $\boxdot$:
+\begin{defin} Semantical translation from $\lambda_{\text{IEL}^{-}}$ to some cartesian closed category $\mathcal{C}$ with an applicative functor $\langle \mathcal{C}, \boxdot, \eta \rangle$:
 
@@
     \AxiomC{$[\![\vec{x} : \vec{A} \vdash N : B]\!] = [\![N]\!] : \prod \limits_{i=1}^n [\![A_i]\!] \rightarrow [\![B]\!]$}
-    \BinaryInfC{$[\![\Gamma \vdash {\bf let \: box \:} \vec{x} = \vec{M} {\: \bf in \: } M : \Box B]\!] = \boxdot ([\![N]\!]) \circ \ast_{[\![A_1]\!],\dots,[\![A_n]\!]} \circ \langle [\![M_1]\!],\dots, [\![M_n]\!] \rangle : [\![\Gamma]\!] \rightarrow \boxdot [\![B]\!]$}
+    \BinaryInfC{$[\![\Gamma \vdash {\bf let \: box \:} \vec{x} = \vec{M} {\: \bf in \: } M : \Box B]\!] = \boxdot ([\![N]\!]) \circ ([\![A_1]\!] \varoast \dots \varoast [\![A_n]\!]) \circ \langle [\![M_1]\!],\dots, [\![M_n]\!] \rangle : [\![\Gamma]\!] \rightarrow \boxdot [\![B]\!]$}
   \end{prooftree}
@@
 \end{defin}
+
+Interpretation for ${\bf let \: box\:}$-rule is similar to interpretation for $\Box$-rule in term calculus for intutionistic {\bf K} \cite{Paiva}.
 
@@
 &\quad\quad\quad\quad\quad\quad\quad  \text{Interpretation for $let_{\Box}$} \\
-&\boxdot ([\![N]\!]) \circ ([\![A_1]\!] \ast \dots \ast [\![A_n]\!]) \circ [\![\Gamma \vdash (\vec{M} [\vec{y} := \vec{P}]) : \Box \vec{A}]\!] =& \\
+&\boxdot ([\![N]\!]) \circ ([\![A_1]\!] \varoast \dots \varoast [\![A_n]\!]) \circ [\![\Gamma \vdash (\vec{M} [\vec{y} := \vec{P}]) : \Box \vec{A}]\!] =& \\
 &\quad\quad\quad\quad\quad\quad\quad \text{Induction hypothesis}& \\
-&\boxdot ([\![N]\!]) \circ ([\![A_1]\!] \ast \dots \ast [\![A_n]\!]) \circ ([\![\vec{M}]\!] \circ \langle [\![P_1]\!],\dots,[\![P_n]\!]\rangle) = & \\
+&\boxdot ([\![N]\!]) \circ ([\![A_1]\!] \varoast \dots \varoast [\![A_n]\!]) \circ ([\![\vec{M}]\!] \circ \langle [\![P_1]\!],\dots,[\![P_n]\!]\rangle) = & \\
 &\quad\quad\quad\quad\quad\quad\quad \text{Associativity of composition}& \\
-&(\boxdot ([\![N]\!]) \circ ([\![A_1]\!] \ast \dots \ast [\![A_n]\!]) \circ [\![\vec{M}]\!]) \circ \langle [\![P_1]\!],\dots,[\![P_n]\!]\rangle = & \\
+&(\boxdot ([\![N]\!]) \circ ([\![A_1]\!] \varoast \dots \varoast [\![A_n]\!]) \circ [\![\vec{M}]\!]) \circ \langle [\![P_1]\!],\dots,[\![P_n]\!]\rangle = & \\
 &\quad\quad\quad\quad\quad\quad\quad \text{Interpretaion}& \\
@@
 &\quad\quad\quad\quad\quad\quad\quad \text{By the definition of an applicative functor}& \\
-&\boxdot {K}([\![M]\!]) \circ \eta_{\mathds{1}} = & \\
+&\boxdot ([\![M]\!]) \circ \eta_{\mathds{1}} = & \\
 &\quad\quad\quad\quad\quad\quad\quad \text{By naturality of $\eta$} & \\
@@
   \item $\widehat{A \to B} = \hat{B}^{\hat{A}}$;
-  \item Evaluation arrow $\epsilon = [x, (\pi_2 x) (\pi_1 x)] \in Hom_{\mathcal{C}(\lambda)(\hat{B}^{\hat{A}} \times \hat{A}, \hat{B})}$.
+  \item Evaluation arrow $\epsilon_{\hat{A},\hat{B}} = [x, (\pi_2 x) (\pi_1 x)] \in Hom_{\mathcal{C}(\lambda)}(\hat{B}^{\hat{A}} \times \hat{A}, \hat{B})$.
 \end{enumerate}
@@
 We define endofunctor $\boxdot$ on $\mathcal{C}(\lambda)$ and natural transformation $\eta$ from $id_{\mathcal{C}(\lambda)}$ to this endofunctor.
-It is sufficient to show $\boxdot$ and $\eta$ form an applicative functor on $\mathcal{C}(\lambda)$ for theorem proof.
+It is sufficient to show $\boxdot$ and $\eta$ form the relevant structure on $\mathcal{C}(\lambda)$.
 
@@
   Let us define an endofunctor $\boxdot : \mathcal{C}(\lambda) \to \mathcal{C}(\lambda)$, such that
-forall $[x,M] \in Hom_{\mathcal{C}(\lambda)}(\hat{A},\hat{B}), \boxdot ([x,M]) = [y, {\bf let \: box \:} x = y {\: \bf in \:} M] \in Hom_{\mathcal{C}(\lambda)}(\boxdot \hat{A}, \boxdot \hat{B})$
-(denotation: $\text{fmap } f$ for an arbitrary arrow $f$).
+forall $[x,M] \in Hom_{\mathcal{C}(\lambda)}(\hat{A},\hat{B}), \boxdot ([x,M]) = [y, {\bf let \: box \:} x = y {\: \bf in \:} M] \in Hom_{\mathcal{C}(\lambda)}(\boxdot \hat{A}, \boxdot \hat{B})$.
 
@@
 \begin{enumerate}
-  \item $\text{fmap }(g \circ f) = \text{fmap }(g) \circ \text{fmap }(f)$;
-  \item $\text{fmap }(id_{\hat{A}}) = id_{\boxdot}\hat{A}$.
+  \item $\boxdot(g \circ f) = \boxdot g \circ \boxdot f$;
+  \item $\boxdot (id_{\hat{A}}) = id_{\boxdot}\hat{A}$.
 \end{enumerate}
@@
 t.
-$\forall \hat{A}, \hat{B} \in Ob_{\mathcal{C}(\lambda)}, \ast_{\hat{A},\hat{B}} = [p, {\bf let \: box \:} x,y = \pi_1 p, \pi_2 p {\: \bf in \:} \langle x, y \rangle] \in Hom_{\mathcal{C}(\lambda)}(\boxdot \hat{A} \times \boxdot \hat{B}, \boxdot (\hat{A} \times \hat{B}))$.
-\end{enumerate}
-\end{defin}
-
-Implementation for $\ast$ in our term model is the instance of ${\text{let}_{\Box}}$-rule:
+$\forall \hat{A}, \hat{B} \in Ob_{\mathcal{C}(\lambda)}, \hat{A} \varoast \hat{B} = [p, {\bf let \: box \:} x,y = \pi_1 p, \pi_2 p {\: \bf in \:} \langle x, y \rangle] \in Hom_{\mathcal{C}(\lambda)}(\boxdot \hat{A} \times \boxdot \hat{B}, \boxdot (\hat{A} \times \hat{B}))$.
+\end{enumerate}
+\end{defin}
+
+Implementation for $\varoast$ in our term model is the instance of ${\text{let}_{\Box}}$-rule:
 
@@
 \begin{enumerate}
-  \item $\text{fmap } f \circ \eta_A = \eta_B \circ f$;
+  \item $\boxdot f \circ \eta_A = \eta_B \circ f$;
   \item $(\hat{A} \varoast \hat{B}) \circ (\eta_{A} \times \eta_{B}) = \eta_{\hat{A} \times \hat{B}}$;
@@
 \begin{enumerate}
-\item $\text{fmap } f \circ \eta_{\hat{A}} = \eta_{\hat{B}} \circ f$ \\
+\item $\boxdot f \circ \eta_{\hat{A}} = \eta_{\hat{B}} \circ f$ \\
 
@@
 &\text{On the other hand:}& \\
-&\text{fmap } f \circ \eta_{\hat{A}} = & \\
+&\boxdot f \circ \eta_{\hat{A}} = & \\
 &[z, {\bf let \: box \:} x = z {\: \bf in \:} M] \circ [x, {\bf box \:} x] = &\\
@@
 
-  Monadic metalanguage is a simply typed lambda calculus with additional typing rules:
+  Monadic metalanguage is a simply typed lambda calculus with additional typing rules \cite{Moggi}:
 
@@
 
-Where ${\bf let \: val \:} \vec{x} = \ulcorner \vec{M} \urcorner {\: \bf in \:} {\bf val \:} \ulcorner N \urcorner$ denotes
-${\bf let \: val \:} x_n = \ulcorner M_n \urcorner {\: \bf in \:} (\dots ({\bf let \: val \:} x_1 = \ulcorner M_1 \urcorner {\bf \: in \: } \ulcorner N \urcorner) \dots)$
-
+Where $\ulcorner \Gamma \urcorner \vdash {\bf let \: val \:} \vec{x} = \ulcorner \vec{M} \urcorner {\: \bf in \:} {\bf val \:} \ulcorner N \urcorner$
+denotes ${\bf let \: val \:} x_1 = \ulcorner M_1 \urcorner {\: \bf in \:} ( \dots {\: \bf in \:} ({\bf let \: val \:} x_n = \ulcorner M_n \urcorner {\: \bf in \:} {\bf val \:} \ulcorner N \urcorner) \dots )$.
 \begin{defin} Interpretation for modal rules
@@
 
+  The research described in this paper was supported by Russian Foundation for Basic Research (grant 16-03-00364).
+
   \addcontentsline{toc}{section}{References}
@@
 
-  \bibitem{Artemov} Artemov S. and Protopopescu T., \/ ``Intuitionistic Epistemic Logic'', \textit{The
-  Review of Symbolic Logic}, 2016, vol. 9, no 2. pp. 266-298.\parskip=1mm
-
-  \bibitem{Krupski} Krupski V. N. and Yatmanov A., \/ ``Sequent Calculus for Intuitionistic Epistemic Logic
-  IEL'', \textit{Logical Foundations of Computer Science: International Symposium, LFCS 2016, Deerfield
-  Beach, FL, USA, January 4-7, 2016. Proceedings}, 2016, pp. 187-201.\parskip=1mm
-
-  \bibitem{Haskell} Haskell Language. // URL: https://www.haskell.org. (Date: 1.08.2017) \parskip=1mm
-
-  \bibitem{Idris} Idris. A Language with Dependent Types.// URL:https://www.idris-lang.org. (Date:
-  1.08.2017) \parskip=1mm
-
-  \bibitem{Purs} Purescript. A strongly-typed functional programming language that compiles to JavaScript.
-  URL: http://www.purescript.org. (Date: 1.08.2017) \parskip=1mm
-
-  \bibitem{Base} Hackage, \/ ``The base package'' // URL: https://hackage.haskell.org/package/base-4.10.0.0
-  (Date: 1.08.2017) \parskip=1mm
-
-  \bibitem{Miran} Lipovaca M, \/ ``Learn you a Haskell for Great Good!''. //URL:
-  http://learnyouahaskell.com/chapters (Date: 1.08.2017) \parskip=1mm
-
-  \bibitem{McP} McBride C. and Paterson R., ``Applicative programming with effects", \textit{Journal of
-  Functional Programming}, 2008, vol. 18, no 01. pp 1-13. \parskip=1mm
-
-  \bibitem{McP2} McBride C. and Paterson R, ``Functional Pearl. Idioms: applicative programming with
-  effects'', \textit{Journal of Functional Programming}, 2005. vol. 18, no 01. pp 1-20. \parskip=1mm
-
-  \bibitem{Neder} R. Nederpelt and H. Geuvers, ``Type Theory and Formal Proof: An Introduction''.
-  \textit{Cambridge University Press}, New York, NY, USA, 2014. pp. 436. \parskip=1mm
-
-  \bibitem{Morten} Sorensen M. H. and Urzyczyn P, ``Lectures on the Curry-Howard isomorphism'',
-  \textit{Studies in Logic and the Foundations of Mathematics}, vol. 149, \textit{Elsevier Science}, 1998.
-  pp 261. \parskip=1mm
-
-  \bibitem{Pierce} Pierce B. C., ``Types and Programming Languages''. \textit{Cambridge, Mass: The MIT
-  Press}, 2002. pp. 605. \parskip=1mm
-
-  \bibitem{Girard} Girard J.-Y., Taylor P. and  Lafont Y, ``Proofs and Types'', \textit{Cambridge University
-  Press}, New York, NY, USA, 1989. pp. 175. \parskip=1mm
-
-  \bibitem{Baren} Barendregt. H. P., ``Lambda calculi with types" // Abramsky S., Gabbay Dov M., and S. E.
-  Maibaum, ``Handbook of logic in computer science (vol. 2), Osborne Handbooks Of Logic In Computer
-  Science'', Vol. 2. \textit{Oxford University Press, Inc.}, New York, NY, USA, 1993. pp 117-309.
-  \parskip=1mm
-
-  \bibitem{Hindley} Hindley J. Roger, ``Basic Simple Type Theory''. \textit{Cambridge University Press}, New
-  York, NY, USA, 1997. pp. 185. \parskip=1mm
-
-  \bibitem{Lax} Pfenning F. and Davies R., ``A judgmental reconstruction of modal logic'',
-  \textit{Mathematical Structures in Computer Science}, vol. 11, no 4, 2001, pp. 511-540. \parskip=1mm
-
-  \bibitem{Baren2} H.P. Barendregt. The Lambda Calculus --- Its Syntax and Semantics. Studies in Logic and
-  the Foundations of Mathematics, vol. 103. Amsterdam: North-Holland, 1985.
-
-  \bibitem{ModalK} Yoshihiko KAKUTANI, A Curry-Howard Correspondence for Intuitionistic Normal Modal Logic, Computer Software, Released February 29, 2008, Online ISSN , Print ISSN 0289-6540.
-
-  \bibitem{ModalK1} Kakutani Y.
-(2007) Call-by-Name and Call-by-Value in Normal Modal Logic.
-In: Shao Z.
-(eds) Programming Languages and Systems.
-APLAS 2007.
-Lecture Notes in Computer Science, vol 4807.
-Springer, Berlin, Heidelberg
-
-  \bibitem{Abe} T. Abe. Completeness of modal proofs in first-order predicate logic. Computer Software, JSSST Journal, 24:165 -- 177, 2007.
-
-  \bibitem{LambekScott} Lambek, J. and Scott P.J. (1986) Introduction to Higher Order Categorical Logic, Cambridge Studies in Advanced Mathematics 7, Cambridge: Cambridge University Press.
-
-  \bibitem{ElKelly} Samuel Eilenberg and Max Kelly, Closed categories. Proc. Conf. Categorical Algebra (La Jolla, Calif., 1965).
-
-  \bibitem{Abramsky} Samson Abramsky and Nikos Tzevelekos, Introduction to Categories and Categorical Logic
-
-  \bibitem{ModalLa} G. A. Kavvos. The Many Worlds of Modal $\Lambda$--calculi: I. Curry-Howard for Necessity, Possibility and Time
-
-  \bibitem{Cons} Ross Paterson. in Mathematics of Program Construction, Madrid, 2012, Lecture Notes in Computer Science, vol. 7342, pp. 300--323, Springer, 2012.
+  \bibitem{Abe} T. Abe, Completeness of modal proofs in first-order predicate logic, Computer Software, JSSST Journal. 24 (2007).
+
+  \bibitem{Abramsky} S. Abramsky and N. Tzevelekos, Introduction to Categories and Categorical Logic, in: B. Coecke, ed., New Structures for Physics.
+  Lecture Notes in Physics, vol. 813 (Springer-Verlag, Berlin, 2011).
+
+  \bibitem{Artemov} S. Artemov and T. Protopopescu, Intuitionistic Epistemic Logic, The
+  Review of Symbolic Logic. 9 (2016).
+
+  \bibitem{Awodey} S. Awodey, Category Theory (2nd ed.), Oxford University Press, 2010.
+
+  \bibitem{Baren2} H. P. Barendregt, The Lambda Calculus --- Its Syntax and Semantics, North-Holland, 1985.
+
+  \bibitem{Baren} H.P. Barendregt. Lambda calculi with types, in: S. Abramsky, Dov M. Gabbay, and S. E.
+  Maibaum, eds., Handbook of logic in computer science (vol. 2), Osborne Handbooks Of Logic In Computer
+  Science, vol. 2. (Oxford University Press, New York, 1993).
+
+  \bibitem{Paiva} G. Bellin, V. C. V. de Paiva, and E. Ritter. Extended Curry-Howard correspondence for a basic constructive modal logic, Proceedings of Methods for Modalities, 2001.
+
+  \bibitem{ElKelly} S. Eilenberg and M. Kelly. Closed categories, Proc. Conf. Categorical Algebra, La Jolla, (1965).
+
+  \bibitem{Girard} J.-Y. Girard, P. Taylor and Y. Lafont, Proofs and Types, Cambridge University
+  Press, 1989.
+
+  \bibitem{Hindley} J. Roger Hindley, Basic Simple Type Theory, Cambridge University Press, 1997.
+
+  \bibitem{ModalK} Y. Kakutani. A Curry-Howard Correspondence for Intuitionistic Normal Modal Logic. // Computer Software, 2008.
+
+  \bibitem{ModalK1} Y. Kakutani, Call-by-Name and Call-by-Value in Normal Modal Logic, in: Shao Z. (eds)
+  Programming Languages and Systems, Lecture Notes in Computer Science, vol 4807 (2016).
+
+  \bibitem{ModalLa} G. A. Kavvos. The Many Worlds of Modal $\lambda$--calculi: I. Curry-Howard for Necessity, Possibility and Time. // [Электронный ресурс] URL: https://arxiv.org/abs/1605.08106
+
+  \bibitem{Krupski} V. N. Krupski and A. Yatmanov. Sequent Calculus for Intuitionistic Epistemic Logic IEL,
+   in: Artemov S., Nerode A. (eds), Logical Foundations of Computer Science. Lecture Notes in Computer Science, vol 9537. Springer, Cham (2016).
+
+  \bibitem{LambekScott} J. Lambek and P.J. Scott, Introduction to Higher Order Categorical Logic, Cambridge University Press, 1986.
+
+  \bibitem{MacLane} S. MacLane. Categories for the Working Mathematician, Graduate Texts in Mathematics, Springer-Verlag, 1998.
+
+  \bibitem{McP} C. McBride and R. Paterson, Applicative programming with effects, Journal of
+  Functional Programming, 18 (2008).
+
+  \bibitem{McP2} C. McBride and R. Paterson, Functional Pearl. Idioms: applicative programming with
+  effects, Journal of Functional Programming, 18 (2005).
+
+  \bibitem{Moggi} E. Moggi, Notions of computation and monads, Inf. Comput., 93 (1) (1991).
+
+  \bibitem{Neder} R. Nederpelt and H. Geuvers, Type Theory and Formal Proof: An Introduction.
+  Cambridge University Press, 2014.
+
+  \bibitem{Cons} R. Paterson, Constructing applicative functors,
+  in: Mathematics of Program Construction, Lecture Notes in Computer Science (Springer, Berlin, 2012).
+
+  \bibitem{Lax} F. Pfenning and R. Davies. A judgmental reconstruction of modal logic.
+  Mathematical Structures in Computer Science, vol. 11 (2001).
+
+  \bibitem{Pierce} B. C. Pierce, Types and Programming Languages, The MIT
+  Press, 2002.
+
+  \bibitem{Morten} M. H. Sorensen and P. Urzyczyn, Lectures on the Curry-Howard isomorphism, Elsevier Science, 1998.
+  pp 261.
+
+  \bibitem{Troelstra} A.S. Troelstra and H. Schwichtenberg, Basic Proof Theory, Cambridge University Press, 1996.
   \end{thebibliography}

### 2018-05-08 article.tex
another one update

@@
       \AxiomC{$\Gamma \vdash A_1 \land A_2$}
-      \RightLabel{$\land_E, i \in {1,2}$}
+      \RightLabel{$\land_E, i = 1,2$}
       \UnaryInfC{$\Gamma \vdash A_i$}
@@
     \AxiomC{ $\Gamma \vdash M : A_1 \times A_2$ }
-    \RightLabel{$\times_e$, $i \in \{ 1, 2 \}$}
+    \RightLabel{$\times_e$, $i = 1, 2$}
     \UnaryInfC{$\Gamma \vdash \pi_i M : A_i$}
@@
 \AxiomC{$x : A \vdash x : A$}
+\RightLabel{$\Box_I$}
 \UnaryInfC{$x : A \vdash {\bf box \:} x : \Box A$}
+\RightLabel{$\to_I$}
 \UnaryInfC{$\vdash (\lambda x. {\bf box \: } x) : A \to \Box A$}
@@
 \TrinaryInfC{$f : \Box (A \to B), x : \Box A \vdash {\bf let \: box \:} g, y  = f, x {\: \bf in \:} g y : \Box B$}
+\RightLabel{$\to_I$}
 \UnaryInfC{$f : \Box (A \to B) \vdash \lambda x. {\bf let \: box \:} g, y = f, x {\: \bf in \:} g y : \Box A \to \Box B$}
+\RightLabel{$\to_I$}
 \UnaryInfC{$\vdash \lambda f. \lambda x. {\bf let \: box \:} g, y = f, x {\: \bf in \:} g y : \Box (A \to B) \to \Box A \to \Box B$}
@@
   \item $FV(\langle M,N \rangle) = FV(M) \cup FV(N)$;
-  \item $FV(\pi_i M) = FV(M)$, $i \in \{ 1, 2\}$;
+  \item $FV(\pi_i M) = FV(M)$, $i = 1, 2$;
   \item $FV(\text{\bf box } M) = FV(M)$;
@@
   \item $(M, N)[x := P] = (M[x := P], N [x := P])$;
-  \item $(\pi_i M) [x := P] = \pi_i (M[x := P])$, $i \in \{ 1, 2\}$;
+  \item $(\pi_i M) [x := P] = \pi_i (M[x := P])$, $i = 1, 2$;
   \item $({\bf box \: } M) [x := P] = {\bf box \: } (M [x := P])$;
@@
   \item ${\bf let \: box \:} \vec{x}, y, \vec{z} = \vec{M}, {\bf let \: box \: } \vec{w} = \vec{N} {\: \bf in \: } Q, \vec{P} {\: in \:} R \rightarrow_{\beta} \\
-  {\bf let \: box \:} \vec{x}, \vec{w}, \vec{z} = \vec{M}, \vec{N}, \vec{P} {\: \bf in \: } R [y := Q]$
-  \item ${\bf let \: box \:} \vec{x} = {\bf box \:} \vec{M} {\: \bf in \:} N \rightarrow_{\beta} {\bf box \:} N [\vec{x} := \vec{M}]$
-  \item ${\bf let \: box \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} M \rightarrow_{\beta} {\bf box \:} M$, where \underline{\quad} is an empty sequence of terms.
+  {\bf let \: box \:} \vec{x}, \vec{w}, \vec{z} = \vec{M}, \vec{N}, \vec{P} {\: \bf in \: } R [y := Q]$;
+  \item ${\bf let \: box \:} \vec{x} = {\bf box \:} \vec{M} {\: \bf in \:} N \rightarrow_{\beta} {\bf box \:} N [\vec{x} := \vec{M}]$;
+  \item ${\bf let \: box \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} M \rightarrow_{\beta} {\bf box \:} M$, where \underline{\quad} is an empty sequence of terms;
   \item $\lambda x. f x \rightarrow_{\eta} f$;
@@
 
-Now we will prove standard lemmas for contexts in type systems\footnote{We will not prove cases with
-$\to$-constructor, they are proved standardly in the same lemmas for simply typed lambda calculus, for
-example see \cite{Neder} \cite{Morten}. We will consider only modal cases}:
+Now we will prove standard lemmas for contexts in type systems\footnote{Cases with
+$\to$-constructor are proven standardly in similar lemmas for simply typed lambda calculus, for
+example see \cite{Neder} \cite{Morten} \cite{Pierce}. We will consider only modal cases}:
 
@@
   \item If $\Gamma \vdash M : A$ and $\Gamma \subseteq \Delta$, then $\Delta \vdash M : A$;
-  \item If $\Gamma \vdash M : A$, then $\Delta \vdash M : A$, where $\Delta = \{ x_i : A_i \: | \: (x_i : A_i) \in \Gamma \: \& \: x_i \in FV(M) \}$
+  \item If $\Gamma \vdash M : A$, then $\Delta \vdash M : A$, where $\Delta = \{ x : A \: | \: (x : A) \in \Gamma \: \& \: x \in FV(M) \}$
   \item If $\Gamma, x : A \vdash M : B$ and $\Gamma \vdash N : A$, then $\Gamma \vdash M [x := N] : B$.
@@
 
-Induction on the derivation of $\Gamma \vdash M : A$.
+1--3) Induction on the derivation of $\Gamma \vdash M : A$.
 
@@
 
-  If $\Gamma \vdash M : A$ and $M \twoheadrightarrow_r N$, then $\Gamma \vdash N : A$
+  If $\Gamma \vdash M : A$ and $M \twoheadrightarrow_r N$, then $\Gamma \vdash N : A$.
 
@@
 Induction on the derivation $\Gamma \vdash M : A$ and on the generation of $\rightarrow_r$.
-
-For cases with axiom, application, abstraction and pairs see \cite{Morten} \cite{Pierce}.
 
@@
   Let us define the transformation from $\lambda_{\text{IEL}^{-}}$ into the simple type theory with
-  $\to$, $\times$ and natural number type $\mathbb{N}$ with additional typing and reduction rules
-  \footnote{Strong normalization for stronger system was shown here \cite{Girard}} :
+  $\to$, $\times$ and natural number type $\mathbb{N}$ with additional typing and reduction rules\footnote{Strong normalization for stronger system was shown here \cite{Girard}}:
 
@@
   \AxiomC{$ $}
+  \RightLabel{{\bf zero}}
   \UnaryInfC{$\Gamma \vdash 0 : \mathbb{N}$}
@@
 
+\begin{minipage}{0.45\textwidth}
+\begin{flushleft}
   \begin{prooftree}
   \AxiomC{$\Gamma \vdash n : \mathbb{N}$}
+  \RightLabel{{\bf succ}}
   \UnaryInfC{$\Gamma \vdash {\bf succ \:} n : \mathbb{N}$}
   \end{prooftree}
-
+\end{flushleft}
+\end{minipage}
+\begin{minipage}{0.45\textwidth}
+\begin{flushright}
   \begin{prooftree}
@@
   \AxiomC{$\Gamma \vdash m : \mathbb{N}$}
+  \RightLabel{{\bf add}}
   \BinaryInfC{$\Gamma \vdash n + m : \mathbb{N}$}
   \end{prooftree}
+\end{flushright}
+\end{minipage}
 
@@
       \item $|\langle M, N \rangle| = \langle |M|, |N| \rangle$;
-      \item $|\pi_i M| = \pi_i |M|$, $i \in \{ 1, 2\}$;
+      \item $|\pi_i M| = \pi_i |M|$, $i = 1, 2$;
       \item $|{\bf box \:} M| = \langle 0, |M| \rangle$;
@@
 \begin{proof}
+  $ $
 
@@
 
-  Monoidal category is a category $\mathcal{C}$ with additional monoidal structure:
+  Monoidal category is a category $\mathcal{C}$ with the following additional structure:
   \begin{enumerate}
@@
   A monoidal functor $\mathcal{F} : \langle \mathcal{C}, \otimes_1, \mathds{1} \rangle \to \langle \mathcal{D}, \otimes_2, \mathds{1}' \rangle$ is a functor
-  $\mathcal{F} : \mathcal{C} \to \mathcal{D}$ with additional natural transformations:
+  $\mathcal{F} : \mathcal{C} \to \mathcal{D}$ with natural transformations:
 
@@
 \BinaryInfC{$[\![\Gamma \vdash (M N) : B]\!] = [\![\Gamma]\!] \xrightarrow{\langle [\![M]\!], [\![N]\!]
-\rangle} [\![B]\!]^{[\![A]\!]} \times [\![A]\!] \xrightarrow{\epsilon} [\![B]\!] $}
+\rangle} [\![B]\!]^{[\![A]\!]} \times [\![A]\!] \xrightarrow{\epsilon_{[\![A]\!],[\![B]\!]}} [\![B]\!] $}
 \end{prooftree}
@@
 [\![A_2]\!]$}
-\RightLabel{$i \in \{1,2\}$}
+\RightLabel{$i = 1,2$}
 \UnaryInfC{$[\![\Gamma \vdash \pi_i M : A_i]\!] = [\![\Gamma]\!] \xrightarrow{[\![M]\!]} [\![A_1]\!] \times
@@
   \item $\widehat{A \times B} = \hat{A} \times \hat{B}$;
-  \item Canonical projection is defined as $[x, \pi_i x] \in Hom_{\mathcal{C}(\lambda)}(\hat{A_1} \times \hat{A_2},\hat{A_i})$ for $i \in \{ 1, 2 \}$;
+  \item Canonical projection is defined as $[x, \pi_i x] \in Hom_{\mathcal{C}(\lambda)}(\hat{A_1} \times \hat{A_2},\hat{A_i})$ for $i = 1, 2$;
   \item $\widehat{A \to B} = \hat{B}^{\hat{A}}$;
-  \item Evaluation arrow $\epsilon_{\hat{A},\hat{B}} = [x, (\pi_2 x) (\pi_1 x)] \in Hom_{\mathcal{C}(\lambda)}(\hat{B}^{\hat{A}} \times \hat{A}, \hat{B})$.
+  \item Evaluation arrow $\epsilon_{\hat{A},\hat{B}} = [x, (\pi_1 x) (\pi_2 x)] \in Hom_{\mathcal{C}(\lambda)}(\hat{B}^{\hat{A}} \times \hat{A}, \hat{B})$.
 \end{enumerate}
@@
   \item $\boxdot(g \circ f) = \boxdot g \circ \boxdot f$;
-  \item $\boxdot (id_{\hat{A}}) = id_{\boxdot}\hat{A}$.
+  \item $\boxdot (id_{\hat{A}}) = id_{\boxdot \hat{A}}$.
 \end{enumerate}
@@
 \begin{enumerate}
-  \item $\eta: Id_{\mathcal{C}(\lambda)} \Rightarrow \boxdot$, s.
+  \item $\eta: Id_{\mathcal{C}(\lambda)} \Rightarrow \boxdot\text{, s.
 t.
-$\forall \hat{A} \in Ob_{\mathcal{C}(\lambda)}$, $\eta_{\hat{A}} = [x, {\bf box \:} x] \in Hom_{\mathcal{C}(\lambda)}(\hat{A}, \boxdot \hat{A})$;
-  \item $\hat{A} \varoast \hat{B} : \boxdot \hat{A} \times \boxdot \hat{B} \to \boxdot (\hat{A} \times \hat{B})$, s.
+}\forall \hat{A} \in Ob_{\mathcal{C}(\lambda)}$, $\eta_{\hat{A}} = [x, {\bf box \:} x] \in Hom_{\mathcal{C}(\lambda)}(\hat{A}, \boxdot \hat{A})$;
+  \item $\hat{A} \varoast \hat{B} : \boxdot \hat{A} \times \boxdot \hat{B} \to \boxdot (\hat{A} \times \hat{B})\text{, s.
 t.
-$\forall \hat{A}, \hat{B} \in Ob_{\mathcal{C}(\lambda)}, \hat{A} \varoast \hat{B} = [p, {\bf let \: box \:} x,y = \pi_1 p, \pi_2 p {\: \bf in \:} \langle x, y \rangle] \in Hom_{\mathcal{C}(\lambda)}(\boxdot \hat{A} \times \boxdot \hat{B}, \boxdot (\hat{A} \times \hat{B}))$.
+} \forall \hat{A}, \hat{B} \in Ob_{\mathcal{C}(\lambda)}, \hat{A} \varoast \hat{B} = [p, {\bf let \: box \:} x,y = \pi_1 p, \pi_2 p {\: \bf in \:} \langle x, y \rangle] \in Hom_{\mathcal{C}(\lambda)}(\boxdot \hat{A} \times \boxdot \hat{B}, \boxdot (\hat{A} \times \hat{B}))$.
 \end{enumerate}
@@
 
+\begin{small}
 \begin{prooftree}
 \AxiomC{$p : \Box A \times \Box B \vdash p : \Box A \times \Box B$}
+\RightLabel{$\times_E$}
 \UnaryInfC{$p : \Box A \times \Box B \vdash \pi_1 p : \Box A$}
 \AxiomC{$p : \Box A \times \Box B \vdash p : \Box A \times \Box B$}
+\RightLabel{$\times_E$}
 \UnaryInfC{$p : \Box A \times \Box B \vdash \pi_2 p : \Box B$}
@@
 \AxiomC{$y : B \vdash y : B$}
+\RightLabel{$\times_I$}
 \BinaryInfC{$x : A, y : B \vdash \langle x, y \rangle : A \times B$}
+\RightLabel{${\bf let}_{\Box}$}
 \TrinaryInfC{$p : \Box A \times \Box B \vdash {\bf let \: box \:} x,y = \pi_1 p, \pi_2 p {\: \bf in \:} \langle x, y \rangle : \Box (A \times B)$}
 \end{prooftree}
+\end{small}
 
